### 2016-08-05 journalSwarmControl/Introduction.tex
Journal added

\section{Introduction}\label{sec:Intro}
%This project studies system models and user interfaces for five multi-robot manipulation tasks with large populations of micro- and nanorobots.  We test several system models with different limitations on controllability and observability of the motion controller, and evaluate several different user interfaces.  We conduct user experiments to understand the impact of these limitations and design choices.

%Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.  
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. These activities require robots that behave intelligently.
Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population.  It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. However, it is often possible to sense global properties of the group, such as mean position and density.  Finally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of robots is a daunting human-swarm interaction (HSI) challenge.

Our previous work with over a hundred hardware robots and thousands of simulated robots~\cite{Becker2013b} demonstrated that direct human control of large swarms is possible. Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. This paper presents a tool for investigating HSI methods through statistically significant numbers of experiments.  There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.

\begin{figure}
\renewcommand{\figwid}{0.32\columnwidth}
\subfloat[][Vary Number]{\label{fig:VaryNum}
\begin{overpic}[width =\figwid]{VaryNum.pdf}\end{overpic}}
%
\subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\end{overpic}}\\
%
\subfloat[][Vary Control]{\label{fig:VaryControl}
\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
%
\subfloat[][Vary Noise]{\label{fig:VaryNoise}
\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}
%
\subfloat[][Control Position]{\label{fig:ControlPos}
\begin{overpic}[width =\figwid]{ControlPos.pdf}\end{overpic}}
%
\caption{\label{fig:5experiments}
Screenshots from our five online experiments controlling multi-robot systems with limited, global control.
\textbf{(a)} Varying the number of robots from 1-500
\textbf{(b)} Comparing 4 levels of visual feedback 
\textbf{(c)} Comparing 3 control architectures
\textbf{(d)} Varying noise from 0 to 200\% of control authority
\textbf{(e)} Controlling the position of 1 to 10 robots.
\href{http://youtu.be/HgNENj3hvEg}{See video overview at \cite{ShivaVideo2015}}
\vspace{-2em}
}
\end{figure}

Our goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size. Towards this end, we created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}, an open-source online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research. Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.  \href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are posted online, \cite{Chris-Ertel2013}. {TODO: link to 2016 code}

Our experiments show that numerous simple robots responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.

%Inspired by our online experiments, because it is not always possible to gather pose information on each robot for feedback control and
Often robots are difficult or impossible to sense individually due to their size and location. 
For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group, such as mean position and variance. 
To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers, inspired by our online experiments, that require only mean and variance measurements of the robot's positions.  
These controllers are used as primitives to perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.

\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{CompleteMazeDescription.pdf}\end{overpic}
\begin{overpic}[width=0.438\columnwidth]{MainExpFig.pdf}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.  Here a swarm of simple robots (blue discs) pushes a green hexagon toward the goal. See video attachment~\cite{ShivaVideo2015}.}
\end{figure}
 %function and implementation function

 Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}. We report the results of our online experiments in Section \ref{sec:expResults}. Inspired by these results, we prove that the mean and variance of a robot swarm are controllable in Section \ref{sec:theory}, and present automatic controllers in Section \ref{sec:simulation}. We conclude with implementations of these controllers on hardware robots and use them to complete an object manipulation task  with 100+ kilobots in Section \ref{sec:realExperiment}.

 This paper is the synthesis of two preliminary conference papers, the first, \cite{swarmcontrol2013}, covering first few months of SwarmControl.net experiments, and the second, \cite{ShahrokhiIROS2015}, with simulations of object manipulation.  This paper presents the final results from SwarmControl.net.  All hardware validation experiments are new.

### 2016-08-05 journalSwarmControl/SwarmControlIJRR.tex
Journal added

% How many words:  $ detex IJRRensembleDifferentialDrive.tex | wc -w
% Submit at http://mc.manuscriptcentral.com/ijrr

\documentclass[11pt]{article}
% IJRR Does not have a template for submissions, but it recommends using
% the standard Latex article format, with computer modern font.
% \usepackage[margin=1in]{geometry}
% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}

% www.SwarmControl.net
%http://mrsl.rice.edu
%http://people.ae.illinois.edu/tbretl/

\usepackage[pdftex]{graphicx}        % standard LaTeX graphics tool
\usepackage[usenames,dvipsnames]{color}
\usepackage{calc}
\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{color}
\usepackage{colortbl} %TODO: lab computers choke on this.  Why?
%\usepackage{cite}
\usepackage[font={small}]{caption}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{overpic}
\graphicspath{{./},{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{bm}   % boldface math type
\usepackage{upgreek}   % non-slanted greek letters
\usepackage{url}
\usepackage{soul} %for highlighting
\usepackage{hyperref}
\hypersetup{
  colorlinks =true,
  urlcolor = black,
  linkcolor = black,
  citecolor= black
}
\usepackage[round]{natbib}
\usepackage{mathtools}
\usepackage{authblk}

%% ABBREVIATIONS
\newcommand\mydefeq{\stackrel{\mathclap{\normalfont\mbox{\tiny def}}}{=}}
\newcommand{\qstart}{\mathbf{q}_{\text{start}}}
\newcommand{\qgoal}{\mathbf{q}_{\text{goal}}}
\newcommand{\pstart}{\mathbf{p}_{\text{start}}}
\newcommand{\pgoal}{\mathbf{p}_{\text{goal}}}
\newcommand{\xstart}{x_{\text{start}}}
\newcommand{\xgoal}{x_{\text{goal}}}
\newcommand{\ystart}{y_{\text{start}}}
\newcommand{\ygoal}{y_{\text{goal}}}
\newcommand{\gammastart}{\gamma_{\text{start}}}
\newcommand{\gammagoal}{\gamma_{\text{goal}}}
\newcommand{\CM}{\textsc{cm}}
\newcommand{\SVM}{\textsc{svm}}
\newcommand{\NN}{\textsc{nn}}
\newcommand{\prm}{\textsc{prm}}
\newcommand{\lemur}{\textsc{lemur}}
\newcommand{\Lemur}{\textsc{Lemur}}
\newcommand{\LP}{\textsc{lp}} 
\newcommand{\SOCP}{\textsc{socp}}
\newcommand{\SDP}{\textsc{sdp}}
\newcommand{\NP}{\textsc{np}}
\newcommand{\SAT}{\textsc{sat}}
\newcommand{\LMI}{\textsc{lmi}}
\newcommand{\hrp}{\textsc{hrp\nobreakdash-2}}
\newcommand{\DOF}{\textsc{dof}}
\newcommand{\UIUC}{\textsc{uiuc}}
\newcommand{\Cspace}{\mathcal{Q}}
\newcommand{\Uspace}{\mathcal{U}}
\newcommand{\D}{\nobreakdash-\textsc{d}}
\newcommand{\contact}[1]{\Cspace_{#1}}
\newcommand{\feasible}[1]{\Fspace_{#1}}
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\ii}{{\rm i}}
\newcommand{\TRUE}{\textbf{true}}
\newcommand{\FALSE}{\textbf{false}}
\newcommand*{\email}[1]{\texttt{#1}}

\newcommand{\figwid}{0.22\columnwidth}
%\newcommand{\topic}[1]{\textcolor{ForestGreen}{\footnotesize \textsf{#1}}}
%\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}

%% MACROS
\providecommand{\proc}[1]{\textsc{#1}}
\providecommand{\ddelta}{\underline{\bm{\updelta}}}  % requires \usepackage{bm,upgreek}
\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
\providecommand{\lscal}{<}
\providecommand{\gscal}{>}
\providecommand{\lvect}{\prec}
\providecommand{\gvect}{\succ}
\providecommand{\leqscal}{\leq}
\providecommand{\geqscal}{\geq}
\providecommand{\leqvect}{\preceq}
\providecommand{\geqvect}{\succeq}
\providecommand{\onevect}{\mathbf{1}}
\providecommand{\zerovect}{\mathbf{0}}
\providecommand{\field}[1]{\mathbb{#1}}
\providecommand{\C}{\field{C}}
\providecommand{\R}{\field{R}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\Hcal}{$\mathcal{H}$}
\providecommand{\Vcal}{$\mathcal{V}$}
\providecommand{\polar}{\triangle}
\providecommand{\ainner}{\underline{a}}
\providecommand{\aouter}{\overline{a}}
\providecommand{\binner}{\underline{b}}
\providecommand{\bouter}{\overline{b}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
\providecommand{\iff}{\Leftrightarrow}
\providecommand{\subinner}[1]{#1_{\text{inner}}}
\providecommand{\subouter}[1]{#1_{\text{outer}}}
\providecommand{\Ppoly}{\mathcal{X}}
\providecommand{\Pproj}{\mathcal{Y}}
\providecommand{\Pinner}{\subinner{\Pproj}}
\providecommand{\Pouter}{\subouter{\Pproj}}
\providecommand{\Aineq}{B}
\providecommand{\Aeq}{A}
\providecommand{\bineq}{u}
\providecommand{\beq}{t}

\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\homog}{homog}
\DeclareMathOperator{\domain}{dom}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\sign}{sgn} 
\DeclareMathOperator{\argmax}{arg\,max}
\DeclareMathOperator{\area}{area}
\DeclareMathOperator{\atan2}{atan2}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}

%useful for highlighting sections that need work
%\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}

\begin{document}
\hyphenation{micro-robot nano-robot nano-robotics micro-robotics micro-robots nano-robots micro-manip-ulators}
%%%%%%%%%%%% For debugging purposes, I like to display the TOC
    \tableofcontents
    \setcounter{tocdepth}{3}
    \newpage
%%%% END TOC %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\LARGE \bf Steering a Swarm Using Global Inputs And Swarm Statistics}

\author[1]{Shiva Shahrokhi}
\author[2]{Chris Ertel}
\author[1]{Mable Wan}
\author[1]{Lillian Lin}
\author[1]{\\ Aaron T.\ Becker}
\affil[1]{Department of Electrical and Computer Engineering, University of Houston\thanks{S.\ Shahrokhi and A.\ Becker are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
}
%\affil[2]{Independent}
%\author{Shiva Shahrokhi, Chris Ertel\thanks{S.\ Shahrokhi and A.\ Becker are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
%, Mable Wan, Lillian Lin, Aaron T. Becker
%
% %end thanks
%} % end author block 
\maketitle

\begin{abstract}
Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion. We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures. We present an alternative.

This paper first examines the perils, pitfalls, and possibilities we discovered by launching \href{http://www.swarmcontrol.net}{SwarmControl.net}, an online game where players steer swarms of up to 500 robots to complete manipulation challenges. We record statistics from thousands of players, and use the game to explore aspects of large-population robot control. We present the game framework as a new, open-source tool for large-scale user experiments. One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback. Inspired by human operators, this paper next investigates controllers that use only the mean and variance of a robot swarm. We prove that the mean position is controllable, then provide conditions under which variance is controllable.  We next derive automatic controllers for these and a hybrid, hysteresis-based switching control to regulate the first two moments of the robot distribution.  Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.
\end{abstract}
%\input{00-Abstract.tex}
\input{Introduction.tex}
\input{relatedWork.tex}
\input{expMethods.tex}
\input{expResults.tex}
\input{theory.tex}
\input{simulation.tex}
\input{experiment.tex}
\input{realExperiment.tex}

%\section*{Acknowledgements}

%\section*{Funding} %Formatted per guidelines at http://www.uk.sagepub.com/journalgateway/pubPolicies.htm#

   %http://www.nsf.gov/awardsearch/showAward?AWD_ID=1035716  James

   % Matlab codes for running the open-loop and closed-loop simulations in this paper are available at:
%
%Aaron Becker. “Range and Bearing Control of an Ensemble of Robots”, September 2012a. URL http://www.mathworks.com/matlabcentral/ fileexchange/38190.
%
%Aaron Becker. “Feedback Control of Many Differential-Drive Robots with Uniform Control Inputs.” MATLAB Central File Exchange, September 2012b. URL http://www.mathworks.com/matlabcentral/ fileexchange/38369.

%\input{11-MultiMediaAppendix.tex}

%\bibliographystyle{IEEEtran}
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,SwarmControlIJRR}

\end{document}

### 2016-08-05 journalSwarmControl/expResults.tex
Journal added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Human-swarm interaction results}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We designed five experiments to investigate human control of large robotic swarms for manipulation tasks.  Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.  Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, three control architectures, different levels of Brownian noise, and position control with 1 to 10 robots. The users could choose which experiment to try, and then our architecture randomly assigned a particular parameter value for each trial.  We recorded the completion time and the participant ID for each successful trial.  As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.  Still, many played multiple games, and their decreasing completion times demonstrates their skills improved.

\begin{figure}
\begin{overpic}[width = 0.48\columnwidth]{gamesPerPlayer.pdf}\end{overpic}
\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
\caption{
\label{fig:Learning}
(Left) The total number of games played per player drops off exponentially. (Right) We are able to show that players skill improves as they retry tests using data from \emph{Varying Number}
}
\end{figure}

% RESULTS
%% For each task, list the hypothesis, then give plot

%\begin{figure}
%\centering
%\begin{overpic}[width = 0.5\columnwidth]{DesignSpace.pdf}\end{overpic}
%\caption{
%\label{fig:DesignSpace} 
%There is a multi-dimensional design space for controlling robot swarms.  In this paper we provide a framework for exploring this design space. To demonstrate the capabilities of this framework, in Section \ref{sec:experiments} we describe experiments that test task completion time as a function of the number of robots, the quality of the state feedback, and the control architecture.}
%\end{figure}

\paragraph{Varying number}
\begin{figure}
\centering
\begin{overpic}[width = 0.5\columnwidth]{ResVaryNum.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.  The data indicates that this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object. Best-fit linear and quadratic lines are overlaid for comparison. 
%\vspace{-1em}
}
\end{figure}

Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems. This experiment varied from 1 to 500 the population of robots used to transport an object. We kept the total area, maximum robot speed, and total net force the swarm could produce constant. The robots pushed a large hexagonal object through an  {\sffamily S}-shaped maze. Our hypothesis was that participants would complete the task faster with more robots. The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating rather that there is a local minima around 130 robots.

\paragraph{Varying control}
Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.

We compared attractive and repulsive control with the global control used for the other experiments. The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g. \cite{Khalil2013} and biological models, e.g. \cite{goodrich2012types}. STMs can be used to arrange atoms and make small assemblies~\cite{avouris1995manipulation}. An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules. In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
The experiment challenged players to assemble a three-block pyramid with a swarm of 16 robots.

%\todo{describe controller model here, use an equation}

The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControlVis}.a: attractive control was the fastest, followed by global control, with repulsive control a distant last.  The median time using repulsive control was four times longer than with attractive control.

%%\begin{figure}
%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
%%\caption{
%%\label{fig:attractiveForce}
%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
%%}
%%\end{figure}

\begin{figure}[b!]
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:ResVaryControlVis} Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid. Right: Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}. Surprisingly, players perform better with limited feedback--subjects with only the mean + variance  outperformed all others.
%\vspace{-1em}
}
\end{figure}

\paragraph{Varying visualization}

Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011}, scientists fasten molecules that fluoresce light when activated by a strong light source. Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all robots; {\bf convex-hull} draws a convex hull around the outermost robots; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same robot swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ robots. Convex-hull requires at worst $2n$, but usually a smaller number.  Mean requires two, and variance three, data points.  Mean and mean + variance are convenient even with millions of robots. Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.

% Additionally, as population characteristics, they are available even if only a percentage of the robots are detected each control cycle.
%Photobleaching: http://www.piercenet.com/browse.cfm?fldID=4DD9D52E-5056-8A76-4E6E-E217FAD0D86B
%
%Photobleaching is caused by the irreversible destruction of fluorophores due to either the prolonged exposure to the excitation source or exposure to high-intensity excitation light. Photobleaching can be minimized or avoided by exposing the fluor(s) to the lowest possible level of excitation light intensity for the shortest length of time that still yields the best signal detection; this requires optimization of the detection method using high sensitivity CCD cameras, high numerical aperture objective and/or the widest bandpass emission filter(s) available. Other approaches include using fluorophores that are more photostable than traditional fluorophores and/or using antifade reagents to protect the fluor(s) against photobleaching. Steps to avoid photobleaching are not feasible for all detection methods and should be optimized for each method used. For example, antifade reagents are toxic to live cells, and therefore they can only be used with fixed cells or tissue. Furthermore, some detection methods, such as flow cytometry, normally do not require steps to avoid photobleaching because of the extremely short exposure time of the fluorophore to the excitation source.

%\begin{figure}
%\centering
%\begin{overpic}[width = \columnwidth]{ResVaryVis.pdf}\end{overpic}
%\vspace{-2em}
%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}. Surprisingly, players perform better with limited feedback--subjects with only the mean + variance  outperformed all others.
%\vspace{-2em}
%}
%\end{figure}

\begin{figure}[b!]
\renewcommand{\figwid}{0.24\columnwidth}
\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
\vspace{-.5em}
\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}. This experiment challenges players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region. We record the completion time and other statistics.
%\vspace{-1em}
}
\end{figure}

To our surprise, our experiment indicates the opposite: players  with just the mean completed the task faster than those with full-state feedback.  As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean + variance, mean, full-state, convex-hull].  Anecdotal evidence from beta-testers who played the game suggests that tracking 100 robots is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator. Our beta-testers found convex-hull feedback confusing and irritating.  A single robot left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
%obscuring what the rest of the swarm is doing.

\paragraph{Varying noise}
Microrobots are affected by turbulence caused by random collisions with molecules. The effect of these collisions is called Brownian motion.

This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep as follows:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
Here $m_i,\psi_i$ are uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$, where $M$ is a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).

We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the robots move erratically.  Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.

\begin{figure}[b!]
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{ResVaryNoise.pdf}\end{overpic}
\begin{overpic}[width =\figwid]{ResPositioning.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time. Right: Increasing the number of robots resulted in longer completion times.  For more than 4 robots the goal pattern contained a void, which may have caused the longer completion times.
%\vspace{-1em}
}
\end{figure}

\paragraph{Position control}
This online experiment examined how completion time scales with the number of robots $n$. Using a single square obstacle, users arranged $n\in[1,10]$ robots into a specified goal pattern.  The goal pattern formed a block `A' with 10 robots, and lesser numbers of robots used a subset of these goal positions. Our hypothesis was that completion time would increase linearly with the number of robots, as with our position control algorithm in \cite{Becker2013b}.  Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.  Though the number of robots presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.

Note there is a bifurcation between $n$=4 and $n$=5 robots. For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.  A better experiment design would randomly place the goal positions.  Initially we tried this, but our beta-testers strongly disliked trying to arrange robots in random patterns.

### 2016-08-05 journalSwarmControl/experiment.tex
Journal added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Object manipulation results}\label{sec:exp}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section analyzes a \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller and by human users.

\subsection{Human-controlled object manipulation}

As we saw in previous section, players with just the mean completed the task faster than those with full-state feedback.  As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].  Interviews with  beta-testers suggests that tracking 100 robots was overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance was like using a ``spongy'' manipulator. Convex-hull feedback was confusing and irritating because a single robot left behind an obstacle would distort the entire hull, obscuring the information about the majority of the swarm.
%obscuring what the rest of the swarm is doing.

\subsection{Automated object manipulation (simulation)}
Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm. To solve this object manipulation task, we discretized the environment. On this discretized grid we used breadth-first search to determine $\mathbf{M}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.  The object's center of mass is at $\mathbf{b}$ and has radius $r_b$. 
Three constants are needed, where $k_1>k_2>1$ and $1>k_2>0$. All experiments used $[k_1,k_2,k_3] = [2.5,1.5,0.1]$.
The robots were directed to assemble behind the object at  $\mathbf{b} - k_2 r_b \nabla \mathbf{M}(\mathbf{b})$, then move to  $\mathbf{b} - k_3 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location. We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit. The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$. 
The control algorithm  for object manipulation is listed in Alg.~\ref{alg:BlockPushing}. 
Experimental results are summarized in Fig.~\ref{fig:AutoVeryParam}.a. It is shown that large number of robots will complete the task faster, although after some point adding more robots would not cause a significant improvement.

\begin{algorithm}
\caption{Object-manipulation controller for a robotic swarm.}\label{alg:BlockPushing}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$,  moveable object's center of mass $\mathbf{b}$, map of the environment, and the locations of all convex corners $\mathbf{C}$
\Require Robot distribution is unimodal
\Require Obstacle-free, straight-line path from swarm to moveable object
\State Compute $\mathbf{M}$, the distance to goal, with breadth-first search
\State Compute the gradient, $\nabla \mathbf{M}$
\State $\mathbf{C} \gets \mathrm{sort(\mathbf{C})}$ according to $-\mathbf{M}$
\While{$\mathbf{b}$ is not in goal region}
\State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
\If {$\sigma^2 > \sigma_{max}^2$}
\While{$\sigma^2 > \sigma_{min}^2$}
\State $\mathbf{c}_i \gets$ the nearest corner in $\mathbf{C}$ to $[\bar{x},\bar{y}]$
\State $ [x_{goal}, y_{goal}] \gets \mathbf{c}_i $
\If {$\mathbf{M}(\mathbf{b}) > \mathbf{M} (\mathbf{c}_i)$}
\State  $[x_{goal}, y_{goal}] \gets  \mathbf{c}_{i-1}$ 
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndIf
\EndWhile
\Else  
\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
	\Else
	\State$r_p \gets k_3 r_b$  \Comment{pushing move}
	\EndIf
\State $[x_{goal}, y_{goal}] \gets \mathbf{b} - r_p \nabla \mathbf{M}(\mathbf{b})$ 
\EndIf
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndWhile
\end{algorithmic}
\end{algorithm}

\begin{figure*}
\centering
%\renewcommand{\figwid}{0.19\columnwidth}
%\href{http://youtu.be/tCej-9e6-4o}{\begin{overpic}[width =\figwid]{story1.png}\put(6,15){T = 5 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story2.png}\put(6,15){T = 12 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story3.png}\put(6,15){T = 20 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story4.png}\put(6,15){T = 25 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story5.png}\put(6,15){T = 33 s}
%\end{overpic}}
\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See the video attachment for an animation~\cite{ShivaVideo2015}.}
%\vspace{-2em}
}
\end{figure*}

\begin{figure}
\centering
\begin{overpic}[scale=0.19]{BFSMode.png}
\end{overpic}
\begin{overpic}[scale=0.19]{GradientView.png}
\end{overpic}
\begin{overpic}[scale=0.19]{PolicyIter.png}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object (left), which is used to compute gradient vectors (middle). Using policy iterations enables encoding penalties for being near obstacles (right).
%\vspace{-2em}
}
\end{figure}

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
%\end{overpic}
\begin{overpic}[width=.9\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
\end{overpic}

\vspace{-1em}
\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control. The automatic controller will see the pink objects as obstacles and finds its path to the goal. See the video attachment for an animation~\cite{ShivaVideo2015}}
%\vspace{-2em}
}
\end{figure}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:AutoVeryParam}Parameter sweep for a) number of robots, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials.
%\vspace{-2em}
}
\end{figure*}

%Algorithm \ref{alg:BlockPushing} is an imperfect solution and has a failure mode if the robot swarm becomes multi-modal with modes separated by an obstacle, as shown in Fig.~\ref{fig:Failure}.  In this case, moving toward a corner will never reduce the variance below $\sigma_{min}^2$.

  The first challenge is to identify when the distribution has become multi-modal.  Measuring just the mean and variance is insufficient to determine if a distribution is no longer unimodal, but if the swarm is being directed to a corner, and the variance does not reduce below $\sigma_{min}^2$, the swarm has become separated. In this case, we must either manipulate with a partial swarm, or run a gathering algorithm.  For the  {\sffamily S}-shaped workspace in this study, an open-loop input that commands the swarm to move in succession \{{\sc west, north, east, south}\} will move the swarm to the bottom right corner.
This is not true for all obstacle fields. In a {\sffamily T}-shaped workspace, it is not possible to find an open-loop input that will move the entire swarm to the bottom of the {\sffamily T}.

  Using only the mean and variance may be overly restrictive.  Many heuristics using high-order moments have been developed to test if a distribution is multimodal~\cite{haldane1951simple}.  Often the sensor data itself, though it may not resolve individual robots, will indicate multi-modality.  For instance CCD images reveal clusters of bacteria, and MRI scans show agglomerations of particles~\cite{stuber2007positive}.  This data can be fitted with $k$-means or expectation maximization algorithms, and manipulation could be performed with the nearest swarm of sufficient size.

### 2016-08-05 journalSwarmControl/realExperiment.tex
Journal added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Object manipulation with hardware robots}\label{sec:realExperiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Environmental setup}
Our experiments use centimeter-scale hardware systems called \emph{kilobots}.  While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot \cite{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available commercially or as an open source platform~\cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and lights: four 50W LED floodlights  at the corners and four 30W LED floodlights on the sides of a 6 m square centered on the workspace and 1.5 m above the table. The lights were controlled using an Arduino Uno board connected to an 8 relay shield board.  Above the table, an overhead machine vision system tracks the position of the swarm. Laser-cut patterns for the neon green fiducial markers and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}.
\begin{figure*}
\begin{center}
	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
	\begin{overpic}[width=0.49\columnwidth]{SetUp.pdf}\end{overpic}
	\begin{overpic}[width=0.35\columnwidth]{MainExpFig.pdf}\end{overpic}
\end{center}

\caption{\label{fig:setup}
Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system (Left). A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.  Here is a swarm of hardware robots (kilobots) that pushes a green hexagon toward the goal (Right). See video attachment~\cite{ShivaVideo2015}
}
\end{figure*}
%\subsection{Mean Control With Real Robots}

\begin{figure*}
\begin{center}
	\begin{overpic}[width=0.6\columnwidth]{Mean_Control_experiment.eps}\end{overpic}
	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
\end{center}
\caption{\label{fig:realMean}
Mean Control plot with kilobots.
}
\end{figure*}

%\begin{figure}
%\begin{center}
%	\includegraphics[width=\columnwidth]{XYMeanControl.eps}
%\end{center}
%\caption{\label{fig:meanRobotFig}
%Mean Control experiment with kilobots.
%}
%\end{figure}

\subsection{Automated object manipulation (hardware experiment)}
\paragraph{Heuristics to handle outliers and improve performance}

The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. The controller in \cite{Shahrokhi2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object. This failure rate increases if  object weight increases or ground-robot friction increases. The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace. Robots that cannot reach the object due to obstacles skew these calculations. The state machine in Fig.\ \ref{fig:Region}.a solves this problem by creating two states for the maze: either main or transfer. Each state has a set of regions representing a discretized visibility polygon. Whenever the object crosses a region boundary the state toggles. The main regions are generated by extending obstacles until they meet another obstacle shown in Fig.~\ref{fig:Region}.b. The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions shown in Fig.~\ref{fig:Region}.c.
This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object. In the example, we want the robots to push the object to the right. Without filtering the robots, the orange star is the mean and the algorithm would instruct the robots to push the object southeast. The filtered mean is at the yellow star and the algorithm instructs the robots to push the object directly east. 
%When the object leaves main region 1 the maze state switches to transfer. 
%The object is in transfer region 0, so only robots in transfer region 0 are included in the mean and covariance calculations.  
This heuristic improves performance by $50\%$ or less regarding the object heaviness.
\begin{figure*}
\begin{center}
	\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf}\end{overpic}
	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}
	\begin{overpic}[width=0.45\columnwidth]{stateMachine.pdf}\end{overpic}
\end{center}
\caption{\label{fig:Region} The state machine and regions.
}
\end{figure*}

\paragraph{Potential fields for swarm management}

Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
The repulsive potential field is centered at object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion as it is shown in Fig.~\ref{fig:potentialField}.
\begin{equation}
F_{att} = -\zeta \Delta \rho / \rho 
\end{equation}

\[ F_{rep} =  \left\{
\begin{array}{ll}
      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho & \rho\leq \rho_0 ~\&~ \theta > \pi/2\\
      0 & otherwise \\
\end{array} 
\right. \]

In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$ . Because the kilobot hardware experiments have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.

\begin{figure}
\centering
\begin{overpic}[width=1\columnwidth]{PotentialField.pdf}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:potentialField} The attractive filed source is centered on (left). Repulsive field source is centered at object's COM(middle). We use a combined forces to use this method to avoid pushing the object backwards(right).}
\end{figure}

### 2016-08-05 journalSwarmControl/SwarmControlIJRR.tex
Added sections for funding and for conclusion and for the algorithm

@@
 \usepackage{mathtools}
-\usepackage{authblk}
+%\usepackage{authblk}
 
 %% ABBREVIATIONS
+\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}
 \newcommand\mydefeq{\stackrel{\mathclap{\normalfont\mbox{\tiny def}}}{=}}
@@
 
-\author[1]{Shiva Shahrokhi}
-\author[2]{Chris Ertel}
-\author[1]{Mable Wan}
-\author[1]{Lillian Lin}
-\author[1]{\\ Aaron T.\ Becker}
-\affil[1]{Department of Electrical and Computer Engineering, University of Houston\thanks{S.\ Shahrokhi and A.\ Becker are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
-}
+%\author{Shiva Shahrokhi, Chris Ertel, Mable Wan,Lillian Lin}
+%\author{Aaron~T.~Becker}
+%\affil[1]{Dept.\ Electrical \& Computer Engineering, University of Houston\thanks{ University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu}}
+%}
 %\affil[2]{Independent}
-%\author{Shiva Shahrokhi, Chris Ertel\thanks{S.\ Shahrokhi and A.\ Becker are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
-%, Mable Wan, Lillian Lin, Aaron T. Becker
-%
-% %end thanks
-%} % end author block 
+\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T.
+Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
+ %end thanks
+} % end author block 
 \maketitle
-
 \begin{abstract}
@@
 \input{realExperiment.tex}
+\input{conclusion.tex}
+
+\section*{Funding} %formatted according to https://us.sagepub.com/en-us/nam/funding-acknowledgements
+%RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
+This work was supported by the National Science Foundation under Grant No.\ 
+\href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
+%Optional disclaimer (mandatory for non-scientific)
+%Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
 
 %\section*{Acknowledgements}
-
-%\section*{Funding} %Formatted per guidelines at http://www.uk.sagepub.com/journalgateway/pubPolicies.htm#
-
-   %http://www.nsf.gov/awardsearch/showAward?AWD_ID=1035716  James
-
-   % Matlab codes for running the open-loop and closed-loop simulations in this paper are available at:
-%
-%Aaron Becker. “Range and Bearing Control of an Ensemble of Robots”, September 2012a. URL http://www.mathworks.com/matlabcentral/ fileexchange/38190.
-%
-%Aaron Becker.
-“Feedback Control of Many Differential-Drive Robots with Uniform Control Inputs.” MATLAB Central File Exchange, September 2012b.
-URL http://www.mathworks.com/matlabcentral/ fileexchange/38369.
 

### 2016-08-05 journalSwarmControl/conclusion.tex
Added sections for funding and for conclusion and for the algorithm

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \todo{rewrite}

    Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion. We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures. We present an alternative.

This paper first examines the perils, pitfalls, and possibilities we discovered by launching \href{http://www.swarmcontrol.net}{SwarmControl.net}, an online game where players steer swarms of up to 500 robots to complete manipulation challenges. We record statistics from thousands of players, and use the game to explore aspects of large-population robot control. We present the game framework as a new, open-source tool for large-scale user experiments. One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback. Inspired by human operators, this paper next investigates controllers that use only the mean and variance of a robot swarm. We prove that the mean position is controllable, then provide conditions under which variance is controllable.  We next derive automatic controllers for these and a hybrid, hysteresis-based switching control to regulate the first two moments of the robot distribution.  Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.

There are many avenues for future work.  Manipulation by large populations of robots is an immature area and there are many open questions. We invite other collaborators to submit their own experiments to SwarmControl.net.
Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, and targeted drug delivery in a vascular network.

%\begin{figure}
%\begin{overpic}[width = 0.48\columnwidth]{Worldbrowsing.pdf}\end{overpic}
%\begin{overpic}[width = 0.48\columnwidth]{USbrowsing.pdf}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:PlayerLocation}Demographic information on game player's location, provided by Google Analytics. Game players from 84 countries and 49 US states visited our site.
%\vspace{-2em}
%}
%\end{figure}

%\paragraph{Site modifications}
%  The current site is optimized for desktop and laptop users, and we currently do not support mobile users. Our IRB allows us to conduct demographic questionnaires, and we will implement these questionnaires in a future release--currently our only source of demographic data is Google Analytics.
%  
%  We are pursuing partnerships to increase the educational content on our website. Our goal is to highlight a variety of the leading micro- and nano-robotics labs and the challenges they are working on.

%\paragraph{Automatic controllers}
%We have compiled a large body of test results.  Our goal is to design automatic controllers using this data. One avenue is to identify the most proficient players and perform inverse optimal control algorithms to learn the cost functions used by the best players.  
%%  demographics questionaires:

%We will use the video game described above as part of our outreach efforts.  The nature of ensemble control and the manipulation tasks will make this a puzzle game, where the user will need to determine the correct actions, and groups of actions, to accomplish the task.
%Discovering when a task changes from `fun' to `frustrating' is an active problem in game design.  Game companies (\emph{e.g.} ReignDesign Fig.~\ref{fig:Flockwork}) depend heavily on beta-testing to discover the point at which a task becomes frustrating to a user.  A model of what makes a task frustrating would revolutionize the game industry. Our estimate of this difficulty metric for massive manipulation is shown in Fig.~\ref{fig:GameDifficulty}.  ReignDesign has experience and current contracts designing educational games.  We can incorporate real-world physics to simulate robot control at the micro and the nano scale.
%
%Currently, we can use our control theoretic results to determine what tasks are possible. Unfortunately, this gives us no metric of human difficulty -- which tasks are easy for a human pilot.  What tasks should be off-loaded for computer control?  Fortunately \emph{gamification} provides built-in tools to gather this data in a transparent manner with the user's consent by measuring the time and number of actions required to complete each task, and through \emph{leaderboards} \cite{Zichermann2011,Kapp2012}, which rank users based on the efficiency of their solutions.  Key to the success of this endeavor is an engaging story.  We have the genesis of the a game in \cite{Becker2012l}, with game play based on the desire to steer many robots equipped with suction-cup darts in a surprise attack against an older sister. Our hope is to use current micro- and nanorobotics research to create an engaging story users will delight to immerse themselves in.

%In order to understand the difficulty of the tasks, we can measure the time and number of actions required to complete each task, as in Fig.~\ref{fig:Flockwork}.  This information will be collected to maintain a `top scores' list on the website, and help us answer questions about task difficulty.

%hopefully a discussion about flockworks, a successful multi-agent simulation (but actually a mobile app) here....

%http://gamification.org/wiki/Game_Features/Leaderboards

%   - Chemistry, Micro/nano education, molecular dynamics

### 2016-08-05 journalSwarmControl/experiment.tex
Newer version
Added sections for funding and for conclusion and for the algorithm

@@
 
-This section analyzes a \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller and by human users.
+This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller and by human users.
 
@@
 
-As we saw in previous section, players with just the mean completed the task faster than those with full-state feedback.
+As reported in the previous section, players with just the mean completed the task faster than those with full-state feedback.
 As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
@@
 
-\subsection{Automated object manipulation (simulation)}
-Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm.
-To solve this object manipulation task, we discretized the environment.
-On this discretized grid we used breadth-first search to determine $\mathbf{M}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
-The object's center of mass is at $\mathbf{b}$ and has radius $r_b$.
-Three constants are needed, where $k_1>k_2>1$ and $1>k_2>0$. All experiments used $[k_1,k_2,k_3] = [2.5,1.5,0.1]$.
+\subsection{Learning a policy for the object}
+The swarm must deliver the object to the goal region.  To solve this object manipulation task, we discretized the environment. 
+In paper (old) we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
+The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$.
 The robots were directed to assemble behind the object at  $\mathbf{b} - k_2 r_b \nabla \mathbf{M}(\mathbf{b})$, then move to  $\mathbf{b} - k_3 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location.
+BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
+
+DISCUSS Policy iteration to learn $\mathbf{M}_{MDP}$, for a book chapter on PI, see \cite{Thrun2005}.
+All grid locations at the goal were assigned a large reward, each grid location was assigned a small cost, and grid locations next to obstacles were assigned a high cost.
+Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal, cite book.
+
+$\mathbf{M}_{BFS}$, and $\mathbf{M}_{MDP}$ are shown in Fig.~\ref{fig:BFSGradient}.
+\todo{report results of 10 simulations with BFS and with PI.  Give mean and std}
+
+\begin{figure}
+\centering
+\begin{overpic}[scale=0.19]{GradientView.png}
+\end{overpic}
+\begin{overpic}[scale=0.19]{PolicyIter.png}
+\end{overpic}
+\begin{overpic}[scale=0.19]{MDPexperiment.png}
+\end{overpic}
+\vspace{-1em}
+\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object  to compute gradient vectors (left).
+Using policy iterations enables encoding penalties for being near obstacles (middle).
+At right, the results of policy iteration on our experimental setup.
+%\vspace{-2em}
+}
+\end{figure}
+
+\subsection{Potential fields for swarm management with a compliant manipulator}
+
+Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
+To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
+The repulsive potential field is centered at object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion as it is shown in Fig.~\ref{fig:potentialField}.
+\begin{equation}
+F_{att} = -\zeta \Delta \rho / \rho 
+\end{equation}
+
+\[ F_{rep} =  \left\{
+\begin{array}{ll}
+      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho & \rho\leq \rho_0 ~\&~ \theta > \pi/2\\
+      0 & otherwise \\
+\end{array} 
+\right. \]
+
+In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$.
+Because the kilobot hardware experiments have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
+
+\todo{report results of 10 simulations with and without Potential fields.  Give mean and std}
+\begin{figure}
+\centering
+\begin{overpic}[width=1\columnwidth]{PotentialField.pdf}\end{overpic}
+%\todo{I like the 'target' symbol, but it is not self-documenting.
+We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.
+I think these are easiest to make in powerpoint.
+%Please use the same color and line style for the variance min and max as you use in Figure 4.
+%}
+%{blockpushingImageWithMeanAndVarianceOverlay.png}
+\caption{\label{fig:potentialField} The attractive filed source is centered on (left).
+Repulsive field source is centered at object's COM (middle).
+We use a combined forces to use this method to avoid pushing the object backwards (right).}
+\end{figure}
+
+\subsection{Outlier rejection}
+
+The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers.
+The controller in \cite{Shahrokhi2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object.
+This failure rate increases if  object weight increases or ground-robot friction increases.
+The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace.
+Robots that cannot reach the object due to obstacles skew these calculations.
+The state machine in Fig.\ \ref{fig:Region}.a solves this problem by creating two states for the maze: either main or transfer.
+Each state has a set of regions representing a discretized visibility polygon.
+Whenever the object crosses a region boundary the state toggles.
+The main regions are generated by extending obstacles until they meet another obstacle shown in Fig.~\ref{fig:Region}.b.
+The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions shown in Fig.~\ref{fig:Region}.c.
+This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object.
+In the example, we want the robots to push the object to the right.
+Without filtering the robots, the orange star is the mean and the algorithm would instruct the robots to push the object southeast.
+The filtered mean is at the yellow star and the algorithm instructs the robots to push the object directly east.
+%When the object leaves main region 1 the maze state switches to transfer. 
+%The object is in transfer region 0, so only robots in transfer region 0 are included in the mean and covariance calculations.  
+%This heuristic improves performance by $50\%$ or less regarding the object heaviness.
+\todo{report results of 10 simulations with and without regions.  Give mean and std}
+
+\begin{figure*}
+\begin{center}
+	\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf}\end{overpic}
+	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}\\
+	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\end{overpic}
+\end{center}
+\caption{\label{fig:Region} The state machine and regions.
+}
+\end{figure*}
+
+\subsection{Algorithm for object manipulation}
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit.
@@
 The control algorithm  for object manipulation is listed in Alg.~\ref{alg:BlockPushing}.
-Experimental results are summarized in Fig.~\ref{fig:AutoVeryParam}.a.
-It is shown that large number of robots will complete the task faster, although after some point adding more robots would not cause a significant improvement.
+
+\todo{add the timeout function}
 
@@
 \Require Obstacle-free, straight-line path from swarm to moveable object
-\State Compute $\mathbf{M}$, the distance to goal, with breadth-first search
+\State Compute $\mathbf{M}$, the distance to goal    \Comment{using $\mathbf{M}_{BFS}$ or $\mathbf{M}_{MDP}$} 
 \State Compute the gradient, $\nabla \mathbf{M}$
@@
 \Else  
-\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
-	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
-	\Else
+%\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
+%	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
+%	\Else
+%	\State$r_p \gets k_3 r_b$  \Comment{pushing move}
+%	\EndIf
 \State$r_p \gets k_3 r_b$  \Comment{pushing move}
-	\EndIf
+
 \State $[x_{goal}, y_{goal}] \gets \mathbf{b} - r_p \nabla \mathbf{M}(\mathbf{b})$ 
@@
 \end{algorithm}
+
+\subsection{Automated object manipulation (simulation)}
+Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation. 
+Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.
+Each trial measured the time to deliver the object to the goal location.
+The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied each simulation step) with $W=10$.
+
+The interaction between the robots and object is impulsive so, like the study of impulsive pulling in  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
+After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success.
+
+Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  Large amounts of noise increase completion time.
+
+The robots have limited force, so increasing the object weight increases completion time.
 
@@
 \vspace{-1em}
-\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.
-See the video attachment for an animation~\cite{ShivaVideo2015}.}
+\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See animation in~\cite{ShivaVideo2015}.}
 %\vspace{-2em}
@@
 
-\begin{figure}
-\centering
-\begin{overpic}[scale=0.19]{BFSMode.png}
-\end{overpic}
-\begin{overpic}[scale=0.19]{GradientView.png}
-\end{overpic}
-\begin{overpic}[scale=0.19]{PolicyIter.png}
-\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object (left), which is used to compute gradient vectors (middle).
-Using policy iterations enables encoding penalties for being near obstacles (right).
-%\vspace{-2em}
-}
-\end{figure}
-
-\begin{figure}
-\centering
-%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
-%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
-%\end{overpic}
-\begin{overpic}[width=.9\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
-\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
-\end{overpic}
-
-\vspace{-1em}
-\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control.
-The automatic controller will see the pink objects as obstacles and finds its path to the goal.
-See the video attachment for an animation~\cite{ShivaVideo2015}}
-%\vspace{-2em}
-}
-\end{figure}
-
 \begin{figure*}
@@
 \renewcommand{\figwid}{0.5\columnwidth}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}\put(1,60){a)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}\put(1,60){b)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}\put(1,65){c)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}\put(1,65){d)}
+
 \end{overpic}
 \vspace{-1em}
-\caption{\label{fig:AutoVeryParam}Parameter sweep for a) number of robots, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials.
+\caption{\label{fig:AutoVeryParam}Parameter sweep for a) number of robots, b) different noise values, c) object weight, and d) object shape.
+Each bar is labelled with the number of trials.
+Completion time is in seconds.
 %\vspace{-2em}
@@
 
-  The first challenge is to identify when the distribution has become multi-modal.
+%  The first challenge is to identify when the distribution has become multi-modal.
 Measuring just the mean and variance is insufficient to determine if a distribution is no longer unimodal, but if the swarm is being directed to a corner, and the variance does not reduce below $\sigma_{min}^2$, the swarm has become separated.
@@
 For the  {\sffamily S}-shaped workspace in this study, an open-loop input that commands the swarm to move in succession \{{\sc west, north, east, south}\} will move the swarm to the bottom right corner.
-This is not true for all obstacle fields. In a {\sffamily T}-shaped workspace, it is not possible to find an open-loop input that will move the entire swarm to the bottom of the {\sffamily T}.
+%This is not true for all obstacle fields. In a {\sffamily T}-shaped workspace, it is not possible to find an open-loop input that will move the entire swarm to the bottom of the {\sffamily T}.
 
-  Using only the mean and variance may be overly restrictive.
+%  Using only the mean and variance may be overly restrictive.
 Many heuristics using high-order moments have been developed to test if a distribution is multimodal~\cite{haldane1951simple}.

### 2016-08-05 journalSwarmControl/realExperiment.tex
Newer version
Added sections for funding and for conclusion and for the algorithm

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Our experiments use centimeter-scale hardware systems called \emph{kilobots}.
+While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
+The kilobot, reported in \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots.
+It is available commercially or as an open source platform from \cite{K-Team2015}.
+Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
+Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.
 
 \subsection{Environmental setup}  
-Our experiments use centimeter-scale hardware systems called \emph{kilobots}.
-While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot \cite{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available commercially or as an open source platform~\cite{K-Team2015}.
-Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
-Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.
 In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and lights: four 50W LED floodlights  at the corners and four 30W LED floodlights on the sides of a 6 m square centered on the workspace and 1.5 m above the table.
 The lights were controlled using an Arduino Uno board connected to an 8 relay shield board.
-Above the table, an overhead machine vision system tracks the position of the swarm.
-Laser-cut patterns for the neon green fiducial markers and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}.
+
+Above the table, an overhead machine vision system tracks the swarm.
+The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance)  by topology, the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
+
+The objects were 3D printed from ABS plastic.
+Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}.
 \begin{figure*}
@@
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[width=0.49\columnwidth]{SetUp.pdf}\end{overpic}
-	\begin{overpic}[width=0.35\columnwidth]{MainExpFig.pdf}\end{overpic}
+	\begin{overpic}[width=0.49\columnwidth]{SetUp.pdf}\put(1,75){a)}\end{overpic}
+	\begin{overpic}[width=0.35\columnwidth]{MainExpFig.pdf}\put(5,92){b)}\end{overpic}
 \end{center}
@@
 \end{figure*}
-%\subsection{Mean Control With Real Robots}
+
+\paragraph{Swarm mean control (hardware experiment)}
+
+Unlike the PD controller \eqref{eq:PDcontrolPosition}, we cannot command a force input to the kilobots.
+Instead, control is given by turning on one of eight lights.
+The kilobots run a phototaxis routine where they search for an orientation that aligns them with the light source, and then move with an approximately constant velocity toward this light.
+In practice, because the kilobots only have one light detector, they oscillate along this orientation.
+
+We use the sign of \eqref{eq:PDcontrolPosition}, and choose the closest orientation among the eight light sources.
+Fig.~\ref{fig:realMean} shows that this limited, discretized control still enables regulating the mean position of a swarm of 100 robots.
 
@@
 \caption{\label{fig:realMean}
-Mean Control plot with kilobots.
+Mean control plot with kilobots.
+\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red crcle in left hand corner of these, and put a scale bar in image}
 }
@@
 \subsection{Automated object manipulation (hardware experiment)}
-\paragraph{Heuristics to handle outliers and improve performance}
 
-The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers.
-The controller in \cite{Shahrokhi2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object.
-This failure rate increases if  object weight increases or ground-robot friction increases.
-The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace.
-Robots that cannot reach the object due to obstacles skew these calculations.
-The state machine in Fig.\ \ref{fig:Region}.a solves this problem by creating two states for the maze: either main or transfer.
-Each state has a set of regions representing a discretized visibility polygon.
-Whenever the object crosses a region boundary the state toggles.
-The main regions are generated by extending obstacles until they meet another obstacle shown in Fig.~\ref{fig:Region}.b.
-The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions shown in Fig.~\ref{fig:Region}.c.
-This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object.
-In the example, we want the robots to push the object to the right.
-Without filtering the robots, the orange star is the mean and the algorithm would instruct the robots to push the object southeast.
-The filtered mean is at the yellow star and the algorithm instructs the robots to push the object directly east.
-%When the object leaves main region 1 the maze state switches to transfer. 
-%The object is in transfer region 0, so only robots in transfer region 0 are included in the mean and covariance calculations.  
-This heuristic improves performance by $50\%$ or less regarding the object heaviness.
-\begin{figure*}
-\begin{center}
-	\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf}\end{overpic}
-	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}
-	\begin{overpic}[width=0.45\columnwidth]{stateMachine.pdf}\end{overpic}
-\end{center}
-\caption{\label{fig:Region} The state machine and regions.
-}
-\end{figure*}
+The kilobots performed five successful runs of object manipulation through the obstacle maze.
+Each trial used 100 kilobots.
+Videos of each are in the multimedia extension.
+Trials 2-5 were performed in a row with no failures in between.
+For each trial, fully charged kilobots were placed in the lower left-hand of the workspace, as shown in Fig.~\ref{fig:expSnapShot}.
+The moveable object was placed in the lower center of the workspace.
+The same controller as in the simulation, Alg.~{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}.
+Trials were run until the object COM entered the goal region.
+The trials ran for \{ \todo{123123,12312,12312,12321,123123} \} (minutes:seconds).
+This is a mean of \todo{34} and a standard deviation of \todo{234}.
 
-\paragraph{Potential fields for swarm management}
-
-Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion as it is shown in Fig.~\ref{fig:potentialField}.
-\begin{equation}
-F_{att} = -\zeta \Delta \rho / \rho 
-\end{equation}
-
-\[ F_{rep} =  \left\{
-\begin{array}{ll}
-      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho & \rho\leq \rho_0 ~\&~ \theta > \pi/2\\
-      0 & otherwise \\
-\end{array} 
-\right. \]
-
-In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$ .
-Because the kilobot hardware experiments have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
+%Rules: http://www.ijrr.org/historic/MMVideoGuideLines.pdf
 
@@
 \centering
-\begin{overpic}[width=1\columnwidth]{PotentialField.pdf}\end{overpic}
-%\todo{I like the 'target' symbol, but it is not self-documenting.
-We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.
-I think these are easiest to make in powerpoint.
-%Please use the same color and line style for the variance min and max as you use in Figure 4.
-%}
-%{blockpushingImageWithMeanAndVarianceOverlay.png}
-\caption{\label{fig:potentialField} The attractive filed source is centered on (left).
-Repulsive field source is centered at object's COM(middle).
-We use a combined forces to use this method to avoid pushing the object backwards(right).}
+%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
+%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
+%\end{overpic}
+\begin{overpic}[width=.9\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
+\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
+\end{overpic}
+\vspace{-1em}
+\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control.
+The automatic controller will see the pink objects as obstacles and finds its path to the goal.
+See the video attachment for an animation~\cite{ShivaVideo2015}}
+%\vspace{-2em}
+}
 \end{figure}

### 2016-08-05 journalSwarmControl/simulation.tex
Added sections for funding and for conclusion and for the algorithm

\section{Simulation of control laws}\label{sec:simulation}
% Simulation:
%%  Mean control results   (image and plot(s) )
%%  Variance Control      (image and plot(s) )
%%  Hybrid control   (image and plot(s) )

Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation~\cite{catto2010box2d}.  All experiments ran on a Chrome web browser on a 2.6 GHz Macbook.  \href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2015}.

\subsection{Controlling the mean position of many robots}

We performed a parameter sweep using the PD controller \eqref{eq:PDcontrolPosition} to identify the best control gains.  Representative experiments are shown in Fig.~\ref{fig:gainvalues}. 100 robots were used and the maximum speed was 3 meters per second. As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52 s with $K_{p}= 4$, and  $K_{d} = 1$. 
\begin{figure}
\centering
\begin{overpic}[width = 0.7\columnwidth ]{gains.eps}
\end{overpic}
%\begin{overpic}[width =0.49\columnwidth]{meanVariance4.eps}
%\end{overpic}
\vspace{-1em}
\caption{\label{fig:gainvalues} Tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots. %b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom). For ease of analysis, only $x$ position and variance are shown.
%\vspace{-2em}
}
\end{figure}

%\begin{figure}
%\centering
%\begin{overpic}[width = \columnwidth* 2/3 ]{meanVideoSnapShot.png}
%\end{overpic}
%\vspace{0 em}
%\caption{\label{fig:meanVideo} A frame from video attachment showing mean position control on a swarm of 200 robots. The mean position of the swarm traces ``SWARM"~\cite{ShivaVideo2015}. 
%%\vspace{-2em}
%}
%\end{figure}

\subsection{Controlling the variance of many robots}
\begin{figure}
\centering
\begin{overpic}[width =0.7 \columnwidth] {brownianWpublish.eps}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:varyBrownian} Increased noise results in more responsive variance control because stronger Brownian noise causes a faster increase of variance.
%\vspace{-2em}
}
\end{figure}

%cite the control law, explain experiment (number of robots, maximum speed, ).

For variance control we use the control law discussed in Section~\ref{sec:VarianceControl}. 
 Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.

%\todo{image showing control x variance and y-variance out of phase}

\subsection{Hybrid control of mean and variance of many robots}

Fig.~\ref{fig:hybrid} shows a simulation run of the hybrid controller in Alg.~\ref{alg:MeanVarianceControl} with 100 robots in a square workspace containing no internal obstacles. 
%\todo{plot showing 1.5 cycles of mean position, and a variance goal.  We might need a longer time}
\begin{figure}
\centering
\begin{overpic}[width =0.7\columnwidth]{meanVariance4.eps}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:hybrid} Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom). For ease of analysis, only $x$ position and variance are shown.
\vspace{-2em}
}
\end{figure}

%\begin{figure}
%\centering
%\begin{overpic}[width = \columnwidth * 2/3]{VideoSnapShot.png}
%\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:videoVar} A frame from video, using Alg.~\ref{alg:MeanVarianceControl} to control variance and mean position of a swarm of 200 robots~\cite{ShivaVideo2015}.
%%\vspace{-2em}
%}
%\end{figure}

### 2016-08-05 journalSwarmControl/theory.tex
Added sections for funding and for conclusion and for the algorithm

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Theory:
%% Models (of each robot, of the global input)
%% Impossibility result
%% Controlling Mean proof
%% Controlling Variance proof (CLF)
%% A Hybrid controller using hysteresis
Consider holonomic robots that move in the 2D plane. We want to control position and velocity of the robots. 
First, assume a noiseless system containing one robot with mass $m$.
 Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t)
\end{align}

and our state space representation as:
\begin{equation}
\begin{bmatrix}
\dot{x}\\ 
\ddot{x}\\
\dot{y}\\
\ddot{y}
\end{bmatrix} = \begin{bmatrix}
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0\\
0 & 0 & 0 & 1\\
0 & 0 & 0 & 0
\end{bmatrix}  \begin{bmatrix}
x\\
\dot{x}\\
y\\
\dot{y}
\end{bmatrix} + \begin{bmatrix}
0 & 0 \\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{bmatrix}  [u_x,u_y]
\end{equation}

We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
\begin{equation}
\mathcal{C} = [ B, AB, A^2B, ... , A^{n-1}B ].
\end{equation}

\begin{equation}
\textrm{Here }
\mathcal{C}=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],
\end{equation}
and thus all four states are controllable. This Section starts by proving independent position control of many robots is not possible, but the mean position can be controlled. We then provide conditions under which the variance of many robots is also controllable.

\subsection{Independent control of many robots is impossible}
A single robot is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
%\begin{equation}\left[ x_{1,1},x_{2,1},\ldots, x_{2n-1},x_{2n}\right]^\intercal = \left[ p_{x,1},v_{x,1},\ldots,p_{x,n},v_{x,n}\right]^\intercal. \nonumber\end{equation}
%\begin{align}
%\dot{p}_{x1} &= v_{x1}\\\nonumber
%\dot{v}_{x1} &= a_{x1}\\\nonumber
%\dot{p}_{x2} &= v_{x2}\\\nonumber
%\dot{v}_{x2} &= a_{x2}\\\nonumber
%&\vdots\\\nonumber
%\dot{p}_{xn} &= v_{xn}\\\nonumber
%\dot{v}_{xn} &= a_{xn}\nonumber
%\end{align}
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because $\mathcal{C}$ has rank two:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,, 
\ldots \right]
\end{equation}  
\subsection{Controlling mean position of many robots}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and $x$ velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} 
+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

We again analyze $\mathcal{C}$:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right]
\end{equation}
This matrix has rank two, and thus the average position and average velocity are controllable.

Due to symmetry of the control input, only the mean position and mean velocity are controllable. However, there are several techniques for breaking symmetry, for example by using obstacles as in \cite{Becker2013b}, or by allowing independent noise sources as in \cite{beckerIJRR2014}.

We control mean position with a PD controller that uses the mean position and mean velocity. Our control input is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\bar{v}_x) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\bar{v}_y)  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

\subsection{Controlling the variance of many robots}\label{sec:VarianceControl}

The variance, $\sigma_x^2,\sigma_y^2$, of $n$ robots' position is computed as:
\begin{align}\label{eq:meanVar}
 \overline{x}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n x_{i}, \qquad  %\nonumber \\ 
\sigma_x^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (x_{i} - \overline{x})^2,  \nonumber \\ 
 \overline{y}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n y_{i}, \qquad  %\nonumber \\ 
\sigma_y^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (y_{i} - \overline{y})^2.  
%NOT 1/(N-1) because we are measuring the actual Variance, not the variance of samples drawn from a distribution
\end{align}

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. 
Microscale systems are affected by unmodelled dynamics. 
These unmodelled dynamics are dominated by Brownian noise, described by~\cite{einstein1956investigations}. 
To model this~\eqref{eq:stdform} must be modified as follows:
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t)
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$. Given a large obstacle-free workspace, a \emph{Brownian noise} process increases the variance linearly with time.
\begin{equation}\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon} \end{equation}
 If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board~\cite{Becker2013b}.

%After some time, Gaussian distribution shapes the outline of the robots because of the Brownian noise feature:\\
%\begin{equation}
%P(x) = \frac{1}{\sigma \sqrt{2\pi } }e^{ -  \frac{\left( x - \mu  \right)^2 } {2\sigma ^2 }}
%\end{equation}
If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

 A flat obstacle can be used to decrease variance. Pushing a group of dispersed robots against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached. For large $n$, ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is 
 \begin{align} \label{eq:optimalvar}
 \sigma^2_{optimal}(n,r) \approx   \frac{\sqrt{3}}{\pi} n r^2\approx 0.55 n r^2.
 \end{align} 
% Sloan minimized second moment U = 1/d^2 \sum_{i=1}^{n} || P_i - P ||^2 = (\sqrt{3} n^2)/(4 \pi), where d = 2r.
% Therefore the variance is d^2 (\sqrt{3} n^2)/(4 \pi), or 4^r^2 (\sqrt{3} n^2)/(4 \pi) = r^2 (\sqrt{3} n^2)/(\pi)If we only need to reduce variance in a single axis, the variance can be reduced to zero, given sufficient space.

We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \citet{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
\begin{align}
\label{eq:LyapunovVariance}
V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2\nonumber\\
\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x})
\end{align}
We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{goal}$.
To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
\begin{align}\label{eq:controlVariance}
u(t) &=   \begin{cases}
	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{goal} \\ 
	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal}.
\end{cases} 
\end{align}
 For such a $u(t)$,
 \begin{align}
\dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{goal}, \sigma^2_{optimal}(n,r))  \\ 
	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal},
\end{cases} 
\end{align} and thus
$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$

A  controller to regulate the variance to $\sigma^2_{ref}$ is:
\begin{align}
u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}) \nonumber\\
u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
\end{align}
In a slight abuse of notation we call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{goal}(\sigma^2_{ref}) = r + \sqrt{3\sigma^2_{ref}}
\end{align}
 If another wall is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.

\subsection{Controlling both mean and variance of many robots}

The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is less than the maximum controlled speed, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.  Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{max}^2$. 
 Variance is reduced until less than $\sigma_{min}^2$, then control again regulates the mean position. 
 This technique satisfies control objectives that evolve at different rates as in~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes. The process is illustrated in Fig.~\ref{fig:hysteresis}.

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ 
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$

\If {$\sigma_x^2 > \sigma_{max}^2$}
\State $x_{goal}  \gets x_{min}$
\State $flag_x  \gets \TRUE$
\Else {\bf~if} {$flag_x$ \textbf{and} $\sigma_x^2 < \sigma_{min}^2$}
\State $x_{goal}  \gets  x_{target}$
\State $flag_x  \gets  \FALSE$
\EndIf

\If{$\sigma_y^2 > \sigma_{max}^2$}
\State $y_{goal}  \gets y_{min}$
\State $flag_y  \gets \TRUE$
\Else {\bf~if} {$flag_y$ \textbf{and} $\sigma_y^2 < \sigma_{min}^2$}
\State $y_{goal}  \gets  y_{target}$
\State $flag_y  \gets \FALSE$
\EndIf
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndLoop
\end{algorithmic}
\end{algorithm}
\begin{figure}
\centering
\begin{overpic}[scale=.3]{hysteresis.png}
\put(39,21){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
\put(39,5){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
\vspace{-1em}
\caption{\label{fig:hysteresis} Two states to control mean and variance of a robot swarm.
%\vspace{-2em}
}
\end{figure}

A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:VarianceMinMaxBand}:
\begin{align}
 \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\
  \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf

%$\sigma^2_{optimal}(n,r)$ for $n$ circles with radius $r$ is $\approx 
 %    \frac{\sqrt{3}}{\pi} (n r)^2\approx 0.55(n r)^2 $. 
% Sloan minimized second moment U = 1/d^2 \sum_{i=1}^{n} || P_i - P ||^2 = (\sqrt{3} n^2)/(4 \pi), where d = 2r.
% Therefore the variance is d^2 (\sqrt{3} n^2)/(4 \pi), or 4^r^2 (\sqrt{3} n^2)/(4 \pi) = r^2 (\sqrt{3} n^2)/(\pi)If we only need to reduce variance in a single axis, the variance can be reduced to zero, given sufficient space.

\begin{figure}
\centering
\begin{overpic}[width = 0.5\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:VarianceMinMaxBand} The switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The above plot uses robot radius $r=1/10$.
}\vspace{-1em}
\end{figure}

%what images should I show here?
%hysteresis control law, \cite{sadra2014}

### 2016-08-05 journalSwarmControl/realExperiment.tex
Added Pics

@@
 Trials were run until the object COM entered the goal region.
-The trials ran for \{ \todo{123123,12312,12312,12321,123123} \} (minutes:seconds).
-This is a mean of \todo{34} and a standard deviation of \todo{234}.
+The trials ran for \{ 24:25, 57:37, 50:00, 36:02, 45:07\} (minutes:seconds).
+This is a mean of \{42:38\} (minutes:seconds) and a standard deviation of \{234\}.
 

### 2016-08-05 journalSwarmControl/experiment.tex
Added Pics
fixed the algorithm

@@
 \subsection{Learning a policy for the object}
-The swarm must deliver the object to the goal region.  To solve this object manipulation task, we discretized the environment. 
-In paper (old) we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
+The swarm must deliver the object to the goal region.  To solve this object manipulation task, we first discretize the environment.
+
+In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
 The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$.
 The robots were directed to assemble behind the object at  $\mathbf{b} - k_2 r_b \nabla \mathbf{M}(\mathbf{b})$, then move to  $\mathbf{b} - k_3 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location.
-BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
+However, BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
 
-DISCUSS Policy iteration to learn $\mathbf{M}_{MDP}$, for a book chapter on PI, see \cite{Thrun2005}.
+\todo{PI stuff: shiva does this} Policy iteration to learn $\mathbf{M}_{MDP}$, for a book chapter on PI, see \cite{Thrun2005}.
 All grid locations at the goal were assigned a large reward, each grid location was assigned a small cost, and grid locations next to obstacles were assigned a high cost.
@@
 \end{overpic}
-\begin{overpic}[scale=0.19]{MDPexperiment.png}
+\begin{overpic}[scale=0.262]{MDPmap.pdf}
 \end{overpic}
-\vspace{-1em}
+\vspace{-0.5em}
 \caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object  to compute gradient vectors (left).
@@
 To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion as it is shown in Fig.~\ref{fig:potentialField}.
-\begin{equation}
-F_{att} = -\zeta \Delta \rho / \rho 
-\end{equation}
-
-\[ F_{rep} =  \left\{
+The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\nabla \mathbf{M}$ as shown in Fig.~\ref{fig:potentialField}.
+\begin{align}
+F_{att} &= -\zeta \Delta \rho / \rho \\
+%\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
+\phi &=\cos^{-1}\left( \frac{ \nabla \mathbf{M} \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\nabla \mathbf{M} } \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
+%\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
+ F_{rep} &=  \left\{
 \begin{array}{ll}
-      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho & \rho\leq \rho_0 ~\&~ \theta > \pi/2\\
-      0 & otherwise \\
+      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\&~\phi <  \theta \\
+      0, & \textrm{ otherwise} \\
 \end{array} 
-\right. \]
+\right.\\
+F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}
+\end{align}
 
@@
 
-\subsection{Outlier rejection}
+\subsection{Outlier rejection}\label{subsec:OutlierRejection}
 
@@
 	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}\\
+	\vspace{0.5em}
 	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\end{overpic}
@@
 \begin{algorithmic}[1]
-\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$,  moveable object's center of mass $\mathbf{b}$, map of the environment, and the locations of all convex corners $\mathbf{C}$
-\Require Robot distribution is unimodal
-\Require Obstacle-free, straight-line path from swarm to moveable object
-\State Compute $\mathbf{M}$, the distance to goal    \Comment{using $\mathbf{M}_{BFS}$ or $\mathbf{M}_{MDP}$} 
+\Require Knowledge of moveable object's center of mass $\mathbf{b}$; swarm mean $[\bar{x},\bar{y}]$ and variance $[\sigma_x^2, \sigma_y^2]$, each calculated using the regions function from \S \ref{subsec:OutlierRejection};  map of the environment
+\State Compute $\mathbf{M}$, the cost to goal    \Comment{using $\mathbf{M}_{BFS}$ or $\mathbf{M}_{MDP}$} 
 \State Compute the gradient, $\nabla \mathbf{M}$
-\State $\mathbf{C} \gets \mathrm{sort(\mathbf{C})}$ according to $-\mathbf{M}$
 \While{$\mathbf{b}$ is not in goal region}
@@
 \While{$\sigma^2 > \sigma_{min}^2$}
-\State $\mathbf{c}_i \gets$ the nearest corner in $\mathbf{C}$ to $[\bar{x},\bar{y}]$
-\State $ [x_{goal}, y_{goal}] \gets \mathbf{c}_i $
-\If {$\mathbf{M}(\mathbf{b}) > \mathbf{M} (\mathbf{c}_i)$}
-\State  $[x_{goal}, y_{goal}] \gets  \mathbf{c}_{i-1}$ 
+\State $ [x_{goal}, y_{goal}] \gets $ the nearest corner in current region
 \State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
-\EndIf
 \EndWhile
@@
 %	\EndIf
-\State$r_p \gets k_3 r_b$  \Comment{pushing move}
-
-\State $[x_{goal}, y_{goal}] \gets \mathbf{b} - r_p \nabla \mathbf{M}(\mathbf{b})$ 
+%\State$r_p \gets k_3 r_b$  \Comment{pushing move}
+\State Apply \eqref{eq:potentialfield}   \Comment{potential field to push object in correct direction}
 \EndIf
-\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
 \EndWhile

### 2016-08-06 journalSwarmControl/experiment.tex
Added Policy Iteration Algorithm

@@
 \subsection{Learning a policy for the object}
+
 The swarm must deliver the object to the goal region.  To solve this object manipulation task, we first discretize the environment. 
-
 In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
@@
 
-\todo{PI stuff: shiva does this} Policy iteration to learn $\mathbf{M}_{MDP}$, for a book chapter on PI, see \cite{Thrun2005}.
-All grid locations at the goal were assigned a large reward, each grid location was assigned a small cost, and grid locations next to obstacles were assigned a high cost.
-Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal, cite book.
+ This paper applies value iteration algorithm mentioned in \cite{Thrun2005} to learn a policy for the Markov Decision Process $\mathbf{M}_{MDP}$.
+At each instance the object can move in eight directions or stay in place.
+The corresponding reward function gives a high reward to the goal location.
+All other free spaces have a small penalty; while obstacles have a large negative reward.
+Value iteration is used to learn an optimal policy which is a function that assigns a control input to every possible state.
+
+The reward function $r(x,u)$ is defined as
+\begin{align}
+r(x,u) &=  \left\{
+\begin{array}{ll}
+     +100, &  \textrm{if } u \textrm{ leads to goal state}\\
+      -100, & \textrm{if } u \textrm{ leads to a state with an obstacle} \\
+      -1, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+
+where $u$ is action and $x$ is the current state.
+The policy iteration algorithm returns $\arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)]$.
+We set $N=200$ for the full convergence and $\hat{V}(x_j) $ is calculated by
+
+\begin{align}
+\hat{V} (x_j) = \gamma \max_u [r(x_j,u) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| u,x_j)],
+\end{align}
+where in our experiments $\gamma = 0.97$.
+%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
 
@@
 Using policy iterations enables encoding penalties for being near obstacles (middle).
-At right, the results of policy iteration on our experimental setup.
+At right, the results of policy iteration generated from the obstacles detected by the vision algorithm on the hardware experimental setup is shown.
 %\vspace{-2em}

### 2016-08-06 journalSwarmControl/realExperiment.tex
Added Policy Iteration Algorithm
Added the mean image

@@
 	\begin{overpic}[width=0.49\columnwidth]{SetUp.pdf}\put(1,75){a)}\end{overpic}
-	\begin{overpic}[width=0.35\columnwidth]{MainExpFig.pdf}\put(5,92){b)}\end{overpic}
+	\begin{overpic}[width=0.3\columnwidth]{TestedShapesR.pdf}\put(5,85){b)}\end{overpic}
 \end{center}
@@
 A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.
-Here is a swarm of hardware robots (kilobots) that pushes a green hexagon toward the goal (Right).
+Here is a swarm of hardware robots (kilobots) that pushes a green hexagon or a green circle toward the goal (Right).
 See video attachment~\cite{ShivaVideo2015}
@@
 \begin{center}
-	\begin{overpic}[width=0.6\columnwidth]{Mean_Control_experiment.eps}\end{overpic}
-	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
+	\begin{overpic}[width=\columnwidth]{RealMean}\end{overpic}
+%	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
 \end{center}
@@
 Mean control plot with kilobots.
-\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red crcle in left hand corner of these, and put a scale bar in image}
+%\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red crcle in left hand corner of these, and put a scale bar in image}
 }

### 2016-08-08 journalSwarmControl/SwarmControlIJRR.tex
Added Policy Iteration Algorithm
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 \DeclareMathOperator{\atan2}{atan2}
+\DeclareMathOperator*{\argmax2}{argmax}
 
@@
 \author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T.
-Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
+Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA, {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
  %end thanks

### 2016-08-08 journalSwarmControl/conclusion.tex
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 We present the game framework as a new, open-source tool for large-scale user experiments.
-One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback.
+One intriguing result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback.
 Inspired by human operators, this paper next investigates controllers that use only the mean and variance of a robot swarm.
@@
 
-There are many avenues for future work.
-Manipulation by large populations of robots is an immature area and there are many open questions.
-We invite other collaborators to submit their own experiments to SwarmControl.net.
+There are many avenues for future work.  Manipulation by large populations of robots has many open questions. We invite other collaborators to submit their own experiments to SwarmControl.net.
 Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, and targeted drug delivery in a vascular network.

### 2016-08-08 journalSwarmControl/expResults.tex
Newer version
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 We recorded the completion time and the participant ID for each successful trial.
-As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.
+%As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.
 Still, many played multiple games, and their decreasing completion times demonstrates their skills improved.
-
-\begin{figure}
-\begin{overpic}[width = 0.48\columnwidth]{gamesPerPlayer.pdf}\end{overpic}
-\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
-\caption{
-\label{fig:Learning}
-(Left) The total number of games played per player drops off exponentially. (Right) We are able to show that players skill improves as they retry tests using data from \emph{Varying Number}
-}
-\end{figure}
-
-% RESULTS
-%% For each task, list the hypothesis, then give plot
-
-%\begin{figure}
-%\centering
-%\begin{overpic}[width = 0.5\columnwidth]{DesignSpace.pdf}\end{overpic}
-%\caption{
-%\label{fig:DesignSpace} 
-%There is a multi-dimensional design space for controlling robot swarms.
-In this paper we provide a framework for exploring this design space.
-To demonstrate the capabilities of this framework, in Section \ref{sec:experiments} we describe experiments that test task completion time as a function of the number of robots, the quality of the state feedback, and the control architecture.}
-%\end{figure}
 
@@
 \begin{figure}
-\centering
 \begin{overpic}[width = 0.5\columnwidth]{ResVaryNum.pdf}\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.
-The data indicates that this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object.
+\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
+\caption{
+\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.
+(Left) Data indicates this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object.
 Best-fit linear and quadratic lines are overlaid for comparison.
-%\vspace{-1em}
+(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
 }
@@
 The median time using repulsive control was four times longer than with attractive control.
+Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
 
@@
 Right: Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.
-Surprisingly, players perform better with limited feedback--subjects with only the mean + variance  outperformed all others.
+Players performed better with limited feedback.
 %\vspace{-1em}
@@
 
-To our surprise, our experiment indicates the opposite: players  with just the mean completed the task faster than those with full-state feedback.
+Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
 As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean + variance, mean, full-state, convex-hull].
+All experiments lasting over 300s were removed, under the assumption that the user stopped playing
+Using ANOVA analysis, we reject the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
 Anecdotal evidence from beta-testers who played the game suggests that tracking 100 robots is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
@@
 Using a single square obstacle, users arranged $n\in[1,10]$ robots into a specified goal pattern.
-The goal pattern formed a block `A' with 10 robots, and lesser numbers of robots used a subset of these goal positions.
+The goal pattern formed a block  {\sffamily A} with 10 robots, and lesser numbers of robots used a subset of these goal positions.
 Our hypothesis was that completion time would increase linearly with the number of robots, as with our position control algorithm in \cite{Becker2013b}.

### 2016-08-08 journalSwarmControl/experiment.tex
updated MDP and policy iteration, did all captions, added new pic, removed extra images

--- 2016-08-05 journalSwarmControl/experiment.tex
@@
 
-This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller and by human users.
+This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in Section~\ref{sec:expResults}.  
+The swarm must deliver the object to the goal region.
+To solve this object manipulation task, we divide the task into two components: designing a policy for the object, and designing a control law for the swarm to push the object according to the policy.
 
-\subsection{Human-controlled object manipulation}
+\subsection{Learning a policy for the object}\label{subsec:objectpolicy}
 
-As reported in the previous section, players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
-Interviews with  beta-testers suggests that tracking 100 robots was overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance was like using a ``spongy'' manipulator.
-Convex-hull feedback was confusing and irritating because a single robot left behind an obstacle would distort the entire hull, obscuring the information about the majority of the swarm.
-%obscuring what the rest of the swarm is doing.
-
-\subsection{Learning a policy for the object}
-The swarm must deliver the object to the goal region.  To solve this object manipulation task, we first discretize the environment.
-
+To design the policy, we first discretize the environment. 
 In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
 The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$. 
-The robots were directed to assemble behind the object at  $\mathbf{b} - k_2 r_b \nabla \mathbf{M}(\mathbf{b})$, then move to  $\mathbf{b} - k_3 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location.
+The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \nabla \mathbf{M}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location.
 However, BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
 
-\todo{PI stuff: shiva does this} Policy iteration to learn $\mathbf{M}_{MDP}$, for a book chapter on PI, see \cite{Thrun2005}.
-All grid locations at the goal were assigned a large reward, each grid location was assigned a small cost, and grid locations next to obstacles were assigned a high cost.
-Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal, cite book.
+ To avoid obstacles, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
+  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}: a function that assigns a control input to every possible state.
+ At each state the object can be commanded to move in one of eight directions with a small probability of moving in the wrong direction.
 
-$\mathbf{M}_{BFS}$, and $\mathbf{M}_{MDP}$ are shown in Fig.~\ref{fig:BFSGradient}.
-\todo{report results of 10 simulations with BFS and with PI.  Give mean and std}
+ A corresponding reward function gives a high reward to the goal state,  
+ a large negative reward to states including obstacles, and a small negative reward to all other states.
+The reward function $r(x,u)$ is defined as
+\begin{align}
+r(x,u) &=  \left\{
+\begin{array}{ll}
+     +100, &  \textrm{if } u \textrm{ leads to goal state}\\
+      -100, & \textrm{if } u \textrm{ leads to a state with an obstacle} \\
+      -1, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+where $u$ is the action and $x$ is the current state. 
+  Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
+   \begin{align} \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
+   \end{align}
+   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
+\begin{align}
+\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
+&\hat{V} (x_j) = \gamma \max_u [r(x_j,u) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| u,x_j)] \label{eq:ValueIteration}\\
+\text{end}& \nonumber
+\end{align}
+In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \todo{link to algorithm}.  
+%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
+
+$\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
+In 10 simulations with 100 robots, pushing the object to goal using $\mathbf{M}_{BFS}$ required $\mu=XX, \sigma=YY$s while  $\mathbf{M}_{MDP}$ required $\mu=XX, \sigma=YY$.
+\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
 
@@
 \caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object  to compute gradient vectors (left).
-Using policy iterations enables encoding penalties for being near obstacles (middle).
-At right, the results of policy iteration on our experimental setup.
+Modeling as an MDP enables encoding penalties for being near obstacles.
+(Middle) The control policy from value iteration.
+(Right) Value function and control policy for the hardware setup.
+The vision algorithm detects obstacles.
 %\vspace{-2em}
@@
 
+In 10 simulations with 100 robots, pushing the object to goal without a repulsive potential field failed in xx of YY runs.
+No failures occurred with the  repulsive potential field.
+Of successful trials, completion time without repulsive potential fields required $\mu=XX, \sigma=YY$s while using repulsive potential fields required $\mu=XX, \sigma=YY$.
 \todo{report results of 10 simulations with and without Potential fields.  Give mean and std}
+
 \begin{figure}
@@
 %{blockpushingImageWithMeanAndVarianceOverlay.png}
-\caption{\label{fig:potentialField} The attractive filed source is centered on (left).
-Repulsive field source is centered at object's COM (middle).
-We use a combined forces to use this method to avoid pushing the object backwards (right).}
+\caption{\label{fig:potentialField} (Left) The attractive field is centered behind the object's COM.
+(Middle) The repulsive field is centered at the object's COM.
+(Right) Combining these forces prevents the swarm from pushing the object backwards.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig:Region} The state machine and regions.
+\caption{\label{fig:Region}  Outlier rejection state machine and regions.
 }
@@
 \Require Knowledge of moveable object's center of mass $\mathbf{b}$; swarm mean $[\bar{x},\bar{y}]$ and variance $[\sigma_x^2, \sigma_y^2]$, each calculated using the regions function from \S \ref{subsec:OutlierRejection};  map of the environment
-\State Compute $\mathbf{M}$, the cost to goal    \Comment{using $\mathbf{M}_{BFS}$ or $\mathbf{M}_{MDP}$} 
-\State Compute the gradient, $\nabla \mathbf{M}$
+\State Compute optimal policy for object, according to \S \ref{subsec:objectpolicy}
 \While{$\mathbf{b}$ is not in goal region}
@@
 %\State$r_p \gets k_3 r_b$  \Comment{pushing move}
+\State Calculate $\nabla \mathbf{M}(\mathbf{b})$  \Comment{policy for object at $\mathbf{b}$}
 \State Apply \eqref{eq:potentialfield}   \Comment{potential field to push object in correct direction}
@@
 The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied each simulation step) with $W=10$.
-
-The interaction between the robots and object is impulsive so, like the study of impulsive pulling in  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
-After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success.
-
-Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  Large amounts of noise increase completion time.
-
-The robots have limited force, so increasing the object weight increases completion time.
 
@@
 \end{overpic}
-\vspace{-1em}
+\vspace{-2em}
 \caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See animation in~\cite{ShivaVideo2015}.}
@@
 \end{figure*}
+
+The interaction between the robots and object is impulsive so, like the study of impulsive pulling in  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
+ After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success. 
+Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  Large amounts of noise increase completion time.
+The robots have limited force, so increasing the object weight increases completion time.  
+ Rectangles and squares tend to get stuck in the 90$^\circ$ workspace corners, and cause longer completion times than circles and hexagons.
 

### 2016-08-08 journalSwarmControl/realExperiment.tex
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 \end{center}
+\vspace{-1em}
 \caption{\label{fig:realMean}
-Mean control plot with kilobots.
-%\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red crcle in left hand corner of these, and put a scale bar in image}
+Regulating average $x$ position of 100 kilobots using control law \eqref{eq:PDcontrolPosition}.
+%Mean control plot with kilobots.
+%\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red circle in left hand corner of these, and put a scale bar in image}
 }
@@
 \caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control.
-The automatic controller will see the pink objects as obstacles and finds its path to the goal.
+The automatic controller classifies pink objects as obstacles and generates a policy to the goal.
 See the video attachment for an animation~\cite{ShivaVideo2015}}

### 2016-08-08 journalSwarmControl/simulation.tex
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 \vspace{-1em}
-\caption{\label{fig:gainvalues} Tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots.
+\caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots.
 %b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom).
@@
 \vspace{-1em}
-\caption{\label{fig:varyBrownian} Increased noise results in more responsive variance control because stronger Brownian noise causes a faster increase of variance.
+\caption{\label{fig:varyBrownian} In simulation, increased noise results in more responsive variance control because stronger Brownian noise causes a faster increase of variance.
 %\vspace{-2em}

### 2016-08-08 journalSwarmControl/theory.tex
updated MDP and policy iteration, did all captions, added new pic, removed extra images

@@
 One option is to control the position and velocity of the $j^{th}$ robot.
-To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and $x$ velocity of the $n$ robots:
+To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and velocity of the $n$ robots:
 \begin{align}

### 2016-08-09 journalSwarmControl/experiment.tex
Changed figures and added the results of comparison

@@
   Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
-   \begin{align} \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
+   \begin{align} \nabla \mathbf{M_{MDP}}= \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
    \end{align}
@@
 \end{align}
-In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \todo{link to algorithm}.  
+In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times.
+A {\sc Matlab} implementation of this algorithm is available at \href{https://www.mathworks.com/matlabcentral/fileexchange/49992-mdp-robot-grid-world-example}{\emph{MDP grid world example}}.%\todo{link to algorithm}.
 %Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
@@
 $\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 robots, pushing the object to goal using $\mathbf{M}_{BFS}$ required $\mu=XX, \sigma=YY$s while  $\mathbf{M}_{MDP}$ required $\mu=XX, \sigma=YY$.
-\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
+In 10 simulations with 100 robots, pushing the object to goal using $\mathbf{M}_{BFS}$ required $\mu=183, \sigma=179$ s while  $\mathbf{M}_{MDP}$ required $\mu=90, \sigma=35$ s.
+%\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
 
@@
 
-In 10 simulations with 100 robots, pushing the object to goal without a repulsive potential field failed in xx of YY runs.
+In 10 simulations with 100 robots, pushing the object to goal without a repulsive potential field failed in two of twelve runs.
 No failures occurred with the repulsive potential field.
-Of successful trials, completion time without repulsive potential fields required $\mu=XX, \sigma=YY$s while using repulsive potential fields required $\mu=XX, \sigma=YY$.
-\todo{report results of 10 simulations with and without Potential fields.  Give mean and std}
+Of successful trials, completion time without repulsive potential fields required $\mu=245, \sigma=135$ s while using repulsive potential fields required $\mu=90, \sigma=35$ s.
+%In 10 simulations with 100 robots, completion time without repulsive potential fields required $\mu=245, \sigma=135$s while using repulsive potential fields required $\mu=90, \sigma=35$.
+
+%\todo{report results of 10 simulations with and without Potential fields.  Give mean and std}
 
@@
 %This heuristic improves performance by $50\%$ or less regarding the object heaviness.
-\todo{report results of 10 simulations with and without regions.  Give mean and std}
+In 10 simulations with 100 robots, completion time without outlier rejection required $\mu=245, \sigma=135$ s while using outlier rejection required $\mu=90, \sigma=35$ s.
+%\todo{report results of 10 simulations with and without regions.  Give mean and std}
 
@@
 The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.
+A timer is also used to monitor the mean position of the swarm to know if the swarm is stuck in one of the corners.
+It gives the previous corner when the swarm is stuck to change the swarm's configuration and getting out of that local minimum.
 The control algorithm  for object manipulation is listed in Alg.~\ref{alg:BlockPushing}.
 
-\todo{add the timeout function}
+\todo{add the timeout function: I added two lines.}
 

### 2016-08-09 journalSwarmControl/realExperiment.tex
Changed figures and added the results of comparison

@@
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[width=0.49\columnwidth]{SetUp.pdf}\put(1,75){a)}\end{overpic}
-	\begin{overpic}[width=0.3\columnwidth]{TestedShapesR.pdf}\put(5,85){b)}\end{overpic}
+	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){a)}
+	\end{overpic}\\
+	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(5,85){b)}
+	\end{overpic}
+	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(5,85){c)}
+	\end{overpic}
+
 \end{center}
@@
 \caption{\label{fig:setup}
-Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system (Left).
-A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.
-Here is a swarm of hardware robots (kilobots) that pushes a green hexagon or a green circle toward the goal (Right).
+(Top)Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by four remotely triggered 30W LED and four 50W LED floodlights, with an overhead machine vision system.
+(Down left) A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.
+Here is a swarm of hardware robots (kilobots) that pushes a green object with different shapes toward the goal.
+(Down right) The potential fields view in hardware experiments.
 See video attachment~\cite{ShivaVideo2015}
@@
 Videos of each are in the multimedia extension.
-Trials 2-5 were performed in a row with no failures in between.
+Trials two-five were performed in a row with no failures in between.
 For each trial, fully charged kilobots were placed in the lower left-hand of the workspace, as shown in Fig.~\ref{fig:expSnapShot}.
 The moveable object was placed in the lower center of the workspace.
-The same controller as in the simulation, Alg.~{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}.
+The same controller as in the simulation, Alg.~\ref{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}.
 Trials were run until the object COM entered the goal region.
-The trials ran for \{ 24:25, 57:37, 50:00, 36:02, 45:07\} (minutes:seconds).
-This is a mean of \{42:38\} (minutes:seconds) and a standard deviation of \{234\}.
+The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} (minutes:seconds).
+This is a mean of \{42:38\} (minutes:seconds) and a standard deviation of \{12:51\}(minutes:seconds).
+Also there was a trial with the circular object for \{52:35 \} (minutes:seconds).
+Another trial with square like object ran for \{114:31 \} (minutes:seconds).
+We also tried triangular object for \{ XX:YY \} (minutes:seconds).
 

### 2016-08-09 journalSwarmControl/Introduction.tex
Added sections for funding and for conclusion and for the algorithm
fixed the algorithm
updated MDP and policy iteration, did all captions, added new pic, removed extra images
added images for foraging

@@
 \subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
-\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\end{overpic}}\\
-%
-\subfloat[][Vary Control]{\label{fig:VaryControl}
-\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
+%\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
+\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\end{overpic}}
 %
 \subfloat[][Vary Noise]{\label{fig:VaryNoise}
-\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}
+\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}\\
 %
@@
 %
+\subfloat[][Vary Control: Assembly]{\label{fig:VaryControl}
+\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
+%
+\subfloat[][Cary Control: Foraging]{\label{fig:Forage}
+\begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
+%
 \caption{\label{fig:5experiments}
-Screenshots from our five online experiments controlling multi-robot systems with limited, global control.
+Screenshots from our online experiments controlling multi-robot systems with limited, global control.
 \textbf{(a)} Varying the number of robots from 1-500
 \textbf{(b)} Comparing 4 levels of visual feedback 
-\textbf{(c)} Comparing 3 control architectures
-\textbf{(d)} Varying noise from 0 to 200\% of control authority
-\textbf{(e)} Controlling the position of 1 to 10 robots.
-\href{http://youtu.be/HgNENj3hvEg}{See video overview at \cite{ShivaVideo2015}}
+\textbf{(c)} Varying noise from 0 to 200\% of control authority
+\textbf{(d)} Controlling the position of 1 to 10 robots
+\textbf{(e)} Comparing 3 control architectures to assemble
+\textbf{(f)} Comparing 3 control architectures to forage.
 \vspace{-2em}
@@
 \href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are posted online, \cite{Chris-Ertel2013}.
-{TODO: link to 2016 code}
+\todo{TODO: Chris, is 2016 code ready to upload?
+link to 2016 code}
 
@@
 Here a swarm of simple robots (blue discs) pushes a green hexagon toward the goal.
+Simulation at left, screen capture form hardware experiment at right.
 See video attachment~\cite{ShivaVideo2015}.}

### 2016-08-09 journalSwarmControl/expResults.tex
added images for foraging

@@
 
-We designed five experiments to investigate human control of large robotic swarms for manipulation tasks.
+We designed six experiments to investigate human control of large robotic swarms for manipulation tasks.
 Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, three control architectures, different levels of Brownian noise, and position control with 1 to 10 robots.
+Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 robots, and three control architectures for both an assembly task and a foraging task.
 The users could choose which experiment to try, and then our architecture randomly assigned a particular parameter value for each trial.
@@
 
-\paragraph{Varying control}
-Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.
-
-We compared attractive and repulsive control with the global control used for the other experiments.
-The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.
-\cite{Khalil2013} and biological models, e.g.
-\cite{goodrich2012types}.
-STMs can be used to arrange atoms and make small assemblies~\cite{avouris1995manipulation}.
-An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules.
-In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
-The experiment challenged players to assemble a three-block pyramid with a swarm of 16 robots.
-
-%\todo{describe controller model here, use an equation}
-
-The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControlVis}.a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
-The median time using repulsive control was four times longer than with attractive control.
-Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
-
-%%\begin{figure}
-%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
-%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
-%%\caption{
-%%\label{fig:attractiveForce}
-%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
-%%}
-%%\end{figure}
-
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.5\columnwidth}
-\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:ResVaryControlVis} Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid.
-Right: Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.
-Players performed better with limited feedback.
-%\vspace{-1em}
-}
-\end{figure}
-
 \paragraph{Varying visualization}
@@
 Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
+
+\begin{figure}[b!]
+\renewcommand{\figwid}{0.5\columnwidth}
+\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
+\vspace{-1em}
+\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
+%\vspace{-1em}
+}
+\end{figure}
 
@@
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryControlVis}.b shows, the levels of feedback arranged by increasing completion time are [mean + variance, mean, full-state, convex-hull].
+As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean + variance, mean, full-state, convex-hull].
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing
@@
 Initially we tried this, but our beta-testers strongly disliked trying to arrange robots in random patterns.
+
+\paragraph{Varying control: Assembly}
+Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.
+
+We compared attractive and repulsive control with the global control used for the other experiments.
+The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.
+\cite{Khalil2013} and biological models, e.g.
+\cite{goodrich2012types}.
+STMs can be used to arrange atoms and make small assemblies~\cite{avouris1995manipulation}.
+An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules.
+In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
+The experiment challenged players to assemble a three-block pyramid with a swarm of 16 robots.
+
+%\todo{describe controller model here, use an equation}
+
+The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
+The median time using repulsive control was four times longer than with attractive control.
+Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
+
+%%\begin{figure}
+%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
+%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
+%%\caption{
+%%\label{fig:attractiveForce}
+%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
+%%}
+%%\end{figure}
+
+\begin{figure}[b!]
+\renewcommand{\figwid}{3.3cm}
+\begin{overpic}[height =\figwid]{ResVaryControl.pdf}\end{overpic}
+\begin{overpic}[height =\figwid]{ResVaryForage.pdf}\end{overpic}
+\vspace{-.5em}
+\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type.
+Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid.
+Right: in the foraging test, global control resulted in the shortest completion time and attractive the longest.
+%\vspace{-1em}
+}
+\end{figure}
+
+\paragraph{Varying control: Foraging}
+Collecting and delivering resources is necessary for drug delivery.
+This experiment also compared attractive and repulsive control with the global control used for the other experiments.
+The experiment challenged players to collect particles using a swarm of 100 robots and return the particles to a home region.
+The robots encapsulate the particles on contact.
+
+%\todo{describe controller model here, use an equation}
+
+The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.
+Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $2.96\times10^{-6}$.

### 2016-08-09 journalSwarmControl/expResults.tex
added captions

@@
 We compared attractive and repulsive control with the global control used for the other experiments.
-The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.
-\cite{Khalil2013} and biological models, e.g.
+The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.\ \cite{Khalil2013} and biological models, e.g.
 \cite{goodrich2012types}.

### 2016-08-09 journalSwarmControl/experiment.tex
added captions

@@
   Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
-   \begin{align} \nabla \mathbf{M_{MDP}}= \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
+   \begin{align} \nabla \mathbf{M_{MDP}}(x) = \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
    \end{align}
@@
 $\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 robots, pushing the object to goal using $\mathbf{M}_{BFS}$ required $\mu=183, \sigma=179$ s while  $\mathbf{M}_{MDP}$ required $\mu=90, \sigma=35$ s.
+In 10 simulations with 100 robots, pushing the object to goal using BFS required a mean and standard deviation of $183\pm 179$ s while Policy Iteration required $90\pm 35$ s.
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.
-A timer is also used to monitor the mean position of the swarm to know if the swarm is stuck in one of the corners.
-It gives the previous corner when the swarm is stuck to change the swarm's configuration and getting out of that local minimum.
 The control algorithm  for object manipulation is listed in Alg.~\ref{alg:BlockPushing}.
 
-\todo{add the timeout function: I added two lines.}
+In rare cases during simulations the swarm may become trapped in a local minima of \eqref{eq:potentialfield}.
+If the swarm mean position does not change for ten seconds, the swarm is assumed to be in a local minima and is commanded to move toward the previous corner.
+As soon as the mean position changes, normal control resumes.
+
+%\todo{add the timeout function: I added two lines.}
 
@@
  After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success. 
-Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  Large amounts of noise increase completion time.
+Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  
+ Increasing noise increases completion time. 
+ Large amounts of noise caused failures, defined as trials lasting longer than 1000s.  With $W=50$, XX of 20 trials failed.  With $W=100$, XX of 20 trials failed.
 The robots have limited force, so increasing the object weight increases completion time.  
- Rectangles and squares tend to get stuck in the 90$^\circ$ workspace corners, and cause longer completion times than circles and hexagons.
+Each shape was designed to have the same mass and area.
+ Rectangles and squares tend to get stuck in the 90$^\circ$ workspace corners, and cause longer completion times than circles, triangles, and hexagons.
 

### 2016-08-09 journalSwarmControl/realExperiment.tex
added captions

@@
 
-The objects were 3D printed from ABS plastic.
-Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}.
+The Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}. 
+The objects were 3D printed from ABS plastic with a paper overlay. 
+Shapes included a 325 cm$^2$ equilateral triangle, 
+%Small Equilateral Triangle 140 cm$^2$,
+ 324 cm$^2$ square,
+ 281 cm$^2$ hexagon,
+254 cm$^2$ circle, 
+and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle.
+
 \begin{figure*}
@@
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){a)}
+	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){A}
 	\end{overpic}\\
-	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(5,85){b)}
+	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
 	\end{overpic}
-	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(5,85){c)}
+	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
 	\end{overpic}
-
 \end{center}
@@
 \caption{\label{fig:setup}
-(Top)Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by four remotely triggered 30W LED and four 50W LED floodlights, with an overhead machine vision system.
-(Down left) A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.
-Here is a swarm of hardware robots (kilobots) that pushes a green object with different shapes toward the goal.
-(Down right) The potential fields view in hardware experiments.
-See video attachment~\cite{ShivaVideo2015}
-}
+(Top) Hardware platform. (Lower left) Four shapes used for hardware experiments. (Lower right) Visualization of potential field. }
 \end{figure*}
@@
 Trials were run until the object COM entered the goal region.
-The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} (minutes:seconds).
-This is a mean of \{42:38\} (minutes:seconds) and a standard deviation of \{12:51\}(minutes:seconds).
-Also there was a trial with the circular object for \{52:35 \} (minutes:seconds).
-Another trial with square like object ran for \{114:31 \} (minutes:seconds).
-We also tried triangular object for \{ XX:YY \} (minutes:seconds).
+The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} minutes.
+This is a mean of 42:38 and a standard deviation of 12:51 minutes.
+A circular object completed in 52:35 minutes. 
+A square object completed in 114:31 minutes. 
+A triangular object completed in XX:YY minutes.
 

### 2016-08-10 journalSwarmControl/conclusion.tex
fixed the conclusion and changed title (lowercase and)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \todo{rewrite}
 
     Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
-The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.
-We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures.
-We present an alternative.
+Their small size makes individual control and autonomy challenging, so currently these robots are steered by global control inputs such as magnetic fields or chemical gradients.
+To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
+Analysis of the game play results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback.
 
-This paper first examines the perils, pitfalls, and possibilities we discovered by launching \href{http://www.swarmcontrol.net}{SwarmControl.net}, an online game where players steer swarms of up to 500 robots to complete manipulation challenges.
-We record statistics from thousands of players, and use the game to explore aspects of large-population robot control.
-We present the game framework as a new, open-source tool for large-scale user experiments.
-One intriguing result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback.
-Inspired by human operators, this paper next investigates controllers that use only the mean and variance of a robot swarm.
-We prove that the mean position is controllable, then provide conditions under which variance is controllable.
-We next derive automatic controllers for these and a hybrid, hysteresis-based switching control to regulate the first two moments of the robot distribution.
-Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.
+Inspired by human operators, this paper designed controllers and controllability results using only the mean and variance of a robot swarm. 
+These controllers were used as primitives for an object manipulation task.
+All automatic controllers were implemented using 100 kilobots steered by the direction of a global light source.
+These experiments culminated in an object manipulation task in a workspace with obstacles.
 
-There are many avenues for future work.  Manipulation by large populations of robots has many open questions. We invite other collaborators to submit their own experiments to SwarmControl.net.
-Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, and targeted drug delivery in a vascular network.
+  Manipulation by large populations of robots has many open questions.
+We invite other collaborators to submit their own experiments for large-scale trials to \href{http://www.swarmcontrol.net}{SwarmControl.net}.
+Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, optimal-control, and targeted drug delivery in a vascular network.
 

### 2016-08-10 journalSwarmControl/experiment.tex
Added the new Images and plots. (Still some plots should change)
Added New Simulation
Deleted noise extra description
Added the newest plots for the newest data
A lot os shapes changed

@@
 
+\begin{figure}
+\begin{center}
+	\begin{overpic}[width=0.45\columnwidth]{AllShapes.pdf}\end{overpic}
+\end{center}
+\caption{\label{fig:Shapes} All the shapes that the simulations ran.
+}
+\end{figure}
+
 %\todo{add the timeout function: I added two lines.}
@@
 Each trial measured the time to deliver the object to the goal location.
-The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied each simulation step) with $W=10$.
+The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied each simulation step) with $W=5$.
 
@@
  Increasing noise increases completion time. 
- Large amounts of noise caused failures, defined as trials lasting longer than 1000s.  With $W=50$, XX of 20 trials failed.  With $W=100$, XX of 20 trials failed.
+ %Large amounts of noise caused failures, defined as trials lasting longer than 1000s.  With $W=50$, XX of 20 trials failed.  With $W=100$, XX of 20 trials failed.
 The robots have limited force, so increasing the object weight increases completion time.  
@@
 \renewcommand{\figwid}{0.5\columnwidth}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}\put(1,60){a)}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}\put(1,55){a)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}\put(1,60){b)}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}\put(1,55){b)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}\put(1,65){c)}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}\put(1,55){c)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}\put(1,65){d)}
+\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}\put(1,55){d)}
 
 \end{overpic}
-\vspace{-1em}
+\vspace{-0.5em}
 \caption{\label{fig:AutoVeryParam}Parameter sweep for a) number of robots, b) different noise values, c) object weight, and d) object shape.

### 2016-08-10 journalSwarmControl/realExperiment.tex
A lot os shapes changed
added wrap figure to 3 figures

@@
 In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and lights: four 50W LED floodlights  at the corners and four 30W LED floodlights on the sides of a 6 m square centered on the workspace and 1.5 m above the table.
-The lights were controlled using an Arduino Uno board connected to an 8 relay shield board.
+An Arduino Uno connected to an 8 relay shield controlled the lights.
 
@@
 
-The Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}. 
+The Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}. 
 The objects were 3D printed from ABS plastic with a paper overlay. 
@@
 254 cm$^2$ circle, 
-and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle.
+and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.
+
+%\begin{figure*}
+%\begin{center}
+%	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
+%	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){A}
+%	\end{overpic}\\
+%	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
+%	\end{overpic}
+%	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
+%	\end{overpic}
+%\end{center}
+%\caption{\label{fig:setup}
+%(Top) Hardware platform. (Lower left) Four shapes used for hardware experiments. (Lower right) Visualization of potential field. }
+%\end{figure*}
 
@@
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){A}
-	\end{overpic}\\
-	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
-	\end{overpic}
-	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
+	\begin{overpic}[width=0.5\columnwidth]{SetUp.pdf}%\put(1,75){A}
+	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
+	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
 	\end{overpic}
 \end{center}
-
 \caption{\label{fig:setup}
-(Top) Hardware platform. (Lower left) Four shapes used for hardware experiments. (Lower right) Visualization of potential field. }
+Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
 \end{figure*}
@@
 \begin{center}
-	\begin{overpic}[width=\columnwidth]{RealMean}\end{overpic}
+	\begin{overpic}[width=\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
 %	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}

### 2016-08-10 journalSwarmControl/theory.tex
added wrap figure to 3 figures

@@
 \end{algorithm}
+
 \begin{figure}
 \centering
-\begin{overpic}[scale=.3]{hysteresis.png}
-\put(39,21){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
-\put(39,5){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:hysteresis} Two states to control mean and variance of a robot swarm.
+\begin{overpic}[width = 0.5\columnwidth]{hysteresis.png}
+\put(33,21){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
+\put(33,5){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
+\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:hysteresis} (Left) Two states to control mean and variance of a robot swarm. 
+(Right) Switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The plot uses robot radius $r=1/10$.
 %\vspace{-2em}
@@
 
+%\begin{figure}
+%\centering
+%\begin{overpic}[width = 0.5\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
+%\vspace{-1em}
+%\caption{\label{fig:VarianceMinMaxBand} The switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density.
+The above plot uses robot radius $r=1/10$.
+%}\vspace{-1em}
+%\end{figure}
+
 A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
-The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:VarianceMinMaxBand}:
+The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:hysteresis}:
 \begin{align}
@@
 
-\begin{figure}
-\centering
-\begin{overpic}[width = 0.5\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:VarianceMinMaxBand} The switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density.
-The above plot uses robot radius $r=1/10$.
-}\vspace{-1em}
-\end{figure}
-
 %what images should I show here?

### 2016-08-10 journalSwarmControl/Introduction.tex
added captions
tried to unify terminology

@@
 %
-\subfloat[][Cary Control: Foraging]{\label{fig:Forage}
+\subfloat[][Vary Control: Foraging]{\label{fig:Forage}
 \begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
@@
  Our paper is organized as follows.
-After a discussion of related work in Section \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.
-We report the results of our online experiments in Section \ref{sec:expResults}.
-Inspired by these results, we prove that the mean and variance of a robot swarm are controllable in Section \ref{sec:theory}, and present automatic controllers in Section \ref{sec:simulation}.
-We conclude with implementations of these controllers on hardware robots and use them to complete an object manipulation task  with 100+ kilobots in Section \ref{sec:realExperiment}.
+After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment in \S \ref{sec:expMethods}.
+We report the results of our online experiments in \S \ref{sec:expResults}.
+Inspired by these results, we prove that the mean and variance of a robot swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
+We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.
+ We conclude with implementations of these controllers in on hardware robots and use them to complete an object manipulation task with 100+ kilobots in \S \ref{sec:realExperiment}.
 

### 2016-08-10 journalSwarmControl/expResults.tex
added wrap figure to 3 figures
tried to unify terminology

@@
 
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.5\columnwidth}
-\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
+%\begin{figure}[b!]
+%\renewcommand{\figwid}{0.5\columnwidth}
+%\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
+%\vspace{-1em}
+%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
+%%\vspace{-1em}
+%}
+%\end{figure}
+
+\begin{wrapfigure}{R}{0.5\textwidth}
+  \vspace{-20pt}
+  \begin{center}
+  \includegraphics[width=0.5\textwidth]{ResVaryVis.pdf}
+  \end{center}
   \vspace{-1em}
 \caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
-%\vspace{-1em}
+\vspace{-1em}
 }
-\end{figure}
+\end{wrapfigure}
 
@@
  \end{align*}
-Here $m_i,\psi_i$ are uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$, where $M$ is a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
+$m_i,\psi_i$ are uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ is a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
 

### 2016-08-10 journalSwarmControl/experiment.tex
added wrap figure to 3 figures
tried to unify terminology

@@
 To design the policy, we first discretize the environment. 
-In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.
-The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$. 
-The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \nabla \mathbf{M}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the object toward the goal location.
+In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$  shown in Fig.~\ref{fig:BFSGradient}.
+The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$, so we define the desired direction for the object as $\mathbf{D}(\mathbf{b}) = \nabla \mathbf{M}(\mathbf{b})$. 
+The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
 However, BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
@@
   Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
-   \begin{align} \nabla \mathbf{M_{MDP}}(x) = \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
+   \begin{align} \mathbf{D}(x) = \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
    \end{align}
@@
 (Middle) The control policy from value iteration.
-(Right) Value function and control policy for the hardware setup.
-The vision algorithm detects obstacles.
+(Right) The vision algorithm detects obstacles, used to produce the
+value function and control policy for the hardware setup. 
 %\vspace{-2em}
@@
 To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\nabla \mathbf{M}$ as shown in Fig.~\ref{fig:potentialField}.
+The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}$ as shown in Fig.~\ref{fig:potentialField}.
 \begin{align}
@@
 %\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
-\phi &=\cos^{-1}\left( \frac{ \nabla \mathbf{M} \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\nabla \mathbf{M} } \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
+\phi &=\cos^{-1}\left( \frac{ \mathbf{D} \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
 %\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
@@
 
-\begin{figure}
+\begin{wrapfigure}{R}{0.3\textwidth}
+  \vspace{-20pt}
   \begin{center}
-	\begin{overpic}[width=0.45\columnwidth]{AllShapes.pdf}\end{overpic}
+\begin{overpic}[width=0.3\columnwidth]{AllShapes.pdf}\end{overpic}
   \end{center}
-\caption{\label{fig:Shapes} All the shapes that the simulations ran.
+  \vspace{-1em}
+\caption{\label{fig:Shapes} Six equal-area objects were tested. %Six object shapes were tested.
+\vspace{-1em}
 }
-\end{figure}
+\end{wrapfigure}
 
@@
 %\State$r_p \gets k_3 r_b$  \Comment{pushing move}
-\State Calculate $\nabla \mathbf{M}(\mathbf{b})$  \Comment{policy for object at $\mathbf{b}$}
+\State Calculate $\mathbf{D}(\mathbf{b})$  \Comment{direction for object at $\mathbf{b}$}
 \State Apply \eqref{eq:potentialfield}   \Comment{potential field to push object in correct direction}

### 2016-08-10 journalSwarmControl/experiment.tex
fixed terminology

@@
 To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}$ as shown in Fig.~\ref{fig:potentialField}.
+The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}.
 \begin{align}
@@
 %\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
-\phi &=\cos^{-1}\left( \frac{ \mathbf{D} \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
+\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
 %\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\

### 2016-08-10 journalSwarmControl/experiment.tex
Changed Sections and u

@@
 
-This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in Section~\ref{sec:expResults}.  
+This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in \S~\ref{sec:expResults}.  
 The swarm must deliver the object to the goal region.
@@
  a large negative reward to states including obstacles, and a small negative reward to all other states.
-The reward function $r(x,u)$ is defined as
+The reward function $r(x,\mathbf{u})$ is defined as
 \begin{align}
@@
 \begin{array}{ll}
-     +100, &  \textrm{if } u \textrm{ leads to goal state}\\
-      -100, & \textrm{if } u \textrm{ leads to a state with an obstacle} \\
+     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
+      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
       -1, & \textrm{otherwise}\\
@@
 \end{align}
-where $u$ is the action and $x$ is the current state. 
+where $\mathbf{u}$ is the action and $x$ is the current state. 
   Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
-   \begin{align} \mathbf{D}(x) = \arg\max_u   [ r(x,u) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| u,x)].  \label{eq:OptimalPolicy}
+   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
    \end{align}
@@
 \text{for }&\text{$i=1$ to $N$ do} \nonumber \\
-&\hat{V} (x_j) = \gamma \max_u [r(x_j,u) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| u,x_j)] \label{eq:ValueIteration}\\
+&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration}\\
 \text{end}& \nonumber

### 2016-08-10 journalSwarmControl/relatedWork.tex
Changed Sections and u

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:RelatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Global-Control of Microrobots
%Block-Pushing and Compliant Manipulation
%Human-Swarm Interaction
%Microrobots can be produced in large numbers, but are generally controlled by global inputs.  This leads to challenges in sensing, 
This section describes global control challenges and reviews highlights of human-swarm interaction, block pushing, and compliant manipulation.

\subsection{Global control of microrobots}
Small robots have been constructed with physical heterogeneity so that they respond differently to a global, broadcast control signal.  Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};   magnetic structures  with different cross-sections that could be independently steered by \cite{Floyd2011,Diller2013};   \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
\cite{Tottori2012} and \cite{Peyer2013}. 
%In our previous work with robots that can be modeled as nonholonomic unicycles~\cite{Becker2012k}, we showed that an inhomogeneity in turning speed is enough to make even an infinite number of robots controllable with regard to position. 
%While these approaches theoretically scale to large populations, they also require both excellent state estimation and heterogeneous robots.

%For micro- and nanorobotics, the marginal cost of producing one additional robot is astonishingly small.  While some microrobots are individually fabricated \cite{Tottori2012}, others are fabricated using MEMS techniques \cite{Donald2008}, and these dies could be tiled to produce multiple copies.  Nanocars are exemplary of this diminishing marginal cost.  Nanocars are synthetic molecules with integrated axles, rolling wheels, and light-driven motors developed by Tour et al.\cite{Chiang2011}.  These  are routinely produced in quantities that are amazing---a batch  the size of an aspirin tablet  contained$~\approx4\times 10^{19}$ nanocars.  This dwarfs the total number of birds on the planet earth---some $3\times10^{11}$\cite{Gaston1997}. 
 %aspirin in US weighs 325 mg.
  %or the number of automobiles produced in the history of humankind
% http://www.virlab.virginia.edu/nanoscience_class/lecture_notes/Lecture_13_Materials/Tour%20on%20nanovehicles.pdf states:
%	On the other hand, nanocars synthesized in our laboratory
%	are each calculated to be approximately 3x6x4 nm in size, and
%	can be produced on the 30 mg scale (3.26x10^18 nanocars)
%	using small 100 ml laboratory reaction flasks, or more
%	nanocars than the number of automobiles made in the
%	history of the world (63 million automobiles were produced
%	in 2005). It would take 30 nanocars, side by side, to span the
%	90 nm width of a small line in the most advanced logic
%	chip being made today
% The motorized nanocar can be synthesized in 12 steps with an overall yield of 5\%,  An electrically driven one is \cite{Kudernac2011}
%Biological agents, including the magnetically controlled ciliate protozoa studied by A. Julius~\cite{Ou2012, Ou2012a} and the magnetotactic bacteria studied by Martel~\cite{Felfoul2011}can be grown or purchased in large numbers.
% According to a scientific study of trace elements in hair, a single strand of hair 4 1/2 inches in length weighs, on average, 0.62 milligram, with weights varying from 0.25 milligram to just under 1 milligram.
%mosquito weight = and weight up to 2.5 milligrams
%Read more: How much does one strand of human hair weigh? | Answerbag http://www.answerbag.com/q_view/2159502#ixzz2EUltJTAl

%
%Consider a collection of $n$ planar robots.  We  describe the configuration of the $i$-th robot by $q_i=[x_i,y_i]^\top$ and its configuration space by $\Cspace = \R^2$.  At the micro/nanoscale viscous effects dominate inertial forces~\cite{Purcell1977}, so our model ignores velocity and assumes in the absence of an external force, the robots stop moving.
% We describe the configuration of the environment by the magnitude of the force field $f = [f_x,f_y]\in\R^2$, which is assumed constant across the workspace.  The control inputs $u = [u_x,u_y]\in\R^2$ change this force field.
%\begin{align}
%\dot{x}_i &= f_x,&
%\dot{f_x} &= u_x,\nonumber \\
%\dot{y}_i &= f_y, &
%\dot{f_y} &= u_y. \label{eq:sysModel}
%\end{align}
%
%\todo{This model is key, because if we use impulsive inputs for $u_x$ and $u_y$, there are a number of tasks we can accomplish with a single input (e.g. 2,3,5,7 in Fig.~\ref{fig:boundaryManipulators}.)}

%In \cite{Becker2012} we proved that systems described by \eqref{eq:ensembleDynamics} can be approximately steered to arbitrary positions in $\R^2$.  In \cite{Becker2012k} we provided closed-loop controllers to drive $n$ robots to $n$ target positions.  For both works, the control laws apply even to an infinite robot continuum.  
%
%Unfortunately, in \cite{Becker2012}, we proved that the heading of the ensemble---the $\theta_i$ values---is not controllable.  The final heading of every robot in the ensemble is the scaled integration of all the turning control inputs, i.e.~ $\theta_i(T) = \theta_i(0) + \epsilon_i\int_0^T \omega(t)\,{\rm d}t$.  Instead of a full-state controller, we provided a control law that steered the position of each robot's center of rotation.  Our goal is to provide control algorithms for many robots with uniform inputs, with  application in trajectory following, manipulation, force-closure, and assembly at the micro- and nanoscale.  Regrettably, 
% being unable to control the heading of the robots makes many tasks difficult or impossible.  For instance, manipulation and assembly work without heading control requires that the robots be able to rotate in place and be circular, since we cannot specify the incident angle of the robots. %In practice, this limits the application of our work to circular or point robots. 
%As Lynch explained in \cite{Lynch1999}, pushing with a round object is inherently unstable.

%  Possible applications include using molecular robots as nanoscale transporters, to break chemical bonds, or to build structures by constructing non-covalent bonds. We derive inspiration from the molecular actuators of Roth et al.~\cite{Minett2002}, the molecular elevators of Badj{\'i}c at al.~\cite{Badjic2004}, and the nanocars of Tour et al.~\cite{Vives2009}. %hydrogen bonds

% From James Tour:
%My motivation has been clear on this topic, and I have written it many times.  Here is a synopsis:
%Transport of goods and materials between points is at the
%heart of all engineering and construction in real-world systems.
%Just as biological systems survive by nanometer-scale
%transport using molecular-sized entities, as we delve into the
%arena of the nanosized world, it beckons that we learn to
%manipulate and transport nanometer-scale materials in a similar
%manner. Nanoscale transporters that are truly moleculesized
%will be required for the enzyme-like fabrication of
%sophisticated integrated structures.
%We want to build things from the bottom up, ex vivo, much like enzyme build in vivo???that is and has been my motivation.  I do not think that targeting these for biological applications is wise.  One would use enzyme-like structures for that, not nanocars.  These are for ex vivo.

Similarly, our previous work, \cite{Becker2012,Becker2012k}, focused on exploiting inhomogeneity between robots.  These control algorithms theoretically apply to any number of robots---even robotic continuums---but in practice process noise cancels the differentiating effects of inhomogeneity for more than tens of robots.  We desire control algorithms that extend to many thousands of robots.

%  \subsection{Three challenges for massive manipulation}
 While it is now possible to create many microrobots, there remain challenges in control, sensing, and computation:

 \paragraph{Control---global inputs:}
 Many micro- and nanorobotic systems, see \cite{Tottori2012,Shirai2005,Chiang2011,Donald2006,Donald2008,Takahashi2006,Floyd2011,Diller2013,Frutiger2008,Peyer2013}
   rely on global inputs, where each robot receives an exact copy of the control signal.  Our experiments follow this global model.
   %Two reasonable questions are ``What tasks are possible with many robots, all under uniform control inputs?'' and ``What tasks are impossible with many robots, all under uniform control inputs?''

 \paragraph{Sensing---large populations:}
 Parallel control of $n$ differential-drive robots in a 2D workspace requires $3n$ state variables. Even holonomic robots require $2n$ state variables. Numerous methods exist for measuring this state in microrobotics.  These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.  These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.   We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.

\paragraph{Computation---calculating the control law:}
In our previous work the controllers required at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
These operations become intractable for large populations of robots. By focusing on \emph{human} control of large robot populations, we accentuate computational difficulties because the controllers are implemented by the unaided human operator. % We present an approach that, for many tasks, bypasses these problems due to large populations by allowing the user to command the entire population as a single unit.  For position control we cannot bypass this problem, but we provide an algorithm that scales linearly in the number of robots.

 \subsection{Human-swarm interaction}
\cite{Jr2004} studied human \emph{fanout}, the number of robots a single human user could control.  %Studied as a key aspect of human-robot interaction, 
They postulated that the optimal number of robots was approximately  the autonomous time  divided by the interaction time required by each robot.  Their sample problem involved a multi-robot search task, where users could assign goals to robots.  Their user interaction studies with simulated planar robots  indicated a \emph{fanout plateau} of about 8 robots, after which there were diminishing returns.   They hypothesize that the location of this plateau is highly dependent on the underlying task.  Indeed, our paper indicates there are some tasks without plateaus. % (Lemmings is an example of this as well, for systems with autonomy
Their research investigated robots with 3 levels of autonomy.  We use robots without autonomy, corresponding with their first-level robots.

% Removed this -- 
%Chen, Barnes, and Harper-Sciarini published a review of supervisory control but emphasized high levels of autonomy, using the 10-level taxonomy given by \cite{Parasuraman2000}.  The direct control techniques this paper examines are the lowest level.

\cite{Squire:2006:HCM:1121241.1121248} designed experiments showing that user-interface design had a high impact on the task effectiveness and the number of robots that could be controlled simultaneously in a multi-robot task.

A number of user studies compare methods for controlling large swarms of simulated robots, for example \cite{bashyal2008human,kolling2012towards,de2012controllability}.  These studies provide insights but are limited by cost to small user studies; have a closed-source code base; and focus on controlling intelligent, programmable agents.  
For instance, the studies \cite{bashyal2008human},  \cite{de2012controllability}, and \cite{kolling2012towards}  were limited to a pool of 5, 18, and 32  participants.
	Using an online testing environment, we conduct similar studies but with sample sizes three orders of magnitude larger.

\subsection{Block pushing and compliant manipulation}
Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object, as in \cite{Sudsang2002,Fink2007}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object. A key difference is that our robots are compliant and tend to flow around the object, making this similar to fluidic trapping as in \cite{Armani2006} and \cite{Becker2009}.

Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators, e.g. \cite{odhner2014compliant,deimel2014novel}.  Like these manipulators, our swarms conform to the object to be manipulated.  However our swarms lack the restoring force provided by flexures in \cite{odhner2014compliant} and the silicone in \cite{deimel2014novel}.  Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},
 to regroup the swarms.

%
%\subsection{Nonprehensile manipulation}
%
%In \emph{nonprehensile manipulation}, a robot affects its environment without grasping objects by pushing and pulling \cite{Lynch1999,Goemans2006,Vose2009a}. In some ways, our problem formulation formulation is the inverse of nonprehensile manipulation. Rather than just use a robot to restructure the environment, we use the environment to restructure a population of robots.
%
%We can also use a large population of robots for traditional nonprehensile tasks, such as transporting objects using the flow of the robots \cite{Sugawara2012}, and manipulating an object too heavy for a single robot. Our control formulation enables efficiently controlling this kind of transport.
%
%%%%
%%OLD CONTENT
%%%%
%
%\subsection{Micro- and Nanoscale Manipulators}
%Micro/nanomanipulation and assembly are the focus of considerable micro and nanorobotics research. Sitti, Yu, and Cecil provide surveys of nanomanipulation and assembly \cite{Sitti2001,Yu2003,Cecil2007}.  Savia and Koivo focus on a survey of contact strategies for micro manipulation \cite{Savia2009}.
% Micro/nanomanipulation refers to manipulating components at the micro- and nanoscale, while assembly describes building structures from smaller components.  The ability to control position and track trajectories enables an ensemble of micro- and nanorobots to be used as a manipulator.  One advantage of using an ensemble control method over other methods, e.g.\ an atomic force microscope tip, is that ensemble methods allows the simultaneous manipulation of  multiple components.
%
%We draw particular inspiration from our colleagues James Tour, Stephan Link, and their associates at Rice University, and their work with nanocars.  They have successfully synthesized a variety of molecular vehicles, or species of nanocars \cite{Vives2009,Vives2009a,Vives2009b,Sasaki2008,Sasaki2008a,Sasaki2008b,Morin2007,Shirai2006,Shirai2006a}.  At the nanoscale simply visualizing the molecules in real-time--key for closed-loop control--has required significant effort \cite{Claytor2009,Khatua2009}. 
%Control thus far is rudimentary, designing nanocars that can either move straight \cite{Shirai2005} or turn \cite{Sasaki2008c} as a function of temperature.  More complex species incorporate a controllable motor, both thermally driven \cite{Morin2006} and light-driven  \cite{Chiang2011}.
%  Due to the challenges and research timescales involved with fabricating and visualizing nanocars, our \href{http://pubs.acs.org/doi/suppl/10.1021/ct7002594/suppl_file/ct7002594-file002.avi}{simulations and game models} will build on \href{http://pubs.acs.org/doi/suppl/10.1021/ct7002594/suppl_file/ct7002594-file001.avi}{molecular dynamics models} by Kolomeisky et al.~\cite{Kimov2008,Konyukhov2010,Akimov2012}.
%%simulation videos:  http://pubs.acs.org/doi/suppl/10.1021/ct7002594/suppl_file/ct7002594-file001.avi and http://pubs.acs.org/doi/suppl/10.1021/ct7002594/suppl_file/ct7002594-file002.avi
% 
% 
%%\subsection{Uniform Control Inputs}
%%Our approach \cite{Becker2012a,Becker2012k,Becker2013} is based on the application of ensemble control, which we used in previous work to derive an approximate (open-loop) steering algorithm for a nonholonomic unicycle despite model perturbation (e.g., unknown wheel size) that scales both the forward speed and turning rate by an unknown but bounded constant \cite{Becker2012}. Rather than steer one unicycle with an unknown parameter, we chose to steer an infinite collection of unicycles, each with a particular value of this parameter in some bounded set. Following the terminology introduced by \cite{Brockett1999,Khaneja2000,Li2009,Li2011}, we called this fictitious collection of unicycles an {\em ensemble} and called our approach to steering {\em ensemble control}. The idea was that if the same control inputs steered the entire ensemble from start to goal, then surely they would steer the particular unicycle of interest from start to goal, regardless of its wheel size.
%%
%%Here, we take advantage of this idea in a slightly different way. Rather than trying to mitigate the effects of bounded model perturbation (i.e., of inhomogeneity), we are trying to exaggerate these effects. Basic controllability results carry over from our previous work. Our main contribution in this paper is to derive a closed-loop feedback policy that guarantees exact asymptotic convergence of the ensemble to any given position. We note that, for single robots, it is possible to build a robust feedback controller that regulates position and orientation \cite{Lucibello2001}. It is not obvious that the same can be done for an infinite collection of robots.
%
%\subsection{Robotic Manipulation}
%There are many challenges to conventional approaches to this problem. Classic form- and force-closure techniques require global
%knowledge about the geometry of the component and the configuration of the robots and their actuators.
%Robotics has derived considerable inspiration from collective transport in social insects \cite{E.-Bonabeau1999, Moffett1988, Berman2010}. 
%Decentralized transport strategies have been developed using inspiration from social insect research or through evolutionary computation \cite{Su2008, Gross2009, Gros2006, Montemayor2005, Mellinger2010, Lindsey2010, Berman2010}. 
%
%
%%\paragraph{gripping:}
%%In this section we primarily focus on transport strategies for autonomous mobile robots that use grippers to move objects, and where collective behavior is required to move objects too heavy for a single robot.
%%Many approaches to this task rely on centralized planners or leader-follower schemes, that require knowledge of
%%the object geometry to optimally position robots for movement \cite{Lindsey2010, Mellinger2010, Su2008}; some decentralized versions of these schemes have also been developed \cite{Montemayor2005}. 
%%Much of this work focuses on a slightly different task called planar manipulation, i.e. controlling object trajectory and orientation for lightweight objects. 
%%An advantage of these control theory based methods is that they are analytically tractable and generalize to complex object shapes.
%%A disadvantage of these strategies is that they often rely on robots knowing object geometry, position, and orientation at all times as well as position of all other team members. 
%%This requires sophisticated robots that can estimate object geometry and track absolute object position and orientation at all times; many schemes also require global communication. 
%%These requirements limit the applicability and scalability of the strategies.
%%A complementary approach has been pursued by swarm robotics research, where decentralized transport strategies have been developed using inspiration from social insect research or through evolutionary computation \cite{Gross2009, Gros2006}. 
%%These algorithms rely on simple local sensing by individual robots, with no explicit knowledge of object shape or explicit coordination between robots. 
%%An advantage of these methods is that they rely on simple robot capabilities, and have been experimentally tested on platforms such as the Swarmbots
%%\cite{Gross2009, Gros2006}.
%%This research has also pursued interesting extensions, such as the ability of robots to form pulling chains to transport an object. 
%%A disadvantage of these methods is that they have not been theoretically analyzed, and evaluation is done only in simulation or through robot experiments. 
%%It is unclear whether the approaches generalize to more complex object geometries and how well the approaches scale for large numbers of robots. 
%%In general, for both types of prior research, experimental verification of the transport strategy has been limited to small numbers of robots (2-6) and simple object shapes (e.g. circles, squares, ``L'' shapes).
%
%Non-prehensile manipulation that maneuvers a component without grasping it, see reviews in \cite{Mason1999, Peshkin1990}.  This approach can eliminate the need for specialized grippers. Two forms of non-prehensile manipulation are pushing and impulsive.
%\paragraph{Manipulation By Pushing:} enjoys a long history, with contributions from Lynch et al.~\cite{Lynch1996,Lynch1999,Bernheisel2006}, Mason et al.~\cite{Akella1999,Akella2000,Akella2000a,Erdmann1988,Mason2001}, Spong et al. \cite{Partridge2000,Spong2001}, and Rus et al.~\cite{rus_moving_1995,donald_analyzing_1994}.  For multi-robot manipulation under uniform inputs, we cannot guarantee robots will remain in contact with the component, so this proposal will focus on other manipulation modalities.
%
%\paragraph{Impulsive Manipulation:} manipulation by striking an component and letting it slide.  Mason et. al~ studied \emph{the inverse sliding problem}, determining the velocities required to send an component to a desired configuration, and \emph{the impact problem}, determining how to strike the component in order to achieve those velocities \cite{Huang1995,Huang1996,Huang1998}.   Excellent reviews of the general theory of impulsive control, e.g. control of a system were at least one state variable can be changed by discrete  ``impulsive'' inputs  are found in \cite{Yang1999,Yang2001}.  % \cite{Gravagne2000}
%In robotics, impulsive manipulation has been studied for vibratory parts feeding, where parts lying on a plate are repeatedly struck as the plate vibrates \cite{Bohringer1995,Reznik1998, Bohringer2000a, Vose2007, Vose2008a,Vose2009,Vose2009a}, and for swinging impact \cite{Mochiyama2007}.
%
%Our proposed contribution is to use large populations of robots to provide the impulsive inputs, under the strong constraint of uniform inputs.
%
%
%\paragraph{Caging Manipulation:}
%A conceptually simple form for manipulation is to grasp a component by encircling it with robots in such a way that the grasp can resist any external force applied to the component. 
% Rimon and Blake introduced caging to robotics \cite{Rimon1996} for non-convex objects and two fingered gripers. Similar centralized planners include Davidson and Blake \cite{Davidson1998}, Ponce et al.~\cite{Sudsang1998,Sudsang1999,Sudsang2000,Sudsang2002}, and Wang and Kumar \cite{Wang2002}.  Kumar et al.~extended this work for decentralized control \cite{Pereira2004} and obstacle avoidance  \cite{Fink2007,Fink2008}
%
%Because ensemble control enables trajectory tracking, all the centralized methods \cite{Rimon1996,Davidson1998,Sudsang1998,Sudsang1999,Sudsang2000,Sudsang2002,Wang2002}
%can be realized by robots with uniform inputs and the decentralized techniques can be emulated.  
%However, ensemble trajectory control is slow, so our proposed contribution will introduce new techniques that exploit the large numbers of potential manipulators.
%
%\paragraph{Towing Manipulation:}
% tethering multiple robots to a component avoids difficulties relating to gripping and maintaining contact with the component.  Robotic manipulation has been studied by Donald and Rus et al.~\cite{Donald2000}, Esposito~\cite{esposito_tugboat_towing_2010}, and with flying robots by Kumar et al.~\cite{Fink2011,Lenarcic2010}.
%One drawback is that towing solutions do not allow the robots to push and pull on the component.  By substituting a spring with a non-zero resting length, our technical approach will implement both towing and pushing with the same mechanism.
%
%\paragraph{Compliant Manipulation:}
% In standard robot arms, contact can be damaging with stiff actuators, even with joint torque sensors (Khatib et al.~\cite{Khatib1986} and  Luo \cite{Luo1993}).
% On macro-scale robotics one solution is the compliant manipulator we presented in~\cite{KongEtAl-MRM-ICRA11}. % shown in Figure~\ref{fig:MRM}
% Its series-elastic actuators (SEA)~\cite{pratt_williamson_1995} provide compliance, safe contact between components and robots, and force sensing.  We will use this setup to measure forces during hardware verification of our algorithms, and implement a miniaturized version on our r-one robots.
%%A general framework for cobot control (Peshkin) \cite{Gillespie2001}, Khatib.

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/MDPgridworldFunction.m
Added All the commented matlab code to the github

@@ -1,0 +1,116 @@
+function [V_hat_prob, DX,DY]  = MDPgridworldExample(map, goalX, goalY)
+%  Applies value iteration to learn a policy for a Markov Decision Process
+%  (MDP) -- a robot in a grid world.
+% 
+% The world is freespaces (0) or obstacles (1). Each turn the robot can
+% move in 8 directions, or stay in place. A reward function gives one
+% freespace, the goal location, a high reward. All other freespaces have a
+% small penalty, and obstacles have a large negative reward. Value
+% iteration is used to learn an optimal 'policy', a function that assigns a
+% control input to every possible location.
+% 
+% This function compares a deterministic robot, one that always executes
+% movements perfectly, with a stochastic robot, that has a small
+% probability of moving +/-45degrees from the commanded move.  The optimal
+% policy for a stochastic robot avoids narrow passages and tries to move to
+% the center of corridors.
+% 
+% From Chapter 14 in 'Probabilistic Robotics', ISBN-13: 978-0262201629,
+% http://www.probabilistic-robotics.org
+% 
+%  Aaron Becker, March 11, 2015, Shiva Shahrokhi Modified Summer 2016.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+set(0,'DefaultAxesFontSize',18)
+format compact
+pauseOn = false;  %setting this to 'true' is useful for teaching, because it pauses between each graph
+World = map;
+
+
+colormap(gray)
+R = -100*ones(size(World)); %-100 reward for obstacles
+R(World==0) = -1; %small negative reward for not being at the goal
+R(goalX,goalY) = 100; %goal state has big reward
+
+
+[V_hat_prob, DX, DY] = MDP_discrete_value_iteration(R,World,true);
+
+    function[V_hat, DX,DY] = MDP_discrete_value_iteration(R,World,prob)
+        % iterates on the value function approximation V_hat until the V_hat converges.
+        V_hat_prev = zeros(size(World));
+        V_hat = -100*ones(size(World));
+        V_hat(World==0) = R(World==0);
+        colormap(jet)
+        gamma = 0.97;
+        
+        xIND = find(World == 0);
+        iteration = 0;
+        
+        [X,Y] = meshgrid(1:size(World,2),1:size(World,1));
+        DX = zeros(size(X));
+        DY = zeros(size(Y));
+        figure(10);
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ObjectManipulation_kilobots.m

@@ -1,0 +1,356 @@
+%% Object Manipulation Experiment With Kilobots
+% In this code we use arduino and our vision system to control a swarm of
+% kilobots to push an object through a maze.
+% By Shiva Shahrokhi, Lillian Lin and Mable Wan, Summer 2016
+
+close all
+clear all
+
+%% Load maps from RegionCode.m
+load('MDPShot', 'movesX', 'movesY','corners');
+load('ThresholdMapsMac','transferRegion','mainRegion'); 
+
+
+%% Define webcam
+webcamShot = true;
+
+if webcamShot
+    cam = webcam(1);
+    %% Using Arduino for our lights, this is how we define arduino in Matlab:
+    if (ispc==1)  
+        a = arduino('Com5','uno');
+    else 
+        a = arduino('/dev/tty.usbmodem1421','uno');
+    end 
+    %% Setup End Goal Position
+    if (ispc==1)  
+        goalX = 5;
+        goalY = 5;
+        goalSize = 4;
+    else 
+        goalX = 5;
+        goalY = 5;
+        goalSize = 3.5;
+    end 
+end
+%% Initalize Variables
+Relay=0;
+delayTime = 10;
+VarCont = false;
+flowDebug = true;
+success = false;
+again = true;
+counter = 1;
+c = 0;
+%Flow Around Variables
+rhoNot=7.5;
+%1 is main regions, 0 is transfer regions
+regionID=1;
+regionNum=3;
+%load regions
+currentRegionMap=mainRegion(:,:,regionNum); %init map
+
+while success == false  
+    %% Read in a webcam snapshot.
+    
+    if webcamShot
+        if (again == true)
+            relayOn(a,0);
+            pause (10);
+        end 
+        rgbIm = snapshot(cam);
+    else
+        rgbIm = imread('PC.jpeg');
+        goalX = 5;
+        goalY = 5;
+        goalSize = 4;
+    end
+%% crop to have just the table view.
+    if (ispc== 1)  
+        originalImage = imcrop(rgbIm,[50 10 500 400]);
+    else 
+        originalImage = imcrop(rgbIm,[345 60 1110 850]);
+    end 
+    s = size(originalImage);
+    scale = floor( size(originalImage,2)/size(mainRegion,2));
+    epsilon = 1* scale;
+    sizeOfMap = floor(s/scale);
+    imshow(originalImage);
+
+%% make HSV scale for robots and object
+    I = rgb2hsv(originalImage);
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ReigonCode.m

@@ -1,0 +1,237 @@
+%% REIGONCODE -- preprocessing for block pushing expirament
+% This code takes a snapshot from the webcam, then processes it to find 
+% obstacles. From these obstacles, the corners, regions, and a policy for MDP 
+% is calculated. It then draws the gradients, regions, and shows the grid
+% view to the user.
+% By Shiva Shahrokhi and Lillian Lin Summer 2016
+close all
+clear all
+
+success = false;
+webcamShot = true;
+obstacles = [];
+if webcamShot
+    cam = webcam(1);
+end
+t0 = tic;
+results3= [];
+if webcamShot
+ originalImage = snapshot(cam);
+     if (ispc)  
+         original = imcrop(originalImage,[50 10 500 400]);
+         img = imcrop(originalImage,[50 10 500 400]);
+         rgbIm = imcrop(originalImage,[50 10 500 400]);
+    else 
+        original = imcrop(originalImage,[345 60 1110 860]);
+         img = imcrop(originalImage,[345 60 1110 860]);
+         rgbIm = imcrop(originalImage,[345 60 1110 860]);
+     end 
+else 
+    rgbIm = imread('PC.jpeg');
+end
+
+%% Obstacle Thresholds
+    I = rgb2hsv(rgbIm);
+    
+    % Define thresholds for channel 1 based on histogram settings
+    channel1Min = 0.862;
+    channel1Max = 0.945;
+
+    % Define thresholds for channel 2 based on histogram settings
+    channel2Min = 0.272;
+    channel2Max = 1.000;
+
+    % Define thresholds for channel 3 based on histogram settings
+    channel3Min = 0.000;
+    channel3Max = 1.000;
+
+% Create mask based on chosen histogram thresholds
+BW = (I(:,:,1) >= channel1Min ) & (I(:,:,1) <= channel1Max) & ...
+    (I(:,:,2) >= channel2Min ) & (I(:,:,2) <= channel2Max) & ...
+    (I(:,:,3) >= channel3Min ) & (I(:,:,3) <= channel3Max);
+
+%% finding Obstacles
+    [B,L] = bwboundaries(BW, 'noholes');
+    stat = regionprops(L,'Centroid','Orientation','MajorAxisLength', 'Area');
+    area = cat(1,stat.Area);
+    centroids = cat(1, stat.Centroid);
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/MDPgridworldFunction.m
had some fun with the code

@@ -1,29 +1,29 @@
-function [V_hat_prob, DX,DY]  = MDPgridworldExample(map, goalX, goalY)
+function [V_hat_prob, DX,DY]  = MDPgridworldFunction(map, goalX, goalY)
 %  Applies value iteration to learn a policy for a Markov Decision Process
 %  (MDP) -- a robot in a grid world.
-% 
+%
 % The world is freespaces (0) or obstacles (1). Each turn the robot can
 % move in 8 directions, or stay in place. A reward function gives one
 % freespace, the goal location, a high reward. All other freespaces have a
 % small penalty, and obstacles have a large negative reward. Value
 % iteration is used to learn an optimal 'policy', a function that assigns a
 % control input to every possible location.
-% 
+%
 % This function compares a deterministic robot, one that always executes
 % movements perfectly, with a stochastic robot, that has a small
 % probability of moving +/-45degrees from the commanded move.  The optimal
 % policy for a stochastic robot avoids narrow passages and tries to move to
 % the center of corridors.
-% 
+%
 % From Chapter 14 in 'Probabilistic Robotics', ISBN-13: 978-0262201629,
 % http://www.probabilistic-robotics.org
-% 
+%
 %  Aaron Becker, March 11, 2015, Shiva Shahrokhi Modified Summer 2016.
-
+%   atbecker@uh.edu,  Shiva Shahrokhi <shiva.shahrokhi@gmail.com>
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 set(0,'DefaultAxesFontSize',18)
 format compact
-pauseOn = false;  %setting this to 'true' is useful for teaching, because it pauses between each graph
+pauseOn = false;  %#ok<NASGU> %setting this to 'true' is useful for teaching, because it pauses between each graph
 World = map;
 
 
@@ -55,9 +55,9 @@ R(goalX,goalY) = 100; %goal state has big reward
         axis tight
         set(gca,'Xtick',[], 'Ytick',[])
         iteration_limit = 200; %value function needs ~600 iterations to converge, but the policy converges after ~100 iterations
-        while ~isequal(V_hat,V_hat_prev) && iteration < iteration_limit 
+        while ~isequal(V_hat,V_hat_prev) && iteration < iteration_limit
             V_hat_prev = V_hat;
-            for i = 1:numel(xIND)  
+            for i = 1:numel(xIND)
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ObjectManipulationKilobots.m

@@ -1,12 +1,19 @@
 %% Object Manipulation Experiment With Kilobots
-% In this code we use arduino and our vision system to control a swarm of
+% Uses an overhead vision system to control a swarm of
 % kilobots to push an object through a maze.
+% Swarm is attracted to the brightest light in the room
+% Lights are controlled by an Arduino relay-sheild
+%
+% requires RegionCode.m and MDPgridworldFunction.m
+% -------------------------
 % By Shiva Shahrokhi, Lillian Lin and Mable Wan, Summer 2016
-
+format compact
 close all
 clear all
 
 %% Load maps from RegionCode.m
+RegionCode
+
 load('MDPShot', 'movesX', 'movesY','corners');
 load('ThresholdMapsMac','transferRegion','mainRegion'); 
 
@@ -60,7 +67,7 @@ while success == false
         end 
         rgbIm = snapshot(cam);
     else
-        rgbIm = imread('PC.jpeg');
+        rgbIm = imread('PC.jpeg'); %#ok<UNRCH>
         goalX = 5;
         goalY = 5;
         goalSize = 4;
@@ -169,8 +176,7 @@ while success == false
             sumX= sumX+centers(i,1);
             sumY= sumY+centers(i,2);
             countMean=countMean+1;
-            cenArray(countMean,1)= centers(i,1); %adding selected robots to array
-            cenArray(countMean,2)= centers(i,2);
+            cenArray(countMean,:) = centers(i,:); %#ok<SAGROW> %adding selected robots to array
         end
     end
     %% Robot Swarm Charactaristics
@@ -227,7 +233,7 @@ while success == false
             VarCont = false;
         end
         %% normal control
-        if ~VarCont & ~NumRobotCont
+        if ~VarCont && ~NumRobotCont
             ep = 5;
             minDistance =  5*scale;
 
@@ -239,7 +245,6 @@ while success == false
             if indOY ==1 || indOY == 0
                 indOY = 2;
             end
-
     % Flow Around Goal Algorithm
             alphaWant=atan2(movesX(indOY,indOX),movesY(indOY,indOX));
             
@@ -310,7 +315,7 @@ while success == false
             txt = int2str(i);
             text(corners(i,1)* scale,corners(i,2)*scale,txt,'HorizontalAlignment','right')
         end
-%%Current Mean and Covariance Ellipse
+%% Current Mean and Covariance Ellipse
         plot_gaussian_ellipsoid(M,C);
         counter = counter+1;
         

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/RegionCode.m

@@ -1,0 +1,246 @@
+%% REGIONCODE -- preprocessing for block pushing experiment
+% Takes a snapshot from the webcam, then processes it to find
+% obstacles. From these obstacles, the corners, regions, and a policy for MDP
+% is calculated. It then draws the gradients, regions, and shows the grid
+% view to the user.
+%
+% saves this data in two files:
+% 1.) MDPShot.mat:
+%    [movesX,movesY]  policy in X and Y  directions
+%    corners,  xy locations of corners in workspace
+% 2.) ThresholdMapsMac.mat:
+%     transferRegion: 3D array. transferRegion(i,:,:) gives a logical
+%                   array for transfer region i
+%     mainRegion:    3D array. mainRegion(i,:,:) gives a logical
+%                   array for main region i
+%
+%  requires MDPgridworldFunction.m
+% ---------------------
+% By Shiva Shahrokhi and Lillian Lin,  Summer 2016
+% Shiva Shahrokhi <shiva.shahrokhi@gmail.com>
+close all
+clear all
+
+success = false;
+webcamShot = true;
+if webcamShot
+    cam = webcam(1);
+end
+t0 = tic;
+if webcamShot
+    originalImage = snapshot(cam);
+    if (ispc)
+        original = imcrop(originalImage,[50 10 500 400]);
+        img = imcrop(originalImage,[50 10 500 400]);
+        rgbIm = imcrop(originalImage,[50 10 500 400]);
+    else
+        original = imcrop(originalImage,[345 60 1110 860]);
+        img = imcrop(originalImage,[345 60 1110 860]);
+        rgbIm = imcrop(originalImage,[345 60 1110 860]);
+    end
+else
+    rgbIm = imread('PC.jpeg'); %#ok<UNRCH>
+end
+
+%% Obstacle Thresholds
+I = rgb2hsv(rgbIm);
+
+% Define thresholds for channel 1 based on histogram settings 
+% these were picked using the Color Thresholding tool
+channel1Min = 0.862;
+channel1Max = 0.945;
+
+% Define thresholds for channel 2 based on histogram settings
+channel2Min = 0.272;
+channel2Max = 1.000;
+
+% Define thresholds for channel 3 based on histogram settings
+channel3Min = 0.000;
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ReigonCode.m

@@ -1,237 +1,0 @@
-%% REIGONCODE -- preprocessing for block pushing expirament
-% This code takes a snapshot from the webcam, then processes it to find 
-% obstacles. From these obstacles, the corners, regions, and a policy for MDP 
-% is calculated. It then draws the gradients, regions, and shows the grid
-% view to the user.
-% By Shiva Shahrokhi and Lillian Lin Summer 2016
-close all
-clear all
-
-success = false;
-webcamShot = true;
-obstacles = [];
-if webcamShot
-    cam = webcam(1);
-end
-t0 = tic;
-results3= [];
-if webcamShot
- originalImage = snapshot(cam);
-     if (ispc)  
-         original = imcrop(originalImage,[50 10 500 400]);
-         img = imcrop(originalImage,[50 10 500 400]);
-         rgbIm = imcrop(originalImage,[50 10 500 400]);
-    else 
-        original = imcrop(originalImage,[345 60 1110 860]);
-         img = imcrop(originalImage,[345 60 1110 860]);
-         rgbIm = imcrop(originalImage,[345 60 1110 860]);
-     end 
-else 
-    rgbIm = imread('PC.jpeg');
-end
-
-%% Obstacle Thresholds
-    I = rgb2hsv(rgbIm);
-    
-    % Define thresholds for channel 1 based on histogram settings
-    channel1Min = 0.862;
-    channel1Max = 0.945;
-
-    % Define thresholds for channel 2 based on histogram settings
-    channel2Min = 0.272;
-    channel2Max = 1.000;
-
-    % Define thresholds for channel 3 based on histogram settings
-    channel3Min = 0.000;
-    channel3Max = 1.000;
-
-% Create mask based on chosen histogram thresholds
-BW = (I(:,:,1) >= channel1Min ) & (I(:,:,1) <= channel1Max) & ...
-    (I(:,:,2) >= channel2Min ) & (I(:,:,2) <= channel2Max) & ...
-    (I(:,:,3) >= channel3Min ) & (I(:,:,3) <= channel3Max);
-
-%% finding Obstacles
-    [B,L] = bwboundaries(BW, 'noholes');
-    stat = regionprops(L,'Centroid','Orientation','MajorAxisLength', 'Area');
-    area = cat(1,stat.Area);
-    centroids = cat(1, stat.Centroid);
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/AngleFix.m
Had to add functions for it to work.

@@ -1,0 +1,26 @@
+function [ angle ] = AngleFix( angle , threshold)
+% ANGLEFIX takes an angle and makes it an equivalent in between a threshold 
+%          in which on the the stated threshold is both the positive and 
+%          neagtive threshold. 
+%   ANGLEFIX(angle,threshold)
+%
+%       x = an angle you would like to change
+%       threshold = the edge values wanted for angles
+%
+%       if threshold = pi/2 or 90 then the angle created will be under the
+%       assumption that you have a half circle and the negative version of  
+%       an angle is the same as the positive version of an angle
+%       
+%       if threshold = pi or 180 then the angle created will be under the
+%       assumption that you have a full circle and the negative version of  
+%       an angle is not the same as the positive version of an angle
+
+    while (angle > threshold||angle<= -threshold)
+        if angle>threshold
+            angle=angle - 2*threshold;
+        elseif angle <= -threshold
+            angle=angle + 2*threshold;
+        end
+    end 
+
+end

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/FlowForce.m

@@ -1,0 +1,33 @@
+
+function [ currgoalX,currgoalY ] = FlowForce(RobotMeanX,RobotMeanY,AttPointX,AttPointY,RepPointX,RepPointY) 
+% FLOWFORCE calculates a vector for the potential field effect on a point
+%   FLOWFORCE(RobotMeanX,RobotMeanY,AttPointX,AttPointY,RepPointX,RepPointY)
+%
+%       RobotMeanX,RobotMeanY = here the robots currently are
+%       AttPointX,AttPointY = Where the robots should move towards
+%       RepPointX,RepPointY = Where the robots should be pushed from
+%
+%       eta and zeta are the ratio between the repulsive and attractive
+%       forces respectively.
+%       rhoNot is the area of effect for the repulsive fields
+%
+% By Lillian Lin Summer 2016
+    eta=50;
+    zeta=1;
+    rhoNot=7.5;
+
+    rho=dist2points(RepPointX,RepPointY,RobotMeanX,RobotMeanY);
+    if (rho<rhoNot)
+        FrepX=eta*((rho^(-1))-(rhoNot^(-1)))*(rho^(-1))^2*(RepPointX-RobotMeanX);
+        FrepY=eta*((rho^(-1))-(rhoNot^(-1)))*(rho^(-1))^2*(RepPointY-RobotMeanY);
+    else 
+       FrepX=0;
+       FrepY=0;
+    end
+    rho=dist2points(AttPointX,AttPointY,RobotMeanX,RobotMeanY);
+    FattX=zeta*(RobotMeanX-AttPointX)/rho;
+    FattY=zeta*(RobotMeanY-AttPointY)/rho;
+        
+    currgoalX=cos(atan2((-FrepY-FattY),(-FattX-FrepX)));
+    currgoalY=sin(atan2((-FrepY-FattY),(-FattX-FrepX)));
+end

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ObjectManipulationKilobots.m

@@ -12,9 +12,9 @@ close all
 clear all
 
 %% Load maps from RegionCode.m
-RegionCode
+%RegionCode
 
-load('MDPShot', 'movesX', 'movesY','corners');
+load('MazeMap', 'movesX', 'movesY','corners');
 load('ThresholdMapsMac','transferRegion','mainRegion'); 
 
 
@@ -202,7 +202,7 @@ while success == false
         %% Current region should at least have a minimum number of robots
         minNumRobD = 10;
         minNumRobU = 20;
-        if countMean<minNumRob
+        if countMean<minNumRobD
             NumRobotCont = true;
             % finding the nearest corner
             for i = 1:size(corners)

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/circle.m

@@ -1,0 +1,10 @@
+function circle(x,y,r)
+%x and y are the coordinates of the center of the circle
+%r is the radius of the circle
+%0.01 is the angle step, bigger values will draw the circle faster but
+%you might notice imperfections (not very smooth)
+ang=0:0.01:2*pi; 
+xp=r*cos(ang);
+yp=r*sin(ang);
+plot(x+xp,y+yp,'Markersize',16,'color','green','linewidth',3);
+end

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/dist2points.m

@@ -1,0 +1,14 @@
+function [ distance ] = dist2points(x1,y1,x2,y2)
+% DIST2POINTS calculates distance between two points
+%   DIST2POINTS(x1,y1,x2,y2)
+%       x1,y1 = start points
+%       x2,y2 = end points
+%
+% By Lillian Lin Summer 2016
+    xs=x2-x1;
+    ys=y2-y1;
+    xs=xs^2;
+    ys=ys^2;
+    distance=sqrt(xs+ys);
+end
+

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/relayOn.m

@@ -1,0 +1,120 @@
+function relayOn(a,which)
+% RELAYON uses an existing arduino connection to turn on the lights
+%   RELAYON(a,which)
+%
+%       which = the relay number you would like to turn on
+%       a = an existing arduino connection
+%
+% By Lillian Lin Summer 2016
+
+% We have 8 Relays.
+%west
+RELAY1 = 9;
+% northwest
+RELAY2 = 8;
+%north
+RELAY3 = 5;
+% northeast
+RELAY4 = 2;
+%east
+RELAY5 = 7;
+%southeast
+RELAY6 = 3;
+%south
+RELAY7 = 6;
+%southwest
+RELAY8 = 4;
+    switch which
+        case 1
+            writeDigitalPin(a,RELAY1,0);
+            writeDigitalPin(a,RELAY2,1);
+            writeDigitalPin(a,RELAY3,1);
+            writeDigitalPin(a,RELAY4,1);
+            writeDigitalPin(a,RELAY5,1);
+            writeDigitalPin(a,RELAY6,1);
+            writeDigitalPin(a,RELAY7,1);
+            writeDigitalPin(a,RELAY8,1);
+        case 2
+            writeDigitalPin(a,RELAY1,1);
+            writeDigitalPin(a,RELAY2,0);
+            writeDigitalPin(a,RELAY3,1);
+            writeDigitalPin(a,RELAY4,1);
+            writeDigitalPin(a,RELAY5,1);
+            writeDigitalPin(a,RELAY6,1);
+            writeDigitalPin(a,RELAY7,1);
+            writeDigitalPin(a,RELAY8,1);
+        case 3
+            writeDigitalPin(a,RELAY1,1);
+            writeDigitalPin(a,RELAY2,1);
+            writeDigitalPin(a,RELAY3,0);
+            writeDigitalPin(a,RELAY4,1);
+            writeDigitalPin(a,RELAY5,1);
+            writeDigitalPin(a,RELAY6,1);
+            writeDigitalPin(a,RELAY7,1);
+            writeDigitalPin(a,RELAY8,1);
+        case 4
+            writeDigitalPin(a,RELAY1,1);
+            writeDigitalPin(a,RELAY2,1);
+            writeDigitalPin(a,RELAY3,1);
+            writeDigitalPin(a,RELAY4,0);
+            writeDigitalPin(a,RELAY5,1);
+            writeDigitalPin(a,RELAY6,1);
[…]

### 2016-08-11 journalSwarmControl/code/ObjectManipulation/ObjectManipulationKilobots.m
Uncommented RegionCode

@@ -12,7 +12,7 @@ close all
 clear all
 
 %% Load maps from RegionCode.m
-%RegionCode
+RegionCode
 
 load('MazeMap', 'movesX', 'movesY','corners');
 load('ThresholdMapsMac','transferRegion','mainRegion'); 

### 2016-08-12 journalSwarmControl/Introduction.tex
updates through page 16

@@
 Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population.
-It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location.
-However, it is often possible to sense global properties of the group, such as mean position and density.
-Finally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of robots is a daunting human-swarm interaction (HSI) challenge.
+Additionally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of robots is a daunting human-swarm interaction (HSI) challenge.
 
@@
 Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
+There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
 This paper presents a tool for investigating HSI methods through statistically significant numbers of experiments.
-There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.
 
@@
 Our goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size.
-Towards this end, we created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}, an open-source online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
+Towards this end, we created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}, an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
 Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.
@@
 %Inspired by our online experiments, because it is not always possible to gather pose information on each robot for feedback control and
+%It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location.
+However, it is often possible to sense global properties of the group, such as mean position and density.
 Often robots are difficult or impossible to sense individually due to their size and location. 
@@
 Simulation at left, screen capture form hardware experiment at right.
-See video attachment~\cite{ShivaVideo2015}.}
+See multimedia extension.}
 \end{figure}
@@
 
- This paper is the synthesis of two preliminary conference papers, the first, \cite{swarmcontrol2013}, covering first few months of SwarmControl.net experiments, and the second, \cite{ShahrokhiIROS2015}, with simulations of object manipulation.
+ This paper is the synthesis of two preliminary conference papers,  \cite{swarmcontrol2013}, covered first few months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015}, presented simulations of object manipulation.
 This paper presents the final results from SwarmControl.net.

### 2016-08-12 journalSwarmControl/SwarmControlIJRR.tex
fixed the conclusion and changed title (lowercase and)
Fixed Olsen's bib, added wrap figure
Had to add functions for it to work.
updates through page 16

@@
 %\usepackage{authblk}
+\usepackage{wrapfig}
 
@@
 
-\title{\LARGE \bf Steering a Swarm Using Global Inputs And Swarm Statistics}
+\title{\LARGE \bf Steering a Swarm Using Global Inputs and Swarm Statistics}
 
@@
 %\affil[2]{Independent}
-\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T.
-Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA, {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
+\author{Shiva Shahrokhi, Chris Ertel, Mable Wan,\\ Lillian Lin, Aaron T.
+Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA, {\tt\small sshahrokhi2@uh.edu, atbecker@uh.edu}}
  %end thanks

### 2016-08-12 journalSwarmControl/expResults.tex
updates through page 16

@@
 Our hypothesis was that participants would complete the task faster with more robots.
-The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating rather that there is a local minima around 130 robots.
+The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minima around 130 robots.
 
@@
 Sensing is expensive, especially on the nanoscale.
-To see nanocars~\cite{Chiang2011}, scientists fasten molecules that fluoresce light when activated by a strong light source.
+To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source.
 Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}.
@@
 Full-state requires $2n$ data points for $n$ robots.
-Convex-hull requires at worst $2n$, but usually a smaller number.
+Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.
 Mean requires two, and variance three, data points.

### 2016-08-12 journalSwarmControl/relatedWork.tex
updates through page 16

@@
 Their sample problem involved a multi-robot search task, where users could assign goals to robots.
-Their user interaction studies with simulated planar robots  indicated a \emph{fanout plateau} of about 8 robots, after which there were diminishing returns.
-They hypothesize that the location of this plateau is highly dependent on the underlying task.
-Indeed, our paper indicates there are some tasks without plateaus.
+Their user interaction studies with simulated planar robots  indicated a \emph{fanout plateau} of about 8 robots, with diminishing returns for more robots.
+They hypothesized the location of this plateau is highly dependent on the underlying task.
+Indeed, our paper indicates there are tasks without plateaus.
 % (Lemmings is an example of this as well, for systems with autonomy
@@
 
-\cite{Squire:2006:HCM:1121241.1121248} designed experiments showing that user-interface design had a high impact on the task effectiveness and the number of robots that could be controlled simultaneously in a multi-robot task.
+\cite{Squire:2006:HCM:1121241.1121248} designed experiments showing user-interface design had a high impact on the task effectiveness and the number of robots that could be controlled simultaneously in a multi-robot task.
 
-A number of user studies compare methods for controlling large swarms of simulated robots, for example \cite{bashyal2008human,kolling2012towards,de2012controllability}.
+Several user studies compare methods for controlling large swarms of simulated robots, for example \cite{bashyal2008human,kolling2012towards,de2012controllability}.
 These studies provide insights but are limited by cost to small user studies; have a closed-source code base; and focus on controlling intelligent, programmable agents.
@@
 
-Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators, e.g.
-\cite{odhner2014compliant,deimel2014novel}.
-Like these manipulators, our swarms conform to the object to be manipulated.
-However our swarms lack the restoring force provided by flexures in \cite{odhner2014compliant} and the silicone in \cite{deimel2014novel}.
+Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.
+Our swarms conform to the object to be manipulated.
+However our swarms lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
 Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},

### 2016-08-12 journalSwarmControl/theory.tex
Changed Sections and u
Deleted some unnecessary files and added hysteresis new fig
updates through page 16

@@
 and thus all four states are controllable.
-This Section starts by proving independent position control of many robots is not possible, but the mean position can be controlled.
+This section starts by proving independent position control of many robots is not possible, but the mean position can be controlled.
 We then provide conditions under which the variance of many robots is also controllable.
@@
 \end{equation}  
-\subsection{Controlling mean position of many robots}\label{sec:controlMeanPosition}
+\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
 This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis.
@@
 \centering
-\begin{overpic}[width = 0.5\columnwidth]{hysteresis.png}
-\put(33,21){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
-\put(33,5){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
+\begin{overpic}[width = 0.5\columnwidth]{hysteresis.pdf}
+\put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
+\put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
 \begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}

### 2016-08-12 journalSwarmControl/Introduction.tex
finished Becker edits to entire paper

@@
 
- This paper is the synthesis of two preliminary conference papers,  \cite{swarmcontrol2013}, covered first few months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015}, presented simulations of object manipulation.
+ This paper is the synthesis of two preliminary conference papers.
+\cite{swarmcontrol2013}, covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015}, presented simulations of object manipulation.
 This paper presents the final results from SwarmControl.net.

### 2016-08-12 journalSwarmControl/expResults.tex
finished Becker edits to entire paper

@@
 This experiment varied from 1 to 500 the population of robots used to transport an object.
-We kept the total area, maximum robot speed, and total net force the swarm could produce constant.
+The total area, maximum robot speed, and total net force the swarm could produce were constant.
 The robots pushed a large hexagonal object through an  {\sffamily S}-shaped maze.
-Our hypothesis was that participants would complete the task faster with more robots.
+We hypothesized participants would complete the task faster with more robots.
 The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minima around 130 robots.
@@
 \paragraph{Varying visualization}
+\begin{figure}[b!]
+\renewcommand{\figwid}{0.24\columnwidth}
+\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
+\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
+\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
+\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
+\vspace{-.5em}
+\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}.
+This experiment challenges players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region.
+We record the completion time and other statistics.
+%\vspace{-1em}
+}
+\end{figure}
+
+\begin{wrapfigure}{R}{0.5\textwidth}
+  \vspace{-20pt}
+  \begin{center}
+  \includegraphics[width=0.5\textwidth]{ResVaryVis.pdf}
+  \end{center}
+  \vspace{-1em}
+\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
+\vspace{-1em}
+}
+\end{wrapfigure}
 
@@
 Mean and mean + variance are convenient even with millions of robots.
-Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
 
@@
 %\end{figure}
-
-\begin{wrapfigure}{R}{0.5\textwidth}
-  \vspace{-20pt}
-  \begin{center}
-  \includegraphics[width=0.5\textwidth]{ResVaryVis.pdf}
-  \end{center}
-  \vspace{-1em}
-\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
-\vspace{-1em}
-}
-\end{wrapfigure}
 
@@
 
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.24\columnwidth}
-\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
-\vspace{-.5em}
-\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}.
-This experiment challenges players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region.
-We record the completion time and other statistics.
-%\vspace{-1em}
-}
-\end{figure}
-
+Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean + variance, mean, full-state, convex-hull].
+As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [ mean, mean + variance, full-state, convex-hull].
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing
@@
 Anecdotal evidence from beta-testers who played the game suggests that tracking 100 robots is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
-Our beta-testers found convex-hull feedback confusing and irritating.
+Our beta-testers described convex-hull feedback as confusing and irritating.
 A single robot left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
@@
 \paragraph{Varying noise}
-Microrobots are affected by turbulence caused by random collisions with molecules. The effect of these collisions is called Brownian motion.
+Microrobots are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
 
@@
 Right: Increasing the number of robots resulted in longer completion times.
-For more than 4 robots the goal pattern contained a void, which may have caused the longer completion times.
+For more than 4 robots the goal pattern contained a void, which may have caused the jump in completion times.
 %\vspace{-1em}

### 2016-08-12 journalSwarmControl/experiment.tex
finished Becker edits to entire paper

@@
 
-This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in \S~\ref{sec:expResults}.  
+This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in \S \ref{sec:expResults}.  
 The swarm must deliver the object to the goal region.
@@
   Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}: a function that assigns a control input to every possible state.
- At each state the object can be commanded to move in one of eight directions with a small probability of moving in the wrong direction.
+ At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
 
@@
 \end{align}
-where $\mathbf{u}$ is the action and $x$ is the current state. 
+ $\mathbf{u}$ is the action and $x$ is the current state. 
   Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
@@
 \end{align}
-In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times.
-A {\sc Matlab} implementation of this algorithm is available at \href{https://www.mathworks.com/matlabcentral/fileexchange/49992-mdp-robot-grid-world-example}{\emph{MDP grid world example}}.%\todo{link to algorithm}.
+In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
 %Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
@@
 $\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 robots, pushing the object to goal using BFS required a mean and standard deviation of $183\pm 179$ s while Policy Iteration required $90\pm 35$ s.
+In 10 simulations with 100 robots, pushing the object to goal using BFS required $\mu=183$, $\sigma=179$ s while Policy Iteration required $\mu=90$, $\sigma=35$ s.
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 (Middle) The control policy from value iteration.
-(Right) The vision algorithm detects obstacles, used to produce the
-value function and control policy for the hardware setup. 
+(Right) The vision algorithm detects obstacles in the hardware setup.
+This map is used to produce the value function and control policy shown.
 %\vspace{-2em}
@@
 Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
+To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
 The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}.
@@
 Robots that cannot reach the object due to obstacles skew these calculations.
-The state machine in Fig.\ \ref{fig:Region}.a solves this problem by creating two states for the maze: either main or transfer.
+The state machine in Fig.\ \ref{fig:Region} solves this problem by creating two states for the maze: either main or transfer.
 Each state has a set of regions representing a discretized visibility polygon.
 Whenever the object crosses a region boundary the state toggles.
-The main regions are generated by extending obstacles until they meet another obstacle shown in Fig.~\ref{fig:Region}.b.
-The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions shown in Fig.~\ref{fig:Region}.c.
+The main regions are generated by extending obstacles until they meet another obstacle.
+The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.
 This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object. 
-In the example, we want the robots to push the object to the right.
+%In the example, we want the robots to push the object to the right.
 Without filtering the robots, the orange star is the mean and the algorithm would instruct the robots to push the object southeast.
@@
 \begin{center}
+	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\end{overpic}\\
 	\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf}\end{overpic}
-	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}\\
-	\vspace{0.5em}
-	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\end{overpic}
+	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}
 \end{center}
+\vspace{-0.5em}
 \caption{\label{fig:Region}  Outlier rejection state machine and regions.
@@
 \vspace{-2em}
-\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See animation in~\cite{ShivaVideo2015}.}
+\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.}
 %\vspace{-2em}

### 2016-08-12 journalSwarmControl/realExperiment.tex
finished Becker edits to entire paper

@@
 The kilobot, reported in \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available commercially or as an open source platform from \cite{K-Team2015}.
+It is available as an open-source platform or commercially from \cite{K-Team2015}.
 Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
@@
 In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and lights: four 50W LED floodlights  at the corners and four 30W LED floodlights on the sides of a 6 m square centered on the workspace and 1.5 m above the table.
-An Arduino Uno connected to an 8 relay shield controlled the lights.
+An Arduino Uno connected to an 8-relay shield controlled the lights.
 
 Above the table, an overhead machine vision system tracks the swarm.
-The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance)  by topology, the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
+The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance), the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
 
-The Laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}. 
+The laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}. 
 The objects were 3D printed from ABS plastic with a paper overlay. 
@@
 The kilobots run a phototaxis routine where they search for an orientation that aligns them with the light source, and then move with an approximately constant velocity toward this light.
-In practice, because the kilobots only have one light detector, they oscillate along this orientation.
+The kilobots oscillate along this orientation because they only have one light detector.
 
-We use the sign of \eqref{eq:PDcontrolPosition}, and choose the closest orientation among the eight light sources.
+We use the sign of \eqref{eq:PDcontrolPosition}, and choose the closest orientation to $\mathbf{D}(\mathbf{b})$ among the eight light sources.
 Fig.~\ref{fig:realMean} shows that this limited, discretized control still enables regulating the mean position of a swarm of 100 robots.
@@
 The moveable object was placed in the lower center of the workspace.
-The same controller as in the simulation, Alg.~\ref{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}.
-Trials were run until the object COM entered the goal region.
-The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} minutes.
-This is a mean of 42:38 and a standard deviation of 12:51 minutes.
+The same controller as in the simulation, Alg.~\ref{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}, available at \todo{matlab link}.
+Trials were run until the object COM entered the goal region.  The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} minutes.  This is a mean of 42:38 and a standard deviation of 12:51 minutes. 
 A circular object completed in 52:35 minutes. 
@@
 The automatic controller classifies pink objects as obstacles and generates a policy to the goal.
-See the video attachment for an animation~\cite{ShivaVideo2015}}
+See  video in multimedia extension.}
 %\vspace{-2em}

### 2016-08-12 journalSwarmControl/relatedWork.tex
finished Becker edits to entire paper

@@
 \subsection{Block pushing and compliant manipulation}
-Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object, as in \cite{Sudsang2002,Fink2007}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object.
+Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object, as in \cite{Sudsang2002,Fink2008}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object.
 A key difference is that our robots are compliant and tend to flow around the object, making this similar to fluidic trapping as in \cite{Armani2006} and \cite{Becker2009}.

### 2016-08-12 journalSwarmControl/simulation.tex
Changed Sections and u
Latest changes
finished Becker edits to entire paper

@@
 100 robots were used and the maximum speed was 3 meters per second.
-As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52 s with $K_{p}= 4$, and  $K_{d} = 1$.
+As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$.
 \begin{figure}
@@
 
-For variance control we use the control law discussed in Section~\ref{sec:VarianceControl}. 
+For variance control we use the control law \eqref{eq:PDcontrolVariance}. 
  Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.

### 2016-08-12 journalSwarmControl/theory.tex
finished Becker edits to entire paper

@@
 
-and our state space representation as:
+and our state space representation is:
 \begin{equation}
@@
 0 & \frac{1}{m}
-\end{bmatrix}  [u_x,u_y]
+\end{bmatrix} 
+ \begin{bmatrix}
+ u_x\\
+ u_y\end{bmatrix} 
 \end{equation}
@@
  \right],
+ \quad \textrm{rank}(\mathcal{C})=4,
 \end{equation}
@@
 Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
-%\begin{equation}\left[ x_{1,1},x_{2,1},\ldots, x_{2n-1},x_{2n}\right]^\intercal = \left[ p_{x,1},v_{x,1},\ldots,p_{x,n},v_{x,n}\right]^\intercal. \nonumber\end{equation}
-%\begin{align}
-%\dot{p}_{x1} &= v_{x1}\\\nonumber
-%\dot{v}_{x1} &= a_{x1}\\\nonumber
-%\dot{p}_{x2} &= v_{x2}\\\nonumber
-%\dot{v}_{x2} &= a_{x2}\\\nonumber
-%&\vdots\\\nonumber
-%\dot{p}_{xn} &= v_{xn}\\\nonumber
-%\dot{v}_{xn} &= a_{xn}\nonumber
-%\end{align}
 Our state-space representation is:
@@
 \end{equation}
- However, just as with one robot, we can only control two states because $\mathcal{C}$ has rank two:
+ However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
 \begin{equation}
-\mathcal{C}=\left[ \begin{matrix} 
+\mathcal{C}_n=\left[ \begin{matrix} 
 0\\
@@
 0
-\end{matrix}\,\middle\vert\,, 
-\ldots \right]
+\end{matrix}\,\middle\vert\,
+\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
 \end{equation}  
@@
 
-We again analyze $\mathcal{C}$:
+We again analyze the controllability matrix $\mathcal{C}_{\mu}$:
 \begin{equation}
-\mathcal{C}=\left[ \begin{matrix} 
+\mathcal{C}_\mu=\left[ \begin{matrix} 
 0\\
@@
 \end{matrix}
- \right]
+ \right],  \quad \textrm{rank}(\mathcal{C}_\mu)=2.
 \end{equation}
-This matrix has rank two, and thus the average position and average velocity are controllable.
+Thus the average position and average velocity are controllable.
 
@@
 
-We control mean position with a PD controller that uses the mean position and mean velocity. Our control input is the global force applied to each robot:
+We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
 \begin{align}
-u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\bar{v}_x) \nonumber\\
-u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\bar{v}_y)  \label{eq:PDcontrolPosition}
+u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}) \nonumber\\
+u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}})  \label{eq:PDcontrolPosition}
 \end{align}
@@
 \begin{align}
-\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t)
+\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t),
 \end{align}
@@
 Given a large obstacle-free workspace, a \emph{Brownian noise} process increases the variance linearly with time.
-\begin{equation}\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon} \end{equation}
- If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board~\cite{Becker2013b}.
+\begin{equation}
+\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon},
+\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t
+\end{equation}
+ If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board as in~\cite{Becker2013b}.
 
-%After some time, Gaussian distribution shapes the outline of the robots because of the Brownian noise feature:\\
-%\begin{equation}
-%P(x) = \frac{1}{\sigma \sqrt{2\pi } }e^{ -  \frac{\left( x - \mu  \right)^2 } {2\sigma ^2 }}
-%\end{equation}
 If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
@@
 \end{align}
-In a slight abuse of notation we call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
+We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
 Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
 \begin{align}
-x_{goal}(\sigma^2_{ref}) = r + \sqrt{3\sigma^2_{ref}}
+x_{goal}(\sigma^2_{ref}) = \ell_x/2 = r + \sqrt{3\sigma^2_{ref}}
 \end{align}
- If another wall is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.
+ If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.
 
@@
 \Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
-\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ 
+\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$  \Comment{initially not in variance control mode}
 \State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$

### 2016-08-12 journalSwarmControl/Introduction.tex
added Mable's edits

@@
 
-Our experiments show that numerous simple robots responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.
+Our online experiments show that numerous simple robots responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.
 
@@
 %{blockpushingImageWithMeanAndVarianceOverlay.png}
-\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.
+\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.
 Here a swarm of simple robots (blue discs) pushes a green hexagon toward the goal.
-Simulation at left, screen capture form hardware experiment at right.
+Simulation at left, screen capture from hardware experiment at right.
 See multimedia extension.}

### 2016-08-12 journalSwarmControl/experiment.tex
added Mable's edits

@@
 
-This section analyzes an \emph{object manipulation} task attempted by both our hybrid, hysteresis-based controller, inspired by the analysis of human users in \S \ref{sec:expResults}.  
+This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers, inspired by the analysis of human users in \S \ref{sec:expResults}.  
 The swarm must deliver the object to the goal region.
@@
 Robots that cannot reach the object due to obstacles skew these calculations.
-The state machine in Fig.\ \ref{fig:Region} solves this problem by creating two states for the maze: either main or transfer.
+The state machine in Fig.\ \ref{fig:Region}A solves this problem by creating two states for the maze: either main or transfer.
 Each state has a set of regions representing a discretized visibility polygon.
@@
 The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.
+
+Fig.\ \ref{fig:Region}B shows the regions for the main state.
+The object is in region 1.
+An indicator function is applied to \eqref{eq:meanVar} so only robots inside region 1 are counted.
 This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object.
-%In the example, we want the robots to push the object to the right.
-Without filtering the robots, the orange star is the mean and the algorithm would instruct the robots to push the object southeast.
-The filtered mean is at the yellow star and the algorithm instructs the robots to push the object directly east.
-%When the object leaves main region 1 the maze state switches to transfer. 
-%The object is in transfer region 0, so only robots in transfer region 0 are included in the mean and covariance calculations.  
-%This heuristic improves performance by $50\%$ or less regarding the object heaviness.
+When the object leaves main region 1 the state switches to transfer.
+The transfer regions are shown in Fig.
+3.
+The object is in transfer region 1, so only robots in transfer region 1 are included in the mean and covariance calculations.
+In Fig.\ \ref{fig:Region}C, the robots should push the object to the left.
+Without filtering using regions, the red circle is the mean and the algorithm would instruct the robots to push the object up.
+The black circle shows the filtered mean and the algorithm instructs the robots to push the object directly left.
+
 In 10 simulations with 100 robots, completion time without outlier rejection required $\mu=245, \sigma=135$ s while using outlier rejection required $\mu=90, \sigma=35$ s.
@@
 \begin{center}
-	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\end{overpic}\\
-	\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf}\end{overpic}
-	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}\end{overpic}
+	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
+	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
+	\put(-6,85){B}
+	\end{overpic}
+	~~
+	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}
+	\put(-6,85){C}
+	\end{overpic}
 \end{center}
@@
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit.
+The min and max variance limits from  \label{eq:sigMaxMin} incremented by $0.003 W$ and $0.006 W$.
  The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.
@@
 In rare cases during simulations the swarm may become trapped in a local minima of \eqref{eq:potentialfield}.
-If the swarm mean position does not change for ten seconds, the swarm is assumed to be in a local minima and is commanded to move toward the previous corner.
+If the swarm mean position does not change for five seconds, the swarm is assumed to be in a local minima and is commanded to move toward the previous corner.
 As soon as the mean position changes, normal control resumes.
@@
 Each trial measured the time to deliver the object to the goal location.
-The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied each simulation step) with $W=5$.
+The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied once each simulation step) with $W=5$.
 

### 2016-08-12 journalSwarmControl/code/ObjectManipulation/MDPgridworldFunction.m
updated the code for sharing

@@ -69,10 +69,13 @@ R(goalX,goalY) = 100; %goal state has big reward
             end
             iteration = iteration+1;
             set(hImageV,'cdata',V_hat);
+            title(['Value iteration step ',num2str(iteration),' of ', num2str(iteration_limit)])
             drawnow
         end
         hold on; hq=quiver(X,Y,DY,DX,0.5,'color',[0,0,0]); hold off
         set(hq,'linewidth',2);
+        title(['Policy after ', num2str(iteration_limit),' iterations'])
+        drawnow
     end
 
     function [bestMove,bestPayoff] = policy_MDP(index,V_hat,prob)

### 2016-08-12 journalSwarmControl/code/ObjectManipulation/ObjectManipulationKilobots.m

@@ -1,32 +1,50 @@
 %% Object Manipulation Experiment With Kilobots
-% Uses an overhead vision system to control a swarm of
-% kilobots to push an object through a maze.
-% Swarm is attracted to the brightest light in the room
-% Lights are controlled by an Arduino relay-sheild
+% Uses an overhead vision system to control a swarm of kilobots to push an
+% object through a maze. Swarm is attracted to the brightest light in the
+% room.
 %
-% requires RegionCode.m and MDPgridworldFunction.m
-% -------------------------
-% By Shiva Shahrokhi, Lillian Lin and Mable Wan, Summer 2016
+% see video: https://youtu.be/tD1JvgRBRVM
+% 
+% In these experiments a whiteboard on a table is the workspace, with four
+% 50W LED floodlights at the corners and four 30W LED floodlights on the
+% sides of a 6 m square centered on the workspace and 1.5 m above the
+% table. An Arduino Uno connected to an 8-relay shield controls the lights.
+% Above the table, an overhead machine vision system tracks the swarm. The
+% vision system identifies obstacles by color segmentation, determines the
+% corners (used to decrease  variance), the object by color segmentation,
+% and identifies robots using color segmentation and circle detection with
+% a circular Hough transform. The path planning uses value iteration for
+% the object and potential fields for the swarm to interact with the
+% object.
+% 
+% REQUIRES RegionCode.m, MDPgridworldFunction.m, RelayOn.m, dist2points.m,
+% AngleFix.m, circle.m, FlowForce.m, plot_gaussian_ellipsoid.m
+% and Arduino drivers:
+% http://www.mathworks.com/hardware-support/arduino-matlab.html
+%
+% SETUP: connect a webcam via USB and an Arduino via USB.
+% ------------------------- 
+% By Shiva Shahrokhi, Lillian Lin, Mable Wan, & Aaron T Becker, Summer 2016
+   
 format compact
 close all
 clear all
 
+%% Define webcam
+webcamShot = true; %if false, uses stored image 'KilobotTableExample.jpeg' if true, uses webcam
+
 %% Load maps from RegionCode.m
-RegionCode
+RegionCode(webcamShot)
 
 load('MazeMap', 'movesX', 'movesY','corners');
 load('ThresholdMapsMac','transferRegion','mainRegion'); 
 
-
-%% Define webcam
-webcamShot = true;
-
 if webcamShot
-    cam = webcam(1);
[…]

### 2016-08-12 journalSwarmControl/code/ObjectManipulation/RegionCode.m

@@ -1,3 +1,4 @@
+function RegionCode(webcamShot)
 %% REGIONCODE -- preprocessing for block pushing experiment
 % Takes a snapshot from the webcam, then processes it to find
 % obstacles. From these obstacles, the corners, regions, and a policy for MDP
@@ -18,34 +19,30 @@
 % ---------------------
 % By Shiva Shahrokhi and Lillian Lin,  Summer 2016
 % Shiva Shahrokhi <shiva.shahrokhi@gmail.com>
-close all
-clear all
 
-success = false;
-webcamShot = true;
+if nargin<1
+    webcamShot = false;
+end
+
+tic;
 if webcamShot
     cam = webcam(1);
-end
-t0 = tic;
-if webcamShot
     originalImage = snapshot(cam);
-    if (ispc)
-        original = imcrop(originalImage,[50 10 500 400]);
-        img = imcrop(originalImage,[50 10 500 400]);
-        rgbIm = imcrop(originalImage,[50 10 500 400]);
-    else
-        original = imcrop(originalImage,[345 60 1110 860]);
-        img = imcrop(originalImage,[345 60 1110 860]);
-        rgbIm = imcrop(originalImage,[345 60 1110 860]);
-    end
+    %imwrite(originalImage,'KilobotTableExample.jpeg'); % to save a new image
 else
-    rgbIm = imread('PC.jpeg'); %#ok<UNRCH>
+    originalImage = imread('KilobotTableExample.jpeg');
+end
+
+if (ispc)  % Code to run on PC platform
+    rgbIm = imcrop(originalImage,[50 10 500 400]);
+else        % Code to run on Mac platform
+    rgbIm = imcrop(originalImage,[345 60 1110 860]);
 end
 
 %% Obstacle Thresholds
 I = rgb2hsv(rgbIm);
 
-% Define thresholds for channel 1 based on histogram settings 
+% Define thresholds for channel 1 based on histogram settings
 % these were picked using the Color Thresholding tool
 channel1Min = 0.862;
 channel1Max = 0.945;
@@ -64,7 +61,7 @@ BW = (I(:,:,1) >= channel1Min ) & (I(:,:,1) <= channel1Max) & ...
     (I(:,:,3) >= channel3Min ) & (I(:,:,3) <= channel3Max);
 
 %% finding Obstacles
-[B,L] = bwboundaries(BW, 'noholes');
[…]

### 2016-08-12 journalSwarmControl/code/ObjectManipulation/relayOn.m

@@ -1,4 +1,4 @@
-function relayOn(a,which)
+function RelayOn(a,which)
 % RELAYON uses an existing arduino connection to turn on the lights
 %   RELAYON(a,which)
 %

### 2016-08-15 journalSwarmControl/experiment.tex
Added reference to matlab code

@@
 
-\subsection{Algorithm for object manipulation}
+\subsection{Algorithm for object manipulation}\label{sec:AlgObjectManipulation}
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit.

### 2016-08-15 journalSwarmControl/realExperiment.tex
Added reference to matlab code

@@
 The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance), the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
+Code for the vision, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central by \cite{Shahrokhi2015MDP}.
 
-The laser-cut patterns for the neon green fiducial markers on the robots and {\sc Matlab} tracking code are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}. 
 The objects were 3D printed from ABS plastic with a paper overlay. 
@@
 254 cm$^2$ circle, 
-and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.
+and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
 

### 2016-08-15 journalSwarmControl/Introduction.tex
Javier's comments fixed

@@
 We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.
- We conclude with implementations of these controllers in on hardware robots and use them to complete an object manipulation task with 100+ kilobots in \S \ref{sec:realExperiment}.
+ We conclude with implementations of these controllers in our hardware robots and use them to complete an object manipulation task with 100+ kilobots in \S \ref{sec:realExperiment}.
 

### 2016-08-15 journalSwarmControl/expResults.tex
Javier's comments fixed

@@
 As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [ mean, mean + variance, full-state, convex-hull].
-All experiments lasting over 300s were removed, under the assumption that the user stopped playing
+All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
 Using ANOVA analysis, we reject the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
@@
 \cite{goodrich2012types}.
-STMs can be used to arrange atoms and make small assemblies~\cite{avouris1995manipulation}.
+STMs can be used to arrange atoms and make small assemblies, as described in \cite{avouris1995manipulation}.
 An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules.

### 2016-08-15 journalSwarmControl/experiment.tex
Javier's comments fixed

@@
 When the object leaves main region 1 the state switches to transfer.
-The transfer regions are shown in Fig.
-3.
+The transfer regions are shown in Fig.\ \ref{fig:Region}C.
 The object is in transfer region 1, so only robots in transfer region 1 are included in the mean and covariance calculations.
-In Fig.\ \ref{fig:Region}C, the robots should push the object to the left.
+ The robots should push the object to the left.
 Without filtering using regions, the red circle is the mean and the algorithm would instruct the robots to push the object up.
@@
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit.
-The min and max variance limits from  \label{eq:sigMaxMin} incremented by $0.003 W$ and $0.006 W$.
+$0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
  The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.

### 2016-08-15 journalSwarmControl/simulation.tex
Javier's comments fixed

@@
 
-Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation~\cite{catto2010box2d}.
+Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation, presented in \cite{catto2010box2d}.
 All experiments ran on a Chrome web browser on a 2.6 GHz Macbook.

### 2016-08-15 journalSwarmControl/Introduction.tex
mary's edits on minimum
mean \pm std for numbers
added new swarm control code

@@
 Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.
-\href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are posted online, \cite{Chris-Ertel2013}.
-\todo{TODO: Chris, is 2016 code ready to upload?
-link to 2016 code}
+\href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are posted online at \cite{Chris-Ertel2016}.
 

### 2016-08-15 journalSwarmControl/experiment.tex
mary's edits on minimum
mean \pm std for numbers
added new swarm control code

@@
 $\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 robots, pushing the object to goal using BFS required $\mu=183$, $\sigma=179$ s while Policy Iteration required $\mu=90$, $\sigma=35$ s.
+In 10 simulations with 100 robots, pushing the object to goal using BFS required 183$\pm$179 s while Policy Iteration required 90$\pm$35 s (mean$\pm$std).
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 No failures occurred with the repulsive potential field.
-Of successful trials, completion time without repulsive potential fields required $\mu=245, \sigma=135$ s while using repulsive potential fields required $\mu=90, \sigma=35$ s.
+Of successful trials, completion time without repulsive potential fields required 245$\pm$135 s while using repulsive potential fields required 90$\pm$35 s (mean$\pm$std).
 %In 10 simulations with 100 robots, completion time without repulsive potential fields required $\mu=245, \sigma=135$s while using repulsive potential fields required $\mu=90, \sigma=35$.
@@
 
-In 10 simulations with 100 robots, completion time without outlier rejection required $\mu=245, \sigma=135$ s while using outlier rejection required $\mu=90, \sigma=35$ s.
+In 10 simulations with 100 robots, completion time without outlier rejection required 245$\pm$135 s while using outlier rejection required 90$\pm$35 s (mean$\pm$std).
 %\todo{report results of 10 simulations with and without regions.  Give mean and std}
@@
 
-In rare cases during simulations the swarm may become trapped in a local minima of \eqref{eq:potentialfield}.
-If the swarm mean position does not change for five seconds, the swarm is assumed to be in a local minima and is commanded to move toward the previous corner.
+In rare cases during simulations the swarm may become trapped in a local minimum of \eqref{eq:potentialfield}.
+If the swarm mean position does not change for five seconds, the swarm is assumed to be in a local minimum and is commanded to move toward the previous corner.
 As soon as the mean position changes, normal control resumes.

### 2016-08-15 journalSwarmControl/realExperiment.tex
mary's edits on minimum
mean \pm std for numbers
added new swarm control code

@@
 The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance), the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
-Code for the vision, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central by \cite{Shahrokhi2015MDP}.
-
-The objects were 3D printed from ABS plastic with a paper overlay. 
-Shapes included a 325 cm$^2$ equilateral triangle, 
-%Small Equilateral Triangle 140 cm$^2$,
- 324 cm$^2$ square,
- 281 cm$^2$ hexagon,
-254 cm$^2$ circle, 
-and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
-
-%\begin{figure*}
-%\begin{center}
-%	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-%	\begin{overpic}[width=0.6\columnwidth]{SetUp.pdf}%\put(1,75){A}
-%	\end{overpic}\\
-%	\begin{overpic}[width=0.35\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
-%	\end{overpic}
-%	\begin{overpic}[width=0.35\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
-%	\end{overpic}
-%\end{center}
-%\caption{\label{fig:setup}
-%(Top) Hardware platform. (Lower left) Four shapes used for hardware experiments. (Lower right) Visualization of potential field. }
-%\end{figure*}
 
@@
 \end{figure*}
+
+The objects were 3D printed from ABS plastic with a paper overlay. 
+Shapes included a 325 cm$^2$ equilateral triangle, 
+%Small Equilateral Triangle 140 cm$^2$,
+ 324 cm$^2$ square,
+ 281 cm$^2$ hexagon,
+254 cm$^2$ circle, 
+and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
 
@@
 
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=\columnwidth]{XYMeanControl.eps}
-%\end{center}
-%\caption{\label{fig:meanRobotFig}
-%Mean Control experiment with kilobots.
-%}
-%\end{figure}
-
 \subsection{Automated object manipulation (hardware experiment)}
@@
 The moveable object was placed in the lower center of the workspace.
-The same controller as in the simulation, Alg.~\ref{alg:BlockPushing}, but with discretized control inputs, was coded in {\sc Matlab}, available at \todo{matlab link}.
-Trials were run until the object COM entered the goal region.  The trials ran for \{24:25, 57:37, 50:00, 36:02, 45:07\} minutes.  This is a mean of 42:38 and a standard deviation of 12:51 minutes. 
-A circular object completed in 52:35 minutes. 
-A square object completed in 114:31 minutes. 
-A triangular object completed in XX:YY minutes.
+{\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central by \cite{Shahrokhi2015MDP}.
+Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std). 
+A circular object completed in 3155 s. 
+A square object completed in 6871 s. 
+\todo{A triangular object completed in ??:?? minutes.}
 

### 2016-08-15 journalSwarmControl/experiment.tex
Made fig 1 smaller, no captions

@@
 The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
-However, BFS has no penalties on object collisions with obstacles, and the resulting collisions made the BFS solution slow.
+This approach using workspace BFS fails to account for the hull of the object and will suggest moves that can cause at least partial collision with the workspace.
+A configuration-space BFS approach avoids that problem but still fails to model uncertain actuation of the object by the swarm.
 
- To avoid obstacles, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
+To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
   Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}: a function that assigns a control input to every possible state.

### 2016-08-15 journalSwarmControl/expResults.tex
mary's edits on minimum
mean \pm std for numbers
added new swarm control code
updated mathematica images and removed 2 figs

@@
 \begin{overpic}[width = 0.5\columnwidth]{ResVaryNum.pdf}\end{overpic}
-\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
+%\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
 \caption{
 \label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.
-(Left) Data indicates this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object.
+Data indicates this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object.
 Best-fit linear and quadratic lines are overlaid for comparison.
-(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
+%(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
 }
@@
 We hypothesized participants would complete the task faster with more robots.
-The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minima around 130 robots.
+The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 robots.
 

### 2016-08-16 journalSwarmControl/code/ObjectManipulation/ObjectManipulationKilobots.m
References are correct now.

@@ -34,7 +34,7 @@ clear all
 webcamShot = true; %if false, uses stored image 'KilobotTableExample.jpeg' if true, uses webcam
 
 %% Load maps from RegionCode.m
-RegionCode(webcamShot)
+%RegionCode(webcamShot)
 
 load('MazeMap', 'movesX', 'movesY','corners');
 load('ThresholdMapsMac','transferRegion','mainRegion'); 
@@ -80,7 +80,7 @@ while success == false
     
     if webcamShot
         if (again == true)
-            RelayOn(a,0);
+            relayOn(a,0);
             pause (10);
         end 
         rgbIm = snapshot(cam);

### 2016-08-16 journalSwarmControl/SwarmControlIJRR.tex
Javier's comments fixed
added conflict of interest
Changed format

--- 2016-08-08 journalSwarmControl/SwarmControlIJRR.tex
@@
-% How many words:  $ detex IJRRensembleDifferentialDrive.tex | wc -w
-% Submit at http://mc.manuscriptcentral.com/ijrr
+%%%%%%%%% IJRR SETTINGS
+%  from https://us.sagepub.com/en-us/nam/manuscript-submission-guidelines#PreparingYourManuscript
+\documentclass[Afour,sageh,times]{sagej} % sage_latex_guidelines.tex V1.01, 11 June 2015
+\usepackage{moreverb,url}
+\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=black,urlcolor=black,linkcolor = black]{hyperref}
+\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
+T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\def\volumeyear{2016}
+%%%%%%%%%%%% END IJRR SETTINGS
 
-\documentclass[11pt]{article}
-% IJRR Does not have a template for submissions, but it recommends using
-% the standard Latex article format, with computer modern font.
-% \usepackage[margin=1in]{geometry}
-% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}
-
-% www.SwarmControl.net
-%http://mrsl.rice.edu
-%http://people.ae.illinois.edu/tbretl/
-
-\usepackage[pdftex]{graphicx}        % standard LaTeX graphics tool
-\usepackage[usenames,dvipsnames]{color}
+%\usepackage[pdftex]{graphicx}        % standard LaTeX graphics tool
+%\usepackage[usenames,dvipsnames]{color}
 \usepackage{calc}
 \usepackage{graphicx}
-\usepackage[cmex10]{amsmath}
+%\usepackage[cmex10]{amsmath}
 \usepackage{amssymb}
@@
 %\usepackage{cite}
-\usepackage[font={small}]{caption}
+%\usepackage[font={small}]{caption}
 \usepackage[caption=false,font=footnotesize]{subfig}
@@
 \usepackage{upgreek}   % non-slanted greek letters
-\usepackage{url}
 \usepackage{soul} %for highlighting
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks =true,
-  urlcolor = black,
-  linkcolor = black,
-  citecolor= black
-}
-\usepackage[round]{natbib}
+
+%\usepackage[round]{natbib}
 \usepackage{mathtools}
 %\usepackage{authblk}
+\usepackage{wrapfig}
 
@@
 \newcommand{\FALSE}{\textbf{false}}
-\newcommand*{\email}[1]{\texttt{#1}}
 
@@
 \begin{document}
+\setcounter{secnumdepth}{3}
+\runninghead{Shahrokhi \textit{et~al.}}
 \hyphenation{micro-robot nano-robot nano-robotics micro-robotics micro-robots nano-robots micro-manip-ulators}
 %%%%%%%%%%%% For debugging purposes, I like to display the TOC
-    \tableofcontents
-    \setcounter{tocdepth}{3}
-    \newpage
+%    \tableofcontents
+%    \setcounter{tocdepth}{3}
+%    \newpage
 %%%% END TOC %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\title{\LARGE \bf Steering a Swarm Using Global Inputs And Swarm Statistics}
+\title{\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics}
 
-%\author{Shiva Shahrokhi, Chris Ertel, Mable Wan,Lillian Lin}
-%\author{Aaron~T.~Becker}
-%\affil[1]{Dept.\ Electrical \& Computer Engineering, University of Houston\thanks{ University of Houston, Houston, TX, 77004 USA {\tt\small sshahrokhi2@uh.edu}}
-%}
-%\affil[2]{Independent}
-\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T.
-Becker\thanks{Authors are with the Department of Electrical and Computer Engineering, University of Houston, Houston, TX, 77004 USA, {\tt\small sshahrokhi2@uh.edu,atbecker@uh.edu}}
- %end thanks
-} % end author block 
-\maketitle
+\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T. Becker}% end author block 
+\affiliation{Electrical and Computer Engineering, University of Houston, USA}
+\corrauth{Shiva Shahrokhi,
+University of Houston,
+4726 Calhoun Rd,
+Houston, TX 77004-4005, USA}
+\email{sshahrokhi2@uh.edu}
+
 \begin{abstract}
-Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
-The same properties that promise breakthrough solutions---small size and large populations---present unique challenges to generating controlled motion.
+Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
+The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
 We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures.
@@
 We present the game framework as a new, open-source tool for large-scale user experiments.
-One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the robot swarm than with full-state feedback.
-Inspired by human operators, this paper next investigates controllers that use only the mean and variance of a robot swarm.
+One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the swarm than with full-state feedback.
+Inspired by human operators, this paper next investigates controllers that use only the mean and variance of the swarm.
 We prove that the mean position is controllable, then provide conditions under which variance is controllable.
-We next derive automatic controllers for these and a hybrid, hysteresis-based switching control to regulate the first two moments of the robot distribution.
+We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the robot distribution.
 Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.
 \end{abstract}
+
+\keywords{micro-robot, global control, human-swarm interaction}
+
+\maketitle
+
 %\input{00-Abstract.tex}
@@
 
-\section*{Funding} %formatted according to https://us.sagepub.com/en-us/nam/funding-acknowledgements
+\begin{dci}
+The Authors declare that there is no conflict of interest.
+To typeset a "Declaration of
+  conflicting interests" section.
+\end{dci}
+\begin{funding}
 %RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
@@
 \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
-%Optional disclaimer (mandatory for non-scientific)
-%Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
-
-%\section*{Acknowledgements}
+\end{funding}
+\begin{sm}
+Video 1:  
+Video 2: 
+\end{sm}
 
@@
 %\bibliographystyle{IEEEtran}
-\bibliographystyle{plainnat}
+\bibliographystyle{SageH}
 \bibliography{IEEEabrv,SwarmControlIJRR}

### 2016-08-16 journalSwarmControl/expResults.tex
added conflict of interest
Changed format

@@
 
-\paragraph{Varying number}
+\paragraph{Varying number:}
 \begin{figure}
-\begin{overpic}[width = 0.5\columnwidth]{ResVaryNum.pdf}\end{overpic}
+\begin{overpic}[width = 0.9\columnwidth]{ResVaryNum.pdf}\end{overpic}
 %\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
 \caption{
-\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.
-Data indicates this task has an optimal number of robots, perhaps due to the relative sizes of the robots, obstacles, and object.
-Best-fit linear and quadratic lines are overlaid for comparison.
+\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.  Best-fit linear and quadratic lines are overlaid for comparison. 
 %(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
@@
 
-\paragraph{Varying visualization}
+\paragraph{Varying visualization:}
 \begin{figure}[b!]
@@
 
-\begin{wrapfigure}{R}{0.5\textwidth}
-  \vspace{-20pt}
+\begin{figure}
+%  \vspace{-20pt}
   \begin{center}
-  \includegraphics[width=0.5\textwidth]{ResVaryVis.pdf}
+  \includegraphics[width=0.8\columnwidth]{ResVaryVis.pdf}
   \end{center}
-  \vspace{-1em}
+%  \vspace{-1em}
 \caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
@@
 }
-\end{wrapfigure}
+\end{figure}
 
@@
 
-\paragraph{Varying noise}
+\paragraph{Varying noise:}
 Microrobots are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
@@
 \begin{figure}[b!]
-\renewcommand{\figwid}{0.5\columnwidth}
+\renewcommand{\figwid}{0.49\columnwidth}
 \begin{overpic}[width =\figwid]{ResVaryNoise.pdf}\end{overpic}
 \begin{overpic}[width =\figwid]{ResPositioning.pdf}\end{overpic}
-\vspace{-1em}
+%\vspace{-1em}
 \caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
@@
 
-\paragraph{Position control}
+\paragraph{Position control:}
 This online experiment examined how completion time scales with the number of robots $n$.
@@
 
-\paragraph{Varying control: Assembly}
+\paragraph{Varying control--Assembly:}
 Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.
@@
 \begin{figure}[b!]
-\renewcommand{\figwid}{3.3cm}
-\begin{overpic}[height =\figwid]{ResVaryControl.pdf}\end{overpic}
-\begin{overpic}[height =\figwid]{ResVaryForage.pdf}\end{overpic}
+\renewcommand{\figwid}{0.45\columnwidth}
+\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
+\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
 \vspace{-.5em}
@@
 
-\paragraph{Varying control: Foraging}
+\paragraph{Varying control--Foraging:}
 Collecting and delivering resources is necessary for drug delivery.

### 2016-08-16 journalSwarmControl/experiment.tex
fixed bf u
Changed format

@@
 \begin{align}
-r(x,u) &=  \left\{
+r(x,\mathbf{u}) &=  \left\{
 \begin{array}{ll}
@@
 \text{for }&\text{$i=1$ to $N$ do} \nonumber \\
-&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration}\\
-\text{end}& \nonumber
+&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration} \nonumber\\
+\text{end}& 
 \end{align}
@@
 \centering
-\begin{overpic}[scale=0.19]{GradientView.png}
+\renewcommand{\figwid}{3cm}
+\begin{overpic}[height=\figwid]{GradientView.png}
 \end{overpic}
-\begin{overpic}[scale=0.19]{PolicyIter.png}
+\begin{overpic}[height=\figwid]{PolicyIter.png}
 \end{overpic}
-\begin{overpic}[scale=0.262]{MDPmap.pdf}
+\begin{overpic}[height=\figwid]{MDPmap.pdf}
 \end{overpic}
 \vspace{-0.5em}
-\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object  to compute gradient vectors (left).
+\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left).
 Modeling as an MDP enables encoding penalties for being near obstacles.
@@
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
-	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
+	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
 	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
@@
 }
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{wrapfigure}{R}{0.3\textwidth}
-  \vspace{-20pt}
+\begin{figure}
+%  \vspace{-20pt}
   \begin{center}
-\begin{overpic}[width=0.3\columnwidth]{AllShapes.pdf}\end{overpic}
+\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
   \end{center}
-  \vspace{-1em}
-\caption{\label{fig:Shapes} Six equal-area objects were tested. %Six object shapes were tested.
+%  \vspace{-1em}
+\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
 \vspace{-1em}
 }
-\end{wrapfigure}
+\end{figure}
 
@@
 \While{$\sigma^2 > \sigma_{min}^2$}
-\State $ [x_{goal}, y_{goal}] \gets $ the nearest corner in current region
+\State $ [x_{goal}, y_{goal}] \gets $ nearest corner in region
 \State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
@@
 \State Calculate $\mathbf{D}(\mathbf{b})$  \Comment{direction for object at $\mathbf{b}$}
-\State Apply \eqref{eq:potentialfield}   \Comment{potential field to push object in correct direction}
+\State Apply \eqref{eq:potentialfield}   \Comment{potential field for swarm}
 \EndIf
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
@@
 \end{overpic}
-\vspace{-2em}
-\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.}
+\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.
 %\vspace{-2em}
 }
-\end{figure*}
+\end{figure}
 
@@
 \renewcommand{\figwid}{0.5\columnwidth}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}\put(1,55){a)}
+\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}\put(1,55){b)}
+\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}\put(1,55){c)}
+\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
 \end{overpic}
-\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}\put(1,55){d)}
+\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
 

### 2016-08-16 journalSwarmControl/realExperiment.tex
Changed format

@@
 \begin{figure*}
+\renewcommand{\figwid}{6cm}
 \begin{center}
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[width=0.5\columnwidth]{SetUp.pdf}%\put(1,75){A}
-	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
-	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
+	\begin{overpic}[height=\figwid]{SetUp.pdf}%\put(1,75){A}
+	\end{overpic}\begin{overpic}[height=\figwid]{AllShapesR.pdf}%\put(-0,85){B}
+	\end{overpic}\begin{overpic}[height=\figwid]{PotentialExp.pdf}%\put(-0,85){C}
 	\end{overpic}
@@
 254 cm$^2$ circle, 
-and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
+and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.
+The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
 
@@
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
-	\begin{overpic}[width=\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
+	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
 %	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
@@
 \caption{\label{fig:realMean}
-Regulating average $x$ position of 100 kilobots using control law \eqref{eq:PDcontrolPosition}.
+Regulating mean $x$ position of 100 kilobots using control law \eqref{eq:PDcontrolPosition}.
 %Mean control plot with kilobots.
@@
 %\end{overpic}
-\begin{overpic}[width=.9\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
+\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
 \put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}

### 2016-08-16 journalSwarmControl/theory.tex
added Mable's edits
Javier's comments fixed
Changed format

@@
 Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
+Without loss of generality, assume $m=1$.
 Our state-space representation is:
@@
 One option is to control the position and velocity of the $j^{th}$ robot.
-To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and velocity of the $n$ robots:
+To find a potentially more useful option, we create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
 \begin{align}
@@
 \dot{x}_{n}
-\end{bmatrix} 
-+ \frac{1}{n}\begin{bmatrix}
+\end{bmatrix} \\
+&+ \frac{1}{n}\begin{bmatrix}
 0& 0&  \ldots &0& 0 \\
@@
 \end{equation}
-Thus the average position and average velocity are controllable.
+Thus the mean position and mean velocity are controllable.
 
@@
 \end{align}
-here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.
+ $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.
 
@@
 \Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
-\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$  \Comment{initially not in variance control mode}
+\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ % \Comment{initially not in variance control mode}
 \State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
@@
 \centering
-\begin{overpic}[width = 0.5\columnwidth]{hysteresis.pdf}
+\begin{overpic}[width = 0.8\columnwidth]{hysteresis.pdf}
 \put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
 \put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
-\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
+%\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
 \vspace{-0.5em}
-\caption{\label{fig:hysteresis} (Left) Two states to control mean and variance of a robot swarm. 
-(Right) Switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The plot uses robot radius $r=1/10$.
+\caption{\label{fig:hysteresis}  Hysteresis to control swarm mean and variance. 
+%(Right) Switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The plot uses robot radius $r=1/10$.
 %\vspace{-2em}
@@
 A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
-The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:hysteresis}:
-\begin{align}
+The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values: 
+%shown in Fig.~\ref{fig:hysteresis}:
+\begin{align} \label{eq:sigMaxMin}
  \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\

### 2016-08-16 journalSwarmControl/realExperiment.tex
fixed some bib entries

@@
 and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.
-The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.
+The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2016blocksimulations}.
 

### 2016-08-16 journalSwarmControl/simulation.tex
Changed format
fixed some bib entries

@@
 All experiments ran on a Chrome web browser on a 2.6 GHz Macbook.
-\href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2015}.
+\href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
 
@@
 \centering
-\begin{overpic}[width = 0.7\columnwidth ]{gains.eps}
+\begin{overpic}[width = \columnwidth ]{gains.pdf}
 \end{overpic}
@@
 \centering
-\begin{overpic}[width =0.7 \columnwidth] {brownianWpublish.eps}
+\begin{overpic}[width = \columnwidth] {brownianWpublish.eps}
 \end{overpic}
@@
 \centering
-\begin{overpic}[width =0.7\columnwidth]{meanVariance4.eps}
+\begin{overpic}[width = \columnwidth]{meanVariance4.eps}
 \end{overpic}

### 2016-08-17 journalSwarmControl/relatedWork.tex
added ayanian

@@
  \subsection{Human-swarm interaction}
-\cite{Jr2004} studied human \emph{fanout}, the number of robots a single human user could control.  %Studied as a key aspect of human-robot interaction, 
+
+ \cite{ayanian2014controlling} designed a user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism.
+Our paper shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.
+
+\cite{Jr2004} studied human \emph{fanout}, the number of robots a single human user could directly control.  %Studied as a key aspect of human-robot interaction, 
 They postulated that the optimal number of robots was approximately  the autonomous time  divided by the interaction time required by each robot.
@@
 
-\cite{Squire:2006:HCM:1121241.1121248} designed experiments showing user-interface design had a high impact on the task effectiveness and the number of robots that could be controlled simultaneously in a multi-robot task.
+%\cite{Squire:2006:HCM:1121241.1121248} designed experiments showing user-interface design had a high impact on the task effectiveness and the number of robots that could be controlled simultaneously in a multi-robot task.
 

### 2016-08-18 journalSwarmControl/experiment.tex
fixed some bib entries
Mable's Comments changed

@@
 The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
-The controller in \cite{Shahrokhi2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object.
+The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object.
 This failure rate increases if  object weight increases or ground-robot friction increases.
@@
 \begin{center}
-	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
+	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-4,25){A}\end{overpic}\\
 	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true

### 2016-08-18 journalSwarmControl/SwarmControlIJRR.tex
Added swarm to keywords
added ayanian
shortened section names

@@
 %%%%%%%%%%%% END IJRR SETTINGS
+
+% How many words:  $ detex IJRRensembleDifferentialDrive.tex | wc -w
 
@@
 
-\keywords{micro-robot, global control, human-swarm interaction}
+\keywords{micro-robot, global control, swarm, human-swarm interaction}
 
@@
 The Authors declare that there is no conflict of interest.
-To typeset a "Declaration of
-  conflicting interests" section.
 \end{dci}
@@
 \begin{sm}
-Video 1:  
-Video 2: 
+\begin{itemize}
+\item Video 1: hardware experiments of object manipulation with a swarm of 100 kilobots.
+\item Video 2:  simulation of parameter sweeps with swarm of robots manipulating objects.
+\end{itemize}
 \end{sm}

### 2016-08-18 journalSwarmControl/realExperiment.tex
changed position of overpic
shortened section names

@@
 \begin{center}
-	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
+	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(6,30){\emph{t}= 90 s}\put(55,30){\emph{t}= 175 s}\end{overpic}
 %	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
@@
 
-\subsection{Automated object manipulation (hardware experiment)}
+\subsection{Automated object manipulation (hardware~experiment)}
 
@@
 A square object completed in 6871 s. 
-\todo{A triangular object completed in ??:?? minutes.}
+Three equilateral triangular objects of varying sizes failed in a total of seven runs.
+Failures occurred when the object was pushed into a corner, requiring torque to be unstuck.
+This paper focuses on force, but not torque control.
+Swarm torque control is the subject of ongoing research in \cite{Shahrokhi2016CASE}.
 
@@
 \begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
-\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
+\put(15,8){\emph{t} = 1374 s}\put(48,8){\emph{t} = 2185 s}\put(80,8){\emph{t} = 2703 s}
 \end{overpic}

### 2016-08-18 journalSwarmControl/theory.tex
shortened section names

@@
 \end{equation}  
-\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
+\subsection{Controlling the mean position}\label{sec:controlMeanPosition}
 This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis.
@@
 
-\subsection{Controlling the variance of many robots}\label{sec:VarianceControl}
+\subsection{Controlling the variance}\label{sec:VarianceControl}
 
@@
 
-\subsection{Controlling both mean and variance of many robots}
+\subsection{Controlling both mean and variance}
 

### 2016-08-18 journalSwarmControl/Introduction.tex
Made fig 1 smaller, no captions
Changed format
edited the last edits

@@
 
-Our previous work with over a hundred hardware robots and thousands of simulated robots~\cite{Becker2013b} demonstrated that direct human control of large swarms is possible. 
+Our previous work with over a hundred hardware robots and thousands of simulated robots demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
 Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
@@
 \begin{figure}
-\renewcommand{\figwid}{0.32\columnwidth}
-\subfloat[][Vary Number]{\label{fig:VaryNum}
-\begin{overpic}[width =\figwid]{VaryNum.pdf}\end{overpic}}
-%
-\subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
-%\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\end{overpic}}
-%
-\subfloat[][Vary Noise]{\label{fig:VaryNoise}
-\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}\\
-%
-\subfloat[][Control Position]{\label{fig:ControlPos}
-\begin{overpic}[width =\figwid]{ControlPos.pdf}\end{overpic}}
-%
-\subfloat[][Vary Control: Assembly]{\label{fig:VaryControl}
-\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
-%
-\subfloat[][Vary Control: Foraging]{\label{fig:Forage}
-\begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
-%
+\centering
+\renewcommand{\figwid}{0.3\columnwidth}
+\begin{overpic}[width =\figwid]{VaryNum.pdf}	\put(10,80){\textbf{a} }\end{overpic}~
+\begin{overpic}[width =\figwid]{VaryVisMV.pdf}	\put(10,80){\textbf{b} }\end{overpic}~
+\begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
+\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
+\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
+\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
 \caption{\label{fig:5experiments}
@@
 \textbf{(f)} Comparing 3 control architectures to forage.
-\vspace{-2em}
 }
@@
 
-Our goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size.
+%\begin{figure}
+%\renewcommand{\figwid}{0.32\columnwidth}
+%\subfloat[][Vary Number]{\label{fig:VaryNum}
+%\begin{overpic}[width =\figwid]{VaryNum.pdf}\end{overpic}}
+%%
+%\subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
+%%\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
+%\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\end{overpic}}
+%%
+%\subfloat[][Vary Noise]{\label{fig:VaryNoise}
+%\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}\\
+%%
+%\subfloat[][Control Position]{\label{fig:ControlPos}
+%\begin{overpic}[width =\figwid]{ControlPos.pdf}\end{overpic}}
+%%
+%\subfloat[][Vary Control: Assembly]{\label{fig:VaryControl}
+%\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
+%%
+%\subfloat[][Vary Control: Foraging]{\label{fig:Forage}
+%\begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
+%%
+%\caption{\label{fig:5experiments}
+%Screenshots from our online experiments controlling multi-robot systems with limited, global control.
+%\textbf{(a)} Varying the number of robots from 1-500
+%\textbf{(b)} Comparing 4 levels of visual feedback 
+%\textbf{(c)} Varying noise from 0 to 200\% of control authority
+%\textbf{(d)} Controlling the position of 1 to 10 robots
+%\textbf{(e)} Comparing 3 control architectures to assemble
+%\textbf{(f)} Comparing 3 control architectures to forage.
+%\vspace{-2em}
+%}
+%\end{figure}
+
+The goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size.
 Towards this end, we created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}, an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
 Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.
-\href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are posted online at \cite{Chris-Ertel2016}.
+\href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are online at \cite{Chris-Ertel2016}.
 
@@
 \centering
-\begin{overpic}[width=0.5\columnwidth]{CompleteMazeDescription.pdf}\end{overpic}
-\begin{overpic}[width=0.438\columnwidth]{MainExpFig.pdf}\end{overpic}
+%\begin{overpic}[height=6cm]{CompleteMazeDescription.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
 %\todo{I like the 'target' symbol, but it is not self-documenting.
@@
 \caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.
-Here a swarm of simple robots (blue discs) pushes a green hexagon toward the goal.
-Simulation at left, screen capture from hardware experiment at right.
-See multimedia extension.}
+Here a swarm of robots pushes a green hexagon toward the goal.
+%Simulation at left, screen capture from hardware experiment at right. 
+See video 1 in supplemental materials.}
 \end{figure}
@@
 We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.
- We conclude with implementations of these controllers in our hardware robots and use them to complete an object manipulation task with 100+ kilobots in \S \ref{sec:realExperiment}.
+ We conclude with implementations of these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}.
 
  This paper is the synthesis of two preliminary conference papers.
-\cite{swarmcontrol2013}, covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015}, presented simulations of object manipulation.
+\cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
 This paper presents the final results from SwarmControl.net.

### 2016-08-18 journalSwarmControl/experiment.tex
edited the last edits

@@
 The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
-This approach using workspace BFS fails to account for the hull of the object and will suggest moves that can cause at least partial collision with the workspace.
+This approach using workspace BFS fails to account for the hull of the object and will suggest moves that can cause collisions with the workspace.
 A configuration-space BFS approach avoids that problem but still fails to model uncertain actuation of the object by the swarm.
@@
 \end{align}
- $\mathbf{u}$ is the action and $x$ is the current state. 
-  Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
+ $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
+  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$.
+The optimal policy is the control action that, in probability, results in the largest value:
    \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
@@
 To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}.
+The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}b.
 \begin{align}
@@
 In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$.
-Because the kilobot hardware experiments have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
+Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
 
@@
 Robots that cannot reach the object due to obstacles skew these calculations.
-The state machine in Fig.\ \ref{fig:Region}A solves this problem by creating two states for the maze: either main or transfer.
+The state machine in Fig.\ \ref{fig:Region}a solves this problem by creating two states for the maze: either main or transfer.
 Each state has a set of regions representing a discretized visibility polygon.
@@
 
-Fig.\ \ref{fig:Region}B shows the regions for the main state.
+Fig.\ \ref{fig:Region}b shows the regions for the main state.
 The object is in region 1.
@@
 When the object leaves main region 1 the state switches to transfer.
-The transfer regions are shown in Fig.\ \ref{fig:Region}C.
+The transfer regions are shown in Fig.\ \ref{fig:Region}c.
 The object is in transfer region 1, so only robots in transfer region 1 are included in the mean and covariance calculations.
@@
 \begin{center}
-	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-4,25){A}\end{overpic}\\
+	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
 	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
-	\put(-6,85){B}
+	\put(-6,85){\textbf{b}}
 	\end{overpic}
@@
 	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}
-	\put(-6,85){C}
+	\put(-6,85){\textbf{c}}
 	\end{overpic}
@@
 \subsection{Algorithm for object manipulation}\label{sec:AlgObjectManipulation}
- We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit.
-$0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
+ We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
  The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.
@@
 %\end{overpic}}
-\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}
-\end{overpic}
-\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.
-%\vspace{-2em}
+\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
+\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control. See supplemental material, video 2.
 }
@@
 \begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
-
 \end{overpic}

### 2016-08-18 journalSwarmControl/realExperiment.tex
edited the last edits

@@
 \end{overpic}
-\vspace{-1em}
-\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control.
-The automatic controller classifies pink objects as obstacles and generates a policy to the goal.
-See  video in multimedia extension.}
-%\vspace{-2em}
-}
+%\vspace{-1em}
+\caption{\label{fig:expSnapShot}Snapshots showing object manipulation experiment with 100 kilobots under automatic control.
+The automatic controller generates a policy to the goal.
+See  supplementary material video 1.}
         \end{figure}

### 2016-08-18 journalSwarmControl/SwarmControlIJRR.tex
edited the last edits
changed order of video numbers

@@
 %%%%%%%%%%%% END IJRR SETTINGS
-
+\def\journalname{The International Journal of Robotics Research}
 % How many words:  $ detex IJRRensembleDifferentialDrive.tex | wc -w
@@
 \begin{itemize}
-\item Video 1: hardware experiments of object manipulation with a swarm of 100 kilobots.
-\item Video 2:  simulation of parameter sweeps with swarm of robots manipulating objects.
+\item Video 1:  simulation of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
+\item Video 2: hardware experiments of object manipulation with a swarm of 100 kilobots.
 \end{itemize}

### 2016-08-19 journalSwarmControl/SwarmControlIJRR.tex
Convex Hall was not correct
The submitted version.

@@
 \end{funding}
+\newpage
 \begin{sm}
 \begin{itemize}
-\item Video 1:  simulation of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
-\item Video 2: hardware experiments of object manipulation with a swarm of 100 kilobots.
+\item Extension 1:  simulation video of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
+\item Extension 2: hardware experiment videos of object manipulation with a swarm of 100 kilobots.
 \end{itemize}

### 2016-08-26 RA-L/experiment.tex
Added RA-L with the first comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiment}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[!htb]
\centering
\renewcommand{\figwid}{0.4\columnwidth}
{
\begin{overpic}[width =0.415\columnwidth]{twoR_1.pdf}\put(15,65){$t$  = 30 s}\end{overpic}\hspace{-.5em}
\begin{overpic}[width =\figwid]{twoR_2.jpg}\put(15,65){$t$  = 60 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR_3.jpg}\put(15,65){$t$  = 90 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR_4.jpg}\put(15,65){$t$  = 120 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR_5.jpg}\put(15,65){$t$  = 150 s}
\end{overpic}}
\vspace{-2em}
\caption{\label{fig:storyReal}{Two robot positioning of two kilobot robots.  The boundary walls have nearly infinite friction, so the blue robot is stopped by the wall from $t = 30$s until the commanded input is directed away form the wall at $t=120$s, while the orange robot in free-space is unhindered.}
%\vspace{-2em}
}
\end{figure*}

%\subsection{Hardware System}

Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.  These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, from \citet{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open source platform or commercially from~\citet{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and eight 30W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace. The lights were controlled using an Arduino Uno board connected to an 8-relay shield.  Above  the table, an overhead machine vision system tracks the position of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:setup}
Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system.
}
\vspace{-1em}
\end{figure}

The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border and the three-legged design of the kilobots. When a kilobot is steered into the zigzag border, they pin themselves to the wall unless the global input directs them away from the wall.  This wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.

To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with a single light source, using friction with the boundary walls to vary the covariance from  -5000 to 10,000 pixel$^2$.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{experiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering 100 kilobot robots to desired covariance. The goal covariance is shown in green, the actual covariance in red. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

### 2016-08-26 RA-L/intro.tex
Added RA-L with the first comments

\section{Introduction}\label{sec:Intro}
Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Donald2013}, \citet{Ghosh2009}, \citet{kim2015imparting}, \citet{qiu2015magnetic}, or \citet{martel2014computer}.
Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection  and 2) ex vivo to assemble structures in parallel. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite high under-actuation.

%\begin{figure}
%\centering
%\begin{overpic}[width=0.9\columnwidth]{Leaf.jpg}\end{overpic}
%\caption{\label{fig:vascularNetwork}
%Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.  Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging
%due to the many obstacles. 
%This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.} 
%\end{figure}

\begin{figure}
\centering
\begin{overpic}[width=0.95\columnwidth]{IntroSwarmPic.pdf}\end{overpic}
\caption{\label{fig:IntroPic}
Swarm of kilobots programmed to move toward the brightest light source. The current covariance ellipse and mean are shown in red, the desired covariance is shown in green.  Navigating a swarm using global inputs is challenging because each member receives the same control inputs. 
This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.} 
\end{figure}

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\cite{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Even without obstacles or boundaries, the mean position of a swarm is controllable.  By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable. 
%The usefulness of these techniques was recently demonstrated by several automatic controllers. One controller steered a swarm of robots to push a larger block through a 2D maze~\citet{ShahrokhiIROS2015}. 
A limitation is that global control can only compress a swarm orthogonal to obstacles.  One implication is that a swarm in an axis-aligned rectangular workspace can not generate a non-zero covariance. This limitation is detrimental to desired applications because the ability to orient the swarm is often useful for navigating narrow passages.
%This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
%Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$.

After a review of recent related work (Section \ref{sec:RelatedWork}), this paper provides analytical position control results in two canonical workspaces with frictionless walls (Section \ref{subsec:FluidInTank}).  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, we introduce wall friction to the system model (Section \ref{subsec:WallFriction}).  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), and extend this proof to prove that a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace (Section \ref{sec:PostionControlnRobots}).
(Section \ref{sec:simulation}) describes implementations of both position control algorithms in simulation and  (Section \ref{sec:expResults}) describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}. We end with directions for future research (Section \ref{sec:conclusion}).
%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2016-08-26 RA-L/simulation.tex
Added RA-L with the first comments

\section{Simulation}\label{sec:simulation}

Two simulations were implemented using wall-friction for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.  All code is available online at [link withheld for review].

Two additional simulations were performed using wall-friction to control variance and covariance.  The first is an open-loop algorithm that demonstrates the effect of varying friction levels.  The second uses a closed-loop controller to achieve desired variance and covariance values.

\subsection{Position Control of Two Robots}

Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, \ref{alg:YControl}, were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  this algorithm for two configurations. 
Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.  Dashed lines show the shortest route if robots could be controlled independently.  The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
Each row has five snapshots taken every quarter second. For the sake of brevity axis-aligned moves were replaced with oblique moves that combine two moves simultaneously. 
 $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
 The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
 $\Delta r_y$ is corrected by $t = 0.75$. 
 Finally, the algorithm moves the robots to their corresponding destinations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
Similarly,  two iterations of Alg. \ref{alg:YControl} are only required if $|\Delta e_y - \Delta s_y|>L$.

%\textcolor{red}{Shiva:  what is the complexity of this algorithm?  How many steps in the worst case?  When does worst case happen?}
%\begin{figure*}
%\centering
%\renewcommand{\figwid}{0.4\columnwidth}
%{\begin{overpic}[width =\figwid]{one_1.jpg}\put(45,75){$t$ = 0 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{one_2.jpg}\put(45,75){$t$ = 0.25 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{one_3.jpg}\put(45,75){$t$  = 0.5 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{one_4.jpg}\put(45,75){$t$  = 0.75 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{one_5.jpg}\put(45,75){$t$  = 1 s}
%\end{overpic}}
%\vspace{-1em}
%\caption{\label{fig:shapeControlMathematica2}{Two robot positioning: switching positions using walls with infinite friction.}
%%\vspace{-2em}
%}
%\end{figure*}

\subsection{Position Control of $n$ Robots}
Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width.
Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
The distance moved grows quadratically with the number of robots $n$. A best-fit line $210 n^2 + 1200n-10,000$ is overlaid by the data..

In Fig.~\ref{fig:4diagramsplots.pdf}, the amount of clearance is $\epsilon=1$.
Control performance is sensitive to the desired clearance.  As $\epsilon$ increases, the total distance decreases asymptotically, as shown in Fig.~\ref{fig:graphrobo.pdf}, because the robots have more room to maneuver and fewer $\operatorname{DriftMove}$s are required.
%We observed that the value of ‘?’, greatly affects the total number of moves of commanded moves on the robots. The smaller the value of ‘?’ the greater is the number of moves required to move a said distance. The second graph plot(fig roboplot) is that of ‘?’ vs ‘total commanded moves’ which is the number of moves the global controller commands onto all the robots. Here we take the pattern 
%‘ROBO’ and calculate the ‘total commanded moves’ for ‘?’=[ 0.25,0.375,0.5,0.625,0.75,0.875,1]. We can notice an exponential fall in the total command moves. It can be inferred that there is a tradeoff between area of bounded region and the number of moves. For smaller ‘?’, smaller is the required area to implement algorithm but the longer is the time taken to complete the algorithm. Fig robo
%Unit of ‘?’ is the ratio of minimum clearance value to the body length of a robot.

%The total commanded distance is plotted, which is approximately -0.15v^3 + 385x^2-35000x+300000
%We compare the total distance moved and commanded  with the  \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.   Because all robots are interchangeable, the LAP distance reduces to \[ \text{LAP} =   \sum_{k=1}^n  \left| f_{k,x}-p_{k,x} \right| +  \left|  f_{k,y}-p_{k,y}  \right| . \]

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{5diagramsplots2.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:4diagramsplots.pdf}
The required number of moves under Alg. \ref{alg:PosControlNRobots}  using wall-friction to rearrange $n$ square-shaped 
robots.  %The plot compares  \emph{Total distance}---the sum of the moves made by every robot, with \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.  
See hardware implementation and simulation at [link withheld for review].
}
\end{figure}

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{graphrobo.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:graphrobo.pdf}
Control performance is sensitive to the desired clearance $\epsilon$.  As $\epsilon$ increases, the total distance decreases asymptotically.
}
\end{figure}

\subsection{Efficient Control of Covariance}
A set of simulations were conducted to demonstrate the importance of boundary friction.  These simulations use  the 2D physics engine Box2D, by \citet{catto2010box2d}.
 144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig. \ref{fig:SimCovarianceFuncFrictionOpenLoop}.  All robots had  the same initial conditions, but in four tests the boundary friction was $F_f = \{0,1/3 F, 2/3F, F\}$.
 Without friction, covariance has minimal variation.  As friction increases, the covariance can be manipulated to greater degrees.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{SimCovarianceFuncFrictionOpenLoopV2.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:SimCovarianceFuncFrictionOpenLoop}
Open-loop simulation with 144 disc robots and varying levels of boundary friction under the same initial conditions.  Without friction, covariance is unchangeable.  As friction increases, the covariance can be manipulated to greater degrees.
}
\end{figure}

 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in Section \ref{subsec:ClosedLoopCovarianceControl}. Fig. \ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ axis were controlled from a set of initial conditions.
\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.eps}
\end{center}
\vspace{-1em}
\caption{\label{fig:CovVarControlPlot}
Closed-loop simulation with 144 disc robots and three sets of initial conditions.  The algorithm tracks goal variance and covariance values (green).  The goal covariance switches sign every 30 s.}
\end{figure}

### 2016-08-26 RA-L/theory.tex
Added RA-L with the first comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=\linewidth]{SquareFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for five values of $A$.  The locus morphs from a square to a circle as $A$ increases.  The covariance ellipse for each $A$ is shown with a dashed line. Center shows two corresponding arrangements of kilobots.  At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values. See online interactive demonstration at [withheld for double-blind review].}
\label{fig:SquareFill}
\end{center}
\end{figure*} 
\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=\linewidth]{CircleFill.pdf} 
\vspace{-1em}
\caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$. Left plot shows locus of possible mean positions for four values of $h$. The locus of possible mean positions are concentric circles. See online interactive demonstration at [withheld for double-blind review].}
\label{fig:CircleFill}
\end{center}
\end{figure*}

\subsection{Controlling Covariance: Fluid Settling In a Tank}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
\paragraph{Square workplace}
This section examines the mean, variance,  covariance and correlation of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together cover an area of constant volume $A$. Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill}.

The range of possible angles for the global input angle $\beta $ is [0,2$\pi $). In this range of angles, the swarm assumes eight different polygonal shapes. The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.

Computing the means, variances, covariance, and correlation requires integrating over the area containing the swarm.  This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows, integrated over the unit square with $x$ and $y$ from 0 to 1:

\begin{equation}\label{eq:meanInSquareWorkspace}
\bar{x} =\frac{\iint x \bm{1}_A(x,y)dxdy}{A}
\text{, }
\bar{y}=\frac{\iint y \bm{1}_A(x,y)dxdy}{A}
\end{equation}

\begin{align}\label{eq:varInSquareWorkspace}
\sigma^2_x &=\frac{\iint \left(x-\bar{x}\right)^2 \bm{1}_A(x,y) dx dy}{A}, \\
\sigma^2_y &=\frac{\iint  \left(y-\bar{y}\right)^2 \bm{1}_A(x,y) dx dy}{A}
\end{align}

\begin{align}\label{eq:covAndcorrInSquareWorkspace}
\sigma_{xy} &= \frac{\iint  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \bm{1}_A(x,y) dx dy}{A},\\
\rho_{xy} &= \frac{\sigma^2_x}{\sigma_x\sigma_y}
\end{align}

Rather than using an indicator function, the region of integration can be changed to only include the polygon containing the swarm. If the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
\begin{align}\label{eq:meanInSquareWorkspaceLL}
\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} x \, dy\right) \, dx}{A} \nonumber \\
	&=\frac{1}{3} \sqrt{2} \sqrt{A \tan (\beta )}\\
\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} y \, dy\right) \, dx}{A} \nonumber\\
	&=\frac{1}{3} \sqrt{2} \sqrt{A \cot (\beta )}
\end{align}
The full equations are included in the appendix, and are summarized in Fig.~\ref{fig:SquareFill}. A few highlights are that the correlation is maximized when the swarm is in a triangular shape, and is $\pm$1/2. The covariance of a triangle is always $\pm(A/18)$. Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape. The range of mean positions are maximized when $A$ is small.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Circular workplace}
Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
Similar calculations can be computed for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
Under a global input, the robot swarm fills the region under a chord with area
\begin{align}
A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
\end{align}
For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
\begin{align}
r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
\end{align}
Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$. Results are summarized in Fig.~\ref{fig:CircleFill}.

%Smashing the swarm to the corner walls will cause some correlation between \emph{x} and \emph{y} axes. Eq. ~\ref{eq:Wall} shows the relation between number of robots and area provided and the correlation we can reach with it.
%%TODO: ADD the equation for showing the correlation Vs area, vs number of robots
% This equation shows that we are not able to reach high correlations bigger than $|0.5|$. Fig. ~\ref{fig:wallCorrelation} shows the different mean positions and correlations we are able to reach. There are ways and corridors that the swarm needs to pass and also to keep the majority of it. If we use the wall corners technique and the way needs a very high correlation to pass, a remarkable number of the swarm members will be lost, and we may lose track of them due to their small size. Meanwhile, the variance of the swarm would get bigger and bigger that it would not anymore be able to deliver sufficient payloads. For this reason we have thought another idea which uses \emph{friction} to make correlations.
%

\subsection{Friction with Boundary Wall}\label{subsec:WallFriction}
Global inputs move a swarm uniformly.  
Controlling covariance requires breaking this uniform symmetry.  A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. Directly controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  An obstacle in the lower-right corner is enough to generate positive covariance.  Generating both positive and negative covariance requires additional obstacles.  Requiring special obstacle configuration also makes covariance control dependent on the local environment. 
  Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  Robots touching a wall have a negative friction force that opposes movement along the boundary.  This  causes robots along the boundary to slow down compared to robots in free-space.

Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary. $N$ is the normal or perpendicular force between the robot and the boundary. The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $|\theta| < \pi/2$. The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by

\begin{align}
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
\text{and } N &= F \cos(\theta) \nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. As illustrated, both experiences different net forces although each receives the same inputs.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, if the control input is parallel or into the wall, the touching robot will not move. 
The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls are sufficient to arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot.}
\label{fig:friction}
\end{center}
\end{figure}

\subsection{Position Control of $2$ Robots Using Wall Friction}\label{sec:PostionControl2Robots}
\begin{figure*}
\centering
\renewcommand{\figwid}{0.4\columnwidth}
{\begin{overpic}[width =\figwid]{one_1.jpg}\put(45,75){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{one_2.jpg}\put(45,75){$t$ = 0.25 s}
\end{overpic}
\begin{overpic}[width =\figwid]{one_3.jpg}\put(45,75){$t$  = 0.5 s}
\end{overpic}
\begin{overpic}[width =\figwid]{one_4.jpg}\put(45,75){$t$  = 0.75 s}
\end{overpic}
\begin{overpic}[width =\figwid]{one_5.jpg}\put(45,75){$t$  = 1 s}
\end{overpic}}\\
%{\begin{overpic}[width =\figwid]{two_1.jpg}\put(45,75){$t$  = 0 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{two_2.jpg}\put(45,75){$t$  = 0.25 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{two_3.jpg}\put(45,75){$t$  = 0.5 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{two_4.jpg}\put(45,75){$t$  = 0.75 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{two_5.jpg}\put(45,75){$t$  = 1 s}
%\end{overpic}}\\
\vspace{.5em}
{\begin{overpic}[width =\figwid]{S3_1.jpg}\put(45,75){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S3_2.jpg}\put(45,75){$t$ = 0.25 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S3_3.jpg}\put(45,75){$t$  = 0.5 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S3_4.jpg}\put(45,75){$t$  = 0.75 s}
\end{overpic}
\begin{overpic}[width =\figwid]{S3_5.jpg}\put(45,75){$t$  = 1 s}
\end{overpic}}
\vspace{-1em}
\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. The algorithm only requires friction along the bottom and left walls.
Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.  Dashed lines show the shortest route if robots could be controlled independently.  The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
The bottom row shows an extreme case where the robots must switch position.
}
%\vspace{-2em}
}
\end{figure*}

This section describes Alg.~\ref{alg:PosControl2Robots}, an algorithm that uses wall-friction to arbitrarily position two robots in a rectangular workspace.  This algorithm  introduces concepts that will be used for multi-robot positioning. It only requires collisions with two orthogonal walls, in this case, the bottom and left walls. Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.

Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
Denote the current positions of the robots  $r_1$ and $r_2$. 
Subscripts $_x$ and $_y$ denote the $x$ and $y$ coordinates, i.e., $s_{1x}$ and $s_{1y}$ denote the $x$ and $y$ locations of $s_1$. 
The algorithm assigns a global control input at every instance.
The goal is to adjust 
 $\Delta r_x = r_{2x}-r_{1x}$ from $\Delta s_x = s_{2x}-s_{1x}$ to $\Delta e_x = e_{2x}-e_{1x}$ and  adjust 
 $\Delta r_y = r_{2y}-r_{1y}$ from $\Delta s_y = s_{2y}-s_{1y}$ to $\Delta e_y = e_{2y}-e_{1y}$ using a shared global control input. 
 This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
 %employing the assumption we have made earlier about the walls' friction.

Our algorithm solves the positioning problem in two steps: 
First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
Second $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant in Alg.~\ref{alg:YControl}.

Step (i): Fixing $\Delta r_x$
\label{theory:step1}
\begin{enumerate}
\item
If $\Delta e_x$ is negative, the robots will be commanded to move toward the left wall and halt $\epsilon$ distance from the left wall.
If $\Delta e_x \ge 0$, the robots will be commanded to move toward the right wall and halt $\epsilon$ distance from the right wall. The epsilon distance prevents the robots from experiencing friction along the vertical wall.
\item  let $y_{\min} = \min_i r_{iy}$, i.e., $y_{\min}$ be the minimum height of the two robots. Move both robots downward  $y_{\min}$ such that one of the robots touches the bottom wall and hence friction force will prevent this robot from moving left or right.
\item friction force holds the lower robot in place while the upper robot may move right and left freely to change $\Delta r_x $ to $\Delta e_x$.
\item  If after the free move of the upper robot $\Delta r_x$ is not $\Delta e_x$, Step (i) will be repeated.  No more than two iterations of step (i) are required.
\end{enumerate}

Step (ii): Fixing $\Delta r_y$
Now that $\Delta r_x$ is corrected, $\Delta r_y$ must be corrected:
\begin{enumerate}
\item 
If $\Delta e_y$ is negative, the robots will be commanded to move toward the top wall and halt $\epsilon$ distance from the top wall.
If $\Delta e_y \ge 0$, the robots will be commanded to move toward the bottom wall and halt $\epsilon$ distance from the bottom wall. The epsilon distance prevents the robots from experiencing friction along the horizontal wall.
\item  let $x_{\min} = \min_i r_{ix}$, i.e., $x_{\min}$ be the minimum distance of the two robots from the origin along the $x$-axis. Move both robots to the left $x_{\min}$ such that one of the robots  touches the left wall and hence friction force will prevent this robot from moving up or down.
\item  friction force holds the left robot in place while the right robot may move up or down freely to change $\Delta r_y $  to $\Delta e_y$.
\item  If after the free move of the right robot $\Delta r_y$ is not $\Delta e_y$, Step (ii) will be repeated.  No more than two iterations of step (ii) are required.
\end{enumerate}
Once $\Delta r_x$ and $\Delta r_y$ are set to $\Delta e_x$ and $\Delta e_y$, a straight-line global input moves both robots from $r_1$ and $r_2$ to $e_1$ and $e_2$.

\begin{algorithm}
\caption{WallFrictionArrange2Robots($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\Require starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. 
 Current position of the robots are $(r_1,r_2)$.

\State ($r_1,r_2$) = GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
\State GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)

\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $\epsilon \gets $ small number
\State $ \Delta s_x  \gets s_{1x} - s_{2x} $
\State $ \Delta e_x \gets e_{1x} - e_{2x} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_x < 0 $ }
\State $ m \gets ( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Move to right wall}
\Else 
\State  $ m \gets ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
\State $ m \gets (\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
\Else
\State $ m \gets (-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move up}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_x = r_{1x} - r_{2x}$
\If {$\Delta r_x \equiv \Delta e_x$} 
%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
\State $ \Delta e_y \gets e_{1y} - e_{2y} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_y < 0 $ }
\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
\Else 
\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
\Else
\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_y = r_{1y} - r_{2y}$
\If {$\Delta r_y \equiv \Delta e_y$} 
\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints. The solution described here is an iterative procedure with $n$ loops. The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}. At the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone. See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:driftmove}
A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. The robot $A$ touching a top wall will move right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
}
\end{figure}
\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{PositionNrobots.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:construction2d}
Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
}
\end{figure}

Assume an open workspace with four axis-aligned walls with infinite friction.
The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots. Without loss of generality, assume the build zone  is above the staging zone. 
Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.  The minimum workspace is then $(\epsilon + 2r + \max(w_f,w_s), 2\epsilon + h_s,h_f)$.

The $n$ robot position control algorithm relies on a $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ control input, shown in Fig.\  \ref{fig:driftmove}.
A drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. The robot touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.

Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot. Label the robots in the staging zone from left-to-right and top-to-bottom, and the $f_k$ configurations right-to-left and top-to-bottom as shown in Fig.~\ref{fig:construction2d}.

\begin{algorithm}
\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
\begin{algorithmic}[1]
\State move( $-\epsilon, r-p_{k,y}$) % move  away from right wall and down till robot k touches bottom

\While{ $p_{k,x} > r$} 
\State $\operatorname{DriftMove}(\epsilon, \min(p_{k,x} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
\EndWhile

\State $m \gets \operatorname{ceil}(\frac{f_{k,y}-r}{\epsilon})$
\State $\beta \gets \frac{f_{k,y}-r}{m}$
\State $\alpha \gets \beta - \frac{r - p_{k,y}-\epsilon}{m}$
\For{ $m$ iterations}
\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
\EndFor

\State move($r+\epsilon-f_{k,x}, 0$)  % move the group to the left until k is in the correct relative x position
\State move($f_{k,x}-r, 0$)

\end{algorithmic}
\end{algorithm}

Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
Second, a set of $\operatorname{DriftMove()}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
Third, a set of $\operatorname{DriftMove()}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.

Finally, all robots are moved right until robot $k$ is in the desired target position.

\subsection{Controlling Covariance Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
Assume an open workspace with infinite boundary friction. Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy}$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$. For our experiments, $c_1 = 0.1$.
%$\sigma_x^2$ decreases when the swarm is pushed into vertical walls, and $\sigma_y^2$ decreases when the swarm is pushed into horizontal walls. Alg.~\ref{alg:CovCont} procedes as follows:
\begin{enumerate}
\item swarm is pushed into the left wall until $\sigma_x^2< c_1\sigma_{goalx}^2$.  
\item swarm's mean position is moved to the center of the workspace
\item swarm is pushed into the bottom wall until $\sigma_y^2 \le \sigma_{goalt}^2$. 
\item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\
else swarm slides left until $\sigma_{xy} \le \sigma_{goalxy}$ 
\item  swarm's mean position is moved to the center of the workspace
\end{enumerate}

%\begin{algorithm}
%\caption{Swarm Covariance Control With Wall Friction}\label{alg:CovCont}
%\begin{algorithmic}[1]
%\Require  goal variances and covariance and mean, variances and covariance of the swarm  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy}, \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$. 
%%Swarm is Gaussian distributed (in limit it is uniformly distributed), 
%$L_w$ is width of the walls and $L_h$ is the height of the wall. $(0,0)$ is top right corner. $\epsilon$ is the error for reaching covariance.
%\If {$\sigma_x^2 > \sigma_{goalx}^2/10 $ }
%\State $ \bar{x} \gets \sigma_{goalx}/10   $ \Comment{Move to left wall}
%\State $MoveToCenter \gets true$
%\Else
%\If {$MoveToCenter$ }
%\State   $ \bar{x} \gets L_w/2 $, $\bar{y} \gets L_h/2 $ \Comment{Move to the center}
%\State $MoveToBottom \gets true$
%\State $MoveToCenter \gets false$
%\Else 
%\If {$MoveToBottom $}
%\State $ \bar{y} \gets L_h - \sigma_{goaly}   $ \Comment{Move to bottom}
%\State $MoveToBottom \gets false$
%\Else
%\If {$\sigma_{xy} -\sigma_{goalxy} > \epsilon$}
%\While  {$\sigma_{xy} -\sigma_{goalxy} > \epsilon$ }
%\State $ \bar{x} \gets \bar{x} + \epsilon  $ \Comment{Move to right wall}
%\EndWhile
%\Else
%\If {$\sigma_{xy} -\sigma_{goalxy} < \epsilon$ }
%\While  {$\sigma_{xy} -\sigma_{goalxy} < \epsilon$ }
%\State $ \bar{x} \gets \bar{x} - \epsilon  $ \Comment{Move to left wall}
%\EndWhile
%\EndIf
%\EndIf
%\EndIf
%\EndIf
%\EndIf
%\State   $ \bar{x} \gets L_w/2 $, $\bar{y} \gets L_h/2 $ \Comment{Move to the center}
%
%\end{algorithmic}
%\end{algorithm}

### 2016-08-26 RA-L/intro.tex
applied some review comments

@@
 Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Donald2013}, \citet{Ghosh2009}, \citet{kim2015imparting}, \citet{qiu2015magnetic}, or \citet{martel2014computer}.
-Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection  and 2) ex vivo to assemble structures in parallel.
-%Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} .
+Someday large swarms of robots will be remotely guided
+ 1)  ex vivo to assemble structures in parallel and 
+2) through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
 Instead, these robots are often controlled by a global, broadcast signal.
- The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite high under-actuation.
+These applications require control techniques that can reliably exploit large populations despite high under-actuation.
 
@@
 \caption{\label{fig:IntroPic}
-Swarm of kilobots programmed to move toward the brightest light source.
+Swarm of kilobots programmed to move toward the brightest light source as explained in \S \ref{sec:expResults}.
 The current covariance ellipse and mean are shown in red, the desired covariance is shown in green.
@@
 One controller steered a swarm of robots to push a larger block through a 2D maze~\citet{ShahrokhiIROS2015}.
-A limitation is that global control can only compress a swarm orthogonal to obstacles.
-One implication is that a swarm in an axis-aligned rectangular workspace can not generate a non-zero covariance.
-This limitation is detrimental to desired applications because the ability to orient the swarm is often useful for navigating narrow passages.
+A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.
 %This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
@@
 
-After a review of recent related work (Section \ref{sec:RelatedWork}), this paper provides analytical position control results in two canonical workspaces with frictionless walls (Section \ref{subsec:FluidInTank}).
+After a review of recent related work \S \ref{sec:RelatedWork}, this paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes, we introduce wall friction to the system model (Section \ref{subsec:WallFriction}).
-We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), and extend this proof to prove that a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace (Section \ref{sec:PostionControlnRobots}).
-(Section \ref{sec:simulation}) describes implementations of both position control algorithms in simulation and  (Section \ref{sec:expResults}) describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-We end with directions for future research (Section \ref{sec:conclusion}).
+To extend the range of possible shapes, we introduce wall friction to the system model \S \ref{subsec:WallFriction}.
+We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this proof to prove that a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
+\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
+We end with directions for future research \S \ref{sec:conclusion}.
 %For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 

### 2016-08-26 RA-L/experiment.tex
Reorganized the paper

@@
 \end{figure}
-
+\subsection{Hardware Experiment: Position Control of Two Robots}
 The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border and the three-legged design of the kilobots.
@@
 
+\subsection{Hardware Experiment: Position Control of n Robots}
+
+\subsection{Hardware Experiment: Control of Covariance}
 To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with a single light source, using friction with the boundary walls to vary the covariance from  -5000 to 10,000 pixel$^2$.

### 2016-08-26 RA-L/intro.tex
Reorganized the paper

@@
 \section{Introduction}\label{sec:Intro}
-Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Donald2013}, \citet{Ghosh2009}, \citet{kim2015imparting}, \citet{qiu2015magnetic}, or \citet{martel2014computer}.
+Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, or \citet{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided

### 2016-08-26 RA-L/simulation.tex
Reorganized the paper

@@
 
-Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, \ref{alg:YControl}, were implemented in Mathematica using point robots (radius = $0$).
-Fig.~\ref{fig:shapeControlMathematica1}  shows  this algorithm for two configurations.
+Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, %\ref{alg:YControl}, 
+were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  this algorithm for two configurations. 
 Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
@@
 In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
-Similarly,  two iterations of Alg. \ref{alg:YControl} are only required if $|\Delta e_y - \Delta s_y|>L$.
+Similarly,  two iterations %of Alg. \ref{alg:YControl} 
+are only required if $|\Delta e_y - \Delta s_y|>L$.
 
@@
 \subsection{Efficient Control of Covariance}
-A set of simulations were conducted to demonstrate the importance of boundary friction.  These simulations use  the 2D physics engine Box2D, by \citet{catto2010box2d}.
+%A set of simulations were conducted to demonstrate the importance of boundary friction.  
+Random disturbances impair the performance of Alg.
+\ref{alg:PosControl2Robots} and Alg.
+\ref{alg:PosControlNRobots}.
+Still, we are able to control covariance of the swarm.
+This section demonstrates simulations of controlling covariance of the swarm.
+These simulations use  the 2D physics engine Box2D, by \citet{catto2010box2d}.
  144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig.

### 2016-08-26 RA-L/theory.tex
applied some review comments
Reorganized the paper

@@
 
-\subsection{Controlling Covariance: Fluid Settling In a Tank}\label{subsec:FluidInTank}
+\subsection{Using Boundaries: Fluid Settling In a Tank}\label{subsec:FluidInTank}
 One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
@@
 This section examines the mean, variance,  covariance and correlation of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together cover an area of constant volume $A$.
+The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$.
 Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill}.
@@
 
-\subsection{Friction with Boundary Wall}\label{subsec:WallFriction}
+\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
 Global inputs move a swarm uniformly.  
@@
 
+\section{Algorithms}\label{sec:algorithms}
 \subsection{Position Control of $2$ Robots Using Wall Friction}\label{sec:PostionControl2Robots}
@@
 First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-Second $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant in Alg.~\ref{alg:YControl}.
+Second $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant.% in Alg.~\ref{alg:YControl}.
 
-Step (i): Fixing $\Delta r_x$
-\label{theory:step1}
-\begin{enumerate}
-\item
-If $\Delta e_x$ is negative, the robots will be commanded to move toward the left wall and halt $\epsilon$ distance from the left wall.
-If $\Delta e_x \ge 0$, the robots will be commanded to move toward the right wall and halt $\epsilon$ distance from the right wall.
+%Step (i): Fixing $\Delta r_x$
+%\label{theory:step1}
+%\begin{enumerate}
+%\item
+%If $\Delta e_x$ is negative, the robots will be commanded to move toward the left wall and halt $\epsilon$ distance from the left wall.
+%If $\Delta e_x \ge 0$, the robots will be commanded to move toward the right wall and halt $\epsilon$ distance from the right wall.
 The epsilon distance prevents the robots from experiencing friction along the vertical wall.
-\item  let $y_{\min} = \min_i r_{iy}$, i.e., $y_{\min}$ be the minimum height of the two robots.
+%\item  let $y_{\min} = \min_i r_{iy}$, i.e., $y_{\min}$ be the minimum height of the two robots.
 Move both robots downward  $y_{\min}$ such that one of the robots touches the bottom wall and hence friction force will prevent this robot from moving left or right.
-\item friction force holds the lower robot in place while the upper robot may move right and left freely to change $\Delta r_x $ to $\Delta e_x$.
-\item  If after the free move of the upper robot $\Delta r_x$ is not $\Delta e_x$, Step (i) will be repeated.  No more than two iterations of step (i) are required.
-\end{enumerate}
-
-Step (ii): Fixing $\Delta r_y$
-Now that $\Delta r_x$ is corrected, $\Delta r_y$ must be corrected:
-\begin{enumerate}
-\item 
-If $\Delta e_y$ is negative, the robots will be commanded to move toward the top wall and halt $\epsilon$ distance from the top wall.
-If $\Delta e_y \ge 0$, the robots will be commanded to move toward the bottom wall and halt $\epsilon$ distance from the bottom wall.
+%\item friction force holds the lower robot in place while the upper robot may move right and left freely to change $\Delta r_x $ to $\Delta e_x$.
+%\item  If after the free move of the upper robot $\Delta r_x$ is not $\Delta e_x$, Step (i) will be repeated.  No more than two iterations of step (i) are required.
+%\end{enumerate}
+%
+%Step (ii): Fixing $\Delta r_y$
+%Now that $\Delta r_x$ is corrected, $\Delta r_y$ must be corrected:
+%\begin{enumerate}
+%\item 
+%If $\Delta e_y$ is negative, the robots will be commanded to move toward the top wall and halt $\epsilon$ distance from the top wall.
+%If $\Delta e_y \ge 0$, the robots will be commanded to move toward the bottom wall and halt $\epsilon$ distance from the bottom wall.
 The epsilon distance prevents the robots from experiencing friction along the horizontal wall.
-\item  let $x_{\min} = \min_i r_{ix}$, i.e., $x_{\min}$ be the minimum distance of the two robots from the origin along the $x$-axis.
+%\item  let $x_{\min} = \min_i r_{ix}$, i.e., $x_{\min}$ be the minimum distance of the two robots from the origin along the $x$-axis.
 Move both robots to the left $x_{\min}$ such that one of the robots  touches the left wall and hence friction force will prevent this robot from moving up or down.
-\item  friction force holds the left robot in place while the right robot may move up or down freely to change $\Delta r_y $  to $\Delta e_y$.
-\item  If after the free move of the right robot $\Delta r_y$ is not $\Delta e_y$, Step (ii) will be repeated.  No more than two iterations of step (ii) are required.
-\end{enumerate}
-Once $\Delta r_x$ and $\Delta r_y$ are set to $\Delta e_x$ and $\Delta e_y$, a straight-line global input moves both robots from $r_1$ and $r_2$ to $e_1$ and $e_2$.
+%\item  friction force holds the left robot in place while the right robot may move up or down freely to change $\Delta r_y $  to $\Delta e_y$.
+%\item  If after the free move of the right robot $\Delta r_y$ is not $\Delta e_y$, Step (ii) will be repeated.  No more than two iterations of step (ii) are required.
+%\end{enumerate}
+%Once $\Delta r_x$ and $\Delta r_y$ are set to $\Delta e_x$ and $\Delta e_y$, a straight-line global input moves both robots from $r_1$ and $r_2$ to $e_1$ and $e_2$. 
+%
 
@@
 
-\begin{algorithm}
-\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
-\begin{algorithmic}[1]
-\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
- $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
-\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
-\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
-\State $ \Delta e_y \gets e_{1y} - e_{2y} $
-\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
-\If {$\Delta e_y < 0 $ }
-\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
-\Else 
-\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
-\EndIf
-\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
-\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
-\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
-\Else
-\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
-\EndIf 
-\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
-\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-\State $\Delta r_y = r_{1y} - r_{2y}$
-\If {$\Delta r_y \equiv \Delta e_y$} 
-\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
-\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-\State  \Return $(r_1,r_2)$
-\Else   
-\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
-\EndIf
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
+%\begin{algorithmic}[1]
+%\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
+%$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
+% $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
+%\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
+%\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
+%\State $ \Delta e_y \gets e_{1y} - e_{2y} $
+%\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
+%\If {$\Delta e_y < 0 $ }
+%\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
+%\Else 
+%\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
+%\EndIf
+%\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
+%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+%\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
+%\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
+%\Else
+%\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
+%\EndIf 
+%\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
+%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+%\State $\Delta r_y = r_{1y} - r_{2y}$
+%\If {$\Delta r_y \equiv \Delta e_y$} 
+%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
+%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+%\State  \Return $(r_1,r_2)$
+%\Else   
+%\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
+%\EndIf
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}.
-At the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
+All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
 See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.

### 2016-08-26 RA-L/intro.tex
Added a citation to Ou's work.

--- 2016-08-26 RA-L/intro.tex
@@
 \section{Introduction}\label{sec:Intro}
-Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Donald2013}, \citet{Ghosh2009}, \citet{kim2015imparting}, \citet{qiu2015magnetic}, or \citet{martel2014computer}.
+Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, \citet{Ou2013} or \citet{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided

### 2016-08-26 RA-L/experiment.tex
Corrected the length of the table

@@
 Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.
-In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and eight 30W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace.
+In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace.
 The lights were controlled using an Arduino Uno board connected to an 8-relay shield.
@@
 \caption{\label{fig:setup}
-Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system.
+Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight remotely triggered LED floodlights, with an overhead machine vision system.
 }

### 2016-08-30 RA-L/intro.tex
Changed the covariance Results!

@@
 
-After a review of recent related work \S \ref{sec:RelatedWork}, this paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.
+This paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.
 These results are limited in the set of shapes that can be generated.
@@
 \S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-We end with directions for future research \S \ref{sec:conclusion}.
+After a review of recent related work \S \ref{sec:RelatedWork}, we end with directions for future research \S \ref{sec:conclusion}.
 %For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 

### 2016-08-30 RA-L/experiment.tex
Changed the covariance Results!
Added the new two robot positioning figures.

@@
 {
-\begin{overpic}[width =0.415\columnwidth]{twoR_1.pdf}\put(15,65){$t$  = 30 s}\end{overpic}\hspace{-.5em}
-\begin{overpic}[width =\figwid]{twoR_2.jpg}\put(15,65){$t$  = 60 s}
+\begin{overpic}[width =\figwid]{twoR1.pdf}\put(12,70){$t$  = 0 s}\end{overpic}
+\begin{overpic}[width =\figwid]{twoR2.pdf}\put(12,70){$t$  = 60 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{twoR_3.jpg}\put(15,65){$t$  = 90 s}
+\begin{overpic}[width =\figwid]{twoR3.pdf}\put(12,70){$t$  = 150 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{twoR_4.jpg}\put(15,65){$t$  = 120 s}
+\begin{overpic}[width =\figwid]{twoR4.pdf}\put(12,70){$t$  = 160 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{twoR_5.jpg}\put(15,65){$t$  = 150 s}
+\begin{overpic}[width =\figwid]{twoR5.pdf}\put(12,70){$t$  = 210 s}
 \end{overpic}}
-\vspace{-2em}
-\caption{\label{fig:storyReal}{Two robot positioning of two kilobot robots.
-The boundary walls have nearly infinite friction, so the blue robot is stopped by the wall from $t = 30$s until the commanded input is directed away form the wall at $t=120$s, while the orange robot in free-space is unhindered.}
-%\vspace{-2em}
+\vspace{-3em}
+\caption{\label{fig:storyReal}{Positioning of two kilobots where they should reach their same color circle.
+The boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
+%\vspace{-1em}
 }
@@
 \subsection{Hardware Experiment: Control of Covariance}
-To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with a single light source, using friction with the boundary walls to vary the covariance from  -5000 to 10,000 pixel$^2$.
+To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 4000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
@@
 \begin{center}
-	\includegraphics[width=\columnwidth]{experiment.pdf}
+	\includegraphics[width=\columnwidth]{newExperiment.pdf}
 \end{center}
@@
 Hardware demonstration steering 100 kilobot robots to desired covariance.
-The goal covariance is shown in green, the actual covariance in red.
+The goal covariance is negative in first 100 seconds and is positive in the second 100 seconds.
+The actual covariance is shown in different trials.
 Frames above the plot show output from machine vision system and an overlaid covariance ellipse.

### 2016-09-01 RA-L/experiment.tex
moved files

@@
 \vspace{-3em}
-\caption{\label{fig:storyReal}{Positioning of two kilobots where they should reach their same color circle.
-The boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
+\caption{\label{fig:storyReal}{
+Position control of two kilobots (Alg.
+\ref{alg:XControl}) steered to corresponding colored circle.
+Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
 %\vspace{-1em}

### 2016-09-02 RA-L/appendix.tex
Added appendix information and references to bib

\documentclass[conference]{IEEEtran}
\usepackage{times}

% numbers option provides compact numerical references in the text. 
\usepackage[numbers]{natbib}
\usepackage{multicol}
\usepackage[bookmarks=true]{hyperref}

\usepackage{bbm}
\usepackage{calc}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{
  colorlinks =true,
  urlcolor = black,
  linkcolor = black
}
\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{rotating}

%\usepackage{xfrac}
\usepackage{nicefrac}
\usepackage{cite}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage[usenames, dvipsnames]{color}
\usepackage{colortbl}
%\usepackage{caption}

%\usepackage{wrapfig}
\usepackage{overpic}
%\usepackage{subfigure}
%\usepackage{textcomp}
\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{todonotes}

%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
% uncomment this to hide all red todos
%\renewcommand{\todo}{}

%% ABBREVIATIONS
\newcommand{\qstart}{q_{\text{start}}}
\newcommand{\qgoal}{q_{\text{goal}}}
\newcommand{\pstart}{p_{\text{start}}}
\newcommand{\pgoal}{p_{\text{goal}}}
\newcommand{\xstart}{x_{\text{start}}}
\newcommand{\xgoal}{x_{\text{goal}}}
\newcommand{\ystart}{y_{\text{start}}}
\newcommand{\ygoal}{y_{\text{goal}}}
\newcommand{\gammastart}{\gamma_{\text{start}}}
\newcommand{\gammagoal}{\gamma_{\text{goal}}}
\providecommand{\proc}[1]{\textsc{#1}}

\newcommand{\ARLfull}{Aero\-space Ro\-bot\-ics La\-bora\-tory }
\newcommand{\ARL}{\textsc{arl}}
\newcommand{\JPL}{\textsc{jpl}}
\newcommand{\PRM}{\textsc{prm}}

\newcommand{\CM}{\textsc{cm}}
\newcommand{\SVM}{\textsc{svm}}
\newcommand{\NN}{\textsc{nn}}
\newcommand{\prm}{\textsc{prm}}
\newcommand{\lemur}{\textsc{lemur}}
\newcommand{\Lemur}{\textsc{Lemur}}
\newcommand{\LP}{\textsc{lp}} 
\newcommand{\SOCP}{\textsc{socp}}
\newcommand{\SDP}{\textsc{sdp}}
\newcommand{\NP}{\textsc{np}}
\newcommand{\SAT}{\textsc{sat}}
\newcommand{\LMI}{\textsc{lmi}}
\newcommand{\hrp}{\textsc{hrp\nobreakdash-2}}
\newcommand{\DOF}{\textsc{dof}}
\newcommand{\UIUC}{\textsc{uiuc}}
%% MACROS

\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
\providecommand{\lscal}{<}
\providecommand{\gscal}{>}
\providecommand{\lvect}{\prec}
\providecommand{\gvect}{\succ}
\providecommand{\leqscal}{\leq}
\providecommand{\geqscal}{\geq}
\providecommand{\leqvect}{\preceq}
\providecommand{\geqvect}{\succeq}
\providecommand{\onevect}{\mathbf{1}}
\providecommand{\zerovect}{\mathbf{0}}
\providecommand{\field}[1]{\mathbb{#1}}
\providecommand{\C}{\field{C}}
\providecommand{\R}{\field{R}}
\newcommand{\Cspace}{\mathcal{Q}}
\newcommand{\Uspace}{\mathcal{U}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\Hcal}{$\mathcal{H}$}
\providecommand{\Vcal}{$\mathcal{V}$}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\homog}{homog}
\DeclareMathOperator{\domain}{dom}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\sign}{sgn}
\providecommand{\polar}{\triangle}
\providecommand{\ainner}{\underline{a}}
\providecommand{\aouter}{\overline{a}}
\providecommand{\binner}{\underline{b}}
\providecommand{\bouter}{\overline{b}}
\newcommand{\D}{\nobreakdash-\textsc{d}}
%\newcommand{\Fspace}{\mathcal{F}}
\providecommand{\Fspace}{\Cspace_\text{free}}
\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
\providecommand{\iff}{\Leftrightarrow}
\providecommand{\subinner}[1]{#1_{\text{inner}}}
\providecommand{\subouter}[1]{#1_{\text{outer}}}
\providecommand{\Ppoly}{\mathcal{X}}
\providecommand{\Pproj}{\mathcal{Y}}
\providecommand{\Pinner}{\subinner{\Pproj}}
\providecommand{\Pouter}{\subouter{\Pproj}}
\DeclareMathOperator{\argmax}{arg\,max}
\providecommand{\Aineq}{B}
\providecommand{\Aeq}{A}
\providecommand{\bineq}{u}
\providecommand{\beq}{t}
\DeclareMathOperator{\area}{area}
\newcommand{\contact}[1]{\Cspace_{#1}}
\newcommand{\feasible}[1]{\Fspace_{#1}}
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\figwid}{0.22\columnwidth}
\newcommand{\TRUE}{\textbf{true}}
\newcommand{\FALSE}{\textbf{false}}
\DeclareMathOperator{\atan2}{atan2}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}

\pdfinfo{
   /Author (Shiva Shahrokhi, Arun Mahadev, and Aaron T. Becker)
   /Title  (Supplement toAlgorithms For Shaping a Particle Swarm With a Shared Control Input Using Boundary Interaction)
   /CreationDate (D:20160129120000)
   /Subject (Simple Robots)
   /Keywords (Robots;Uniform Control Inputs)
}

\begin{document}

% paper title
\title{\huge{ \emph{Supplement to} 
Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}}

\author{Shiva Shahrokhi, Arun Mahadev, and Aaron T. Becker}

\maketitle

\begin{abstract}
%Also 
Includes algorithms and equations too lengthy for main paper, but potentially useful for the community.
Also links to videos and demonstration code for the algorithms.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
This supplement
\S \ref{sec:Videos}
\S \ref{sec:2robotWallFriction}
\S \ref{sec:fluidInPlanarRegion}

\section{Supplementary Videos}\label{sec:Videos}
Five videos animate the key algorithms in this paper.

\subsection{Robot Swarm in a Circle under Gravity}
The video \emph{Robot Swarm in a Circle under Gravity} shows the stable configuration of a swarm under a constant global input.  Animated plots show mean, variance, covariance, and correlation for a swarm in a circular workspace.
Full resolution video: \url{https://youtu.be/nPFAjVIOxYc}.
An online demonstration and source code of the algorithm are at \citet{Zhao2016mathematicaSquare}

\subsection{Distribution of Robot Swarm in Square under Gravity }
The video \emph{Distribution of Robot Swarm in Square under Gravity } shows the stable configuration of a swarm under a constant global input.  Animated plots show mean, variance, covariance, and correlation for a swarm in a square workspace.
Full resolution video: \url{https://youtu.be/ZEksDxLpAzg}.
An online demonstration and source code of the algorithm are at \citet{Zhao2016mathematica}.

\subsection{TwoRobotPosition.mp4}
Animates Algs. 1, 2, 3 in Mathematica to show how two robots can be arbitrarily positioned in a square workspace. In this video the desired initial and ending positions of the two robots are manipulated, and the path that the robots should follow is drawn. The video ends with an extreme case where the robots must exchange positions. 
Full resolution video: \url{https://youtu.be/5TWlw7vThsM}.
An online demonstration and source code of the algorithm are at \citet{Shahrokhi2015mathematicaParticle}.

\subsection{Arranging a robot swarm with global inputs and wall friction [discrete] }
An implementation of Alg. 4  in {\sc Matlab} that illustrates how the two robots positioning algorithm is extendable to $n$ robots. In this video all  robots gets the same input, but by exploiting wall friction each robot reaches its goal, the formation "UH".
Full resolution video: \url{https://youtu.be/uhpsAyPwKeI}.
Full code is available at \citet{Arun2015}.

\subsection{AutomaticCovControl.mp4}
A closed-loop controller that steers a swarm of particles to a desired covariance,  implemented with a box2D simulator. In this video the green ellipse is the desired covariance ellipse, the red ellipse is the current covariance ellipse of the swarm and the red dot is the mean position of the robots. Robots follow the algorithm to achieve the desired values for $\sigma_{goalxy}$, $\sigma_x^2$ and $\sigma_y^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ Algorithm for generating desired $y$ spacing between two robots using wall friction}\label{sec:2robotWallFriction}
\begin{algorithm}
\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. Current position of the robots are $(r_1,r_2)$.
\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
\State $ \Delta e_y \gets e_{1y} - e_{2y} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_y < 0 $ }
\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
\Else 
\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
\Else
\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_y = r_{1y} - r_{2y}$
\If {$\Delta r_y \equiv \Delta e_y$} 
\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Calculations for modeling swarm as fluid in a simple planar workspace}\label{sec:fluidInPlanarRegion}
Two workspaces are used, a square and a circular workspace.

\subsection{Square Workspace}
This section provides formulas for the mean, variance,  covariance and correlation of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together cover an area of constant volume $A$. Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape. The workspace is

The range of possible angles for the global input angle $\beta $ is [0,2$\pi $). In this range of angles, the swarm assumes eight different polygonal shapes. The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.

Two representative formulas are attached, the outline of the swarm shapes in \eqref{tab:SquareRobotRegions} and $\bar{x}(\beta,A)$ in \eqref{tab:SquareXMean}.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{SquarePlotPositions.pdf} 
\caption{A swarm in a square workspace under a constant global input assumes either a triangular or a trapezoidal shape if $A<1/2$.  If $A>1/2$ the swarm is either a squares with one corner removed or a trapezoidal  shape.}
\label{fig:friction}
\end{center}
\end{figure}

\begin{table*}
\begin{align}
\bar{x}(\beta,A) = A\leq \frac{1}{2}: &\begin{cases}
 -\frac{\tan ^2(\beta )}{24 A}-\frac{A}{2}+1 & 0\leq \beta \leq \tan ^{-1}(2 A)\lor 2 \pi -\tan ^{-1}(2 A)<\beta \leq 2 \pi  \\
 1-\frac{1}{3} \sqrt{2} \sqrt{A \tan (\beta )} & \tan ^{-1}(2 A)<\beta \leq \frac{\pi }{2}-\tan ^{-1}(2 A) \\
 \frac{\cot (\beta )}{12 A}+\frac{1}{2} & \frac{\pi }{2}-\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\frac{\pi }{2} \\
 \frac{1}{3} \sqrt{2} \sqrt{-A \tan (\beta )} & \tan ^{-1}(2 A)+\frac{\pi }{2}<\beta \leq \pi -\tan ^{-1}(2 A) \\
 \frac{\tan ^2(\beta )}{24 A}+\frac{A}{2} & \pi -\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\pi  \\
 \frac{1}{3} \sqrt{2} \sqrt{A \tan (\beta )} & \tan ^{-1}(2 A)+\pi <\beta \leq \frac{3 \pi }{2}-\tan ^{-1}(2 A) \\
 \frac{1}{2}-\frac{\cot (\beta )}{12 A} & \frac{3 \pi }{2}-\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\frac{3 \pi }{2} \\
 1-\frac{1}{3} \sqrt{2} \sqrt{-A \tan (\beta )} & \tan ^{-1}(2 A)+\frac{3 \pi }{2}<\beta \leq 2 \pi -\tan ^{-1}(2 A) \\
\end{cases} \nonumber\\
\frac{1}{2}<A<1:&\begin{cases}
 -\frac{\tan ^2(\beta )}{24 A}-\frac{A}{2}+1 & 0\leq \beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)\lor 2 \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq 2 \pi  \\
 \frac{2 \sqrt{2} \sqrt{(1-A) \tan (\beta )} (A-1)+3}{6 A} & \tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \frac{\pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \frac{6 A+\cot (\beta )}{12 A} & \frac{\pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{\pi }{2} \\
 \frac{-2 \sqrt{2} \sqrt{(A-1) \tan (\beta )} (A-1)+6 A-3}{6 A} & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{\pi }{2}<\beta \leq \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \frac{\tan ^2(\beta )}{24 A}+\frac{A}{2} & \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\pi  \\
 \frac{2 \sqrt{2} \sqrt{(1-A) \tan (\beta )} (1-A)+6 A-3}{6 A} & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\pi <\beta \leq \frac{3 \pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \frac{1}{2}-\frac{\cot (\beta )}{12 A} & \frac{3 \pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{3 \pi }{2} \\
 \frac{2 \sqrt{2} \sqrt{(A-1) \tan (\beta )} (A-1)+3}{6 A} & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{3 \pi }{2}<\beta \leq 2 \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
\end{cases}
 \nonumber \\
A=1: &\frac{1}{2}
\end{align}
\protect\caption{$\bar{x}$ in a unit-square workspace}
\label{tab:SquareXMean}
\end{table*}

\begin{table*}
\tiny
\begin{align}
\text{RobotRegion}(\beta,A)= \nonumber 
A\leq \frac{1}{2}:&
\begin{cases}
 \left(
\begin{array}{cc}
 1 & 0 \\
 1 & 1 \\
 -A-\frac{\tan (\beta )}{2}+1 & 1 \\
 -A+\frac{\tan (\beta )}{2}+1 & 0 \\
\end{array}
\right) & 0\leq \beta \leq \tan ^{-1}(2 A)\lor 2 \pi -\tan ^{-1}(2 A)<\beta \leq 2 \pi  \\
 \left(
\begin{array}{cc}
 1 & 1 \\
 1-\sqrt{2} \sqrt{A \tan (\beta )} & 1 \\
 1 & 1-\sqrt{2} \sqrt{A \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}(2 A)<\beta \leq \frac{\pi }{2}-\tan ^{-1}(2 A) \\
 \left(
\begin{array}{cc}
 1 & 1 \\
 0 & 1 \\
 0 & -A+\frac{\cot (\beta )}{2}+1 \\
 1 & -A-\frac{\cot (\beta )}{2}+1 \\
\end{array}
\right) & \frac{\pi }{2}-\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\frac{\pi }{2} \\
 \left(
\begin{array}{cc}
 0 & 1 \\
 \sqrt{2} \sqrt{-A \tan (\beta )} & 1 \\
 0 & 1-\sqrt{2} \sqrt{-A \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}(2 A)+\frac{\pi }{2}<\beta \leq \pi -\tan ^{-1}(2 A) \\
 \left(
\begin{array}{cc}
 0 & 0 \\
 0 & 1 \\
 A-\frac{\tan (\beta )}{2} & 1 \\
 A+\frac{\tan (\beta )}{2} & 0 \\
\end{array}
\right) & \pi -\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\pi  \\
 \left(
\begin{array}{cc}
 0 & 0 \\
 0 & \sqrt{2} \sqrt{A \cot (\beta )} \\
 \sqrt{2} \sqrt{A \tan (\beta )} & 0 \\
\end{array}
\right) & \tan ^{-1}(2 A)+\pi <\beta \leq \frac{3 \pi }{2}-\tan ^{-1}(2 A) \\
 \left(
\begin{array}{cc}
 0 & 0 \\
 1 & 0 \\
 1 & A-\frac{\cot (\beta )}{2} \\
 0 & A+\frac{\cot (\beta )}{2} \\
\end{array}
\right) & \frac{3 \pi }{2}-\tan ^{-1}(2 A)<\beta \leq \tan ^{-1}(2 A)+\frac{3 \pi }{2} \\
 \left(
\begin{array}{cc}
 1 & 0 \\
 1-\sqrt{2} \sqrt{-A \tan (\beta )} & 0 \\
 1 & \sqrt{2} \sqrt{-A \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}(2 A)+\frac{3 \pi }{2}<\beta \leq 2 \pi -\tan ^{-1}(2 A) \\
\end{cases}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nonumber \\
\frac{1}{2}<A<1:&
\begin{cases}
 \left(
\begin{array}{cc}
 1 & 0 \\
 1 & 1 \\
 (1-A)-\frac{\tan (\beta )}{2} & 1 \\
 (1-A)+\frac{\tan (\beta )}{2} & 0 \\
\end{array}
\right) & 0\leq \beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)\lor 2 \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq 2 \pi  \\
 \left(
\begin{array}{cc}
 1 & 0 \\
 1 & 1 \\
 0 & 1 \\
 0 & \sqrt{2} \sqrt{(1-A) \cot (\beta )} \\
 \sqrt{2} \sqrt{(1-A) \tan (\beta )} & 0 \\
\end{array}
\right) & \tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \frac{\pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \left(
\begin{array}{cc}
 0 & 1 \\
 1 & 1 \\
 1 & (1-A)-\frac{\cot (\beta )}{2} \\
 0 & (1-A)+\frac{\cot (\beta )}{2} \\
\end{array}
\right) & \frac{\pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{\pi }{2} \\
 \left(
\begin{array}{cc}
 1 & 1 \\
 0 & 1 \\
 0 & 0 \\
 1-\sqrt{2} \sqrt{-(1-A) \tan (\beta )} & 0 \\
 1 & \sqrt{2} \sqrt{-(1-A) \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{\pi }{2}<\beta \leq \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \left(
\begin{array}{cc}
 0 & 0 \\
 0 & 1 \\
 -(1-A)-\frac{\tan (\beta )}{2}+1 & 1 \\
 -(1-A)+\frac{\tan (\beta )}{2}+1 & 0 \\
\end{array}
\right) & \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\pi  \\
 \left(
\begin{array}{cc}
 1 & 0 \\
 0 & 0 \\
 0 & 1 \\
 1-\sqrt{2} \sqrt{(1-A) \tan (\beta )} & 1 \\
 1 & 1-\sqrt{2} \sqrt{(1-A) \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\pi <\beta \leq \frac{3 \pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
 \left(
\begin{array}{cc}
 1 & 0 \\
 0 & 0 \\
 0 & -(1-A)+\frac{\cot (\beta )}{2}+1 \\
 1 & -(1-A)-\frac{\cot (\beta )}{2}+1 \\
\end{array}
\right) & \frac{3 \pi }{2}-\tan ^{-1}\left(\frac{1}{2},1-A\right)<\beta \leq \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{3 \pi }{2} \\
 \left(
\begin{array}{cc}
 0 & 0 \\
 1 & 0 \\
 1 & 1 \\
 \sqrt{2} \sqrt{-(1-A) \tan (\beta )} & 1 \\
 0 & 1-\sqrt{2} \sqrt{-(1-A) \cot (\beta )} \\
\end{array}
\right) & \tan ^{-1}\left(\frac{1}{2},1-A\right)+\frac{3 \pi }{2}<\beta \leq 2 \pi -\tan ^{-1}\left(\frac{1}{2},1-A\right) \\
\end{cases},\nonumber\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A=1:&\left(
\begin{array}{cc}
 1 & 0 \\
 0 & 0 \\
 0 & 1 \\
 1 & 1 \\
\end{array}
\right)
\end{align}
\protect\caption{RobotRegions in a unit-square workspace}
\label{tab:SquareRobotRegions}
\end{table*}

\subsection{Circle Workspace}
The area under a chord of a circle is the area of a sector less the area of the triangle originating at the circle center: 
$A=S(sector)-S(triangle)=1/2 LR-1/2 C(1-h)$, thus
\begin{align}
A=(1/2)\left[LR-c(R-h)\right]
\end{align}
where $L$ is arc length, $c$ is chord length, $R$ is radius and $h$ is height. Solving for $L$ and $C$ gives
\begin{align}
L&=2 \cos ^{-1}(1-h)\\
C&=2\sqrt{h(2-h)}
\end{align}
Therefore the area under a chord is
\begin{align}
\cos ^{-1}(1-h)-(1-h) \sqrt{(2-h) h}
\end{align}

For a circular workspace, with $\beta = 0$, the variance of $x$ and $y$ are:
{\tiny
\begin{align}
&\sigma_x^2(h)=\frac{64 (h-2)^3 h^3}{144 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)^2} +\nonumber\\
&\frac{9 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right) \left(\sin \left(4 \arcsin(1-h)\right)+4 \arccos(1-h)\right)}{144 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)^2}
\end{align}}

{\tiny
\begin{align}
\sigma_y^2(h)=
\frac{12 \arccos(1-h)-8 \sin \left(2 \arccos(1-h)\right)+\sin \left(4 \arccos(1-h)\right)}{48 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
\end{align}}

For $\beta = 0$, $\sigma_{xy}=0$. These values can be rotated to calculate $\sigma_x^2(\beta,h),\sigma_y^2(\beta,h),$ and $\sigma_{xy}(\beta,h)$.

%%%%%%%%%%%%%%%
%% Use plainnat to work nicely with natbib. 
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,ShapingSwarmFrictionSharedInput}
\end{document}

### 2016-09-02 RA-L/simulation.tex
One small edit to section
added references to code

@@
 The first controls the position of two robots, the second controls the position of $n$ robots.
-All code is available online at [link withheld for review].
+All code is available online at \citet{Zhao2016mathematica,Zhao2016mathematicaSquare}.
 
@@
 Similarly,  two iterations %of Alg. \ref{alg:YControl} 
-are only required if $|\Delta e_y - \Delta s_y|>L$.
+are only required if $|\Delta e_y - \Delta s_y|>L$. An online interactive demonstration and source code of the algorithm are available at \citet{Shahrokhi2015mathematicaParticle}.
 
@@
 \subsection{Position Control of $n$ Robots}
-Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width.
+Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \citet{Arun2015}.
 Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
@@
 
- 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in Section \ref{subsec:ClosedLoopCovarianceControl}.
+ 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl}.
 Fig.

### 2016-09-02 RA-L/appendix.tex
Updated Appendix

@@
 Also links to videos and demonstration code for the algorithms.
+
+Consider a swarm of agents that are controlled by the same global inputs and have no autonomy. This paper presents algorithms for shaping such swarms in 2D.
+
+This model is common for current micro- and nano-robots, whose small size makes it difficult to perform onboard computation or contain a power and propulsion source.
+For this reason these robots are usually powered and controlled by global inputs, such as a uniform external electric or magnetic field, and every robot receives exactly the same control inputs.
+Due to their small size, large numbers of micro-robots are required to deliver sufficient payloads.
+ Nevertheless, these applications require precision control of the shape and position of the robot swarm. Precision control requires breaking the symmetry caused by the global input.
+
+A promising technique uses collisions with boundary walls to shape the swarm, however, the range of configurations created by conforming a swarm to a boundary wall is limited.
+This paper describes the set of stable configurations of a swarm in two canonical workspaces, a circle and a square.
+
+To increase the diversity of configurations, we add boundary interaction to our model.
+We provide algorithms using friction with walls to place two robots at arbitrary locations in a rectangular workspace.
+Next, we extend this algorithm to place $n$ robots at desired locations.
+We conclude with efficient techniques to control the covariance of a swarm not possible without wall-friction.
+Simulations and hardware implementations with 100 robots validate these results.
+
 \end{abstract}
@@
 \section{Introduction}
-This supplement
-\S \ref{sec:Videos}
-\S \ref{sec:2robotWallFriction}
-\S \ref{sec:fluidInPlanarRegion}
+This supplement gives overviews of the videos and code in 
+\S \ref{sec:Videos}, 
+provides the algorithm for $y$ position control of two robots in
+\S \ref{sec:2robotWallFriction},
+and gives gull analytical models for fluid settling in square-shaped tanks in
+\S \ref{sec:fluidInPlanarRegion}.
 
@@
 Full resolution video: \url{https://youtu.be/nPFAjVIOxYc}.
-An online demonstration and source code of the algorithm are at \citet{Zhao2016mathematicaSquare}
+An online demonstration and source code of the algorithm are at \citet{Zhao2016mathematicaSquare}.
 
@@
 
-\subsection{TwoRobotPosition.mp4}
+\subsection{Steering 2 Particles with Shared Controls Using Wall Friction}
 Animates Algs.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgments}
+This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
+
 %%%%%%%%%%%%%%%
@@
 \bibliographystyle{plainnat}
+\footnotesize
 \bibliography{IEEEabrv,ShapingSwarmFrictionSharedInput}

### 2016-09-02 RA-L/theory.tex
Replaced Shiva's files and changes images for pink borders

@@
 \vspace{-1em}
-\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot.}
+\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot.
+At left there is a fluid and inside it the velocity of the robot is almost zero at the surface and is more and more until it is the same as the free space on the threshold.}
 \label{fig:friction}
@@
 \renewcommand{\figwid}{0.4\columnwidth}
-{\begin{overpic}[width =\figwid]{one_1.jpg}\put(45,75){$t$ = 0 s}
+{\begin{overpic}[width =\figwid]{S1_1.png}\put(55,15){$t$ = 0 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{one_2.jpg}\put(45,75){$t$ = 0.25 s}
+\begin{overpic}[width =\figwid]{S1_2.png}\put(55,15){$t$ = 0.14 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{one_3.jpg}\put(45,75){$t$  = 0.5 s}
+\begin{overpic}[width =\figwid]{S1_3.png}\put(55,15){$t$  = 0.29 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{one_4.jpg}\put(45,75){$t$  = 0.75 s}
+\begin{overpic}[width =\figwid]{S1_4.png}\put(55,15){$t$  = 0.45 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{one_5.jpg}\put(45,75){$t$  = 1 s}
+\begin{overpic}[width =\figwid]{S1_5.png}\put(55,15){$t$  = 0.7 s}
 \end{overpic}}\\
@@
 %\end{overpic}}\\
-\vspace{.5em}
-{\begin{overpic}[width =\figwid]{S3_1.jpg}\put(45,75){$t$ = 0 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S3_2.jpg}\put(45,75){$t$ = 0.25 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S3_3.jpg}\put(45,75){$t$  = 0.5 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S3_4.jpg}\put(45,75){$t$  = 0.75 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S3_5.jpg}\put(45,75){$t$  = 1 s}
-\end{overpic}}
-\vspace{-1em}
+%\vspace{.5em}
+%{\begin{overpic}[width =\figwid]{S3_1.jpg}\put(45,75){$t$ = 0 s}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S3_2.jpg}\put(45,75){$t$ = 0.25 s}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S3_3.jpg}\put(45,75){$t$  = 0.5 s}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S3_4.jpg}\put(45,75){$t$  = 0.75 s}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S3_5.jpg}\put(45,75){$t$  = 1 s}
+%\end{overpic}}
+%\vspace{-1em}
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
@@
 The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
-The bottom row shows an extreme case where the robots must switch position.
+%The bottom row shows an extreme case where the robots must switch position.
 }

### 2016-09-02 RA-L/intro.tex
Rewrote the abstract and introduction
Is it better?

@@
 \section{Introduction}\label{sec:Intro}
-Micro- and nano-robots can be manufactured in large numbers, see recent examples in~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, \citet{Ou2013} or \citet{qiu2015magnetic}.
+Particle swarms steered by a global force is common in applied mathematics, biology, and computer graphics. 
+\begin{align}
+[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
+\end{align}
+The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
+As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, \citet{Ou2013} or \citet{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
@@
 Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze.
-Even without obstacles or boundaries, the mean position of a swarm is controllable.
+Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
 By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable.

### 2016-09-02 RA-L/simulation.tex
Rewrote the abstract and introduction
Is it better?

@@
 Similarly,  two iterations %of Alg. \ref{alg:YControl} 
-are only required if $|\Delta e_y - \Delta s_y|>L$. An online interactive demonstration and source code of the algorithm are available at \citet{Shahrokhi2015mathematicaParticle}.
+are only required if $|\Delta e_y - \Delta s_y|>L$. An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
 
@@
 \subsection{Position Control of $n$ Robots}
-Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \citet{Arun2015}.
+Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \citep{Arun2015}.
 Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
@@
 %The plot compares  \emph{Total distance}---the sum of the moves made by every robot, with \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.
-See hardware implementation and simulation at [link withheld for review].
+See hardware implementation and simulation at \citep{Arun2015}.
 }

### 2016-09-02 RA-L/theory.tex
Rewrote the abstract and introduction
Is it better?

@@
 At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
-See online interactive demonstration at [withheld for double-blind review].}
+See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
 \label{fig:SquareFill}
@@
 The locus of possible mean positions are concentric circles.
-See online interactive demonstration at [withheld for double-blind review].}
+See online interactive demonstration at \citep{Shahrokhi2015mathematicaParticle}.}
 \label{fig:CircleFill}
@@
 
-Computing the means, variances, covariance, and correlation requires integrating over the area containing the swarm.
-This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
-The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows, integrated over the unit square with $x$ and $y$ from 0 to 1:
+Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm.
+%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
 
-\begin{equation}\label{eq:meanInSquareWorkspace}
-\bar{x} =\frac{\iint x \bm{1}_A(x,y)dxdy}{A}
-\text{, }
-\bar{y}=\frac{\iint y \bm{1}_A(x,y)dxdy}{A}
-\end{equation}
-
-\begin{align}\label{eq:varInSquareWorkspace}
-\sigma^2_x &=\frac{\iint \left(x-\bar{x}\right)^2 \bm{1}_A(x,y) dx dy}{A}, \\
-\sigma^2_y &=\frac{\iint  \left(y-\bar{y}\right)^2 \bm{1}_A(x,y) dx dy}{A}
+\begin{align}
+\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
+\text{, } \bar{y}=\frac{\iint_R y \,dx\,dy}{A}&\\
+%\end{equation}
+%\begin{align}
+\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}&\\
+%\end{align}
+%\begin{align}
+\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
+\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}&
 \end{align}
 
-\begin{align}\label{eq:covAndcorrInSquareWorkspace}
-\sigma_{xy} &= \frac{\iint  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \bm{1}_A(x,y) dx dy}{A},\\
-\rho_{xy} &= \frac{\sigma^2_x}{\sigma_x\sigma_y}
-\end{align}
-
-Rather than using an indicator function, the region of integration can be changed to only include the polygon containing the swarm.
-If the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
+The region of integration $R$ is the polygon containing the swarm. If the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
 \begin{align}\label{eq:meanInSquareWorkspaceLL}

### 2016-09-02 RA-L/intro.tex
adding missing refs

@@
 To extend the range of possible shapes, we introduce wall friction to the system model \S \ref{subsec:WallFriction}.
-We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this proof to prove that a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
+We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this to prove a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
 \S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.

### 2016-09-02 RA-L/theory.tex
adding missing refs

@@
 \paragraph{Square workplace}
-This section examines the mean, variance,  covariance and correlation of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
+%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
+This section examines the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$.
@@
 
-The range of possible angles for the global input angle $\beta $ is [0,2$\pi $).
-In this range of angles, the swarm assumes eight different polygonal shapes.
+The range for the global input angle $\beta $ is [0,2$\pi $).
+In this range, the swarm assumes eight different polygonal shapes.
 The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.
 
-Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm.
-%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
-The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
+Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
 
@@
 \bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, } \bar{y}=\frac{\iint_R y \,dx\,dy}{A}&\\
-%\end{equation}
-%\begin{align}
+\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A}
+\end{align}
+\begin{align}
 \sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}&\\
-%\end{align}
-%\begin{align}
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}
+\end{align}
+\begin{align}
 \sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}&
+\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
 \end{align}

### 2016-09-05 RA-L/experiment.tex
Fixed a conflict.
fixed another conflict,,,
Worked on intro

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiment}\label{sec:expResults}
+\section{experiment}\label{sec:expResults}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-\renewcommand{\figwid}{0.4\columnwidth}
+\renewcommand{\figwid}{0.38\columnwidth}
 {
-\begin{overpic}[width =\figwid]{twoR1.pdf}\put(12,70){$t$  = 0 s}\end{overpic}
+\begin{overpic}[width =\figwid]{twoR1.pdf}%\put(12,70){$t$  = 0 s}
+\end{overpic}
 \begin{overpic}[width =\figwid]{twoR2.pdf}\put(12,70){$t$  = 60 s}
@@
 \end{overpic}}
-\vspace{-3em}
-\caption{\label{fig:storyReal}{
-Position control of two kilobots (Alg.
+%\vspace{-3em}
+\caption{\label{fig:storyReal}Position control of two kilobots (Alg.
 \ref{alg:XControl}) steered to corresponding colored circle.
 Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
-%\vspace{-1em}
-}
 \end{figure*}
@@
 \caption{\label{fig:setup}
-Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight remotely triggered LED floodlights, with an overhead machine vision system.
+Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight  triggered LED floodlights, with an overhead machine vision system.
 }

### 2016-09-05 RA-L/intro.tex
Worked on intro

@@
 Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
-By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable.
+By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}.
 %The usefulness of these techniques was recently demonstrated by several automatic controllers.
@@
 
-This paper provides analytical position control results in two canonical workspaces with frictionless walls \S\ref{subsec:FluidInTank}.
+\S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes, we introduce wall friction to the system model \S \ref{subsec:WallFriction}.
-We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots \S \ref{sec:PostionControl2Robots}, and extend this to prove a rectangular workspace with high friction boundaries is sufficient to position a swarm of $n$ robots arbitrarily within a subset of the workspace \S \ref{sec:PostionControlnRobots}.
+To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces wall friction to the system model.
+We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in \S \ref{sec:PostionControl2Robots}, and \S \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
 \S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.

### 2016-09-05 RA-L/theory.tex
Worked on intro

@@
 
-Computing the means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+Computing means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
@@
 \bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A}
-\end{align}
-\begin{align}
+\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \\
+%\end{align}
+%\begin{align}
 \sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}
-\end{align}
-\begin{align}
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\\
+%\end{align}
+%\begin{align}
 \sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
@@
 \vspace{-1em}
-\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot.
-At left there is a fluid and inside it the velocity of the robot is almost zero at the surface and is more and more until it is the same as the free space on the threshold.}
+\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (Right) velocity of a fluid reduces to zero at the boundary.}
 \label{fig:friction}
@@
 Dashed lines show the shortest route if robots could be controlled independently.
-The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
+Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
+Online demonstration and source code at \citet{Shahrokhi2015mathematicaParticle}.
 %The bottom row shows an extreme case where the robots must switch position.
@@
 \caption{\label{fig:driftmove}
-A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-The robot $A$ touching a top wall will move right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
+A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
+Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
 }
@@
 \begin{center}
-	\includegraphics[width=1.0\columnwidth]{PositionNrobots.pdf}
+	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
 \end{center}

### 2016-09-06 RA-L/experiment.tex

@@
 \subsection{Hardware Experiment: Position Control of n Robots}
+The hardware setup has a bounded platform, magnetic sliders, and a magnetic guide board.
+Designs for each are available at \citep{arunhardware}.
+The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
+Only discrete, 1 cm moves in the $x$ and $y$ directions are used.
+The goal configuration highlighted in the top right corner represents a `U' made of seven sliders.
+The dark red configuration is the current position of the sliders.
+Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.
+Free robots return to their start positions but robots on the boundary to move laterally, generating a net sliding motion.
+
+Figure \ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2 and 7.
+All robots receive the same control inputs, but boundary interactions breaks the control symmetry.
+Robots reach their respective goal position in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.
 

### 2016-09-06 RA-L/experiment.tex
edits to the experiment

@@
 \subsection{Hardware Experiment: Position Control of Two Robots}
-The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border and the three-legged design of the kilobots.
+The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border.
 When a kilobot is steered into the zigzag border, they pin themselves to the wall unless the global input directs them away from the wall.
@@
 
-Figure \ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2 and 7.
+Fig.
+\ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2 and 7.
 All robots receive the same control inputs, but boundary interactions breaks the control symmetry.
@@
 \subsection{Hardware Experiment: Control of Covariance}
-To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 4000 cm$^2$.
+To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
@@
 \caption{\label{fig:covExperiment}
-Hardware demonstration steering 100 kilobot robots to desired covariance.
-The goal covariance is negative in first 100 seconds and is positive in the second 100 seconds.
+Hardware demonstration steering up to 100 kilobot robots to desired covariance.
+The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.
 The actual covariance is shown in different trials.

### 2016-09-06 RA-L/theory.tex
Changed coloring
Fixed reference, added spot for equation

@@
 The locus of possible mean positions are concentric circles.
-See online interactive demonstration at \citep{Shahrokhi2015mathematicaParticle}.}
+See online interactive demonstration at \citep{Zhao2016mathematica}.}
 \label{fig:CircleFill}
@@
 This means that if one robot is touching the wall and another robot is free, if the control input is parallel or into the wall, the touching robot will not move. 
-The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls are sufficient to arbitrarily position two robots. 
+There are many alternate models of friction that also break control symmetry.
+Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
+Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer.
+\begin{align}
+INSERT FORMULA \label{eq:boundarylayerflow}
+\end{align}
+
+The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls can arbitrarily position two robots. 
 \begin{figure}[h]
@@
 \renewcommand{\figwid}{0.4\columnwidth}
-{\begin{overpic}[width =\figwid]{S1_1.png}\put(55,15){$t$ = 0 s}
+{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_2.png}\put(55,15){$t$ = 0.14 s}
+\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_3.png}\put(55,15){$t$  = 0.29 s}
+\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_4.png}\put(55,15){$t$  = 0.45 s}
+\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_5.png}\put(55,15){$t$  = 0.7 s}
+\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
 \end{overpic}}\\
@@
 Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
-Online demonstration and source code at \citet{Shahrokhi2015mathematicaParticle}.
+Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
 %The bottom row shows an extreme case where the robots must switch position.
@@
 
-This section describes Alg.~\ref{alg:PosControl2Robots}, an algorithm that uses wall-friction to arbitrarily position two robots in a rectangular workspace.
+This section describes Alg.~\ref{alg:PosControl2Robots}, which uses wall-friction to arbitrarily position two robots in a rectangular workspace.
 This algorithm  introduces concepts that will be used for multi-robot positioning.
@@
 Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
-
 Finally, all robots are moved right until robot $k$ is in the desired target position.

### 2016-09-06 RA-L/intro.tex
Updated he intro
changed abstract

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms steered by a global force is common in applied mathematics, biology, and computer graphics. 
+Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
 \begin{align}
@@
 Someday large swarms of robots will be remotely guided
- 1)  ex vivo to assemble structures in parallel and 
-2) through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
+  ex vivo to assemble structures in parallel and 
+ through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
@@
 By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}.
-%The usefulness of these techniques was recently demonstrated by several automatic controllers.
-One controller steered a swarm of robots to push a larger block through a 2D maze~\citet{ShahrokhiIROS2015}.
 A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.
-%This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
-%Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$.
 
+The paper is arranged as follows.
 \S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.

### 2016-09-06 RA-L/theory.tex
changed abstract

@@
 \vspace{-1em}
-\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (Right) velocity of a fluid reduces to zero at the boundary.}
+\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.}
 \label{fig:friction}

### 2016-09-07 RA-L/appendix.tex
Worked on intro
compressed some images, made appendix able to be added easily.

--- 2016-09-02 RA-L/appendix.tex
@@
-\documentclass[conference]{IEEEtran}
-\usepackage{times}
-
-% numbers option provides compact numerical references in the text. 
-\usepackage[numbers]{natbib}
-\usepackage{multicol}
-\usepackage[bookmarks=true]{hyperref}
-
-\usepackage{bbm}
-\usepackage{calc}
-\usepackage{url}
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks =true,
-  urlcolor = black,
-  linkcolor = black
-}
-\usepackage{graphicx}
-\usepackage[cmex10]{amsmath}
-\usepackage{bm}
-\usepackage{amssymb}
-\usepackage{rotating}
-
-%\usepackage{xfrac}
-\usepackage{nicefrac}
-\usepackage{cite}
-\usepackage[caption=false,font=footnotesize]{subfig}
-\usepackage[usenames, dvipsnames]{color}
-\usepackage{colortbl}
-%\usepackage{caption}
-
-%\usepackage{wrapfig}
-\usepackage{overpic}
-%\usepackage{subfigure}
-%\usepackage{textcomp}
-\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
-\usepackage{breqn} %for breaking equations automatically
-\usepackage[ruled]{algorithm}
-\usepackage{algpseudocode}
-%\usepackage{algorithmic}
-\usepackage{multirow}
-\usepackage{todonotes}
-
-%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
-% uncomment this to hide all red todos
-%\renewcommand{\todo}{}
-
-%% ABBREVIATIONS
-\newcommand{\qstart}{q_{\text{start}}}
-\newcommand{\qgoal}{q_{\text{goal}}}
-\newcommand{\pstart}{p_{\text{start}}}
-\newcommand{\pgoal}{p_{\text{goal}}}
-\newcommand{\xstart}{x_{\text{start}}}
-\newcommand{\xgoal}{x_{\text{goal}}}
-\newcommand{\ystart}{y_{\text{start}}}
-\newcommand{\ygoal}{y_{\text{goal}}}
-\newcommand{\gammastart}{\gamma_{\text{start}}}
-\newcommand{\gammagoal}{\gamma_{\text{goal}}}
-\providecommand{\proc}[1]{\textsc{#1}}
-
-\newcommand{\ARLfull}{Aero\-space Ro\-bot\-ics La\-bora\-tory }
-\newcommand{\ARL}{\textsc{arl}}
-\newcommand{\JPL}{\textsc{jpl}}
-\newcommand{\PRM}{\textsc{prm}}
-
-\newcommand{\CM}{\textsc{cm}}
-\newcommand{\SVM}{\textsc{svm}}
-\newcommand{\NN}{\textsc{nn}}
-\newcommand{\prm}{\textsc{prm}}
-\newcommand{\lemur}{\textsc{lemur}}
-\newcommand{\Lemur}{\textsc{Lemur}}
-\newcommand{\LP}{\textsc{lp}} 
-\newcommand{\SOCP}{\textsc{socp}}
-\newcommand{\SDP}{\textsc{sdp}}
-\newcommand{\NP}{\textsc{np}}
-\newcommand{\SAT}{\textsc{sat}}
-\newcommand{\LMI}{\textsc{lmi}}
-\newcommand{\hrp}{\textsc{hrp\nobreakdash-2}}
-\newcommand{\DOF}{\textsc{dof}}
-\newcommand{\UIUC}{\textsc{uiuc}}
-%% MACROS
-
-\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
-\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
-\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
-\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
-\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
-\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
-\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
-\providecommand{\lscal}{<}
-\providecommand{\gscal}{>}
-\providecommand{\lvect}{\prec}
-\providecommand{\gvect}{\succ}
-\providecommand{\leqscal}{\leq}
-\providecommand{\geqscal}{\geq}
-\providecommand{\leqvect}{\preceq}
-\providecommand{\geqvect}{\succeq}
-\providecommand{\onevect}{\mathbf{1}}
-\providecommand{\zerovect}{\mathbf{0}}
-\providecommand{\field}[1]{\mathbb{#1}}
-\providecommand{\C}{\field{C}}
-\providecommand{\R}{\field{R}}
-\newcommand{\Cspace}{\mathcal{Q}}
-\newcommand{\Uspace}{\mathcal{U}}
-\providecommand{\Fspace}{\Cspace_\text{free}}
-\providecommand{\Hcal}{$\mathcal{H}$}
-\providecommand{\Vcal}{$\mathcal{V}$}
-\DeclareMathOperator{\conv}{conv}
-\DeclareMathOperator{\cone}{cone}
-\DeclareMathOperator{\homog}{homog}
-\DeclareMathOperator{\domain}{dom}
-\DeclareMathOperator{\range}{range}
-\DeclareMathOperator{\sign}{sgn}
-\providecommand{\polar}{\triangle}
-\providecommand{\ainner}{\underline{a}}
-\providecommand{\aouter}{\overline{a}}
-\providecommand{\binner}{\underline{b}}
-\providecommand{\bouter}{\overline{b}}
-\newcommand{\D}{\nobreakdash-\textsc{d}}
-%\newcommand{\Fspace}{\mathcal{F}}
-\providecommand{\Fspace}{\Cspace_\text{free}}
-\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
-\providecommand{\iff}{\Leftrightarrow}
-\providecommand{\subinner}[1]{#1_{\text{inner}}}
-\providecommand{\subouter}[1]{#1_{\text{outer}}}
-\providecommand{\Ppoly}{\mathcal{X}}
-\providecommand{\Pproj}{\mathcal{Y}}
-\providecommand{\Pinner}{\subinner{\Pproj}}
-\providecommand{\Pouter}{\subouter{\Pproj}}
-\DeclareMathOperator{\argmax}{arg\,max}
-\providecommand{\Aineq}{B}
-\providecommand{\Aeq}{A}
-\providecommand{\bineq}{u}
-\providecommand{\beq}{t}
-\DeclareMathOperator{\area}{area}
-\newcommand{\contact}[1]{\Cspace_{#1}}
-\newcommand{\feasible}[1]{\Fspace_{#1}}
-\newcommand{\dd}{\; \mathrm{d}}
-\newcommand{\figwid}{0.22\columnwidth}
-\newcommand{\TRUE}{\textbf{true}}
-\newcommand{\FALSE}{\textbf{false}}
-\DeclareMathOperator{\atan2}{atan2}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{lemma}[theorem]{Lemma}
-
-\pdfinfo{
-   /Author (Shiva Shahrokhi, Arun Mahadev, and Aaron T. Becker)
-   /Title  (Supplement toAlgorithms For Shaping a Particle Swarm With a Shared Control Input Using Boundary Interaction)
-   /CreationDate (D:20160129120000)
-   /Subject (Simple Robots)
-   /Keywords (Robots;Uniform Control Inputs)
-}
-
-\begin{document}
+%
+%\documentclass[conference]{IEEEtran}
+%\usepackage{times}
+%
+%% numbers option provides compact numerical references in the text. 
+%\usepackage[numbers]{natbib}
+%\usepackage{multicol}
+%\usepackage[bookmarks=true]{hyperref}
+%
+%\usepackage{bbm}
+%\usepackage{calc}
+%\usepackage{url}
+%\usepackage{hyperref}
+%\hypersetup{
+%  colorlinks =true,
+%  urlcolor = black,
+%  linkcolor = black
+%}
+%\usepackage{graphicx}
+%\usepackage[cmex10]{amsmath}
+%\usepackage{bm}
+%\usepackage{amssymb}
+%\usepackage{rotating}
+%
+%
+%%\usepackage{xfrac}
+%\usepackage{nicefrac}
+%\usepackage{cite}
+%\usepackage[caption=false,font=footnotesize]{subfig}
+%\usepackage[usenames, dvipsnames]{color}
+%\usepackage{colortbl}
+%%\usepackage{caption}
+%
+%%\usepackage{wrapfig}
+%\usepackage{overpic}
+%%\usepackage{subfigure}
+%%\usepackage{textcomp}
+%\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
+%\usepackage{breqn} %for breaking equations automatically
+%\usepackage[ruled]{algorithm}
+%\usepackage{algpseudocode}
+%%\usepackage{algorithmic}
+%\usepackage{multirow}
+%\usepackage{todonotes}
+%
+%%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
+%% uncomment this to hide all red todos
+%%\renewcommand{\todo}{}
+%
+%%% ABBREVIATIONS
+%\newcommand{\qstart}{q_{\text{start}}}
+%\newcommand{\qgoal}{q_{\text{goal}}}
+%\newcommand{\pstart}{p_{\text{start}}}
+%\newcommand{\pgoal}{p_{\text{goal}}}
+%\newcommand{\xstart}{x_{\text{start}}}
+%\newcommand{\xgoal}{x_{\text{goal}}}
+%\newcommand{\ystart}{y_{\text{start}}}
+%\newcommand{\ygoal}{y_{\text{goal}}}
+%\newcommand{\gammastart}{\gamma_{\text{start}}}
+%\newcommand{\gammagoal}{\gamma_{\text{goal}}}
+%\providecommand{\proc}[1]{\textsc{#1}}
+%
+%
+%\newcommand{\ARLfull}{Aero\-space Ro\-bot\-ics La\-bora\-tory }
+%\newcommand{\ARL}{\textsc{arl}}
+%\newcommand{\JPL}{\textsc{jpl}}
+%\newcommand{\PRM}{\textsc{prm}}
+%
+%\newcommand{\CM}{\textsc{cm}}
+%\newcommand{\SVM}{\textsc{svm}}
+%\newcommand{\NN}{\textsc{nn}}
+%\newcommand{\prm}{\textsc{prm}}
+%\newcommand{\lemur}{\textsc{lemur}}
+%\newcommand{\Lemur}{\textsc{Lemur}}
+%\newcommand{\LP}{\textsc{lp}} 
+%\newcommand{\SOCP}{\textsc{socp}}
+%\newcommand{\SDP}{\textsc{sdp}}
+%\newcommand{\NP}{\textsc{np}}
+%\newcommand{\SAT}{\textsc{sat}}
+%\newcommand{\LMI}{\textsc{lmi}}
+%\newcommand{\hrp}{\textsc{hrp\nobreakdash-2}}
+%\newcommand{\DOF}{\textsc{dof}}
+%\newcommand{\UIUC}{\textsc{uiuc}}
+%%% MACROS
+%
+%
+%\providecommand{\abs}[1]{\left\lvert#1\right\rvert}
+%\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
+%\providecommand{\normn}[2]{\left\lVert#1\right\rVert_#2}
+%\providecommand{\dualnorm}[1]{\norm{#1}_\ast}
+%\providecommand{\dualnormn}[2]{\norm{#1}_{#2\ast}}
+%\providecommand{\set}[1]{\lbrace\,#1\,\rbrace}
+%\providecommand{\cset}[2]{\lbrace\,{#1}\nobreak\mid\nobreak{#2}\,\rbrace}
+%\providecommand{\lscal}{<}
+%\providecommand{\gscal}{>}
+%\providecommand{\lvect}{\prec}
+%\providecommand{\gvect}{\succ}
+%\providecommand{\leqscal}{\leq}
+%\providecommand{\geqscal}{\geq}
+%\providecommand{\leqvect}{\preceq}
+%\providecommand{\geqvect}{\succeq}
+%\providecommand{\onevect}{\mathbf{1}}
+%\providecommand{\zerovect}{\mathbf{0}}
+%\providecommand{\field}[1]{\mathbb{#1}}
+%\providecommand{\C}{\field{C}}
+%\providecommand{\R}{\field{R}}
+%\newcommand{\Cspace}{\mathcal{Q}}
+%\newcommand{\Uspace}{\mathcal{U}}
+%\providecommand{\Fspace}{\Cspace_\text{free}}
+%\providecommand{\Hcal}{$\mathcal{H}$}
+%\providecommand{\Vcal}{$\mathcal{V}$}
+%\DeclareMathOperator{\conv}{conv}
+%\DeclareMathOperator{\cone}{cone}
+%\DeclareMathOperator{\homog}{homog}
+%\DeclareMathOperator{\domain}{dom}
+%\DeclareMathOperator{\range}{range}
+%\DeclareMathOperator{\sign}{sgn}
+%\providecommand{\polar}{\triangle}
+%\providecommand{\ainner}{\underline{a}}
+%\providecommand{\aouter}{\overline{a}}
+%\providecommand{\binner}{\underline{b}}
+%\providecommand{\bouter}{\overline{b}}
+%\newcommand{\D}{\nobreakdash-\textsc{d}}
+%%\newcommand{\Fspace}{\mathcal{F}}
+%\providecommand{\Fspace}{\Cspace_\text{free}}
+%\providecommand{\free}{\text{\{}\mathsf{free}\text{\}}}
+%\providecommand{\iff}{\Leftrightarrow}
+%\providecommand{\subinner}[1]{#1_{\text{inner}}}
+%\providecommand{\subouter}[1]{#1_{\text{outer}}}
+%\providecommand{\Ppoly}{\mathcal{X}}
+%\providecommand{\Pproj}{\mathcal{Y}}
+%\providecommand{\Pinner}{\subinner{\Pproj}}
+%\providecommand{\Pouter}{\subouter{\Pproj}}
+%\DeclareMathOperator{\argmax}{arg\,max}
+%\providecommand{\Aineq}{B}
+%\providecommand{\Aeq}{A}
+%\providecommand{\bineq}{u}
+%\providecommand{\beq}{t}
+%\DeclareMathOperator{\area}{area}
+%\newcommand{\contact}[1]{\Cspace_{#1}}
+%\newcommand{\feasible}[1]{\Fspace_{#1}}
+%\newcommand{\dd}{\; \mathrm{d}}
+%\newcommand{\figwid}{0.22\columnwidth}
+%\newcommand{\TRUE}{\textbf{true}}
+%\newcommand{\FALSE}{\textbf{false}}
+%\DeclareMathOperator{\atan2}{atan2}
+%
+%
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{definition}[theorem]{Definition}
+%\newtheorem{lemma}[theorem]{Lemma}
+%
+%
+%\pdfinfo{
+%   /Author (Shiva Shahrokhi, Arun Mahadev, and Aaron T. Becker)
+%   /Title  (Supplement toAlgorithms For Shaping a Particle Swarm With a Shared Control Input Using Boundary Interaction)
+%   /CreationDate (D:20160129120000)
+%   /Subject (Simple Robots)
+%   /Keywords (Robots;Uniform Control Inputs)
+%}
+%
+%\begin{document}
 
@@
 
+\newpage
 \maketitle
+
+{\huge{ \emph{Supplement to} 
+Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}}\\
 
@@
 Also links to videos and demonstration code for the algorithms.
+
 \end{abstract}
@@
 Full code is available at \citet{Arun2015}.
+Note that this code uses discretized version of Algorithm 3.  The continuous-movement version is illustrated in Fig.\ref{fig:PositionNrobotsCont}.
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{PositionNrobotsComp.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:PositionNrobotsCont}
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
+}
+\end{figure}
 
@@
 \bibliographystyle{plainnat}
+\footnotesize
 \bibliography{IEEEabrv,ShapingSwarmFrictionSharedInput}
-\end{document}
+%\end{document}

### 2016-09-07 RA-L/experiment.tex
New Reference style

@@
 These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot, from \citet{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available as an open source platform or commercially from~\citet{K-Team2015}.
+The kilobot, from \citep{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots.
+It is available as an open source platform or commercially from~\citep{K-Team2015}.
 Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
@@
 \caption{\label{fig:covExperiment}
-Hardware demonstration steering up to 100 kilobot robots to desired covariance.
+Hardware demonstration steering 100 kilobot robots to desired covariance.
 The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.

### 2016-09-07 RA-L/intro.tex
Trying to make introduction flow
New Reference style

@@
 The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
-As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citet{Chowdhury2015}, \citet{martel2014computer}, \citet{kim2015imparting}, \citet{Donald2013}, \citet{Ghosh2009}, \citet{Ou2013} or \citet{qiu2015magnetic}.
+As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
   ex vivo to assemble structures in parallel and 
- through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citet{munoz2014review} \citet{sitti2015biomedical} . 
+ through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
 Instead, these robots are often controlled by a global, broadcast signal.
-These applications require control techniques that can reliably exploit large populations despite high under-actuation.
+These applications require techniques to shape the swarm that can reliably exploit large populations despite high under-actuation.
 
@@
 
-%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\cite{Becker2013b}.
+%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}.
 Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze.

### 2016-09-07 RA-L/simulation.tex
moved unused pics
New Reference style

@@
 The first controls the position of two robots, the second controls the position of $n$ robots.
-All code is available online at \citet{Zhao2016mathematica,Zhao2016mathematicaSquare}.
+All code is available online at \citep{Zhao2016mathematica,Zhao2016mathematicaSquare}.
 
@@
 This section demonstrates simulations of controlling covariance of the swarm.
-These simulations use  the 2D physics engine Box2D, by \citet{catto2010box2d}.
+These simulations use  the 2D physics engine Box2D, by \citep{catto2010box2d}.
  144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig.
@@
 \begin{center}
-	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.eps}
+	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.pdf}
 \end{center}

### 2016-09-08 RA-L/experiment.tex
Edits made to the paper

@@
 \subsection{Hardware Experiment: Position Control of Two Robots}
-The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border.
-When a kilobot is steered into the zigzag border, they pin themselves to the wall unless the global input directs them away from the wall.
-This wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.
+The walls of the hardware platform have almost infinite friction, due to a laser-cut, slotted border.
+When a kilobot is steered into the slotted border, they pin themselves to the wall unless the global input directs them away from the wall.
+Even though the kilobots are stochastic, this wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.
 
 \subsection{Hardware Experiment: Position Control of n Robots}
-The hardware setup has a bounded platform, magnetic sliders, and a magnetic guide board.
+Kilobots have too much stochasticity to implement Alg.
+\ref{alg:PosControlNRobots}.
+Instead, a hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used.
 Designs for each are available at \citep{arunhardware}.
@@
 \subsection{Hardware Experiment: Control of Covariance}
-To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
+To demonstrate covariance control of up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
@@
 \caption{\label{fig:covExperiment}
-Hardware demonstration steering 100 kilobot robots to desired covariance.
+Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance.
 The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.

### 2016-09-08 RA-L/intro.tex
Edits made to the paper

@@
   ex vivo to assemble structures in parallel and 
- through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
+ through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
@@
 Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
-By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}.
-A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.
+By adding rectangular boundary walls, some higher-order moments, such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls), are also controllable~\citep{ShahrokhiIROS2015}.
+A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.
 
 The paper is arranged as follows.
-\S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
+First \S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
@@
 \S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-After a review of recent related work \S \ref{sec:RelatedWork}, we end with directions for future research \S \ref{sec:conclusion}.
+After a review of recent related work \S \ref{sec:RelatedWork}, we end by suggesting directions for future research in \S \ref{sec:conclusion}.
 %For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 

### 2016-09-08 RA-L/theory.tex
Reducing size of pdf to less than 10 MB
compressed some images, made appendix able to be added easily.
some of Mary's Comments

@@
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
-This section examines the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
+We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$.
@@
 In this range, the swarm assumes eight different polygonal shapes.
-The shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and alternate between squares with one corner removed and trapezoids when $A$$>$1/2.
+Therse shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and between squares with one corner removed and trapezoids when $A$$>$1/2.
 
@@
 \bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} x \, dy\right) \, dx}{A} \nonumber \\
-	&=\frac{1}{3} \sqrt{2} \sqrt{A \tan (\beta )}\\
+	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
 \bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} y \, dy\right) \, dx}{A} \nonumber\\
-	&=\frac{1}{3} \sqrt{2} \sqrt{A \cot (\beta )}
+	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
 \end{align}
 The full equations are included in the appendix, and are summarized in Fig.~\ref{fig:SquareFill}.
-A few highlights are that the correlation is maximized when the swarm is in a triangular shape, and is $\pm$1/2.
+A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
 The covariance of a triangle is always $\pm(A/18)$.
@@
 Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
-Similar calculations can be computed for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
+Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
 For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
@@
 Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are summarized in Fig.~\ref{fig:CircleFill}.
+Results are displayed in Fig.~\ref{fig:CircleFill}.
 
@@
 Directly controlling covariance by pushing the swarm into a boundary requires changes to the boundary.
-An obstacle in the lower-right corner is enough to generate positive covariance.
-Generating both positive and negative covariance requires additional obstacles.
-Requiring special obstacle configuration also makes covariance control dependent on the local environment.
+An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.
+Requiring a special obstacle configuration also makes covariance control dependent on the local environment.
   Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.
@@
 Robots touching a wall have a negative friction force that opposes movement along the boundary.
-This  causes robots along the boundary to slow down compared to robots in free-space.
+This  causes robots along the boundary to move more slowly than robots in free-space.
 
@@
 \end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. As illustrated, both experiences different net forces although each receives the same inputs.
+ Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. As illustrated, both experience different net forces although each receives the same inputs.
   For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
-This means that if one robot is touching the wall and another robot is free, if the control input is parallel or into the wall, the touching robot will not move. 
+This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
 There are many alternate models of friction that also break control symmetry.
@@
 \begin{align}
-INSERT FORMULA \label{eq:boundarylayerflow}
+u(y) = u_0 [1- \frac{(y-h)^2}{h^2} = u_0 \frac{y}{h} [2- \frac{y}{h}] \label{eq:boundarylayerflow}
 \end{align}
@@
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
-The algorithm only requires friction along the bottom and left walls.
+The algorithm only requires friction along the bottom and right walls.
 Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
@@
 
-This section describes Alg.~\ref{alg:PosControl2Robots}, which uses wall-friction to arbitrarily position two robots in a rectangular workspace.
+Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
 This algorithm  introduces concepts that will be used for multi-robot positioning.
@@
 \begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
+	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardwareComp.pdf}
 \end{center}

### 2016-09-09 RA-L/experiment.tex
edits, cut to 8 pages

@@
 Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.
-In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace.
-The lights were controlled using an Arduino Uno board connected to an 8-relay shield.
-Above  the table, an overhead machine vision system tracks the position of the swarm.
+In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the midpoints and vertices of a 6 m square centered on the workspace.
+An Arduino Uno connected to an 8-relay shield controlled the lights.
+Above  the table, an overhead machine vision system tracked the position of the swarm.
 
@@
 \subsection{Hardware Experiment: Position Control of Two Robots}
-The walls of the hardware platform have almost infinite friction, due to a laser-cut, slotted border.
+The hardware platform walls have almost infinite friction, due to a laser-cut, slotted border.
 When a kilobot is steered into the slotted border, they pin themselves to the wall unless the global input directs them away from the wall.
@@
 \subsection{Hardware Experiment: Control of Covariance}
-To demonstrate covariance control of up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
+To demonstrate covariance control, up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.

### 2016-09-09 RA-L/intro.tex
some of Mary's Comments
edits, cut to 8 pages

@@
 By adding rectangular boundary walls, some higher-order moments, such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls), are also controllable~\citep{ShahrokhiIROS2015}.
-A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.
+%A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.
+These techniques could only compress a swarm orthogonal to obstacles.
+Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
 
@@
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces wall friction to the system model.
+To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces general models for boundary interaction.
 We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in \S \ref{sec:PostionControl2Robots}, and \S \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
-\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-After a review of recent related work \S \ref{sec:RelatedWork}, we end by suggesting directions for future research in \S \ref{sec:conclusion}.
+Both algorithms require robots with low process noise and require time proportional to the number of robots, so \S \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm covariance.
+\S \ref{sec:simulation} describes implementations of the algorithms in simulation and  \S \ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
+After a review of recent related work in \S \ref{sec:RelatedWork}, we end with directions for future research in \S \ref{sec:conclusion}.
 %For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 

### 2016-09-09 RA-L/simulation.tex
Edits made to the paper
edits, cut to 8 pages

@@
 
-Two simulations were implemented using wall-friction for position control.
-The first controls the position of two robots, the second controls the position of $n$ robots.
-All code is available online at \citep{Zhao2016mathematica,Zhao2016mathematicaSquare}.
+Two simulations were implemented using wall-friction for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.
 
@@
 
-Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, %\ref{alg:YControl}, 
-were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  this algorithm for two configurations. 
+Algorithms \ref{alg:PosControl2Robots} and \ref{alg:XControl}, %\ref{alg:YControl}, 
+were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
 Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
@@
 Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
-The distance moved grows quadratically with the number of robots $n$. A best-fit line $210 n^2 + 1200n-10,000$ is overlaid by the data..
+The distance moved grows quadratically with the number of robots $n$. A best-fit line $210 n^2 + 1200n-10,000$ is overlaid by the data.
 
@@
 \ref{alg:PosControlNRobots}.
-Still, we are able to control covariance of the swarm.
-This section demonstrates simulations of controlling covariance of the swarm.
-These simulations use  the 2D physics engine Box2D, by \citep{catto2010box2d}.
+Due to the central limit theorem, we can control the covariance of large swarms.
+This section demonstrates controlling swarm covariance in simulation, using the 2D physics engine Box2D~\citep{catto2010box2d}.
  144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig.

### 2016-09-09 RA-L/theory.tex
edits, cut to 8 pages

@@
 \begin{center}
+\vspace{-1em}
 \includegraphics[width=\linewidth]{CircleFill.pdf} 
@@
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$.
-Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill}.
+Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
 
-The range for the global input angle $\beta $ is [0,2$\pi $).
-In this range, the swarm assumes eight different polygonal shapes.
-Therse shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and between squares with one corner removed and trapezoids when $A$$>$1/2.
+The range for the global input angle $\beta $ is [0,2$\pi $). In this range, the swarm assumes eight different polygonal shapes. 
+These shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and between squares with one corner removed and trapezoids when $A$$>$1/2.
 
@@
 Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}.
+Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.
 
@@
 The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
-
 \begin{align}
@@
 \end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. As illustrated, both experience different net forces although each receives the same inputs.
+ Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
+Though each receives the same inputs,  they experience different net forces.
   For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.

### 2016-09-09 RA-L/experiment.tex
small other edits

@@
 \subsection{Hardware Experiment: Position Control of Two Robots}
-The hardware platform walls have almost infinite friction, due to a laser-cut, slotted border.
-When a kilobot is steered into the slotted border, they pin themselves to the wall unless the global input directs them away from the wall.
+The hardware platform walls have almost infinite friction, due to a laser-cut, scalloped border.
+When a kilobot is steered into the scalloped border, they are pinned to the wall unless the global input directs them away from the wall.
 Even though the kilobots are stochastic, this wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.
@@
 \ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2 and 7.
-All robots receive the same control inputs, but boundary interactions breaks the control symmetry.
+All robots receive the same control inputs, but boundary interactions break the control symmetry.
 Robots reach their respective goal position in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.

### 2016-09-09 RA-L/simulation.tex
small other edits

@@
 were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
-Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
+Robot initial positions are shown by crosshairs, and final positions by circled crosshairs.
 Dashed lines show the shortest route if robots could be controlled independently.
@@
 Fig.
-\ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ axis were controlled from a set of initial conditions.
+\ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ directions were controlled from a set of initial conditions.
 \begin{figure}

### 2016-09-09 RA-L/theory.tex
Added appendix link
small other edits

@@
 \end{align}
-The full equations are included in the appendix, and are summarized in Fig.~\ref{fig:SquareFill}.
+The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
 A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
@@
 First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
-Second, a set of $\operatorname{DriftMove()}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
-Third, a set of $\operatorname{DriftMove()}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
+Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
+Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
 Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
@@
 Assume an open workspace with infinite boundary friction.
-Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy}$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$.
+Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$.
 For our experiments, $c_1 = 0.1$.
@@
 \item swarm's mean position is moved to the center of the workspace
-\item swarm is pushed into the bottom wall until $\sigma_y^2 \le \sigma_{goalt}^2$. 
+\item swarm is pushed into the bottom wall until $\sigma_y^2 \le \sigma_{goaly}^2$. 
 \item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\

### 2016-09-10 RA-L/experiment.tex
added thinigverse
Full revision pass.

@@
 Instead, a hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used.
-Designs for each are available at \citep{arunhardware}.
+Designs for each are available at \citep{Arun2016Thingiverse}.
 The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
@@
 Fig.
-\ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2 and 7.
+\ref{fig:construction2d} follows the motion of the sliders through iterations  $k$=1, 2, and 7.
 All robots receive the same control inputs, but boundary interactions break the control symmetry.

### 2016-09-10 RA-L/intro.tex
Full revision pass.

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
+Particle swarms propelled by a global force are common in applied mathematics, biology, and computer graphics. 
 \begin{align}
@@
 \end{align}
-The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
+Our control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
 As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
-  ex vivo to assemble structures in parallel and 
- through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
+(1) to assemble structures in parallel and 
+ (2) through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.

### 2016-09-10 RA-L/simulation.tex
Full revision pass.

@@
 The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
-Each row has five snapshots taken every quarter second. For the sake of brevity axis-aligned moves were replaced with oblique moves that combine two moves simultaneously. 
- $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
- The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
- $\Delta r_y$ is corrected by $t = 0.75$. 
- Finally, the algorithm moves the robots to their corresponding destinations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
-Similarly,  two iterations %of Alg. \ref{alg:YControl} 
-are only required if $|\Delta e_y - \Delta s_y|>L$. An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
+%Each row has five snapshots taken every quarter second. 
+For the sake of brevity axis-aligned moves were replaced with oblique moves that combine two moves simultaneously.
+% THES IMAGES DON'T DO THIS ANYMORE 
+% $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
+% The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
+% $\Delta r_y$ is corrected by $t = 0.75$. 
+% Finally, the algorithm moves the robots to their corresponding destinations.
+An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
 
@@
 \ref{alg:PosControlNRobots}.
-Due to the central limit theorem, we can control the covariance of large swarms.
+However, due to the central limit theorem, we can control the covariance of large swarms.
 This section demonstrates controlling swarm covariance in simulation, using the 2D physics engine Box2D~\citep{catto2010box2d}.
- 144 disc-shaped robots were controlled by an open-loop control input as illustrated in Fig.
+ 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow} were controlled by an open-loop control input as illustrated in Fig.
 \ref{fig:SimCovarianceFuncFrictionOpenLoop}.
@@
 
- 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl}.
+ 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl} with $c_1=0.1$.
 Fig.

### 2016-09-10 RA-L/theory.tex
Changed some wrong notations
Full revision pass.

@@
 
-The region of integration $R$ is the polygon containing the swarm. If the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
+The region of integration $R$ is the polygon containing the swarm. If $A<1/2$ and the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
 \begin{align}\label{eq:meanInSquareWorkspaceLL}
@@
 Global inputs move a swarm uniformly.  
-Controlling covariance requires breaking this uniform symmetry.
+Shape control requires breaking this uniform symmetry.  
 A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
-Directly controlling covariance by pushing the swarm into a boundary requires changes to the boundary.
+Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
 An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
-Requiring a special obstacle configuration also makes covariance control dependent on the local environment.
+Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
 Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.
 These frictional forces are  sufficient to break the symmetry caused by uniform inputs.
-Robots touching a wall have a negative friction force that opposes movement along the boundary.
+Robots touching a wall have a friction force that opposes movement along the boundary.
 This causes robots along the boundary to move more slowly than robots in free-space.
@@
 \begin{align}
-u(y) = u_0 [1- \frac{(y-h)^2}{h^2} = u_0 \frac{y}{h} [2- \frac{y}{h}] \label{eq:boundarylayerflow}
+%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
+F_{forward}(y) &= F - F_f\begin{cases}  \left( \frac{y-h}{h} \right)^2  , &  y<h \label{eq:boundarylayerflow} \\
+0, & \text{else} \end{cases}
 \end{align}
@@
 \vspace{-1em}
-\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.}
+\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
 \label{fig:friction}
+}
 \end{center}
@@
 
-Our algorithm solves the positioning problem in two steps: 
+Our algorithm solves the positioning problem in three steps: 
 First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-Second $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant.% in Alg.~\ref{alg:YControl}.
+Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
+Third, the robots, now correctly positioned relative to each other, are moved to their goal locations.
+
+In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
+Similarly,  two iterations %of Alg. \ref{alg:YControl} 
+are only required if $|\Delta e_y - \Delta s_y|>L$.
 
@@
  $L$ is length of the walls. 
- Current position of the robots are $(r_1,r_2)$.
+ Current robot positions are $(r_1,r_2)$.
 
-\State ($r_1,r_2$) = GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
+\State  GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
 \State  GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
+\State Move robots $e_1 - r_1$
 
@@
 \begin{algorithmic}[1]
-\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
+\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
 $(0,0)$ is bottom corner, $s_1$ is topmost robot, 
@@
 \State $ \Delta e_x \gets e_{1x} - e_{2x} $
-\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
+%\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
 \If {$\Delta e_x < 0 $ }
-\State $ m \gets ( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Move to right wall}
+\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{To right wall}
 \Else 
-\State  $ m \gets ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
+\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
 \EndIf
-\State $m  \gets  m + (0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
-\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+\State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
+%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
 \If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
-\State $ m \gets (\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
+\State move$(\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
 \Else
-\State $ m \gets (-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
+\State move$(-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
 \EndIf 
-\State $m  \gets  m + (0, \epsilon)$ \Comment{Move up}
-\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-\State $\Delta r_x = r_{1x} - r_{2x}$
-\If {$\Delta r_x \equiv \Delta e_x$} 
+\State move$(0, \epsilon)$ \Comment{Move up}
+%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+%\State $\Delta r_x = r_{1x} - r_{2x}$
+\If {$r_{1x} - r_{2x} \equiv \Delta e_x$} 
 %\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
@@
 \begin{center}
-	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:driftmove}
-A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
-}
-\end{figure}
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardwareComp.pdf}
+	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
 \end{center}
@@
 Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.
-The minimum workspace is then $(\epsilon + 2r + \max(w_f,w_s), 2\epsilon + h_s,h_f)$.
+The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.
 
@@
 
+\begin{figure}
+\begin{center}
+	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:driftmove}
+A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
+Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
+\end{figure}
+
 Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot.
-Label the robots in the staging zone from left-to-right and top-to-bottom, and the $f_k$ configurations right-to-left and top-to-bottom as shown in Fig.~\ref{fig:construction2d}.
+Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.
 
@@
 \begin{algorithmic}[1]
-\State move( $-\epsilon, r-p_{k,y}$) % move  away from right wall and down till robot k touches bottom
+\State move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
 
-\While{ $p_{k,x} > r$} 
-\State $\operatorname{DriftMove}(\epsilon, \min(p_{k,x} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
+\While{ $p_{kx} > r$} 
+\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
 \EndWhile
 
-\State $m \gets \operatorname{ceil}(\frac{f_{k,y}-r}{\epsilon})$
-\State $\beta \gets \frac{f_{k,y}-r}{m}$
-\State $\alpha \gets \beta - \frac{r - p_{k,y}-\epsilon}{m}$
+\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
+\State $\beta \gets \frac{f_{ky}-r}{m}$
+\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
 \For{ $m$ iterations}
@@
 
-\State move($r+\epsilon-f_{k,x}, 0$)  % move the group to the left until k is in the correct relative x position
-\State move($f_{k,x}-r, 0$)
+\State move($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
+\State move($f_{kx}-r, 0$)
 
@@
 Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
-Finally, all robots are moved right until robot $k$ is in the desired target position.
+Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.
 
@@
 Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$.
-For our experiments, $c_1 = 0.1$.
 %$\sigma_x^2$ decreases when the swarm is pushed into vertical walls, and $\sigma_y^2$ decreases when the swarm is pushed into horizontal walls. Alg.~\ref{alg:CovCont} procedes as follows:
 \begin{enumerate}
-\item swarm is pushed into the left wall until $\sigma_x^2< c_1\sigma_{goalx}^2$.  
+\item swarm pushed into the left wall until $\sigma_x^2 \le c_1\sigma_{goalx}^2$.  
 \item swarm's mean position is moved to the center of the workspace
-\item swarm is pushed into the bottom wall until $\sigma_y^2 \le \sigma_{goaly}^2$. 
+\item swarm pushed into the bottom wall until $\sigma_y^2 \le  \sigma_{goaly}^2$. 
 \item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\
@@
 \end{enumerate}
+The constant $c_1$ in step 1 should be less than 1 to ensure $\sigma_x^2$ stays within tolerance during the subsequent moves.
 

### 2016-09-10 RA-L/intro.tex
added reference to multimedia

@@
 Navigating a swarm using global inputs is challenging because each member receives the same control inputs.
-This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.} 
+This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.
+See multimedia attachment for overview of the experiments.}
 \end{figure}

### 2016-09-21 JournalTorqueControl/code/Data/DrawTorquePlot.m
Added the folder for torque control journal

@@ -1,0 +1,173 @@
+% load TSC10D.txt;
+% x = TSC10D(:,1);
+% y = TSC10D(:,2);
+% plot(x,y)
+% hold on
+% load TSC10S.txt;
+% x2 = TSC10S(:,1);
+% y2 = TSC10S(:,2);
+% plot(x2,y2)
+% 
+% load TSC10Up.txt;
+% x3 = TSC10Up(:,1);
+% y3 = TSC10Up(:,2);
+% plot(x3,y3)
+% 
+% 
+% % load Straight4.txt;
+% % x4 = Straight4(:,1);
+% % y4 = Straight4(:,2);
+% % plot(x4,y4)
+% % 
+% % load Straight5.txt;
+% % x5 = Straight5(:,1);
+% % y5 = Straight5(:,2);
+% % plot(x5,y5)
+% line([11 30] , [0 0]);
+% hold off
+% load OrientationCont2.txt;
+% x = OrientationCont2(:,1);
+% y = OrientationCont2(:,2);
+% z = OrientationCont2(:,3);
+% plot(x,y)
+% hold on
+% plot (x,z)
+% 
+% load OrientationCont3.txt;
+% x2 = OrientationCont3(:,1);
+% y2 = OrientationCont3(:,2);
+% z2 = OrientationCont3(:,3);
+% plot(x2,y2)
+% 
+% plot (x2,z2)
+% load OrientationCont4.txt;
+% x3 = OrientationCont4(:,1);
+% y3 = OrientationCont4(:,2);
+% z3 = OrientationCont4(:,3);
+% plot(x3,y3)
+% 
+% plot (x3,z3)
+% hold off
+
+
+% load PivotFull.txt;
+% x3 = PivotFull(:,1);
+% y3 = PivotFull(:,2);
+% plot(x3,y3, 'color', 'red')
+% hold on
+% load PivotHalf.txt;
+% x2 = PivotHalf(:,1);
+% y2 = PivotHalf(:,2);
+% plot(x2,y2, 'color', 'blue')
+% 
+% 
+% 
+% load PivotQuart.txt;
+% x = PivotQuart(:,1);
+% y = PivotQuart(:,2);
+% 
+% plot(x,y, 'color','green')
+% 
+% load PivotZero.txt;
+% x4 = PivotZero(:,1);
+% y4 = PivotZero(:,2);
+% 
+% plot(x4,y4, 'color', 'cyan')
[…]

### 2016-09-26 JournalTorqueControl/theory.tex
small changes to the mathematica and theory

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torque Control}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Controlling Torque}

The orientation of an object's major axis is important when a swarm is manipulating a non-symmetric object through narrow corridors. 
Orientation is controllable by applying torque to the object. 
To change the output torque $\tau$ in Eq.~\eqref{eq:torque}, we can choose the direction and magnitude of the force applied $F$, and the moment arm from the object's center of mass (COM) to the point of contact $r$.

\begin{equation}
\tau = F \times r\label{eq:torque}
\end{equation}
The swarm version of \eqref{eq:torque} is the summation of the forces contributed by individual robots.

\begin{align}
\tau_{total} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O )   \label{eq:swarmtorque}\\
F_{total} &= \sum\limits_{i=1}^n \rho_i F_i  \label{eq:swarmforce}
\end{align}

Here $F_i$ is the force that the $i$th robot applies.  If all robots are identical and the control input is uniform, the force is equivalent for every robot and $F_i = F_c$.
Not all robots are in contact with the object.  $\rho_i$ is an indicator variable: $\rho_i$ is 1 if the robot is in direct contact with the object or touching a chain of robots where at least one robot is in contact with the object. Otherwise $\rho_i = 0$.
The moment arm is the robot's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$.

\begin{equation}
P(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right. \kern-\nulldelimiterspace} {2\sigma ^2 }}}
\end{equation}
\begin{equation}
\tau = \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right. \kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx
\end{equation}

\begin{equation}
\frac{1}{2} \mu  (\text{erf}(1.41421 (1-\mu ))+\text{erf}(1.41421 \mu ))+0.199471 \left(e^{-2. \mu
   ^2}-e^{-2. (\mu -1)^2}\right)
\end{equation}

### 2016-10-05 JournalTorqueControl/theory.tex
Added some written theory
Adding not working triangle
Added a bunch of plots and equations.

@@
 The moment arm is the robot's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$.
-
+Consider a normal distribution, 
+%normal distribution
 \begin{equation}
-P(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+P(x)_{n} = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
 \kern-\nulldelimiterspace} {2\sigma ^2 }}}
 \end{equation}
-\begin{equation}
-\tau = \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx
-\end{equation}
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{Normal.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:pdfNorm}
+PDF of Normal distributions with the same $\mu$ and different $\sigma$s.
+}
+\vspace{-1em}
+\end{figure}
 
-\begin{equation}
-\frac{1}{2} \mu  (\text{erf}(1.41421 (1-\mu ))+\text{erf}(1.41421 \mu ))+0.199471 \left(e^{-2. \mu
-   ^2}-e^{-2. (\mu -1)^2}\right)
-\end{equation}
+%defining torque
+\begin{align} \nonumber
+\tau &= \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx\\ \nonumber
+&= \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-\mu^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}\\ 
+&+ \frac{1}{2}\mu(erf(\frac{1-\mu}{\sqrt{2}\sigma})+erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\end{align}
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{TorqueNormal.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:torqueNorm}
+Torque as a function of the mean position, $\mu$ which is the pushing location when the swarm is normally distributed with different standard deviations.
+}
+\vspace{-1em}
+\end{figure}
+%taking derivative of torque
+\begin{align} 
+\frac{d\tau}{d\mu} &= -\frac{e^{-\frac{{-1+\mu}^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma} + \frac{1}{2}(erf(\frac{1-\mu}{\sqrt{2}\sigma})+erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\end{align}
+
+%triangular distribution
+\begin{align}
+P(x)_{t} &=  \left\{
+\begin{array}{ll}
+    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
+     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{Triangular.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:pdfTri}
+PDF of Triangular distributions with the same $\mu$ and different $\sigma$s.
+}
+\vspace{-1em}
+\end{figure}
+%defining torque
+\begin{align}
+\tau_t &=  \left\{
+\begin{array}{ll}
+    \mu, &  \mu<1 \& \sqrt{6} \mu + 6 \sigma < \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
+    \frac{2-3\mu+3\sqrt{6}\sigma}{36\sigma^2}, & (\mu>1\&\sqrt{6}\mu \leq 6\sigma) \| \mu =1 \& \sigma \geq \frac{1}{\sqrt{6}})\\
+    \frac{-2+3\mu-2\mu^3+3\sqrt{6}\sigma}{36\sigma^2}, &  \mu<1 \& \sqrt{6}\mu+6\sigma\geq \sqrt{6} \& \sqrt{6} \mu \leq 6\sigma\\
+    \frac{1}{2}-\frac{\sigma}{\sqrt{6}}, &   \mu = 1 \& \sigma < \frac{1}{\sqrt{6}}\\
+    -\frac{2+\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)-3\mu(1+6\sigma^2}{36\sigma^2} & \mu<1 \& \sqrt{6} \mu + 6 \sigma \geq \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
+    \frac{2+\mu^3-3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(1-2\sigma^2)+3\mu(-1+6\sigma^2}{36\sigma^2} & \mu>1 \& 0<\sqrt{6}\mu-6\sigma<\sqrt{6}\\
+    \frac{1}{36}(18\mu-\frac{\mu^3}{\sigma^2} + \frac{3\sqrt{6}\mu^2}{\sigma}+6\sqrt{6}\sigma) &  \sqrt{6} \mu+6\sigma<\sqrt{6}\&\sqrt{6}\mu\leq 6\sigma\\
+     0, & \textrm{otherwise}\\ 
+\end{array} 
+\right.
+\end{align}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{TorqueTri.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:torqueTri}
+Torque as a function of the mean position, $\mu$ which is the pushing location when the swarm is distributed by triangular distribution with different standard deviations.
+}
+\vspace{-1em}
+\end{figure}
+
+%uniform distribution
+\begin{align}
+P(x)_{u} &=  \left\{
+\begin{array}{ll}
+    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu-\sqrt{3}\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{Uniform.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:pdfUni}
+The Uniform distribution with different standard deviations and the same mean.
+}
+\vspace{-1em}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{TorqueUni.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:torqueUni}
+Torque as a function of the mean position, $\mu$ which is the pushing location when the swarm is uniformly distributed with different standard deviations.
+}
+\vspace{-1em}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{BestLocationPivot.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:besLocPiv}
+Comparison for three different distributions for the best location to push the object.
+}
+\vspace{-1em}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{MaxTorquePivot.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:maxTorPiv}
+Comparison for three different distributions for the maximum torque possible.
+}
+\vspace{-1em}
+\end{figure}

### 2016-10-05 JournalTorqueControl/theory.tex
tiny changes
Added remaining plots and figures
Added text, changed plots

@@
 The moment arm is the robot's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$.
-Consider a normal distribution, 
+
+\section{Instantiation of torque from robot distribution}
+Consider a swarm of robots with probability density $p(x)$.
+This section examines where to steer the mean of the probability distribution to maximize torque.
+We examine two problems.
+First, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is:
+\begin{equation}
+\tau_{pivot} = \int_0^1 xp(x)\, dx
+\end{equation}
+Second, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
+\begin{equation}
+\tau_{free} = \int_{-1}^1 xp(x)\, dx
+\end{equation}
+This section considers three canonical probability distributions: uniform, triangular, and normal.
+
+%uniform distribution
+\begin{align}
+P(x)_{u} &=  \left\{
+\begin{array}{ll}
+    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu-\sqrt{3}\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+
 %normal distribution
 \begin{equation}
-P(x)_{n} = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+p_n(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
 \kern-\nulldelimiterspace} {2\sigma ^2 }}}
@@
 %defining torque
+
+\begin{strip}
 \begin{align}
@@
 \begin{array}{ll}
-    \mu, &  \mu<1 \& \sqrt{6} \mu + 6 \sigma < \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
-    \frac{2-3\mu+3\sqrt{6}\sigma}{36\sigma^2}, & (\mu>1\&\sqrt{6}\mu \leq 6\sigma) \| \mu =1 \& \sigma \geq \frac{1}{\sqrt{6}})\\
-    \frac{-2+3\mu-2\mu^3+3\sqrt{6}\sigma}{36\sigma^2}, &  \mu<1 \& \sqrt{6}\mu+6\sigma\geq \sqrt{6} \& \sqrt{6} \mu \leq 6\sigma\\
-    \frac{1}{2}-\frac{\sigma}{\sqrt{6}}, &   \mu = 1 \& \sigma < \frac{1}{\sqrt{6}}\\
-    -\frac{2+\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)-3\mu(1+6\sigma^2}{36\sigma^2} & \mu<1 \& \sqrt{6} \mu + 6 \sigma \geq \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
-    \frac{2+\mu^3-3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(1-2\sigma^2)+3\mu(-1+6\sigma^2}{36\sigma^2} & \mu>1 \& 0<\sqrt{6}\mu-6\sigma<\sqrt{6}\\
-    \frac{1}{36}(18\mu-\frac{\mu^3}{\sigma^2} + \frac{3\sqrt{6}\mu^2}{\sigma}+6\sqrt{6}\sigma) &  \sqrt{6} \mu+6\sigma<\sqrt{6}\&\sqrt{6}\mu\leq 6\sigma\\
+    \mu, &   \textrm{for     }\mu<1 \& \sqrt{6} \mu + 6 \sigma < \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
+    \frac{2-3\mu+3\sqrt{6}\sigma}{36\sigma^2}, & \textrm{for     } (\mu>1\&\sqrt{6}\mu \leq 6\sigma) | \mu =1 \& \sigma \geq \frac{1}{\sqrt{6}})\\
+    \frac{-2+3\mu-2\mu^3+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } \mu<1 \& \sqrt{6}\mu+6\sigma\geq \sqrt{6} \& \sqrt{6} \mu \leq 6\sigma\\
+    \frac{1}{2}-\frac{\sigma}{\sqrt{6}}, &   \textrm{for     } \mu = 1 \& \sigma < \frac{1}{\sqrt{6}}\\
+    -\frac{2+\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)-3\mu(1+6\sigma^2}{36\sigma^2} &  \textrm{for     } \mu<1 \& \sqrt{6} \mu + 6 \sigma \geq \sqrt{6} \& \sqrt{6} \mu > 6 \sigma\\
+    \frac{2+\mu^3-3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(1-2\sigma^2)+3\mu(-1+6\sigma^2}{36\sigma^2} &  \textrm{for     } \mu>1 \& 0<\sqrt{6}\mu-6\sigma<\sqrt{6}\\
+    \frac{1}{36}(18\mu-\frac{\mu^3}{\sigma^2} + \frac{3\sqrt{6}\mu^2}{\sigma}+6\sqrt{6}\sigma) &  \textrm{for     }  \sqrt{6} \mu+6\sigma<\sqrt{6}\&\sqrt{6}\mu\leq 6\sigma\\
      0, &  \textrm{otherwise}\\ 
@@
 \end{align}
+\end{strip}
 
@@
 }
-\vspace{-1em}
+\vspace{2em}
 \end{figure}
 
-%uniform distribution
-\begin{align}
-P(x)_{u} &=  \left\{
-\begin{array}{ll}
-    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu-\sqrt{3}\\
-     0, & \textrm{otherwise}\\
-\end{array} 
-\right.
-\end{align}
-
+\newpage
 \begin{figure}
@@
 \end{center}
-\vspace{-1em}
+%\vspace{-1em}
 \caption{\label{fig:pdfUni}
@@
 \end{figure}
-
+%uniform torque
+\begin{align}
+\mu = \textrm{Max}[1-\sqrt{3}\sigma,\sqrt{3}\sigma]
+\end{align}
 \begin{figure}
@@
 
+\newpage
+
+\section{Free object torque}
+%torque normal free
+
+\begin{strip}
+\begin{align} \nonumber
+\tau = \int_{-1}^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx = \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-(1+\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}+ \frac{1}{2}\mu(erf(\frac{1-\mu}{\sqrt{2}\sigma})+erf(\frac{1+\mu}{\sqrt{2}\sigma}))
+\end{align}
+\end{strip}
+%torque triangular free
+
+\begin{strip}
+\begin{align}
+\tau_t &=  \left\{
+\begin{array}{ll}
+    \mu, &   \textrm{for     }\sqrt{6}(1+\mu)\geq 6\sigma \& \sqrt{6}\mu + 6 \sigma < \sqrt{6} \& \sigma>0\\
+     -\frac{1}{9\sigma^2},   & \textrm{for     } (\mu=-1\&3\sigma \geq \sqrt{6} | \mu <-1 \& \sqrt{6}\mu+6\sigma > \sqrt{6})\\
+   \frac{1}{9\sigma^2}, & \textrm{for     } (\mu>1\&\sqrt{6}(1+\mu) \leq 6\sigma) | \mu =1 \& 3\sigma \geq \sqrt{6})\\
+    -\frac{\mu(-3+\mu^2)}{18\sigma^2}, & -1<\mu<1 \& \sqrt{6}(1+\mu) <6\sigma \& \sqrt{6}\mu + 6\sigma\geq\sqrt{6}\\
+    \frac{1}{2}-\frac{\sigma}{\sqrt{6}}, & \textrm{for     } \mu = 1 \& 0< \sigma < \sqrt{\frac{2}{\sqrt{3}}}\\
+    -\frac{1}{2}+\frac{\sigma}{\sqrt{6}}, & \textrm{for     } \mu = -1 \&0< \sigma < \sqrt{\frac{2}{\sqrt{3}}}\\
+    -\frac{2+\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)-3\mu(1+6\sigma^2}{36\sigma^2} &  \textrm{for     }  \sqrt{6} \mu+6\sigma\geq\sqrt{6}\&\mu<1\&\sqrt{6}(1+\mu)\geq 6\sigma\\
+   \frac{2+\mu^3-3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(1-2\sigma^2)+3\mu(-1+6\sigma^2}{36\sigma^2} &  \textrm{for     } \mu>1 \& -\sqrt{6}< \sqrt{6}\mu-6\sigma< \sqrt{6} \\
+     \frac{2-\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)+3\mu(1+6\sigma^2}{36\sigma^2} &  \textrm{for     } \mu>-1 \& -\sqrt{6}(1+\mu)<6\sigma\& \sqrt{6}\mu+6\sigma< \sqrt{6} \\
+     \frac{-2+\mu^3+3\sqrt{6}\mu^2\sigma+3\sqrt{6}\sigma(-1+2\sigma^2)+3\mu(-1+6\sigma^2}{36\sigma^2} &  \textrm{for     } \mu<-1 \& -\sqrt{6}< \sqrt{6}\mu+6\sigma< \sqrt{6} \\
+     0, &  \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+\end{strip}
+
+%torque uniform free
+
+%pivoted comparisons:
+
 \begin{figure}
@@
 }
-\vspace{-1em}
+\vspace{0em}
 \end{figure}
@@
 \end{center}
-\vspace{-1em}
+%\vspace{0em}
 \caption{\label{fig:maxTorPiv}
@@
 }
+\vspace{0em}
+\end{figure}
+
+%free comparisons
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{TorqueFreeComparison.pdf}
+\end{center}
 \vspace{-1em}
+\caption{\label{fig:maxTorqueFree}
+Comparison for three different distributions for the maximum torque possible.
+}
+\vspace{0em}
 \end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{TorqueFreeBestPush.pdf}
+\end{center}
+%\vspace{0em}
+\caption{\label{fig:bestLocFree}
+Comparison for three different distributions for the best location to push the object.
+}
+\vspace{0em}
+\end{figure}

### 2016-10-07 JournalTorqueControl/theory.tex
Added limits and equations and plots.

--- 2016-09-26 JournalTorqueControl/theory.tex
@@
 
+\section{Instantiation of torque from robot distribution}
+Consider a swarm of robots with probability density $p(x)$.
+This section examines where to steer the mean of the probability distribution to maximize torque.
+We examine two problems.
+First, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is:
 \begin{equation}
-P(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\tau_{pivot} = \int_0^1 xp(x)\, dx
+\end{equation}
+Second, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
+\begin{equation}
+\tau_{free} = \int_{-1}^1 xp(x)\, dx
+\end{equation}
+This section considers three canonical probability distributions: uniform, triangular, and normal.
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.66\columnwidth}
+\begin{overpic}[width =\figwid]{Uniform.pdf}%\put(1,55){a)}
+\end{overpic}
+\begin{overpic}[width =\figwid]{Triangular.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{Normal.pdf}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:PDF} Three different distributions are examined in this work: uniform, triangular, and normal.
+%\vspace{-2em}
+}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.66\columnwidth}
+\begin{overpic}[width =\figwid]{TorqueUni.pdf}%\put(1,55){a)}
+\end{overpic}
+\begin{overpic}[width =\figwid]{TorqueTri.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{TorqueNormal.pdf}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:torque} Torque with regard to mean position, $\mu$. Mean position is the pushing location.
+%\vspace{-2em}
+}
+\end{figure*}
+%uniform distribution
+\begin{align}
+p_u(x) &=  \left\{
+\begin{array}{ll}
+    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu+\sqrt{3}\sigma\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+
+%triangular distribution
+\begin{align}
+p_t(x) &=  \left\{
+\begin{array}{ll}
+    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
+     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+%normal distribution
+\begin{equation}
+p_n(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
 \kern-\nulldelimiterspace} {2\sigma ^2 }}}
 \end{equation}
-\begin{equation}
-\tau = \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx
-\end{equation}
 
-\begin{equation}
-\frac{1}{2} \mu  (\text{erf}(1.41421 (1-\mu ))+\text{erf}(1.41421 \mu ))+0.199471 \left(e^{-2. \mu
-   ^2}-e^{-2. (\mu -1)^2}\right)
-\end{equation}
+%defining torque uniform distribution
+\begin{align}
+l &= \max{0,\mu -\sqrt{3} \sigma}\\
+u &= \min{1,\mu+\sqrt{3}\sigma}\\
+\tau_u &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
+\end{align}
+
+%define torque for triangular:
+\begin{align}
+l_1 &= \max{0,\mu-\sqrt{6}\sigma}\\ \nonumber
+l_2 &= \max{0,\mu}\\ \nonumber
+u_1 &= \min{1,\mu}\\ \nonumber
+u_2 &= \min{1,\mu+\sqrt{6}\sigma} \nonumber
+\end{align}
+\begin{align}
+\tau_t &=  \left\{
+\begin{array}{ll}
+\frac{-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_1 > l_1\\
+\frac{2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)+{u_2}^2(-2u_1 + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_2 > l_2\\
+\end{array} 
+\right.
+\end{align}
+
+%defining torque normal distribution
+\begin{align} \nonumber
+\tau_n &= \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx\\ \nonumber
+&= \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-\mu^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}\\ 
+&+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\end{align}
+
+%taking derivative of torque
+\begin{align} 
+\frac{d\tau}{d\mu} &= -\frac{e^{-\frac{{-1+\mu}^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma} + \frac{1}{2}(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\end{align}
+
+%%uniform torque
+%\begin{align}
+%\mu = \textrm{Max}[1-\sqrt{3}\sigma,\sqrt{3}\sigma]
+%\end{align}
+
+\section{Free object torque}
+%torque uniform free
+\begin{align}
+l &= \max{-1,\mu -\sqrt{3} \sigma}\\
+u &= \min{1,\mu+\sqrt{3}\sigma}\\
+\tau_{uf} &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
+\end{align}
+%torque triangular free
+\begin{align}
+l_1 &= \max{-1,\mu-\sqrt{6}\sigma}\\ \nonumber
+l_2 &= \max{-1,\mu}\\ \nonumber
+u_1 &= \min{1,\mu}\\ \nonumber
+u_2 &= \min{1,\mu+\sqrt{6}\sigma} \nonumber
+\end{align}
+\begin{align}
+\tau_{tf} &=  \left\{
+\begin{array}{ll}
+\frac{-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_1 > l_1\\
+\frac{2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)+{u_2}^2(-2u_1 + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_2 > l_2\\
+\end{array} 
+\right.
+\end{align}
+
+%torque normal free
+
+\begin{align} \nonumber
+\tau_{nf} &= \int_{-1}^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx\\
+ &= \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-(1+\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}\\
+ &+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{1+\mu}{\sqrt{2}\sigma})) 
+\end{align}
+
+%limits and equations
+
+Consider robots are uniformly distributed. Finding the best place to guide the mean position:
+% best place to push for uniform
+\begin{align}
+\left\{
+\begin{array}{ll}
+\mu = 1-\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
+\mu = \sqrt{3}\sigma &   \textrm{otherwise}\\
+\end{array} 
+\right.
+\end{align}
+%best place to push for triangular:
+
+%best place to push for normal distribution:
+For triangular and normal distributions, the best place to push if $\sigma = 0$ is 1.
+The best place to push moves in the $x$ direction when $\sigma$ gets bigger.
+We used L'hopital rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
+It was because the first derivative had undefined numerator and denominator and we were not able to calculate it.
+The result is:
+
+\begin{align}
+\lim_{\sigma\to\infty} \frac{d^3x}{d\mu}(\tau_n)= -1+\frac{3\mu}{2}
+\end{align}
+
+which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
+
+%pivoted comparisons:
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{BestLocationPivot.pdf}%\put(1,55){a)}
+\end{overpic}
+\begin{overpic}[width =\figwid]{TorqueFreeBestPush.pdf}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:bestLoc} Best location to push in two different situations: when the object is pivoted, and when the object is free.
+%\vspace{-2em}
+}
+\end{figure*}
+
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{BestLocationPivot.pdf}
+%\end{center}
+%\vspace{-1em}
+%\caption{\label{fig:besLocPiv}
+%Comparison for three different distributions for the best location to push the object.
+%}
+%\vspace{0em}
+%\end{figure}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{MaxTorquePivot.pdf}%\put(1,55){a)}
+\end{overpic}
+\begin{overpic}[width =\figwid]{TorqueFreeComparison.pdf}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:maxTorque} Maximum torque possible in two situations: pivoted object and free object.
+%\vspace{-2em}
+}
+\end{figure*}
+
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{MaxTorquePivot.pdf}
+%\end{center}
+%%\vspace{0em}
+%\caption{\label{fig:maxTorPiv}
+%Comparison for three different distributions for the maximum torque possible.
+%}
+%\vspace{0em}
+%\end{figure}
+%
+%%free comparisons
+%
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{TorqueFreeComparison.pdf}
+%\end{center}
+%\vspace{-1em}
+%\caption{\label{fig:maxTorqueFree}
+%Comparison for three different distributions for the maximum torque possible.
+%}
+%\vspace{0em}
+%\end{figure}
+
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{TorqueFreeBestPush.pdf}
+%\end{center}
+%%\vspace{0em}
+%\caption{\label{fig:bestLocFree}
+%Comparison for three different distributions for the best location to push the object.
+%}
+%\vspace{0em}
+%\end{figure}

### 2016-10-11 JournalTorqueControl/theory.tex
Pivot Torque written

@@
 
-\section{Instantiation of torque from robot distribution}
+\section{Instantiating torque from swarm distribution}
 Consider a swarm of robots with probability density $p(x)$.
@@
 \begin{equation}
-\tau_{pivot} = \int_0^1 xp(x)\, dx
+\tau_{pivot} = \int_0^1 x\,p(x)\, dx
 \end{equation}
@@
 \begin{equation}
-\tau_{free} = \int_{-1}^1 xp(x)\, dx
+\tau_{free} = \int_{-1}^1 x\,p(x)\, dx
 \end{equation}
-This section considers three canonical probability distributions: uniform, triangular, and normal.
+This section considers three canonical probability distributions: uniform, triangular, and normal, all parameterized by mean, $\mu$ and standard deviation, $\sigma$. They are described by:
 
@@
 \end{array} 
-\right.
-\end{align}
-
+\right.\\
 %triangular distribution
-\begin{align}
 p_t(x) &=  \left\{
@@
 \end{array} 
-\right.
+\right.\\
+%normal distribution
+p_n(x) &= \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}}
 \end{align}
-%normal distribution
-\begin{equation}
-p_n(x) = \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}}
-\end{equation}
 
+\subsection{Pivoted object torque}
+For calculating torque in pivoted object when the swarm has a uniform distribution, we defined $l$ and $u$ for simplicity of the following derivations. Therefore, the torque applied is:
 %defining torque uniform distribution
 \begin{align}
-l &= \max{0,\mu -\sqrt{3} \sigma}\\
-u &= \min{1,\mu+\sqrt{3}\sigma}\\
+l &= \max(0,\mu -\sqrt{3} \sigma) \, , u = \min({1,\mu+\sqrt{3}\sigma})\\
 \tau_u &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
 \end{align}
-
+To simplify the following derivations for triangular distribution we use:
 %define torque for triangular:
 \begin{align}
-l_1 &= \max{0,\mu-\sqrt{6}\sigma}\\ \nonumber
-l_2 &= \max{0,\mu}\\ \nonumber
-u_1 &= \min{1,\mu}\\ \nonumber
-u_2 &= \min{1,\mu+\sqrt{6}\sigma} \nonumber
+l_1 &= \max({0,\mu-\sqrt{6}\sigma})\, , l_2 = \max({0,\mu})\\ \nonumber
+u_1 &= \min({1,\mu})\, , u_2 = \min({1,\mu+\sqrt{6}\sigma})\\ \nonumber
+\tau_t &= \frac{\tau_{left} + \tau_{right}}{36\sigma^2}\nonumber
 \end{align}
+where $\tau_{left}$ and $\tau_{right}$ are defined as:
 \begin{align}
-\tau_t &=  \left\{
-\begin{array}{ll}
-\frac{-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_1 > l_1\\
-\frac{2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)+{u_2}^2(-2u_1 + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_2 > l_2\\
-\end{array} 
-\right.
+\tau_{left} &=  -2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)\\ \nonumber
+&+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma), & \textrm{for     } u_1 > l_1\\ \nonumber
+\tau_{right} &= 2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
+&+{u_2}^2(-2u_2 + 3\mu+3\sqrt{6}\sigma),  & \textrm{for     } u_2 > l_2\\ \nonumber
 \end{align}
-
+Also, torque when swarm is normally distributed is defined by:
 %defining torque normal distribution
 \begin{align} \nonumber
-\tau_n &= \int_0^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx\\ \nonumber
-&= \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-\mu^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}\\ 
-&+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
-\end{align}
-
-%taking derivative of torque
-\begin{align} 
-\frac{d\tau}{d\mu} &= -\frac{e^{-\frac{{-1+\mu}^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma} + \frac{1}{2}(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\tau_n &= \frac{(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
 \end{align}
@@
 
-\section{Free object torque}
+\subsection{Free object torque}
 %torque uniform free
 \begin{align}
-l &= \max{-1,\mu -\sqrt{3} \sigma}\\
-u &= \min{1,\mu+\sqrt{3}\sigma}\\
+l &= \max({-1,\mu -\sqrt{3} \sigma})\\
+u &= \min({1,\mu+\sqrt{3}\sigma})\\
 \tau_{uf} &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
@@
 \begin{align}
-l_1 &= \max{-1,\mu-\sqrt{6}\sigma}\\ \nonumber
-l_2 &= \max{-1,\mu}\\ \nonumber
-u_1 &= \min{1,\mu}\\ \nonumber
-u_2 &= \min{1,\mu+\sqrt{6}\sigma} \nonumber
+l_1 &= \max({-1,\mu-\sqrt{6}\sigma})\\ \nonumber
+l_2 &= \max({-1,\mu})\\ \nonumber
+u_1 &= \min({1,\mu})\\ \nonumber
+u_2 &= \min({1,\mu+\sqrt{6}\sigma}) \nonumber
 \end{align}
@@
 \end{align}
+
+Thus the torque would be:
+\begin{align}
+\tau_{umax} =\left\{
+\begin{array}{ll}
+\sqrt{3}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
+\frac{1}{4 \sqrt{3}\sigma} &   \sigma > \frac{1}{2\sqrt{3}}\\
+\end{array} 
+\right.
+\end{align}
 %best place to push for triangular:
@@
 %best place to push for normal distribution:
-For triangular and normal distributions, the best place to push if $\sigma = 0$ is 1.
+For all distributions, the optimal place to push if $\sigma = 0$ is 1.
 The best place to push moves in the $x$ direction when $\sigma$ gets bigger.
-We used L'hopital rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
-It was because the first derivative had undefined numerator and denominator and we were not able to calculate it.
+We used L'Hopital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
+It was because in the first derivative numerator and denominator were both going to zero and we were not able to calculate the limit.
 The result is:
@@
 %\end{figure}
+
+%taking derivative of torque
+\begin{align} 
+\frac{d\tau}{d\mu} &= -\frac{e^{-\frac{{-1+\mu}^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma} + \frac{1}{2}(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\end{align}

### 2016-10-19 JournalTorqueControl/theory.tex
Added torque max for triangular, edited the draft.

@@
 \begin{align} \nonumber
-\tau_n &= \frac{(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
+\tau_n &= \frac{(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}+ \frac{1}{2}\mu(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right))
 \end{align}
 
-%%uniform torque
-%\begin{align}
-%\mu = \textrm{Max}[1-\sqrt{3}\sigma,\sqrt{3}\sigma]
-%\end{align}
-
-\subsection{Free object torque}
-%torque uniform free
-\begin{align}
-l &= \max({-1,\mu -\sqrt{3} \sigma})\\
-u &= \min({1,\mu+\sqrt{3}\sigma})\\
-\tau_{uf} &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
-\end{align}
-%torque triangular free
-\begin{align}
-l_1 &= \max({-1,\mu-\sqrt{6}\sigma})\\ \nonumber
-l_2 &= \max({-1,\mu})\\ \nonumber
-u_1 &= \min({1,\mu})\\ \nonumber
-u_2 &= \min({1,\mu+\sqrt{6}\sigma}) \nonumber
-\end{align}
-\begin{align}
-\tau_{tf} &=  \left\{
-\begin{array}{ll}
-\frac{-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_1 > l_1\\
-\frac{2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)+{u_2}^2(-2u_1 + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_2 > l_2\\
-\end{array} 
-\right.
-\end{align}
-
-%torque normal free
-
-\begin{align} \nonumber
-\tau_{nf} &= \int_{-1}^1\frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}} x \, dx\\
- &= \frac{(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-(1+\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}\\
- &+ \frac{1}{2}\mu(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{1+\mu}{\sqrt{2}\sigma})) 
-\end{align}
-
-%limits and equations
-
-Consider robots are uniformly distributed. Finding the best place to guide the mean position:
+Consider robots are uniformly distributed. The mean position that maximizes torque is:
 % best place to push for uniform
@@
 \begin{array}{ll}
-\mu = 1-\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
+\mu = 1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
 \mu = \sqrt{3}\sigma &   \textrm{otherwise}\\
@@
 \begin{array}{ll}
-\sqrt{3}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
+1-\sqrt{3}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
 \frac{1}{4 \sqrt{3}\sigma} &   \sigma > \frac{1}{2\sqrt{3}}\\
@@
 %best place to push for triangular:
+For the swarm that has triangularly distributed, the mean position that maximizes torque is:
+
+\begin{align}
+\left\{
+\begin{array}{ll}
+\mu = -\sqrt{6} + \sqrt{12\sigma^2 +1} &   \textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
+\mu = \frac{\sqrt{2}}{2} &   \textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
+\end{array} 
+\right.
+\end{align}
+
+Thus the torque would be:
+\begin{align}
+\tau_{tmax} =\left\{
+\begin{array}{ll}
+\frac{-1 + \sqrt{1+12\sigma^2}^{\frac{3}{2}}- 18\sigma^3\sqrt{6}}{18\sigma^2} &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
+\frac{\sqrt{2}-2+3\sqrt{6}}{36\sigma^2} &   \sigma > \frac{1}{2\sqrt{3}}\\
+\end{array} 
+\right.
+\end{align}
 
 %best place to push for normal distribution:
-For all distributions, the optimal place to push if $\sigma = 0$ is 1.
-The best place to push moves in the $x$ direction when $\sigma$ gets bigger.
-We used L'Hopital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
+For all distributions, the $\mu$ that maximizes torque if $\sigma = 0$ is 1.
+For the $\tau_{pivot}$ case, the optimal $\mu$ moves in the $-x$ direction as $\sigma$ increases and approaches a limit.
+We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
 It was because in the first derivative numerator and denominator were both going to zero and we were not able to calculate the limit.
@@
 which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
+
+%%uniform torque
+%\begin{align}
+%\mu = \textrm{Max}[1-\sqrt{3}\sigma,\sqrt{3}\sigma]
+%\end{align}
+
+\subsection{Free object torque}
+
+For calculating torque in free object when the swarm has a uniform distribution, we again defined $l_f$ and $u_f$ for simplicity of the following derivations. Therefore, the torque applied is:
+%torque uniform free
+\begin{align}
+l_f &= \max({-1,\mu -\sqrt{3} \sigma}),\, u_f = \min({1,\mu+\sqrt{3}\sigma})\\
+\tau_{uf} &= -\frac{l_f^2}{4\sqrt{3}\sigma}+ \frac{u_f^2}{4\sqrt{3}\sigma} \textrm{  for    }  u_f>l_f
+\end{align}
+
+%torque triangular free
+To simplify the following derivations for triangular distribution we use:
+\begin{align}
+l_{f1} &= \max({-1,\mu-\sqrt{6}\sigma}), \,l_{f2} = \max({-1,\mu})\\ \nonumber
+u_{f1} &= \min({1,\mu}), \, u_{f2} = \min({1,\mu+\sqrt{6}\sigma}) \nonumber
+\end{align}
+\begin{align}
+\tau_{tf} &=  \left\{
+\begin{array}{ll}
+\frac{-2{l_{f1}}^3+3{l_{f1}}^2(\mu-\sqrt{6}\sigma)+{u_{f1}}^2(2u_{f1} - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_{f1} > l_{f1}\\
+\frac{2{l_{f2}}^3-3{l_{f2}}^2(\mu+\sqrt{6}\sigma)+{u_{f2}}^2(-2u_{f2} + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_{f2} > l_{f2}\\
+\end{array} 
+\right.
+\end{align}
+
+%torque normal free
+Also, torque when swarm is normally distributed is defined by:
+\begin{align} \nonumber
+\tau_{nf} &= \frac{\left(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-(1+\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\
+ &+ \frac{1}{2}\mu\left(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{1+\mu}{\sqrt{2}\sigma})\right) 
+\end{align}
+
+%limits and equations
 

### 2016-10-26 JournalTorqueControl/theory.tex
Added free object torque triangular

@@
 \section{Instantiating torque from swarm distribution}
+
 Consider a swarm of robots with probability density $p(x)$.
-This section examines where to steer the mean of the probability distribution to maximize torque.
-We examine two problems.
-First, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is:
+This section considers three canonical probability distributions: uniform, triangular, and normal, all parameterized by mean, $\mu$ and standard deviation, $\sigma$.
+They are described by:
+
+%uniform distribution
+\begin{align}
+p_u(x) &=  \left\{
+\begin{array}{ll}
+    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu+\sqrt{3}\sigma\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.\\
+%triangular distribution
+p_t(x) &=  \left\{
+\begin{array}{ll}
+    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
+     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
+     0, & \textrm{otherwise}\\
+\end{array} 
+\right.\\
+%normal distribution
+p_n(x) &= \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
+\kern-\nulldelimiterspace} {2\sigma ^2 }}}
+\end{align}
+This section examines where to steer the mean of the probability distribution to maximize torque. We discuss two problems: pivoted object torque and free object torque
+\subsection{Pivoted object torque}
+
+In this problem, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is:
 \begin{equation}
@@
 \end{equation}
-Second, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
-\begin{equation}
-\tau_{free} = \int_{-1}^1 x\,p(x)\, dx
-\end{equation}
-This section considers three canonical probability distributions: uniform, triangular, and normal, all parameterized by mean, $\mu$ and standard deviation, $\sigma$. They are described by:
 
@@
 \end{figure*}
-%uniform distribution
-\begin{align}
-p_u(x) &=  \left\{
-\begin{array}{ll}
-    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu+\sqrt{3}\sigma\\
-     0, & \textrm{otherwise}\\
-\end{array} 
-\right.\\
-%triangular distribution
-p_t(x) &=  \left\{
-\begin{array}{ll}
-    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
-     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
-     0, & \textrm{otherwise}\\
-\end{array} 
-\right.\\
-%normal distribution
-p_n(x) &= \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}}
-\end{align}
 
-\subsection{Pivoted object torque}
-For calculating torque in pivoted object when the swarm has a uniform distribution, we defined $l$ and $u$ for simplicity of the following derivations. Therefore, the torque applied is:
+First,  $l$ and $u$ for simplicity of the following derivations are defined for the uniform distribution. 
 %defining torque uniform distribution
 \begin{align}
-l &= \max(0,\mu -\sqrt{3} \sigma) \, , u = \min({1,\mu+\sqrt{3}\sigma})\\
-\tau_u &= -\frac{l^2}{4\sqrt{3}\sigma}+ \frac{u^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
+l &= \max(0,\mu -\sqrt{3} \sigma) \, , u = \min({1,\mu+\sqrt{3}\sigma})
 \end{align}
-To simplify the following derivations for triangular distribution we use:
+Therefore, the torque applied is:
+\begin{align}
+\tau_u &=  \frac{u^2-l^2}{4\sqrt{3}\sigma} \textrm{  for    }  u>l
+\end{align}
+To simplify derivations for triangular distribution we define bounds of integration as:
 %define torque for triangular:
@@
 l_1 &= \max({0,\mu-\sqrt{6}\sigma})\, , l_2 = \max({0,\mu})\\ \nonumber
-u_1 &= \min({1,\mu})\, , u_2 = \min({1,\mu+\sqrt{6}\sigma})\\ \nonumber
-\tau_t &= \frac{\tau_{left} + \tau_{right}}{36\sigma^2}\nonumber
+u_1 &= \min({1,\mu})\, , u_2 = \min({1,\mu+\sqrt{6}\sigma}) \nonumber
+\end{align}
+Thus, torque is:
+\begin{align}
+\tau_t &= \frac{\tau_{left} + \tau_{right}}{36\sigma^2}
 \end{align}
@@
 \begin{align}
-\tau_{left} &=  -2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma)\\ \nonumber
-&+{u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma), & \textrm{for     } u_1 > l_1\\ \nonumber
+\tau_{left} &=  {u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
+&-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma), & \textrm{for     } u_1 > l_1\\ \nonumber
 \tau_{right} &= 2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
-&+{u_2}^2(-2u_2 + 3\mu+3\sqrt{6}\sigma),  & \textrm{for     } u_2 > l_2\\ \nonumber
+&-{u_2}^2(2u_2 - 3\mu-3\sqrt{6}\sigma),  & \textrm{for     } u_2 > l_2\\ \nonumber
 \end{align}
-Also, torque when swarm is normally distributed is defined by:
+Torque by a uniformly distributed swarm is defined by:
 %defining torque normal distribution
 \begin{align} \nonumber
-\tau_n &= \frac{(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}})\sigma}{\sqrt{2\pi}}+ \frac{1}{2}\mu(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right))
+\tau_n &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right)
 \end{align}
 
-Consider robots are uniformly distributed. The mean position that maximizes torque is:
+Consider a uniformly distributed swarm. The mean position that maximizes torque is computed.
 % best place to push for uniform
 \begin{align}
-\left\{
+\mu=\left\{
 \begin{array}{ll}
-\mu = 1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
-\mu = \sqrt{3}\sigma &   \textrm{otherwise}\\
+1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
+\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
 \end{array} 
@@
 
-Thus the torque would be:
+The resulting torque is:
 \begin{align}
@@
 %best place to push for triangular:
-For the swarm that has triangularly distributed, the mean position that maximizes torque is:
+For a triangularly distributed swarm, the mean position that maximizes torque is:
 
 \begin{align}
-\left\{
+\mu= \left\{
 \begin{array}{ll}
-\mu = -\sqrt{6} + \sqrt{12\sigma^2 +1} &   \textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
-\mu = \frac{\sqrt{2}}{2} &   \textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
+ \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
+ \frac{\sqrt{2}}{2} &   \textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
 \end{array} 
@@
 
-Thus the torque would be:
-\begin{align}
+Thus the torque is computed by:
+\begin{align}\label{eq:triMaxTorque}
 \tau_{tmax} =\left\{
 \begin{array}{ll}
-\frac{-1 + \sqrt{1+12\sigma^2}^{\frac{3}{2}}- 18\sigma^3\sqrt{6}}{18\sigma^2} &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
+\frac{(1+12\sigma^2)^{\frac{3}{2}}- 18\sigma^3\sqrt{6}-1}{18\sigma^2} &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
 \frac{\sqrt{2}-2+3\sqrt{6}}{36\sigma^2} &   \sigma > \frac{1}{2\sqrt{3}}\\
@@
 For the $\tau_{pivot}$ case, the optimal $\mu$ moves in the $-x$ direction as $\sigma$ increases and approaches a limit.
+This limit is $\mu = 0$ for a uniformly distributed swarm and $\mu = \frac{\sqrt{2}}{2}$ for a triangularly distributed swarm as we can see from Eq.\ref{eq:triMaxTorque}.
 We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
-It was because in the first derivative numerator and denominator were both going to zero and we were not able to calculate the limit.
+It was because in the first derivative numerator and denominator were both going to infinity and we were not able to calculate the limit.
 The result is:
@@
 \begin{align}
-\lim_{\sigma\to\infty} \frac{d^3x}{d\mu}(\tau_n)= -1+\frac{3\mu}{2}
+\lim_{\sigma\to\infty} \frac{d^3\tau_n}{d\mu}= \frac{3\mu}{2}-1
 \end{align}
@@
 
-%%uniform torque
-%\begin{align}
-%\mu = \textrm{Max}[1-\sqrt{3}\sigma,\sqrt{3}\sigma]
-%\end{align}
-
 \subsection{Free object torque}
-
-For calculating torque in free object when the swarm has a uniform distribution, we again defined $l_f$ and $u_f$ for simplicity of the following derivations. Therefore, the torque applied is:
+In this problem, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
+\begin{equation}
+\tau_{free} = \int_{-1}^1 x\,p(x)\, dx
+\end{equation}
+For calculating free object torque for a uniformly distributed swarm, $l_f$ and $u_f$ are defined for simplicity of derivations. 
 %torque uniform free
 \begin{align}
-l_f &= \max({-1,\mu -\sqrt{3} \sigma}),\, u_f = \min({1,\mu+\sqrt{3}\sigma})\\
-\tau_{uf} &= -\frac{l_f^2}{4\sqrt{3}\sigma}+ \frac{u_f^2}{4\sqrt{3}\sigma} \textrm{  for    }  u_f>l_f
+l_f &= \max({-1,\mu -\sqrt{3} \sigma}),\, u_f = \min({1,\mu+\sqrt{3}\sigma})
+\end{align}
+
+Therefore, the torque applied is:
+\begin{align}
+\tau_{uf} &= \frac{u_f^2-l_f^2}{4\sqrt{3}\sigma} \textrm{  for    }  u_f>l_f
 \end{align}
@@
 %torque triangular free
-To simplify the following derivations for triangular distribution we use:
+To simplify derivations for a triangularly distributed swarm, we define bounds of integration as:
 \begin{align}
@@
 \end{align}
+Torque is:
 \begin{align}
-\tau_{tf} &=  \left\{
+\tau_{tf} = \frac{\tau_{fleft}+ \tau_{fright}}{36\sigma^2}
+\end{align}
+where $\tau_{fleft}$ and $\tau_{fright}$ is defined as
+\begin{align}\nonumber
+\tau_{fleft} &=  {u_{f1}}^2(2u_{f1} - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
+&-2{l_{f1}}^3+3{l_{f1}}^2(\mu-\sqrt{6}\sigma) &   \textrm{for     } u_{f1} > l_{f1}\\ \nonumber
+\tau_{fright} &= 2{l_{f2}}^3-3{l_{f2}}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
+&+{u_{f2}}^2( 3\mu-2u_{f2}+3\sqrt{6}\sigma) &   \textrm{for     } u_{f2} > l_{f2}\\ \nonumber
+\end{align}
+
+%torque normal free
+Also, torque of a normally distributed swarm is defined by:
+\begin{align} \nonumber
+\tau_{nf} &= \frac{\left(e^{\frac{-(1+\mu)^2}{2\sigma^2}}-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\
+ &+ \frac{1}{2}\mu\left(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{1+\mu}{\sqrt{2}\sigma})\right) 
+\end{align}
+
+Maximum torque of a uniformly distributed swarm for a free object has the same boundary as the pivoted object.
+The difference is because the pivoted object even when the swarm is pushing all the length of the object, they are applying torque, but for the free object, if the swarm pushes the whole length of the object, they would just apply force.
+
+For triangularly distributed swarm we have:
+
+\begin{align}
+\mu= \left\{
 \begin{array}{ll}
-\frac{-2{l_{f1}}^3+3{l_{f1}}^2(\mu-\sqrt{6}\sigma)+{u_{f1}}^2(2u_{f1} - 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_{f1} > l_{f1}\\
-\frac{2{l_{f2}}^3-3{l_{f2}}^2(\mu+\sqrt{6}\sigma)+{u_{f2}}^2(-2u_{f2} + 3\mu+3\sqrt{6}\sigma}{36\sigma^2}, &   \textrm{for     } u_{f2} > l_{f2}\\
+ \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     }  \sigma  <\frac{2}{\sqrt{6}} \\
+1 &   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}} 
 \end{array} 
@@
 \end{align}
-
-%torque normal free
-Also, torque when swarm is normally distributed is defined by:
-\begin{align} \nonumber
-\tau_{nf} &= \frac{\left(-e^{-\frac{(-1+\mu)^2}{2\sigma^2}}+e^{-\frac{-(1+\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\
- &+ \frac{1}{2}\mu\left(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{1+\mu}{\sqrt{2}\sigma})\right) 
-\end{align}
-
 %limits and equations
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{TorqueFreeBestPush.pdf}
+\begin{overpic}[width =\figwid]{MaxTorquePivot.pdf}%\put(1,55){a)}
 \end{overpic}
@@
 
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=0.8\columnwidth]{BestLocationPivot.pdf}
-%\end{center}
-%\vspace{-1em}
-%\caption{\label{fig:besLocPiv}
-%Comparison for three different distributions for the best location to push the object.
-%}
-%\vspace{0em}
-%\end{figure}
-
 \begin{figure*}
@@
 \renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{MaxTorquePivot.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{TorqueFreeBestPush.pdf}
 \end{overpic}
@@
 \end{figure*}
-
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=0.8\columnwidth]{MaxTorquePivot.pdf}
-%\end{center}
-%%\vspace{0em}
-%\caption{\label{fig:maxTorPiv}
-%Comparison for three different distributions for the maximum torque possible.
-%}
-%\vspace{0em}
-%\end{figure}
-%
-%%free comparisons
-%
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=0.8\columnwidth]{TorqueFreeComparison.pdf}
-%\end{center}
-%\vspace{-1em}
-%\caption{\label{fig:maxTorqueFree}
-%Comparison for three different distributions for the maximum torque possible.
-%}
-%\vspace{0em}
-%\end{figure}
-
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=0.8\columnwidth]{TorqueFreeBestPush.pdf}
-%\end{center}
-%%\vspace{0em}
-%\caption{\label{fig:bestLocFree}
-%Comparison for three different distributions for the best location to push the object.
-%}
-%\vspace{0em}
-%\end{figure}
-
-%taking derivative of torque
-\begin{align} 
-\frac{d\tau}{d\mu} &= -\frac{e^{-\frac{{-1+\mu}^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma} + \frac{1}{2}(\erf(\frac{1-\mu}{\sqrt{2}\sigma})+\erf(\frac{\mu}{\sqrt{2}\sigma})) 
-\end{align}

### 2016-10-26 JournalTorqueControl/theory.tex
changed captions

@@
 
+%pivoted comparisons:
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{BestLocationPivot.pdf}%\put(1,55){a)}
+\end{overpic}
+\begin{overpic}[width =\figwid]{MaxTorquePivot.pdf}%\put(1,55){a)}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:bestLoc} Best location to push and maximum torque plot in pivoted object.
+%\vspace{-2em}
+}
+\end{figure*}
+
 \subsection{Free object torque}
@@
 
-%pivoted comparisons:
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{BestLocationPivot.pdf}%\put(1,55){a)}
-\end{overpic}
-\begin{overpic}[width =\figwid]{MaxTorquePivot.pdf}%\put(1,55){a)}
-\end{overpic}
-\vspace{-0.5em}
-\caption{\label{fig:bestLoc} Best location to push in two different situations: when the object is pivoted, and when the object is free.
-%\vspace{-2em}
-}
-\end{figure*}
-
+%free torque comparison
 \begin{figure*}
@@
 \vspace{-0.5em}
-\caption{\label{fig:maxTorque} Maximum torque possible in two situations: pivoted object and free object.
+\caption{\label{fig:maxTorque} Best location to push and maximum torque plot in free object.
 %\vspace{-2em}

### 2016-11-04 JournalTorqueControl/theory.tex
Added the limit

@@
 %defining torque normal distribution
-\begin{align} \nonumber
+\begin{align} \label{eq:normalTorque}
 \tau_n &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right)
@@
 This limit is $\mu = 0$ for a uniformly distributed swarm and $\mu = \frac{\sqrt{2}}{2}$ for a triangularly distributed swarm as we can see from Eq.\ref{eq:triMaxTorque}.
-We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
+%We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
 It was because in the first derivative numerator and denominator were both going to infinity and we were not able to calculate the limit.
 The result is:
-
+The first derivative of Eq.\ref{eq:normalTorque} is:
 \begin{align}
-\lim_{\sigma\to\infty} \frac{d^3\tau_n}{d\mu}= \frac{3\mu}{2}-1
+\frac{d_{\tau}}{d_{\mu}}= \frac{1}{2}\left(\erf{\frac{1-\mu}{\sqrt{2}\sigma}}+ \erf{\frac{\mu}{\sqrt{2}\sigma}}\right)- \frac{e^{\frac{-(\mu-1)^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma}
 \end{align}
 
-which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
+The limit as $\sigma\to\infty$ is $\frac{2 - 3 \mu}{\sigma^3 6 \sqrt{2 \pi}}$ which goes to zero as the swarm variance increases.
+However, the zero of the numerator is $\mu = \frac{2}{3}$ which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
+
+%\begin{align}
+%\lim_{\sigma\to\infty} \frac{d^3\tau_n}{d\mu}= \frac{3\mu}{2}-1
+%\end{align}
+
+%which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
 

### 2016-11-18 JournalTorqueControl/simulation.tex
Corrected some mistakes regarding reviews.

\section{Simulation}\label{sec:simulation}
This section examines four main challenges for pose and torque control of an object, arranged in increasing difficulty.  Each task uses a PD controller that regulates the swarm's mean position,  as in \cite{ShahrokhiIROS2015}. The control input is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(goal_x - \bar{x}) + K_{d}(0-\bar{v}_x) \nonumber\\
u_y &= K_{p}(goal_y  - \bar{y}) + K_{d}(0-\bar{v}_y)  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.  
The swarm's average position is $[\bar{x},\bar{y}]^T$ and mean velocity is $[\bar{v}_x,\bar{v}_y]^T$.  
Each task uses a different algorithm to select the swarm's goal position $[goal_x,goal_y]^T$. The derivative gain $K_d$ limits overshoot.

\paragraph{Pure torque control} \label{sec:simulationPureTorque}
An object with a pivot point can rotate, but not translate. A door is an example.
 If there was only one robot touching the object, the robot should push at the point which maximizes the moment arm, at the extreme end of the object furthest from the pivot point.
The optimal pushing location provides the maximum force, because it maximizes  $r$ in \eqref{eq:torque}.
However, given a swarm of robots, maximizing $r$ is no longer the optimal solution.  
If the mean of the swarm hits the object at the extreme edge, half of the robots will miss the object and  the swarm will be split.
Because few robots remain,  the force is significantly decreased and torque is not maximized.
 In our simulation, the swarm applies torque until the swarm's mean position is beyond the object. Then the swarm will regather in a corner before returning to torque control, a time-consuming task. 
 The key parameter of interest for a hinged door of length $L$ is $C$, the position along the door where the mean of the swarm will push.  The goal position in ~\eqref{eq:TorqueControl} is set to:

\begin{align}\nonumber
goal_x &= O_x + C \cos(O_{\theta}) \\
goal_y &= O_y + C \sin(O_{\theta}),  \label{eq:TorqueControl}
\end{align}
$O_{\theta}$ = the orientation of the object's major axis, measured from the world $x$-axis.
 Fig. \ref{fig:LFig} illustrates how different values of $C$ result in different rates of turning. These simulations tested $C = \{1/2, 3/4, 5/8, 1\}L$. The fastest turning rates occurred with  $C =  5/8L$. Code is available at \cite{Shahrokhi16torque}.

\begin{figure}
\begin{center}
	\includegraphics[width=0.8\columnwidth]{LFig.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:LFig}
Simulation results from a swarm applying force to a hinged door. 
The swarm mean is steered toward a point $C$ units along the object from the pivot point. The red dashed line indicates the times that the swarm was in variance control mode.
 Simulation used 144 robots of diameter $0.2$ m with a standard deviation of less than $1.5$ m and an object length of $6$ m.
}
\vspace{-1em}
\end{figure}

\paragraph{Orientation of the object}
These simulations used a uniform density rectangle as the object. This object was 30$\times$ larger than the robots.
Using the pure torque control discussed in the previous paragraph, the orientation of the object can be controlled by applying force. 
The rectangular object is not pivoted, so it moves in addition to rotating. 
 The swarm still may split into multiple components.
  We use the hysteresis variance control from \cite{ShahrokhiIROS2015}  to gather the swarm when its variance grows too large. 
  The following control law chooses a goal position to regulate the orientation of the object.

 \begin{align}\nonumber
goal_x = O_x +  K_{orient}  (O_{\theta} - goal_\theta ) \cos(O_{\theta}) \\
goal_y = O_y +  K_{orient}  ( O_{\theta} -goal_\theta  ) \sin(O_{\theta})
\end{align}
Here $K_{orient}$ is a positive gain on the control input.

\begin{figure}
\begin{center}
	\includegraphics[width=0.8\columnwidth]{Orientation.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:OrientCont}
Plot demonstrating  orientation control of a rectangular object. The green line is the goal orientation. Other lines show different random starting position of the swarm. When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace until the variance is below a desired threshold. 
}
\vspace{-1em}
\end{figure}

Fig. \ref{fig:OrientCont} illustrates this controller with different starting positions. When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace. Code is available at \cite{Shahrokhi16Orient}.

### 2016-11-18 RA-L/intro.tex
severe under actuation
Corrected some mistakes regarding reviews.

@@
-\section{Introduction}\label{sec:Intro}
+Sec. ection{Introduction}\label{sec:Intro}
 Particle swarms propelled by a global force are common in applied mathematics, biology, and computer graphics. 
@@
 \end{align}
-Our control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
+Here $x_i$ and $y_i$ regard to the position and respectively $\dot{x}_i, \dot{y}_i$ are velocities of the particles.
+Our control problem is to design the control inputs $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
 As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
-(1) to assemble structures in parallel and 
- (2) through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
+to assemble structures in parallel and 
+ through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
  For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
 Instead, these robots are often controlled by a global, broadcast signal.
-These applications require techniques to shape the swarm that can reliably exploit large populations despite high under-actuation.
+They require techniques to shape the swarm that can reliably exploit large populations despite severe underactuation ($2 \ll2n$).
 
@@
 \caption{\label{fig:IntroPic}
-Swarm of kilobots programmed to move toward the brightest light source as explained in \S \ref{sec:expResults}.
+Swarm of kilobots programmed to move toward the brightest light source as explained in Sec.
+\ref{sec:expResults}.
 The current covariance ellipse and mean are shown in red, the desired covariance is shown in green.
@@
 The paper is arranged as follows.
-First \S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
+First Sec.
+\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces general models for boundary interaction.
-We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in \S \ref{sec:PostionControl2Robots}, and \S \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
-Both algorithms require robots with low process noise and require time proportional to the number of robots, so \S \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm covariance.
-\S \ref{sec:simulation} describes implementations of the algorithms in simulation and  \S \ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-After a review of recent related work in \S \ref{sec:RelatedWork}, we end with directions for future research in \S \ref{sec:conclusion}.
-%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
+To extend the range of possible shapes, Sec.
+\ref{subsec:WallFriction} introduces general models for boundary interaction.
+We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in Sec.
+\ref{sec:PostionControl2Robots}, and Sec.
+\ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
+Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.
+\ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm covariance.
+Sec.
+\ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.
+\ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
+After a review of recent related work in Sec.
+\ref{sec:RelatedWork}, we end with directions for future research in Sec.
+\ref{sec:conclusion}.
+%For controlling $Sec. igma_{xy}$, we prove that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction. 
 %We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
-%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
+%We conclude by designing controllers that efficiently regulate $Sec. igma_{xy}$.
 %This paper
-%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
+%(1) proves that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
 %(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
@@
 %(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
-%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
-%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
+%TODO JOURNAL: design controllers that efficiently regulate $Sec. igma_{xy}$.
+%TODO JOURNAL: We will design Lyapunov-inspired controllers for $Sec. igma_{xy}$ to prove controllability. 
 %TODO JOURNAL:  and rank controllability as a function of friction.

### 2016-11-18 RA-L/theory.tex
added worst case
adde
Corrected some mistakes regarding reviews.

@@
 
-\subsection{Using Boundaries: Fluid Settling In a Tank}\label{subsec:FluidInTank}
+\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
 One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
@@
 We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$.
+The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1].
 Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
@@
 
-Computing means, variances, covariance, and correlation requires integrating over the region $R$ containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+Computing means, variances, covariance, and correlation requires integrating over the region $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
@@
 %\begin{align}
-\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}_x\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
+\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
 \text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
@@
 
-The region of integration $R$ is the polygon containing the swarm. If $A<1/2$ and the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
+The region of integration $R$ is the polygon containing the swarm. If $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
 \begin{align}\label{eq:meanInSquareWorkspaceLL}
-\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} x \, dy\right) \, dx}{A} \nonumber \\
+\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} x \, dy\right) \, dx}{A} \nonumber \\
 	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
-\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} y \, dy\right) \, dx}{A} \nonumber\\
+\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
 	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
@@
 Similarly,  two iterations %of Alg. \ref{alg:YControl} 
-are only required if $|\Delta e_y - \Delta s_y|>L$.
+are only required if $|\Delta e_y - \Delta s_y|>L$.  The worst case path length is 6$L$.
 

### 2016-11-18 RA-L/theory.tex
Added Changes

@@
 One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
+Like fluid settling in a tank, the stable final configuration minimizes potential energy.
 \paragraph{Square workplace}
@@
 
-The region of integration $R$ is the polygon containing the swarm. If $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
+The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
 \begin{align}\label{eq:meanInSquareWorkspaceLL}
-\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} x \, dy\right) \, dx}{A} \nonumber \\
+\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
 	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
@@
 $N$ is the normal or perpendicular force between the robot and the boundary.
-The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $|\theta| < \pi/2$.
+The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$.
 The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
@@
 \end{overpic}}\\
-%{\begin{overpic}[width =\figwid]{two_1.jpg}\put(45,75){$t$  = 0 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{two_2.jpg}\put(45,75){$t$  = 0.25 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{two_3.jpg}\put(45,75){$t$  = 0.5 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{two_4.jpg}\put(45,75){$t$  = 0.75 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{two_5.jpg}\put(45,75){$t$  = 1 s}
-%\end{overpic}}\\
-%\vspace{.5em}
-%{\begin{overpic}[width =\figwid]{S3_1.jpg}\put(45,75){$t$ = 0 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S3_2.jpg}\put(45,75){$t$ = 0.25 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S3_3.jpg}\put(45,75){$t$  = 0.5 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S3_4.jpg}\put(45,75){$t$  = 0.75 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S3_5.jpg}\put(45,75){$t$  = 1 s}
-%\end{overpic}}
-%\vspace{-1em}
+
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
@@
 
-%Step (i): Fixing $\Delta r_x$
-%\label{theory:step1}
-%\begin{enumerate}
-%\item
-%If $\Delta e_x$ is negative, the robots will be commanded to move toward the left wall and halt $\epsilon$ distance from the left wall.
-%If $\Delta e_x \ge 0$, the robots will be commanded to move toward the right wall and halt $\epsilon$ distance from the right wall.
-The epsilon distance prevents the robots from experiencing friction along the vertical wall.
-%\item  let $y_{\min} = \min_i r_{iy}$, i.e., $y_{\min}$ be the minimum height of the two robots.
-Move both robots downward  $y_{\min}$ such that one of the robots touches the bottom wall and hence friction force will prevent this robot from moving left or right.
-%\item friction force holds the lower robot in place while the upper robot may move right and left freely to change $\Delta r_x $ to $\Delta e_x$.
-%\item  If after the free move of the upper robot $\Delta r_x$ is not $\Delta e_x$, Step (i) will be repeated.  No more than two iterations of step (i) are required.
-%\end{enumerate}
-%
-%Step (ii): Fixing $\Delta r_y$
-%Now that $\Delta r_x$ is corrected, $\Delta r_y$ must be corrected:
-%\begin{enumerate}
-%\item 
-%If $\Delta e_y$ is negative, the robots will be commanded to move toward the top wall and halt $\epsilon$ distance from the top wall.
-%If $\Delta e_y \ge 0$, the robots will be commanded to move toward the bottom wall and halt $\epsilon$ distance from the bottom wall.
-The epsilon distance prevents the robots from experiencing friction along the horizontal wall.
-%\item  let $x_{\min} = \min_i r_{ix}$, i.e., $x_{\min}$ be the minimum distance of the two robots from the origin along the $x$-axis.
-Move both robots to the left $x_{\min}$ such that one of the robots  touches the left wall and hence friction force will prevent this robot from moving up or down.
-%\item  friction force holds the left robot in place while the right robot may move up or down freely to change $\Delta r_y $  to $\Delta e_y$.
-%\item  If after the free move of the right robot $\Delta r_y$ is not $\Delta e_y$, Step (ii) will be repeated.  No more than two iterations of step (ii) are required.
-%\end{enumerate}
-%Once $\Delta r_x$ and $\Delta r_y$ are set to $\Delta e_x$ and $\Delta e_y$, a straight-line global input moves both robots from $r_1$ and $r_2$ to $e_1$ and $e_2$. 
-%
-
 \begin{algorithm}
@@
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
-\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
+\State flag $\gets 0$
+ \If {$ r_{1y}  \equiv r_{2y}$} 
+\State perturb $x$ values
+\State flag $\gets 1$
+\EndIf
+%\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
 \State $\epsilon \gets $ small number
-\State $ \Delta s_x  \gets s_{1x} - s_{2x} $
+\State $ \Delta s_x  \gets r_{1x} - r_{2x} $
 \State $ \Delta e_x \gets e_{1x} - e_{2x} $
-%\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
 \If {$\Delta e_x < 0 $ }
-\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{To right wall}
+\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Near right wall}
 \Else 
-\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
+\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Near left wall}
 \EndIf
 \State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
 \If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
@@
 \State move$(0, \epsilon)$ \Comment{Move up}
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-%\State $\Delta r_x = r_{1x} - r_{2x}$
-\If {$r_{1x} - r_{2x} \equiv \Delta e_x$} 
-%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
+\If {$r_{1x} - r_{2x} \not\equiv \Delta e_x$} 
+\State GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
+\EndIf
+
+ \If {$ flag == 1$} 
+\State unperturb $x$ values
+\EndIf
+
 \State  \Return $(r_1,r_2)$
-\Else   
-\State \Return GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
-\EndIf
 \end{algorithmic}
 \end{algorithm}
-
-%\begin{algorithm}
-%\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
-%\begin{algorithmic}[1]
-%\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-%$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
-% $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
-%\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
-%\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
-%\State $ \Delta e_y \gets e_{1y} - e_{2y} $
-%\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
-%\If {$\Delta e_y < 0 $ }
-%\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
-%\Else 
-%\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
-%\EndIf
-%\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-%\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
-%\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
-%\Else
-%\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
-%\EndIf 
-%\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-%\State $\Delta r_y = r_{1y} - r_{2y}$
-%\If {$\Delta r_y \equiv \Delta e_y$} 
-%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
-%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
-%\State  \Return $(r_1,r_2)$
-%\Else   
-%\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
-%\EndIf
-%\end{algorithmic}
-%\end{algorithm}
 

### 2016-11-20 RA-L/intro.tex
Added Changes
Reworked intro and summary.

@@
-Sec. ection{Introduction}\label{sec:Intro}
+\section{Introduction}\label{sec:Intro}
 Particle swarms propelled by a global force are common in applied mathematics, biology, and computer graphics. 
+If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
 \begin{align}
-[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
+[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
 \end{align}
-Here $x_i$ and $y_i$ regard to the position and respectively $\dot{x}_i, \dot{y}_i$ are velocities of the particles.
-Our control problem is to design the control inputs $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
+These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
+ In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
+  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
+ Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
 As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
@@
 The paper is arranged as follows. 
-First Sec.
-\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.
+After a review of recent related work in Sec.  \ref{sec:RelatedWork},
+ Sec.
+\ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
@@
 \ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
-After a review of recent related work in Sec.
-\ref{sec:RelatedWork}, we end with directions for future research in Sec.
+We end with directions for future research in Sec.
 \ref{sec:conclusion}.

### 2016-12-03 RA-L/theory.tex
Added the algorithm

@@
 
+%\begin{algorithm}
+%\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
+%\begin{algorithmic}[1]
+%\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
+%$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
+% $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
+%\State flag $\gets 0$
+% \If {$ r_{1y}  \equiv r_{2y}$} 
+%\State perturb $x$ values
+%\State flag $\gets 1$
+%\EndIf
+%%\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
+%\State $\epsilon \gets $ small number
+%\State $ \Delta s_x  \gets r_{1x} - r_{2x} $
+%\State $ \Delta e_x \gets e_{1x} - e_{2x} $
+%\If {$\Delta e_x < 0 $ }
+%\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Near right wall}
+%\Else 
+%\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Near left wall}
+%\EndIf
+%\State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
+%\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
+%\State move$(\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
+%\Else
+%\State move$(-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
+%\EndIf 
+%\State move$(0, \epsilon)$ \Comment{Move up}
+%\If {$r_{1x} - r_{2x} \not\equiv \Delta e_x$} 
+%\State GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
+%\EndIf
+%
+% \If {$ flag == 1$} 
+%\State unperturb $x$ values
+%\EndIf
+%
+%\State  \Return $(r_1,r_2)$
+%\end{algorithmic}
+%\end{algorithm}
 \begin{algorithm}
-\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
+\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
 \begin{algorithmic}[1]
 \Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
+$(0,0)$ is bottom corner, $s_1$ is topmost, leftmost robot, 
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
-\State flag $\gets 0$
- \If {$ r_{1y}  \equiv r_{2y}$} 
-\State perturb $x$ values
-\State flag $\gets 1$
+\State $\Delta e_y = e_2.y - e_1.y$
+\State $\Delta e_x = e_2.x - e_1.x$
+\If {$\Delta e_y \neq 0$}
+\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
+\Else
+\State $sign \gets 0$
 \EndIf
-%\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
-\State $\epsilon \gets $ small number
-\State $ \Delta s_x  \gets r_{1x} - r_{2x} $
-\State $ \Delta e_x \gets e_{1x} - e_{2x} $
-\If {$\Delta e_x < 0 $ }
-\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Near right wall}
-\Else 
-\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Near left wall}
+\State Move to $(-r_1.x , -\min(R_2.y , R_1.y+ \Delta e_y)+\epsilon)$
+\State Move to $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$
+\If {$r_2.y \neq r_1.y$}
+\State Move to $(\max(\epsilon, \Delta e_x- (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$
+\State Move to $\Delta e_x , (|sign|-1) \epsilon$
 \EndIf
-\State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
-\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
-\State move$(\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
-\Else
-\State move$(-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
-\EndIf 
-\State move$(0, \epsilon)$ \Comment{Move up}
-\If {$r_{1x} - r_{2x} \not\equiv \Delta e_x$} 
-\State GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
-\EndIf
-
- \If {$ flag == 1$} 
-\State unperturb $x$ values
-\EndIf
-
-\State  \Return $(r_1,r_2)$
+\State Move to $(e_2.x - r_2.x, e_2.y - r_2.y)$
 \end{algorithmic}
@@
 \subsection{Controlling Covariance Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
-Assume an open workspace with infinite boundary friction.
+Assume an obstacle-free, bounded workspace with infinite boundary friction.
 Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$.

### 2016-12-04 RA-L/theory.tex
some tiny typos
Updated Algorithm

@@
 \Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is topmost, leftmost robot, 
+$(0,0)$ is bottom corner, $s_1$ is top-leftmost robot, 
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
+\Ensure $(e_1, e_2) , (s_1, s_2)$ all have $\epsilon$ distance with the walls.
+\State $epsilon \gets$ diameter of the robot
 \State $\Delta e_y = e_2.y - e_1.y$
@@
 \EndIf
-\State Move to $(-r_1.x , -\min(R_2.y , R_1.y+ \Delta e_y)+\epsilon)$
-\State Move to $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$
-\If {$r_2.y \neq r_1.y$}
-\State Move to $(\max(\epsilon, \Delta e_x- (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$
-\State Move to $\Delta e_x , (|sign|-1) \epsilon$
+\State Move by $(-r_1.x , -\min(r_2.y , r_1.y+ \Delta e_y)+\epsilon)$ \Comment{Touch left wall}
+\State Move by $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
+%\If {$r_2.y \neq r_1.y$}
+\State Move by $(\max(\epsilon, \Delta e_x- (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
+\State Move by $(\Delta e_x , (|sign|-1) \epsilon)$ \Comment{Adjust $x$}
+%\EndIf
+\If {$\Delta e_y \neq r_2.y - r_1.y$}
+\State Move by $(L- r_1.x , \epsilon)$ \Comment{Touch Right wall}
+\State Move by $(0, \Delta e_y - r_2.y)$ \Comment{Adjust $y$}
 \EndIf
-\State Move to $(e_2.x - r_2.x, e_2.y - r_2.y)$
+\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to the goal position}
 \end{algorithmic}

### 2016-12-05 RA-L/theory.tex
Algorithm completed.
Added a comment to the algorithm
Another typo
last typo for tonight.
small edits to typography and grammar for also

--- 2016-12-03 RA-L/theory.tex
@@
 \begin{algorithmic}[1]
+\scriptsize
 \Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is topmost, leftmost robot, 
+$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
+\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
+\Ensure $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
+\State $\epsilon \gets$ robot diameter
 \State $\Delta e_y = e_2.y - e_1.y$
@@
 \EndIf
-\State Move to $(-r_1.x , -\min(R_2.y , R_1.y+ \Delta e_y)+\epsilon)$
-\State Move to $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$
+\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
+\State Move $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
 \If {$r_2.y \neq r_1.y$}
-\State Move to $(\max(\epsilon, \Delta e_x- (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$
-\State Move to $\Delta e_x , (|sign|-1) \epsilon$
+\State Move $(\max(\epsilon, -\Delta e_x+ (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
+\State Move $(\Delta e_x , (|sign|-1) \epsilon)$ \Comment{Adjust $x$}
 \EndIf
-\State Move to $(e_2.x - r_2.x, e_2.y - r_2.y)$
+\If {$\Delta e_y \neq r_2.y - r_1.y$}
+\State Move $(L- r_1.x , \epsilon)$ \Comment{Touch right wall}
+\State Move $(0, \Delta e_y - r_2.y)$ \Comment{Adjust $y$}
+\EndIf
+\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
 \end{algorithmic}

### 2016-12-06 RA-L/intro.tex
Small changes and added Algorithm description

@@
  Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
-As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
+As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided

### 2016-12-06 RA-L/simulation.tex
Added Changes
Small changes and added Algorithm description

@@
 
-Algorithms \ref{alg:PosControl2Robots} and \ref{alg:XControl}, %\ref{alg:YControl}, 
-were implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
+Algorithms \ref{alg:PosControl2Robots} %\ref{alg:YControl}, 
+was implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
 Robot initial positions are shown by crosshairs, and final positions by circled crosshairs.
@@
 
-\subsection{Efficient Control of Covariance}
+\subsection{Efficient Control of Correlation}
 %A set of simulations were conducted to demonstrate the importance of boundary friction.  
@@
 \ref{alg:PosControlNRobots}.
-However, due to the central limit theorem, we can control the covariance of large swarms.
-This section demonstrates controlling swarm covariance in simulation, using the 2D physics engine Box2D~\citep{catto2010box2d}.
+However, due to the central limit theorem, we can control the correlation of large numbers of particles.
+This section demonstrates controlling swarm correlation in simulation, using the 2D physics engine Box2D~\citep{catto2010box2d}.
  144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow} were controlled by an open-loop control input as illustrated in Fig.
@@
 All robots had  the same initial conditions, but in four tests the boundary friction was $F_f = \{0,1/3 F, 2/3F, F\}$.
- Without friction, covariance has minimal variation.  As friction increases, the covariance can be manipulated to greater degrees.
+ Without friction, covariance has minimal variation.  As friction increases, the correlation can be manipulated to greater degrees.
 
@@
 
- 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in \S \ref{subsec:ClosedLoopCovarianceControl} with $c_1=0.1$.
+ 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in Sec.
+\ref{subsec:ClosedLoopCovarianceControl} with $c_1=0.1$.
 Fig.

### 2016-12-06 RA-L/theory.tex
Small changes and added Algorithm description

@@
 %u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \left( \frac{y-h}{h} \right)^2  , &  y<h \label{eq:boundarylayerflow} \\
+F_{forward}(y) &= F - F_f\begin{cases}  \frac{y-h}{h}   , &  y<h \label{eq:boundarylayerflow} \\
 0, & \text{else} \end{cases}
@@
 This algorithm  introduces concepts that will be used for multi-robot positioning.
-It only requires collisions with two orthogonal walls, in this case, the bottom and left walls.
 Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
@@
 Denote the current positions of the robots  $r_1$ and $r_2$. 
-Subscripts $_x$ and $_y$ denote the $x$ and $y$ coordinates, i.e., $s_{1x}$ and $s_{1y}$ denote the $x$ and $y$ locations of $s_1$. 
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Also the sign function is:
+\begin{align}
+\sign (x) = \begin{cases}  1, &  x > 0 \\
+0, & x=0\\
+-1, & x<0 \end{cases} 
+\end{align}
 The algorithm assigns a global control input at every instance.
 The goal is to adjust 
- $\Delta r_x = r_{2x}-r_{1x}$ from $\Delta s_x = s_{2x}-s_{1x}$ to $\Delta e_x = e_{2x}-e_{1x}$ and  adjust 
- $\Delta r_y = r_{2y}-r_{1y}$ from $\Delta s_y = s_{2y}-s_{1y}$ to $\Delta e_y = e_{2y}-e_{1y}$ using a shared global control input. 
+ $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
+ $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
  This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
@@
 
-Our algorithm solves the positioning problem in three steps: 
-First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Third, the robots, now correctly positioned relative to each other, are moved to their goal locations.
+Our algorithm solves the positioning problem in four steps: 
+First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
+Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
+Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
+%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
+%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
+Last, the robots, now correctly positioned relative to each other, are moved to their goal locations.
 
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
-Similarly,  two iterations %of Alg. \ref{alg:YControl} 
-are only required if $|\Delta e_y - \Delta s_y|>L$.  The worst case path length is 6$L$.
+In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all the steps. The worst case path length is $2(\sqrt{2}+1)L$.
 
 \begin{algorithm}
-\caption{WallFrictionArrange2Robots($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
-\begin{algorithmic}[1]
-\Require starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
- $L$ is length of the walls. 
- Current robot positions are $(r_1,r_2)$.
-
-\State  GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
-\State  GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
-\State Move robots $e_1 - r_1$
-
-\end{algorithmic}
-\end{algorithm}
-
-%\begin{algorithm}
-%\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
-%\begin{algorithmic}[1]
-%\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-%$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
-% $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
-%\State flag $\gets 0$
-% \If {$ r_{1y}  \equiv r_{2y}$} 
-%\State perturb $x$ values
-%\State flag $\gets 1$
-%\EndIf
-%%\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
-%\State $\epsilon \gets $ small number
-%\State $ \Delta s_x  \gets r_{1x} - r_{2x} $
-%\State $ \Delta e_x \gets e_{1x} - e_{2x} $
-%\If {$\Delta e_x < 0 $ }
-%\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Near right wall}
-%\Else 
-%\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Near left wall}
-%\EndIf
-%\State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
-%\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
-%\State move$(\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
-%\Else
-%\State move$(-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
-%\EndIf 
-%\State move$(0, \epsilon)$ \Comment{Move up}
-%\If {$r_{1x} - r_{2x} \not\equiv \Delta e_x$} 
-%\State GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
-%\EndIf
-%
-% \If {$ flag == 1$} 
-%\State unperturb $x$ values
-%\EndIf
-%
-%\State  \Return $(r_1,r_2)$
-%\end{algorithmic}
-%\end{algorithm}
-\begin{algorithm}
-\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
+\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
 \begin{algorithmic}[1]
@@
 \State $\Delta e_x = e_2.x - e_1.x$
-\If {$\Delta e_y \neq 0$}
-\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
-\Else
-\State $sign \gets 0$
-\EndIf
+%\If {$\Delta e_y \neq 0$}
+%\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
+%\Else
+%\State $sign \gets 0$
+%\EndIf
 \State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-\State Move $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), sign \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
+\State Move $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), \sign(\Delta e_y) \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
 \If {$r_2.y \neq r_1.y$}
 \State Move $(\max(\epsilon, -\Delta e_x+ (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
-\State Move $(\Delta e_x , (|sign|-1) \epsilon)$ \Comment{Adjust $x$}
+\State Move $(\Delta e_x , (|\sign(\Delta e_y)|-1) \epsilon)$ \Comment{Adjust $x$}
 \EndIf

### 2016-12-06 RA-L/theoryCorrelation.tex
Added plots and text on correlation

\subsection{Maximizing Correlation Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
Using the environment to individually move $n$ particles to goal positions in Alg. \ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg. \ref{alg:PosControl2Robots} can only control two particles. The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited. This section presents an efficient technique to generate desired correlations.

Assume an obstacle-free, bounded, unit-size square workspace. 
 As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
 This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1. By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
\begin{align} \label{eq:GravityCorrelation}
\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
 \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
 , & \frac{1}{2} \le A\le 1
\end{cases}
\end{align}
 If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate much larger correlations.
Specifically, we can generate correlations larger than 1/2 by using boundary friction if the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick.

 Assume that the swarm is initialized in the lower-left corner, in a rectangle of width $w$ and height $A/w$. 
 Such a rectangular configuration can be accomplished using the variance controllers from \citep{ShahrokhiIROS2015}. 
  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
   The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top. These regions can be defined by the rectangle's left side, bottom, and top:
\begin{align}
r_{\text{left}} &= \min (L,1-w)\nonumber \\
r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right)\nonumber \\
r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right)
\end{align}
If $\frac{A}{w} \le r_{\text{top}}$ the top parallelogram has no area. 
 Similarly, if $r_{\text{top}} \le r_{\text{bottom}}$ the rectangle has no area. 
The mean, variance, and correlation are calculated  using \eqref{eq:meanInSquareWorkspace}, \eqref{eq:varInSquareWorkspace}, and \eqref{eq:covAndcorrInSquareWorkspace} over the region $R$:
\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{h}y}^{\frac{L}{h}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
&+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy \nonumber
\end{align}

 Given an environment parameterized by $A$ and $h$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
  Negative correlations can be generated by initializing the swarm in the upper left, or lower right.

%freedom is not the freedom to do what we want, but freedom to do what we ought.

% PLACE IN SIMULATION

\subsection{Efficient Control of Correlation}
%A set of simulations were conducted to demonstrate the importance of boundary friction.  
This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.  As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
Larger boundary layers $h$ enable more control of correlation.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:ControlledCorrelation}
Analytical results comparing maximum correlation under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
}
\end{figure}

### 2016-12-06 RA-L/theoryCorrelation.tex
slight updates to plot

@@
 Larger boundary layers $h$ enable more control of correlation.
+The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\citep{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
 
@@
 \caption{\label{fig:ControlledCorrelation}
-Analytical results comparing maximum correlation under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+Analytical results comparing maximum correlation $\rho_{xy}$ under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
 }

### 2016-12-07 RA-L/simulation.tex
Merged stuff together and added reference.

@@
 An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
-
-%\textcolor{red}{Shiva:  what is the complexity of this algorithm?  How many steps in the worst case?  When does worst case happen?}
-%\begin{figure*}
-%\centering
-%\renewcommand{\figwid}{0.4\columnwidth}
-%{\begin{overpic}[width =\figwid]{one_1.jpg}\put(45,75){$t$ = 0 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{one_2.jpg}\put(45,75){$t$ = 0.25 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{one_3.jpg}\put(45,75){$t$  = 0.5 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{one_4.jpg}\put(45,75){$t$  = 0.75 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{one_5.jpg}\put(45,75){$t$  = 1 s}
-%\end{overpic}}
-%\vspace{-1em}
-%\caption{\label{fig:shapeControlMathematica2}{Two robot positioning: switching positions using walls with infinite friction.}
-%%\vspace{-2em}
-%}
-%\end{figure*}
 
@@
 %A set of simulations were conducted to demonstrate the importance of boundary friction.  
-Random disturbances impair the performance of Alg.
-\ref{alg:PosControl2Robots} and Alg.
-\ref{alg:PosControlNRobots}.
-However, due to the central limit theorem, we can control the correlation of large numbers of particles.
-This section demonstrates controlling swarm correlation in simulation, using the 2D physics engine Box2D~\citep{catto2010box2d}.
- 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow} were controlled by an open-loop control input as illustrated in Fig.
-\ref{fig:SimCovarianceFuncFrictionOpenLoop}.
-All robots had  the same initial conditions, but in four tests the boundary friction was $F_f = \{0,1/3 F, 2/3F, F\}$.
- Without friction, covariance has minimal variation.  As friction increases, the correlation can be manipulated to greater degrees.
+This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
+from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
+The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
+For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
+As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
+Larger boundary layers $h$ enable more control of correlation.
+The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\citep{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
 
@@
 \begin{center}
-	\includegraphics[width=\columnwidth]{SimCovarianceFuncFrictionOpenLoopV2.pdf}
+	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
 \end{center}
 \vspace{-1em}
-\caption{\label{fig:SimCovarianceFuncFrictionOpenLoop}
-Open-loop simulation with 144 disc robots and varying levels of boundary friction under the same initial conditions.
-Without friction, covariance is unchangeable.
-As friction increases, the covariance can be manipulated to greater degrees.
+\caption{\label{fig:ControlledCorrelation}
+Analytical results comparing maximum correlation $\rho_{xy}$ under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
 }
 \end{figure}
-
- 144 disc-shaped robots were also controlled by a closed-loop controller using the procedure in Sec.
-\ref{subsec:ClosedLoopCovarianceControl} with $c_1=0.1$.
-Fig.
-\ref{fig:CovVarControlPlot} illustrates that covariance and variances in $x$ and $y$ directions were controlled from a set of initial conditions.
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{CovVarControlPlot2.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:CovVarControlPlot}
-Closed-loop simulation with 144 disc robots and three sets of initial conditions.  The algorithm tracks goal variance and covariance values (green).  The goal covariance switches sign every 30 s.}
-\end{figure}

### 2016-12-07 RA-L/theory.tex
Merged stuff together and added reference.

@@
 Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer.
+Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \citep{fluidMechanics}.
 \begin{align}
 %u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \frac{y-h}{h}   , &  y<h \label{eq:boundarylayerflow} \\
+F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
 0, & \text{else} \end{cases}
@@
 
-\subsection{Controlling Covariance Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
-Assume an obstacle-free, bounded workspace with infinite boundary friction.
-Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$.
-%$\sigma_x^2$ decreases when the swarm is pushed into vertical walls, and $\sigma_y^2$ decreases when the swarm is pushed into horizontal walls. Alg.~\ref{alg:CovCont} procedes as follows:
-\begin{enumerate}
-\item swarm pushed into the left wall until $\sigma_x^2 \le c_1\sigma_{goalx}^2$.  
-\item swarm's mean position is moved to the center of the workspace
-\item swarm pushed into the bottom wall until $\sigma_y^2 \le  \sigma_{goaly}^2$. 
-\item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\
-else swarm slides left until $\sigma_{xy} \le \sigma_{goalxy}$ 
-\item  swarm's mean position is moved to the center of the workspace
-\end{enumerate}
-The constant $c_1$ in step 1 should be less than 1 to ensure $\sigma_x^2$ stays within tolerance during the subsequent moves.
+\subsection{Maximizing Correlation Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
+Using the environment to individually move $n$ particles to goal positions in Alg.
+\ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg.
+\ref{alg:PosControl2Robots} can only control two particles.
+The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited.
+This section presents an efficient technique to generate desired correlations.
 
-%\begin{algorithm}
-%\caption{Swarm Covariance Control With Wall Friction}\label{alg:CovCont}
-%\begin{algorithmic}[1]
-%\Require  goal variances and covariance and mean, variances and covariance of the swarm  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy}, \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$. 
-%%Swarm is Gaussian distributed (in limit it is uniformly distributed), 
-%$L_w$ is width of the walls and $L_h$ is the height of the wall. $(0,0)$ is top right corner. $\epsilon$ is the error for reaching covariance.
-%\If {$\sigma_x^2 > \sigma_{goalx}^2/10 $ }
-%\State $ \bar{x} \gets \sigma_{goalx}/10   $ \Comment{Move to left wall}
-%\State $MoveToCenter \gets true$
-%\Else
-%\If {$MoveToCenter$ }
-%\State   $ \bar{x} \gets L_w/2 $, $\bar{y} \gets L_h/2 $ \Comment{Move to the center}
-%\State $MoveToBottom \gets true$
-%\State $MoveToCenter \gets false$
-%\Else 
-%\If {$MoveToBottom $}
-%\State $ \bar{y} \gets L_h - \sigma_{goaly}   $ \Comment{Move to bottom}
-%\State $MoveToBottom \gets false$
-%\Else
-%\If {$\sigma_{xy} -\sigma_{goalxy} > \epsilon$}
-%\While  {$\sigma_{xy} -\sigma_{goalxy} > \epsilon$ }
-%\State $ \bar{x} \gets \bar{x} + \epsilon  $ \Comment{Move to right wall}
-%\EndWhile
-%\Else
-%\If {$\sigma_{xy} -\sigma_{goalxy} < \epsilon$ }
-%\While  {$\sigma_{xy} -\sigma_{goalxy} < \epsilon$ }
-%\State $ \bar{x} \gets \bar{x} - \epsilon  $ \Comment{Move to left wall}
-%\EndWhile
-%\EndIf
-%\EndIf
-%\EndIf
-%\EndIf
-%\EndIf
-%\State   $ \bar{x} \gets L_w/2 $, $\bar{y} \gets L_h/2 $ \Comment{Move to the center}
-%
-%\end{algorithmic}
-%\end{algorithm}
+Assume an obstacle-free, bounded, unit-size square workspace. 
+ As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
+ This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1.
+By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
+\begin{align} \label{eq:GravityCorrelation}
+\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
+% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
+ \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
+ , & \frac{1}{2} \le A\le 1
+\end{cases}
+\end{align}
+ If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate much larger correlations.
+Specifically, we can generate correlations larger than 1/2 by using boundary friction if the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick.
+
+ Assume that the swarm is initialized in the lower-left corner, in a rectangle of width $w$ and height $A/w$. 
+ Such a rectangular configuration can be accomplished using the variance controllers from \citep{ShahrokhiIROS2015}. 
+  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
+   The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top.
+These regions can be defined by the rectangle's left side, bottom, and top:
+\begin{align}
+r_{\text{left}} &= \min (L,1-w)\nonumber \\
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right)\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right)
+\end{align}
+If $\frac{A}{w} \le r_{\text{top}}$ the top parallelogram has no area. 
+ Similarly, if $r_{\text{top}} \le r_{\text{bottom}}$ the rectangle has no area. 
+The mean, variance, and correlation are calculated  using \eqref{eq:meanInSquareWorkspace}, \eqref{eq:varInSquareWorkspace}, and \eqref{eq:covAndcorrInSquareWorkspace} over the region $R$:
+\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{h}y}^{\frac{L}{h}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
+&+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
+&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy \nonumber
+\end{align}
+
+ Given an environment parameterized by $A$ and $h$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
+  Negative correlations can be generated by initializing the swarm in the upper left, or lower right.

### 2016-12-07 RA-L/intro.tex
Small changes and the newest version to review

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a global force are common in applied mathematics, biology, and computer graphics. 
+Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
 If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then

### 2016-12-08 RA-L/theory.tex
Newest version of the algorithm(I found a bug in it.)
workspace, not workplace

@@
 Like fluid settling in a tank, the stable final configuration minimizes potential energy.
-\paragraph{Square workplace}
+\paragraph{Square workspace}
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
-We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a very large swarm of robots as they move inside a square workplace under the influence of gravity pointing in the direction $\beta$.
+We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1].
-Under a global input such as gravity, they flow like water, moving to a side of the workplace and forming a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
+Under a global input, they move to a side of the workspace and form a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Circular workplace}
+\paragraph{Circular workspace}
 Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
@@
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
+ Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
 \Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\Ensure $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
 \State $\epsilon \gets$ robot diameter
@@
 %\EndIf
+%First Move
 \State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-\State Move $(-\min(|-r_2.x+ \Delta e_x|, r_2.x), \sign(\Delta e_y) \, \min(\max(r_1.y + \Delta e_y, r_2.y)-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
-\If {$r_2.y \neq r_1.y$}
-\State Move $(\max(\epsilon, -\Delta e_x+ (r_2.x - r_1.x)), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
-\State Move $(\Delta e_x , (|\sign(\Delta e_y)|-1) \epsilon)$ \Comment{Adjust $x$}
+%Second Move
+%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
+\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
+%Third Move
+\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
+%Fourth Move
+\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
+\If {$\Delta e_y \neq r_2.y - r_1.y$}
+%Fifth Move
+%TODO This step is not correct.
+\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
+%Sixth Move
+\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ \Comment{Adjust $y$}
 \EndIf
-\If {$\Delta e_y \neq r_2.y - r_1.y$}
-\State Move $(L- r_1.x , \epsilon)$ \Comment{Touch right wall}
-\State Move $(0, \Delta e_y - r_2.y)$ \Comment{Adjust $y$}
-\EndIf
+%Seventh Move
 \State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}

### 2016-12-09 RA-L/experiment.tex
Small changes and the newest version to review
grammar check, force to 8 pages, and revised review doc

@@
 \caption{\label{fig:storyReal}Position control of two kilobots (Alg.
-\ref{alg:XControl}) steered to corresponding colored circle.
+\ref{alg:PosControl2Robots}) steered to corresponding colored circle.
 Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
@@
 All robots receive the same control inputs, but boundary interactions break the control symmetry.
-Robots reach their respective goal position in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.
+Robots reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.
 
@@
 Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
-\vspace{-1em}
+\vspace{-2.5 em}
 }

### 2016-12-09 RA-L/intro.tex
grammar check, force to 8 pages, and revised review doc

@@
  through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
- For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
+ For each task, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
 Instead, these robots are often controlled by a global, broadcast signal.
@@
 Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.
-\ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm covariance.
+\ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm correlation.
 Sec.

### 2016-12-09 RA-L/simulation.tex
grammar check, force to 8 pages, and revised review doc

@@
 
-Algorithms \ref{alg:PosControl2Robots} %\ref{alg:YControl}, 
-was implemented in Mathematica using point robots (radius = $0$).  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm. 
-Robot initial positions are shown by crosshairs, and final positions by circled crosshairs.
-Dashed lines show the shortest route if robots could be controlled independently.
-The path given by  Alg.\ \ref{alg:PosControl2Robots} is shown with solid arrows.
+Algorithm \ref{alg:PosControl2Robots}  was implemented in Mathematica using point robots (radius = $0$). 
+An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
+  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm wiht robot initial positions shown by hollow squares and final positions by circles. 
+ Dashed lines show the shortest route if robots could be controlled independently, solid lines the generated path.
+
+   %The path given by our algorithm is shown with solid arrows.
 %Each row has five snapshots taken every quarter second. 
-For the sake of brevity axis-aligned moves were replaced with oblique moves that combine two moves simultaneously.
+%For the sake of brevity axis-aligned moves were replaced with oblique moves that combine 
+%jjkjk klklk klklklk klklk lklkl lklklklkl  klklklklk lklklklklk lklklkl klklklkl klklklk klklklk klklk klk klkl
 % THES IMAGES DON'T DO THIS ANYMORE 
@@
 % Finally, the algorithm moves the robots to their corresponding destinations.
-An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
 

### 2016-12-09 RA-L/theory.tex
grammar check, force to 8 pages, and revised review doc

@@
 We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$ ranging from [0,1].
-Under a global input, they move to a side of the workspace and form a polygonal shape, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
+The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
+Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
 
@@
 
-Computing means, variances, covariance, and correlation requires integrating over the region $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
@@
 A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
-Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
-An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
-Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
-Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.
+If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
+%Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
+%An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
+%Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
+Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
@@
 
-The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls can arbitrarily position two robots. 
+The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
 \begin{figure}[h]
@@
 \section{Algorithms}\label{sec:algorithms}
-\subsection{Position Control of $2$ Robots Using Wall Friction}\label{sec:PostionControl2Robots}
+\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
 \begin{figure*}
@@
 %Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Last, the robots, now correctly positioned relative to each other, are moved to their goal locations.
+Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
 
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all the steps. The worst case path length is $2(\sqrt{2}+1)L$.
+In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.
 
 \begin{algorithm}
-\caption{GenerateDesired spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
+\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
 \begin{algorithmic}[1]
@@
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$ if not, rotate workspace coordinates $90^{\circ}$.
+ Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
 \Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
@@
 \State $\Delta e_x = e_2.x - e_1.x$
-%\If {$\Delta e_y \neq 0$}
-%\State $sign \gets \frac{|\Delta e_y|}{\Delta e_y}$
-%\Else
-%\State $sign \gets 0$
-%\EndIf
-%First Move
 \State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
@@
 %(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
-\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$ \Comment{Adjust  $y$}
+\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
+\Statex \Comment{Adjust  $y$}
 %Third Move
-\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ \Comment{Touch bottom wall}
+\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
+\Statex \Comment{Touch bottom wall}
 %Fourth Move
@@
 %Sixth Move
-\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ \Comment{Adjust $y$}
+\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
+\Statex \Comment{Adjust $y$}
 \EndIf
 %Seventh Move
-\State Move by $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
+\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
 \end{algorithmic}
@@
 
-\State move($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State move($f_{kx}-r, 0$)
+\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
+\State Move ($f_{kx}-r, 0$)
 
@@
 
-Assume an obstacle-free, bounded, unit-size square workspace. 
+Assume an obstacle-free, bounded, unit-size, square workspace. 
  As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
@@
 \end{align}
- If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate much larger correlations.
-Specifically, we can generate correlations larger than 1/2 by using boundary friction if the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick.
+ If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
+ If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
 

### 2016-12-09 RA-L/theory.tex
added description of Shiva's code

@@
 \end{center}
+\vspace{-1em}
 \end{figure*}
@@
 \label{fig:friction}
-}
+}\vspace{-1em}
 \end{center}
@@
 The algorithm only requires friction along the bottom and right walls.
-Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
+Robot initial positions are shown by a square, and final positions by a circle.
 Dashed lines show the shortest route if robots could be controlled independently.
@@
 }
-%\vspace{-2em}
+\vspace{-1em}
 }
@@
 Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
+We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.
+If $s_1$ is not also the topmost robot, we rotate the coordinate frame by 90$^\circ$.
 Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Also the sign function is:
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Define the sign function as:
 \begin{align}
@@
 Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
+\vspace{-1em}
 \end{figure}

### 2017-02-08 IROS17/experiment.tex
Added IROS first draft!

--- 2016-12-09 RA-L/experiment.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}[!htb]
-\centering
-\renewcommand{\figwid}{0.38\columnwidth}
-{
-\begin{overpic}[width =\figwid]{twoR1.pdf}%\put(12,70){$t$  = 0 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{twoR2.pdf}\put(12,70){$t$  = 60 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{twoR3.pdf}\put(12,70){$t$  = 150 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{twoR4.pdf}\put(12,70){$t$  = 160 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{twoR5.pdf}\put(12,70){$t$  = 210 s}
-\end{overpic}}
-%\vspace{-3em}
-\caption{\label{fig:storyReal}Position control of two kilobots (Alg.
-\ref{alg:PosControl2Robots}) steered to corresponding colored circle.
-Boundary walls have nearly infinite friction, so the green robot is stopped by the wall from $t = 60$s until the commanded input is directed away form the wall at $t=150$s, while the pink robot in free-space is unhindered.}
-\end{figure*}
-
-%\subsection{Hardware System}
-
-Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.
-These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot, from \citep{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available as an open source platform or commercially from~\citep{K-Team2015}.
-Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
-Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.
-In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1 m$\times$1 m whiteboard as the workspace, four 30W and four 50W LED floodlights arranged 1.5 m above the plane of the table at the midpoints and vertices of a 6 m square centered on the workspace.
-An Arduino Uno connected to an 8-relay shield controlled the lights.
-Above  the table, an overhead machine vision system tracked the position of the swarm.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=.9\columnwidth]{SetUp.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:setup}
-Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight  triggered LED floodlights, with an overhead machine vision system.
-}
-\vspace{-1em}
-\end{figure}
-\subsection{Hardware Experiment: Position Control of Two Robots}
-The hardware platform walls have almost infinite friction, due to a laser-cut, scalloped border.
-When a kilobot is steered into the scalloped border, they are pinned to the wall unless the global input directs them away from the wall.
-Even though the kilobots are stochastic, this wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.
 
@@
 Robots reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left robot from the staging zone occupying the top-right position of the build zone.
-
-\subsection{Hardware Experiment: Control of Covariance}
-To demonstrate covariance control, up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
-The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=\columnwidth]{newExperiment.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:covExperiment}
-Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance.
-The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.
-The actual covariance is shown in different trials.
-Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
-\vspace{-2.5 em}
-}
-\end{figure}

### 2017-02-08 IROS17/intro.tex
Added IROS first draft!

--- 2016-12-09 RA-L/intro.tex
@@
 Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
-If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
-\begin{align}
-[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
-\end{align}
-These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
- In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
-  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
- Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
+%If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
+%\begin{align}
+%[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
+%\end{align}
+%These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
+% In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
+%  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
+% Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
 As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
@@
 This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.
-See multimedia attachment for overview of the experiments.}
+See multimedia attachment for overview of the experiments.} \vspace{-1em}
 \end{figure}
@@
 Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze.
-Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
+%Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
 By adding rectangular boundary walls, some higher-order moments, such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls), are also controllable~\citep{ShahrokhiIROS2015}.
 %A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.
-These techniques could only compress a swarm orthogonal to obstacles.
+%These techniques could only compress a swarm orthogonal to obstacles.
 Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
@@
 After a review of recent related work in Sec.  \ref{sec:RelatedWork},
- Sec.
+ %Sec.
 \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes, Sec.
+To extend the range of possible shapes,
+  Sec.
 \ref{subsec:WallFriction} introduces general models for boundary interaction.
-We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in Sec.
+We introduce algorithms with two orthogonal boundaries with high friction sufficient to arbitrarily position two robots in Sec.
 \ref{sec:PostionControl2Robots}, and Sec.
 \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
-Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.
+%Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.
 \ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm correlation.
@@
 \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.
-\ref{sec:expResults} describes hardware experiments with up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}.
+\ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}.
 We end with directions for future research in Sec.

### 2017-02-08 IROS17/simulation.tex
Added IROS first draft!

--- 2016-12-09 RA-L/simulation.tex
@@
 Control performance is sensitive to the desired clearance $\epsilon$.  As $\epsilon$ increases, the total distance decreases asymptotically.
-}
+}\vspace{-1em}
 \end{figure}
-
-\subsection{Efficient Control of Correlation}
-%A set of simulations were conducted to demonstrate the importance of boundary friction.  
-This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
-from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
-The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
-For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
-As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
-Larger boundary layers $h$ enable more control of correlation.
-The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\citep{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:ControlledCorrelation}
-Analytical results comparing maximum correlation $\rho_{xy}$ under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
-}
-\end{figure}

### 2017-02-08 IROS17/theory.tex
Added IROS first draft!

--- 2016-12-09 RA-L/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=\linewidth]{SquareFill.pdf} 
-\vspace{-1em}
-\caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$.
-Left plot shows locus of possible mean positions for five values of $A$.
-The locus morphs from a square to a circle as $A$ increases.
-The covariance ellipse for each $A$ is shown with a dashed line.
-Center shows two corresponding arrangements of kilobots.
-At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
-See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
-\label{fig:SquareFill}
-\end{center}
-\end{figure*} 
-\begin{figure*}[!htb]
-\begin{center}
-\vspace{-1em}
-\includegraphics[width=\linewidth]{CircleFill.pdf} 
-\vspace{-1em}
-\caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$.
-Left plot shows locus of possible mean positions for four values of $h$.
-The locus of possible mean positions are concentric circles.
-See online interactive demonstration at \citep{Zhao2016mathematica}.}
-\label{fig:CircleFill}
-\end{center}
-\vspace{-1em}
-\end{figure*}
-
-\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
-One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
-Like fluid settling in a tank, the stable final configuration minimizes potential energy.
-\paragraph{Square workspace}
-%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
-We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
-Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
-
-The range for the global input angle $\beta $ is [0,2$\pi $). In this range, the swarm assumes eight different polygonal shapes. 
-These shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and between squares with one corner removed and trapezoids when $A$$>$1/2.
-
-Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
-%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
-
-\begin{align}
-\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \\
-%\end{align}
-%\begin{align}
-\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\\
-%\end{align}
-%\begin{align}
-\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
-\end{align}
-
-The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
-\begin{align}\label{eq:meanInSquareWorkspaceLL}
-\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
-	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
-\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
-	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
-\end{align}
-The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
-A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
-The covariance of a triangle is always $\pm(A/18)$.
-Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape.
-The range of mean positions are maximized when $A$ is small.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Circular workspace}
-Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
-Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
-For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
-Under a global input, the robot swarm fills the region under a chord with area
-\begin{align}
-A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
-\end{align}
-For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
-\begin{align}
-r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
-\end{align}
-Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.
-
-%Smashing the swarm to the corner walls will cause some correlation between \emph{x} and \emph{y} axes.
-Eq.
-~\ref{eq:Wall} shows the relation between number of robots and area provided and the correlation we can reach with it.
-%%TODO: ADD the equation for showing the correlation Vs area, vs number of robots
-% This equation shows that we are not able to reach high correlations bigger than $|0.5|$.
-Fig.
-~\ref{fig:wallCorrelation} shows the different mean positions and correlations we are able to reach.
-There are ways and corridors that the swarm needs to pass and also to keep the majority of it.
-If we use the wall corners technique and the way needs a very high correlation to pass, a remarkable number of the swarm members will be lost, and we may lose track of them due to their small size.
-Meanwhile, the variance of the swarm would get bigger and bigger that it would not anymore be able to deliver sufficient payloads.
-For this reason we have thought another idea which uses \emph{friction} to make correlations.
-%
 
@@
 \centering
-\renewcommand{\figwid}{0.4\columnwidth}
-{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
+\renewcommand{\figwid}{0.32\columnwidth}
+{\begin{overpic}[width =\figwid]{S1.pdf}\put(65,80){Start}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
+\begin{overpic}[width =\figwid]{S2.pdf}\put(50,80){Move 1}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
+\begin{overpic}[width =\figwid]{S3.pdf}\put(50,80){Move 2}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
+\begin{overpic}[width =\figwid]{S4.pdf}\put(50,80){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
+\begin{overpic}[width =\figwid]{S5.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
 \end{overpic}}\\
@@
  Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
+$\epsilon $ is a small, nonzero, user-specified value.
+
 \Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\epsilon \gets$ robot diameter
 \State $\Delta e_y = e_2.y - e_1.y$
@@
 %Fifth Move
-%TODO This step is not correct.
-\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
+\If {$\Delta e_x \neq 0$}
+\State Move $(\sign(r_2.y- r_1.y)\Delta e_x, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
+\Else
+\State Move  $(|\sign(\Delta e_y -r_2.y+ r_1.y)| \epsilon, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
+\EndIf
 %Sixth Move
-\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
+\If {$\Delta e_x < 0$}
+\State Move $(L - \max(r_1.x, r_2.x), \epsilon)$
+\Else 
+\State Move $(L - \max(r_1.x, r_2.x),\Delta e_y -\max(r_1.y, r_2.y)$ 
 \Statex \Comment{Adjust $y$}
+\EndIf
 \EndIf
@@
 Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.
-
-\subsection{Maximizing Correlation Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
-Using the environment to individually move $n$ particles to goal positions in Alg.
-\ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg.
-\ref{alg:PosControl2Robots} can only control two particles.
-The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited.
-This section presents an efficient technique to generate desired correlations.
-
-Assume an obstacle-free, bounded, unit-size, square workspace. 
- As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
- This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1.
-By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
-\begin{align} \label{eq:GravityCorrelation}
-\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
-% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
- \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
- , & \frac{1}{2} \le A\le 1
-\end{cases}
-\end{align}
- If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
- If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
-
- Assume that the swarm is initialized in the lower-left corner, in a rectangle of width $w$ and height $A/w$. 
- Such a rectangular configuration can be accomplished using the variance controllers from \citep{ShahrokhiIROS2015}. 
-  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
-   The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top.
-These regions can be defined by the rectangle's left side, bottom, and top:
-\begin{align}
-r_{\text{left}} &= \min (L,1-w)\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right)\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right)
-\end{align}
-If $\frac{A}{w} \le r_{\text{top}}$ the top parallelogram has no area. 
- Similarly, if $r_{\text{top}} \le r_{\text{bottom}}$ the rectangle has no area. 
-The mean, variance, and correlation are calculated  using \eqref{eq:meanInSquareWorkspace}, \eqref{eq:varInSquareWorkspace}, and \eqref{eq:covAndcorrInSquareWorkspace} over the region $R$:
-\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{h}y}^{\frac{L}{h}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
-&+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
-&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy \nonumber
-\end{align}
-
- Given an environment parameterized by $A$ and $h$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
-  Negative correlations can be generated by initializing the swarm in the upper left, or lower right.

### 2017-02-17 IROS17/theory.tex
Comments, and some edits to the draft
some new figures and the code.

@@
 See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{PositionNrobotsComp.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:simulationNrobot}
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
+}
+\end{figure}
 
@@
 \begin{algorithmic}[1]
-\State move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
+\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
 
 \While{ $p_{kx} > r$} 
-\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
+\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon,left)$    %drift move left until kth robot touches left wall
 \EndWhile
@@
 \For{ $m$ iterations}
-\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
+\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon,up)$    %move kth robot to f_{ky} and leave the rest in position.
 \EndFor
 
+\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
+\State Move ($f_{kx}-r, 0$)
+
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{ {\sc DriftMove}($\alpha,\beta,\epsilon,direction$)}\label{alg:DriftMove}
+\begin{algorithmic}[1]
+\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
 \State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position

### 2017-02-20 IROS17/intro.tex
Added a lot of images to the paper.

@@
 
-\begin{figure}
-\centering
-\begin{overpic}[width=0.95\columnwidth]{IntroSwarmPic.pdf}\end{overpic}
-\caption{\label{fig:IntroPic}
-Swarm of kilobots programmed to move toward the brightest light source as explained in Sec.
-\ref{sec:expResults}.
-The current covariance ellipse and mean are shown in red, the desired covariance is shown in green.
-Navigating a swarm using global inputs is challenging because each member receives the same control inputs.
-This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.
-See multimedia attachment for overview of the experiments.} \vspace{-1em}
-\end{figure}
-
 %One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}.
@@
 Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
+
+\begin{figure}
+\centering
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegion.pdf}\end{overpic}
+\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\caption{\label{fig:IntroPic}
+All the robots get shared control input.
+Here in the left picture the wall has infinite friction.
+Therefore, when the first robot is touching the wall, the second robot can move to anywhere in the shaded area.
+We introduce a dynamic algorithm to find the shortest path the robots can move to get to their goal positions.
+Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots.
+In the right picture, it is shown that the shortest path with three moves is the straight line between the starting position of the robot and the reflection of the goal.
+} \vspace{-1em}
+\end{figure}
 

### 2017-02-20 IROS17/theory.tex
Added a lot of images to the paper.

@@
 
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.64\columnwidth}
+{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{RegionMove2.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{RegionMove3.pdf}
+\end{overpic}
+}\\
+
+\caption{\label{fig:regionMove}{Two rectangles representing the reachable regions for the current start and goal positions.
+The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
+The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
+}
+\vspace{-1em}
+}
+\end{figure*}
+
 \section{Algorithms}\label{sec:algorithms}
@@
 \renewcommand{\figwid}{0.32\columnwidth}
-{\begin{overpic}[width =\figwid]{S1.pdf}\put(65,80){Start}
+{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(65,80){Start}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S2.pdf}\put(50,80){Move 1}
+\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(50,80){Move 1}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S3.pdf}\put(50,80){Move 2}
+\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(50,80){Move 2}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S4.pdf}\put(50,80){Move 3}
+\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(50,80){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S5.pdf}\put(50,80){Move 4}
+%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
+%\end{overpic}
+}\\
+{\begin{overpic}[width =\figwid]{story5move0.pdf}\put(10,80){Start}
 \end{overpic}
-\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
-\end{overpic}}\\
+\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,80){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,80){Move 5}
+\end{overpic}
+}\\
+{\begin{overpic}[width =\figwid]{story7move0.pdf}\put(65,80){Start}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move3.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move4.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
 
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. 
-The algorithm only requires friction along the bottom and right walls.
-Robot initial positions are shown by a square, and final positions by a circle.
+Robot start positions are shown by a square, and goal positions by a circle.
 Dashed lines show the shortest route if robots could be controlled independently.
@@
 
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
+Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
 We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.
-If $s_1$ is not also the topmost robot, we rotate the coordinate frame by 90$^\circ$.
+If $s_1$ is not also the topmost robot, we rotate the coordinate frame by 90$^\circ$ counterclockwise.
 Denote the current positions of the robots  $r_1$ and $r_2$. 
@@
  $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
- This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
+ %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
  %employing the assumption we have made earlier about the walls' friction.
@@
 \begin{algorithm}
-\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
+\caption{PoseControl2Robots($s_1,s_2,g_1,g_2,L$)}\label{alg:PosControl2Robots}
 \begin{algorithmic}[1]
 \scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
  $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
+ Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates counterclockwise $90^{\circ}$.
 $\epsilon $ is a small, nonzero, user-specified value.
 
-\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\Delta e_y = e_2.y - e_1.y$
-\State $\Delta e_x = e_2.x - e_1.x$
+\Ensure $(g_1, g_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
+\State $\Delta g_y = e_2.y - e_1.y$
+\State $\Delta g_x = e_2.x - e_1.x$
 \State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}

### 2017-02-23 IROS17/theory.tex
Changed drift move images as well
Updated the paper part
Added Arun's stuff

@@
 
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:contour}
+The contour plot keeps 3 points steady.
+}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Number.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:numberPlot}
+Number of moves needed.
+}
+\end{figure}
+
+The setup in ??
+is an assembled practical setup that assumes that $\epsilon=2r = 1$ cm.
+It is a $7\times 7cm$ grid space.
+The wall friction is emulated by the jagged teeth design to keep rollers from moving out of place while implementing the drift move.
+The rollers are magnetic and can be controlled by magnetic steering.
+See video ``refer the video which shows moving rollers"  to see the algorithm at work.
+In the demonstration, the algorithm is a fixed step algorithm and all drift moves are an counterclockwise square move  of size $1cm\times1cm$.
+Once the kth roller gets to it?s designated location in every loop, a correction step is implemented.
+This makes the overall number of moves required for the fixed step algorithm more than that required for the Algorithm ?.
+??
+shows how there are only 6 stages per roller involved in Algorithm ??.
+The fixed step algorithm would go through 8 stages per roller as shown in ???.
+The fixed step algorithm can be found online.
+One significant difference between Algorithm ??
+and the fixed move implementation of it is that, we can achieve higher accuracy in placing rollers in their precise location if Algorithm 2 is used.
+
 Assume an open workspace with four axis-aligned walls with infinite friction.
@@
 \begin{center}
-	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
+	\includegraphics[width=.3\columnwidth]{driftmove.pdf}
+	\includegraphics[width=.3\columnwidth]{driftmove2.pdf}
+	\includegraphics[width=.3\columnwidth]{driftmove3.pdf}
 \end{center}

### 2017-02-24 IROS17/intro.tex
Updated the paper part
All the codes and pictures added(some pictures to be added)

--- 2017-02-08 IROS17/intro.tex
@@
 % Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
-As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
+As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}-\citep{qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
@@
 Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
+\begin{figure}
+\centering
+\vspace{1.5em}
+\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){n=2}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){Optimal shortest path}\end{overpic}\\
+\vspace{1em}
+\begin{overpic}[width=0.43\columnwidth]{monalizaBegin.pdf}\put(30, 100){n=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
+\begin{overpic}[width=0.38\columnwidth]{monalisa.pdf}\put(-3, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
+\caption{\label{fig:IntroPic}
+Dynamic algorithm to find the shortest path for two robots is designed.
+The two top pictures shows the initial position and goal position of the robots and the optimal shortest path by making use of wall friction when shared control input is applied.
+The two bottom pictures shows the n robot positioning using shared control input and boundary walls.
+} \vspace{-1em}
+\end{figure}
 

### 2017-02-24 IROS17/theory.tex
All the codes and pictures added(some pictures to be added)

@@
 
-\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
+\emph{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
 Global inputs move a swarm uniformly.  
@@
 A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
-If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
+%If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
 %Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
@@
 %Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
-Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
- These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
-Robots touching a wall have a friction force that opposes movement along the boundary.  
-This causes robots along the boundary to move more slowly than robots in free-space.
+%Instead of pushing our robots directly into a wall, 
+The following sections examine using boundaries that stop the particles if they are pushed into the wall.
+ These forces are  sufficient to break the symmetry caused by uniform inputs.  
+%Robots touching a wall have a friction force that opposes movement along the boundary.  
+%This causes robots along the boundary to move more slowly than robots in free-space.
 
-Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary.
+\begin{figure}
+\centering
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+\caption{\label{fig:TwoRegions}
+All the robots get shared control input.
+Here in the left picture the wall has infinite friction.
+Therefore, when the first robot is touching the wall, the second robot can move to anywhere in the shaded area.
+We introduce a dynamic algorithm to find the shortest path the robots can move to get to their goal positions.
+Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots.
+In the right picture, it is shown that the shortest path with three moves is the straight line between the starting position of the robot and the reflection of the goal.
+} \vspace{-1em}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\caption{\label{fig:TwoRegions}
+All the robots get shared control input.
+Here in the left picture the wall has infinite friction.
+Therefore, when the first robot is touching the wall, the second robot can move to anywhere in the shaded area.
+We introduce a dynamic algorithm to find the shortest path the robots can move to get to their goal positions.
+Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots.
+In the right picture, it is shown that the shortest path with three moves is the straight line between the starting position of the robot and the reflection of the goal.
+} \vspace{-1em}
+\end{figure}
+
+%Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary.
 $N$ is the normal or perpendicular force between the robot and the boundary.
@@
 The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
-\begin{align}
-F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
-\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
-F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
-\text{and } N &= F \cos(\theta) \nonumber
-\end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
-Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
-This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
-There are many alternate models of friction that also break control symmetry.
+%\begin{align}
+%F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
+%\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
+%F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
+%\text{and } N &= F \cos(\theta) \nonumber
+%\end{align}
+% Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
+%Though each receives the same inputs,  they experience different net forces.
+%  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
+%This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
+%There are many alternate models of friction that also break control symmetry.
 Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
 Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \citep{fluidMechanics}.
-\begin{align}
-%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
-0, & \text{else} \end{cases}
-\end{align}
-
-The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.8\columnwidth]{friction.pdf} 
-\vspace{-1em}
-\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
-\label{fig:friction}
-}\vspace{-1em}
-\end{center}
-\end{figure}
+%\begin{align}
+%%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
+%F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
+%0, & \text{else} \end{cases}
+%\end{align}
+%
+%The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=0.8\columnwidth]{friction.pdf} 
+%\vspace{-1em}
+%\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
+%\label{fig:friction}
+%}\vspace{-1em}
+%\end{center}
+%\end{figure}
 
@@
 \centering
-\renewcommand{\figwid}{0.64\columnwidth}
+\renewcommand{\figwid}{0.67\columnwidth}
 {\begin{overpic}[width =\figwid]{RegionMove1.pdf}
@@
 
-\caption{\label{fig:regionMove}{Two rectangles representing the reachable regions for the current start and goal positions.
+\caption{\label{fig:regionMove}{Workspace and $\Delta$ configuration space for three sets of robot configurations with the same final goal.
 The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
@@
 
-\section{Algorithms}\label{sec:algorithms}
-\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
+%\section{Algorithms}\label{sec:algorithms}
+\section{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
 \begin{figure*}
-\centering
-\renewcommand{\figwid}{0.32\columnwidth}
-{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(65,80){Start}
+\renewcommand{\figwid}{0.4\columnwidth}
+{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(50,80){Move 1}
+\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(50,80){Move 2}
+\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(50,80){Move 3}
+\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
 \end{overpic}
@@
 }\\
-{\begin{overpic}[width =\figwid]{story5move0.pdf}\put(10,80){Start}
+
+%\vspace{1em}
+{
+\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,80){Move 1}
+\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,80){Move 2}
+\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,80){Move 3}
+\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,80){Move 5}
+\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
 \end{overpic}
 }\\
-{\begin{overpic}[width =\figwid]{story7move0.pdf}\put(65,80){Start}
-\end{overpic}
+
+%\vspace{1em}
+{
 \begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move3.pdf}\put(50,80){Move 3}
+\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 4}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move4.pdf}\put(50,80){Move 4}
+\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 6}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 5}
+\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 7}
 \end{overpic}
@@
 
-\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
+\section{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
 Alg.
@@
 \begin{center}
-	\includegraphics[width=1.0\columnwidth]{PositionNrobotsComp.pdf}
+	\includegraphics[width=1.0\columnwidth]{PositionNrobots.pdf}
 \end{center}
@@
 \begin{center}
-	\includegraphics[width=.3\columnwidth]{driftmove.pdf}
-	\includegraphics[width=.3\columnwidth]{driftmove2.pdf}
-	\includegraphics[width=.3\columnwidth]{driftmove3.pdf}
+\includegraphics[width=.47\columnwidth]{driftmove0.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove1.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove2.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove3.pdf}
 \end{center}

### 2017-02-24 IROS17/theory.tex
started writing the algorithm

@@
 \emph{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
+
 Global inputs move a swarm uniformly.  
 Shape control requires breaking this uniform symmetry.
-A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
+%A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
 %If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
@@
 We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.
-If $s_1$ is not also the topmost robot, we rotate the coordinate frame by 90$^\circ$ counterclockwise.
+If $s_1$ is not also the topmost robot, we mirror the coordinate frame by the right wall.
+To give an example to understand mirroring, consider the two starting position as $s_1 =  (0.2, 0.2) $ and $s_2 = (0.8, 0.8)$.
+Because the leftmost robot is not the topmost robot, we mirror the coordinate frame and by mirroring we will have new $s_1 = (0.2, 0.8)$ and new $s_2 = (0.8,0.2)$.
+We do this to reuse the main algorithm's logic and after the path is found, we unmirror the coordinate frame and translate the path.
+
 Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Define the sign function as:
-\begin{align}
-\sign (x) = \begin{cases}  1, &  x > 0 \\
-0, & x=0\\
--1, & x<0 \end{cases} 
-\end{align}
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. 
 The algorithm assigns a global control input at every instance.
-The goal is to adjust 
- $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
- $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
+The goal is to adjust the error in both coordinates,
+ $\Delta e =(\Delta e_x, \Delta e_y) =(g_2-g_1)- (s_2-s_1)$to $(0,0)$ using a shared global control input. 
  %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
  %employing the assumption we have made earlier about the walls' friction. 
+Our algorithm uses a dynamic search A* like method to find the shortest path in each move.
+Base case occurs when $\Delta e= (0,0)$.
+In this base case, simply the shortest path is a straight line between current position of each robot to its goal position.
+If  $\Delta e\ineq (0,0)$ then we consider the reachable set as shown in Fig.~\ref{fig:regionMove}.
+If a robot is touching a wall, the other robot can go where ever in the reachable set while the first robot is not moving.
+Two reachable sets are possible, horizontal and vertical.
+%Base case 1:  deltaX and deltaY are 0, simply move robots to final position
+%Base case 2:  In 2 moves we can make both deltaX and deltaY 0, implement this move, then call Base case 1:
+%
+%We build a list like in an A* algorithm.  Each item in the list has 4 entries, and the list is sorted based on the first entry:
+%{estimatedPathLength,  r1, r2, MOVES}
+%
+%MOVES is the moves that have been applied to the robots: {{x1,y1},{x2,y2}} for a two move sequence.
+%r1 and r2 are the positions of robot 1 and 2 after applying MOVES to them.
+%
+%estimatedPathLength is the sum of the euclidean distances of MOVES + an admissible heuristic (the maximum distance of r1 or r2 from their goal positions).
+%
+%Our algorithm solves the positioning problem in four steps: 
+%First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
+%Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
+%Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
+%%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
+%%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
+%Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
 
-Our algorithm solves the positioning problem in four steps: 
-First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
-Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
-Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
-%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.
+In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps.
+If we consider the length of each side of the square workspace as L, the worst case path length is $(\sqrt{2}+2)L$.
 

### 2017-02-24 IROS17/experiment.tex
Split the theory section so
Shiva can work on it separately

@@
 \subsection{Hardware Experiment: Position Control of n Robots}
-Kilobots have too much stochasticity to implement Alg.
-\ref{alg:PosControlNRobots}.
-Instead, a hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used.
+A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
   Designs for each are available at \citep{Arun2016Thingiverse}. 

### 2017-02-24 IROS17/theory.tex
Split the theory section so
Shiva can work on it separately

--- 2017-02-23 IROS17/theory.tex
@@
 
-\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
+\emph{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
+
 Global inputs move a swarm uniformly.  
 Shape control requires breaking this uniform symmetry.
-A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
-If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
-%Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
-%An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
-%Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
-Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
- These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
-Robots touching a wall have a friction force that opposes movement along the boundary.  
-This causes robots along the boundary to move more slowly than robots in free-space.
 
-Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary.
-$N$ is the normal or perpendicular force between the robot and the boundary.
-The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$.
-The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
-\begin{align}
-F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
-\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
-F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
-\text{and } N &= F \cos(\theta) \nonumber
-\end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
-Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
-This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
-There are many alternate models of friction that also break control symmetry.
-Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \citep{fluidMechanics}.
-\begin{align}
-%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
-0, & \text{else} \end{cases}
-\end{align}
-
-The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.8\columnwidth]{friction.pdf} 
-\vspace{-1em}
-\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
-\label{fig:friction}
-}\vspace{-1em}
-\end{center}
-\end{figure}
-
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.64\columnwidth}
-{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove2.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove3.pdf}
-\end{overpic}
-}\\
-
-\caption{\label{fig:regionMove}{Two rectangles representing the reachable regions for the current start and goal positions.
-The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
-The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
-}
-\vspace{-1em}
-}
-\end{figure*}
-
-\section{Algorithms}\label{sec:algorithms}
-\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.32\columnwidth}
-{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(65,80){Start}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(50,80){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(50,80){Move 3}
-\end{overpic}
-%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
-%\end{overpic}
-}\\
-{\begin{overpic}[width =\figwid]{story5move0.pdf}\put(10,80){Start}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,80){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,80){Move 5}
-\end{overpic}
-}\\
-{\begin{overpic}[width =\figwid]{story7move0.pdf}\put(65,80){Start}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move3.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move4.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. 
-Robot start positions are shown by a square, and goal positions by a circle.
-Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
-Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-\vspace{-1em}
-}
-\end{figure*}
-
-Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
-This algorithm  introduces concepts that will be used for multi-robot positioning.
-Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.
-If $s_1$ is not also the topmost robot, we rotate the coordinate frame by 90$^\circ$ counterclockwise.
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Define the sign function as:
-\begin{align}
-\sign (x) = \begin{cases}  1, &  x > 0 \\
-0, & x=0\\
--1, & x<0 \end{cases} 
-\end{align}
-The algorithm assigns a global control input at every instance.
-The goal is to adjust 
- $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
- $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
- %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction.
-
-Our algorithm solves the positioning problem in four steps: 
-First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
-Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
-Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
-%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.
-
-\begin{algorithm}
-\caption{PoseControl2Robots($s_1,s_2,g_1,g_2,L$)}\label{alg:PosControl2Robots}
-\begin{algorithmic}[1]
-\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates counterclockwise $90^{\circ}$.
-$\epsilon $ is a small, nonzero, user-specified value.
-
-\Ensure $(g_1, g_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\Delta g_y = e_2.y - e_1.y$
-\State $\Delta g_x = e_2.x - e_1.x$
-\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-%Second Move
-%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
-\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
-\Statex \Comment{Adjust  $y$}
-%Third Move
-\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
-\Statex \Comment{Touch bottom wall}
-%Fourth Move
-\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
-\If {$\Delta e_y \neq r_2.y - r_1.y$}
-%Fifth Move
-\If {$\Delta e_x \neq 0$}
-\State Move $(\sign(r_2.y- r_1.y)\Delta e_x, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
-\Else
-\State Move  $(|\sign(\Delta e_y -r_2.y+ r_1.y)| \epsilon, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
-\EndIf
-%Sixth Move
-\If {$\Delta e_x < 0$}
-\State Move $(L - \max(r_1.x, r_2.x), \epsilon)$
-\Else 
-\State Move $(L - \max(r_1.x, r_2.x),\Delta e_y -\max(r_1.y, r_2.y)$ 
-\Statex \Comment{Adjust $y$}
-\EndIf
-\EndIf
-%Seventh Move
-\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
-Alg.
-\ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints.
-The solution described here is an iterative procedure with $n$ loops.
-The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}.
-All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
-See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{PositionNrobotsComp.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:simulationNrobot}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:construction2d}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:contour}
-The contour plot keeps 3 points steady.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Number.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:numberPlot}
-Number of moves needed.
-}
-\end{figure}
-
-The setup in ??
-is an assembled practical setup that assumes that $\epsilon=2r = 1$ cm.
-It is a $7\times 7cm$ grid space.
-The wall friction is emulated by the jagged teeth design to keep rollers from moving out of place while implementing the drift move.
-The rollers are magnetic and can be controlled by magnetic steering.
-See video ``refer the video which shows moving rollers"  to see the algorithm at work.
-In the demonstration, the algorithm is a fixed step algorithm and all drift moves are an counterclockwise square move  of size $1cm\times1cm$.
-Once the kth roller gets to it?s designated location in every loop, a correction step is implemented.
-This makes the overall number of moves required for the fixed step algorithm more than that required for the Algorithm ?.
-??
-shows how there are only 6 stages per roller involved in Algorithm ??.
-The fixed step algorithm would go through 8 stages per roller as shown in ???.
-The fixed step algorithm can be found online.
-One significant difference between Algorithm ??
-and the fixed move implementation of it is that, we can achieve higher accuracy in placing rollers in their precise location if Algorithm 2 is used.
-
-Assume an open workspace with four axis-aligned walls with infinite friction.
-The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots.
-Without loss of generality, assume the build zone  is above the staging zone.
-Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.
-The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.
-
-The $n$ robot position control algorithm relies on a $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ control input, shown in Fig.\  \ref{fig:driftmove}.
-A drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-The robot touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=.3\columnwidth]{driftmove.pdf}
-	\includegraphics[width=.3\columnwidth]{driftmove2.pdf}
-	\includegraphics[width=.3\columnwidth]{driftmove3.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:driftmove}
-A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
-\vspace{-1em}
-\end{figure}
-
-Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot.
-Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.
-
-\begin{algorithm}
-\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
-\begin{algorithmic}[1]
-\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
-
-\While{ $p_{kx} > r$} 
-\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon,left)$    %drift move left until kth robot touches left wall
-\EndWhile
-
-\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
-\State $\beta \gets \frac{f_{ky}-r}{m}$
-\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
-\For{ $m$ iterations}
-\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon,up)$    %move kth robot to f_{ky} and leave the rest in position.
-\EndFor
-
-\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State Move ($f_{kx}-r, 0$)
-
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{ {\sc DriftMove}($\alpha,\beta,\epsilon,direction$)}\label{alg:DriftMove}
-\begin{algorithmic}[1]
-\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
-\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State Move ($f_{kx}-r, 0$)
-
-\end{algorithmic}
-\end{algorithm}
-
-Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
-First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
-Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
-Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
-Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
-Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.
+The following sections examine using boundaries that stop the particles if they are pushed into the wall.
+ These forces are  sufficient to break the symmetry caused by uniform inputs.

### 2017-02-24 IROS17/theory2robAlgs.tex
Split the theory section so
Shiva can work on it separately

\section{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}

\begin{figure}
\centering
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
\caption{\label{fig:TwoRegions}
Each robot gets the same control input. 
(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
} \vspace{-1em}
\end{figure}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.67\columnwidth}
{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{RegionMove2.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{RegionMove3.pdf}
\end{overpic}
}\\

\caption{\label{fig:regionMove}{Workspace and $\Delta$ configuration space for three sets of robot configurations with the same final goal. The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$. The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
}
\vspace{-1em}
}
\end{figure*}

\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
\caption{\label{fig:TwoRegions}
Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots. In the right picture, it is shown that the shortest path with three moves is the straight line between the starting position of the robot and the reflection of the goal.
} \vspace{-1em}
\end{figure}
\begin{figure*}
\renewcommand{\figwid}{0.4\columnwidth}
{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
\end{overpic}
\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
\end{overpic}
\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
\end{overpic}
\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
\end{overpic}
%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
%\end{overpic}
%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
%\end{overpic}
}\\

%\vspace{1em}
{
\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
\end{overpic}
\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
\end{overpic}
\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
\end{overpic}
\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
\end{overpic}
\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
\end{overpic}
}\\

%\vspace{1em}
{
\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 4}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 6}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 7}
\end{overpic}
}\\

\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. 
Robot start positions are shown by a square, and goal positions by a circle.  Dashed lines show the shortest route if robots could be controlled independently.  Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
%The bottom row shows an extreme case where the robots must switch position.
}
\vspace{-1em}
}
\end{figure*}

Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.  This algorithm  introduces concepts that will be used for multi-robot positioning. Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.

Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.  If $s_1$ is not also the topmost robot, we mirror the coordinate frame by the right wall. To give an example to understand mirroring, consider the two starting position as $s_1 =  (0.2, 0.2) $ and $s_2 = (0.8, 0.8)$. Because the leftmost robot is not the topmost robot, we mirror the coordinate frame and by mirroring we will have new $s_1 = (0.2, 0.8)$ and new $s_2 = (0.8,0.2)$. We do this to reuse the main algorithm's logic and after the path is found, we unmirror the coordinate frame and translate the path.

Denote the current positions of the robots  $r_1$ and $r_2$. 
Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. 
The algorithm assigns a global control input at every instance.
The goal is to adjust the error in both coordinates,
 $\Delta e =(\Delta e_x, \Delta e_y) =(g_2-g_1)- (s_2-s_1)$to $(0,0)$ using a shared global control input. 
 %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
 %employing the assumption we have made earlier about the walls' friction. 
Our algorithm uses a dynamic search A* like method to find the shortest path in each move. Base case occurs when $\Delta e= (0,0)$. In this base case, simply the shortest path is a straight line between current position of each robot to its goal position. 
If  $\Delta e\neq (0,0)$ then we consider the reachable set as shown in Fig.~\ref{fig:regionMove}. If a robot is touching a wall, the other robot can go where ever in the reachable set while the first robot is not moving. Two reachable sets are possible, horizontal and vertical. 
%Base case 1:  deltaX and deltaY are 0, simply move robots to final position
%Base case 2:  In 2 moves we can make both deltaX and deltaY 0, implement this move, then call Base case 1:
%
%We build a list like in an A* algorithm.  Each item in the list has 4 entries, and the list is sorted based on the first entry:
%{estimatedPathLength,  r1, r2, MOVES}
%
%MOVES is the moves that have been applied to the robots: {{x1,y1},{x2,y2}} for a two move sequence.
%r1 and r2 are the positions of robot 1 and 2 after applying MOVES to them.
%
%estimatedPathLength is the sum of the euclidean distances of MOVES + an admissible heuristic (the maximum distance of r1 or r2 from their goal positions).
%
%Our algorithm solves the positioning problem in four steps: 
%First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
%Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
%Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
%%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
%%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
%Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. If we consider the length of each side of the square workspace as L, the worst case path length is $(\sqrt{2}+2)L$.

\begin{algorithm}
\caption{PoseControl2Robots($s_1,s_2,g_1,g_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\scriptsize
\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
$(0,0)$ is bottom corner,
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
 Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates counterclockwise $90^{\circ}$.
$\epsilon $ is a small, nonzero, user-specified value.

\Ensure $(g_1, g_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
\State $\Delta g_y = e_2.y - e_1.y$
\State $\Delta g_x = e_2.x - e_1.x$
\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
%Second Move
%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
\Statex \Comment{Adjust  $y$}
%Third Move
\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
\Statex \Comment{Touch bottom wall}
%Fourth Move
\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
\If {$\Delta e_y \neq r_2.y - r_1.y$}
%Fifth Move
\If {$\Delta e_x \neq 0$}
\State Move $(\sign(r_2.y- r_1.y)\Delta e_x, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
\Else
\State Move  $(|\sign(\Delta e_y -r_2.y+ r_1.y)| \epsilon, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
\EndIf
%Sixth Move
\If {$\Delta e_x < 0$}
\State Move $(L - \max(r_1.x, r_2.x), \epsilon)$
\Else 
\State Move $(L - \max(r_1.x, r_2.x),\Delta e_y -\max(r_1.y, r_2.y)$ 
\Statex \Comment{Adjust $y$}
\EndIf
\EndIf
%Seventh Move
\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
\end{algorithmic}
\end{algorithm}

### 2017-02-24 IROS17/theorynrobAlgs.tex
Split the theory section so
Shiva can work on it separately

--- 2016-12-09 RA-L/theory.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
-\label{sec:theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=\linewidth]{SquareFill.pdf} 
-\vspace{-1em}
-\caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$.
-Left plot shows locus of possible mean positions for five values of $A$.
-The locus morphs from a square to a circle as $A$ increases.
-The covariance ellipse for each $A$ is shown with a dashed line.
-Center shows two corresponding arrangements of kilobots.
-At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
-See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
-\label{fig:SquareFill}
-\end{center}
-\end{figure*} 
-\begin{figure*}[!htb]
-\begin{center}
-\vspace{-1em}
-\includegraphics[width=\linewidth]{CircleFill.pdf} 
-\vspace{-1em}
-\caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$.
-Left plot shows locus of possible mean positions for four values of $h$.
-The locus of possible mean positions are concentric circles.
-See online interactive demonstration at \citep{Zhao2016mathematica}.}
-\label{fig:CircleFill}
-\end{center}
-\end{figure*}
-
-\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
-One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
-Like fluid settling in a tank, the stable final configuration minimizes potential energy.
-\paragraph{Square workspace}
-%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
-We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
-Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
-
-The range for the global input angle $\beta $ is [0,2$\pi $). In this range, the swarm assumes eight different polygonal shapes. 
-These shapes alternate between triangles and trapezoids when the area $A$$<$1/2, and between squares with one corner removed and trapezoids when $A$$>$1/2.
-
-Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
-%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
-
-\begin{align}
-\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \\
-%\end{align}
-%\begin{align}
-\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\\
-%\end{align}
-%\begin{align}
-\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
-\end{align}
-
-The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
-\begin{align}\label{eq:meanInSquareWorkspaceLL}
-\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
-	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
-\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
-	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
-\end{align}
-The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
-A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
-The covariance of a triangle is always $\pm(A/18)$.
-Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape.
-The range of mean positions are maximized when $A$ is small.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Circular workspace}
-Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
-Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
-For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
-Under a global input, the robot swarm fills the region under a chord with area
-\begin{align}
-A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
-\end{align}
-For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
-\begin{align}
-r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
-\end{align}
-Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.
-
-%Smashing the swarm to the corner walls will cause some correlation between \emph{x} and \emph{y} axes.
-Eq.
-~\ref{eq:Wall} shows the relation between number of robots and area provided and the correlation we can reach with it.
-%%TODO: ADD the equation for showing the correlation Vs area, vs number of robots
-% This equation shows that we are not able to reach high correlations bigger than $|0.5|$.
-Fig.
-~\ref{fig:wallCorrelation} shows the different mean positions and correlations we are able to reach.
-There are ways and corridors that the swarm needs to pass and also to keep the majority of it.
-If we use the wall corners technique and the way needs a very high correlation to pass, a remarkable number of the swarm members will be lost, and we may lose track of them due to their small size.
-Meanwhile, the variance of the swarm would get bigger and bigger that it would not anymore be able to deliver sufficient payloads.
-For this reason we have thought another idea which uses \emph{friction} to make correlations.
-%
-
-\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
-Global inputs move a swarm uniformly.  
-Shape control requires breaking this uniform symmetry.  
-A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
-If the swarm can flow around each other, pushing the swarm into a boundary produces the limited set of configurations presented in Sec.~\ref{subsec:FluidInTank}.
-%Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
-%An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
-%Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
-Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
- These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
-Robots touching a wall have a friction force that opposes movement along the boundary.  
-This causes robots along the boundary to move more slowly than robots in free-space.
-
-Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary.
-$N$ is the normal or perpendicular force between the robot and the boundary.
-The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$.
-The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
-\begin{align}
-F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
-\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
-F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
-\text{and } N &= F \cos(\theta) \nonumber
-\end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
-Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
-This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
-There are many alternate models of friction that also break control symmetry.
-Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \citep{fluidMechanics}.
-\begin{align}
-%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
-0, & \text{else} \end{cases}
-\end{align}
-
-The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.8\columnwidth]{friction.pdf} 
-\vspace{-1em}
-\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
-\label{fig:friction}
-}
-\end{center}
-\end{figure}
-
-\section{Algorithms}\label{sec:algorithms}
-\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.4\columnwidth}
-{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
-\end{overpic}}\\
-
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
-The algorithm only requires friction along the bottom and right walls.
-Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
-Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
-Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-%\vspace{-2em}
-}
-\end{figure*}
-
-Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
-This algorithm  introduces concepts that will be used for multi-robot positioning.
-Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Also the sign function is:
-\begin{align}
-\sign (x) = \begin{cases}  1, &  x > 0 \\
-0, & x=0\\
--1, & x<0 \end{cases} 
-\end{align}
-The algorithm assigns a global control input at every instance.
-The goal is to adjust 
- $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
- $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
- This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction.
-
-Our algorithm solves the positioning problem in four steps: 
-First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
-Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
-Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
-%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.
-
-\begin{algorithm}
-\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
-\begin{algorithmic}[1]
-\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
- $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
-\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\epsilon \gets$ robot diameter
-\State $\Delta e_y = e_2.y - e_1.y$
-\State $\Delta e_x = e_2.x - e_1.x$
-\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-%Second Move
-%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
-\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
-\Statex \Comment{Adjust  $y$}
-%Third Move
-\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
-\Statex \Comment{Touch bottom wall}
-%Fourth Move
-\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
-\If {$\Delta e_y \neq r_2.y - r_1.y$}
-%Fifth Move
-%TODO This step is not correct.
-\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
-%Sixth Move
-\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
-\Statex \Comment{Adjust $y$}
-\EndIf
-%Seventh Move
-\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
+\section{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
 Alg.
@@
 See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{PositionNrobots.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:simulationNrobot}
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
+}
+\end{figure}
 
@@
 \end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:contour}
+The contour plot keeps 3 points steady.
+}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Number.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:numberPlot}
+Number of moves needed.
+}
+\end{figure}
+
+The setup in ??
+is an assembled practical setup that assumes that $\epsilon=2r = 1$ cm.
+It is a $7\times 7cm$ grid space.
+The wall friction is emulated by the jagged teeth design to keep rollers from moving out of place while implementing the drift move.
+The rollers are magnetic and can be controlled by magnetic steering.
+See video ``refer the video which shows moving rollers"  to see the algorithm at work.
+In the demonstration, the algorithm is a fixed step algorithm and all drift moves are an counterclockwise square move  of size $1cm\times1cm$.
+Once the kth roller gets to it?s designated location in every loop, a correction step is implemented.
+This makes the overall number of moves required for the fixed step algorithm more than that required for the Algorithm ?.
+??
+shows how there are only 6 stages per roller involved in Algorithm ??.
+The fixed step algorithm would go through 8 stages per roller as shown in ???.
+The fixed step algorithm can be found online.
+One significant difference between Algorithm ??
+and the fixed move implementation of it is that, we can achieve higher accuracy in placing rollers in their precise location if Algorithm 2 is used.
 
@@
 \begin{center}
-	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
+\includegraphics[width=.47\columnwidth]{driftmove0.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove1.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove2.pdf}
+	\includegraphics[width=.47\columnwidth]{driftmove3.pdf}
 \end{center}

### 2017-02-25 IROS17/intro.tex
Split the theory section so
Shiva can work on it separately
Updated citation style
I dislike cited

@@
 % Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
-As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}-\citep{qiu2015magnetic}.
+As a current example, micro- and nano-robots can be manufactured in large numbers, see~\cite{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
 Someday large swarms of robots will be remotely guided
@@
 
-%\begin{figure}
-%\centering
-%\begin{overpic}[width=0.9\columnwidth]{Leaf.jpg}\end{overpic}
-%\caption{\label{fig:vascularNetwork}
-%Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.
-Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging
-%due to the many obstacles. 
-%This paper focuses on using boundary walls and wall friction to break the symmetry caused by the global input and control the shape of a swarm.} 
-%\end{figure}
-
-%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}.
-Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze.
-%Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.
-By adding rectangular boundary walls, some higher-order moments, such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls), are also controllable~\citep{ShahrokhiIROS2015}.
-%A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.
-%These techniques could only compress a swarm orthogonal to obstacles.
 Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
@@
 \ref{sec:conclusion}.
-%For controlling $Sec. igma_{xy}$, we prove that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction. 
-%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
-%We conclude by designing controllers that efficiently regulate $Sec. igma_{xy}$.
-%This paper
-%(1) proves that the swarm position covariance $Sec. igma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
-%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
-%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
-%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
-%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
-%TODO JOURNAL: design controllers that efficiently regulate $Sec. igma_{xy}$.
-%TODO JOURNAL: We will design Lyapunov-inspired controllers for $Sec. igma_{xy}$ to prove controllability. 
-%TODO JOURNAL:  and rank controllability as a function of friction.
-% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
-%\begin{figure}[t]
-%\centering
-%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
-%\vspace{-1em}
-%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
-%}\vspace{-1em}
-%\end{figure}
-% Our paper is organized as follows.
-After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.
-We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2017-02-25 IROS17/simulation.tex
Updated citation style
I dislike cited

@@
 Algorithm \ref{alg:PosControl2Robots}  was implemented in Mathematica using point robots (radius = $0$). 
-An online interactive demonstration and source code of the algorithm are available at \citep{Shahrokhi2015mathematicaParticle}.
+An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
   Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm wiht robot initial positions shown by hollow squares and final positions by circles. 
@@
 \subsection{Position Control of $n$ Robots}
-Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \citep{Arun2015}.
+Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \cite{Arun2015}.
 Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
@@
 %The plot compares  \emph{Total distance}---the sum of the moves made by every robot, with \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.
-See hardware implementation and simulation at \citep{Arun2015}.
+See hardware implementation and simulation at \cite{Arun2015}.
 }

### 2017-02-25 IROS17/intro.tex
worked on intro, theory, and experiment
mostly grammar and continuity

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a global field are common in applied mathematics, biology, and computer graphics. 
+Particle swarms propelled by a global field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics. 
 %If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
@@
 Instead, these robots are often controlled by a global, broadcast signal.
-They require techniques to shape the swarm that can reliably exploit large populations despite severe underactuation ($2 \ll2n$).
+ The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a global control field, and sensing, as the robot.
+Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the particle swarm.
+ Techniques are needed that can handle this underactuation. 
+ In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.
 
-Swarm shape control is necessary for navigation through narrow passages, for self-assembly, and for manipulating objects.
+Positioning is a foundational capability for a robotic system, but requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
+This paper relaxes that constraint, and provides position control algorithms that only require interactions with the boundaries.
+We assume that particles in contact with the boundaries have zero velocity if the global control input pushes the particle into the wall.
+
 \begin{figure}
@@
 \vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){n=2}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){Optimal shortest path}\end{overpic}\\
+\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){$n$=2}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){optimal shortest path}\end{overpic}\\
 \vspace{1em}
-\begin{overpic}[width=0.43\columnwidth]{monalizaBegin.pdf}\put(30, 100){n=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
+\begin{overpic}[width=0.43\columnwidth]{monalizaBegin.pdf}\put(30, 100){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
 \begin{overpic}[width=0.38\columnwidth]{monalisa.pdf}\put(-3, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
 \caption{\label{fig:IntroPic}
-Dynamic algorithm to find the shortest path for two robots is designed.
-The two top pictures shows the initial position and goal position of the robots and the optimal shortest path by making use of wall friction when shared control input is applied.
-The two bottom pictures shows the n robot positioning using shared control input and boundary walls.
+Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+Top row shows the output of a dynamic algorithm that finds the shortest path for two particles.
+ Top left shows the initial position and goal position of the particles.  
+ The shortest path consists of moving at angle 50$^\circ$ until the blue robot contacts the top wall, then moving the  magenta robot at angle 165$^\circ$ until the particles reach the desired relative spacing, then moving -60$^\circ$ to the goal positions.
+ The two bottom pictures shows $n$-particle positioning using shared control inputs and the boundary walls. 
 } \vspace{-1em}
@@
 To extend the range of possible shapes,
-  Sec.
-\ref{subsec:WallFriction} introduces general models for boundary interaction.
-We introduce algorithms with two orthogonal boundaries with high friction sufficient to arbitrarily position two robots in Sec.
+  Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
+We provide an optimal shortest-path algorithm for  arbitrarily position two robots in Sec.
 \ref{sec:PostionControl2Robots}, and Sec.
 \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
-%Both algorithms require robots with low process noise and require time proportional to the number of robots, so Sec.
-\ref{subsec:ClosedLoopCovarianceControl} introduces a robust and efficient technique to control swarm correlation.
-Sec.
-\ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.
-\ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}.
-We end with directions for future research in Sec.
-\ref{sec:conclusion}.
+Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
+ We end with directions for future research in Sec.  \ref{sec:conclusion}.

### 2017-02-25 IROS17/simulation.tex
worked on intro, theory, and experiment
mostly grammar and continuity

@@
 
-Two additional simulations were performed using wall-friction to control variance and covariance.
-The first is an open-loop algorithm that demonstrates the effect of varying friction levels.
-The second uses a closed-loop controller to achieve desired variance and covariance values.
-
 \subsection{Position Control of Two Robots}
@@
 Algorithm \ref{alg:PosControl2Robots}  was implemented in Mathematica using point robots (radius = $0$). 
-An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
-  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm wiht robot initial positions shown by hollow squares and final positions by circles. 
- Dashed lines show the shortest route if robots could be controlled independently, solid lines the generated path.
+%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
+  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
+ Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.
+
+ The contour plots in Fig.~\ref{fig:contour} show the length of the shortest path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace.
+This plot clearly shows the nonlinear nature of the path planning, with multiple isolate islands showing regions that are difficult to reach.
+ The contour plots in Fig.~\ref{fig:numberPlot} show the same configurations, but plot the required number of moves.
+There is never more than one $g_2$ position reachable in one move.
+If two-move solutions exist, they only appear along the boundary.
+If $g_2$ results in a contraction of $(\delta x, \delta y)$, there are many three move sequences.
+Four-move and six-move sequences are absent, but five and seven moves are sometimes required.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:contour}
+The contour plot keeps 3 points steady.
+}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{Number.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:numberPlot}
+Number of moves needed.
+}
+\end{figure}
 
@@
 As $\epsilon$ increases, the total distance decreases asymptotically, as shown in Fig.~\ref{fig:graphrobo.pdf}, because the robots have more room to maneuver and fewer $\operatorname{DriftMove}$s are required.
-%We observed that the value of ‘?’, greatly affects the total number of moves of commanded moves on the robots.
+%We observed that the value of $\epsilon$, greatly affects the total number of moves of commanded moves on the robots.
 The smaller the value of ‘?’ the greater is the number of moves required to move a said distance.
@@
 
-%The total commanded distance is plotted, which is approximately -0.15v^3 + 385x^2-35000x+300000
-%We compare the total distance moved and commanded  with the  \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.
-Because all robots are interchangeable, the LAP distance reduces to \[ \text{LAP} =   \sum_{k=1}^n  \left| f_{k,x}-p_{k,x} \right| +  \left|  f_{k,y}-p_{k,y}  \right| .
-\]
-
 \begin{figure}
@@
 robots.  
-%The plot compares  \emph{Total distance}---the sum of the moves made by every robot, with \emph{LAP distance}---the shortest distance according to the Linear Assignment Problem using Manhattan distance.
 See hardware implementation and simulation at \cite{Arun2015}.

### 2017-02-25 IROS17/theory.tex
worked on intro, theory, and experiment
mostly grammar and continuity

@@
 
-Global inputs move a swarm uniformly.  
+Global inputs in the absence of obstacles move a swarm uniformly.  
 Shape control requires breaking this uniform symmetry.
-
 The following sections examine using boundaries that stop the particles if they are pushed into the wall.
  These forces are  sufficient to break the symmetry caused by uniform inputs.
+
+ If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then we assume the following system model:
+ \begin{align*}\label{eq:swarmDynamicsAndFric} 
+ \begin{bmatrix}
+ \dot{x}_i\\
+ \dot{y}_i
+ \end{bmatrix}
+ &=
+ \mathbf{u}(t)
+ +F(x_i,y_i, \mathbf{u}(t)), \qquad i \in [1,n].\\
+ F(x_i,y_i, \mathbf{u}(t)) &= \begin{cases}
+  - \mathbf{u}(t) & \begin{matrix}(x_i,y_i) \in  \textrm{boundary \textbf{and}}\\
+\mathbf{N}(\textrm{boundary$(x_i,y_i)$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
+ \\
+ 0 & \textrm{else} 
+ \end{cases}
+ \end{align*}
+ Here $\mathbf{N}(\textrm{boundary$(x_i,y_i)$})$ is the normal to the boundary at position $(x_i,y_i)$, and
+ $F(x_i,y_i, \mathbf{u}(t)) $ is the frictional force provided by the boundary.
+
+These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
+ In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
+  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
+ Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.

### 2017-02-25 IROS17/theorynrobAlgs.tex
worked on intro, theory, and experiment
mostly grammar and continuity

@@
 \section{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
-Alg.
-\ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints. 
 The solution described here is an iterative procedure with $n$ loops. 
@@
   All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
-See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.
+   See Fig.~\ref{fig:simulationNrobot} for a schematic of the build and staging zones.
+
 \begin{figure}
@@
 
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:construction2d}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:contour}
-The contour plot keeps 3 points steady.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Number.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:numberPlot}
-Number of moves needed.
-}
-\end{figure}
-
-The setup in ??
-is an assembled practical setup that assumes that $\epsilon=2r = 1$ cm.
-It is a $7\times 7cm$ grid space.
-The wall friction is emulated by the jagged teeth design to keep rollers from moving out of place while implementing the drift move.
-The rollers are magnetic and can be controlled by magnetic steering.
-See video ``refer the video which shows moving rollers"  to see the algorithm at work.
-In the demonstration, the algorithm is a fixed step algorithm and all drift moves are an counterclockwise square move  of size $1cm\times1cm$.
-Once the kth roller gets to it?s designated location in every loop, a correction step is implemented.
-This makes the overall number of moves required for the fixed step algorithm more than that required for the Algorithm ?.
-??
-shows how there are only 6 stages per roller involved in Algorithm ??.
-The fixed step algorithm would go through 8 stages per roller as shown in ???.
-The fixed step algorithm can be found online.
-One significant difference between Algorithm ??
-and the fixed move implementation of it is that, we can achieve higher accuracy in placing rollers in their precise location if Algorithm 2 is used.
-
 Assume an open workspace with four axis-aligned walls with infinite friction.
@@
 
-The $n$ robot position control algorithm relies on a $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ control input, shown in Fig.\  \ref{fig:driftmove}.
-A drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-The robot touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
+The $n$ robot position control algorithm relies on a $\text{\sc DriftMove}(\alpha, \beta, \epsilon,\theta)$ control input, described in Alg.~\ref{alg:DriftMove} and shown in Fig.\  \ref{fig:driftmove}.
+For $\theta = 0^\circ$, a drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
+Any particle touching a top wall moves right $\beta$ units, while every particle not touching the top move right $\beta-\alpha$.
 
@@
 \begin{center}
-\includegraphics[width=.47\columnwidth]{driftmove0.pdf}
-	\includegraphics[width=.47\columnwidth]{driftmove1.pdf}
-	\includegraphics[width=.47\columnwidth]{driftmove2.pdf}
-	\includegraphics[width=.47\columnwidth]{driftmove3.pdf}
+%\includegraphics[width=.47\columnwidth]{driftmove0.pdf}
+%	\includegraphics[width=.47\columnwidth]{driftmove1.pdf}
+%	\includegraphics[width=.47\columnwidth]{driftmove2.pdf}
+	\includegraphics[width=\columnwidth]{driftmovefullres.pdf}
 \end{center}
@@
 \caption{\label{fig:driftmove}
-A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
+A  $\text{\sc DriftMove}(\alpha, \beta, \epsilon,0^\circ)$ repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
 Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
@@
 \While{ $p_{kx} > r$} 
-\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon,left)$    %drift move left until kth robot touches left wall
+\State $\text{\sc DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon,180^\circ)$    %drift move left until kth robot touches left wall
 \EndWhile
@@
 \For{ $m$ iterations}
-\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon,up)$    %move kth robot to f_{ky} and leave the rest in position.
+\State $\text{\sc DriftMove}(\alpha, \beta, \epsilon,90^\circ)$    %move kth robot to f_{ky} and leave the rest in position.
 \EndFor
@@
 \begin{algorithm}
-\caption{ {\sc DriftMove}($\alpha,\beta,\epsilon,direction$)}\label{alg:DriftMove}
+\caption{ {\sc DriftMove}($\alpha,\beta,\epsilon,\theta$)
+}\label{alg:DriftMove}
+particles touching the wall move $\beta$ units, while particles not touching the wall move $\beta-\alpha$ units.
 \begin{algorithmic}[1]
-\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
-\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State Move ($f_{kx}-r, 0$)
-
+\State $R = \begin{bmatrix} \cos(\theta) & -\sin(\theta) \\
+ \sin(\theta) & \cos(\theta)  \end{bmatrix}$
+\State {\sc Move}$(R \cdot [\beta/2,-\epsilon]^\top)$ 
+\State  {\sc Move}$(R \cdot [\beta/2,\epsilon]^\top)$ 
+\State  {\sc Move}$(R \cdot [-\alpha,0]^\top)$ 
 \end{algorithmic}
@@
 Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.
+
+The hardware platform depicted in \ref{fig:construction2d} is an assembled practical setup that assumes that $\epsilon= 1$ cm. 
+The workspace is a $7\times 7$ cm grid space. 
+All particles are 3D-printed plastic whose top is a 1cm diameter cylinder with a narrower base that encapsulates a steel bearing ball.
+Wall friction is emulated by a toothed wall design to keep particles from moving out of place while implementing the drift move. 
+The workspace boundary is mounted on top of a white sheet of cardboard.
+Underneath the cardboard, a grid of 3mm diameter magnets glued with 1 cm spacing to a thin board generates the global control input.
+ A video attachment  shows the algorithm at work. 
+This discretized setup requires several modification to Alg.~\ref{alg:PosControlNRobots}.
+ In this demonstration, all moves are 1 cm in length.
+   All drift moves are an counterclockwise \emph{square} move  of size 1 cm$\times$1 cm. 
+   Once the $k$th roller gets to its designated location in each loop, a correction step is implemented. 
+   This correction step increases by two the total number of moves required per particle.
+   Fig.~\ref{fig:simulationNrobot} shows there are only 6 stages per particle involved in Alg.~\ref{alg:PosControlNRobots}.
+The fixed step algorithm requires 8 stages per particle as shown in Fig.~\ref{fig:construction2d}.
+
+A significant difference between Alg.~\ref{alg:PosControlNRobots} and the fixed move implementation of it is that Alg.~\ref{alg:PosControlNRobots}
+enables placing particles at arbitrary, non-overlapping locations, while the fixed move implementation requires goal locations at the center of grid cells.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:construction2d}
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, discretized $n$ robot position control  using wall friction.
+}
+\end{figure}

### 2017-02-26 IROS17/theory.tex
small grammatical errors fixed

@@
 
- If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then we assume the following system model:
+ If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
  \begin{align*}\label{eq:swarmDynamicsAndFric} 
- \begin{bmatrix}
- \dot{x}_i\\
- \dot{y}_i
- \end{bmatrix}
+\dot{\mathbf{x}}_i(t)
  &=
  \mathbf{u}(t)
- +F(x_i,y_i, \mathbf{u}(t)), \qquad i \in [1,n].\\
- F(x_i,y_i, \mathbf{u}(t)) &= \begin{cases}
-  - \mathbf{u}(t) & \begin{matrix}(x_i,y_i) \in  \textrm{boundary \textbf{and}}\\
-\mathbf{N}(\textrm{boundary$(x_i,y_i)$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
+ +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), \qquad i \in [1,n].\\
+ F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
+  - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
+\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
  \\
@@
  \end{align*}
- Here $\mathbf{N}(\textrm{boundary$(x_i,y_i)$})$ is the normal to the boundary at position $(x_i,y_i)$, and
- $F(x_i,y_i, \mathbf{u}(t)) $ is the frictional force provided by the boundary.
+ Here $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$, and
+ $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary.
 

### 2017-02-26 IROS17/theorynrobAlgs.tex
small grammatical errors fixed

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints. 
+The ideas from Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ particles using wall friction.
+The solution is not optimal, and requires the starting and final configurations of particles to be disjoint.
 The solution described here is an iterative procedure with $n$ loops. 
  The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
-  All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
+  All robots move according to the global input, but due to wall friction, at the end of the $k^{\text{th}}$ loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
    See Fig.~\ref{fig:simulationNrobot} for a schematic of the build and staging zones.
@@
 For $\theta = 0^\circ$, a drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. 
-Any particle touching a top wall moves right $\beta$ units, while every particle not touching the top move right $\beta-\alpha$.
+ Any particle touching a top wall moves right $\beta$ units, while every particle not touching the top moves right $\beta-\alpha$.
 
@@
 Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
-First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
-Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
-Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
+First, the robots are moved left away from the right wall, and down so all robots in $k$'s row touch the bottom wall.
+Second, a set of $\operatorname{DriftMove}$s are executed that move all robots in $k$'s row left until $k$ touches the left wall, with no net movement of the other robots.
+Third, a set of $\operatorname{DriftMove}$s are executed that move only robot $k$ to its target height and return the other robots to their initial heights. 
 Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
@@
  A video attachment  shows the algorithm at work. 
-This discretized setup requires several modification to Alg.~\ref{alg:PosControlNRobots}.
+This discretized setup requires several modifications to Alg.~\ref{alg:PosControlNRobots}.
  In this demonstration, all moves are 1 cm in length.

### 2017-02-26 IROS17/intro.tex
Added tons of images, still working on writing the algorithm.

@@
 \vspace{1em}
-\begin{overpic}[width=0.43\columnwidth]{monalizaBegin.pdf}\put(30, 100){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
-\begin{overpic}[width=0.38\columnwidth]{monalisa.pdf}\put(-3, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(30, 102){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
 \caption{\label{fig:IntroPic}

### 2017-02-26 IROS17/theory2robAlgs.tex
Updated citation style
I dislike cited
Added tons of images, still working on writing the algorithm.

@@
 
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.67\columnwidth}
+{\begin{overpic}[width =\figwid]{epsilonMin.pdf}\put(20,20){$\epsilon = 0.001$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{epsilonMed.pdf}\put(20,20){$\epsilon = 0.05$}
+
+\end{overpic}
+\begin{overpic}[width =\figwid]{epsilonMax.pdf}\put(20,20){$\epsilon = 0.1$}
+
+\end{overpic}
+}\\
+
+\caption{\label{fig:epsilon}{By changing epsilon, the path changes and will ensure the minimum distance of $\epsilon$ between two robots.}
+\vspace{-1em}
+}
+\end{figure*}
+
 \begin{figure}
@@
 \begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
-\caption{\label{fig:TwoRegions}
+\caption{\label{fig:reflection}
 Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots.
@@
 \renewcommand{\figwid}{0.4\columnwidth}
+{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
 {\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
@@
 
-%\vspace{1em}
+\vspace{-0.75em}
 {
@@
 
+\vspace{-0.75em}
 %\vspace{1em}
@@
 
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 6}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 7}
+\end{overpic}
+}\\
+
+%\vspace{-0.75em}
+%{
+%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
+%\end{overpic}
+%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
+%%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
+%\end{overpic}
+%}\\
+
+\vspace{-1em}
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. 
@@
 Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
-Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
+Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
 %The bottom row shows an extreme case where the robots must switch position.
 }
+\vspace{-1em}
+}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{2\columnwidth}
+{\begin{overpic}[width =\figwid]{contourPlots.pdf}%\put(20,20){$\epsilon = 0.001$}
+\end{overpic}
+}\\
+\caption{\label{fig:contourPlots}{Starting positions of the robots and goal position of one robot is fixed, these contour plots calculate the distance if the other robot's goal position is at any other location.
+Number of moves are also showed in the bottom plots for the same configuration.}
 \vspace{-1em}
@@
 We do this to reuse the main algorithm's logic and after the path is found, we unmirror the coordinate frame and translate the path.
+To reuse the algorithm even more, we have implemented just the up wall and rotate the coordinate frame by 90, 180 or 270 degrees counterclockwise to get other walls paths and translate the path after the algorithm returns the path for the rotated coordinate frame.
 
@@
 Two reachable sets are possible, horizontal and vertical. 
+If $(\Delta g_x, \Delta g_y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+It is shown as $m_2$ in Fig.
+\ref{fig:reflection}.
+For each wall to find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
+The minimum distance to get to the goal in two move when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall.
+That point is selected when possible, and if it makes $m_2$ reach out of the workspace, the closest point to this point where the moving robot would not get out of the workspace is selected.
+
+If  $(\Delta g_x, \Delta g_y)$ is not in the reachable set, we choose the nearest $\Delta x$ and $\Delta y$ to $\Delta g_x$ and $\Delta g_y$ and reach there.
+In this step for simplicity we choose the straight line from the touching robot to the wall for the first move.
+This may cause the algorithm not to return the shortest path, but the difference is not significant.
+Alg.
+\ref{alg:wallup} returns the total distance of the path added to the euclidian distance of the final position of the robots to their goal.
+We use euclidian distance as a heuristic to find the path faster.
+
+Alg.
+\ref{alg:dynamicAlg} is a dynamic search to find the shortest path by all the walls.
+It keeps a list containing all the possible paths with their current and final position of the robots.
+At each iteration, the algorithm sorts all the paths in the list by their total distance plus heuristic.
+It picks the shortest path, and calls Alg.
+\ref{alg:Wallup} for all the four walls.
+It keeps calling and sorting until the goal has reached.
+When the goal is reached, the algorithm returns the shortest path and the final solution.
 %Base case 1:  deltaX and deltaY are 0, simply move robots to final position
@@
 \begin{algorithm}
-\caption{PoseControl2Robots($s_1,s_2,g_1,g_2,L$)}\label{alg:PosControl2Robots}
+\caption{WallFrictionMoveUp($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
 \begin{algorithmic}[1]
 \scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
 $(0,0)$ is bottom corner,
- $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates counterclockwise $90^{\circ}$.
+ $L$ is length of the walls. 
+ The array \emph{moves} is the current moves up to the current position.
+ Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots.
  $\epsilon $ is a small, nonzero, user-specified value.
 
-\Ensure $(g_1, g_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\Delta g_y = e_2.y - e_1.y$
-\State $\Delta g_x = e_2.x - e_1.x$
+\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least epsilon distance from each other.
+\State $\Delta e = (g_2 - g_1)- (r_2- r_1)$
+%\State $\Delta e_x = (g_2.x - g_1.x)- (r_2.x - r_1.x)$
+\If $\Delta e = (0,0)$
+\State 
+\EndIf
 \State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}

### 2017-02-27 IROS17/intro.tex
adjusted captions and over pic, changed r to d/2

@@
 \vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){$n$=2}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){$n$=2}\put(20, 4){{\tiny$\updownarrow$}~$\epsilon$}\end{overpic}
 \begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){optimal shortest path}\end{overpic}\\
-\vspace{1em}
-\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(30, 102){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 102){position control algorithm}\put(7,25){final configuration}\end{overpic}
+\vspace{1.5em}
+\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 102){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 102){position control algorithm}\put(12,25){final configuration}\end{overpic}
 \caption{\label{fig:IntroPic}

### 2017-02-27 IROS17/theory2robAlgs.tex
adjusted captions and over pic, changed r to d/2

@@
 \renewcommand{\figwid}{0.67\columnwidth}
-{\begin{overpic}[width =\figwid]{epsilonMin.pdf}\put(20,20){$\epsilon = 0.001$}
+{\begin{overpic}[width =\figwid]{epsilonMin.pdf}%\put(20,20){$\epsilon = 0.001$}
+\put(20, 3.1){{\scriptsize$\swarrow$}~$\epsilon = 0.001$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMed.pdf}\put(20,20){$\epsilon = 0.05$}
-
+\begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
+\put(20, 4){{\scriptsize$\updownarrow$}~$\epsilon = 0.05$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMax.pdf}\put(20,20){$\epsilon = 0.1$}
-
+\begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
+\put(20, 5){{\Large$\updownarrow$}~$\epsilon = 0.1$}
 \end{overpic}
 }\\
-
-\caption{\label{fig:epsilon}{By changing epsilon, the path changes and will ensure the minimum distance of $\epsilon$ between two robots.}
+\caption{\label{fig:epsilon}{Changing the minimum spacing $\epsilon$ changes the path.   $\epsilon$ is the minimum spacing between two robots and the minimum separation from the boundaries.}
 \vspace{-1em}
@@
 \begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
+\put(110,50){If $[\Delta x, \Delta y] = [0,0]$, only one move is needed}
 \end{overpic}
@@
 }\\
-\caption{\label{fig:contourPlots}{Starting positions of the robots and goal position of one robot is fixed, these contour plots calculate the distance if the other robot's goal position is at any other location.
-Number of moves are also showed in the bottom plots for the same configuration.}
+\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of  robot $2$ are fixed, and 
+ the contour plots calculate the distance if robot $1$'s goal position is varied in $x$ and $y$. The second row shows the number of moves required for the same configurations.}
 \vspace{-1em}

### 2017-02-27 IROS17/theorynrobAlgs.tex
adjusted captions and over pic, changed r to d/2

@@
 The ideas from Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ particles using wall friction.
-The solution is not optimal, and requires the starting and final configurations of particles to be disjoint.
+The solution is complete, but not optimal, and requires the starting and final configurations of particles to be disjoint.
 The solution described here is an iterative procedure with $n$ loops. 
@@
 The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots.
- Without loss of generality, assume the build zone  is above the staging zone. 
-Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.
-The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.
+ Without loss of generality, assume the build zone  is above the staging zone.  Let $d$ be the diameter the particles.
+Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + d$ to the left of the build and staging zone.
+The minimum workspace is then $(\epsilon + d + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.
 
@@
 \begin{algorithmic}[1]
-\State Move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
+\State Move( $-\epsilon, d/2-p_{ky}$) % move  away from right wall and down till robot k touches bottom
 
-\While{ $p_{kx} > r$} 
-\State $\text{\sc DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon,180^\circ)$    %drift move left until kth robot touches left wall
+\While{ $p_{kx} > d/2$} 
+\State $\text{\sc DriftMove}(\epsilon, \min(p_{kx} - d/2,\epsilon), \epsilon,180^\circ)$    %drift move left until kth robot touches left wall
 \EndWhile
 
-\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
-\State $\beta \gets \frac{f_{ky}-r}{m}$
-\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
+\State $m \gets \operatorname{ceil}(\frac{f_{ky}-d/2}{\epsilon})$
+\State $\beta \gets \frac{f_{ky}-d/2}{m}$
+\State $\alpha \gets \beta - \frac{d/2 - p_{ky}-\epsilon}{m}$
 \For{ $m$ iterations}
@@
 
-\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State Move ($f_{kx}-r, 0$)
+\State Move ($d/2+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
+\State Move ($f_{kx}-d/2, 0$)
 

### 2017-02-27 IROS17/simulation.tex
Added tons of images, still working on writing the algorithm.
added to caption

@@
 Four-move and six-move sequences are absent, but five and seven moves are sometimes required.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Contour.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:contour}
-The contour plot keeps 3 points steady.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{Number.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:numberPlot}
-Number of moves needed.
-}
-\end{figure}
 
@@
 The required number of moves under Alg. \ref{alg:PosControlNRobots}  using wall-friction to rearrange $n$ square-shaped 
-robots.  
+robots grows quadratically with $n$.  
 See hardware implementation and simulation at \cite{Arun2015}.

### 2017-02-27 IROS17/theory2robAlgs.tex
Added algorithms and edited one more image.

@@
 \begin{algorithmic}[1]
-\scriptsize
+%\scriptsize
 \Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
@@
  The array \emph{moves} is the current moves up to the current position.
+ $m_1$ is the first move toward the wall or goal.
+ $m_2$ is the second move adjusting $\Delta e$.
  Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots.
@@
 \State $\Delta e = (g_2 - g_1)- (r_2- r_1)$
+\State Admissible heuristic = $\max$(euclidian distance from goal)
 %\State $\Delta e_x = (g_2.x - g_1.x)- (r_2.x - r_1.x)$
-\If $\Delta e = (0,0)$
-\State 
+\If {$\Delta e = (0,0)$}
+\State $m_1 = (g_2.x -  r_2.x, g_2.y - r_2.y)$
+\State path $\gets$ path + $m_1$
+\State Return (total distance of path, path)
 \EndIf
-\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-%Second Move
-%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
-\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
-\Statex \Comment{Adjust  $y$}
-%Third Move
-\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
-\Statex \Comment{Touch bottom wall}
-%Fourth Move
-\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
-\If {$\Delta e_y \neq r_2.y - r_1.y$}
-%Fifth Move
-\If {$\Delta e_x \neq 0$}
-\State Move $(\sign(r_2.y- r_1.y)\Delta e_x, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
+\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g_x \leq  1 ~ \&\&~ r_2.y - r_1,y \leq \Delta g_y \leq  0$} %\Comment{Reachable region}
+\State $m_1 = (\frac{1-r_1.y}{2-g_1.y-r_1.y} g_1.x -r_1.x, 1-r_1.y)$
+\If {$r_2.x + m_1.x >L$}
+\State $m_1.x = 1- r_2.x$
+\EndIf
+\If {$r_2.x + m_1.x <0$}
+\State $m_1.x = -r_2.x$
+\EndIf
 \Else 
-\State Move  $(|\sign(\Delta e_y -r_2.y+ r_1.y)| \epsilon, |\sign(\Delta e_y)-1| |r_1.y-r_2.y|)$ \Comment{Touch right wall}
+\State $m_1 = (0, 1-r_1.y)$
+\State Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
+%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
 \EndIf
-%Sixth Move
-\If {$\Delta e_x < 0$}
-\State Move $(L - \max(r_1.x, r_2.x), \epsilon)$
-\Else 
-\State Move $(L - \max(r_1.x, r_2.x),\Delta e_y -\max(r_1.y, r_2.y)$ 
-\Statex \Comment{Adjust $y$}
+\State path $\gets$ path + $m_1$
+\State $r_1 = r_1 + m_1$ and $r_2 = r_2 + m_1$
+\State $m_2 = (\Delta g_x - (r_2.x- r_1.x), \Delta g_y - (r_2.y-r_1.y))$
+\If {Robots on each other or on the wall}
+\State Add $\pm \epsilon$ to $m_2.x$
 \EndIf
-\EndIf
-%Seventh Move
-\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
+\State path $\gets$ path + $m_2$
+%\State  $r_2 = r_2 + m_2$
+\State Unmirror if necessary
+\State Return(total distance of path + Admissible heuristic, path)
 \end{algorithmic}
 \end{algorithm}
+
+\begin{algorithm}
+\caption{DynamicOptimalSolutionFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:Wallup}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
+ $L$ is length of the walls. 
+ $OptimalPathList$ is a List containing all the paths sorted by their total distance. 
+\State Pop the first element in OptimalPathList
+\While {We have not reached the goal in the first element}
+\State Call Alg. \ref{alg:Wallup} for all the four walls.
+\State Add the results to OptimalPathList
+\State Sort OptimalPathList
+\EndWhile
+\State Return(path)
+\end{algorithmic}
+\end{algorithm}

### 2017-02-27 IROS17/theory2robAlgs.tex
added a caption

@@
 \caption{\label{fig:reflection}
-Consider the reflection of the goal position to the wall that is aimed to be touched by one of the robots.
-In the right picture, it is shown that the shortest path with three moves is the straight line between the starting position of the robot and the reflection of the goal.
+If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
 } \vspace{-1em}

### 2017-02-28 IROS17/theory2robAlgs.tex
couple of changes

@@
 \caption{\label{fig:TwoRegions}
-Each robot gets the same control input. 
+Boundary interaction is used to change relative position. Each robot gets the same control input. 
 (left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
@@
 \begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
-\put(110,50){If $[\Delta x, \Delta y] = [0,0]$, only one move is needed}
+\put(110,50){If $(\Delta e.x, \Delta e.y) = (0,0)$, only one move is needed}
 \end{overpic}
@@
 \vspace{-1em}
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction. 
+\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two robot positioning using walls with infinite friction. 
 Robot start positions are shown by a square, and goal positions by a circle.
 Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
+Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
 Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
@@
 
-Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
+Alg.~\ref{alg:Wallup} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
 This algorithm introduces concepts that will be used for multi-robot positioning.
@@
 
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $s_1$.
-If $s_1$ is not also the topmost robot, we mirror the coordinate frame by the right wall.
-To give an example to understand mirroring, consider the two starting position as $s_1 =  (0.2, 0.2) $ and $s_2 = (0.8, 0.8)$.
-Because the leftmost robot is not the topmost robot, we mirror the coordinate frame and by mirroring we will have new $s_1 = (0.2, 0.8)$ and new $s_2 = (0.8,0.2)$.
-We do this to reuse the main algorithm's logic and after the path is found, we unmirror the coordinate frame and translate the path.
-To reuse the algorithm even more, we have implemented just the up wall and rotate the coordinate frame by 90, 180 or 270 degrees counterclockwise to get other walls paths and translate the path after the algorithm returns the path for the rotated coordinate frame.
+Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $r_1$.
+If $r_1$ is not also the topmost robot, we mirror the coordinate frame across the right wall.
+To give an example to understand mirroring, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$.
+Because the leftmost robot is not the topmost robot, we mirror the coordinate frame giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$.
+We do this to reuse the main algorithm's logic and after the path is found, we undo the mirrorring to the output path.
+Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
+If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$ , 180$^{\circ}$  or 270$^{\circ}$ counterclockwise to get other walls paths and translate the path after the algorithm returns the path for the rotated coordinate frame.
 
 Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. 
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
 The algorithm assigns a global control input at every instance.
 The goal is to adjust the error in both coordinates,
- $\Delta e =(\Delta e_x, \Delta e_y) =(g_2-g_1)- (s_2-s_1)$to $(0,0)$ using a shared global control input. 
+ $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (s_2-s_1)$to $(0,0)$ using a shared global control input. 
  %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
  %employing the assumption we have made earlier about the walls' friction. 
-Our algorithm uses a dynamic search A* like method to find the shortest path in each move.
-Base case occurs when $\Delta e= (0,0)$.
-In this base case, simply the shortest path is a straight line between current position of each robot to its goal position.
-If  $\Delta e\neq (0,0)$ then we consider the reachable set as shown in Fig.~\ref{fig:regionMove}.
-If a robot is touching a wall, the other robot can go where ever in the reachable set while the first robot is not moving.
+Our algorithm uses a A* like method of dynamic search to find the shortest path in each move.
+The base case occurs when $\Delta e= (0,0)$.
+In this base case, the shortest path is a straight line between current position of each robot to its goal position, see Fig.
+\ref{fig:shapeControlMathematica1} a.
+If  $\Delta e\neq (0,0)$, we consult the two-move reachable sets shown in Fig.~\ref{fig:regionMove}.
+If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile.
 Two reachable sets are possible, horizontal and vertical. 
-If $(\Delta g_x, \Delta g_y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
-It is shown as $m_2$ in Fig.
+If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+This is shown as $m_2$ in Fig.
 \ref{fig:reflection}.
-For each wall to find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
+To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
 The minimum distance to get to the goal in two move when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall.
-That point is selected when possible, and if it makes $m_2$ reach out of the workspace, the closest point to this point where the moving robot would not get out of the workspace is selected.
+That point is selected when possible, but if it makes $m_2$ reach out of the workspace, the closest point to this point where the moving robot would not get out of the workspace is selected.
 
@@
 Alg.
-\ref{alg:wallup} returns the total distance of the path added to the euclidian distance of the final position of the robots to their goal.
+\ref{alg:Wallup} returns the total distance of the path added to the euclidian distance of the final position of the robots to their goal.
 We use euclidian distance as a heuristic to find the path faster.
@@
 Alg.
-\ref{alg:dynamicAlg} is a dynamic search to find the shortest path by all the walls.
+\ref{alg:optimalAlg} is a dynamic search to find the shortest path by all the walls.
 It keeps a list containing all the possible paths with their current and final position of the robots.
@@
 \begin{algorithm}
-\caption{DynamicOptimalSolutionFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:Wallup}
+\caption{DynamicOptimalSolutionFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
 \begin{algorithmic}[1]

### 2017-02-28 IROS17/theory2robAlgs.tex
Some critical errors in the code is fixed, text is updated

@@
 Two reachable sets are possible, horizontal and vertical. 
-If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
 This is shown as $m_2$ in Fig.
@@
 The minimum distance to get to the goal in two move when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall.
+If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
 That point is selected when possible, but if it makes $m_2$ reach out of the workspace, the closest point to this point where the moving robot would not get out of the workspace is selected.
 
-If  $(\Delta g_x, \Delta g_y)$ is not in the reachable set, we choose the nearest $\Delta x$ and $\Delta y$ to $\Delta g_x$ and $\Delta g_y$ and reach there.
-In this step for simplicity we choose the straight line from the touching robot to the wall for the first move.
+If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$.
+For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
 This may cause the algorithm not to return the shortest path, but the difference is not significant.
-Alg.
-\ref{alg:Wallup} returns the total distance of the path added to the euclidian distance of the final position of the robots to their goal.
-We use euclidian distance as a heuristic to find the path faster.
+Alg. \ref{alg:Wallup} uses an admissible heuristic that adds the current path length to the max Euclidian distance $||g-r||$ to direct exploration by expanding favorable routes first.
 
 Alg.
-\ref{alg:optimalAlg} is a dynamic search to find the shortest path by all the walls.
+\ref{alg:optimalAlg} is a dynamic search to find the shortest path by trying out collisions with all the walls.
 It keeps a list containing all the possible paths with their current and final position of the robots.
-At each iteration, the algorithm sorts all the paths in the list by their total distance plus heuristic.
-It picks the shortest path, and calls Alg.
+At each iteration, the algorithm sorts all the paths in the list by their total distance plus admissible heuristic.
+It selects the shortest path, and calls Alg.
 \ref{alg:Wallup} for all the four walls.
-It keeps calling and sorting until the goal has reached.
-When the goal is reached, the algorithm returns the shortest path and the final solution.
+It keeps calling and sorting until the goal is reached.
+When the goal is reached, the algorithm returns this shortest path.
 %Base case 1:  deltaX and deltaY are 0, simply move robots to final position
@@
 
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps.
+%In the worse case, adjusting both $\Delta r.x$ and $\Delta r.y$ needs all four steps. 
 If we consider the length of each side of the square workspace as L, the worst case path length is $(\sqrt{2}+2)L$.

### 2017-02-28 IROS17/intro.tex
couple of changes
Worked on algorithm2 and the description

@@
 Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
-Top row shows the output of a dynamic algorithm that finds the shortest path for two particles.
+Top row shows the output of a best-first-search algorithm that finds the shortest path for two particles.
  Top left shows the initial position and goal position of the particles.  
  The shortest path consists of moving at angle 50$^\circ$ until the blue robot contacts the top wall, then moving the  magenta robot at angle 165$^\circ$ until the particles reach the desired relative spacing, then moving -60$^\circ$ to the goal positions.
- The two bottom pictures shows $n$-particle positioning using shared control inputs and the boundary walls. 
+ The two bottom pictures shows $n$-particle positioning using shared control inputs and boundary interaction. 
 } \vspace{-1em}

### 2017-02-28 IROS17/theory.tex
Worked on algorithm2 and the description

@@
  If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
- \begin{align*}\label{eq:swarmDynamicsAndFric} 
+ \begin{align}\label{eq:swarmDynamicsAndFric} 
 \dot{\mathbf{x}}_i(t)
@@
  \mathbf{u}(t)
- +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), \qquad i \in [1,n].\\
+ +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), ~i \in [1,n].\\
  F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
@@
  0 & \textrm{else} 
- \end{cases}
- \end{align*}
+ \end{cases}\nonumber
+ \end{align}
  Here $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$, and

### 2017-02-28 IROS17/theory2robAlgs.tex
Added to one of the pictures.
Small Changes
Worked on algorithm2 and the description

--- 2017-02-28 IROS17/theory2robAlgs.tex
@@
-\section{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
+\section{Position Control of Two Robots Using Boundary Interaction}\label{sec:PostionControl2Robots}
 
-\begin{figure}
-\centering
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
-\caption{\label{fig:TwoRegions}
-Boundary interaction is used to change relative position. Each robot gets the same control input. 
-(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-} \vspace{-1em}
-\end{figure}
+Alg.~\ref{alg:optimalAlg} uses wall-friction to arbitrarily position two robots in a rectangular workspace. 
+ Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
 
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.67\columnwidth}
-{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove2.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove3.pdf}
-\end{overpic}
-}\\
+Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+The solution is a best-first search algorithm that maintains a list of possible paths ($moves$), sorted according to an admissible heuristic on path length.
+The algorithm works by expanding the path with the shortest estimated length.
+Expanding a path means either moving directly to the goal, or pushing one robot to a wall and adjusting the relative position of the other robot.
+As soon as the goal is reached, the algorithm returns this shortest path.
 
-\caption{\label{fig:regionMove}{Workspace and $\Delta$ configuration space for three sets of robot configurations with the same final goal.
-The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
-The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
-}
-\vspace{-1em}
-}
-\end{figure*}
+Denote the current positions of the robots  $r_1$ and $r_2$. 
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
+The algorithm assigns a global control input at every instance.
+The goal is to zero the error in both coordinates,
+ $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared global control input. 
+ The base case occurs when $\Delta e = (0,0)$.
+In this base case, the shortest path is a straight line between current position of each robot to its goal position, as shown in Fig.
+\ref{fig:shapeControlMathematica1}a.
+If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets shown in Fig.~\ref{fig:TwoRegions}.
+ If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile. 
+Two reachable sets are possible, horizontal and vertical. 
+The reachable sets may change after each move, as shown in  Fig.~\ref{fig:regionMove}.
 
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.67\columnwidth}
-{\begin{overpic}[width =\figwid]{epsilonMin.pdf}%\put(20,20){$\epsilon = 0.001$}
-\put(20, 3.1){{\scriptsize$\swarrow$}~$\epsilon = 0.001$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
-\put(20, 4){{\scriptsize$\updownarrow$}~$\epsilon = 0.05$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
-\put(20, 5){{\Large$\updownarrow$}~$\epsilon = 0.1$}
-\end{overpic}
-}\\
-\caption{\label{fig:epsilon}{Changing the minimum spacing $\epsilon$ changes the path.   $\epsilon$ is the minimum spacing between two robots and the minimum separation from the boundaries.}
-\vspace{-1em}
-}
-\end{figure*}
+ %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
+ %employing the assumption we have made earlier about the walls' friction. 
+%Our algorithm uses an A*-like method to find the shortest path in each move. 
+%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+This is shown as $m_2$ in Fig.
+\ref{fig:reflection}.
+To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
+%The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
+If the goal configuration can be reached in three moves:  $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
+\ref{fig:shapeControlMathematica1}b.
+$m_2$ cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
+\ref{fig:reflection}.
+That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
 
-\begin{figure}
-\centering
-\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
-\caption{\label{fig:reflection}
-If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
-The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
-} \vspace{-1em}
-\end{figure}
+If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
+%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
+This may cause the algorithm not to return the shortest path, but the difference is not significant.
+
+Alg.~\ref{alg:optimalAlg} uses an admissible heuristic that adds the current path length to the greatest distance from each robot to their goal.
+This heuristic directs exploration by expanding favorable routes first.
+\begin{align}\label{eq:admissibleHeuristic}
+h(moves, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|moves|} \norm{moves_i}  \\
+&+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
+\end{align}
+
+We  exploit symmetry in the solution by labeling the leftmost (or, if they have the same $x$ coordinate, the topmost) robot $r_1$. 
+ If $r_1$ is not also the topmost robot, we mirror the coordinate frame about the right boundary. 
+ As an example, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$. 
+  Because the leftmost robot is not the topmost robot, we mirror the coordinate frame about the right boundary giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$. 
+ After the path is found, we undo the mirroring to the output path. 
+  Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
+   If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$, 180$^{\circ}$, or 270$^{\circ}$ counterclockwise and then push the robot to hit the top wall.
+After the path is found, we undo the rotation.
+   This symmetry allows us to use a single function, Alg. \ref{alg:Wallup},  for collisions with all four walls.
+
+\begin{algorithm}[htb]
+\caption{2ParticlePathFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
+ $L$ is length of the walls. 
+ $OptimalPathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
+ \State $OptimalPathList \gets \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\}$ 
+\While {We have not reached the goal in the first element}
+\State Pop the first element in $OptimalPathList$
+\State Call Alg. \ref{alg:Wallup} for all the four walls.
+\State Add the results to $OptimalPathList$
+\State Sort($OptimalPathList$)
+\EndWhile
+\State \Return $moves$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{WallFrictionMoveUp($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
+ $L$ is length of the walls. 
+ The array \emph{moves} is the current sequence of moves up to the current position.
+ Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots, then undo the mirroring before returning.
+ $\epsilon $ is a small, nonzero, user-specified value.
+
+\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least $\epsilon$ distance from each other.  
+$m_1$ is the first move toward the wall or goal.
+ $m_2$ is the second move adjusting $\Delta e$.
+\State $\Delta e \gets (g_2 - g_1)- (r_2- r_1)$
+\If {$\Delta e = (0,0)$} \Comment{base case}
+\State $m_1  \gets g_2 -  r_2$
+\State $moves \gets \{ moves,   m_1 \}$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves\}$
+\EndIf
+\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
+\State $m_1 = \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
+\If {$r_2.x + m_1.x >L$}
+\State $m_1.x = 1- r_2.x$
+\EndIf
+\If {$r_2.x + m_1.x <0$}
+\State $m_1.x = -r_2.x$
+\EndIf
+\Else 
+\State $m_1 = (0, 1-r_1.y)$
+\State $\Delta g \gets $ closest reachable $(\Delta x,\Delta y)$.
+%Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
+%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
+\EndIf
+\State $moves \gets \{ moves,   m_1\}$
+\State $r_1 = r_1 + m_1$ \textbf{and} $r_2 = r_2 + m_1$
+\State $m_2 = \Delta g -  (r_2- r_1)$
+\If {robots on each other or on the wall}
+\State Add $\pm \epsilon$ to $m_2.x$
+\EndIf
+\State $moves \gets \{ moves,   m_2\}$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves\}$
+\end{algorithmic}
+\end{algorithm}
+
 \begin{figure*}
@@
 {\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
+\put(10,85){a)}
+\put(30,70){$r_1$}
+\put(48,88){$g_1$}
+\put(65,17){$r_2$}
+\put(87,52){$g_2$}
 \end{overpic}
@@
 {\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
+\put(10,85){b)}
+\put(28,70){$r_1$}
+\put(50,80){$g_1$}
+\put(62,15){$r_2$}
+\put(30,22){$g_2$}
 \end{overpic}
@@
 \begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
+\put(10,85){c)}
+\put(28,70){$r_1$}
+\put(62,28){$g_1$}
+\put(87,50){$r_2$}
+\put(30,22){$g_2$}
 \end{overpic}
@@
 \begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
+\put(10,85){d)}
+\put(25,30){$r_1$}
+\put(89,70){$g_1$}
+\put(67,20){$r_2$}
+\put(28,75){$g_2$}
 \end{overpic}
@@
 \begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
+\put(10,85){e)}
+\put(15,35){$r_1,g_2$}
+\put(65,35){$r_2,g_1$}
 \end{overpic}
@@
 \begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
+\put(10,85){f)}
+\put(15,10){$r_1,g_2$}
+\put(65,88){$r_2,g_1$}
 \end{overpic}
@@
 
+\begin{figure}
+\centering
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+\caption{\label{fig:TwoRegions}
+Boundary interaction is used to change the relative positions of the robots. Each robot gets the same control input. 
+(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
+(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
+} \vspace{-1em}
+\end{figure}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.67\columnwidth}
+{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{RegionMove2.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{RegionMove3.pdf}
+\end{overpic}
+}\\
+
+\caption{\label{fig:regionMove}{Workspace and $\Delta$ configuration space for three sets of robot configurations with the same final goal.
+The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
+The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
+}
+\vspace{-1em}
+}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.5\columnwidth}
+{\begin{overpic}[width =\figwid]{epsilonMin.pdf}%\put(20,20){$\epsilon = 0.001$}
+\put(20, 3.1){{\scriptsize$\swarrow$}~$\epsilon = 0.001$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
+\put(20, 4){{\scriptsize$\updownarrow$}~$\epsilon = 0.05$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
+\put(20, 5){{\Large$\updownarrow$}~$\epsilon = 0.1$}
+\end{overpic}
+}\\
+\caption{\label{fig:epsilon}{Changing the minimum spacing $\epsilon$ changes the path.   $\epsilon$ is the minimum spacing between two robots and the minimum separation from the boundaries.}
+\vspace{-1em}
+}
+\end{figure*}
+
+\begin{figure}
+\centering
+\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\caption{\label{fig:reflection}
+If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
+} \vspace{-1em}
+\end{figure}
+
 \begin{figure*}
@@
 \end{figure*}
-
-Alg.~\ref{alg:Wallup} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
-This algorithm introduces concepts that will be used for multi-robot positioning.
-Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-We can exploit symmetry in the solution by labeling the leftmost  (or, if they have the same $x$ coordinate, the topmost) robot $r_1$.
-If $r_1$ is not also the topmost robot, we mirror the coordinate frame across the right wall.
-To give an example to understand mirroring, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$.
-Because the leftmost robot is not the topmost robot, we mirror the coordinate frame giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$.
-We do this to reuse the main algorithm's logic and after the path is found, we undo the mirrorring to the output path.
-Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
-If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$ , 180$^{\circ}$  or 270$^{\circ}$ counterclockwise to get other walls paths and translate the path after the algorithm returns the path for the rotated coordinate frame.
-
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
-The algorithm assigns a global control input at every instance.
-The goal is to adjust the error in both coordinates,
- $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (s_2-s_1)$to $(0,0)$ using a shared global control input. 
- %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction. 
-Our algorithm uses a A* like method of dynamic search to find the shortest path in each move.
-The base case occurs when $\Delta e= (0,0)$.
-In this base case, the shortest path is a straight line between current position of each robot to its goal position, see Fig.
-\ref{fig:shapeControlMathematica1} a.
-If  $\Delta e\neq (0,0)$, we consult the two-move reachable sets shown in Fig.~\ref{fig:regionMove}.
-If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile.
-Two reachable sets are possible, horizontal and vertical. 
-If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
-This is shown as $m_2$ in Fig.
-\ref{fig:reflection}.
-To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
-The minimum distance to get to the goal in two move when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall.
-That point is selected when possible, but if it makes $m_2$ reach out of the workspace, the closest point to this point where the moving robot would not get out of the workspace is selected.
-
-If  $(\Delta g_x, \Delta g_y)$ is not in the reachable set, we choose the nearest $\Delta x$ and $\Delta y$ to $\Delta g_x$ and $\Delta g_y$ and reach there.
-In this step for simplicity we choose the straight line from the touching robot to the wall for the first move.
-This may cause the algorithm not to return the shortest path, but the difference is not significant.
-Alg.
-\ref{alg:Wallup} returns the total distance of the path added to the euclidian distance of the final position of the robots to their goal.
-We use euclidian distance as a heuristic to find the path faster.
-
-Alg.
-\ref{alg:optimalAlg} is a dynamic search to find the shortest path by all the walls.
-It keeps a list containing all the possible paths with their current and final position of the robots.
-At each iteration, the algorithm sorts all the paths in the list by their total distance plus heuristic.
-It picks the shortest path, and calls Alg.
-\ref{alg:Wallup} for all the four walls.
-It keeps calling and sorting until the goal has reached.
-When the goal is reached, the algorithm returns the shortest path and the final solution.
-%Base case 1:  deltaX and deltaY are 0, simply move robots to final position
-%Base case 2:  In 2 moves we can make both deltaX and deltaY 0, implement this move, then call Base case 1:
-%
-%We build a list like in an A* algorithm.  Each item in the list has 4 entries, and the list is sorted based on the first entry:
-%{estimatedPathLength,  r1, r2, MOVES}
-%
-%MOVES is the moves that have been applied to the robots: {{x1,y1},{x2,y2}} for a two move sequence.
-%r1 and r2 are the positions of robot 1 and 2 after applying MOVES to them.
-%
-%estimatedPathLength is the sum of the euclidean distances of MOVES + an admissible heuristic (the maximum distance of r1 or r2 from their goal positions).
-%
-%Our algorithm solves the positioning problem in four steps: 
-%First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
-%Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
-%Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
-%%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-%%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-%Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps.
-If we consider the length of each side of the square workspace as L, the worst case path length is $(\sqrt{2}+2)L$.
-
-\begin{algorithm}
-\caption{WallFrictionMoveUp($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- The array \emph{moves} is the current moves up to the current position.
- $m_1$ is the first move toward the wall or goal.
- $m_2$ is the second move adjusting $\Delta e$.
- Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots.
- $\epsilon $ is a small, nonzero, user-specified value.
-
-\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least epsilon distance from each other.
-\State $\Delta e = (g_2 - g_1)- (r_2- r_1)$
-\State Admissible heuristic = $\max$(euclidian distance from goal)
-%\State $\Delta e_x = (g_2.x - g_1.x)- (r_2.x - r_1.x)$
-\If {$\Delta e = (0,0)$}
-\State $m_1 = (g_2.x -  r_2.x, g_2.y - r_2.y)$
-\State path $\gets$ path + $m_1$
-\State Return (total distance of path, path)
-\EndIf
-\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g_x \leq  1 ~ \&\&~ r_2.y - r_1,y \leq \Delta g_y \leq  0$} %\Comment{Reachable region}
-\State $m_1 = (\frac{1-r_1.y}{2-g_1.y-r_1.y} g_1.x -r_1.x, 1-r_1.y)$
-\If {$r_2.x + m_1.x >L$}
-\State $m_1.x = 1- r_2.x$
-\EndIf
-\If {$r_2.x + m_1.x <0$}
-\State $m_1.x = -r_2.x$
-\EndIf
-\Else 
-\State $m_1 = (0, 1-r_1.y)$
-\State Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
-%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
-\EndIf
-\State path $\gets$ path + $m_1$
-\State $r_1 = r_1 + m_1$ and $r_2 = r_2 + m_1$
-\State $m_2 = (\Delta g_x - (r_2.x- r_1.x), \Delta g_y - (r_2.y-r_1.y))$
-\If {Robots on each other or on the wall}
-\State Add $\pm \epsilon$ to $m_2.x$
-\EndIf
-\State path $\gets$ path + $m_2$
-%\State  $r_2 = r_2 + m_2$
-\State Unmirror if necessary
-\State Return(total distance of path + Admissible heuristic, path)
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{DynamicOptimalSolutionFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- $OptimalPathList$ is a List containing all the paths sorted by their total distance. 
-\State Pop the first element in OptimalPathList
-\While {We have not reached the goal in the first element}
-\State Call Alg. \ref{alg:Wallup} for all the four walls.
-\State Add the results to OptimalPathList
-\State Sort OptimalPathList
-\EndWhile
-\State Return(path)
-\end{algorithmic}
-\end{algorithm}

### 2017-02-28 IROS17/theorynrobAlgs.tex
couple of changes
Worked on algorithm2 and the description

@@
-\section{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
+\section{Position Control of $n$ Robots Using Boundary Interaction}\label{sec:PostionControlnRobots}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The ideas from Alg. \ref{alg:PosControl2Robots}  can be extended to control the position of $n$ particles using wall friction.
+The ideas from Alg. \ref{alg:optimalAlg}  can be extended to control the position of $n$ particles using wall friction.
 The solution is complete, but not optimal, and requires the starting and final configurations of particles to be disjoint.

### 2017-02-28 IROS17/simulation.tex
Worked on algorithm2 and the description
All the ? are gone.

@@
 
-Algorithm \ref{alg:PosControl2Robots}  was implemented in Mathematica using point robots (radius = $0$). 
+Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using point robots (radius = $0$). 
 %An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
@@
 
- The contour plots in Fig.~\ref{fig:contour} show the length of the shortest path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace.
+ The contour plots in Fig.~\ref{fig:contourPlots} top row show the length of the shortest path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace.
 This plot clearly shows the nonlinear nature of the path planning, with multiple isolate islands showing regions that are difficult to reach.
- The contour plots in Fig.~\ref{fig:numberPlot} show the same configurations, but plot the required number of moves.
+ If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
+
+ The contour plots in Fig.~\ref{fig:contourPlots} bottom row show the same configurations, but plot the required number of moves.
 There is never more than one $g_2$ position reachable in one move.

### 2017-03-01 IROS17/theory2robAlgs.tex
updated algorithm

@@
 \begin{algorithm}[htb]
-\caption{2ParticlePathFinder($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
+\caption{ { \sc 2-ParticlePathFinder}($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
 \begin{algorithmic}[1]
@@
  $L$ is length of the walls. 
- $OptimalPathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
- \State $OptimalPathList \gets \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\}$ 
-\While {We have not reached the goal in the first element}
-\State Pop the first element in $OptimalPathList$
-\State Call Alg. \ref{alg:Wallup} for all the four walls.
-\State Add the results to $OptimalPathList$
-\State Sort($OptimalPathList$)
+ $PathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
+ \State  $PathList \gets \{\}$
+ \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains the admissible heuristic, the move sequence, and the current robot positions
+\While {$P.r_1 \ne g_1$ \textbf{and} $P.r_2 \ne g_2$}
+\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
+\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($P.r_1,P.r_2,g_1,g_2,\theta$)
+\State $\{d, $ \parbox[t]{.3\linewidth}{%
+ $moves, r_1,r_2\} \gets$\\
+ {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.moves$)}
+\State $(moves, r_1,r_2) \gets$ {\sc Rotate}($moves, r_1,r_2,-\theta$)
+\State   {\sc Push} $\{d, moves, r_1,r_2\} $ onto $PathList$
+\EndFor
+\State {\sc Sort}($PathList$) \Comment sort by admissible heuristic
+\State $P \gets $ {\sc Pop} first element of $PathList$
 \EndWhile
@@
 \begin{algorithm}
-\caption{WallFrictionMoveUp($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
+\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
 \begin{algorithmic}[1]
@@
 \State $moves \gets \{ moves,   m_1 \}$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves\}$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
 \EndIf
 \If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
-\State $m_1 = \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
+\State $m_1 \gets \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
 \If {$r_2.x + m_1.x >L$}
-\State $m_1.x = 1- r_2.x$
-\EndIf
-\If {$r_2.x + m_1.x <0$}
-\State $m_1.x = -r_2.x$
+\State $m_1.x \gets 1- r_2.x$
+%\EndIf
+\Else \textbf{ if} {$r_2.x + m_1.x < 0$}  \textbf{then}
+\State $m_1.x \gets -r_2.x$
 \EndIf
@@
 \State $moves \gets \{ moves,   m_1\}$
-\State $r_1 = r_1 + m_1$ \textbf{and} $r_2 = r_2 + m_1$
-\State $m_2 = \Delta g -  (r_2- r_1)$
-\If {robots on each other or on the wall}
-\State Add $\pm \epsilon$ to $m_2.x$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+\State $m_2 \gets \Delta g -  (r_2- r_1)$
+\If {robots on each other \textbf{or} on the wall}
+\State Add $\pm \epsilon$ to $m_2.x$ to avoid collision
 \EndIf
 \State $moves \gets \{ moves,   m_2\}$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves\}$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
 \end{algorithmic}

### 2017-03-01 IROS17/intro.tex
robot in white

@@
 \begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){$n$=2}\put(20, 4){{\tiny$\updownarrow$}~$\epsilon$}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(10, 102){optimal shortest path}\end{overpic}\\
+\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(30, 102){shortest path}\end{overpic}\\
 \vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 102){$n$=862}\put(16,60){goal}\put(30,20){robots}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 102){$n$=862}\put(16,60){goal}
+%\put(29.9,20){\scalebox{.91}[1.0]{\textcolor{white}{\textbf{robots}}}}
+% \put(30,20){\textcolor{black}{robots}}
+ \put(30,5){\textcolor{white}{\textbf{robots}}}
+\end{overpic}
 \begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 102){position control algorithm}\put(12,25){final configuration}\end{overpic}

### 2017-03-02 IROS17/theory2robAlgs.tex
Added a lot of things!
Deleted a couple of bad images, changed one word in abstract
a tiny change
Added references to all the figs.
Changed order of the figures.

@@
 That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
-
-If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
-%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
-This may cause the algorithm not to return the shortest path, but the difference is not significant.
-
-Alg.~\ref{alg:optimalAlg} uses an admissible heuristic that adds the current path length to the greatest distance from each robot to their goal.
-This heuristic directs exploration by expanding favorable routes first.
-\begin{align}\label{eq:admissibleHeuristic}
-h(moves, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|moves|} \norm{moves_i}  \\
-&+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
-\end{align}
-
-We  exploit symmetry in the solution by labeling the leftmost (or, if they have the same $x$ coordinate, the topmost) robot $r_1$. 
- If $r_1$ is not also the topmost robot, we mirror the coordinate frame about the right boundary. 
- As an example, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$. 
-  Because the leftmost robot is not the topmost robot, we mirror the coordinate frame about the right boundary giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$. 
- After the path is found, we undo the mirroring to the output path. 
-  Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
-   If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$, 180$^{\circ}$, or 270$^{\circ}$ counterclockwise and then push the robot to hit the top wall.
-After the path is found, we undo the rotation.
-   This symmetry allows us to use a single function, Alg. \ref{alg:Wallup},  for collisions with all four walls.
-
-\begin{algorithm}[htb]
-\caption{ { \sc 2-ParticlePathFinder}($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- $PathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
- \State  $PathList \gets \{\}$
- \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains the admissible heuristic, the move sequence, and the current robot positions
-\While {$P.r_1 \ne g_1$ \textbf{and} $P.r_2 \ne g_2$}
-\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
-\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($P.r_1,P.r_2,g_1,g_2,\theta$)
-\State $\{d, $ \parbox[t]{.3\linewidth}{%
- $moves, r_1,r_2\} \gets$\\
- {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.moves$)}
-\State $(moves, r_1,r_2) \gets$ {\sc Rotate}($moves, r_1,r_2,-\theta$)
-\State   {\sc Push} $\{d, moves, r_1,r_2\} $ onto $PathList$
-\EndFor
-\State {\sc Sort}($PathList$) \Comment sort by admissible heuristic
-\State $P \gets $ {\sc Pop} first element of $PathList$
-\EndWhile
-\State \Return $moves$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- The array \emph{moves} is the current sequence of moves up to the current position.
- Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots, then undo the mirroring before returning.
- $\epsilon $ is a small, nonzero, user-specified value.
-
-\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least $\epsilon$ distance from each other.  
-$m_1$ is the first move toward the wall or goal.
- $m_2$ is the second move adjusting $\Delta e$.
-\State $\Delta e \gets (g_2 - g_1)- (r_2- r_1)$
-\If {$\Delta e = (0,0)$} \Comment{base case}
-\State $m_1  \gets g_2 -  r_2$
-\State $moves \gets \{ moves,   m_1 \}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
-\EndIf
-\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
-\State $m_1 \gets \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
-\If {$r_2.x + m_1.x >L$}
-\State $m_1.x \gets 1- r_2.x$
-%\EndIf
-\Else \textbf{ if} {$r_2.x + m_1.x < 0$}  \textbf{then}
-\State $m_1.x \gets -r_2.x$
-\EndIf
-\Else 
-\State $m_1 = (0, 1-r_1.y)$
-\State $\Delta g \gets $ closest reachable $(\Delta x,\Delta y)$.
-%Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
-%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
-\EndIf
-\State $moves \gets \{ moves,   m_1\}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
-\State $m_2 \gets \Delta g -  (r_2- r_1)$
-\If {robots on each other \textbf{or} on the wall}
-\State Add $\pm \epsilon$ to $m_2.x$ to avoid collision
-\EndIf
-\State $moves \gets \{ moves,   m_2\}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
-\end{algorithmic}
-\end{algorithm}
+If $m_2$ causes the two particles to overlap, we add or subtract $\epsilon$ to $m_2.x$ to avoid collisions.
+This is shown in Fig.
+~\ref{fig:epsilon} with three different $\epsilon$ values.
 
@@
 \end{figure*}
-
 \begin{figure}
@@
 \end{figure}
+If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
+%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
+This may cause the algorithm not to return the shortest path, but the difference is not significant.
+
+Alg.~\ref{alg:optimalAlg} uses an admissible heuristic that adds the current path length to the greatest distance from each robot to their goal.
+This heuristic directs exploration by expanding favorable routes first.
+\begin{align}\label{eq:admissibleHeuristic}
+h(moves, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|moves|} \norm{moves_i}  \\
+&+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
+\end{align}
+
+We  exploit symmetry in the solution by labeling the leftmost (or, if they have the same $x$ coordinate, the topmost) robot $r_1$. 
+ If $r_1$ is not also the topmost robot, we mirror the coordinate frame about the right boundary. 
+ As an example, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$. 
+  Because the leftmost robot is not the topmost robot, we mirror the coordinate frame about the right boundary giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$. 
+ After the path is found, we undo the mirroring to the output path. 
+  Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
+   If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$, 180$^{\circ}$, or 270$^{\circ}$ counterclockwise and then push the robot to hit the top wall.
+After the path is found, we undo the rotation.
+   This symmetry allows us to use a single function, Alg. \ref{alg:Wallup},  for collisions with all four walls.
+
+\begin{algorithm}[htb]
+\caption{ { \sc 2-ParticlePathFinder}($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
+ $L$ is length of the walls. 
+ $PathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
+ \State  $PathList \gets \{\}$
+ \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains the admissible heuristic, the move sequence, and the current robot positions
+\While {$P.r_1 \ne g_1$ \textbf{and} $P.r_2 \ne g_2$}
+\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
+\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($P.r_1,P.r_2,g_1,g_2,\theta$)
+\State $\{d, $ \parbox[t]{.3\linewidth}{%
+ $moves, r_1,r_2\} \gets$\\
+ {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.moves$)}
+\State $(moves, r_1,r_2) \gets$ {\sc Rotate}($moves, r_1,r_2,-\theta$)
+\State   {\sc Push} $\{d, moves, r_1,r_2\} $ onto $PathList$
+\EndFor
+\State {\sc Sort}($PathList$) \Comment sort by admissible heuristic
+\State $P \gets $ {\sc Pop} first element of $PathList$
+\EndWhile
+\State \Return $moves$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$(0,0)$ is bottom corner,
+ $L$ is length of the walls. 
+ The array \emph{moves} is the current sequence of moves up to the current position.
+ Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots, then undo the mirroring before returning.
+ $\epsilon $ is a small, nonzero, user-specified value.
+
+\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least $\epsilon$ distance from each other.  
+$m_1$ is the first move toward the wall or goal.
+ $m_2$ is the second move adjusting $\Delta e$.
+\State $\Delta e \gets (g_2 - g_1)- (r_2- r_1)$
+\If {$\Delta e = (0,0)$} \Comment{base case}
+\State $m_1  \gets g_2 -  r_2$
+\State $moves \gets \{ moves,   m_1 \}$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
+\EndIf
+\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
+\State $m_1 \gets \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
+\If {$r_2.x + m_1.x >L$}
+\State $m_1.x \gets 1- r_2.x$
+%\EndIf
+\Else \textbf{ if} {$r_2.x + m_1.x < 0$}  \textbf{then}
+\State $m_1.x \gets -r_2.x$
+\EndIf
+\Else 
+\State $m_1 = (0, 1-r_1.y)$
+\State $\Delta g \gets $ closest reachable $(\Delta x,\Delta y)$.
+%Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
+%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
+\EndIf
+\State $moves \gets \{ moves,   m_1\}$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+\State $m_2 \gets \Delta g -  (r_2- r_1)$
+\If {robots on each other \textbf{or} on the wall}
+\State Add $\pm \epsilon$ to $m_2.x$ to avoid collision
+\EndIf
+\State $moves \gets \{ moves,   m_2\}$
+\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
+\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
+\end{algorithmic}
+\end{algorithm}
 
@@
 \end{figure*}
+\begin{figure}
+\centering
+\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\caption{\label{fig:reflection}
+If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
+} \vspace{-1em}
+\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
-\caption{\label{fig:reflection}
-If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
-The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
-} \vspace{-1em}
-\end{figure}
-
 \begin{figure*}
@@
 \renewcommand{\figwid}{2\columnwidth}
-{\begin{overpic}[width =\figwid]{contourPlots.pdf}%\put(20,20){$\epsilon = 0.001$}
+{\begin{overpic}[width =\figwid]{contourPlotDist.pdf}
+\end{overpic}
+
+\vspace{1em}
+\begin{overpic}[width =\figwid]{contourPlotNum.pdf}\put(97.5,21){7}
+\put(97.5,18.5){6}
+\put(97.5,16){5}
+\put(97.5,13.5){4}
+\put(97.5,11){3}
 \end{overpic}

### 2017-03-02 IROS17/theorynrobAlgs.tex
Added a lot of things!
added missing word "Fig."

@@
 A  $\text{\sc DriftMove}(\alpha, \beta, \epsilon,0^\circ)$ repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
+At the sequence end, robot $A$ has moved $\beta$ units right, and robot $B$ has moved $\beta-\alpha$ units right.}
 \vspace{-1em}
@@
 
-The hardware platform depicted in \ref{fig:construction2d} is an assembled practical setup that assumes that $\epsilon= 1$ cm. 
+The hardware platform depicted in Fig.~\ref{fig:construction2d} is an assembled practical setup that assumes that $\epsilon= 1$ cm. 
 The workspace is a $7\times 7$ cm grid space. 

### 2017-03-02 IROS17/experiment.tex
Updated citation style
I dislike cited
Added "with" to abstract, removed thingiverse made it 8 pages
edits, semi final version

@@
 
-\subsection{Hardware Experiment: Position Control of n Robots}
+\emph{Hardware Experiment: Position Control of n Robots}\\
 A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
-  Designs for each are available at \citep{Arun2016Thingiverse}. 
+ % Designs for each are available at \cite{Arun2016Thingiverse}. 
   The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.

### 2017-03-02 IROS17/simulation.tex
A very small modification.
Added references to all the figs.
edits, semi final version

@@
  The contour plots in Fig.~\ref{fig:contourPlots} top row show the length of the shortest path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace.
-This plot clearly shows the nonlinear nature of the path planning, with multiple isolate islands showing regions that are difficult to reach.
+This plot clearly shows the nonlinear nature of the path planning, with multiple isolated islands showing regions that are difficult to reach.
  If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
 
- The contour plots in Fig.~\ref{fig:contourPlots} bottom row show the same configurations, but plot the required number of moves.
-There is never more than one $g_2$ position reachable in one move.
-If two-move solutions exist, they only appear along the boundary.
-If $g_2$ results in a contraction of $(\delta x, \delta y)$, there are many three move sequences.
-Four-move and six-move sequences are absent, but five and seven moves are sometimes required.
+ The contour plots in Fig.~\ref{fig:contourPlots} bottom row show the same configurations, but plot the required number of moves.  There is never more than one $g_2$ position reachable in one move.  
+ %If two-move solutions exist, they only appear along the boundary.
+  If $g_2$ results in a contraction of $(\delta x, \delta y)$, there are many three move sequences. Four, five and seven moves are sometimes required.
 

### 2017-03-02 IROS17/theorynrobAlgs.tex
edits, semi final version

@@
 The solution described here is an iterative procedure with $n$ loops. 
- The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
+ The $k^{th}$ loop moves the $k^{th}$ robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
   All robots move according to the global input, but due to wall friction, at the end of the $k^{\text{th}}$ loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
@@
 The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots.
- Without loss of generality, assume the build zone  is above the staging zone.  Let $d$ be the diameter the particles.
+ Without loss of generality, assume the build zone  is above the staging zone.  Let $d$ be the diameter of the particles.
 Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + d$ to the left of the build and staging zone.
@@
 
-The $n$ robot position control algorithm relies on a $\text{\sc DriftMove}(\alpha, \beta, \epsilon,\theta)$ control input, described in Alg.~\ref{alg:DriftMove} and shown in Fig.\  \ref{fig:driftmove}.
+The $n$ robots position control algorithm relies on a $\text{\sc DriftMove}(\alpha, \beta, \epsilon,\theta)$ control input, described in Alg.~\ref{alg:DriftMove} and shown in Fig.\  \ref{fig:driftmove}.
 For $\theta = 0^\circ$, a drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. 
@@
 
-Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot.
+Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k^{th}$ robot, and $f_k$ the final $x,y$ position of the $k^{th}$ robot.
 Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.
@@
    All drift moves are an counterclockwise \emph{square} move  of size 1 cm$\times$1 cm. 
-   Once the $k$th roller gets to its designated location in each loop, a correction step is implemented. 
+   Once the $k^{th}$ roller gets to its designated location in each loop, a correction step is implemented. 
    This correction step increases by two the total number of moves required per particle.

### 2017-03-02 IROS17/theorynrobAlgs.tex
kith row

@@
 Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
-First, the robots are moved left away from the right wall, and down so all robots in $k$'s row touch the bottom wall.
-Second, a set of $\operatorname{DriftMove}$s are executed that move all robots in $k$'s row left until $k$ touches the left wall, with no net movement of the other robots.
+First, the robots are moved left away from the right wall, and down so all robots in $k^{th}$ row touch the bottom wall.
+Second, a set of $\operatorname{DriftMove}$s are executed that move all robots in $k^{th}$ row left until $k$ touches the left wall, with no net movement of the other robots.
 Third, a set of $\operatorname{DriftMove}$s are executed that move only robot $k$ to its target height and return the other robots to their initial heights. 

### 2017-03-02 IROS17/theory2robAlgs.tex
nicer kth and caption

@@
 \caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and 
- the contour plots calculate the distance if robot $1$'s goal position is varied in $x$ and $y$. The second row shows the number of moves required for the same configurations.}
+ the top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
 \vspace{-1em}

### 2017-03-02 IROS17/theorynrobAlgs.tex
nicer kth and caption

--- 2017-03-02 IROS17/theorynrobAlgs.tex
@@
 The solution described here is an iterative procedure with $n$ loops. 
- The $k^{th}$ loop moves the $k^{th}$ robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
+ The $k^{\text{th}}$ loop moves the $k^{\text{th}}$ robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
   All robots move according to the global input, but due to wall friction, at the end of the $k^{\text{th}}$ loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
@@
 
-Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k^{th}$ robot, and $f_k$ the final $x,y$ position of the $k^{th}$ robot.
+Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k^{\text{th}}$ robot, and $f_k$ the final $x,y$ position of the $k^{\text{th}}$ robot.
 Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.
@@
 Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
-First, the robots are moved left away from the right wall, and down so all robots in $k$'s row touch the bottom wall.
-Second, a set of $\operatorname{DriftMove}$s are executed that move all robots in $k$'s row left until $k$ touches the left wall, with no net movement of the other robots.
+First, the robots are moved left away from the right wall, and down so all robots in $k^{\text{th}}$ row touch the bottom wall.
+Second, a set of $\operatorname{DriftMove}$s are executed that move all robots in $k^{\text{th}}$ row left until $k$ touches the left wall, with no net movement of the other robots.
 Third, a set of $\operatorname{DriftMove}$s are executed that move only robot $k$ to its target height and return the other robots to their initial heights. 
@@
    All drift moves are an counterclockwise \emph{square} move  of size 1 cm$\times$1 cm. 
-   Once the $k^{th}$ roller gets to its designated location in each loop, a correction step is implemented. 
+   Once the $k^{\text{th}}$ roller gets to its designated location in each loop, a correction step is implemented. 
    This correction step increases by two the total number of moves required per particle.

### 2017-03-02 IROS17/theory2robAlgs.tex
epsilon!
Added epsilon to caption

@@
  Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-The solution is a best-first search algorithm that maintains a list of possible paths ($moves$), sorted according to an admissible heuristic on path length.
-The algorithm works by expanding the path with the shortest estimated length.
-Expanding a path means either moving directly to the goal, or pushing one robot to a wall and adjusting the relative position of the other robot.
-As soon as the goal is reached, the algorithm returns this shortest path.
-
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
-The algorithm assigns a global control input at every instance.
-The goal is to zero the error in both coordinates,
- $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared global control input. 
- The base case occurs when $\Delta e = (0,0)$.
-In this base case, the shortest path is a straight line between current position of each robot to its goal position, as shown in Fig.
-\ref{fig:shapeControlMathematica1}a.
-If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets shown in Fig.~\ref{fig:TwoRegions}.
- If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile. 
-Two reachable sets are possible, horizontal and vertical. 
-The reachable sets may change after each move, as shown in  Fig.~\ref{fig:regionMove}.
-
- %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction. 
-%Our algorithm uses an A*-like method to find the shortest path in each move. 
-%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
-This is shown as $m_2$ in Fig.
-\ref{fig:reflection}.
-To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
-%The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
-If the goal configuration can be reached in three moves:  $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
-\ref{fig:shapeControlMathematica1}b.
-$m_2$ cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
-\ref{fig:reflection}.
-That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
-If $m_2$ causes the two particles to overlap, we add or subtract $\epsilon$ to $m_2.x$ to avoid collisions.
-This is shown in Fig.
-~\ref{fig:epsilon} with three different $\epsilon$ values.
 
@@
 \end{figure*}
+Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+The solution is a best-first search algorithm that maintains a list of possible paths ($moves$), sorted according to an admissible heuristic on path length.
+The algorithm works by expanding the path with the shortest estimated length.
+Expanding a path means either moving directly to the goal, or pushing one robot to a wall and adjusting the relative position of the other robot.
+As soon as the goal is reached, the algorithm returns this shortest path.
+
+Denote the current positions of the robots  $r_1$ and $r_2$. 
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
+The algorithm assigns a global control input at every instance.
+The goal is to zero the error in both coordinates,
+ $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared global control input. 
+ The base case occurs when $\Delta e = (0,0)$.
+In this base case, the shortest path is a straight line between current position of each robot to its goal position, as shown in Fig.
+\ref{fig:shapeControlMathematica1}a.
+If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets shown in Fig.~\ref{fig:TwoRegions}.
+ If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile. 
+Two reachable sets are possible, horizontal and vertical. 
+The reachable sets may change after each move, as shown in  Fig.~\ref{fig:regionMove}.
+
+ %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
+ %employing the assumption we have made earlier about the walls' friction. 
+%Our algorithm uses an A*-like method to find the shortest path in each move. 
+%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+This is shown as $m_2$ in Fig.
+\ref{fig:reflection}.
+To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
+%The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
+If the goal configuration can be reached in three moves:  $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
+\ref{fig:shapeControlMathematica1}b.
+$m_2$ cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
+\ref{fig:reflection}.
+That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
+If $m_2$ causes the two particles to overlap, we add or subtract $\epsilon$ to $m_2.x$ to avoid collisions.
+This is shown in Fig.
+~\ref{fig:epsilon} with three different $\epsilon$ values.
+
 \begin{figure}
@@
 (right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-} \vspace{-1em}
+}
 \end{figure}
@@
 \begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
-\put(20, 4){{\scriptsize$\updownarrow$}~$\epsilon = 0.05$}
+\put(20, 3.9){{\tiny$\updownarrow$}~$\epsilon = 0.05$}
 \end{overpic}
 \begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
-\put(20, 5){{\Large$\updownarrow$}~$\epsilon = 0.1$}
+\put(20, 5){{\large$\updownarrow$}~$\epsilon = 0.1$}
 \end{overpic}
@@
 }\\
-\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and 
- the top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
+\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and $\epsilon=0.001$.
+ The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
 \vspace{-1em}

### 2017-03-22 journalSwarmControl/IJRR Archive/realExperiment.tex
First round on TASE paper (format, changes)

--- 2016-08-18 journalSwarmControl/realExperiment.tex
@@
 
-The kilobots performed five successful runs of object manipulation through the obstacle maze.
+The kilobots performed five successful runs of object manipulation through the obstacle maze, see Extension 2.
 Each trial used 100 kilobots.
-Videos of each are in the multimedia extension.
 Trials two-five were performed in a row with no failures in between.
@@
 A square object completed in 6871 s. 
-Three equilateral triangular objects of varying sizes failed in a total of seven runs.
+A rectangle and three equilateral triangle objects of varying sizes failed in a total of nine runs.
 Failures occurred when the object was pushed into a corner, requiring torque to be unstuck.
@@
 %\vspace{-1em}
-\caption{\label{fig:expSnapShot}Snapshots showing object manipulation experiment with 100 kilobots under automatic control.
-The automatic controller generates a policy to the goal.
-See  supplementary material video 1.}
+\caption{\label{fig:expSnapShot}Snapshots showing object manipulation experiment with 100 kilobots under automatic control. The automatic controller generates a policy to the goal, (see Extension 2).}
         \end{figure}

### 2017-03-22 journalSwarmControl/Introduction.tex
changed order of video numbers
The submitted version.
First round on TASE paper (format, changes)

@@
 These activities require robots that behave intelligently.
-Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population.
-Additionally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of robots is a daunting human-swarm interaction (HSI) challenge.
+Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire particle population.
+Additionally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge.
 
-Our previous work with over a hundred hardware robots and thousands of simulated robots demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
+Our previous work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
 Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
@@
 \centering
+%\begin{overpic}[height=6cm]{CompleteMazeDescription.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
+%\todo{I like the 'target' symbol, but it is not self-documenting.
+We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.
+I think these are easiest to make in powerpoint.
+%Please use the same color and line style for the variance min and max as you use in Figure 4.
+%}
+%{blockpushingImageWithMeanAndVarianceOverlay.png}
+\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of particles, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.
+Here a swarm of particles pushes a green hexagon toward the goal
+%Simulation at left, screen capture from hardware experiment at right. 
+(see also Extension 2).}
+\end{figure}
+
+\begin{figure}
+%\centering
 \renewcommand{\figwid}{0.3\columnwidth}
@@
 \begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
+\\
 \begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
@@
 \caption{\label{fig:5experiments}
-Screenshots from our online experiments controlling multi-robot systems with limited, global control.
-\textbf{(a)} Varying the number of robots from 1-500
+Screenshots from our online experiments controlling multi-particle systems with limited, global control.
+\textbf{(a)} Varying the number of particles from 1-500
 \textbf{(b)} Comparing 4 levels of visual feedback 
 \textbf{(c)} Varying noise from 0 to 200\% of control authority
-\textbf{(d)} Controlling the position of 1 to 10 robots
+\textbf{(d)} Controlling the position of 1 to 10 particles
 \textbf{(e)} Comparing 3 control architectures to assemble
@@
 
-Our online experiments show that numerous simple robots responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.
+Our online experiments show that numerous particles responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.
 
@@
 However, it is often possible to sense global properties of the group, such as mean position and density.
-Often robots are difficult or impossible to sense individually due to their size and location. 
+Often particles are difficult or impossible to sense individually due to their size and location. 
 For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group, such as mean position and variance.
-To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers, inspired by our online experiments, that require only mean and variance measurements of the robot's positions.
+To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers, inspired by our online experiments, that require only mean and variance measurements of the particle's positions.
 These controllers are used as primitives to perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.
 
-\begin{figure}
-\centering
-%\begin{overpic}[height=6cm]{CompleteMazeDescription.pdf}\end{overpic}
-\begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
-%\todo{I like the 'target' symbol, but it is not self-documenting.
-We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.
-I think these are easiest to make in powerpoint.
-%Please use the same color and line style for the variance min and max as you use in Figure 4.
-%}
-%{blockpushingImageWithMeanAndVarianceOverlay.png}
-\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.
-Here a swarm of robots pushes a green hexagon toward the goal.
-%Simulation at left, screen capture from hardware experiment at right. 
-See video 1 in supplemental materials.}
-\end{figure}
  %function and implementation function
@@
 We report the results of our online experiments in \S \ref{sec:expResults}.
-Inspired by these results, we prove that the mean and variance of a robot swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
+Inspired by these results, we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
 We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.

### 2017-03-22 journalSwarmControl/TASE/SwarmControlIJRR.tex
First round on TASE paper (format, changes)

--- 2016-08-18 journalSwarmControl/SwarmControlIJRR.tex
@@
+% compress:
+% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=CompressedSwarmControlIJRR.pdf SwarmControlIJRR.pdf
 %%%%%%%%% IJRR SETTINGS
 %  from https://us.sagepub.com/en-us/nam/manuscript-submission-guidelines#PreparingYourManuscript
-\documentclass[Afour,sageh,times]{sagej} % sage_latex_guidelines.tex V1.01, 11 June 2015
-\usepackage{moreverb,url}
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=black,urlcolor=black,linkcolor = black]{hyperref}
-\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
-T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\def\volumeyear{2016}
-%%%%%%%%%%%% END IJRR SETTINGS
+%\documentclass[Afour,sageh,times]{sagej} % sage_latex_guidelines.tex V1.01, 11 June 2015
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts% This command is only needed if 
+                                                          % you want to use the \thanks command
+%\overrideIEEEmargins                                      % Needed to meet printer requirements.
+\usepackage{times}
 
-% How many words:  $ detex IJRRensembleDifferentialDrive.tex | wc -w
+\makeatletter 
+\let\NAT@parse\undefined
+\makeatother
 
-%\usepackage[pdftex]{graphicx}        % standard LaTeX graphics tool
-%\usepackage[usenames,dvipsnames]{color}
+% numbers option provides compact numerical references in the text. 
+%\usepackage[numbers]{natbib}
+\usepackage{multicol}
+\usepackage[bookmarks=true]{hyperref}
+
+\usepackage{bbm}
 \usepackage{calc}
+\usepackage{url}
+\usepackage{transparent}
+\usepackage{hyperref}
+\hypersetup{
+  colorlinks =false,
+  urlcolor = black,
+  linkcolor = black
+}
 \usepackage{graphicx}
-%\usepackage[cmex10]{amsmath}
+\usepackage[cmex10]{amsmath}
+\usepackage{bm}
 \usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{color}
-\usepackage{colortbl} %TODO: lab computers choke on this.  Why?
-%\usepackage{cite}
-%\usepackage[font={small}]{caption}
+\usepackage{rotating}
+
+\usepackage{chngcntr}
+\counterwithin{paragraph}{subsection} % makes paragraph depend on subsection
+
+%\usepackage{xfrac}
+\usepackage{nicefrac}
+\usepackage{cite}
 \usepackage[caption=false,font=footnotesize]{subfig}
+\usepackage[usenames, dvipsnames]{color}
+\usepackage{colortbl}
+%\usepackage{caption}
+
+%\usepackage{wrapfig}
 \usepackage{overpic}
-\graphicspath{{./},{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
+%\usepackage{subfigure}
+%\usepackage{textcomp}
+\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
 \usepackage{breqn} %for breaking equations automatically
@@
 %\usepackage{algorithmic}
-\usepackage{bm}   % boldface math type
-\usepackage{upgreek}   % non-slanted greek letters
-\usepackage{soul} %for highlighting
+\usepackage{multirow}
+\usepackage{todonotes}
+\usepackage{authblk}
 
-%\usepackage[round]{natbib}
-\usepackage{mathtools}
-%\usepackage{authblk}
-\usepackage{wrapfig}
+%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
+% uncomment this to hide all red todos
+%\renewcommand{\todo}{}
 
 %% ABBREVIATIONS
-\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}
-\newcommand\mydefeq{\stackrel{\mathclap{\normalfont\mbox{\tiny def}}}{=}}
-\newcommand{\qstart}{\mathbf{q}_{\text{start}}}
-\newcommand{\qgoal}{\mathbf{q}_{\text{goal}}}
-\newcommand{\pstart}{\mathbf{p}_{\text{start}}}
-\newcommand{\pgoal}{\mathbf{p}_{\text{goal}}}
-\newcommand{\xstart}{x_{\text{start}}}
-\newcommand{\xgoal}{x_{\text{goal}}}
-\newcommand{\ystart}{y_{\text{start}}}
-\newcommand{\ygoal}{y_{\text{goal}}}
-\newcommand{\gammastart}{\gamma_{\text{start}}}
-\newcommand{\gammagoal}{\gamma_{\text{goal}}}
-\newcommand{\CM}{\textsc{cm}}
-\newcommand{\SVM}{\textsc{svm}}
-\newcommand{\NN}{\textsc{nn}}
-\newcommand{\prm}{\textsc{prm}}
-\newcommand{\lemur}{\textsc{lemur}}
-\newcommand{\Lemur}{\textsc{Lemur}}
-\newcommand{\LP}{\textsc{lp}} 
-\newcommand{\SOCP}{\textsc{socp}}
-\newcommand{\SDP}{\textsc{sdp}}
-\newcommand{\NP}{\textsc{np}}
-\newcommand{\SAT}{\textsc{sat}}
-\newcommand{\LMI}{\textsc{lmi}}
-\newcommand{\hrp}{\textsc{hrp\nobreakdash-2}}
-\newcommand{\DOF}{\textsc{dof}}
-\newcommand{\UIUC}{\textsc{uiuc}}
-\newcommand{\Cspace}{\mathcal{Q}}
-\newcommand{\Uspace}{\mathcal{U}}
-\newcommand{\D}{\nobreakdash-\textsc{d}}
-\newcommand{\contact}[1]{\Cspace_{#1}}
-\newcommand{\feasible}[1]{\Fspace_{#1}}
-\newcommand{\dd}{\; \mathrm{d}}
-\newcommand{\ii}{{\rm i}}
-\newcommand{\TRUE}{\textbf{true}}
-\newcommand{\FALSE}{\textbf{false}}
-
-\newcommand{\figwid}{0.22\columnwidth}
-%\newcommand{\topic}[1]{\textcolor{ForestGreen}{\footnotesize \textsf{#1}}}
-%\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}
+\newcommand{\qstart}{q_{\text{start}}}
 
 %% MACROS
-\providecommand{\proc}[1]{\textsc{#1}}
-\providecommand{\ddelta}{\underline{\bm{\updelta}}}  % requires \usepackage{bm,upgreek}
+
 \providecommand{\abs}[1]{\left\lvert#1\right\rvert}
@@
 \providecommand{\R}{\field{R}}
+\newcommand{\Cspace}{\mathcal{Q}}
+\newcommand{\Uspace}{\mathcal{U}}
 \providecommand{\Fspace}{\Cspace_\text{free}}
@@
 \providecommand{\Vcal}{$\mathcal{V}$}
+\DeclareMathOperator{\conv}{conv}
+\DeclareMathOperator{\cone}{cone}
+\DeclareMathOperator{\homog}{homog}
+\DeclareMathOperator{\domain}{dom}
+\DeclareMathOperator{\range}{range}
+\DeclareMathOperator{\sign}{sgn}
 \providecommand{\polar}{\triangle}
@@
 \providecommand{\bouter}{\overline{b}}
+\newcommand{\D}{\nobreakdash-\textsc{d}}
+%\newcommand{\Fspace}{\mathcal{F}}
 \providecommand{\Fspace}{\Cspace_\text{free}}
@@
 \providecommand{\Pouter}{\subouter{\Pproj}}
+\DeclareMathOperator{\argmax}{arg\,max}
 \providecommand{\Aineq}{B}
@@
 \providecommand{\beq}{t}
-
-\DeclareMathOperator{\conv}{conv}
-\DeclareMathOperator{\cone}{cone}
-\DeclareMathOperator{\homog}{homog}
-\DeclareMathOperator{\domain}{dom}
-\DeclareMathOperator{\range}{range}
-\DeclareMathOperator{\sign}{sgn} 
-\DeclareMathOperator{\argmax}{arg\,max}
 \DeclareMathOperator{\area}{area}
+\newcommand{\contact}[1]{\Cspace_{#1}}
+\newcommand{\feasible}[1]{\Fspace_{#1}}
+\newcommand{\dd}{\; \mathrm{d}}
+\newcommand{\figwid}{0.22\columnwidth}
+\newcommand{\TRUE}{\textbf{true}}
+\newcommand{\FALSE}{\textbf{false}}
 \DeclareMathOperator{\atan2}{atan2}
-\DeclareMathOperator*{\argmax2}{argmax}
+\allowdisplaybreaks
 
@@
 
-\begin{document}
-\setcounter{secnumdepth}{3}
-\runninghead{Shahrokhi \textit{et~al.}}
-\hyphenation{micro-robot nano-robot nano-robotics micro-robotics micro-robots nano-robots micro-manip-ulators}
+%\setcounter{secnumdepth}{3}
+%\runninghead{Shahrokhi \textit{et~al.}}
+%\hyphenation{micro-robot nano-robot nano-robotics micro-robotics micro-robots nano-robots micro-manip-ulators}
 %%%%%%%%%%%% For debugging purposes, I like to display the TOC
@@
 
-\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T. Becker}% end author block 
-\affiliation{Electrical and Computer Engineering, University of Houston, USA}
-\corrauth{Shiva Shahrokhi,
-University of Houston,
-4726 Calhoun Rd,
-Houston, TX 77004-4005, USA}
-\email{sshahrokhi2@uh.edu}
+\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T. Becker% end author block 
+\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2, atbecker\}@uh.edu}}
+}
 
+\begin{document}
+
+\maketitle
+\thispagestyle{empty}
+\pagestyle{empty}
 \begin{abstract}
@@
 
-\keywords{micro-robot, global control, swarm, human-swarm interaction}
-
-\maketitle
+%\keywords{micro-robot, global control, swarm, human-swarm interaction}
 
@@
 
-\begin{dci}
-The Authors declare that there is no conflict of interest.
-\end{dci}
-\begin{funding}
-%RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
-This work was supported by the National Science Foundation under Grant No.\ 
-\href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
-\end{funding}
-\begin{sm}
-\begin{itemize}
-\item Video 1: hardware experiments of object manipulation with a swarm of 100 kilobots.
-\item Video 2:  simulation of parameter sweeps with swarm of robots manipulating objects.
-\end{itemize}
-\end{sm}
+%\begin{dci}
+%The Authors declare that there is no conflict of interest.
+%\end{dci}
+%\begin{funding}
+%%RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
+%This work was supported by the National Science Foundation under Grant No.\ 
+%\href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
+%\end{funding}
+%\newpage
+%\begin{sm}
+%\begin{itemize}
+%\item Extension 1:  simulation video of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
+%\item Extension 2: hardware experiment videos of object manipulation with a swarm of 100 kilobots.
+%\end{itemize}
+%\end{sm}
 
@@
 
-%\bibliographystyle{IEEEtran}
-\bibliographystyle{SageH}
+\bibliographystyle{IEEEtran}
+%\bibliographystyle{SageH}
 \bibliography{IEEEabrv,SwarmControlIJRR}

### 2017-03-22 journalSwarmControl/TASE/expResults.tex
First round on TASE paper (format, changes)

--- 2016-08-16 journalSwarmControl/expResults.tex
@@
 Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 robots, and three control architectures for both an assembly task and a foraging task.
-The users could choose which experiment to try, and then our architecture randomly assigned a particular parameter value for each trial.
+Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 robots, and three control architectures for both assembly and foraging tasks.
+The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.
 We recorded the completion time and the participant ID for each successful trial.
@@
 \caption{
-\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.  Best-fit linear and quadratic lines are overlaid for comparison. 
+\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location. 
+% Best-fit linear and quadratic lines are overlaid for comparison. 
 %(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
@@
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [ mean, mean + variance, full-state, convex-hull].
+As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
@@
 Microrobots are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
-
-This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep as follows:
+This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
 \begin{align*}
@@
  \end{align*}
-$m_i,\psi_i$ are uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ is a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
+$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
 
@@
 \caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
-Right: Increasing the number of robots resulted in longer completion times.
+Right: For position control, increasing the number of robots resulted in longer completion times.
 For more than 4 robots the goal pattern contained a void, which may have caused the jump in completion times.
@@
 
-Note there is a bifurcation between $n$=4 and $n$=5 robots.
+Time increases for $n>4$ robots.
 For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.
 A better experiment design would randomly place the goal positions.
-Initially we tried this, but our beta-testers strongly disliked trying to arrange robots in random patterns.
+Initially we tried this, but beta-testers strongly disliked trying to arrange robots in random patterns.
 
@@
 
-The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
+The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
 The median time using repulsive control was four times longer than with attractive control.

### 2017-03-22 journalSwarmControl/expResults.tex
removed text for linear fit lines
First round on TASE paper (format, changes)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Human-swarm interaction results}\label{sec:expResults}
+%\emph{Human-swarm interaction results}\label{sec:expResults}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We designed six experiments to investigate human control of large robotic swarms for manipulation tasks.
+We designed six experiments to investigate human control of large swarms for manipulation tasks.
 Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 robots, and three control architectures for both an assembly task and a foraging task.
-The users could choose which experiment to try, and then our architecture randomly assigned a particular parameter value for each trial.
+Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
+The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.
 We recorded the completion time and the participant ID for each successful trial.
@@
 \caption{
-\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location.  Best-fit linear and quadratic lines are overlaid for comparison. 
+\label{fig:ResVaryNu}Data from \emph{Varying Number} using particles to push an object through a maze to a goal location. 
+% Best-fit linear and quadratic lines are overlaid for comparison. 
 %(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
@@
 Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems.
-This experiment varied from 1 to 500 the population of robots used to transport an object.
-The total area, maximum robot speed, and total net force the swarm could produce were constant.
-The robots pushed a large hexagonal object through an  {\sffamily S}-shaped maze.
-We hypothesized participants would complete the task faster with more robots.
-The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 robots.
+This experiment varied from 1 to 500 the population of particles used to transport an object.
+The total area, maximum particle speed, and total net force the swarm could produce were constant.
+The particles pushed a large hexagonal object through an  {\sffamily S}-shaped maze.
+We hypothesized participants would complete the task faster with more particles.
+The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 particles.
 
@@
 \caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}.
-This experiment challenges players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region.
+This experiment challenges players to quickly steer 100 particles (blue discs) to push an object (green hexagon) into a goal region.
 We record the completion time and other statistics.
@@
 Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.
-This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all robots; {\bf convex-hull} draws a convex hull around the outermost robots; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse.
-Fig.~\ref{fig:Visualization} shows screenshots of the same robot swarm with each type of visual feedback.
-Full-state requires $2n$ data points for $n$ robots.
+This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse.
+Fig.~\ref{fig:Visualization} shows screenshots of the same particle swarm with each type of visual feedback.
+Full-state requires $2n$ data points for $n$ particles.
 Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.
 Mean requires two, and variance three, data points.
-Mean and mean + variance are convenient even with millions of robots.
+Mean and mean + variance are convenient even with millions of particles.
 
@@
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [ mean, mean + variance, full-state, convex-hull].
+As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
 Using ANOVA analysis, we reject the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
-Anecdotal evidence from beta-testers who played the game suggests that tracking 100 robots is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
+Anecdotal evidence from beta-testers who played the game suggests that tracking 100 particles is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
 Our beta-testers described convex-hull feedback as confusing and irritating.
-A single robot left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
+A single particle left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
 %obscuring what the rest of the swarm is doing.
@@
 \paragraph{Varying noise:}
-Microrobots are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
-
-This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep as follows:
+Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
+This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
 \begin{align*}
@@
  \end{align*}
-$m_i,\psi_i$ are uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ is a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
+$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
 
-We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the robots move erratically.
+We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.
 Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.
@@
 \caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
-Right: Increasing the number of robots resulted in longer completion times.
-For more than 4 robots the goal pattern contained a void, which may have caused the jump in completion times.
+Right: For position control, increasing the number of particles resulted in longer completion times.
+For more than 4 particles the goal pattern contained a void, which may have caused the jump in completion times.
 %\vspace{-1em}
@@
 \paragraph{Position control:}
-This online experiment examined how completion time scales with the number of robots $n$.
-Using a single square obstacle, users arranged $n\in[1,10]$ robots into a specified goal pattern.
-The goal pattern formed a block  {\sffamily A} with 10 robots, and lesser numbers of robots used a subset of these goal positions.
-Our hypothesis was that completion time would increase linearly with the number of robots, as with our position control algorithm in \cite{Becker2013b}.
+This online experiment examined how completion time scales with the number of particles $n$.
+Using a single square obstacle, users arranged $n\in[1,10]$ particles into a specified goal pattern.
+The goal pattern formed a block  {\sffamily A} with 10 particles, and lesser numbers of particles used a subset of these goal positions.
+Our hypothesis was that completion time would increase linearly with the number of particles, as with our position control algorithm in \cite{Becker2013b}.
 Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.
-Though the number of robots presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
+Though the number of particles presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
 
-Note there is a bifurcation between $n$=4 and $n$=5 robots.
+Time increases for $n>4$ particles.
 For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.
 A better experiment design would randomly place the goal positions.
-Initially we tried this, but our beta-testers strongly disliked trying to arrange robots in random patterns.
+Initially we tried this, but beta-testers strongly disliked trying to arrange particles in random patterns.
 
 \paragraph{Varying control--Assembly:}
-Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.
+Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
 
@@
 In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
-The experiment challenged players to assemble a three-block pyramid with a swarm of 16 robots.
+The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.
 
@@
 This experiment also compared attractive and repulsive control with the global control used for the other experiments.
-The experiment challenged players to collect particles using a swarm of 100 robots and return the particles to a home region.
-The robots encapsulate the particles on contact.
+The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region.
+The particles encapsulate the particles on contact.
 

### 2017-03-22 journalSwarmControl/experiment.tex
changed order of video numbers
First round on TASE paper (format, changes)

@@
 The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
-The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, often because workspace obstacles made some robots unable to reach the object.
+The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, often because workspace obstacles made some particles unable to reach the object.
 This failure rate increases if  object weight increases or ground-robot friction increases.
@@
 \begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
-\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control. See supplemental material, video 2.
+\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control (see also Extension 1).
 }

### 2017-03-22 journalSwarmControl/simulation.tex
First round on TASE paper (format, changes)

@@
 
-\subsection{Controlling the mean position of many robots}
+\subsection{Controlling the mean position}
 
@@
 Representative experiments are shown in Fig.~\ref{fig:gainvalues}.
-100 robots were used and the maximum speed was 3 meters per second.
+100 particles were used and the maximum speed was 3 meters per second.
 As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$.
@@
 \vspace{-1em} 
-\caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots.
+\caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ particles.
 %b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom).
@@
 
-\subsection{Controlling the variance of many robots}
+\subsection{Controlling the variance}
 \begin{figure}
@@
 
-\subsection{Hybrid control of mean and variance of many robots}
+\subsection{Hybrid control of mean and variance}
 
-Fig.~\ref{fig:hybrid} shows a simulation run of the hybrid controller in Alg.~\ref{alg:MeanVarianceControl} with 100 robots in a square workspace containing no internal obstacles. 
+Fig.~\ref{fig:hybrid} shows a simulation run of the hybrid controller in Alg.~\ref{alg:MeanVarianceControl} with 100 particles in a square workspace containing no internal obstacles. 
 %\todo{plot showing 1.5 cycles of mean position, and a variance goal.  We might need a longer time}
@@
 \vspace{-1em}
-\caption{\label{fig:hybrid} Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom).
+\caption{\label{fig:hybrid} Simulation result with 100 particles under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom).
 For ease of analysis, only $x$ position and variance are shown.

### 2017-03-22 journalSwarmControl/IJRR Archive/pictures/pdf/Weight.m

@@ -1,0 +1,30 @@
+clear all;
+
+%....Results taken from 15 simulations...%
+result050 = [93.733;128.257;90.424;82.174;104.578;160.626;148.221;82.222;129.733;99.704]; 
+
+result075 = [122.759;77.254;112.602;97.494;81.712;89.443;87.103;83.894;143.415;125.442];
+result100 = [77.492;75.84;88.577;106.791;105.187;121.288;107.253;84.696;74.841;154.204];
+%result100 = [76.868;93.446;68.473;87.333;84.375;116.218;108.54;106.594;81.217;74.186];
+result125 = [137.918;76.143;89.689;130.424;100.966;119.111;83.886;72.423;182.625;109.818];
+result150 = [96.93;64.471;92.351;114.401;100.407;75.588;101.824;89.568;143.857;112.561];
+
+%...to plot the boxplot...
+Time = [result050; result075; result100; result125; result150];
+
+
+ group = ['result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';
+     %'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';
+     'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';
+      %'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';
+    'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';
+    %'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';
+    'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';
+    %'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';
+    %'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';
+     'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';];
+
+boxplot(Time, group);
+title('How Object Density will affect time');
+xlabel('Algorithm');
+ylabel('Time (s)');

### 2017-03-24 journalSwarmControl/Introduction.tex
reworked abstract and intro

@@
 Additionally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge.
+This paper examines object manipulation by a swarm of particles, each actuated by the same, shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}.
 
-Our previous work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
+ This paper is the synthesis of two preliminary conference papers.
+\cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
+This paper presents three years of results from new games at SwarmControl.net.
+For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parametric studies.
+All hardware validation experiments are new.
+
+Our early work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}. 
 Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
@@
  Our paper is organized as follows.
-After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment in \S \ref{sec:expMethods}.
-We report the results of our online experiments in \S \ref{sec:expResults}.
+After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in \S \ref{sec:expMethods}.
 Inspired by these results, we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
@@
  We conclude with implementations of these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}.
-
- This paper is the synthesis of two preliminary conference papers.
-\cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
-This paper presents the final results from SwarmControl.net.
-All hardware validation experiments are new.

### 2017-03-25 journalSwarmControl/Introduction.tex
cleaned up intro

@@
 Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire particle population.
-Additionally, many promising applications will require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge.
 This paper examines object manipulation by a swarm of particles, each actuated by the same, shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}.
 
- This paper is the synthesis of two preliminary conference papers.
+This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a
+single journal paper.
   \cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
@@
 
+Many promising applications for particle swarms require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge.
 Our early work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}.
@@
 \centering
-%\begin{overpic}[height=6cm]{CompleteMazeDescription.pdf}\end{overpic}
 \begin{overpic}[width=\columnwidth]{MainExpFig.pdf}\end{overpic}
-%\todo{I like the 'target' symbol, but it is not self-documenting.
-We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.
-I think these are easiest to make in powerpoint.
-%Please use the same color and line style for the variance min and max as you use in Figure 4.
-%}
-%{blockpushingImageWithMeanAndVarianceOverlay.png}
-\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of particles, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.
-Here a swarm of particles pushes a green hexagon toward the goal
-%Simulation at left, screen capture from hardware experiment at right. 
-(see also Extension 2).}
+\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} 
+A swarm of particles, all controlled by a uniform control input, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.  
+Here a swarm of particles (kilobot robots) pushes a green hexagon toward the goal (see video attachment).
+}
 \end{figure}

### 2017-03-27 journalSwarmControl/expResults.tex
shortened the paper to 12 pages.
Worked on related work and cut some more of the paper

@@
 
-Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems.
+%Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems. 
 This experiment varied from 1 to 500 the population of particles used to transport an object.
@@
 
-Sensing is expensive, especially on the nanoscale.
-To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source.
-Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}.
-Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.
+%Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source. 
+%Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. 
+%Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  
 This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse.
@@
 \paragraph{Varying noise:}
-Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
-This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
+%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
+This experiment varied the strength of %these 
+disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
 \begin{align*}
@@
 
-Time increases for $n>4$ particles.
+%Time increases for $n>4$ particles.
 For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.

### 2017-03-27 journalSwarmControl/relatedWork.tex
Worked on related work and cut some more of the paper

@@
 \cite{Tottori2012} and \cite{Peyer2013}.
+
 %In our previous work with robots that can be modeled as nonholonomic unicycles~\cite{Becker2012k}, we showed that an inhomogeneity in turning speed is enough to make even an infinite number of robots controllable with regard to position.
@@
  \paragraph{Sensing---large populations:}
- Parallel control of $n$ differential-drive robots in a 2D workspace requires $3n$ state variables.
+$n$ differential-drive robots in a 2D workspace require $3n$ state variables.
 Even holonomic robots require $2n$ state variables.
@@
  \subsection{Human-swarm interaction}
+
+  Most humans are able, with practice, steer a swarm of robots controlled by a global input.
+Prior to our paper, no algorithm existed.
+Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
 
@@
 Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.
-Our swarms conform to the object to be manipulated.
-However our swarms lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
+Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
 Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},

### 2017-03-27 journalSwarmControl/theory.tex
Worked on related work and cut some more of the paper

@@
 We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.
-One option is to control the position and velocity of the $j^{th}$ robot.
-To find a potentially more useful option, we create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
+%One option is to control the position and velocity of the $j^{th}$ robot.
+To find a potentially more useful option,
+We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
 \begin{align}
@@
 
-We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \citet{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
+We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
 \begin{align}

### 2017-03-31 journalSwarmControl/Introduction.tex
The newest version of the TASE paper

@@
 Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire particle population.
-This paper examines object manipulation by a swarm of particles, each actuated by the same, shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}.
+This paper examines object manipulation by a swarm of particles, each actuated by the same shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}.
 
@@
 The goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size.
-Towards this end, we created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}, an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
+Towards this end, we have created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}: an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
 Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.
@@
 Often particles are difficult or impossible to sense individually due to their size and location. 
-For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group, such as mean position and variance.
-To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers, inspired by our online experiments, that require only mean and variance measurements of the particle's positions.
+For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
+To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions.
 These controllers are used as primitives to perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.

### 2017-03-31 journalSwarmControl/expResults.tex
The newest version of the TASE paper

@@
 
-\paragraph{Varying number:}
+\paragraph{Varying number}
 \begin{figure}
@@
 
-\paragraph{Varying visualization:}
+\paragraph{Varying visualization}
 \begin{figure}[b!]
@@
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
-Using ANOVA analysis, we reject the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
+Using ANOVA analysis, we rejected the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
 Anecdotal evidence from beta-testers who played the game suggests that tracking 100 particles is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
@@
 
-\paragraph{Varying noise:}
+\paragraph{Varying noise}
 %Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
@@
 
-\paragraph{Position control:}
+\paragraph{Position control}
 This online experiment examined how completion time scales with the number of particles $n$.
@@
 
-\paragraph{Varying control--Assembly:}
+\paragraph{Varying control--Assembly}
 Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
@@
 
-\paragraph{Varying control--Foraging:}
+\paragraph{Varying control--Foraging}
 Collecting and delivering resources is necessary for drug delivery.

### 2017-03-31 journalSwarmControl/experiment.tex
The newest version of the TASE paper

@@
 
-This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers, inspired by the analysis of human users in \S \ref{sec:expResults}.  
+This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers.%, inspired by the analysis of human users in \S \ref{sec:expResults}.  
 The swarm must deliver the object to the goal region.
-To solve this object manipulation task, we divide the task into two components: designing a policy for the object, and designing a control law for the swarm to push the object according to the policy.
+To solve this object manipulation task we divide the task into two components: designing a policy for the object, and designing a control law for the swarm to push the object according to the policy.
 
@@
 
-To design the policy, we first discretize the environment. 
+To design the policy we first discretize the environment. 
 In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$  shown in Fig.~\ref{fig:BFSGradient}.
@@
 \vspace{-0.5em}
-\caption{\label{fig:AutoVeryParam}Parameter sweep for a) number of robots, b) different noise values, c) object weight, and d) object shape.
+\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of robots, b) different noise values, c) object weight, and d) object shape.
 Each bar is labelled with the number of trials.

### 2017-03-31 journalSwarmControl/realExperiment.tex
changed order of video numbers
Added rectangle
The newest version of the TASE paper

@@
 
-The kilobots performed five successful runs of object manipulation through the obstacle maze.
+The kilobots performed five successful runs manipulating a hexagonal object through an obstacle maze. Videos of these runs are in Extension 2.
 Each trial used 100 kilobots.
-Videos of each are in the multimedia extension.
-Trials two-five were performed in a row with no failures in between.
+Trials two through five were performed in a row with no failures in between.
 For each trial, fully charged kilobots were placed in the lower left-hand of the workspace, as shown in Fig.~\ref{fig:expSnapShot}.
@@
 Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).
+
+We also tested other object shapes. 
 A circular object completed in 3155 s.    
 A square object completed in 6871 s. 
-Three equilateral triangular objects of varying sizes failed in a total of seven runs.
-Failures occurred when the object was pushed into a corner, requiring torque to be unstuck.
-This paper focuses on force, but not torque control.
-Swarm torque control is the subject of ongoing research in \cite{Shahrokhi2016CASE}.
+A rectangle and three equilateral triangle objects of varying sizes failed in a total of nine runs. 
+Manipulation failures occurred when the object was pushed into a corner, requiring torque to be unstuck.  
+%This paper focuses on force, but not torque control. 
+Swarm torque control is the subject of our ongoing research begun in \cite{Shahrokhi2016CASE}.
 

### 2017-03-31 journalSwarmControl/relatedWork.tex
The newest version of the TASE paper

@@
 \subsection{Global control of microrobots}
-Small robots have been constructed with physical heterogeneity so that they respond differently to a global, broadcast control signal.
-Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};   magnetic structures  with different cross-sections that could be independently steered by \cite{Floyd2011,Diller2013};   \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
+Small robots have been constructed with physical heterogeneity so that they respond differently to a global broadcast control signal.
+Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};   magnetic structures  with different cross-sections that can be independently steered by \cite{Floyd2011,Diller2013};   \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
 \cite{Tottori2012} and \cite{Peyer2013}.
@@
 
-  Most humans are able, with practice, steer a swarm of robots controlled by a global input.
+  Most humans are able to, with practice, steer a swarm of robots controlled by a global input.
 Prior to our paper, no algorithm existed.
@@
 Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
-Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},
- to regroup the swarms.
+Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in\S \ref{sec:VarianceControl}.
 

### 2017-04-25 journalSwarmControl/Introduction.tex
Some of Dylan's comments are affected

@@
 
-This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a
-single journal paper.
-  \cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
+%%%% Move it to the cover letter.
+%This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper.
+ % \cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.
 This paper presents three years of results from new games at SwarmControl.net.
@@
 \begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
-\\
-\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
-\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
-\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
+%\\
+%\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
+%\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
+%\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
 \caption{\label{fig:5experiments}
@@
 \textbf{(c)} Varying noise from 0 to 200\% of control authority
-\textbf{(d)} Controlling the position of 1 to 10 particles
-\textbf{(e)} Comparing 3 control architectures to assemble
-\textbf{(f)} Comparing 3 control architectures to forage.
+%\textbf{(d)} Controlling the position of 1 to 10 particles
+%\textbf{(e)} Comparing 3 control architectures to assemble
+%\textbf{(f)} Comparing 3 control architectures to forage.
 }

### 2017-04-25 journalSwarmControl/expResults.tex
Some of Dylan's comments are affected

@@
 We designed six experiments to investigate human control of large swarms for manipulation tasks.
-Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.
+Screenshots of representative experiment are shown in Fig.~\ref{fig:5experiments}.
 Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
@@
 
-\paragraph{Varying number}
+\subsection{Varying number}
 \begin{figure}
@@
 
-\paragraph{Varying visualization}
+\subsection{Varying visualization}
 \begin{figure}[b!]
@@
 
-\paragraph{Varying noise}
+\subsection{Varying noise}
 %Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
@@
 Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.
+This indicates swarm control is robust to IID noise.
 
@@
 
-\paragraph{Position control}
-This online experiment examined how completion time scales with the number of particles $n$.
+%\paragraph{Position control}
+%This online experiment examined how completion time scales with the number of particles $n$.
 Using a single square obstacle, users arranged $n\in[1,10]$ particles into a specified goal pattern.
@@
 Though the number of particles presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
+%
 
@@
 
-\paragraph{Varying control--Assembly}
-Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
-
-We compared attractive and repulsive control with the global control used for the other experiments.
+%\paragraph{Varying control--Assembly}
+%Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
+%
+%We compared attractive and repulsive control with the global control used for the other experiments.
 The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.\ \cite{Khalil2013} and biological models, e.g.
@@
 In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
-The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.
-
+%The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.
+%
 %\todo{describe controller model here, use an equation}
 
-The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
+%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
 The median time using repulsive control was four times longer than with attractive control.
-Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
-
+%Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
+%
 %%\begin{figure}
@@
 
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.45\columnwidth}
-\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
-\vspace{-.5em}
-\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type.
+%\begin{figure}[b!]
+%\renewcommand{\figwid}{0.45\columnwidth}
+%\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
+%\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
+%\vspace{-.5em}
+%\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type.
 Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid.
 Right: in the foraging test, global control resulted in the shortest completion time and attractive the longest.
-%\vspace{-1em}
-}
-\end{figure}
-
-\paragraph{Varying control--Foraging}
-Collecting and delivering resources is necessary for drug delivery.
-This experiment also compared attractive and repulsive control with the global control used for the other experiments.
+%%\vspace{-1em}
+%}
+%\end{figure}
+%
+%\paragraph{Varying control--Foraging}
+%Collecting and delivering resources is necessary for drug delivery.
+%This experiment also compared attractive and repulsive control with the global control used for the other experiments.
 The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region.
 The particles encapsulate the particles on contact.
-
-%\todo{describe controller model here, use an equation}
-
-The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.
+%
+%%\todo{describe controller model here, use an equation}
+%
+%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.
 Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $2.96\times10^{-6}$.
+%
+%

### 2017-04-25 journalSwarmControl/experiment.tex
Some of Dylan's comments are affected

@@
 To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
-  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}: a function that assigns a control input to every possible state.
+  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}.
    At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
 
- A corresponding reward function gives a high reward to the goal state,  
- a large negative reward to states including obstacles, and a small negative reward to all other states.
+% A corresponding reward function gives a high reward to the goal state,  
+% a large negative reward to states including obstacles, and a small negative reward to all other states.
 The reward function $r(x,\mathbf{u})$ is defined as
@@
 \end{align}
- $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
+ where $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
   Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$.
-The optimal policy is the control action that, in probability, results in the largest value:
-   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
+The optimal policy is %the control action that, in probability, results in the largest value:
+   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j|x, \mathbf{u})].  \label{eq:OptimalPolicy}
    \end{align}
@@
 \text{for }&\text{$i=1$ to $N$ do} \nonumber \\
-&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration} \nonumber\\
+&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
 \text{end}& 
@@
 \begin{align}
-F_{att} &= -\zeta \Delta \rho / \rho \\
+F_{att} &= -\zeta \Delta \rho / \rho, \\
 %\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
-\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
+\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right), \\
 %\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
@@
 \begin{array}{ll}
-      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\&~\phi <  \theta \\
+      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\wedge~\phi <  \theta \\
       0, & \textrm{ otherwise} \\
-\end{array} 
+\end{array},
 \right.\\
-F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}
+F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}.
 \end{align}
@@
 The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
-The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, often because workspace obstacles made some particles unable to reach the object.
+The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them.
 This failure rate increases if  object weight increases or ground-robot friction increases.

### 2017-04-25 journalSwarmControl/relatedWork.tex
Some of Dylan's comments are affected

@@
 
- \paragraph{Control---global inputs:}
+ \paragraph{Control---global inputs}
  Many micro- and nanorobotic systems, see \cite{Tottori2012,Shirai2005,Chiang2011,Donald2006,Donald2008,Takahashi2006,Floyd2011,Diller2013,Frutiger2008,Peyer2013}
@@
 
- \paragraph{Sensing---large populations:}
+ \paragraph{Sensing---large populations}
 $n$ differential-drive robots in a 2D workspace require $3n$ state variables.
@@
 
-\paragraph{Computation---calculating the control law:}
+\paragraph{Computation---calculating the control law}
 In our previous work the controllers required at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
@@
 Prior to our paper, no algorithm existed.
-Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
+Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.
+This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
 
- \cite{ayanian2014controlling} designed a user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism.
+ A user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism is designed in \cite{ayanian2014controlling}.
 Our paper shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.
 
-\cite{Jr2004} studied human \emph{fanout}, the number of robots a single human user could directly control.  %Studied as a key aspect of human-robot interaction, 
+Human \emph{fanout}, the number of robots a single human user could directly control is studied in \cite{Jr2004}.  %Studied as a key aspect of human-robot interaction, 
 They postulated that the optimal number of robots was approximately  the autonomous time  divided by the interaction time required by each robot.
@@
 Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
-Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in\S \ref{sec:VarianceControl}.
+Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in \S \ref{sec:VarianceControl}.
 

### 2017-04-25 journalSwarmControl/theory.tex
Some of Dylan's comments are affected

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
+\section{Global Control Laws for a holonomic swarm}
 \label{sec:theory}
@@
 %% A Hybrid controller using hysteresis
-Consider holonomic robots that move in the 2D plane. We want to control position and velocity of the robots. 
+
+Emboldened by the three lessons from our online experiments, this section presents automatic controllers for large numbers of particles that only rely on the first two moments of the swarm position distribution.
+
+We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
 First, assume a noiseless system containing one robot with mass $m$.
@@
 \begin{align}\label{eq:stdform}
-\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t)
+\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t).
 \end{align}
@@
  u_x\\
- u_y\end{bmatrix} 
+ u_y\end{bmatrix}.
 \end{equation}
@@
 
-\subsection{Independent control of many robots is impossible}
-A single robot is fully controllable.
+\subsection{Independent control of many particles is impossible}
+In this model, a single particle is fully controllable.
 For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis.
-Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
+Given $n$ particles to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
 Without loss of generality, assume $m=1$.
@@
 1
-\end{bmatrix} u_x
+\end{bmatrix} u_x.
 \end{equation}
- However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
+ However, just as with one particle, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
 \begin{equation}
@@
 \subsection{Controlling the mean position}\label{sec:controlMeanPosition}
-This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis.
+This means \emph{any} number of particles controlled by a global command have just two controllable states in each axis.
 We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.
@@
 To find a potentially more useful option,
-We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
+We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ particles:
 \begin{align}
@@
 1
-\end{bmatrix} u_x
+\end{bmatrix} u_x.
 \end{align}
@@
 1
-\end{bmatrix} u_x
+\end{bmatrix} u_x.
 \end{equation}
@@
 \begin{align}
-u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}) \nonumber\\
-u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}})  \label{eq:PDcontrolPosition}
+u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}), \nonumber\\
+u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}}).  \label{eq:PDcontrolPosition}
 \end{align}
@@
 \dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon},
-\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t
+\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t.
 \end{equation}
@@
 
-We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
+We will prove the goal is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
 \begin{align}
 \label{eq:LyapunovVariance}
-V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2\nonumber\\
-\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x})
+V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2,\nonumber\\
+\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x}).
 \end{align}
@@
 \begin{align}
-u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}) \nonumber\\
+u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}), \nonumber\\
 u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
@@
 \begin{align}
-x_{goal}(\sigma^2_{ref}) = \ell_x/2 = r + \sqrt{3\sigma^2_{ref}}
+x_{goal}(\sigma^2_{ref}) = \ell_x/2 = r + \sqrt{3\sigma^2_{ref}}.
 \end{align}
@@
 \begin{align} \label{eq:sigMaxMin}
- \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\
+ \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r), \nonumber\\
   \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).

### 2017-04-26 journalSwarmControl/Introduction.tex
Edited the paper

@@
 There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
-This paper presents a tool for investigating HSI methods through statistically significant numbers of experiments.
+One contribution of this paper is a tool for investigating HSI methods through statistically significant numbers of experiments.
 
@@
 \end{figure}
-
-\begin{figure}
-%\centering
-\renewcommand{\figwid}{0.3\columnwidth}
-\begin{overpic}[width =\figwid]{VaryNum.pdf}	\put(10,80){\textbf{a} }\end{overpic}~
-\begin{overpic}[width =\figwid]{VaryVisMV.pdf}	\put(10,80){\textbf{b} }\end{overpic}~
-\begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
-%\\
-%\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
-%\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
-%\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
-\caption{\label{fig:5experiments}
-Screenshots from our online experiments controlling multi-particle systems with limited, global control.
-\textbf{(a)} Varying the number of particles from 1-500
-\textbf{(b)} Comparing 4 levels of visual feedback 
-\textbf{(c)} Varying noise from 0 to 200\% of control authority
-%\textbf{(d)} Controlling the position of 1 to 10 particles
-%\textbf{(e)} Comparing 3 control architectures to assemble
-%\textbf{(f)} Comparing 3 control architectures to forage.
-}
-\end{figure}
-
-%\begin{figure}
-%\renewcommand{\figwid}{0.32\columnwidth}
-%\subfloat[][Vary Number]{\label{fig:VaryNum}
-%\begin{overpic}[width =\figwid]{VaryNum.pdf}\end{overpic}}
-%%
-%\subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
-%%\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
-%\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\end{overpic}}
-%%
-%\subfloat[][Vary Noise]{\label{fig:VaryNoise}
-%\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}\\
-%%
-%\subfloat[][Control Position]{\label{fig:ControlPos}
-%\begin{overpic}[width =\figwid]{ControlPos.pdf}\end{overpic}}
-%%
-%\subfloat[][Vary Control: Assembly]{\label{fig:VaryControl}
-%\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
-%%
-%\subfloat[][Vary Control: Foraging]{\label{fig:Forage}
-%\begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
-%%
-%\caption{\label{fig:5experiments}
-%Screenshots from our online experiments controlling multi-robot systems with limited, global control.
-%\textbf{(a)} Varying the number of robots from 1-500
-%\textbf{(b)} Comparing 4 levels of visual feedback 
-%\textbf{(c)} Varying noise from 0 to 200\% of control authority
-%\textbf{(d)} Controlling the position of 1 to 10 robots
-%\textbf{(e)} Comparing 3 control architectures to assemble
-%\textbf{(f)} Comparing 3 control architectures to forage.
-%\vspace{-2em}
-%}
-%\end{figure}
-
-The goal was to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size.
-Towards this end, we have created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}: an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research.
-Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.
-\href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are online at \cite{Chris-Ertel2016}.
-
-Our online experiments show that numerous particles responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.
 

### 2017-04-26 journalSwarmControl/conclusion.tex
Edited the paper

@@
 
-    Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
-Their small size makes individual control and autonomy challenging, so currently these robots are steered by global control inputs such as magnetic fields or chemical gradients.
+    %Micro- and nanorobotics have the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  
+  The small size of micro and nano particles makes individual control and autonomy challenging, so currently these particles are steered by global control inputs such as magnetic fields or chemical gradients.
 To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
@@
 
-Inspired by human operators, this paper designed controllers and controllability results using only the mean and variance of a robot swarm. 
+Inspired by the three lessons from swarmcontrol.net, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
+We developed a hysteresis based controller to regulate the position and variance of a swarm.
+We designed a controller for object manipulation using policy iteration for path planning, regions for outlier rejection and potential fields for minimizing moving the object backwards.
 These controllers were used as primitives for an object manipulation task.
@@
 
-  Manipulation by large populations of robots has many open questions.
+ % Manipulation by large populations of robots has many open questions.
 We invite other collaborators to submit their own experiments for large-scale trials to \href{http://www.swarmcontrol.net}{SwarmControl.net}.
+
+Our main goal is to perform assembly using particle swarms to manipulate and attach components.
+This task requires controlling the position and orientation of components, manipulate them through obstacles and other components, and apply force and torque to components.
+This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.
+
 Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, optimal-control, and targeted drug delivery in a vascular network.

### 2017-04-26 journalSwarmControl/experiment.tex
affected more comments

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Object manipulation results}\label{sec:exp}
+\section{Particle Swarm Object manipulation}\label{sec:exp}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   The swarm must deliver the object to the goal region.
-To solve this object manipulation task we divide the task into two components: designing a policy for the object, and designing a control law for the swarm to push the object according to the policy.
-
+To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.
 \subsection{Learning a policy for the object}\label{subsec:objectpolicy}
@@
 To design the policy we first discretize the environment. 
-In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$  shown in Fig.~\ref{fig:BFSGradient}.
-The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$, so we define the desired direction for the object as $\mathbf{D}(\mathbf{b}) = \nabla \mathbf{M}(\mathbf{b})$. 
-The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
-This approach using workspace BFS fails to account for the hull of the object and will suggest moves that can cause collisions with the workspace.
+In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid,% to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$  shown in Fig.~\ref{fig:BFSGradient}.
+%The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$, so we define the desired direction for the object as $\mathbf{D}(\mathbf{b}) = \nabla \mathbf{M}(\mathbf{b})$. 
+%The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
+but using workspace BFS fails to account for the hull of the object and will suggest moves that can cause collisions with the workspace.
 A configuration-space BFS approach avoids that problem but still fails to model uncertain actuation of the object by the swarm.
@@
 
-Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
+When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
 To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
@@
 
-\subsection{Algorithm for object manipulation}\label{sec:AlgObjectManipulation}
+\subsection{Simulation results}\label{sec:AlgObjectManipulation}
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
@@
 
-\subsection{Automated object manipulation (simulation)}
+%\subsection{Automated object manipulation (simulation)}
 Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation. 

### 2017-04-27 journalSwarmControl/Introduction.tex
All the comments are solved.

@@
 \caption{\label{fig:bigPictureMeanAndVarianceForSwarm} 
-A swarm of particles, all controlled by a uniform control input, can be effectively controlled by a hybrid controller that knows only the mean and variance of the robot distribution.  
+A swarm of particles, all actuated by a uniform control input where each particle gets the same control input, can be effectively manipulated by a control law that uses only the mean and variance of the robot distribution.
 Here a swarm of particles (kilobot robots) pushes a green hexagon toward the goal (see video attachment).
@@
 After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in \S \ref{sec:expMethods}.
-Inspired by these results, we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
+Next we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
 We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.

### 2017-04-27 journalSwarmControl/conclusion.tex
All the comments are solved.

@@
 To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
-Analysis of the game play results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback.
+Analysis of the game play results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.
 
@@
 We developed a hysteresis based controller to regulate the position and variance of a swarm.
-We designed a controller for object manipulation using policy iteration for path planning, regions for outlier rejection and potential fields for minimizing moving the object backwards.
-These controllers were used as primitives for an object manipulation task.
+We designed a controller for object manipulation using policy iteration for path planning, regions for outlier rejection, and potential fields for minimizing moving the object backwards.
 All automatic controllers were implemented using 100 kilobots steered by the direction of a global light source.
@@
 
-Our main goal is to perform assembly using particle swarms to manipulate and attach components.
-This task requires controlling the position and orientation of components, manipulate them through obstacles and other components, and apply force and torque to components.
+Our future goal is to perform assembly using particle swarms to manipulate and attach components.
+This task requires controlling the position and orientation of components, manipulating them through obstacles and other components, and applying force and torque to components.
 This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.

### 2017-04-27 journalSwarmControl/expResults.tex
All the comments are solved.

@@
 Screenshots of representative experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
+Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise.%, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
  The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.
@@
 We hypothesized participants would complete the task faster with more particles.
-The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 particles.
+The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 particles, but only a gradual increase in completion time from 50 to 500.
 
@@
 A single particle left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
+Similarly, our algorithm must be robust to outliers.
 %obscuring what the rest of the swarm is doing.
+
+\begin{figure}[b!]
+\renewcommand{\figwid}{0.49\columnwidth}
+\begin{overpic}[width =\figwid]{ResVaryNoise.pdf}\end{overpic}
+\begin{overpic}[width =\figwid]{ResPositioning.pdf}\end{overpic}
+%\vspace{-1em}
+\caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
+Right: For position control, increasing the number of particles resulted in longer completion times.
+For more than 4 particles the goal pattern contained a void, which may have caused the jump in completion times.
+%\vspace{-1em}
+}
+\end{figure}
 
@@
 This indicates swarm control is robust to IID noise.
-
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.49\columnwidth}
-\begin{overpic}[width =\figwid]{ResVaryNoise.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{ResPositioning.pdf}\end{overpic}
-%\vspace{-1em}
-\caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
-Right: For position control, increasing the number of particles resulted in longer completion times.
-For more than 4 particles the goal pattern contained a void, which may have caused the jump in completion times.
-%\vspace{-1em}
-}
-\end{figure}
 

### 2017-04-27 journalSwarmControl/experiment.tex
All the comments are solved.

@@
 
-This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers.%, inspired by the analysis of human users in \S \ref{sec:expResults}.  
+This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers.
 The swarm must deliver the object to the goal region.
@@
 To design the policy we first discretize the environment. 
-In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid,% to determine $\mathbf{M}_{BFS}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$  shown in Fig.~\ref{fig:BFSGradient}.
-%The object's center of mass is at $\mathbf{b}$ and has average radius $r_b$, so we define the desired direction for the object as $\mathbf{D}(\mathbf{b}) = \nabla \mathbf{M}(\mathbf{b})$. 
-%The robots were directed behind the object at  $\mathbf{b} - 1.5 r_b \mathbf{D}(\mathbf{b})$, then directed to  $\mathbf{b} - 0.1 r_b \mathbf{D}(\mathbf{b})$ to push the object toward the goal location.
-but using workspace BFS fails to account for the hull of the object and will suggest moves that can cause collisions with the workspace.
+In \cite{ShahrokhiIROS2015}, we used breadth-first search (BFS) on this discretized grid, but using workspace BFS fails to account for the hull of the object and will suggest moves that can cause collisions with the workspace.
 A configuration-space BFS approach avoids that problem but still fails to model uncertain actuation of the object by the swarm.
@@
      +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
-      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
+      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
       -1, & \textrm{otherwise}\\
@@
 \end{align}
- where $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
+ where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
   Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$.
@@
 $\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 robots, pushing the object to goal using BFS required 183$\pm$179 s while Policy Iteration required 90$\pm$35 s (mean$\pm$std).
+In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while Policy Iteration required 90$\pm$35 s (mean$\pm$std).
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 
-In 10 simulations with 100 robots, pushing the object to goal without a repulsive potential field failed in two of twelve runs.
+In 10 simulations with 100 particles, pushing the object to goal without a repulsive potential field failed in two of twelve runs.
 No failures occurred with the repulsive potential field.
@@
 This failure rate increases if  object weight increases or ground-robot friction increases.
-The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace.
-Robots that cannot reach the object due to obstacles skew these calculations.
-The state machine in Fig.\ \ref{fig:Region}a solves this problem by creating two states for the maze: either main or transfer.
+The mean and covariance calculations \eqref{eq:meanVar} included all particles in the workspace.
+Particles that cannot reach the object due to obstacles skew these calculations.
+The state machine in Fig.
+\ref{fig:Region}a solves this problem by creating two states for the maze: either main or transfer.
 Each state has a set of regions representing a discretized visibility polygon.
 Whenever the object crosses a region boundary the state toggles.
-The main regions are generated by extending obstacles until they meet another obstacle.
-The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.
+The \emph{main} regions are generated by extending obstacles until they meet another obstacle.
+The \emph{transfer} regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.
 
@@
 
-In 10 simulations with 100 robots, completion time without outlier rejection required 245$\pm$135 s while using outlier rejection required 90$\pm$35 s (mean$\pm$std).
+In 10 simulations with 100 robots, completion time without outlier rejection required 271$\pm$267 s while using outlier rejection required 90$\pm$35 s (mean$\pm$std).
 %\todo{report results of 10 simulations with and without regions.  Give mean and std}
@@
  We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
- The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$.
+ The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than the object with an obstacle-free straight-line path to the object.
 The control algorithm  for object manipulation is listed in Alg.~\ref{alg:BlockPushing}.
@@
 \State Compute optimal policy for object, according to \S \ref{subsec:objectpolicy}
-\While{$\mathbf{b}$ is not in goal region}
+\While{is not in goal region}
 \State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$

### 2017-04-27 journalSwarmControl/realExperiment.tex
Changed the swarm control Journal, added code.
All the comments are solved.

@@
 
-The kilobots performed five successful runs manipulating a hexagonal object through an obstacle maze. Videos of these runs are in Extension 2.
+The kilobots performed five successful runs manipulating a hexagonal object through an obstacle maze.
+Videos of these runs are in Extension 2.
+These hardware experiments represent the results of over 100 hours of trials.
 Each trial used 100 kilobots.
@@
 The moveable object was placed in the lower center of the workspace.
-{\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central by \cite{Shahrokhi2015MDP}.
+{\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
 Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

### 2017-04-27 journalSwarmControl/relatedWork.tex
All the comments are solved.

@@
 Even holonomic robots require $2n$ state variables.
-Numerous methods exist for measuring this state in microrobotics.
+Numerous methods exist for measuring this state in microrobotics \cite{Peyer2013,Chiang2011,martel2014computer}.
 These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.
-These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.
+These challenges increase at small scales where sensing competes with control for communication bandwidth.
 We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.
@@
 
-Several user studies compare methods for controlling large swarms of simulated robots, for example \cite{bashyal2008human,kolling2012towards,de2012controllability}.
+Several user studies compare methods for controlling large swarms of simulated robots, for example \cite{bashyal2008human,de2012controllability,kolling2012towards}.
 These studies provide insights but are limited by cost to small user studies; have a closed-source code base; and focus on controlling intelligent, programmable agents.

### 2017-04-27 journalSwarmControl/theory.tex
affected more comments
All the comments are solved.

@@
 We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
-\begin{equation}
-\mathcal{C} = [ B, AB, A^2B, ... , A^{n-1}B ].
-\end{equation}
-
-\begin{equation}
+\begin{align}
+\mathcal{C} &= [ B, AB, A^2B, ... , A^{n-1}B ].\\
 \textrm{Here }
-\mathcal{C}=\left[
+\mathcal{C}&=\left[
 \begin{matrix} 
@@
 \end{matrix}
- \right],
- \quad \textrm{rank}(\mathcal{C})=4,
-\end{equation}
+ \right],\\
+ \quad \textrm{rank}(\mathcal{C})&=4,
+\end{align}
 and thus all four states are controllable.
@@
 
-Due to symmetry of the control input, only the mean position and mean velocity are controllable.
-However, there are several techniques for breaking symmetry, for example by using obstacles as in \cite{Becker2013b}, or by allowing independent noise sources as in \cite{beckerIJRR2014}.
+%Due to symmetry of the control input, only the mean position and mean velocity are controllable. However, 
+There are several techniques for breaking the symmetry of the control control input to allow controlling more states, for example by using obstacles as in \cite{Becker2013b}, or by allowing independent noise sources as in \cite{beckerIJRR2014}.
 
@@
 where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$.
-Given a large obstacle-free workspace, a \emph{Brownian noise} process increases the variance linearly with time.
+Given a large obstacle-free workspace with $\mathbf{u}(t)= 0$, a \emph{Brownian noise} process increases the variance linearly with time.
 \begin{equation}
-\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon},
+\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t))  = W \bm{\varepsilon},
 \quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t.
@@
 
-A  controller to regulate the variance to $\sigma^2_{ref}$ is:
+A PID controller to regulate the variance to $\sigma^2_{ref}$ is:
 \begin{align}
@@
 \Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
-\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ % \Comment{initially not in variance control mode}
 \State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
@@
 \State $x_{goal}  \gets x_{min}$
-\State $flag_x  \gets \TRUE$
-\Else {\bf~if} {$flag_x$ \textbf{and} $\sigma_x^2 < \sigma_{min}^2$}
+\ElsIf { $\sigma_x^2 < \sigma_{min}^2$}
 \State $x_{goal}  \gets  x_{target}$
-\State $flag_x  \gets  \FALSE$
 \EndIf
@@
 \State $y_{goal}  \gets y_{min}$
-\State $flag_y  \gets \TRUE$
-\Else {\bf~if} {$flag_y$ \textbf{and} $\sigma_y^2 < \sigma_{min}^2$}
+\ElsIf { $\sigma_y^2 < \sigma_{min}^2$}
 \State $y_{goal}  \gets  y_{target}$
-\State $flag_y  \gets \FALSE$
 \EndIf

### 2017-05-02 journalSwarmControl/experiment.tex
Changed the notations and edited

@@
 To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.
+
+The table below summarizes the results:
+
+\begin{center}
+\begin{tabular}{ c c }
+\hline
+Method & Result (s)\\ [0.5ex] 
+\hline \hline
+Value Iteration (VI) & 367 $\pm$ 253 \\ 
+\hline
+VI + Potential Fields (PF) & 271 $\pm$ 267 \\  
+\hline
+VI + Outlier Rejection (OR) & 245 $\pm$ 135\\
+\hline
+BFS + PF + OR & 183 $\pm$ 179 \\
+\hline
+VI + PF + OR & 90 $\pm$ 35 \\[0.5ex]
+\hline
+\end{tabular}
+\end{center}
 \subsection{Learning a policy for the object}\label{subsec:objectpolicy}
@@
 
-$\mathbf{M}_{BFS}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
+$\mathbf{M}_{\rm{BFS}}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
 In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while Policy Iteration required 90$\pm$35 s (mean$\pm$std).
@@
 \begin{align}
-F_{att} &= -\zeta \Delta \rho / \rho, \\
+F_{\rm{att}} &= -\zeta \Delta \rho / \rho, \\
 %\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
@@
 %\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
- F_{rep} &=  \left\{
+ F_{\rm{rep}} &=  \left\{
 \begin{array}{ll}
@@
 \right.\\
-F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}.
+F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}.
 \end{align}
@@
 \State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
-\If {$\sigma^2 > \sigma_{max}^2$}
-\While{$\sigma^2 > \sigma_{min}^2$}
-\State $ [x_{goal}, y_{goal}] \gets $ nearest corner in region
-\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
+\If {$\sigma^2 > \sigma_{\rm{max}}^2$}
+\While{$\sigma^2 > \sigma_{\rm{min}}^2$}
+\State $ [x_{\rm{goal}}, y_{\rm{goal}}] \gets $ nearest corner in region
+\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{\rm{goal}}, y_{\rm{goal}}]$
 \EndWhile

### 2017-05-02 journalSwarmControl/theory.tex
Changed the notations and edited

@@
 \begin{align}
-u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}), \nonumber\\
-u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}}).  \label{eq:PDcontrolPosition}
+u_x &= K_{p}(x_{\rm{goal}} - \bar{x}) + K_{d}(0-\dot{\bar{x}}), \nonumber\\
+u_y &= K_{p}(y_{\rm{goal}}  - \bar{y}) + K_{d}(0-\dot{\bar{y}}).  \label{eq:PDcontrolPosition}
 \end{align}
@@
  \begin{align} \label{eq:optimalvar}
- \sigma^2_{optimal}(n,r) \approx   \frac{\sqrt{3}}{\pi} n r^2\approx 0.55 n r^2.
+ \sigma^2_{\rm{optimal}}(n,r) \approx   \frac{\sqrt{3}}{\pi} n r^2\approx 0.55 n r^2.
  \end{align} 
@@
 \label{eq:LyapunovVariance}
-V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2,\nonumber\\
-\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x}).
+V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{\rm{goal}})^2,\nonumber\\
+\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{\rm{goal}})\dot{\sigma}^2(\mathbf{x}).
 \end{align}
-We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{goal}$.
+We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{\rm{goal}}$.
 To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
@@
 u(t) &=   \begin{cases}
-	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{goal} \\ 
-	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal}.
+	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{\rm{goal}} \\ 
+	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{\rm{goal}}.
 \end{cases} 
@@
 \dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
-	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{goal}, \sigma^2_{optimal}(n,r))  \\ 
-	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal},
+	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{\rm{goal}}, \sigma^2_{\rm{optimal}}(n,r))  \\ 
+	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{\rm{goal}},
 \end{cases} 
@@
 
-A PID controller to regulate the variance to $\sigma^2_{ref}$ is:
+A PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}
-u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}), \nonumber\\
-u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
+u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}), \nonumber\\
+u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
 \end{align}
@@
 \begin{align}
-x_{goal}(\sigma^2_{ref}) = \ell_x/2 = r + \sqrt{3\sigma^2_{ref}}.
+x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
 \end{align}
- If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.
+ If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.
 
@@
 The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is less than the maximum controlled speed, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.
-Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{max}^2$.
- Variance is reduced until less than $\sigma_{min}^2$, then control again regulates the mean position. 
+Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{\rm{max}}^2$.
+ Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
  This technique satisfies control objectives that evolve at different rates as in~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
@@
 \begin{algorithmic}[1]
-\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
-\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
+\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{\rm{target}},y_{\rm{target}}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
+\State $x_{goal} \gets  x_{\rm{target}}$, $y_{\rm{goal}} \gets y_{\rm{target}}$
 \Loop
@@
 
-\If {$\sigma_x^2 > \sigma_{max}^2$}
-\State $x_{goal}  \gets x_{min}$
-\ElsIf { $\sigma_x^2 < \sigma_{min}^2$}
-\State $x_{goal}  \gets  x_{target}$
+\If {$\sigma_x^2 > \sigma_{\rm{max}}^2$}
+\State $x_{\rm{goal}}  \gets x_{\rm{min}}$
+\ElsIf { $\sigma_x^2 < \sigma_{\rm{min}}^2$}
+\State $x_{\rm{goal}}  \gets  x_{\rm{target}}$
 \EndIf
 
-\If{$\sigma_y^2 > \sigma_{max}^2$}
-\State $y_{goal}  \gets y_{min}$
-\ElsIf { $\sigma_y^2 < \sigma_{min}^2$}
-\State $y_{goal}  \gets  y_{target}$
+\If{$\sigma_y^2 > \sigma_{\rm{max}}^2$}
+\State $y_{\rm{goal}}  \gets y_{\rm{min}}$
+\ElsIf { $\sigma_y^2 < \sigma_{\rm{min}}^2$}
+\State $y_{\rm{goal}}  \gets  y_{\rm{target}}$
 \EndIf
-\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
+\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{\rm{goal}}, y_{\rm{goal}}]$
 \EndLoop
@@
 \begin{overpic}[width = 0.8\columnwidth]{hysteresis.pdf}
-\put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
-\put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
+\put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{\rm{min}}$ }
+\put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{\rm{max}}$}\end{overpic}
 %\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
@@
 
-A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
+A key challenge is to select proper values for $\sigma_{\rm{min}}^2$ and $\sigma_{\rm{max}}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
 The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values: 
@@
 \begin{align} \label{eq:sigMaxMin}
- \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r), \nonumber\\
-  \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).
+ \sigma^2_{\rm{min}} &= 2.5r+ \sigma^2_{\rm{optimal}}(n,r), \nonumber\\
+  \sigma^2_{\rm{max}} &= 15r+ \sigma^2_{\rm{optimal}}(n,r).
   \end{align}

### 2017-05-02 journalSwarmControl/theory.tex
Edited some texts

@@
 \begin{algorithmic}[1]
-\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{\rm{target}},y_{\rm{target}}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
+\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{\rm{min}}, x_{\rm{max}}, y_{\rm{min}}, y_{\rm{max}}\}$, and the target mean position $[x_{\rm{target}},y_{\rm{target}}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
 \State $x_{goal} \gets  x_{\rm{target}}$, $y_{\rm{goal}} \gets y_{\rm{target}}$

### 2017-05-02 journalSwarmControl/coverletter/EditorNoteTROparticleSwarmManipulation.tex
added cover letter and other edits to the paper (mean = std, previous work by author)

Dear Editors,
 We are submitting our manuscript “Steering a Particle Swarm Using Global Inputs and Swarm Statistics”, to be considered for publishing in IEEE Transactions on Robotics (TRO). Authors are Shiva Shahrokhi, Lillian Lin, Chris Ertel, Mable Wan, and Aaron T. Becker.

This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper. These papers are in IEEE International Conference on Robotics and Automation (ICRA) 2014 and IEEE International Conference on Intelligent Robots and Systems (IROS) 2015. The first conference paper covered the first three months of SwarmControl.net experiments [1], and the second one presented simulations of object manipulation [2]. This paper presents three years of results from SwarmControl.net. For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parameter sweeps over key variables. All hardware validation experiments are new, and feature a set of 100+ kilobots programmed to move toward a bright light source. Changing the location of the light source steers the kilobot swarm and uses them as a compliant manipulator.

We believe that the topic is a perfect fit for the scope of your journal and are hopeful that editors and reviewers share the opinion that this submission will be of interest to the readers of IEEE Transactions on Robotics.

With Best Regards,
Aaron T. Becker (on behalf of all authors)
713 743 6671, atbecker@uh.edu.
Assistant Professor, Electrical and Computer Engineering, University of Houston

[1] A. T. Becker, C. Ertel, and J. McLurkin, “Crowdsourcing swarm manipulation experiments: A massive online user study with large swarms of simple robots,” in IEEE International Conference on Robotics and Automation (ICRA), 2014, pp. 2825–2830.
 [2] S. Shahrokhi and A. T. Becker, “Stochastic swarm control with global inputs,” in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Sep. 2015, pp. 421–427.

### 2017-05-02 journalSwarmControl/relatedWork.tex
added cover letter and other edits to the paper (mean = std, previous work by author)

@@
 
+\subsection{Relationship to authors' prior work}
+ This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper.
+One paper covered the first three months of SwarmControl.net experiments~\cite{swarmcontrol2013}, and the second presented simulations of object manipulation~\cite{ShahrokhiIROS2015}.
+This paper presents three years of results from SwarmControl.net.
+For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parameter sweeps over key variables.
+All hardware validation experiments are new.
+
 %

### 2017-05-02 journalSwarmControl/simulation.tex
Some of Dylan's comments are affected
added cover letter and other edits to the paper (mean = std, previous work by author)

@@
 Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation, presented in \cite{catto2010box2d}.
-All experiments ran on a Chrome web browser on a 2.6 GHz Macbook.
+All experiments in this section ran on a Chrome web browser on a 2.6 GHz Macbook.
 \href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
@@
 \subsection{Controlling the mean position}
-
-We performed a parameter sweep using the PD controller \eqref{eq:PDcontrolPosition} to identify the best control gains.
-Representative experiments are shown in Fig.~\ref{fig:gainvalues}.
-100 particles were used and the maximum speed was 3 meters per second.
-As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$.
 \begin{figure}
@@
 \end{figure}
+We performed a parameter sweep using the PD controller \eqref{eq:PDcontrolPosition} to identify the best control gains.
+Representative experiments are shown in Fig.~\ref{fig:gainvalues}.
+100 particles were used and the maximum speed was 3 meters per second.
+As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$.
 

### 2017-05-05 journalSwarmControl/conclusion.tex
Thanked reviewers

@@
 
-Inspired by the three lessons from swarmcontrol.net, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
+Inspired by the three lessons from \href{www.swarmcontrol.net}{swarmcontrol.net}, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
 We developed a hysteresis based controller to regulate the position and variance of a swarm.

### 2017-05-08 journalSwarmControl/Reviews/responsetoIJRRreview2017-03-06.tex
added 'balance' option

1. Change the names of papers in the beginning of the paper.  This helps editor select better reviewers.
2. check that we cite the papers

06-Mar-2017

Dear Ms. Shahrokhi:

I write you in regards to manuscript # IJR-16-2710 entitled "Steering a Swarm Using Global Inputs and Swarm Statistics" which you submitted to the International Journal of Robotics Research.

In view of the criticisms of the reviewer(s) found at the bottom of this letter, your manuscript has been denied publication in the International Journal of Robotics Research.

Thank you for considering the International Journal of Robotics Research for the publication of your research.  I hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.

Sincerely,
Prof. John Hollerbach
Editor, International Journal of Robotics Research
Ijrr.admin@sagepub.co.uk

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author
The control of robot swarms using global input, and the underlying requirements in terms of human-swarm interactions, is an important area of research. The results presented here are exciting and impressive.

However, the theory and all figures up until Figure 12 were taken directly from, or are closely based on the following papers by the authors:

“Crowdsourcing Swarm Manipulation Experiments: A Massive Online User Study with Large Swarms of Simple Robots”

“Stochastic Swarm Control with Global Inputs”

This leaves only 3 pages (10-12) with original unpublished content and therefore does not warrant a full journal publication. The paper should be rewritten to focus on the original material.

Reviewer: 2

Comments to the Author
This paper describes controllers for steering a swarm using global inputs for object transportation. The paper is well written and its contribution is important. In particular, experimentally demonstrating that control of a large swarm is indeed possible using simple global control input has a number of applications in micro and nano robotics for drug delivery.

Unfortunately, there are two major problems with the paper, which have to be addressed prior to publication. First, the control of a robot swarm using simple global inputs and the resulting distributions from biased random walk are actually already quite well understood and described in the references below. The novelty of this paper lies therefore in leveraging these techniques for object transport. I recommend the authors to ground their theoretical models and controllers - which might have indeed advantages - as much as possible in the above literature. Second, although I really, really like the human swarm interaction experiments, their connection to the robot controller and using them as a general method for design inspiration is questionable from a scientific perspective. It is akin to let humans estimate the square root of a number and then extract heuristics for doing so using a computer. What might be hard for people could be very easy for a computer and the other way round. Reducing the dimensionality of the problem is actually the first thing a control designer would do, possibly in a way that can be mathematically justified. Such an analysis might then reveal that using the median would be even better than using the mean as it is more robust to outliers. The approach is also at odds with the principled method provided by Milutinovic (see references below) and should probably not be in the same paper as any formal method as long as there cannot be made a clear connection between the two. My suggestion to resolve this problem would be to motivate your choice of information reduction options using the existing literature, and then focus exclusively on the object transportation task. Another option could be to use the simulation results to tune parameters of the controller.

REFERENCES

Milutinovic, Dejan, and Pedro Lima. "Modeling and optimal centralized control of a large-size robotic population." IEEE Transactions on Robotics 22.6 (2006): 1280-1285.

Bretl, Timothy. "Control of Many Agents Using Few Instructions." Robotics: Science and Systems. 2007.

Palmer, Aaron, and Dejan Milutinovi?. "A hamiltonian approach using partial differential equations for open-loop stochastic optimal control." American Control Conference (ACC), 2011. IEEE, 2011.

Prorok, Amanda, Nikolaus Correll, and Alcherio Martinoli. "Multi-level spatial modeling for stochastic distributed robotic systems." International Journal of Robotics Research 30.5 (2011): 574-589.

Demir, Nazlı, and Behçet Açıkme?e. "Probabilistic density control for swarm of decentralized ON-OFF agents with safety constraints." American Control Conference (ACC), 2015. IEEE, 2015.

Associate Editor Comments:

Associate Editor
Comments to the Author:
There is agreement among both reviewers that the work presented in this manuscript is of high quality and of interest to the readership of IJRR. However, the reviewers also agree in their assessment of the manuscript's novelty with respect to prior work by the authors and by others in the community. Reviewer 1 indicates that the novelty beyond the author's prior work is insufficient to warrant a journal publication. Reviewer 2 is enthusiastic about the work but suggests that the novelty must be presented in context of related work by others to be assessed properly. Some of the references provided by reviewer 2 closely relate to theoretical contributions in the manuscript. It would therefore be necessary to clearly explain how the proposed methods extend the state of the art, as represented in those additional references. As the correction of both criticisms will require substantial changes to the manuscript and will probably necessitate the inclusion of additional theoretical and experimental results, the IJRR review process requires that the manuscript be rejected at this time. However, after you have carefully addressed the reviewers' concerns, you are encouraged to submit a new manuscript to IJRR. This new manuscript will be handled independently of this submission.

### 2017-05-08 journalSwarmControl/Reviews/IJRRreview2017-03-06.tex
added 'balance' option

06-Mar-2017

Dear Ms. Shahrokhi:

I write you in regards to manuscript # IJR-16-2710 entitled "Steering a Swarm Using Global Inputs and Swarm Statistics" which you submitted to the International Journal of Robotics Research.

In view of the criticisms of the reviewer(s) found at the bottom of this letter, your manuscript has been denied publication in the International Journal of Robotics Research.

Thank you for considering the International Journal of Robotics Research for the publication of your research.  I hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.

Sincerely,
Prof. John Hollerbach
Editor, International Journal of Robotics Research
Ijrr.admin@sagepub.co.uk

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author
The control of robot swarms using global input, and the underlying requirements in terms of human-swarm interactions, is an important area of research. The results presented here are exciting and impressive.

However, the theory and all figures up until Figure 12 were taken directly from, or are closely based on the following papers by the authors:

“Crowdsourcing Swarm Manipulation Experiments: A Massive Online User Study with Large Swarms of Simple Robots”

“Stochastic Swarm Control with Global Inputs”

This leaves only 3 pages (10-12) with original unpublished content and therefore does not warrant a full journal publication. The paper should be rewritten to focus on the original material.

Reviewer: 2

Comments to the Author
This paper describes controllers for steering a swarm using global inputs for object transportation. The paper is well written and its contribution is important. In particular, experimentally demonstrating that control of a large swarm is indeed possible using simple global control input has a number of applications in micro and nano robotics for drug delivery.

Unfortunately, there are two major problems with the paper, which have to be addressed prior to publication. First, the control of a robot swarm using simple global inputs and the resulting distributions from biased random walk are actually already quite well understood and described in the references below. The novelty of this paper lies therefore in leveraging these techniques for object transport. I recommend the authors to ground their theoretical models and controllers - which might have indeed advantages - as much as possible in the above literature. Second, although I really, really like the human swarm interaction experiments, their connection to the robot controller and using them as a general method for design inspiration is questionable from a scientific perspective. It is akin to let humans estimate the square root of a number and then extract heuristics for doing so using a computer. What might be hard for people could be very easy for a computer and the other way round. Reducing the dimensionality of the problem is actually the first thing a control designer would do, possibly in a way that can be mathematically justified. Such an analysis might then reveal that using the median would be even better than using the mean as it is more robust to outliers. The approach is also at odds with the principled method provided by Milutinovic (see references below) and should probably not be in the same paper as any formal method as long as there cannot be made a clear connection between the two. My suggestion to resolve this problem would be to motivate your choice of information reduction options using the existing literature, and then focus exclusively on the object transportation task. Another option could be to use the simulation results to tune parameters of the controller.

REFERENCES

Milutinovic, Dejan, and Pedro Lima. "Modeling and optimal centralized control of a large-size robotic population." IEEE Transactions on Robotics 22.6 (2006): 1280-1285.

Bretl, Timothy. "Control of Many Agents Using Few Instructions." Robotics: Science and Systems. 2007.

Palmer, Aaron, and Dejan Milutinovi?. "A hamiltonian approach using partial differential equations for open-loop stochastic optimal control." American Control Conference (ACC), 2011. IEEE, 2011.

Prorok, Amanda, Nikolaus Correll, and Alcherio Martinoli. "Multi-level spatial modeling for stochastic distributed robotic systems." International Journal of Robotics Research 30.5 (2011): 574-589.

Demir, Nazlı, and Behçet Açıkme?e. "Probabilistic density control for swarm of decentralized ON-OFF agents with safety constraints." American Control Conference (ACC), 2015. IEEE, 2015.

Associate Editor Comments:

Associate Editor
Comments to the Author:
There is agreement among both reviewers that the work presented in this manuscript is of high quality and of interest to the readership of IJRR. However, the reviewers also agree in their assessment of the manuscript's novelty with respect to prior work by the authors and by others in the community. Reviewer 1 indicates that the novelty beyond the author's prior work is insufficient to warrant a journal publication. Reviewer 2 is enthusiastic about the work but suggests that the novelty must be presented in context of related work by others to be assessed properly. Some of the references provided by reviewer 2 closely relate to theoretical contributions in the manuscript. It would therefore be necessary to clearly explain how the proposed methods extend the state of the art, as represented in those additional references. As the correction of both criticisms will require substantial changes to the manuscript and will probably necessitate the inclusion of additional theoretical and experimental results, the IJRR review process requires that the manuscript be rejected at this time. However, after you have carefully addressed the reviewers' concerns, you are encouraged to submit a new manuscript to IJRR. This new manuscript will be handled independently of this submission.

### 2017-05-15 MARSS17/expResults.tex
Added MARSS. Terminated.

--- 2017-03-22 journalSwarmControl/TASE/expResults.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Human-swarm interaction results}\label{sec:expResults}
+\section{Human-swarm interaction results}\label{sec:expResults}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We designed six experiments to investigate human control of large robotic swarms for manipulation tasks.
-Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of robots for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 robots, and three control architectures for both assembly and foraging tasks.
-The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.
-We recorded the completion time and the participant ID for each successful trial.
-%As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.
-Still, many played multiple games, and their decreasing completion times demonstrates their skills improved.
-
-\paragraph{Varying number:}
-\begin{figure}
-\begin{overpic}[width = 0.9\columnwidth]{ResVaryNum.pdf}\end{overpic}
-%\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
-\caption{
-\label{fig:ResVaryNu}Data from \emph{Varying Number} using robots to push an object through a maze to a goal location. 
-% Best-fit linear and quadratic lines are overlaid for comparison. 
-%(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
-}
-\end{figure}
-
-Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems.
-This experiment varied from 1 to 500 the population of robots used to transport an object.
-The total area, maximum robot speed, and total net force the swarm could produce were constant.
-The robots pushed a large hexagonal object through an  {\sffamily S}-shaped maze.
-We hypothesized participants would complete the task faster with more robots.
-The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 robots.
-
-\paragraph{Varying visualization:}
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.24\columnwidth}
-\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
-\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
-\vspace{-.5em}
-\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}.
-This experiment challenges players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region.
-We record the completion time and other statistics.
-%\vspace{-1em}
-}
-\end{figure}
-
-\begin{figure}
-%  \vspace{-20pt}
-  \begin{center}
-  \includegraphics[width=0.8\columnwidth]{ResVaryVis.pdf}
-  \end{center}
-%  \vspace{-1em}
-\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
-\vspace{-1em}
-}
-\end{figure}
-
-Sensing is expensive, especially on the nanoscale.
-To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source.
-Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}.
-Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.
-This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all robots; {\bf convex-hull} draws a convex hull around the outermost robots; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse.
-Fig.~\ref{fig:Visualization} shows screenshots of the same robot swarm with each type of visual feedback.
-Full-state requires $2n$ data points for $n$ robots.
-Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.
-Mean requires two, and variance three, data points.
-Mean and mean + variance are convenient even with millions of robots.
-
-%\begin{figure}[b!]
-%\renewcommand{\figwid}{0.5\columnwidth}
-%\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
-%\vspace{-1em}
-%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
-%%\vspace{-1em}
-%}
-%\end{figure}
-
-% Additionally, as population characteristics, they are available even if only a percentage of the robots are detected each control cycle.
-%Photobleaching: http://www.piercenet.com/browse.cfm?fldID=4DD9D52E-5056-8A76-4E6E-E217FAD0D86B
-%
-%Photobleaching is caused by the irreversible destruction of fluorophores due to either the prolonged exposure to the excitation source or exposure to high-intensity excitation light.
-Photobleaching can be minimized or avoided by exposing the fluor(s) to the lowest possible level of excitation light intensity for the shortest length of time that still yields the best signal detection; this requires optimization of the detection method using high sensitivity CCD cameras, high numerical aperture objective and/or the widest bandpass emission filter(s) available.
-Other approaches include using fluorophores that are more photostable than traditional fluorophores and/or using antifade reagents to protect the fluor(s) against photobleaching.
-Steps to avoid photobleaching are not feasible for all detection methods and should be optimized for each method used.
-For example, antifade reagents are toxic to live cells, and therefore they can only be used with fixed cells or tissue.
-Furthermore, some detection methods, such as flow cytometry, normally do not require steps to avoid photobleaching because of the extremely short exposure time of the fluorophore to the excitation source.
-
-%\begin{figure}
-%\centering
-%\begin{overpic}[width = \columnwidth]{ResVaryVis.pdf}\end{overpic}
-%\vspace{-2em}
-%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.
-Surprisingly, players perform better with limited feedback--subjects with only the mean + variance  outperformed all others.
-%\vspace{-2em}
-%}
-%\end{figure}
-
-Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
-Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
-All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
-Using ANOVA analysis, we reject the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
-Anecdotal evidence from beta-testers who played the game suggests that tracking 100 robots is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
-Our beta-testers described convex-hull feedback as confusing and irritating.
-A single robot left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
-%obscuring what the rest of the swarm is doing.
-
-\paragraph{Varying noise:}
-Microrobots are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
-This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
-\begin{align*}
-\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
- \dot{y}_i &= u_y + m_i\sin(\psi_i).
- \end{align*}
-$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
-
-We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the robots move erratically.
-Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.
-
-\begin{figure}[b!]
-\renewcommand{\figwid}{0.49\columnwidth}
-\begin{overpic}[width =\figwid]{ResVaryNoise.pdf}\end{overpic}
-\begin{overpic}[width =\figwid]{ResPositioning.pdf}\end{overpic}
-%\vspace{-1em}
-\caption{\label{fig:ResVaryNoisePosition} Left: Varying the noise from 0 to 200\% of the maximum control input resulted in only a small increase in completion time.
-Right: For position control, increasing the number of robots resulted in longer completion times.
-For more than 4 robots the goal pattern contained a void, which may have caused the jump in completion times.
-%\vspace{-1em}
-}
-\end{figure}
-
-\paragraph{Position control:}
-This online experiment examined how completion time scales with the number of robots $n$.
-Using a single square obstacle, users arranged $n\in[1,10]$ robots into a specified goal pattern.
-The goal pattern formed a block  {\sffamily A} with 10 robots, and lesser numbers of robots used a subset of these goal positions.
-Our hypothesis was that completion time would increase linearly with the number of robots, as with our position control algorithm in \cite{Becker2013b}.
-Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.
-Though the number of robots presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
-
-Time increases for $n>4$ robots.
-For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.
-A better experiment design would randomly place the goal positions.
-Initially we tried this, but beta-testers strongly disliked trying to arrange robots in random patterns.
-
-\paragraph{Varying control--Assembly:}
-Ultimately, we want to use swarms of robots to build things. This experiment compared different control architectures modeled after real-world devices.
+\paragraph{Varying control--Assembly}
+Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
 
@@
 Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
-
-%%\begin{figure}
-%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
-%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
-%%\caption{
-%%\label{fig:attractiveForce}
-%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
-%%}
-%%\end{figure}
+%
+%\begin{figure}
+%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
+%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
+%\caption{
+%\label{fig:attractiveForce}
+%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
+%}
+%\end{figure}
 
 \begin{figure}[b!]
-\renewcommand{\figwid}{0.45\columnwidth}
+\renewcommand{\figwid}{0.9\columnwidth}
 \begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
@@
 
-\paragraph{Varying control--Foraging:}
+\paragraph{Varying control--Foraging}
 Collecting and delivering resources is necessary for drug delivery.
 This experiment also compared attractive and repulsive control with the global control used for the other experiments.
-The experiment challenged players to collect particles using a swarm of 100 robots and return the particles to a home region.
-The robots encapsulate the particles on contact.
+The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region.
+The particles encapsulate the particles on contact.
 

### 2017-06-05 journalSwarmControl/coverletter/EditorNoteTROparticleSwarmManipulation.tex
added 'balance' option
final edits to paper (abstract too long)

@@
 The first conference paper covered the first three months of SwarmControl.net experiments [1], and the second one presented simulations of object manipulation [2].
+We have incorporated numerous suggestions from external reviewers to increase the scope of this work.
 This paper presents three years of results from SwarmControl.net.
@@
  [2] S. Shahrokhi and A. T. Becker, “Stochastic swarm control with global inputs,” in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Sep. 2015, pp. 421–427.
+
+Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
+The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
+Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
+To identify the key parameters for particle manipulation, we used a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges.
+We recorded statistics from over ten thousand players.
+Inspired by techniques where human operators performed well, we investigate controllers that use only the mean and variance of the swarm.
+We prove the mean position is controllable and provide conditions under which variance is controllable.
+We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution.
+Finally, we employ these controllers as primitives for an object manipulation task and implement all controllers on 100 kilobots controlled by the direction of a global light source.

### 2017-06-05 journalSwarmControl/relatedWork.tex
Added bib and second reference of IJRR reviewer to the paper
Added MARSS. Terminated.
final edits to paper (abstract too long)

@@
 \subsection{Global control of microrobots}
-Small robots have been constructed with physical heterogeneity so that they respond differently to a global broadcast control signal.
+
+This paper investigates global control of particles that have no onboard computation.
+This prevents us from applying controllers that require computation on the agents, as in \cite{milutinovic2006modeling,prorok2011multi,demir2015probabilistic}.
+%\cite{milutinovic2006modeling} requires robots that can execute task primitives.  Our particles simply move in the direction of the global control unless they are stopped by an obstacle. 
+%\cite{palmer2011hamiltonian} uses only Open-Loop control.  Our submission relies on state feedback 
+%\cite{prorok2011multi} requires collective intelligence of a group of miniature robots
+%\cite{demir2015probabilistic} requires robots that can individually decide when to actuate and when to remain anchored, can perform onboard computation (generate a random number, calculate its bin and the next bin).
+Another control paradigm is to construct robots with physical heterogeneity so that they respond differently to a global broadcast control signal.
 Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};  magnetic structures  with different cross-sections that can be independently steered by \cite{Floyd2011,Diller2013};  \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
 \cite{Tottori2012} and \cite{Peyer2013}. 
+Similarly, our previous work, \cite{Becker2012,Becker2012k}, focused on exploiting inhomogeneity between robots.
+These control algorithms theoretically apply to any number of robots, even robotic continuums.
 
-%In our previous work with robots that can be modeled as nonholonomic unicycles~\cite{Becker2012k}, we showed that an inhomogeneity in turning speed is enough to make even an infinite number of robots controllable with regard to position.
-%While these approaches theoretically scale to large populations, they also require both excellent state estimation and heterogeneous robots.
-
-%For micro- and nanorobotics, the marginal cost of producing one additional robot is astonishingly small.
-While some microrobots are individually fabricated \cite{Tottori2012}, others are fabricated using MEMS techniques \cite{Donald2008}, and these dies could be tiled to produce multiple copies.
-Nanocars are exemplary of this diminishing marginal cost.
-Nanocars are synthetic molecules with integrated axles, rolling wheels, and light-driven motors developed by Tour et al.\cite{Chiang2011}.
-These  are routinely produced in quantities that are amazing---a batch  the size of an aspirin tablet  contained$~\approx4\times 10^{19}$ nanocars.
-This dwarfs the total number of birds on the planet earth---some $3\times10^{11}$\cite{Gaston1997}.
- %aspirin in US weighs 325 mg.
-  %or the number of automobiles produced in the history of humankind
-% http://www.virlab.virginia.edu/nanoscience_class/lecture_notes/Lecture_13_Materials/Tour%20on%20nanovehicles.pdf states:
-%	On the other hand, nanocars synthesized in our laboratory
-%	are each calculated to be approximately 3x6x4 nm in size, and
-%	can be produced on the 30 mg scale (3.26x10^18 nanocars)
-%	using small 100 ml laboratory reaction flasks, or more
-%	nanocars than the number of automobiles made in the
-%	history of the world (63 million automobiles were produced
-%	in 2005). It would take 30 nanocars, side by side, to span the
-%	90 nm width of a small line in the most advanced logic
-%	chip being made today
-% The motorized nanocar can be synthesized in 12 steps with an overall yield of 5\%,  An electrically driven one is \cite{Kudernac2011}
-%Biological agents, including the magnetically controlled ciliate protozoa studied by A.
-Julius~\cite{Ou2012, Ou2012a} and the magnetotactic bacteria studied by Martel~\cite{Felfoul2011}can be grown or purchased in large numbers.
-% According to a scientific study of trace elements in hair, a single strand of hair 4 1/2 inches in length weighs, on average, 0.62 milligram, with weights varying from 0.25 milligram to just under 1 milligram.
-%mosquito weight = and weight up to 2.5 milligrams
-%Read more: How much does one strand of human hair weigh? | Answerbag http://www.answerbag.com/q_view/2159502#ixzz2EUltJTAl
-
-%
-%Consider a collection of $n$ planar robots.
-We  describe the configuration of the $i$-th robot by $q_i=[x_i,y_i]^\top$ and its configuration space by $\Cspace = \R^2$.
-At the micro/nanoscale viscous effects dominate inertial forces~\cite{Purcell1977}, so our model ignores velocity and assumes in the absence of an external force, the robots stop moving.
-% We describe the configuration of the environment by the magnitude of the force field $f = [f_x,f_y]\in\R^2$, which is assumed constant across the workspace.
-The control inputs $u = [u_x,u_y]\in\R^2$ change this force field.
-%\begin{align}
-%\dot{x}_i &= f_x,&
-%\dot{f_x} &= u_x,\nonumber \\
-%\dot{y}_i &= f_y, &
-%\dot{f_y} &= u_y. \label{eq:sysModel}
-%\end{align}
-%
-%\todo{This model is key, because if we use impulsive inputs for $u_x$ and $u_y$, there are a number of tasks we can accomplish with a single input (e.g.
-2,3,5,7 in Fig.~\ref{fig:boundaryManipulators}.)}
-
-%In \cite{Becker2012} we proved that systems described by \eqref{eq:ensembleDynamics} can be approximately steered to arbitrary positions in $\R^2$.
-In \cite{Becker2012k} we provided closed-loop controllers to drive $n$ robots to $n$ target positions.
-For both works, the control laws apply even to an infinite robot continuum.
-%
-%Unfortunately, in \cite{Becker2012}, we proved that the heading of the ensemble---the $\theta_i$ values---is not controllable.
-The final heading of every robot in the ensemble is the scaled integration of all the turning control inputs, i.e.~ $\theta_i(T) = \theta_i(0) + \epsilon_i\int_0^T \omega(t)\,{\rm d}t$.
-Instead of a full-state controller, we provided a control law that steered the position of each robot's center of rotation.
-Our goal is to provide control algorithms for many robots with uniform inputs, with  application in trajectory following, manipulation, force-closure, and assembly at the micro- and nanoscale.
-Regrettably,
-% being unable to control the heading of the robots makes many tasks difficult or impossible.
-For instance, manipulation and assembly work without heading control requires that the robots be able to rotate in place and be circular, since we cannot specify the incident angle of the robots.
-%In practice, this limits the application of our work to circular or point robots.
-%As Lynch explained in \cite{Lynch1999}, pushing with a round object is inherently unstable.
-
-%  Possible applications include using molecular robots as nanoscale transporters, to break chemical bonds, or to build structures by constructing non-covalent bonds.
-We derive inspiration from the molecular actuators of Roth et al.~\cite{Minett2002}, the molecular elevators of Badj{\'i}c at al.~\cite{Badjic2004}, and the nanocars of Tour et al.~\cite{Vives2009}.
-%hydrogen bonds
-
-% From James Tour:
-%My motivation has been clear on this topic, and I have written it many times.  Here is a synopsis:
-%Transport of goods and materials between points is at the
-%heart of all engineering and construction in real-world systems.
-%Just as biological systems survive by nanometer-scale
-%transport using molecular-sized entities, as we delve into the
-%arena of the nanosized world, it beckons that we learn to
-%manipulate and transport nanometer-scale materials in a similar
-%manner. Nanoscale transporters that are truly moleculesized
-%will be required for the enzyme-like fabrication of
-%sophisticated integrated structures.
-%We want to build things from the bottom up, ex vivo, much like enzyme build in vivo???that is and has been my motivation.
-I do not think that targeting these for biological applications is wise.
-One would use enzyme-like structures for that, not nanocars.
-These are for ex vivo.
-
-Similarly, our previous work, \cite{Becker2012,Becker2012k}, focused on exploiting inhomogeneity between robots.
-These control algorithms theoretically apply to any number of robots---even robotic continuums---but in practice process noise cancels the differentiating effects of inhomogeneity for more than tens of robots.
-We desire control algorithms that extend to many thousands of robots.
+However, process noise cancels the differentiating effects of inhomogeneity for more than tens of robots.  We desire control algorithms that extend to many thousands of robots.
+Limited position control was achieved by \cite{bretl2007control} and our previous work \cite{beckerIJRR2014}, but both used robots commanded in their local coordinate frame.
+Our new submission focuses on a more common paradigm: particles commanded in a global coordinate frame
 
@@
 Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.
-This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
+This paper exploits insights gained from \href{http://www.swarmcontrol.net}{\emph{SwarmControl.net}}, particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
 
@@
 	Using an online testing environment, we conduct similar studies but with sample sizes three orders of magnitude larger.
+%Crowdsourcing human-swarm interaction is gaining considerable attention, see \cite{tavakoli2016crowdsourced} and \cite{hauert2014mechanisms}.
 

### 2017-06-06 journalSwarmControl/pictures/pdf/Weight.m
moved unneeded images and made the arXiv submission

@@ -1,30 +1,0 @@
-clear all;
-
-%....Results taken from 15 simulations...%
-result050 = [93.733;128.257;90.424;82.174;104.578;160.626;148.221;82.222;129.733;99.704]; 
-
-result075 = [122.759;77.254;112.602;97.494;81.712;89.443;87.103;83.894;143.415;125.442];
-result100 = [77.492;75.84;88.577;106.791;105.187;121.288;107.253;84.696;74.841;154.204];
-%result100 = [76.868;93.446;68.473;87.333;84.375;116.218;108.54;106.594;81.217;74.186];
-result125 = [137.918;76.143;89.689;130.424;100.966;119.111;83.886;72.423;182.625;109.818];
-result150 = [96.93;64.471;92.351;114.401;100.407;75.588;101.824;89.568;143.857;112.561];
-
-%...to plot the boxplot...
-Time = [result050; result075; result100; result125; result150];
-
-
- group = ['result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';
-     %'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';'result050';
-     'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';
-      %'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';'result075';
-    'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';
-    %'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';'result100';
-    'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';
-    %'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';'result125';
-    %'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';
-     'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';'result150';];
-
-boxplot(Time, group);
-title('How Object Density will affect time');
-xlabel('Algorithm');
-ylabel('Time (s)');

### 2017-07-27 ICRA-RA-L2018/experiment.tex
Final version of IROS is added. New paper started

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Demonstration}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\emph{Hardware Demonstration: Position Control of n Robots}\\
A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
 % Designs for each are available at \cite{Arun2016Thingiverse}. 
  The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
    Only discrete, 1 cm moves in the $x$ and $y$ directions are used. The goal configuration highlighted in the top right corner represents a `U' made of seven particles. The dark red configuration is the current position of the particles. 
Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.  Free particles return to their start positions but particles on the boundary to move laterally, generating a net sliding motion.

Fig. \ref{fig:construction2d} follows the motion of the particles through iterations  $k$=1, 2, and 7. All particles receive the same control inputs, but boundary interactions break the control symmetry.  Particles reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left particle from the staging zone occupying the top-right position of the build zone.

### 2017-07-27 ICRA-RA-L2018/intro.tex
Final version of IROS is added. New paper started

--- 2017-03-01 IROS17/intro.tex
@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a global field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics. 
+Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics. 
 %If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
@@
  For each task, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
-Instead, these robots are often controlled by a global, broadcast signal.
- The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a global control field, and sensing, as the robot.
+Instead, these robots are often controlled by a broadcast signal.
+ The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
 Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the particle swarm.
@@
 
-Positioning is a foundational capability for a robotic system, but requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
-This paper relaxes that constraint, and provides position control algorithms that only require interactions with the boundaries.
-We assume that particles in contact with the boundaries have zero velocity if the global control input pushes the particle into the wall.
+Positioning is a foundational capability for a robotic system, e.g.
+placement of brachytherapy seeds.
+However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
+This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contacts.
+We assume that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
 
@@
 \vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 102){$n$=862}\put(16,60){goal}
+\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 101){$n$=862}\put(16,60){goal}
 %\put(29.9,20){\scalebox{.91}[1.0]{\textcolor{white}{\textbf{robots}}}}
 % \put(30,20){\textcolor{black}{robots}}
- \put(30,5){\textcolor{white}{\textbf{robots}}}
+ \put(20,45){\textcolor{white}{\textbf{goal positions}}}
+  \put(25,5){\textcolor{white}{\textbf{particles}}}
 \end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 102){position control algorithm}\put(12,25){final configuration}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 101){position control algorithm}\put(12,25){final configuration}\end{overpic}
 \caption{\label{fig:IntroPic}
@@
  Top left shows the initial position and goal position of the particles.  
- The shortest path consists of moving at angle 50$^\circ$ until the blue robot contacts the top wall, then moving the  magenta robot at angle 165$^\circ$ until the particles reach the desired relative spacing, then moving -60$^\circ$ to the goal positions.
- The two bottom pictures shows $n$-particle positioning using shared control inputs and boundary interaction. 
+ The shortest path consists of moving at angle 50$^\circ$ until the blue robot contacts the top wall, then moving the  magenta robot at angle 165$^\circ$ until the particles reach the desired relative spacing, then moving $-$60$^\circ$ to the goal positions.
+ The two bottom pictures show $n$-particle positioning using shared control inputs and boundary interaction. 
 } \vspace{-1em}
@@
   Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
-We provide an optimal shortest-path algorithm for  arbitrarily position two robots in Sec.
+We provide a shortest-path algorithm to arbitrarily position two robots in Sec.
 \ref{sec:PostionControl2Robots}, and Sec.
-\ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
+\ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with non-slip boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
 Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 

### 2017-07-27 ICRA-RA-L2018/simulation.tex
Final version of IROS is added. New paper started

--- 2017-03-02 IROS17/simulation.tex
@@
 
-Two simulations were implemented using wall-friction for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.
+Two simulations were implemented using non-slip contact walls for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.
 
@@
  %If two-move solutions exist, they only appear along the boundary.
-  If $g_2$ results in a contraction of $(\delta x, \delta y)$, there are many three move sequences. Four, five and seven moves are sometimes required.
+  If $g_2$ results in a contraction of $(\delta x, \delta y)$, there are many three move sequences. Four and five moves are sometimes required.
 
@@
 \subsection{Position Control of $n$ Robots}
-Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. Code is available at \cite{Arun2015}.
+Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. %Code is available at \cite{Arun2015}.
  Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
@@
 \caption{\label{fig:4diagramsplots.pdf}
-The required number of moves under Alg. \ref{alg:PosControlNRobots}  using wall-friction to rearrange $n$ square-shaped 
-robots grows quadratically with $n$.  
-See hardware implementation and simulation at \cite{Arun2015}.
+The required number of moves under Alg.~\ref{alg:PosControlNRobots} using non-slip contact walls to rearrange $n$ square-shaped robots grows quadratically with $n$. % \cite{Arun2015}.
+ See code at \cite{Arun2015}.
 }

### 2017-07-27 ICRA-RA-L2018/theory.tex
Final version of IROS is added. New paper started

--- 2017-02-28 IROS17/theory.tex
@@
 
-Global inputs in the absence of obstacles move a swarm uniformly.  
-Shape control requires breaking this uniform symmetry.
-The following sections examine using boundaries that stop the particles if they are pushed into the wall.
- These forces are  sufficient to break the symmetry caused by uniform inputs.
+In the absence of obstacles uniform inputs move a swarm identically.  
+Shape control requires breaking this symmetry. 
+The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.
 
@@
  In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
-  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
+  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
+  As in our model, fluid flowing in a pipe has zero velocity along the boundary.
+Similar mechanical systems exist at larger scales, e.g.
+all tumblers of a combination look move uniformly unless obstructed by an obstacle.
  Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.

### 2017-07-27 ICRA-RA-L2018/theory2robAlgs.tex
Final version of IROS is added. New paper started

--- 2017-03-02 IROS17/theory2robAlgs.tex
@@
 
-Alg.~\ref{alg:optimalAlg} uses wall-friction to arbitrarily position two robots in a rectangular workspace. 
+Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a rectangular workspace. 
  Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
@@
 \renewcommand{\figwid}{0.4\columnwidth}
+
 {\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
@@
 \put(87,52){$g_2$}
+\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
 \end{overpic}
@@
 \put(110,50){If $(\Delta e.x, \Delta e.y) = (0,0)$, only one move is needed}
+\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
+%\put(47, 65){$\xleftrightarrow{1.8cm}$}
 \end{overpic}
@@
 \begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
+\put(110,50){Three move sequence}
+\put(12,38){$m_3$}
+\put(48,52){$m_2$}
+\put(65,30){$m_1$}
 \end{overpic}
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 4}
+\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 6}
+\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
 \end{overpic}
-\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 7}
+\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
 \end{overpic}
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 4}
+\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 6}
+\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 7}
+\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
+\put(15, 85){worst case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
 \end{overpic}
@@
 \vspace{-1em}
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two robot positioning using walls with infinite friction. 
+\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two robot positioning using walls with non-slip contacts. 
 Robot start positions are shown by a square, and goal positions by a circle.
@@
 Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-The solution is a best-first search algorithm that maintains a list of possible paths ($moves$), sorted according to an admissible heuristic on path length.
+The solution is a best-first search algorithm that maintains a list of possible paths (\emph{moves}), sorted according to an admissible heuristic on path length.
 The algorithm works by expanding the path with the shortest estimated length.
@@
 Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
-The algorithm assigns a global control input at every instance.
+The algorithm assigns a uniform control input at every instance.
 The goal is to zero the error in both coordinates,
- $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared global control input. 
+ $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared control input. 
  The base case occurs when $\Delta e = (0,0)$.
-In this base case, the shortest path is a straight line between current position of each robot to its goal position, as shown in Fig.
+In this base case, the shortest path is a straight line between the current position of each robot to its goal position, as shown in Fig.
 \ref{fig:shapeControlMathematica1}a.
-If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets shown in Fig.~\ref{fig:TwoRegions}.
- If a robot is touching a wall, the other robot can move to any position in the reachable set while the first robot is immobile. 
-Two reachable sets are possible, horizontal and vertical. 
-The reachable sets may change after each move, as shown in  Fig.~\ref{fig:regionMove}.
+If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets generated by first immobilizing one particle by contacting the wall and then repositioning the second particle.
+There are four reachable sets, as shown in Fig.~\ref{fig:TwoRegions}.
+The horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose to immobilize the particle closest to a wall. 
+% When a particle contacts a wall, the other particle can move to any position in the reachable set while the first robot is immobile. 
+%Two reachable sets are possible, horizontal and vertical. 
+Each non-slip move changes the $\Delta$ configuration space, as shown in  Fig.~\ref{fig:regionMove}.
 
@@
 %The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
-If the goal configuration can be reached in three moves:  $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
+If the goal configuration can be reached in three moves, then $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
 \ref{fig:shapeControlMathematica1}b.
@@
 \begin{align}\label{eq:admissibleHeuristic}
-h(moves, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|moves|} \norm{moves_i}  \\
+h(\text{\emph{moves}}, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|\text{\emph{moves}}|} \norm{\text{\emph{moves}}_i}  \\
 &+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
@@
  $L$ is length of the walls. 
- $PathList$ contains all the paths sorted by their path length plus an admissible heuristic. 
- \State  $PathList \gets \{\}$
+ \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
+ \State  \emph{PathList} $\gets \{\}$
  \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains the admissible heuristic, the move sequence, and the current robot positions
@@
 \State $\{d, $ \parbox[t]{.3\linewidth}{%
- $moves, r_1,r_2\} \gets$\\
- {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.moves$)}
-\State $(moves, r_1,r_2) \gets$ {\sc Rotate}($moves, r_1,r_2,-\theta$)
-\State   {\sc Push} $\{d, moves, r_1,r_2\} $ onto $PathList$
+ \emph{moves,}$ r_1,r_2\} \gets$\\
+ {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.\text{\emph{moves}}$)}
+\State $(\text{\emph{moves}}, r_1,r_2) \gets$ {\sc Rotate}(\emph{moves}$, r_1,r_2,-\theta$)
+\State   {\sc Push} $\{d, \text{\emph{moves}}, r_1,r_2\} $ onto \emph{PathList}
 \EndFor
-\State {\sc Sort}($PathList$) \Comment sort by admissible heuristic
-\State $P \gets $ {\sc Pop} first element of $PathList$
+\State {\sc Sort}(\emph{PathList}) \Comment sort by admissible heuristic
+\State $P \gets $ {\sc Pop} first element of \emph{PathList}
 \EndWhile
-\State \Return $moves$
+\State \Return \emph{moves}
 \end{algorithmic}
@@
 \begin{algorithm}
-\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, moves$)}\label{alg:Wallup}
+\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, $\emph{moves})}\label{alg:Wallup}
 \begin{algorithmic}[1]
@@
 \State $m_1  \gets g_2 -  r_2$
-\State $moves \gets \{ moves,   m_1 \}$
+\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1 \}$
 \State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
+\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
 \EndIf
@@
 \EndIf
-\State $moves \gets \{ moves,   m_1\}$
+\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1\}$
 \State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
@@
 \EndIf
-\State $moves \gets \{ moves,   m_2\}$
+\State \emph{moves} $ \gets \{ \text{\emph{moves}},   m_2\}$
 \State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
-\State \Return $\{h(moves, r_1,r_2,g_1,g_2), moves,r_1,r_2\}$
+\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
 \end{algorithmic}
@@
 The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
-The green rectangle illustrates one move reachable $\Delta x$ and $\Delta y$ by horizontal walls and the blue rectangle illustrates the vertical walls reachable region.
+The green rectangle illustrates reachable $\Delta x$ and $\Delta y$ when one particle is in contact with a horizontal wall and the blue rectangle illustrates the reachable region when in contact with a vertical wall.
 }
@@
 \renewcommand{\figwid}{2\columnwidth}
-{\begin{overpic}[width =\figwid]{contourPlotDist.pdf}
+{\begin{overpic}[width =\figwid]{contourDistnew.png}
 \end{overpic}
@@
 \vspace{1em}
-\begin{overpic}[width =\figwid]{contourPlotNum.pdf}\put(97.5,21){7}
-\put(97.5,18.5){6}
-\put(97.5,16){5}
-\put(97.5,13.5){4}
-\put(97.5,11){3}
+\begin{overpic}[width =\figwid]{JustSimulationV6.png}
+%\put(97.5,21){5}
+%\put(97.5,17.25){4}
+%\put(97.5,13.5){3}
+%\put(97.5,24){6}
+%\put(97.5,27){7}
+%\put(97.5,10){2}
+%\put(97.5,7.5){1}
+%\put(97.5,5){0}
 \end{overpic}

### 2017-07-27 ICRA-RA-L2018/theorynrobAlgs.tex
Final version of IROS is added. New paper started

--- 2017-03-02 IROS17/theorynrobAlgs.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The ideas from Alg. \ref{alg:optimalAlg}  can be extended to control the position of $n$ particles using wall friction.
+The ideas from Alg. \ref{alg:optimalAlg}  can be extended to control the position of $n$ particles using walls with non-slip contact.
 The solution is complete, but not optimal, and requires the starting and final configurations of particles to be disjoint.
@@
  The $k^{\text{th}}$ loop moves the $k^{\text{th}}$ robot from a \emph{staging zone} to the desired position in a \emph{build zone}. 
-  All robots move according to the global input, but due to wall friction, at the end of the $k^{\text{th}}$ loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
+  All robots move according to the uniform input, but due to non-slip wall contacts, at the end of the $k^{\text{th}}$ loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
    See Fig.~\ref{fig:simulationNrobot} for a schematic of the build and staging zones.
@@
 \caption{\label{fig:simulationNrobot}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using walls with non-slip contact.
 }
@@
 
-Assume an open workspace with four axis-aligned walls with infinite friction.
+Assume an open workspace with four axis-aligned walls with non-slip contact.
 The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots.
@@
 \begin{algorithm}
-\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
+\caption{PositionControl$n$Robots($k$)}\label{alg:PosControlNRobots}
 \begin{algorithmic}[1]
@@
 All particles are 3D-printed plastic whose top is a 1cm diameter cylinder with a narrower base that encapsulates a steel bearing ball.
-Wall friction is emulated by a toothed wall design to keep particles from moving out of place while implementing the drift move. 
+Non-slip wall contact is generated by a toothed wall design to keep particles from moving out of place while implementing the drift move. 
 The workspace boundary is mounted on top of a white sheet of cardboard.
-Underneath the cardboard, a grid of 3mm diameter magnets glued with 1 cm spacing to a thin board generates the global control input.
+Underneath the cardboard, a grid of 3mm diameter magnets glued with 1 cm spacing to a thin board generates the uniform control input.
  A video attachment  shows the algorithm at work. 
@@
  In this demonstration, all moves are 1 cm in length.
-   All drift moves are an counterclockwise \emph{square} move  of size 1 cm$\times$1 cm. 
-   Once the $k^{\text{th}}$ roller gets to its designated location in each loop, a correction step is implemented. 
+   All drift moves are a counterclockwise \emph{square} move  of size 1 cm$\times$1 cm. 
+   Once the $k^{\text{th}}$ particle gets to its designated location in each loop, a correction step is implemented. 
    This correction step increases by two the total number of moves required per particle.
@@
 \begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
+	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware2.pdf}
 \end{center}
@@
 \caption{\label{fig:construction2d}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, discretized $n$ robot position control  using wall friction.
+Illustration of Alg.\ \ref{alg:PosControlNRobots}, discretized $n$ robot position control  using walls that enforce non-slip contact.
 }

### 2017-07-27 JournalTorqueControl/theory.tex
Some of Dylan's comments are affected
Edited some texts
Final version of IROS is added. New paper started

@@
 
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{Uniform.pdf}%\put(1,55){a)}
-\end{overpic}
-\begin{overpic}[width =\figwid]{Triangular.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{Normal.pdf}
-\end{overpic}
-\vspace{-0.5em}
-\caption{\label{fig:PDF} Three different distributions are examined in this work: uniform, triangular, and normal.
-%\vspace{-2em}
-}
-\end{figure*}
+%\begin{figure*}
+%\centering
+%\renewcommand{\figwid}{0.66\columnwidth}
+%\begin{overpic}[width =\figwid]{Uniform.pdf}%\put(1,55){a)}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{Triangular.pdf}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{Normal.pdf}
+%\end{overpic}
+%\vspace{-0.5em}
+%\caption{\label{fig:PDF} Three different distributions are examined in this work: uniform, triangular, and normal.
+%%\vspace{-2em}
+%}
+%\end{figure*}
 
@@
 \vspace{-0.5em}
-\caption{\label{fig:torque} Torque with regard to mean position, $\mu$. Mean position is the pushing location.
+\caption{\label{fig:torque} Torque with regard to mean position, $\mu$. Mean position is the pushing location along a rod extending from 0 to 1.
 %\vspace{-2em}
@@
 \begin{align}
-\tau_{left} &=  {u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
+\tau_{l} &=  {u_1}^2(2u_1 - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
 &-2{l_1}^3+3{l_1}^2(\mu-\sqrt{6}\sigma), & \textrm{for     } u_1 > l_1\\ \nonumber
-\tau_{right} &= 2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
+\tau_{r} &= 2{l_2}^3-3{l_2}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
 &-{u_2}^2(2u_2 - 3\mu-3\sqrt{6}\sigma),  & \textrm{for     } u_2 > l_2\\ \nonumber
@@
 \begin{align} \label{eq:normalTorque}
-\tau_n &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right)
+\tau_n &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\ \nonumber
+&+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right) 
 \end{align}
@@
 \begin{equation}
-\tau_{free} = \int_{-1}^1 x\,p(x)\, dx
+\tau_{f} = \int_{-1}^1 x\,p(x)\, dx
 \end{equation}
@@
 \begin{align}
-\tau_{tf} = \frac{\tau_{fleft}+ \tau_{fright}}{36\sigma^2}
+\tau_{tf} = \frac{\tau_{fl}+ \tau_{fr}}{36\sigma^2}
 \end{align}
-where $\tau_{fleft}$ and $\tau_{fright}$ is defined as
+where $\tau_{fl}$ and $\tau_{fr}$ is defined as
 \begin{align}\nonumber
-\tau_{fleft} &=  {u_{f1}}^2(2u_{f1} - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
+\tau_{fl} &=  {u_{f1}}^2(2u_{f1} - 3\mu+3\sqrt{6}\sigma)\\ \nonumber
 &-2{l_{f1}}^3+3{l_{f1}}^2(\mu-\sqrt{6}\sigma) &   \textrm{for     } u_{f1} > l_{f1}\\ \nonumber
-\tau_{fright} &= 2{l_{f2}}^3-3{l_{f2}}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
+\tau_{fr} &= 2{l_{f2}}^3-3{l_{f2}}^2(\mu+\sqrt{6}\sigma)\\ \nonumber
 &+{u_{f2}}^2( 3\mu-2u_{f2}+3\sqrt{6}\sigma) &   \textrm{for     } u_{f2} > l_{f2}\\ \nonumber

### 2017-07-27 journalSwarmControl/Introduction.tex

@@
 This paper examines object manipulation by a swarm of particles, each actuated by the same shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
+The transportation methodology is similar to that in \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
 
@@
 To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions.
-These controllers are used as primitives to perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.
+ To perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}, we use these controllers   as primitives, policy iteration for path planning, handle outliers by partitioning the workspace, and minimize pushing the object backwards with potential field navigation.
 

### 2017-07-27 journalSwarmControl/experiment.tex
added cover letter and other edits to the paper (mean = std, previous work by author)

@@
 
-The table below summarizes the results:
-
+The table below summarizes the simulation results for each 10 successful trials:
 \begin{center}
@@
 \hline
-Method & Result (s)\\ [0.5ex] 
+Method & Result, mean$\pm$std (s)\\ [0.5ex] 
 \hline \hline
@@
 \begin{center}
-	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
-	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
+	\begin{overpic}[width=0.9\columnwidth]{StateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
+	~~\begin{overpic}[width=0.45\columnwidth]{MainRegions.pdf} %,grid=true
 	\put(-6,85){\textbf{b}}
@@
 	~~
-	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}
+	\begin{overpic}[width=0.45\columnwidth]{TransferRegions.pdf}
 	\put(-6,85){\textbf{c}}

### 2017-08-01 ICRA-RA-L2018/intro.tex
The first draft?

@@
 \vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves1.pdf}\put(40, 102){$n$=2}\put(20, 4){{\tiny$\updownarrow$}~$\epsilon$}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{story3Moves4.pdf}\put(30, 102){shortest path}\end{overpic}\\
-\vspace{1.5em}
-\begin{overpic}[width=0.47\columnwidth]{monalisaBegin.pdf}\put(25, 101){$n$=862}\put(16,60){goal}
-%\put(29.9,20){\scalebox{.91}[1.0]{\textcolor{white}{\textbf{robots}}}}
-% \put(30,20){\textcolor{black}{robots}}
- \put(20,45){\textcolor{white}{\textbf{goal positions}}}
-  \put(25,5){\textcolor{white}{\textbf{particles}}}
-\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{monalisa.pdf}\put(2, 101){position control algorithm}\put(12,25){final configuration}\end{overpic}
+\begin{overpic}[width=\columnwidth]{circleWallFrictionReachableRegions.pdf}\end{overpic}
 \caption{\label{fig:IntroPic}

### 2017-08-01 ICRA-RA-L2018/theory2robAlgs.tex
The first draft?

@@
 
-\begin{figure*}
-\renewcommand{\figwid}{0.4\columnwidth}
-
-{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
-\put(10,85){a)}
-\put(30,70){$r_1$}
-\put(48,88){$g_1$}
-\put(65,17){$r_2$}
-\put(87,52){$g_2$}
-\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
-\put(110,50){If $(\Delta e.x, \Delta e.y) = (0,0)$, only one move is needed}
-\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
-%\put(47, 65){$\xleftrightarrow{1.8cm}$}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
-\put(10,85){b)}
-\put(28,70){$r_1$}
-\put(50,80){$g_1$}
-\put(62,15){$r_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
-\put(110,50){Three move sequence}
-\put(12,38){$m_3$}
-\put(48,52){$m_2$}
-\put(65,30){$m_1$}
-\end{overpic}
-%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
-%\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
-\put(10,85){c)}
-\put(28,70){$r_1$}
-\put(62,28){$g_1$}
-\put(87,50){$r_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-%\vspace{1em}
-{
-\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
-\put(10,85){d)}
-\put(25,30){$r_1$}
-\put(89,70){$g_1$}
-\put(67,20){$r_2$}
-\put(28,75){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
-\put(10,85){e)}
-\put(15,35){$r_1,g_2$}
-\put(65,35){$r_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
-\put(10,85){f)}
-\put(15,10){$r_1,g_2$}
-\put(65,88){$r_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
-\put(15, 85){worst case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
-\end{overpic}
-}\\
-
-%\vspace{-0.75em}
-%{
-%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
-%\end{overpic}
-%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
-%%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
-%\end{overpic}
-%}\\
-
-\vspace{-1em}
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two robot positioning using walls with non-slip contacts. 
-Robot start positions are shown by a square, and goal positions by a circle.
-Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
-Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-\vspace{-1em}
-}
-\end{figure*}
 Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
@@
 \centering
-\renewcommand{\figwid}{0.67\columnwidth}
-{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove2.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove3.pdf}
+\renewcommand{\figwid}{\columnwidth}
+{\begin{overpic}[width =\figwid]{circleWallFrictionReachableRegions.pdf}
 \end{overpic}
@@
 \renewcommand{\figwid}{2\columnwidth}
-{\begin{overpic}[width =\figwid]{contourDistnew.png}
+{\begin{overpic}[width =\figwid]{identical.png}
 \end{overpic}

### 2017-08-15 journalSwarmControl/Reviews/Response.tex
Added Reviews from T-RO as well as making a few revisions according to those reviews

\documentclass[conference, onecolumn, 10pt]{../IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{times}

\usepackage{microtype,xparse,tcolorbox}
\usepackage{mathtools, cuted}
\usepackage{todonotes}

\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

\setlength\parindent{0pt}

\renewcommand{\thepoint}{[R\,\thereviewer.\arabic{point}] }

\newcommand{\response}[1]{\refstepcounter{point}  \smallbreak \noindent
	\hangindent=3.5em
	\hangafter=1
	\thepoint \space ``\textcolor{blue}{#1}" \par\smallskip}

\NewDocumentCommand \Reviewer { m } {
	\stepcounter{reviewer}
	\subsection*{Comments~by~Reviewer~#1}
}

\begin{document}

	\vspace*{\fill} 
	\begin{quote}  
		\todo[inline]{I still need to format this page to look like the one in the word document}

		Steering a Particle Swarm Using Global Inputs and Swarm Statistics

		Dear IEEE Transactions on Robotics Editorial Office,

		Please find attached the revised paper, Steering a Particle Swarm Using Global Inputs and Swarm Statistics
		along with the document containing response to the reviewers. We are grateful to the reviewers for helping us in improving our manuscript through their comments and questions. Please let us know if further information is required.

		Sincerely,

		\todo[inline]{Insert Signature}

		Aaron T Becker (on behalf of all the authors) 
	\end{quote}
	\vspace*{\fill}
	\break

	\section*{Response to Reviewers}
	In the following document, we’ve provided detailed responses to the comments and questions of the reviewers. Comments and questions by reviewers are in \textcolor{blue}{blue}, our responses are in black.

	\Reviewer{\#1}\response{the paper is very interesting, well written and well organized. It presents an important research topic which is of great interest of most people interested in swarm robotics. The video is also very nice and useful to better understand the work.}
	Thank you!

	\response{What are the assumptions on the manipulated block properties? (Friction? Necessary force to move it? Is one robot enough to move it?)}
		\todo[inline]{This is a question we were trying to avoid right? Since we can't seem to measure the strength of a single kilobot. Should we give simulation specs? What about the real world experiments?}

	\response{Regading the hybrid controller, what are the necessary assumptions to reach convergence? Are there practical assumptions to guarantee the claim (18) =$>$ (19)? How about the geometry of the environment? What happens if the environment is sparse so that corners are too far? Is it possible to control variance in this case?}
	This is an interesting problem that we can expand on in a later work.

	\response{In section VI the authors should give more detils regarding the Value Iteration approach. How did the authors set the probabilities p($x_j$ | x,u)? It seems it should be j instead of i in the loop "for i = ..." in (25).}
	The probabilities for set p($x_j$ | x,u) come from the reward function in equation (23). The loop in equation (25) is looping the probability p($x_i$ | $x_j$,u) such that for a specific $x_j$ it iterates through all $x_i$.

	\response{It is also necessary to give more details of the potential fields in VI-b. The parameters $\Delta \rho$, $\rho$, b in (26) and (27) were not defined.}
	A sentence has been added and a second sentence has been altered for clarification, now reading, "... The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b. ..."

	\response{In the outlier rejection, how are the regions computed?}
		\todo[inline]{We already state, "The \emph{main} regions are generated by extending obstacles until they meet another obstacle. The \emph{transfer} regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions." Are they looking for a formula?}

	\response{In algorithm 2, how is (29) applied? Is it used directly as ux and uy? The composition of the potential field and the plan obtained with the value iteration algorithm should be better clarified.}
	Yes, We have amended the paragraph prior with "These fields operate in place of $\mathbf{D}(\mathbf{b})$ only when the repulsive field is active."

	\response{In the simulations, what is the model for collisions? }
	We use a Javascript port called Box2D to calculate our physics. This has been clarified in Section 5.

	\Reviewer{\#2}Is there not a Reviewer 2?

	\Reviewer{\#3}\response{In the paper the terms: micro, nano robots, and particles are used interchangeably, that is perfectly fine, however, it would be better in the very beginning to make that clear possible with a footnote, especially the use of the term particles. }
		\todo[inline]{I thought in the first paragraph we state that our robots are the particles?}

	\response{In page two, column 1 , the authors state that "process noise cancels the differentiating effects... tens of robots." The term tens of robots is not clear. It would also help a citation if it exists.}
		\todo[inline]{Doesn't the term "tens of..." mean something of magnitude X$*10$?}

	\response{Throughout the paper citations [?] are used as nouns, while that is quite common practice the paper reads better when the authors are used. e.g. "[123] shows that..." instead use: "Joe and Bob [123] show that..."  }
		\todo[inline]{Should I change all of our citations then?}

	\response{When the walls are immovable then the proposed technique to push the swarm against a corner works as is demonstrated. In the case of the human body where the walls could be more elastic, or even permeable, how that would affect the proposed strategy.}
	That is a good question which we would love to observe further in a future work.

	\response{In the same line, if pushing against a "wall" could increase friction or even absorption of the transported medicine, have you consider a strategy that envelops the object with a small group of particles shielding the object from the walls?}
	That is a good idea however we are unable to implement such an idea in the current iteration of this paper.

	\response{The reviewer will recommend in simulation to create bigger more complex environments emulating the major veins or the digestive tract of the human body. Definitely something with curvy walls. }
	That is a good idea however we are unable to implement such an idea in the current iteration of this paper.

	\response{The paper, and in general the proposed approach, will gain by demonstrating the proposed approach in environments where there are junctions where the object has to be pushed in one of two, or more, options. }
	We are currently working on attempting to do just this! Look out for our future work.

	\response{Furthermore, it would be interesting to see if the proposed method will work in an open environment with multiple obstacles.}
	We would love to expand on our work in this fashion and are looking for collaborators to do so.

	\Reviewer{\#4}
	\response{One major concern is about the Lyapunov derivative in eq. (17) which looks wrong. I found $\dot{V}=2 \sigma \dot{\sigma} (\sigma^2-\sigma_{goal}^2)$. This demands to revise your justification as your result on variance stability conditions might be totally different. Please clarify.}
	\todo[inline]{I will look into this one more later, maybe.}

	\response{The global control law is presented for holonomic robots, while in small-scales often particles obey nonholonomic motion. Is your approach still applicable? }
		\todo[inline]{I don't know and if I did, is s/he asking for us to add in the paper?}

	\response{In eq. (17) how can be $\dot{\sigma}^2$ be negative? Please clarify.}
		\todo[inline]{That.... is a good question.}

	\response{Eq. (20) is a little bit vague. A clearer statement is required. Perhaps to say, eq. (20) is an extended version of eq. (11), which allows to control variance as well.}
	Great idea! The paper now reads, "A PID controller based on the original PD controller in Eq. (11) to regulate the variance to $\sigma^2_{\rm{ref}}$ is:".

	\response{Page 3 section 3 on the lesson list number 2, you have to define IID?}
	This issue has been remedied entirely by changing IID to "independent and identically distributed".

	\response{Page 3 in section 3A, please follow paragraph style in the middle of the text.}
		\todo[inline]{Dr Becker? I don't understand what isn't in paragraph format?}

	\response{In section 3C and D, the writing can be improved to make it easier for general readership.}
		\todo[inline]{It doesn't give much in terms of what it lacks though.}

	\response{Correct the repetition (control control) on the last line on the left in page 5.}
	Corrected, the line now reads, "There are several techniques for breaking the symmetry of the control input to allow controlling more states, ..."

\end{document}

### 2017-08-15 journalSwarmControl/experiment.tex
Added Reviews from T-RO as well as making a few revisions according to those reviews

@@
 When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
-The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}b.
+To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal but repulses the swarm from in front of the object.
+The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b.
+These fields operate in place of $\mathbf{D}(\mathbf{b})$ only when the repulsive field is active.
 \begin{align}
@@
 Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
+In both cases, $\Delta \rho$ is the position of the swarm.
 

### 2017-08-15 journalSwarmControl/simulation.tex
Added Reviews from T-RO as well as making a few revisions according to those reviews

@@
 
-Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation, presented in \cite{catto2010box2d}.
+Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics, including collision, density, and friction, and fixed-time step simulation, presented in \cite{catto2010box2d}.
 All experiments in this section ran on a Chrome web browser on a 2.6 GHz Macbook.

### 2017-08-15 journalSwarmControl/theory.tex
Added Reviews from T-RO as well as making a few revisions according to those reviews

@@
 %Due to symmetry of the control input, only the mean position and mean velocity are controllable. However, 
-There are several techniques for breaking the symmetry of the control control input to allow controlling more states, for example by using obstacles as in \cite{Becker2013b}, or by allowing independent noise sources as in \cite{beckerIJRR2014}.
+There are several techniques for breaking the symmetry of the control input to allow controlling more states, for example by using obstacles as in \cite{Becker2013b}, or by allowing independent noise sources as in \cite{beckerIJRR2014}.
 
@@
 
-A PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+A PID controller based on the original PD controller in Eq.~\eqref{eq:PDcontrolPosition} to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}

### 2017-09-05 journalSwarmControl/Reviews/Response.tex
Updated Comments

@@
 \usepackage{todonotes}
+\usepackage{graphicx}
+\usepackage{wrapfig}
 
@@
 
+\newcommand{\letter}[2]{\vspace*{\fill}
+
+	\begin{quote}
+
+		\includegraphics[width=8cm]{engineeringLogo.eps}
+
+		\begin{flushright}
+		Aaron T Becker\\Department of Electrical Engineering\\University of Houston\\Houston, TX 77004, USA\\Email: atbecker@central.uh.edu\\Web: youTube.com/aabecker5\\
+
+		\today \vspace{2cm} \\
+		\end{flushright}
+
+		#1\medskip\\ 
+		Dear #2 Editorial Office,\medskip\\
+		Please find attached the revised paper, #1
+		along with the document containing response to the reviewers.
+We are grateful to the reviewers for helping us in improving our manuscript through their comments and questions.
+Please let us know if further information is required.\medskip\\
+		Sincerely,
+			\todo[inline]{Insert Signature}
+		Aaron T Becker (on behalf of all the authors) 
+	\end{quote}
+	\vspace*{\fill}
+	\break }
+
 \NewDocumentCommand \Reviewer { m } {
@@
 \begin{document}
-
-	\vspace*{\fill} 
-	\begin{quote}  
-		\todo[inline]{I still need to format this page to look like the one in the word document}
-
-		Steering a Particle Swarm Using Global Inputs and Swarm Statistics
-
-		Dear IEEE Transactions on Robotics Editorial Office,
-
-		Please find attached the revised paper, Steering a Particle Swarm Using Global Inputs and Swarm Statistics
-		along with the document containing response to the reviewers.
-We are grateful to the reviewers for helping us in improving our manuscript through their comments and questions.
-Please let us know if further information is required.
-
-		Sincerely,
-
-		\todo[inline]{Insert Signature}
-
-		Aaron T Becker (on behalf of all the authors) 
-	\end{quote}
-	\vspace*{\fill}
-	\break
+	\letter{Steering a Particle Swarm Using Global Inputs and Swarm Statistics}{IEEE Transactions on Robotics}
 
@@
 It seems it should be j instead of i in the loop "for i = ..." in (25).}
+	\todo[inline]{Check the equation in the book to clarify if this should be i or j before the equation}
+
 	The probabilities for set p($x_j$ | x,u) come from the reward function in equation (23).
@@
 	A sentence has been added and a second sentence has been altered for clarification, now reading, "...
-The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b.
-..."
+The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown ..."
 
@@
 
-	\Reviewer{\#2}Is there not a Reviewer 2?
+	\Reviewer{\#2}No comments were submitted from Reviewer 2.
 
@@
 It would also help a citation if it exists.}
-		\todo[inline]{Doesn't the term "tens of..." mean something of magnitude X$*10$?}
+	This has been clarified. The paper now reads "process noise cancels the differentiating effects ... large numbers of robots."
 
@@
 "[123] shows that..." instead use: "Joe and Bob [123] show that..."  }
-		\todo[inline]{Should I change all of our citations then?}
+	That's a good idea. This has been changed. %I changed a lot of our citations to be in this format but some of them seem kind of weird to change.
 
@@
 }
-	We are currently working on attempting to do just this! Look out for our future work.
+		\todo[inline]{Refer to Li's work!}
 
@@
 This demands to revise your justification as your result on variance stability conditions might be totally different.
+Please clarify.
+...
+In eq.
+(17) how can be $\dot{\sigma}^2$ be negative?
 Please clarify.}
-	\todo[inline]{I will look into this one more later, maybe.}
+	Our Lyapunov controller was frivolous as it is obvious that the variance can be controlled. This section detracted from the paper as a whole and as a result was removed entirely.
 
@@
 
-	\response{In eq. (17) how can be $\dot{\sigma}^2$ be negative? Please clarify.}
-		\todo[inline]{That.... is a good question.}
-
 	\response{Eq. (20) is a little bit vague. A clearer statement is required. Perhaps to say, eq. (20) is an extended version of eq. (11), which allows to control variance as well.}
-	Great idea! The paper now reads, "A PID controller based on the original PD controller in Eq. (11) to regulate the variance to $\sigma^2_{\rm{ref}}$ is:".
+	Good point!
+This has now been clarified in our paper, reading, "Just as we used a PD controller in Eq.
+(11) to control the robots, a PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:".
 

### 2017-09-05 journalSwarmControl/Introduction.tex
Updated Comments

@@
 This paper examines object manipulation by a swarm of particles, each actuated by the same shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
-The transportation methodology is similar to that in \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
+The transportation methodology is similar to that in Sugawara's work \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
 
@@
 Often particles are difficult or impossible to sense individually due to their size and location. 
-For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
+For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see Martel \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
 To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions.

### 2017-09-05 journalSwarmControl/expResults.tex
Changed the notations and edited
Updated Comments

@@
 Full-state requires $2n$ data points for $n$ particles.
-Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.
+Convex-hull requires at worst $2n$, but according to Har \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.
 Mean requires two, and variance three, data points.
@@
  \end{align*}
-$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).
+$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{\rm{max}}$).
 

### 2017-09-05 journalSwarmControl/experiment.tex
Updated Comments

@@
 To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
-  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}.
+  Value iteration,  as described by Thrun \cite{Thrun2005}, is used to learn an \emph{optimal policy}.
    At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
@@
 When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal but repulses the swarm from in front of the object.
+To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal but repulses the swarm from in front of the object.
 The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b.
@@
 The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
-The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them.
+Our controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them.
 This failure rate increases if  object weight increases or ground-robot friction increases.
@@
 
-The interaction between the robots and object is impulsive so, like the study of impulsive pulling in  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
+The interaction between the robots and object is impulsive so, like the study of impulsive pulling by Christensen  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
  After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success. 

### 2017-09-05 journalSwarmControl/realExperiment.tex
Updated Comments

@@
 While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot, reported in \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available as an open-source platform or commercially from \cite{K-Team2015}.
+The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots.
+It is available as an open-source platform or commercially from K-team  \cite{K-Team2015}.
 Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.

### 2017-09-05 journalSwarmControl/relatedWork.tex
Updated Comments

@@
 
-However, process noise cancels the differentiating effects of inhomogeneity for more than tens of robots.  We desire control algorithms that extend to many thousands of robots.
+However, process noise cancels the differentiating effects of inhomogeneity for large numbers of robots.  We desire control algorithms that extend to many thousands of robots.
 Limited position control was achieved by \cite{bretl2007control} and our previous work \cite{beckerIJRR2014}, but both used robots commanded in their local coordinate frame.
@@
 Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.
-Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
+Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in Odhner's work \cite{odhner2014compliant} or silicone in Deimel's work \cite{deimel2014novel}.
 Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in \S \ref{sec:VarianceControl}.

### 2017-09-05 journalSwarmControl/simulation.tex
Updated Comments

--- 2017-05-02 journalSwarmControl/simulation.tex
@@
 
-Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation, presented in \cite{catto2010box2d}.
+Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics, including collision, density, and friction, and fixed-time step simulation, presented by Catto \cite{catto2010box2d}.
 All experiments in this section ran on a Chrome web browser on a 2.6 GHz Macbook.

### 2017-09-05 journalSwarmControl/theory.tex
Updated Comments

@@
 Microscale systems are affected by unmodelled dynamics. 
-These unmodelled dynamics are dominated by Brownian noise, described by~\cite{einstein1956investigations}. 
+These unmodelled dynamics are dominated by Brownian noise, described by Einstein~\cite{einstein1956investigations}. 
 To model this~\eqref{eq:stdform} must be modified as follows:
@@
 Pushing a group of dispersed robots against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached.
-For large $n$, ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is
+For large $n$, Graham ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is
  \begin{align} \label{eq:optimalvar}
@@
 
-We will prove the goal is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
-\begin{align}
-\label{eq:LyapunovVariance}
-V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{\rm{goal}})^2,\nonumber\\
-\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{\rm{goal}})\dot{\sigma}^2(\mathbf{x}).
-\end{align}
-We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{\rm{goal}}$.
-To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
+%We will prove the goal is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
+%\begin{align}
+%\label{eq:LyapunovVariance}
+%V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{\rm{goal}})^2,\nonumber\\
+%\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{\rm{goal}})\dot{\sigma}^2(\mathbf{x}).
+%\end{align}
+%We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{\rm{goal}}$.
+%To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
+Thus, to control this variance, we choose
+
 \begin{align}\label{eq:controlVariance}
@@
 \end{align}
- For such a $u(t)$,
- \begin{align}
-\dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
-	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{\rm{goal}}, \sigma^2_{\rm{optimal}}(n,r))  \\ 
-	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{\rm{goal}},
-\end{cases} 
-\end{align} and thus
-$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$
+% For such a $u(t)$,
+% \begin{align}
+%\dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
+%	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{\rm{goal}}, \sigma^2_{\rm{optimal}}(n,r))  \\ 
+%	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{\rm{goal}},
+%\end{cases} 
+%\end{align} and thus
+%$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$
 
-A PID controller based on the original PD controller in Eq.~\eqref{eq:PDcontrolPosition} to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+Just as we used a PD controller in Eq.~\eqref{eq:PDcontrolPosition} to control the robots, a PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}
@@
  Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
- This technique satisfies control objectives that evolve at different rates as in~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
+ This technique satisfies control objectives that evolve at different rates as Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
 The process is illustrated in Fig.~\ref{fig:hysteresis}.

### 2017-09-20 journalSwarmControl/Introduction.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

@@
 These activities require robots that behave intelligently.
-Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire particle population.
-This paper examines object manipulation by a swarm of particles, each actuated by the same shared global input, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
+Limited computation and communication at small scales makes autonomous operation or direct control over individual robots difficult. 
+Instead, this paper treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
+This paper examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
 The transportation methodology is similar to that in Sugawara's work \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.

### 2017-09-20 journalSwarmControl/expResults.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

@@
 Mean requires two, and variance three, data points.
-Mean and mean + variance are convenient even with millions of particles.
+Because they do not increase with population size, mean and mean + variance are convenient even with millions of particles.
 
@@
 
-Our hypothesis predicted a steady decay in performance as the amount of visual feedback decreased.
+Our hypothesis predicted a steady decrease in performance as the amount of visual feedback decreased.
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}.b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
+As Fig.~\ref{fig:ResVaryVis}b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
 All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
@@
 Anecdotal evidence from beta-testers who played the game suggests that tracking 100 particles is overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance is like using a ``spongy'' manipulator.
-Our beta-testers described convex-hull feedback as confusing and irritating.
+However, our beta-testers described convex-hull feedback as confusing and irritating since it is not robust to outliers.
 A single particle left behind an obstacle will stretch the entire hull, obscuring the majority of the swarm.
-Similarly, our algorithm must be robust to outliers.
 %obscuring what the rest of the swarm is doing.
@@
 This experiment varied the strength of %these 
-disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
+disturbances to study how noise (disturbance inputs) affects human control of large swarms. Noise was applied at every time step:
 \begin{align*}
@@
 We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.
-Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.
+Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}a show only a 40\% increase in completion time for the maximum noise.
 This indicates swarm control is robust to IID noise.

### 2017-09-20 journalSwarmControl/experiment.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

@@
 
-This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers.
-The swarm must deliver the object to the goal region.
+This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
+%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
+We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
+%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
  To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.
@@
 The optimal policy is %the control action that, in probability, results in the largest value:
-   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j|x, \mathbf{u})].  \label{eq:OptimalPolicy}
+   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i|x, \mathbf{u})].  \label{eq:OptimalPolicy}
    \end{align}
@@
 \begin{align}
-\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
+\text{for }&\text{$j=1$ to $N$ do} \nonumber \\
 &\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
@@
 When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal but repulses the swarm from in front of the object.
-The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown in Fig.~\ref{fig:potentialField}b.
-These fields operate in place of $\mathbf{D}(\mathbf{b})$ only when the repulsive field is active.
+To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}.
+The repulsive potential field is centered at $\mathbf{b}$, the object's COM, and is active in a circular sector of angular width 2$\theta$ and radius $d_0$ aligned with $\mathbf{D}(\mathbf{b})$.
+$\mathbf{D}(\mathbf{b})$ is the desired direction of motion from~\eqref{eq:OptimalPolicy}.
 \begin{align}
-F_{\rm{att}} &= -\zeta \Delta \rho / \rho, \\
-%\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
-\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right), \\
-%\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
+\mathbf{d} &=  [\bar{x},\bar{y}] - \mathbf{b}, \qquad   \phi =\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot  \mathbf{d}  }{\norm{\mathbf{D}(\mathbf{b})} \cdot \norm{\mathbf{d}} } \right), \nonumber \\
+F_{\rm{att}} &= -\zeta \frac{\mathbf{d}  }{ \norm{\mathbf{d}}}, \nonumber\\
  F_{\rm{rep}} &=  \left\{
 \begin{array}{ll}
-      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\wedge~\phi <  \theta \\
-      0, & \textrm{ otherwise} \\
+      \eta( \frac{1}{\norm{\mathbf{d}}}- \frac{1}{d_0}) \frac{1}{\norm{\mathbf{d}}^2} \mathbf{d}, ~& \norm{\mathbf{d}} \leq d_0~\wedge~\phi <  \theta \\
+      0, & \textrm{ otherwise} \nonumber \\
 \end{array},
 \right.\\
-F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}.
+F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}, \nonumber\\
+[x_{\rm{goal}}&, y_{\rm{goal}}] =  [\bar{x},\bar{y}]  + \frac{F_{\rm{pot}}}{\norm{F_{\rm{pot}}}}.
 \end{align}
-
-In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $\rho_0 = 3$.
-Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $\rho_0 = 7.5$.
-In both cases, $\Delta \rho$ is the position of the swarm.
+Here  $\eta$, $\zeta$, and $d_0$ are all positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
+In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $d_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $d_0 = 7.5$.
 

### 2017-09-20 journalSwarmControl/theory.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes

@@
 
-Just as we used a PD controller in Eq.~\eqref{eq:PDcontrolPosition} to control the robots, a PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean robot position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}
-u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}), \nonumber\\
-u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
+u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
+u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
 \end{align}
-We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
+We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
 Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
@@
 \end{align}
- If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.
+ If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.
 
@@
  Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
- This technique satisfies control objectives that evolve at different rates as Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
+ This technique satisfies control objectives that evolve at different rates as in Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
 The process is illustrated in Fig.~\ref{fig:hysteresis}.

### 2017-10-03 journalSwarmControl/experiment.tex
New plots have been added
Added the plots

@@
 Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation.
+ We also have tested two other mazes with ``E'' shape and a ``Spiral'' without changing the algorithm to illustrate the capability of the algorithm in Fig.~\ref{fig:Emaze}.
+Fig.~\ref{fig:DifferentMazes} shows the results of all the three different mazes.
+E-shape maze has the least average time because the path to the goal is shorter.
  Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.
@@
 \caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control (see also Extension 1).
+}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}
+\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
+\vspace{0.5em}
+\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}
+\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
+\caption{\label{fig:Emaze} We have tested multiple mazes to illustrate that the algorithm works for different environments.
+}
+\end{figure}
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{DifferentShapes.pdf}\end{overpic}
+\vspace{-2em}
+\caption{\label{fig:DifferentMazes} Different maps and mazes were tested to validate the algorithm.
+E-shape maze has the least completion time average because it has the shortest path to the goal from the starting position.
 }

### 2017-10-03 journalSwarmControl/Reviews/Response.tex
added my signature

@@
 		\begin{flushright}
-		Aaron T Becker\\Department of Electrical Engineering\\University of Houston\\Houston, TX 77004, USA\\Email: atbecker@central.uh.edu\\Web: youTube.com/aabecker5\\
+		Aaron T.\ Becker\\Department of Electrical Engineering\\University of Houston\\Houston, TX 77004, USA\\Email: atbecker@central.uh.edu\\Web: youTube.com/aabecker5\\
 
@@
 		Please find attached the revised paper, #1
-		along with the document containing response to the reviewers.
+		along with the document containing a response to the reviewers.
 We are grateful to the reviewers for helping us in improving our manuscript through their comments and questions.
 Please let us know if further information is required.\medskip\\
-		Sincerely,
-			\todo[inline]{Insert Signature}
-		Aaron T Becker (on behalf of all the authors) 
+
+		Sincerely,\\
+
+			\includegraphics[width=4cm]{SignatureAaronTBeckerWhite.png}
+			\medskip\\
+		Aaron T.\ Becker (on behalf of all the authors) 
 	\end{quote}
@@
 \begin{document}
-	\letter{Steering a Particle Swarm Using Global Inputs and Swarm Statistics}{IEEE Transactions on Robotics}
+	\letter{\emph{Steering a Particle Swarm Using Global Inputs and Swarm Statistics}}{IEEE Transactions on Robotics}
 
 	\section*{Response to Reviewers}
-	In the following document, we’ve provided detailed responses to the comments and questions of the reviewers.
+	In the following document, weâ have provided detailed responses to the comments and questions of the reviewers.
 Comments and questions by reviewers are in \textcolor{blue}{blue}, our responses are in black.
@@
 	\response{What are the assumptions on the manipulated block properties? (Friction? Necessary force to move it? Is one robot enough to move it?)}
-		\todo[inline]{This is a question we were trying to avoid right?
-Since we can't seem to measure the strength of a single kilobot.
-Should we give simulation specs?
-What about the real world experiments?}
+	This is now addressed in \S VI:
+	``We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion. Increasing the number of pushing particles increases the object speed.''
 
-	\response{Regading the hybrid controller, what are the necessary assumptions to reach convergence?
+	\response{Regarding the hybrid controller, what are the necessary assumptions to reach convergence?
 Are there practical assumptions to guarantee the claim (18) =$>$ (19)?
@@
 Is it possible to control variance in this case?}
-	This is an interesting problem that we can expand on in a later work.
+	We removed the discussion on Lyapunov stability, but now point out in \S IV-D that
+	``The mean and variance of the swarm cannot be controlled simultaneously.
+However, if the variance gained while moving from a corner to the target position is less than some $\sigma_{\rm{max}}^2-\sigma_{\rm{min}}^2$, we can adopt the hybrid, hysteresis-based controller shown in Alg.~1 to regulate the mean and variance.''
 
-	\response{In section VI the authors should give more detils regarding the Value Iteration approach.
+	\response{In section VI the authors should give more details regarding the Value Iteration approach.
 How did the authors set the probabilities p($x_j$ | x,u)?
 It seems it should be j instead of i in the loop "for i = ..." in (25).}
-	\todo[inline]{Check the equation in the book to clarify if this should be i or j before the equation}
-
-	The probabilities for set p($x_j$ | x,u) come from the reward function in equation (23).
+	Thank you. The probabilities to set $p(x_j | x,u)$ come from the reward function in equation (23). 
+	We are using the same notation as Thrun et al.~[45] for hte $i$ and $j$ on page 502.
 	The loop in equation (25) is looping the probability p($x_i$ | $x_j$,u) such that for a specific $x_j$ it iterates through all $x_i$.
@@
 	\response{In algorithm 2, how is (29) applied?
-Is it used directly as ux and uy?
+Is it used directly as $u_x$ and $u_y$?
 The composition of the potential field and the plan obtained with the value iteration algorithm should be better clarified.}

### 2017-10-03 journalSwarmControl/theory.tex
added my signature

@@
 
-The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is less than the maximum controlled speed, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.
-Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds some $\sigma_{\rm{max}}^2$.
+The mean and variance of the swarm cannot be controlled simultaneously.
+However, if the variance gained while moving from a corner to the target position is less than some $\sigma_{\rm{max}}^2-\sigma_{\rm{min}}^2$, we can adopt the hybrid, hysteresis-based controller shown in Alg.~\ref{alg:MeanVarianceControl} to regulate the mean and variance.
+ Such a controller normally controls the mean position, but switches to minimizing variance if the variance exceeds $\sigma_{\rm{max}}^2$. 
  Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 

### 2017-10-03 journalSwarmControl/Reviews/Response.tex
preparing to read through the paper!

@@
 \usepackage{wrapfig}
+\usepackage{hyperref}
+\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
 
@@
 
+	\emph{Comments by Associate Editor}
+	\response{The paper proposes a theoretical contribution in the swarm robotics
+field. Controllability properties are obtained for a swarm of robots
+subject to common inputs and it results that only two variables are
+controllable. In the paper is is also shown that variance is
+stabilizable. The idea of controlling mean and variance comes from a
+large set of online tests performed by humans controlling swarms in
+manipulation activities and collected by the authors. Based on the
+theoretical results, a hybrid mean and variance control law is obtained
+to manipulate objects in a scenario with obstacles. Both simulation
+and, more importantly, experiment results are reported.}
+Thank you.
+
+\response{The paper deals with an interesting problem for the community and it is
+very well written. Some improvements may be achieved following
+reviewers comments in providing more insights or details on several
+aspects. For example, as reported by Reviewer 3, simulation scenarios
+could be of higher complexity with respect to the shown experiments so
+to prove other potentialities of the proposed approach. In addition to the corrections requested by reviewers I suggest also
+the following:}
+
+ To illustrate the flexibility of the algorithm we tested two additional workspaces, \emph{{\sffamily E}-shaped} and  \emph{Spiral}, without changing the algorithm. 
+ More complicated workspaces could be generated by composing these workspaces. See Figs. 16 and 17.
+
+\response{- Simplify equation 3 removing dots and reporting $A^3 B$ instead of
+$A^{n-1}B$ since the dimension of the system is 4. Moreover, variable $n$
+is later used as the number of robots in the swarm.}
+This has been changed.
+
+\response{- In Section V would avoid the subdivision in subsections since they
+are very short and in this case the subdivision does not help the flow
+of reading.}
+Quite right.  We changed these to paragraph headings because they
+
 	\Reviewer{\#1}\response{the paper is very interesting, well written and well organized.
@@
 It seems it should be j instead of i in the loop "for i = ..." in (25).}
-	Thank you. The probabilities to set $p(x_j | x,u)$ come from the reward function in equation (23). 
-	We are using the same notation as Thrun et al.~[45] for hte $i$ and $j$ on page 502.
-	The loop in equation (25) is looping the probability p($x_i$ | $x_j$,u) such that for a specific $x_j$ it iterates through all $x_i$.
+	We added text ``Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25.''
+	Thank you. We are now using the same notation as Thrun et al.~[45] for the $i$ and $j$ on page 502.
 
 	\response{It is also necessary to give more details of the potential fields in VI-b. The parameters $\Delta \rho$, $\rho$, b in (26) and (27) were not defined.}
-	A sentence has been added and a second sentence has been altered for clarification, now reading, "...
-The repulsive potential field is centered at the object's COM and is active when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ and $\rho$, the distance between the swarm mean and the object's COM, is less than $\rho_0$ as shown ..."
+
+	We rewrote section VI-B to define the parameters and explain the potential fields. This allowed us to simplify the notation. Each parameter is now defined.
 
 	\response{In the outlier rejection, how are the regions computed?}
-		\todo[inline]{We already state, "The \emph{main} regions are generated by extending obstacles until they meet another obstacle.
-The \emph{transfer} regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions." Are they looking for a formula?}
+		In \S VI-B, we now explain,``The \emph{main} regions are generated by extending obstacles until they meet another obstacle.
+The \emph{transfer} regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.''
 
@@
 The composition of the potential field and the plan obtained with the value iteration algorithm should be better clarified.}
-	Yes, We have amended the paragraph prior with "These fields operate in place of $\mathbf{D}(\mathbf{b})$ only when the repulsive field is active."
+	We added equation (25), which is
+	\begin{align}
+	[x_{\rm{goal}}&, y_{\rm{goal}}] =  [\bar{x},\bar{y}]  + \frac{F_{\rm{pot}}}{\norm{F_{\rm{pot}}}}.\nonumber
+\end{align}
 
 	\response{In the simulations, what is the model for collisions? }
-	We use a Javascript port called Box2D to calculate our physics. This has been clarified in Section 5.
+	We use a Javascript port called Box2D to calculate our physics. This has been clarified in \S V.
 
@@
 }
-		\todo[inline]{I thought in the first paragraph we state that our robots are the particles?}
+
+	Good point. We have edited the paper so that references to micro and nano robots are only in the introduction, related work, and conclusion. We have replaced 
+	``robot'' with ``particle'' unless the object to be controlled is a hardware robot (such as a kilobot).
+	Additionally, in paragraph 1 we now state ``Instead, this paper treats the robots as particles that are steered by a global control signal broadcast to the entire population."
 
@@
 It would also help a citation if it exists.}
-	This has been clarified. The paper now reads "process noise cancels the differentiating effects ... large numbers of robots."
+	This has been clarified by modifying text and adding references:
+	``Similarly, our previous work, [16], [17], focused on exploiting inhomogeneity between robots.
+These control algorithms theoretically apply to any number of robots, even robotic continuums.
+However, all these works never controlled more than twelve robots at a time because process noise cancels the differentiating effects of robot inhomogeneity.
+We desire control algorithms that extend to many thousands of robots.''
 
@@
 "[123] shows that..." instead use: "Joe and Bob [123] show that..."  }
-	That's a good idea. This has been changed. %I changed a lot of our citations to be in this format but some of them seem kind of weird to change.
+	That is a good idea. This has been changed as you suggested.
 
@@
 In the case of the human body where the walls could be more elastic, or even permeable, how that would affect the proposed strategy.}
-	That is a good question which we would love to observe further in a future work.
+	How exciting! Elastic walls are a realm of soft robotics that our lab has just started modeling.  Unfortunately space constraints prevent us from addressing this in the current paper.
 
 	\response{In the same line, if pushing against a "wall" could increase friction or even absorption of the transported medicine, have you consider a strategy that envelops the object with a small group of particles shielding the object from the walls?}
-	That is a good idea however we are unable to implement such an idea in the current iteration of this paper.
+	Another worthy idea worth further investigation.
+We are unable to implement such an idea in the current iteration of this paper, but please see our recent IROS 2017 paper that investigates shaping swarms using boundary friction:
+		{\bf Shiva Shahrokhi}, {\bf Arun Mahadev}, {\bf Aaron Becker} ``Algorithms for Shaping a Particle Swarm with a Shared Input by Exploiting Non-Slip Wall Contacts",  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
+28, 2017.
+	\href{	https://youtu.be/MxudWbNZdC4}{[video]}
 
@@
 }
-	That is a good idea however we are unable to implement such an idea in the current iteration of this paper.
+
+	That is a good idea.  We added two additional maps for the simulation tests to this paper (Figs. 16 and 17). Our lab has several papers in progress using swarms in curved environments.
 
@@
 }
-		\todo[inline]{Refer to Li's work!}
+	Also a good idea.  We added an E-shaped maze in (Figs. 16 and 17). Also see preliminary work in this direction in our recent IROS papers
+
+	{\bf Li Huang}, Louis Rogowski, MinJun Kim, {\bf Aaron Becker}.
+``Path Planning and Aggregation for a Microrobot Swarm in Vascular Networks Using a Global Input'',  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
+28, 2017.
+
+and
+
+ {\bf Arun Mahadev}, Dominik Michael Krupke, S{\'a}ndor Fekete, {\bf Aaron T.
+Becker}, ``Mapping and Coverage with a Particle Swarm Controlled by Uniform Inputs",  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
+28, 2017.
+		\href{https://youtu.be/77AaoEnd5b4}{[video]}
 
 	\response{Furthermore, it would be interesting to see if the proposed method will work in an open environment with multiple obstacles.}
+	The current algorithm requires corners (see Alg. 2). 
 	We would love to expand on our work in this fashion and are looking for collaborators to do so.
@@
 Please clarify.}
+
 	Our Lyapunov controller was frivolous as it is obvious that the variance can be controlled. This section detracted from the paper as a whole and as a result was removed entirely.
@@
 	\response{The global control law is presented for holonomic robots, while in small-scales often particles obey nonholonomic motion. Is your approach still applicable? }
-		\todo[inline]{I don't know and if I did, is s/he asking for us to add in the paper?}
+
+		The kilobots are nonholonomic, but the holonomic controller is still able to use them to manipulate the object.
+\S VII-B now starts with ``Even though kilobots are nonholonomic, they performed five successful runs manipulating a hexagonal object through an obstacle maze.
+Videos of these runs are in Extension 2.
+These hardware experiments represent the results of over 100 hours of trials.''
 
 	\response{Eq. (20) is a little bit vague. A clearer statement is required. Perhaps to say, eq. (20) is an extended version of eq. (11), which allows to control variance as well.}
-	Good point!
-This has now been clarified in our paper, reading, "Just as we used a PD controller in Eq.
-(11) to control the robots, a PID controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:".
+	This has now been clarified in our paper:
+
+	``Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+\begin{align}
+u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \nonumber \\
+u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \nonumber
+\end{align}
+We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.''
 
 	\response{Page 3 section 3 on the lesson list number 2, you have to define IID?}
-	This issue has been remedied entirely by changing IID to "independent and identically distributed".
+	This issue has been remedied by defining IID as "independent and identically distributed".
 
 	\response{Page 3 in section 3A, please follow paragraph style in the middle of the text.}
-		\todo[inline]{Dr Becker? I don't understand what isn't in paragraph format?}
+		We incorporated this sentence into the previous paragraph.
 
 	\response{In section 3C and D, the writing can be improved to make it easier for general readership.}
-		\todo[inline]{It doesn't give much in terms of what it lacks though.}
+		We rewrote \S III-C and D.
 
 	\response{Correct the repetition (control control) on the last line on the left in page 5.}
-	Corrected, the line now reads, "There are several techniques for breaking the symmetry of the control input to allow controlling more states, ..."
+	Corrected, the line now reads, ``There are several techniques for breaking the symmetry of the control input to allow controlling more states, \dots''.
 

### 2017-10-03 journalSwarmControl/experiment.tex
preparing to read through the paper!

--- 2017-09-20 journalSwarmControl/experiment.tex
@@
 To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
-  Value iteration,  as described by Thrun \cite{Thrun2005}, is used to learn an \emph{optimal policy}.
+  Value iteration,  as described by Thrun et al.~\cite{Thrun2005}, is used to learn an \emph{optimal policy}.
    At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
@@
 The optimal policy is %the control action that, in probability, results in the largest value:
-   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i|x, \mathbf{u})].  \label{eq:OptimalPolicy}
+   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u})].  \label{eq:OptimalPolicy}
    \end{align}
-   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
+   The value function $\hat{V}(x_i) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
 \begin{align}
-\text{for }&\text{$j=1$ to $N$ do} \nonumber \\
-&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
-\text{end}& 
+\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
+&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} [r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
+\text{end}&  %page 502
 \end{align}
-In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
+Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25.
+In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
 %Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
@@
 \end{align}
-Here  $\eta$, $\zeta$, and $d_0$ are all positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
+Here  $\eta$ and $\zeta$ are positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
 In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $d_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $d_0 = 7.5$.
@@
 Our controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them.
-This failure rate increases if  object weight increases or ground-robot friction increases.
+This failure rate increases if  object weight increases or ground-particle friction increases.
 The mean and covariance calculations \eqref{eq:meanVar} included all particles in the workspace.
@@
 The object is in region 1.
-An indicator function is applied to \eqref{eq:meanVar} so only robots inside region 1 are counted.
-This filtering increases experimental success because the mean calculation only includes nearby robots that can directly interact with the object.
+An indicator function is applied to \eqref{eq:meanVar} so only particles inside region 1 are counted.
+This filtering increases experimental success because the mean calculation only includes nearby particles that can directly interact with the object.
 When the object leaves main region 1 the state switches to transfer.
 The transfer regions are shown in Fig.\ \ref{fig:Region}c.
-The object is in transfer region 1, so only robots in transfer region 1 are included in the mean and covariance calculations.
- The robots should push the object to the left.
-Without filtering using regions, the red circle is the mean and the algorithm would instruct the robots to push the object up.
-The black circle shows the filtered mean and the algorithm instructs the robots to push the object directly left.
+The object is in transfer region 1, so only particles in transfer region 1 are included in the mean and covariance calculations.
+ The particles should push the object to the left.
+Without filtering using regions, the red circle is the mean and the algorithm would instruct the particles to push the object up.
+The black circle shows the filtered mean and the algorithm instructs the particles to push the object directly left.
 
-In 10 simulations with 100 robots, completion time without outlier rejection required 271$\pm$267 s while using outlier rejection required 90$\pm$35 s (mean$\pm$std).
+In 10 simulations with 100 particles, completion time without outlier rejection required 271$\pm$267 s while using outlier rejection required 90$\pm$35 s (mean$\pm$std).
 %\todo{report results of 10 simulations with and without regions.  Give mean and std}
@@
 \subsection{Simulation results}\label{sec:AlgObjectManipulation}
- We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
+ We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient particle density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
  The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than the object with an obstacle-free straight-line path to the object.
@@
 \begin{algorithm}
-\caption{Object-manipulation controller for a robotic swarm.}\label{alg:BlockPushing}
+\caption{Object-manipulation controller for a particle swarm.}\label{alg:BlockPushing}
 \begin{algorithmic}[1]
@@
 Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation.
+ To illustrate the flexibility of the algorithm we tested two additional workspaces, \emph{{\sffamily E}-shaped} and  \emph{Spiral}, without changing the algorithm.
+These are shown in Fig.~\ref{fig:Emaze}.
+ More complicated workspaces could be generated by composing these workspaces.
+ Fig.~\ref{fig:DifferentMazes} shows the results of all  three mazes. The {\sffamily E}-shaped maze required the least average time because the path to the goal is shorter. 
    Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.
 Each trial measured the time to deliver the object to the goal location.
-The default parameter settings used 100 robots, a normalized weight of 1, a hexagon shape, and Brownian noise (applied once each simulation step) with $W=5$.
+The default parameter settings used 100 particles, a normalized weight of 1, a hexagon shape, and Brownian noise (applied once each simulation step) with $W=5$.
 
@@
 \centering
-%\renewcommand{\figwid}{0.19\columnwidth}
-%\href{http://youtu.be/tCej-9e6-4o}{\begin{overpic}[width =\figwid]{story1.png}\put(6,15){T = 5 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{story2.png}\put(6,15){T = 12 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{story3.png}\put(6,15){T = 20 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{story4.png}\put(6,15){T = 25 s}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{story5.png}\put(6,15){T = 33 s}
-%\end{overpic}}
+
 \begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
-\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control (see also Extension 1).
+\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 particles under automatic control (see also Extension 1).
 }
@@
 
-The interaction between the robots and object is impulsive so, like the study of impulsive pulling by Christensen  \cite{christensen2016let},  adding additional robots decreases completion time, but with diminishing returns.
- After 75 robots, additional robots no longer can interact with the object and do not contribute to the task success. 
+\begin{figure}\begin{center}\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
+\vspace{.25em}
+\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
+\end{center}
+\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
+}
+\end{figure}
+\begin{figure}
+\centering
+\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\end{overpic}
+\vspace{-1em}
+\caption{\label{fig:DifferentMazes} Completion times for three workspaces.}
+\end{figure}
+
+The interaction between the particles and object is impulsive so, like the study of impulsive pulling by Christensen  \cite{christensen2016let},  adding additional particles decreases completion time, but with diminishing returns.
+ After 75 particles, additional particles no longer can interact with the object and do not contribute to the task success. 
 Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  
@@
  %Large amounts of noise caused failures, defined as trials lasting longer than 1000s.  With $W=50$, XX of 20 trials failed.  With $W=100$, XX of 20 trials failed.
-The robots have limited force, so increasing the object weight increases completion time.  
+The particles have limited force, so increasing the object weight increases completion time.  
 Each shape was designed to have the same mass and area.
@@
 \vspace{-0.5em}
-\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of robots, b) different noise values, c) object weight, and d) object shape.
+\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.
 Each bar is labelled with the number of trials.
@@
 \end{figure*}
-
-%Algorithm \ref{alg:BlockPushing} is an imperfect solution and has a failure mode if the robot swarm becomes multi-modal with modes separated by an obstacle, as shown in Fig.~\ref{fig:Failure}.
-In this case, moving toward a corner will never reduce the variance below $\sigma_{min}^2$.
-
-%  The first challenge is to identify when the distribution has become multi-modal.
-Measuring just the mean and variance is insufficient to determine if a distribution is no longer unimodal, but if the swarm is being directed to a corner, and the variance does not reduce below $\sigma_{min}^2$, the swarm has become separated.
-In this case, we must either manipulate with a partial swarm, or run a gathering algorithm.
-For the  {\sffamily S}-shaped workspace in this study, an open-loop input that commands the swarm to move in succession \{{\sc west, north, east, south}\} will move the swarm to the bottom right corner.
-%This is not true for all obstacle fields. In a {\sffamily T}-shaped workspace, it is not possible to find an open-loop input that will move the entire swarm to the bottom of the {\sffamily T}.
-
-%  Using only the mean and variance may be overly restrictive.
-Many heuristics using high-order moments have been developed to test if a distribution is multimodal~\cite{haldane1951simple}.
-Often the sensor data itself, though it may not resolve individual robots, will indicate multi-modality.
-For instance CCD images reveal clusters of bacteria, and MRI scans show agglomerations of particles~\cite{stuber2007positive}.
-This data can be fitted with $k$-means or expectation maximization algorithms, and manipulation could be performed with the nearest swarm of sufficient size.

### 2017-10-03 journalSwarmControl/realExperiment.tex
preparing to read through the paper!

@@
 While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots.
+The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a nonholonomic, low-cost robot designed for testing collective algorithms with large numbers of robots.
 It is available as an open-source platform or commercially from K-team  \cite{K-Team2015}.
@@
 
-The kilobots performed five successful runs manipulating a hexagonal object through an obstacle maze.
+Even though kilobots are nonholonomic, they performed five successful runs manipulating a hexagonal object through an obstacle maze.
 Videos of these runs are in Extension 2.

### 2017-10-03 journalSwarmControl/theory.tex
preparing to read through the paper!

@@
 We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
-First, assume a noiseless system containing one robot with mass $m$.
+First, assume a noiseless system containing one particle with mass $m$.
  Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
@@
 \begin{align}
-\mathcal{C} &= [ B, AB, A^2B, ... , A^{n-1}B ].\\
+\mathcal{C} &= [ B, AB, A^2B, A^{3}B ].\\
 \textrm{Here }
@@
 and thus all four states are controllable.
-This section starts by proving independent position control of many robots is not possible, but the mean position can be controlled.
-We then provide conditions under which the variance of many robots is also controllable.
+This section starts by proving independent position control of many particles is not possible, but the mean position can be controlled.
+We then provide conditions under which the variance of many particles is also controllable.
 
@@
 In this model, a single particle is fully controllable.
-For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis.
+For holonomic particles, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis.
 Given $n$ particles to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
@@
 This means \emph{any} number of particles controlled by a global command have just two controllable states in each axis.
-We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.
+We cannot arbitrarily control the position and velocity of two or more particles, but have options on which states to control.
 %One option is to control the position and velocity of the $j^{th}$ robot.
@@
 
-We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
+We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each particle:
 \begin{align}
@@
 
-The variance, $\sigma_x^2,\sigma_y^2$, of $n$ robots' position is computed as:
+The variance, $\sigma_x^2,\sigma_y^2$, of $n$ particles' position is:
 \begin{align}\label{eq:meanVar}
@@
 
-If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
+If particles with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
 \begin{align}
@@
  A flat obstacle can be used to decrease variance.
-Pushing a group of dispersed robots against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached.
+Pushing a group of dispersed particles against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached.
 For large $n$, Graham ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is
@@
 
-Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean robot position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}
@@
 We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
-Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
+Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
 \begin{align}
@@
 A key challenge is to select proper values for $\sigma_{\rm{min}}^2$ and $\sigma_{\rm{max}}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
-The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values: 
+The random packings generated by pushing our particles into corners are suboptimal, so we choose the conservative values: 
 %shown in Fig.~\ref{fig:hysteresis}:

### 2017-10-03 journalSwarmControl/Reviews/Response.tex
Handing over to Shiva

@@
 "[123] shows that..." instead use: "Joe and Bob [123] show that..."  }
-	That is a good idea. This has been changed as you suggested.
+	That is a good idea, but made the paper over the page limit.  Instead we have worked to minimize the number of times references are used as nouns.
 

### 2017-10-03 journalSwarmControl/conclusion.tex
Handing over to Shiva

--- 2017-04-27 journalSwarmControl/conclusion.tex
@@
 
-Inspired by the three lessons from swarmcontrol.net, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
+Inspired by the three lessons from \href{www.swarmcontrol.net}{SwarmControl.net}, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
 We developed a hysteresis based controller to regulate the position and variance of a swarm.

### 2017-10-03 journalSwarmControl/expResults.tex
Handing over to Shiva

@@
 We designed six experiments to investigate human control of large swarms for manipulation tasks.
-Screenshots of representative experiment are shown in Fig.~\ref{fig:5experiments}.
-Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise.%, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
+Screenshots of representative experiments are shown in Fig.~\ref{fig:5experiments}.
+Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise.
+%, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks.
  The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.

### 2017-10-03 journalSwarmControl/relatedWork.tex
Shiva and Aaron worked on the reviewers' comments.
Only thing left is E and snail mazes
Handing over to Shiva

--- 2017-06-05 journalSwarmControl/relatedWork.tex
@@
 These control algorithms theoretically apply to any number of robots, even robotic continuums.
-
-However, process noise cancels the differentiating effects of inhomogeneity for more than tens of robots.  We desire control algorithms that extend to many thousands of robots.
+However, all these works never controlled more than twelve robots at a time because process noise cancels the differentiating effects of robot inhomogeneity.
+%However, a consequence of the central limit theorem is that independent process noise cancels the differentiating effects of inhomogeneity for large numbers of robots. 
+ We desire control algorithms that extend to many thousands of robots.
 Limited position control was achieved by \cite{bretl2007control} and our previous work \cite{beckerIJRR2014}, but both used robots commanded in their local coordinate frame.
-Our new submission focuses on a more common paradigm: particles commanded in a global coordinate frame
+Our new submission focuses on a more common paradigm: particles commanded in a global coordinate frame.
 
@@
 % (Lemmings is an example of this as well, for systems with autonomy
-Their research investigated robots with 3 levels of autonomy.  We use robots without autonomy, corresponding with their first-level robots.
+Their research investigated robots with three levels of autonomy.  We use robots without autonomy, corresponding with their first-level robots.
 
@@
 Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.
-Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.
+Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in~\cite{odhner2014compliant} or silicone in~\cite{deimel2014novel}.
 Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in \S \ref{sec:VarianceControl}.

### 2017-10-03 journalSwarmControl/simulation.tex
bib file is changed

@@
 All experiments in this section ran on a Chrome web browser on a 2.6 GHz Macbook.
-\href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
+% \href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
 
-\subsection{Controlling the mean position}
+\paragraph{Controlling the mean position}
 \begin{figure}
@@
 \end{overpic}
-%\begin{overpic}[width =0.49\columnwidth]{meanVariance4.eps} 
-%\end{overpic} 
-\vspace{-1em} 
+\vspace{-2em} 
 \caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ particles.
-%b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom).
-For ease of analysis, only $x$ position and variance are shown.
-%\vspace{-2em}
 }
@@
 
-%\begin{figure}
-%\centering
-%\begin{overpic}[width = \columnwidth* 2/3 ]{meanVideoSnapShot.png}
-%\end{overpic}
-%\vspace{0 em}
-%\caption{\label{fig:meanVideo} A frame from video attachment showing mean position control on a swarm of 200 robots. The mean position of the swarm traces ``SWARM"~\cite{ShivaVideo2015}. 
-%%\vspace{-2em}
-%}
-%\end{figure}
-
-\subsection{Controlling the variance}
+\paragraph{Controlling the variance}
 \begin{figure}
@@
 
-%cite the control law, explain experiment (number of robots, maximum speed, ).
+Variance control uses the control law \eqref{eq:PDcontrolVariance} with $K_{p,i,d} = [4,1,1]$. 
+ Results are in Fig.~\ref{fig:varyBrownian}.
 
-For variance control we use the control law \eqref{eq:PDcontrolVariance}. 
- Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.
-
-%\todo{image showing control x variance and y-variance out of phase}
-
-\subsection{Hybrid control of mean and variance}
+\paragraph{Hybrid control of mean and variance}
 
@@
 \end{figure}
-
-%\begin{figure}
-%\centering
-%\begin{overpic}[width = \columnwidth * 2/3]{VideoSnapShot.png}
-%\end{overpic}
-%\vspace{-1em}
-%\caption{\label{fig:videoVar} A frame from video, using Alg.~\ref{alg:MeanVarianceControl} to control variance and mean position of a swarm of 200 robots~\cite{ShivaVideo2015}.
-%%\vspace{-2em}
-%}
-%\end{figure}

### 2017-10-03 journalSwarmControl/expResults.tex

@@
 
-%\begin{figure}[b!]
-%\renewcommand{\figwid}{0.5\columnwidth}
-%\begin{overpic}[width =\figwid]{ResVaryVis.pdf}\end{overpic}
-%\vspace{-1em}
-%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.  Players performed better with limited feedback.
-%%\vspace{-1em}
-%}
-%\end{figure}
-
-% Additionally, as population characteristics, they are available even if only a percentage of the robots are detected each control cycle.
-%Photobleaching: http://www.piercenet.com/browse.cfm?fldID=4DD9D52E-5056-8A76-4E6E-E217FAD0D86B
-%
-%Photobleaching is caused by the irreversible destruction of fluorophores due to either the prolonged exposure to the excitation source or exposure to high-intensity excitation light.
-Photobleaching can be minimized or avoided by exposing the fluor(s) to the lowest possible level of excitation light intensity for the shortest length of time that still yields the best signal detection; this requires optimization of the detection method using high sensitivity CCD cameras, high numerical aperture objective and/or the widest bandpass emission filter(s) available.
-Other approaches include using fluorophores that are more photostable than traditional fluorophores and/or using antifade reagents to protect the fluor(s) against photobleaching.
-Steps to avoid photobleaching are not feasible for all detection methods and should be optimized for each method used.
-For example, antifade reagents are toxic to live cells, and therefore they can only be used with fixed cells or tissue.
-Furthermore, some detection methods, such as flow cytometry, normally do not require steps to avoid photobleaching because of the extremely short exposure time of the fluorophore to the excitation source.
-
-%\begin{figure}
-%\centering
-%\begin{overpic}[width = \columnwidth]{ResVaryVis.pdf}\end{overpic}
-%\vspace{-2em}
-%\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}.
-Surprisingly, players perform better with limited feedback--subjects with only the mean + variance  outperformed all others.
-%\vspace{-2em}
-%}
-%\end{figure}
-
 Our hypothesis predicted a steady decrease in performance as the amount of visual feedback decreased.
 Our experiment indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.
-As Fig.~\ref{fig:ResVaryVis}b shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
-All experiments lasting over 300s were removed, under the assumption that the user stopped playing.
+As Fig.~\ref{fig:ResVaryVis} shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].
+All experiments lasting over 300 s were removed, under the assumption that the user stopped playing.
 Using ANOVA analysis, we rejected the null hypothesis that all visualization methods are equivalent, with $p$-value $2.69\times10^{-19}$.
@@
 This experiment varied the strength of %these 
-disturbances to study how noise (disturbance inputs) affects human control of large swarms. Noise was applied at every time step:
+disturbances to study how noise (disturbance inputs) affects human control of large swarms. Noise was applied at every time step as follows:
 \begin{align*}
@@
  \end{align*}
-$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{\rm{max}}$).
+Here, $m_i$ and $\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{\rm{max}}$).
 
@@
 This indicates swarm control is robust to IID noise.
-
-%\paragraph{Position control}
-%This online experiment examined how completion time scales with the number of particles $n$.
-Using a single square obstacle, users arranged $n\in[1,10]$ particles into a specified goal pattern.
-The goal pattern formed a block  {\sffamily A} with 10 particles, and lesser numbers of particles used a subset of these goal positions.
-Our hypothesis was that completion time would increase linearly with the number of particles, as with our position control algorithm in \cite{Becker2013b}.
-Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.
-Though the number of particles presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
-%
-
-%Time increases for $n>4$ particles.
-For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.
-A better experiment design would randomly place the goal positions.
-Initially we tried this, but beta-testers strongly disliked trying to arrange particles in random patterns.
-
-%\paragraph{Varying control--Assembly}
-%Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
-%
-%We compared attractive and repulsive control with the global control used for the other experiments.
-The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.\ \cite{Khalil2013} and biological models, e.g.
-\cite{goodrich2012types}.
-STMs can be used to arrange atoms and make small assemblies, as described in \cite{avouris1995manipulation}.
-An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules.
-In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
-%The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.
-%
-%\todo{describe controller model here, use an equation}
-
-%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.
-The median time using repulsive control was four times longer than with attractive control.
-%Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
-%
-%%\begin{figure}
-%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
-%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
-%%\caption{
-%%\label{fig:attractiveForce}
-%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
-%%}
-%%\end{figure}
-
-%\begin{figure}[b!]
-%\renewcommand{\figwid}{0.45\columnwidth}
-%\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
-%\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
-%\vspace{-.5em}
-%\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type.
-Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid.
-Right: in the foraging test, global control resulted in the shortest completion time and attractive the longest.
-%%\vspace{-1em}
-%}
-%\end{figure}
-%
-%\paragraph{Varying control--Foraging}
-%Collecting and delivering resources is necessary for drug delivery.
-%This experiment also compared attractive and repulsive control with the global control used for the other experiments.
-The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region.
-The particles encapsulate the particles on contact.
-%
-%%\todo{describe controller model here, use an equation}
-%
-%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.
-Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $2.96\times10^{-6}$.
-%
-%

### 2017-10-03 journalSwarmControl/experiment.tex

@@
 To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
-  Value iteration,  as described by Thrun et al.~\cite{Thrun2005}, is used to learn an \emph{optimal policy}.
+  Value Iteration is used to learn an \emph{optimal policy}~\cite{Thrun2005}.
    At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
@@
 The optimal policy is %the control action that, in probability, results in the largest value:
-   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u})].  \label{eq:OptimalPolicy}
+   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}  \left[ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u}) \right].  \label{eq:OptimalPolicy}
    \end{align}
@@
 \text{for }&\text{$i=1$ to $N$ do} \nonumber \\
-&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} [r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
+&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
 \text{end}&  %page 502
@@
 $\mathbf{M}_{\rm{BFS}}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while Policy Iteration required 90$\pm$35 s (mean$\pm$std).
+In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while Value Iteration required 90$\pm$35 s (mean$\pm$std).
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.
-To fix this, we implement a potential field approach inspired by Spong \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}.
+To fix this, we implement a potential field approach that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}.
+The approach is similar to \cite{spong2008robot}, chapter 5.
 The repulsive potential field is centered at $\mathbf{b}$, the object's COM, and is active in a circular sector of angular width 2$\theta$ and radius $d_0$ aligned with $\mathbf{D}(\mathbf{b})$.
@@
 The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
-Our controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them.
+Our controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, in each failure some particles were unable to reach the object because workspace obstacles were blocking them.
 This failure rate increases if  object weight increases or ground-particle friction increases.
@@
 \subsection{Simulation results}\label{sec:AlgObjectManipulation}
- We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient particle density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}.
+ We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient particle density to move the object by switching to minimize variance whenever variance exceeds a set limit:  $0.003 W$ and $0.006 W$ were added to the min and max variance limits from \eqref{eq:sigMaxMin}, where $W$ is the magnitude of the Brownian noise.
  The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than the object with an obstacle-free straight-line path to the object.
@@
 \State Compute optimal policy for object, according to \S \ref{subsec:objectpolicy}
-\While{is not in goal region}
+\While{ $\mathbf{b}$ is not in goal region}
 \State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
@@
 \end{center}
-\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
+\caption{\label{fig:Emaze} We tested three workspaces. The control policies from Value Iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
 }
@@
 
-The interaction between the particles and object is impulsive so, like the study of impulsive pulling by Christensen  \cite{christensen2016let},  adding additional particles decreases completion time, but with diminishing returns.
- After 75 particles, additional particles no longer can interact with the object and do not contribute to the task success. 
+The interaction between the particles and object is impulsive so, like the study of impulsive pulling in \cite{christensen2016let},  adding additional particles decreases completion time, but with diminishing returns.
+ The effect of adding particles diminishes asymptotically because additional particles have difficulty interacting with the object. 
 Brownian noise adds stochasticity.  This randomness can break the object free if it is stuck, but diminishes the effect of the control input.  

### 2017-10-03 journalSwarmControl/realExperiment.tex
Handing over to Shiva

@@
 While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection.
-The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a nonholonomic, low-cost robot designed for testing collective algorithms with large numbers of robots.
-It is available as an open-source platform or commercially from K-team  \cite{K-Team2015}.
+The kilobot is a nonholonomic, low-cost robot designed for testing collective algorithms with large numbers of robot~\cite{Rubenstein2012,rubenstein2014programmable}.
+It is available as an open-source platform or commercially~\cite{K-Team2015}.
 Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.
@@
 \subsection{Environmental setup}  
-In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and lights: four 50W LED floodlights  at the corners and four 30W LED floodlights on the sides of a 6 m square centered on the workspace and 1.5 m above the table.
-An Arduino Uno connected to an 8-relay shield controlled the lights.
+In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots and a 1.5 m$\times$1.2 m whiteboard as the workspace. 
+LED floodlights were placed 1.5 m above the table on the sides and corners of a square with 6 m sides. An Arduino Uno connected to an 8-relay shield controlled the lights.
 
 Above the table, an overhead machine vision system tracks the swarm.
-The vision system identifies obstacles by color segmentation, determines the corners  (used to decrease  variance), the object by color segmentation, and identifies robots using color segmentation and circle detection with a circular Hough transform.
+The vision system identifies obstacles and the object by color segmentation, determines the corners of the maze, and identifies robots using a circular Hough transform.
 
 \begin{figure*}
-\renewcommand{\figwid}{6cm}
+\renewcommand{\figwid}{5.25cm}
 \begin{center}
@@
 	\begin{overpic}[height=\figwid]{SetUp.pdf}%\put(1,75){A}
-	\end{overpic}\begin{overpic}[height=\figwid]{AllShapesR.pdf}%\put(-0,85){B}
+	\end{overpic}\begin{overpic}[height=5.5cm]{AllShapesR.pdf}%\put(-0,85){B}
 	\end{overpic}\begin{overpic}[height=\figwid]{PotentialExp.pdf}%\put(-0,85){C}
@@
 254 cm$^2$ circle, 
-and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.
-The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2016blocksimulations}.
+and a 486 cm$^2$ %(18 cm$\times$27 cm) 
+ rectangle, all shown in Fig.~\ref{fig:setup}.
+The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository~\cite{Shahrokhi2016blocksimulations}.
 
@@
 \begin{center}
-	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(6,30){\emph{t}= 90 s}\put(55,30){\emph{t}= 175 s}\end{overpic}
+	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(6,30){\emph{t} = 90 s}\put(55,30){\emph{t} = 175 s}\end{overpic}
 %	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
@@
 The moveable object was placed in the lower center of the workspace.
-{\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
+{\sc Matlab} code for vision processing, the Value Iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
 Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

### 2017-10-03 journalSwarmControl/theory.tex

@@
 First, assume a noiseless system containing one particle with mass $m$.
- Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
+ Our inputs are global forces $[u_x,u_y]^\top$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
 The state-space representation in standard form is: 
@@
 
-We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
+We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}:
 \begin{align}
@@
 Microscale systems are affected by unmodelled dynamics. 
-These unmodelled dynamics are dominated by Brownian noise, described by Einstein~\cite{einstein1956investigations}. 
+These unmodelled dynamics are dominated by Brownian noise, as described in~\cite{einstein1956investigations}. 
 To model this~\eqref{eq:stdform} must be modified as follows:
@@
 Pushing a group of dispersed particles against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached.
-For large $n$, Graham ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is
+For large $n$, Graham and Sloane~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is
  \begin{align} \label{eq:optimalvar}
@@
 We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
-Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
+Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position such that the swarm, if uniformly distributed between 0 and $\ell$, would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
 \begin{align}
-x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
+x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
 \end{align}
@@
  Variance is reduced until less than $\sigma_{\rm{min}}^2$, then control again regulates the mean position. 
- This technique satisfies control objectives that evolve at different rates as in Kloetzer's work ~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
+ This technique satisfies control objectives that evolve at different rates as in~\cite{kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes.
 The process is illustrated in Fig.~\ref{fig:hysteresis}.
@@
 \Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{\rm{min}}, x_{\rm{max}}, y_{\rm{min}}, y_{\rm{max}}\}$, and the target mean position $[x_{\rm{target}},y_{\rm{target}}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
-\State $x_{goal} \gets  x_{\rm{target}}$, $y_{\rm{goal}} \gets y_{\rm{target}}$
+\State $x_{\rm{goal}} \gets  x_{\rm{target}}$, $y_{\rm{goal}} \gets y_{\rm{target}}$
 \Loop

### 2017-10-03 journalSwarmControl/Reviews/Response.tex
updated the capitalization value iteration

@@
 		#1\medskip\\ 
+		\medskip\\
 		Dear #2 Editorial Office,\medskip\\
-		Please find attached the revised paper, #1
+		Please find attached the revised paper, #1,
 		along with the document containing a response to the reviewers.
-We are grateful to the reviewers for helping us in improving our manuscript through their comments and questions.
+We are grateful to the reviewers for helping us improve our manuscript through their comments and questions.
 Please let us know if further information is required.\medskip\\
@@
 of reading.}
-Quite right.  We changed these to paragraph headings because they
+Quite right.  We changed these subdivisions to be inline paragraph headings.
 
@@
 
-	\response{In page two, column 1 , the authors state that "process noise cancels the differentiating effects...
+	\response{In page two, column 1, the authors state that "process noise cancels the differentiating effects...
 tens of robots." The term tens of robots is not clear.
@@
 	This has been clarified by modifying text and adding references:
-	``Similarly, our previous work, [16], [17], focused on exploiting inhomogeneity between robots.
+	``Similarly, our previous work focused on exploiting inhomogeneity between robots [16], [17].
 These control algorithms theoretically apply to any number of robots, even robotic continuums.

### 2017-10-03 journalSwarmControl/experiment.tex
updated the capitalization value iteration

@@
 To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
-  Value Iteration is used to learn an \emph{optimal policy}~\cite{Thrun2005}.
+  Value iteration is used to learn an \emph{optimal policy}~\cite{Thrun2005}.
    At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.
@@
 $\mathbf{M}_{\rm{BFS}}$ and the value function are shown in Fig.~\ref{fig:BFSGradient}. 
-In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while Value Iteration required 90$\pm$35 s (mean$\pm$std).
+In 10 simulations with 100 particles, pushing the object to goal using BFS required 183$\pm$179 s while value iteration required 90$\pm$35 s (mean$\pm$std).
 %\todo{(report results of 10 simulations with BFS and with PI.  Give mean and std)}
@@
 \end{center}
-\caption{\label{fig:Emaze} We tested three workspaces. The control policies from Value Iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
+\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
 }

### 2017-10-03 journalSwarmControl/Reviews/Response.tex
trying to compress

@@
 
-	\response{In the same line, if pushing against a "wall" could increase friction or even absorption of the transported medicine, have you consider a strategy that envelops the object with a small group of particles shielding the object from the walls?}
+	\response{In the same line, if pushing against a  ``wall'' could increase friction or even absorption of the transported medicine, have you consider a strategy that envelops the object with a small group of particles shielding the object from the walls?}
 	Another worthy idea worth further investigation.
 We are unable to implement such an idea in the current iteration of this paper, but please see our recent IROS 2017 paper that investigates shaping swarms using boundary friction:
-		{\bf Shiva Shahrokhi}, {\bf Arun Mahadev}, {\bf Aaron Becker} ``Algorithms for Shaping a Particle Swarm with a Shared Input by Exploiting Non-Slip Wall Contacts",  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
+		{ Shiva Shahrokhi}, { Arun Mahadev}, { Aaron Becker} ``Algorithms for Shaping a Particle Swarm with a Shared Input by Exploiting Non-Slip Wall Contacts",  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
 28, 2017.
@@
 }
-	Also a good idea.  We added an E-shaped maze in (Figs. 16 and 17). Also see preliminary work in this direction in our recent IROS papers
+	Also a good idea.  We added an E-shaped maze in (Figs. 16 and 17). Also see preliminary work in this direction in our recent IROS papers:
 
-	{\bf Li Huang}, Louis Rogowski, MinJun Kim, {\bf Aaron Becker}.
+	{ Li Huang}, Louis Rogowski, MinJun Kim, { Aaron Becker}.
 ``Path Planning and Aggregation for a Microrobot Swarm in Vascular Networks Using a Global Input'',  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
@@
 
- {\bf Arun Mahadev}, Dominik Michael Krupke, S{\'a}ndor Fekete, {\bf Aaron T.
+ { Arun Mahadev}, Dominik Michael Krupke, S{\'a}ndor Fekete, {Aaron T.
 Becker}, ``Mapping and Coverage with a Particle Swarm Controlled by Uniform Inputs",  in IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, Canada, Sept.
@@
 
-	Our Lyapunov controller was frivolous as it is obvious that the variance can be controlled. This section detracted from the paper as a whole and as a result was removed entirely.
+	Our Lyapunov controller was frivolous as it is obvious that the variance can be controlled.
+This section detracted from the paper as a whole and as a result was removed entirely.
+While the swarm variance $\sigma^2$ is always positive, the derivative of the variance can be positive or negative, depending on if the variance is growing or decreasing.
 
@@
 
-	``Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
+	``Similar to the PD controller in Eq. (11) that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
 \begin{align}
@@
 	\response{Page 3 section 3 on the lesson list number 2, you have to define IID?}
-	This issue has been remedied by defining IID as "independent and identically distributed".
+	This issue has been remedied by defining IID as ``independent and identically distributed".
 

### 2017-10-26 2017-tro-SteeringSwarmParticles/Introduction.tex
made a clean directory.  Need to shrink figures

--- 2017-09-20 journalSwarmControl/Introduction.tex
@@
 This paper examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
-The transportation methodology is similar to that in Sugawara's work \cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
+The transportation methodology is similar to that in~\cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
 
@@
 Often particles are difficult or impossible to sense individually due to their size and location. 
-For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see Martel \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
+For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
 To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions.
@@
 We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.
- We conclude with implementations of these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}.
+ We implement these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}, and conclude in \S \ref{sec:conclusion}.

### 2017-10-26 2017-tro-SteeringSwarmParticles/SwarmControlJournal.tex
made a clean directory.  Need to shrink figures

--- 2017-03-22 journalSwarmControl/TASE/SwarmControlIJRR.tex
@@
 % compress:
-% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=CompressedSwarmControlIJRR.pdf SwarmControlIJRR.pdf
+% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=CompressedSwarmControlJournal.pdf SwarmControlJournal.pdf
 %%%%%%%%% IJRR SETTINGS
@@
 \usepackage{multicol}
-\usepackage[bookmarks=true]{hyperref}
 
@@
 \usepackage{transparent}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \hypersetup{
@@
   urlcolor = black,
-  linkcolor = black
+  linkcolor = black,
+  citecolor = black
 }
@@
 
-%\usepackage{xfrac}
 \usepackage{nicefrac}
@@
 \usepackage{colortbl}
-%\usepackage{caption}
-
-%\usepackage{wrapfig}
 \usepackage{overpic}
-%\usepackage{subfigure}
-%\usepackage{textcomp}
-\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
+\graphicspath{{./}}
 \usepackage{breqn} %for breaking equations automatically
@@
 \usepackage{algpseudocode}
-%\usepackage{algorithmic}
 \usepackage{multirow}
@@
 \usepackage{authblk}
+\usepackage{flushend}
 
@@
 \newcommand{\D}{\nobreakdash-\textsc{d}}
-%\newcommand{\Fspace}{\mathcal{F}}
 \providecommand{\Fspace}{\Cspace_\text{free}}
@@
 
-\title{\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics}
+\title{\LARGE \bf Steering a Swarm of Particles\\ Using Global Inputs and Swarm Statistics}
 
-\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T. Becker% end author block 
+\author{Shiva Shahrokhi, Lillian Lin, Chris Ertel, Mable Wan,  Aaron T. Becker% end author block 
+\thanks{This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{[IIS-1553063]}.}% <-this % stops a space
 \thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2, atbecker\}@uh.edu}}
@@
 The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
-We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures.
-We present an alternative.
 
-This paper first examines the perils, pitfalls, and possibilities we discovered by launching \href{http://www.swarmcontrol.net}{SwarmControl.net}, an online game where players steer swarms of up to 500 robots to complete manipulation challenges.
-We record statistics from thousands of players, and use the game to explore aspects of large-population robot control.
-We present the game framework as a new, open-source tool for large-scale user experiments.
-One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the swarm than with full-state feedback.
-Inspired by human operators, this paper next investigates controllers that use only the mean and variance of the swarm.
-We prove that the mean position is controllable, then provide conditions under which variance is controllable.
-We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the robot distribution.
-Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.
+Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
+To identify the key parameters for particle manipulation, we used a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges.
+We recorded statistics from over ten thousand players.
+Inspired by techniques where human operators performed well, we investigate controllers that use only the mean and variance of the swarm.
+We prove the mean position is controllable and provide conditions under which variance is controllable.
+We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution.
+Finally, we employ these controllers as primitives for an object manipulation task and implement all controllers on 100 kilobots controlled by the direction of a global light source.
 \end{abstract}
 
-%\keywords{micro-robot, global control, swarm, human-swarm interaction}
+%  TRO       Manipulation Planning, 	Underactuated Robots,   human-swarm interaction,   swarm
+%  maybe TRO	Manipulation and Compliant Assembly, 	Path Planning for Multiple Mobile Robot Systems
 
-%\input{00-Abstract.tex}
+%Previously presented*  
+% [1] A.
+T.
+Becker, C.
+Ertel, and J.
+McLurkin, ÒCrowdsourcing swarm manipulation experiments: A massive online user study with large swarms of simple robots,Ó in IEEE International Conference on Robotics and Automation (ICRA), 2014, pp.
+2825Ð2830.
+% [2] S. Shahrokhi and A. T. Becker, ÒStochastic swarm control with global inputs,Ó in IEEE/RSJ International Conference on  Intelligent Robots and Systems (IROS), Sep. 2015, pp. 421Ð427.
+
 \input{Introduction.tex}
@@
 \input{conclusion.tex}
-
-%\begin{dci}
-%The Authors declare that there is no conflict of interest.
-%\end{dci}
-%\begin{funding}
-%%RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
-%This work was supported by the National Science Foundation under Grant No.\ 
-%\href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
-%\end{funding}
-%\newpage
-%\begin{sm}
-%\begin{itemize}
-%\item Extension 1:  simulation video of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
-%\item Extension 2: hardware experiment videos of object manipulation with a swarm of 100 kilobots.
-%\end{itemize}
-%\end{sm}
-
-%\input{11-MultiMediaAppendix.tex}
+\section*{Acknowledgements}\label{sec:Acknowledgements}
+We thank our anonymous reviewers and Dylan Shell for their substantive comments which improved the presentation and content of this paper.
 
 \bibliographystyle{IEEEtran}
-%\bibliographystyle{SageH}
-\bibliography{IEEEabrv,SwarmControlIJRR}
+\bibliography{IEEEabrv,SwarmControlJournal}
 

### 2017-10-26 2017-tro-SteeringSwarmParticles/conclusion.tex
made a clean directory.  Need to shrink figures

--- 2017-10-03 journalSwarmControl/conclusion.tex
@@
 To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
-Analysis of the game play results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.
+Analysis of the gameplay results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.
 
 Inspired by the three lessons from \href{www.swarmcontrol.net}{SwarmControl.net}, this paper designed controllers and controllability results using only the mean and variance of a particle swarm. 
-We developed a hysteresis based controller to regulate the position and variance of a swarm.
-We designed a controller for object manipulation using policy iteration for path planning, regions for outlier rejection, and potential fields for minimizing moving the object backwards.
+We developed a hysteresis-based controller to regulate the position and variance of a swarm.
+We designed a controller for object manipulation using value iteration for path planning, regions for outlier rejection, and potential fields for minimizing moving the object backwards.
 All automatic controllers were implemented using 100 kilobots steered by the direction of a global light source.
@@
 Our future goal is to perform assembly using particle swarms to manipulate and attach components.
-This task requires controlling the position and orientation of components, manipulating them through obstacles and other components, and applying force and torque to components.
+This task requires applying force and torque to components and manipulating them through obstacles and each other.
 This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.
 
-Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, optimal-control, and targeted drug delivery in a vascular network.
+Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network.
 

### 2017-10-26 2017-tro-SteeringSwarmParticles/realExperiment.tex
made a clean directory.  Need to shrink figures

--- 2017-10-03 journalSwarmControl/realExperiment.tex
@@
 The moveable object was placed in the lower center of the workspace.
-{\sc Matlab} code for vision processing, the Value Iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
+{\sc Matlab} code for vision processing, the value iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
 Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

### 2017-10-26 2017-tro-SteeringSwarmParticles/simulation.tex
made a clean directory.  Need to shrink figures

--- 2017-10-03 journalSwarmControl/simulation.tex
@@
 
-Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics, including collision, density, and friction, and fixed-time step simulation, presented by Catto \cite{catto2010box2d}.
+Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics, including collision, density, and friction, and fixed time step simulation~\cite{catto2010box2d}.
 All experiments in this section ran on a Chrome web browser on a 2.6 GHz Macbook.
-% \href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
+\href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2016blocksimulations}.
 
@@
 Representative experiments are shown in Fig.~\ref{fig:gainvalues}.
-100 particles were used and the maximum speed was 3 meters per second.
+100 particles were used and the maximum speed was 3 $m/s$.
 As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$.

### 2017-10-26 2017-tro-SteeringSwarmParticles/experiment.tex
switched to jpg, made compressed version (less than 5 mB!)

--- 2017-10-03 journalSwarmControl/experiment.tex
@@
 \centering
-\begin{overpic}[width=1\columnwidth]{PotentialField.pdf}\end{overpic}
+\begin{overpic}[width=1\columnwidth]{PotentialField.jpg}\end{overpic}
 %\todo{I like the 'target' symbol, but it is not self-documenting.
@@
 	\begin{overpic}[width=0.9\columnwidth]{StateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
-	~~\begin{overpic}[width=0.45\columnwidth]{MainRegions.pdf} %,grid=true
+	~~\begin{overpic}[width=0.44\columnwidth]{MainRegions.pdf} %,grid=true
 	\put(-6,85){\textbf{b}}
@@
 	~~
-	\begin{overpic}[width=0.45\columnwidth]{TransferRegions.pdf}
+	\begin{overpic}[width=0.48\columnwidth]{TransferRegions.pdf}
 	\put(-6,85){\textbf{c}}

### 2017-11-03 journalSwarmControl/realExperiment.tex
Simplified math and renamed to ifInChord

@@
 \centering
-%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
-%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
-%\end{overpic}
 \begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}

### 2017-11-16 2017-tro-SteeringSwarmParticles/SwarmControlJournal.tex
Added bios and pictures.

@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) is a PhD Candidate in the Electrical and Computer Engineering Department at the University of Houston expected to graduate in 2018.
+
+She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
+Tehran, Iran. She is involved in studies of swarm manipulation and control for robotics applications.
+These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.
+
+\end{IEEEbiography}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Lillian.jpg}}]{Lillian Lin}(S`14) is currently pursuing a B.S.
+degree in Electrical Engineering from the University of Houston, Houston, Texas, and is expected to graduate in 2018.
+She is currently the president of the student IEEE branch on campus.
+She works as an undergraduate researcher at the University of Houston at the Robotics Swarm Control lab.
+Her main research interests include robotic control systems.
+
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} Chris Ertel received his B.S.
+in mechanical engineering from Rice University in 2010, and has since worked in industry and startups developing games, computer-aided design applications, medical telemetry systems, and other software.
+Since 2013 he's collaborated with Dr. Becker on SwarmControl.net and related projects at the University of Houston Robotic Swarm Control Lab and the Rice Multi-Robot Systems Lab.
+
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} Mable Wan received a B.S.
+degree in electrical and computer engineering and a B.B.A.
+in entrepreneurship from the University of Houston in 2017.
+Since May 2017, she has been working with Northrop Grumman Corporation in Baltimore, MD as a software engineer.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T.
+Becker}(M`06) received M.S.
+and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012.
+He was a post doctoral research scholar at the Multi-Robot Systems Lab at Rice University and a post doctoral research fellow at Boston Children's Hospital, Harvard Medical School, before joining the Electrical and Computer Engineering Department at the University of Houston as an Assistant Professor in 2014.
+He won the best paper at IROS 2013 and received the NSF CAREER award in 2016.
+He is active in developing control laws and algorithms for robot swarms.
+
+\end{IEEEbiography}
+
 \end{document}

### 2017-11-16 2017-tro-SteeringSwarmParticles/SwarmControlJournal.tex
tiny change

@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} Chris Ertel received his B.S.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} received his B.S.
 in mechanical engineering from Rice University in 2010, and has since worked in industry and startups developing games, computer-aided design applications, medical telemetry systems, and other software.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} Mable Wan received a B.S.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} received a B.S.
 degree in electrical and computer engineering and a B.B.A.

### 2017-11-16 2017-tro-SteeringSwarmParticles/SwarmControlJournal.tex
Added all the pictures and bios and shrinked the size.etc

@@
 The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
-
 Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
@@
 \end{abstract}
+
+\begin{IEEEkeywords}
+Multi-robot systems, Robot motion 
+\end{IEEEkeywords}
 
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) is a PhD Candidate in the Electrical and Computer Engineering Department at the University of Houston expected to graduate in 2018.
-
 She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
@@
 These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.
-
 \end{IEEEbiography}
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Lillian.jpg}}]{Lillian Lin}(S`14) is currently pursuing a B.S.
+\vspace{-2em}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Lillian.jpeg}}]{Lillian Lin}(S`14) is currently pursuing a B.S.
 degree in Electrical Engineering from the University of Houston, Houston, Texas, and is expected to graduate in 2018.
@@
 Her main research interests include robotic control systems.
-
 \end{IEEEbiography}
-
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} received his B.S.
@@
 Since 2013 he's collaborated with Dr. Becker on SwarmControl.net and related projects at the University of Houston Robotic Swarm Control Lab and the Rice Multi-Robot Systems Lab. 
-
 \end{IEEEbiography}
-
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} received a B.S.
@@
 \end{IEEEbiography}
-
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T.
-Becker}(M`06) received M.S.
+Becker}(S`06-M`12-SM`17) received M.S.
 and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012.

### 2017-11-16 2017-tro-SteeringSwarmParticles/expMethods.tex
Added all the pictures and bios and shrinked the size.etc

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Online experiment}
\label{sec:expMethods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Experimental Methods
%%  Platform  
%%  Human Subjects
%%% recruited through social media
%%% IRB form  Protocol Number: 14-012E Protocol Title: Massive Manipulation: A n online user study on controlling large swarms of simple robot sApproval Date: 7/26/2013Expiration Date: 7/26/2014
%%% Costs  for experiment:  ??
%%% Instrumenting:
%%%% Google analytics, airbrake, etc.

%% wherein we describe our framework
\begin{figure}
%\centering
\renewcommand{\figwid}{0.3\columnwidth}
\begin{overpic}[width =\figwid]{VaryNum.pdf}	\put(10,80){\textbf{a} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}	\put(10,80){\textbf{b} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}
%\\
%\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
%\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
%\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
\caption{\label{fig:5experiments}
Screenshots from our online experiments controlling multi-particle systems with limited, global control.
\textbf{(a)} Varying the number of particles from 1-500
\textbf{(b)} Comparing 4 levels of visual feedback 
\textbf{(c)} Varying noise from 0 to 200\% of control authority.
%\textbf{(d)} Controlling the position of 1 to 10 particles
%\textbf{(e)} Comparing 3 control architectures to assemble
%\textbf{(f)} Comparing 3 control architectures to forage.
}
\end{figure}

%\begin{figure}
%\renewcommand{\figwid}{0.32\columnwidth}
%\subfloat[][Vary Number]{\label{fig:VaryNum}
%\begin{overpic}[width =\figwid]{VaryNum.pdf}\end{overpic}}
%%
%\subfloat[][Vary Visual Feedback]{\label{fig:VaryVis}
%%\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\end{overpic}
%\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\end{overpic}}
%%
%\subfloat[][Vary Noise]{\label{fig:VaryNoise}
%\begin{overpic}[width =\figwid]{VaryNoise.pdf}\end{overpic}}\\
%%
%\subfloat[][Control Position]{\label{fig:ControlPos}
%\begin{overpic}[width =\figwid]{ControlPos.pdf}\end{overpic}}
%%
%\subfloat[][Vary Control: Assembly]{\label{fig:VaryControl}
%\begin{overpic}[width =\figwid]{VaryControl.pdf}\end{overpic}}
%%
%\subfloat[][Vary Control: Foraging]{\label{fig:Forage}
%\begin{overpic}[width =\figwid]{VaryForage.pdf}\end{overpic}}
%%
%\caption{\label{fig:5experiments}
%Screenshots from our online experiments controlling multi-robot systems with limited, global control.
%\textbf{(a)} Varying the number of robots from 1-500
%\textbf{(b)} Comparing 4 levels of visual feedback 
%\textbf{(c)} Varying noise from 0 to 200\% of control authority
%\textbf{(d)} Controlling the position of 1 to 10 robots
%\textbf{(e)} Comparing 3 control architectures to assemble
%\textbf{(f)} Comparing 3 control architectures to forage.
%\vspace{-2em}
%}
%\end{figure}

The goal of these online experiments is to test several scenarios involving large-scale human-swarm interaction (HSI), and to do so with a statistically-significant sample size. Towards this end, we have created \href{http://www.swarmcontrol.net/show_results}{SwarmControl.net}: an open-source, online testing platform suitable for inexpensive deployment and data collection on a scale not yet seen in swarm robotics research. Screenshots from this platform are shown in Fig.~\ref{fig:5experiments}.  \href{https://github.com/crertel/swarmmanipulate.git}{All code} \href{http://www.swarmcontrol.net/show_results}{and experimental results} are online at \cite{Chris-Ertel2016}.

%Our online experiments show that numerous particles responding to global control inputs are directly controllable by a human operator without special training, that the visual feedback of the swarm state should be simple to increase task performance, and that humans perform swarm-object manipulation faster using attractive control schemes than repulsive control schemes.

We developed a flexible testing framework for online human-swarm interaction studies. Over 5,000 humans performed over 20,000 swarm-robotics experiments with this framework, logging almost 700 hours of experiments.
These experiments indicated three lessons used for designing automatic controllers for object manipulation with particle swarms:

1) When the number of particles is large ($>50$), varying the number of particles does not significantly affect the performance.

2) Swarm control is robust to independent and identically distributed (IID) noise.

3) Controllers that only use the mean and variance of the swarm can perform better than controllers with full feedback.
%There are two halves to our framework: the server backend and the client-side (in-browser) frontend. The server backend is responsible for tabulating results, serving webpages containing the frontend code, and for issuing unique identifiers to each experiment participant. The in-browser frontend is responsible for running an experiment---that is to say, accepting user input, updating the state of the robot swarm, and ultimately evaluating task completion.

%% wherein we outline the process that a user takes to participate in an experiment
%\subsection{Methods}

\subsection{Implementation}

Our web server generates a unique identifier for each participant and sends it along with the landing page to the participant. 
A script on the participant's browser runs the experiment and posts the experiment data to the server. 
Anonymized human subject data was collected under IRB \#14357-01.

### 2017-11-16 2017-tro-SteeringSwarmParticles/experiment.tex
modified the publishing version
Added all the pictures and bios and shrinked the size.etc

@@
 
-The table below summarizes the simulation results for each 10 successful trials:
+The table below summarizes the simulation results for each 10 successful trials.
 \begin{center}
@@
 Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25.
-In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
+In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. %A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
 %Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.
@@
 \end{overpic}
-\vspace{-0.5em}
+\vspace{-0.em}
 \caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left).
@@
 %{blockpushingImageWithMeanAndVarianceOverlay.png}
+\vspace{-1.5em}
 \caption{\label{fig:potentialField} (Left) The attractive field is centered behind the object's COM.
@@
 (Right) Combining these forces prevents the swarm from pushing the object backwards.}
+
 \end{figure}
@@
 	\begin{overpic}[width=0.48\columnwidth]{TransferRegions.pdf}
-	\put(-6,85){\textbf{c}}
+	\put(-6,80){\textbf{c}}
 	\end{overpic}
 \end{center}
-\vspace{-0.5em}
+\vspace{-1.em}
 \caption{\label{fig:Region}  Outlier rejection state machine and regions.
 }
+\vspace{-1.5em}
 \end{figure}
@@
 
-\begin{figure}
-%  \vspace{-20pt}
-  \begin{center}
-\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
-  \end{center}
-%  \vspace{-1em}
-\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
-\vspace{-1em}
-}
-\end{figure}
+%\begin{figure}
+%%  \vspace{-20pt}
+%  \begin{center}
+%\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
+%  \end{center}
+%%  \vspace{-1em}
+%\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
+%\vspace{-1em}
+%}
+%\end{figure}
 
@@
 %\subsection{Automated object manipulation (simulation)}
-Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation.
+ Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation. Fig.~\ref{fig:DifferentMazes}a shows the six equal-area object shapes tested in simulations.
  To illustrate the flexibility of the algorithm we tested two additional workspaces, \emph{{\sffamily E}-shaped} and  \emph{Spiral}, without changing the algorithm.
@@
  More complicated workspaces could be generated by composing these workspaces. 
- Fig.~\ref{fig:DifferentMazes} shows the results of all  three mazes. The {\sffamily E}-shaped maze required the least average time because the path to the goal is shorter. 
+ Fig.~\ref{fig:DifferentMazes}b shows the results of all  three mazes. The {\sffamily E}-shaped maze required the least average time because the path to the goal is shorter. 
    Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.
@@
 \begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
+\vspace{-2em}
 \caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 particles under automatic control (see also Extension 1).
+\vspace{-2em}
 }
@@
 
-\begin{figure}\begin{center}\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
+\begin{figure}\begin{center}\begin{overpic}[width=0.3\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.3\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
 \vspace{.25em}
-\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
+\begin{overpic}[width=0.3\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.3\columnwidth]{SpiralSuccess.pdf}\end{overpic}
 \end{center}
+\vspace{-1.em}
 \caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
@@
 \centering
-\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\end{overpic}
-\vspace{-1em}
-\caption{\label{fig:DifferentMazes} Completion times for three workspaces.}
+\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\put(-6,85){\textbf{a}}\end{overpic}
+\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\put(0,68){\textbf{b}}\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:DifferentMazes} a. The six equal-area objects tested in simulation. b. Completion times for three workspaces.}
 \end{figure}
@@
 \end{overpic}
-\vspace{-0.5em}
+\vspace{-2em}
 \caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.

### 2018-01-18 IROS17/theory2robAlgs.tex
A new version of the journal.

@@
 \vspace{1em}
-\begin{overpic}[width =\figwid]{contourPlotNum.pdf}\put(97.5,21){7}
-\put(97.5,18.5){6}
-\put(97.5,16){5}
-\put(97.5,13.5){4}
-\put(97.5,11){3}
+\begin{overpic}[width =\figwid]{JustSimulationV6.png}
+%\put(97.5,21){5}
+%\put(97.5,17.25){4}
+%\put(97.5,13.5){3}
+%\put(97.5,24){6}
+%\put(97.5,27){7}
+%\put(97.5,10){2}
+%\put(97.5,7.5){1}
+%\put(97.5,5){0}
 \end{overpic}
-}\\
+}
 \caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and $\epsilon=0.001$.

### 2018-01-18 journalWallFriction/SwarmShapeControl.tex
A new version of the journal.

--- 2017-03-22 journalSwarmControl/TASE/SwarmControlIJRR.tex
@@
-% compress:
-% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=CompressedSwarmControlIJRR.pdf SwarmControlIJRR.pdf
-%%%%%%%%% IJRR SETTINGS
-%  from https://us.sagepub.com/en-us/nam/manuscript-submission-guidelines#PreparingYourManuscript
-%\documentclass[Afour,sageh,times]{sagej} % sage_latex_guidelines.tex V1.01, 11 June 2015
-\documentclass[conference]{IEEEtran}
+%  compress using: gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=foo-ShapingSwarmCompressed.pdf ShapingSwarmFrictionSharedInput.pdf
+
+%\documentclass[conference]{IEEEtran}
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
 \IEEEoverridecommandlockouts% This command is only needed if 
@@
 \newcommand{\FALSE}{\textbf{false}}
+\newcommand{\xupdownarrow}[1]{%
+  {\left\updownarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
+}
+%\newcommand{\xleftrightarrow}[1]{%
+%  {\left\leftrightarrow\vbox to #1{}\right.\kern-\nulldelimiterspace}
+%}
 \DeclareMathOperator{\atan2}{atan2}
@@
 
-%useful for highlighting sections that need work
-%\newcommand{\todo}[1]{\textcolor{red}{\footnotesize \textsf{#1}}}
+\pdfinfo{
+   /Author (Shiva Shahrokhi and Aaron T. Becker)
+   /Title  (Algorithms for Shaping a Particle Swarm with a Shared Input by Exploiting Non-Slip Wall Contacts)
+   /CreationDate (D:20160129120000)
+   /Subject (Simple Robots)
+   /Keywords (Robots;Uniform Control Inputs)
+}
 
-%\setcounter{secnumdepth}{3}
-%\runninghead{Shahrokhi \textit{et~al.}}
-%\hyphenation{micro-robot nano-robot nano-robotics micro-robotics micro-robots nano-robots micro-manip-ulators}
-%%%%%%%%%%%% For debugging purposes, I like to display the TOC
-%    \tableofcontents
-%    \setcounter{tocdepth}{3}
-%    \newpage
-%%%% END TOC %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% paper title
+\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using a Shared Input}
 
-\title{\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics}
-
-\author{Shiva Shahrokhi, Chris Ertel, Mable Wan, Lillian Lin, Aaron T. Becker% end author block 
-\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2, atbecker\}@uh.edu}}
+% You will get a Paper-ID when submitting a pdf file to the conference system
+\author{Shiva Shahrokhi, Jingang Shi, and Aaron T. Becker% <-this % stops a space
+\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
+\thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
 }
+%\affil{Department of Electrical and Computer Engineering, \\
+% University of Houston, Houston, TX 77204-4005 USA\\
+% {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}}
+%\thanks{S.
+Shahrokhi, A.
+Mahadev and  A.
+Becker are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204-4005 USA {\tt\small  \{sshahrokhi2, aviswanathanmahadev, atbecker\}@uh.edu}
+%}
+%} %\end thanks
 
@@
 \pagestyle{empty}
+
 \begin{abstract}
-Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
-The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
-We want to use large swarms of robots to perform manipulation tasks; unfortunately, human-swarm interaction studies as conducted today are limited in sample size, are difficult to reproduce, and are prone to hardware failures.
-We present an alternative.
+There are driving applications for large populations of tiny robots in robotics, biology, and chemistry.
+These robots often lack onboard computation, actuation, and communication.
+Instead, these ``robots''  are  particles carrying some payload and the particle swarm is controlled by a shared control input such as a uniform magnetic gradient or electric field.
+%In previous works, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.
+In previous works, we showed that the individual 2D position of two particles in such a swarm is controllable in a rectangular workspace requiring non-slip wall contact.
+However, both in vivo and artificial environments are usually not rectangular.
+%This work extends that work to any polygonal workspace, and focus on extending the algorithm for circular workspaces.
+Then it extends the algorithm to 3D.
+Results are validated with simulations and with a magnetic setup using two steel half a millimeter balls.
+This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
+%Requiring a small, rigid obstacle suspended in the middle of the workspace is a strong constraint, especially in 3D.
+%Often particles are placed into an em
+%This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contact in 2D.
+%Both in vivo and artificial environments often have such boundaries.
+%We assume that particles in contact with the boundaries have zero velocity if the shared control input pushes the particle into the wall.
+%This paper provides a shortest-path algorithm for positioning a two-particle swarm, and a generalization to positioning an $n$-particle swarm.
+%Results are validated with simulations and a hardware demonstration.
 
-This paper first examines the perils, pitfalls, and possibilities we discovered by launching \href{http://www.swarmcontrol.net}{SwarmControl.net}, an online game where players steer swarms of up to 500 robots to complete manipulation challenges.
-We record statistics from thousands of players, and use the game to explore aspects of large-population robot control.
-We present the game framework as a new, open-source tool for large-scale user experiments.
-One surprising result was that humans completed an object manipulation task \emph{faster} when provided with only the mean and variance of the swarm than with full-state feedback.
-Inspired by human operators, this paper next investigates controllers that use only the mean and variance of the swarm.
-We prove that the mean position is controllable, then provide conditions under which variance is controllable.
-We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the robot distribution.
-Finally, we employ these controllers as primitives for an object manipulation task and implement all the automatic controllers on 100 kilobots controlled by the direction of a global light source.
+%Consider a swarm of particles controlled by global inputs. 
+%This paper presents algorithms for shaping such swarms in 2D using boundary walls.
+%The range of configurations created by conforming a swarm to a boundary wall is limited. 
+%We describe the set of stable configurations of a particle swarm in two canonical workspaces, a circle and a square. 
+%To increase the diversity of configurations, we add boundary interaction to our model.  
+%We provide algorithms using friction with walls to place two robots at arbitrary locations in a rectangular workspace.
+%Next, we extend this algorithm to place $n$ agents at desired locations. 
+%We conclude with efficient techniques to generate correlations of a particle swarm not possible without wall-friction. 
+%Simulations %and hardware implementations with 100 robots 
+%validate these results.
+
+%These methods may have particular relevance for micro- and nano-robots controlled by global inputs.
+%, whose small size limits onboard computation and power.
+For this reason they are usually powered and controlled by global inputs, such as a uniform external electric or magnetic field, and every robot receives the same control inputs.
+%Due to their small size, large numbers of micro-robots are required to deliver sufficient payloads.
+% Nevertheless, these applications require precision control of the shape and position of the robot swarm. Precision control requires breaking the symmetry caused by the global input.
+
+% KEYWORDS:   uniform control, under-actuation, particle swarm
 \end{abstract}
 
-%\keywords{micro-robot, global control, swarm, human-swarm interaction}
+\IEEEpeerreviewmaketitle
 
-%\input{00-Abstract.tex}
-\input{Introduction.tex}
-\input{relatedWork.tex}
-\input{expMethods.tex}
-\input{expResults.tex}
-\input{theory.tex}
-\input{simulation.tex}
-\input{experiment.tex}
-\input{realExperiment.tex}
-\input{conclusion.tex}
+%%%%%%%%%%%%%%%
+\input{intro}
+%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%
+\input{relatedWork}
+%%%%%%%%%%%%%%%
+\input{theory}
+\input{theory2robAlgs}
+%\input{theorynrobAlgs}
+%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%
+\input{simulation}
+%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%
+\input{experiment}
+%%%%%%%%%%%%%%%
+\input{conclusion}
+%%%%%%%%%%%%%%%
+%\section*{Acknowledgments}
+%%%%%%%%%%%%%%%
+%% Use plainnat to work nicely with natbib. 
+{%\footnotesize
+%\bibliographystyle{plainnat}
+%\bibliographystyle{SageH}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,SwarmShapeControl}
+}
 
-%\begin{dci}
-%The Authors declare that there is no conflict of interest.
-%\end{dci}
-%\begin{funding}
-%%RULES: http://www.nsf.gov/pubs/policydocs/pappguide/nsf16001/aag_6.jsp
-%This work was supported by the National Science Foundation under Grant No.\ 
-%\href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}.
-%\end{funding}
-%\newpage
-%\begin{sm}
-%\begin{itemize}
-%\item Extension 1:  simulation video of the parameter sweeps in Fig.~\ref{fig:AutoVeryParam} with robot swarms manipulating objects.
-%\item Extension 2: hardware experiment videos of object manipulation with a swarm of 100 kilobots.
-%\end{itemize}
-%\end{sm}
-
-%\input{11-MultiMediaAppendix.tex}
-
-\bibliographystyle{IEEEtran}
-%\bibliographystyle{SageH}
-\bibliography{IEEEabrv,SwarmControlIJRR}
+% Uncomment to add appendix:
+%\input{appendix}
 

### 2018-01-18 journalWallFriction/experiment.tex
A new version of the journal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Results}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\emph{Hardware Demonstration: Position Control of n Robots}\\
%A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
% % Designs for each are available at \cite{Arun2016Thingiverse}. 
%  The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
%    Only discrete, 1 cm moves in the $x$ and $y$ directions are used. The goal configuration highlighted in the top right corner represents a `U' made of seven particles. The dark red configuration is the current position of the particles. 
%Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.  Free particles return to their start positions but particles on the boundary to move laterally, generating a net sliding motion.
%
%Fig. \ref{fig:construction2d} follows the motion of the particles through iterations  $k$=1, 2, and 7. All particles receive the same control inputs, but boundary interactions break the control symmetry.  Particles reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left particle from the staging zone occupying the top-right position of the build zone.
%

### 2018-01-18 journalWallFriction/intro.tex
A new version of the journal.

--- 2017-08-01 ICRA-RA-L2018/intro.tex
@@
 \vspace{1.5em}
-\begin{overpic}[width=\columnwidth]{circleWallFrictionReachableRegions.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
 \caption{\label{fig:IntroPic}
 Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
-Top row shows the output of a best-first-search algorithm that finds the shortest path for two particles.
- Top left shows the initial position and goal position of the particles.  
- The shortest path consists of moving at angle 50$^\circ$ until the blue robot contacts the top wall, then moving the  magenta robot at angle 165$^\circ$ until the particles reach the desired relative spacing, then moving $-$60$^\circ$ to the goal positions.
- The two bottom pictures show $n$-particle positioning using shared control inputs and boundary interaction. 
 } \vspace{-1em}
@@
   Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
-We provide a shortest-path algorithm to arbitrarily position two robots in Sec.
-\ref{sec:PostionControl2Robots}, and Sec.
-\ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with non-slip boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
+We provide a shortest-path algorithm to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}.
 Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
  We end with directions for future research in Sec.  \ref{sec:conclusion}.
+
+This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which consider only square workspaces.
+This work extends the analysis to convex workspaces and 3D positioning.
+This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.

### 2018-01-18 journalWallFriction/simulation.tex
A new version of the journal.

\section{Simulation}\label{sec:simulation}

%Two simulations were implemented using non-slip contact walls for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.

\subsection{Position Control of Two Robots}

Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using point robots (radius = $0$). 
%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
 %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.

 The contour plots in Fig.~\ref{fig:contour} left shows the length of the path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace. Fig.~\ref{fig:contour} right shows the total distance of the path. This plot clearly shows the nonlinear nature of the path planning. The hardest point to achieve is the when the goals have $\pi$ difference and are very close to the boundary.
 %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.

 The plots in Fig.~\ref{fig:deltanumdist} show the exponentially increasing number of moves and distance when the accuracy of reaching to the goal ($\delta$) is getting to zero when the goal positions have $\pi$ difference with each on the boundaries.

\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{numcontour.pdf}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{distcontour.pdf}\end{overpic}
\caption{\label{fig:contour}
Plots showing the algorithm with one goal on the boundary.
} \vspace{-1em}
\end{figure}

\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{deltanum.pdf}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{deltadist.pdf}\end{overpic}
\caption{\label{fig:deltanumdist}
Plots showing decreasing error when the number of moves grows.
} \vspace{-1em}
\end{figure}

\begin{figure}
\centering
\renewcommand{\figwid}{1\columnwidth}
{\begin{overpic}[width =\figwid]{identical.png}
\end{overpic}
\begin{overpic}[width =\figwid]{contourDistnew.png}
\end{overpic}
\vspace{1em}
\begin{overpic}[width =\figwid]{JustSimulationV6.png}
%\put(97.5,21){5}
%\put(97.5,17.25){4}
%\put(97.5,13.5){3}
%\put(97.5,24){6}
%\put(97.5,27){7}
%\put(97.5,10){2}
%\put(97.5,7.5){1}
%\put(97.5,5){0}
\end{overpic}
}\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and $\epsilon=0.001$.
 The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
\vspace{-1em}
}
\end{figure}

   %The path given by our algorithm is shown with solid arrows.
%Each row has five snapshots taken every quarter second. 
%For the sake of brevity axis-aligned moves were replaced with oblique moves that combine 
%jjkjk klklk klklklk klklk lklkl lklklklkl  klklklklk lklklklklk lklklkl klklklkl klklklk klklklk klklk klk klkl
% THES IMAGES DON'T DO THIS ANYMORE 
% $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
% The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
% $\Delta r_y$ is corrected by $t = 0.75$. 
% Finally, the algorithm moves the robots to their corresponding destinations.

%\subsection{Position Control of $n$ Robots}
%Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. %Code is available at \cite{Arun2015}.
% Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
%The distance moved grows quadratically with the number of robots $n$. A best-fit line $210 n^2 + 1200n-10,000$ is overlaid by the data.
%
%In Fig.~\ref{fig:4diagramsplots.pdf}, the amount of clearance is $\epsilon=1$.
%Control performance is sensitive to the desired clearance.  As $\epsilon$ increases, the total distance decreases asymptotically, as shown in Fig.~\ref{fig:graphrobo.pdf}, because the robots have more room to maneuver and fewer $\operatorname{DriftMove}$s are required.
%%We observed that the value of $\epsilon$, greatly affects the total number of moves of commanded moves on the robots. The smaller the value of ‘?’ the greater is the number of moves required to move a said distance. The second graph plot(fig roboplot) is that of ‘?’ vs ‘total commanded moves’ which is the number of moves the global controller commands onto all the robots. Here we take the pattern 
%%‘ROBO’ and calculate the ‘total commanded moves’ for ‘?’=[ 0.25,0.375,0.5,0.625,0.75,0.875,1]. We can notice an exponential fall in the total command moves. It can be inferred that there is a tradeoff between area of bounded region and the number of moves. For smaller ‘?’, smaller is the required area to implement algorithm but the longer is the time taken to complete the algorithm. Fig robo
%%Unit of ‘?’ is the ratio of minimum clearance value to the body length of a robot.
%
%
%
%
%\begin{figure}
%\begin{center}
%	\includegraphics[width=\columnwidth]{5diagramsplots2.pdf}
%\end{center}
%\vspace{-1em}
%\caption{\label{fig:4diagramsplots.pdf}
%The required number of moves under Alg.~\ref{alg:PosControlNRobots} using non-slip contact walls to rearrange $n$ square-shaped robots grows quadratically with $n$. % \cite{Arun2015}.
% See code at \cite{Arun2015}.
%}
%\end{figure}
%
%\begin{figure}
%\begin{center}
%	\includegraphics[width=\columnwidth]{graphrobo.pdf}
%\end{center}
%\vspace{-1em}
%\caption{\label{fig:graphrobo.pdf}
%Control performance is sensitive to the desired clearance $\epsilon$.  As $\epsilon$ increases, the total distance decreases asymptotically.
%}\vspace{-1em}
%\end{figure}

### 2018-01-18 journalWallFriction/theory.tex
A new version of the journal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Boundary Interaction Model}\label{subsec:WallFriction}

In the absence of obstacles uniform inputs move a swarm identically.  
Independent control requires breaking this symmetry. 
The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.

 If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
 \begin{align}\label{eq:swarmDynamicsAndFric} 
\dot{\mathbf{x}}_i(t)
 &=
 \mathbf{u}(t)
 +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), ~i \in [1,n].\\
 F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
  - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
 \\
 0 & \textrm{else} 
 \end{cases}\nonumber
 \end{align}
 Here $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$, and
 $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary.

These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
  As in our model, fluid flowing in a pipe has zero velocity along the boundary. Similar mechanical systems exist at larger scales, e.g. all tumblers of a combination look move uniformly unless obstructed by an obstacle.
 Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.

 \subsection{Shortest Path}
 The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments shown in Fig.~\ref{fig:shortestpath}. The problem can be simplified by choosing the coordinate system carefully. We define the x axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point by a radius e and angle $\beta$, $E=e(\cos \beta,\sin \beta)$. Then define a symmetry point about S of line OP named T. Then the length of the two line segments is 
 \begin{equation}
\scriptsize{ \sqrt{(s-\cos \theta)^2+(-\sin \theta)^2}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}}
 \end{equation}
 which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to

 \begin{equation}
 \sqrt{1+e^2-2 e \cos(\beta-\theta)}+\sqrt{1+s^2-2 s \cos \theta}. 
 \end{equation}

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.}
%\vspace{-1em}
}
\end{figure}

 The length of the two line segments as a function of $\theta$ is drawn in the right plot. There are several simple solutions. If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0. If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$. Label the origin $O$. The optimal solution shows that the angle $\angle{OPS}$ (from the origin to P to S) is the same as the angle $\angle{OPE}$ (from the origin to P to E). We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are S and E. When the ellipse is tangent to the circle, the point of tangency is exactly P. 
  Since the distance from the origin to P is always 1, we can set up three equalities using the law of sines:
 From triangle OSP: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{SP}$, and from triangle OEP: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{EP}$. If we mirror the point S about the $\theta$ axis and label this point C, from triangle CEO: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{CE}$.

 Simplifying this system of equations results in: $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi<=\beta<=\pi$. We evaluate each and select the solution that results in the shortest length path. Note that the optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.

 \subsection{Different Polygonal Workspaces}

 Fig. \ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces. Consider one robot touching each vertex of the workspace. For each pair of vertices, compute where the other robot will be if the first robot goes to that vertex. If the final position of the second robot is still inside the workspace, then its position is one of the vertices of the reachable set. If the point is not inside the polygon, then the intersection of the line that point and the second robot's position with the polygon is one vertex of the reachable set. Compute the distance to all the vertices of the workspace from this point. By subtracting the relative distance of the robots, then all the vertices of one reachable set are found. Doing this for all the vertices of the workspace will give us all the reachable sets.
 \begin{figure}
\centering
\renewcommand{\figwid}{0.8\columnwidth}
{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
\end{overpic}
}
\caption{\label{fig:polygon}{Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets. As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
}
\vspace{-1em}
}
\end{figure}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.66\columnwidth}
\begin{overpic}[width =\figwid]{anypolygon1.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{anypolygon2.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{anypolygon3.pdf}
\end{overpic}
\caption{\label{fig:polygonAlg}{ blah blah blah}
\vspace{-1em}
}
\end{figure*}

### 2018-01-18 journalWallFriction/theory2robAlgs.tex
A new version of the journal.

--- 2018-01-18 IROS17/theory2robAlgs.tex
@@
 
-Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a rectangular workspace. 
- Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
+Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a circular workspace.
+In our previous work we used a rectangular workspace.
+We use the same idea and algorithm here but we have to modify the algorithm to adjust it for a circular workspace.
+% Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
+
+Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+Denote the current positions of the robots  $p_1$ and $p_2$ and the current distance between the robots is $d$.
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
+The algorithm assigns a uniform control input at every instance.
+The goal is to move the particles to the goal positions using a shared control input.
+We do this by first moving them to the correct relative position and then translating the particles to the goal.
+The first step minimizes
+ $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
+
+ We define a $\Delta$ configuration space as a circular shape that considers all possible $\Delta p$s.
+We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:deltaSpace}.
+Reachable set is the part of $\Delta$ configuration space where if one robot touches a wall in a specific location, the other robot can make the required relative distance without causing touching robot to move.
+To compute reachable set for circular workspace, first we considered all passible hitting point locations in the workspace.
+ The set of boundary points that a robot can touch before the  other robot  touches is an arc of angle $2(\pi - \frac{\arcsin{d}}{r})$, where $d= |s_1 - s_2|$ and $r$ is the radius of the circle.
+ %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
+ We define the angle between two particles as $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
+
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}\end{overpic}
+\caption{\label{fig:deltaSpace}Left:  the set of points where the red robot is the first to contact the boundary are drawn with a red arc.
+The  set of points where the blue robot is the first to contact the boundary are drawn with a blue arc.
+The possible points for the blue and pink particles to touch the boundary is shown in blue and pink arcs.
+Right: When the blue particle is touching a wall (blue square) the other particle (pink square) can go anywhere in the reachable set (blue region).}
+\end{figure}
 
 \begin{figure*}
-\renewcommand{\figwid}{0.4\columnwidth}
-
-{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
-\put(10,85){a)}
-\put(30,70){$r_1$}
-\put(48,88){$g_1$}
-\put(65,17){$r_2$}
-\put(87,52){$g_2$}
-\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
-\put(110,50){If $(\Delta e.x, \Delta e.y) = (0,0)$, only one move is needed}
-\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
-%\put(47, 65){$\xleftrightarrow{1.8cm}$}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
-\put(10,85){b)}
-\put(28,70){$r_1$}
-\put(50,80){$g_1$}
-\put(62,15){$r_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
-\put(110,50){Three move sequence}
-\put(12,38){$m_3$}
-\put(48,52){$m_2$}
-\put(65,30){$m_1$}
-\end{overpic}
-%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
-%\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
-\put(10,85){c)}
-\put(28,70){$r_1$}
-\put(62,28){$g_1$}
-\put(87,50){$r_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-%\vspace{1em}
-{
-\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
-\put(10,85){d)}
-\put(25,30){$r_1$}
-\put(89,70){$g_1$}
-\put(67,20){$r_2$}
-\put(28,75){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
-\put(10,85){e)}
-\put(15,35){$r_1,g_2$}
-\put(65,35){$r_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
-\put(10,85){f)}
-\put(15,10){$r_1,g_2$}
-\put(65,88){$r_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
-\put(15, 85){worst case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
-\end{overpic}
-}\\
-
-%\vspace{-0.75em}
-%{
-%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
-%\end{overpic}
-%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
-%%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
-%\end{overpic}
-%}\\
-
-\vspace{-1em}
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two robot positioning using walls with non-slip contacts. 
-Robot start positions are shown by a square, and goal positions by a circle.
-Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
-Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-\vspace{-1em}
+\centering
+%\begin{overpic}[width=0.5\columnwidth]{startPath.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{finalPath.pdf}\end{overpic}
+\caption{\label{fig:reachableSet}
+Left circle shows the workspace. %, and the corresponding reachable set when blue robot is touching the wall is shown as a blue region. 
+Right shows the $\Delta$ configuration space and the reachable set that is shown in red is representative of the point we need to go to get to the goal relative distance in one move.%representative of the blue and pink reachable sets shown in the workspace.
 }
 \end{figure*}
-Assume two robots are initialized at $r_1$ and $r_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-The solution is a best-first search algorithm that maintains a list of possible paths (\emph{moves}), sorted according to an admissible heuristic on path length.
-The algorithm works by expanding the path with the shortest estimated length.
+
+%\begin{figure}
+%\centering
+%
+%\caption{\label{fig:chord}
+%When the blue particle is touching a wall (blue square) the other particle (pink square) can go anywhere in the reachable set (blue region).
+%}
+%\end{figure}
+%The algorithm works by expanding the path with the shortest estimated length.
 Expanding a path means either moving directly to the goal, or pushing one robot to a wall and adjusting the relative position of the other robot.
-As soon as the goal is reached, the algorithm returns this shortest path.
+As soon as the goal is reached, the algorithm returns this path.
 
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $r_1.x$ and $r_1.y$ denote the $x$ and $y$ locations of $r_1$. 
-The algorithm assigns a uniform control input at every instance.
-The goal is to zero the error in both coordinates,
- $\Delta e =(\Delta e.x, \Delta e.y) =\Delta g - \Delta r = (g_2-g_1)- (r_2-r_1)$ using a shared control input. 
- The base case occurs when $\Delta e = (0,0)$.
-In this base case, the shortest path is a straight line between the current position of each robot to its goal position, as shown in Fig.
-\ref{fig:shapeControlMathematica1}a.
-If  $\Delta e\neq (0,0)$, we compute the two-move reachable sets generated by first immobilizing one particle by contacting the wall and then repositioning the second particle.
-There are four reachable sets, as shown in Fig.~\ref{fig:TwoRegions}.
-The horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose to immobilize the particle closest to a wall. 
+There are infinite reachable sets, parameterized by first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove}.
+%We make the reachable sets by knowing the angle of the chord, $\psi$, shown in Fig.~\ref{fig:psigamma}. Minimum and maximum values of $\psi$ can be calculated in the following equation.
+\begin{align}
+ \psi \in [\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}]
+% \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
+\end{align}
+$\gamma$ is half the angle of the arc that the reachable set's chord has shown in Fig.~\ref{fig:deltaSpace} and is calculated by:
+\begin{align}
+p_\psi &= r[\cos(\psi ), \sin(\psi )]\\
+d_\perp&= 2 \abs{(s_1.p_\psi - s_2.p_\psi)}\\
+\gamma &= \cos^{-1} (1-\frac{d_\perp}{r} )
+\end{align}
+Reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose to immobilize the particle closest to a wall.
+
+The equation for the four lines outlining the reachable set can be found as follows:
+\todo{Insert the equations}
+%\begin{align}
+%l_1 = r{\cos{\psi_{\min}, 
+%   Sin[\[Theta] + \[Alpha]/2 + \[Pi]/2 ]} - 
+% 1/2 {Cos[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2], 
+%   Sin[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2]}
+%\end{align}
 % When a particle contacts a wall, the other particle can move to any position in the reachable set while the first robot is immobile. 
 %Two reachable sets are possible, horizontal and vertical. 
-Each non-slip move changes the $\Delta$ configuration space, as shown in  Fig.~\ref{fig:regionMove}.
+%Each non-slip move changes the $\Delta$ configuration space, as shown in  Fig.~\ref{fig:regionMove}.
 
@@
  %employing the assumption we have made earlier about the walls' friction. 
-%Our algorithm uses an A*-like method to find the shortest path in each move. 
-%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+ \subsection{Square workspace}
+Our algorithm uses an A*-like method to find the shortest path in each move. 
+If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
 This is shown as $m_2$ in Fig.
@@
 To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
-%The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
+The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
 If the goal configuration can be reached in three moves, then $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
@@
  \State  \emph{PathList} $\gets \{\}$
- \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains the admissible heuristic, the move sequence, and the current robot positions
+ \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains $h$, the admissible heuristic \eqref{eq:admissibleHeuristic}, the move sequence, and the current robot positions
 \While {$P.r_1 \ne g_1$ \textbf{and} $P.r_2 \ne g_2$}
@@
 
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.67\columnwidth}
-{\begin{overpic}[width =\figwid]{RegionMove1.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove2.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{RegionMove3.pdf}
-\end{overpic}
-}\\
-
-\caption{\label{fig:regionMove}{Workspace and $\Delta$ configuration space for three sets of robot configurations with the same final goal.
-The red square represents the starting $\Delta x$ and $\Delta y$ and the green circle represents the goal $\Delta x$ and $\Delta y$.
-The green rectangle illustrates reachable $\Delta x$ and $\Delta y$ when one particle is in contact with a horizontal wall and the blue rectangle illustrates the reachable region when in contact with a vertical wall.
-}
-\vspace{-1em}
-}
-\end{figure*}
 \begin{figure}
@@
 
-\begin{figure*}
+%\begin{figure*}
+%\centering
+%\renewcommand{\figwid}{0.5\columnwidth}
+%{\begin{overpic}[width =\figwid]{epsilonMin.pdf}%\put(20,20){$\epsilon = 0.001$}
+%\put(20, 3.1){{\scriptsize$\swarrow$}~$\epsilon = 0.001$}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
+%\put(20, 3.9){{\tiny$\updownarrow$}~$\epsilon = 0.05$}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
+%\put(20, 5){{\large$\updownarrow$}~$\epsilon = 0.1$}
+%\end{overpic}
+%}\\
+%\caption{\label{fig:epsilon}{Changing the minimum spacing $\epsilon$ changes the path.   $\epsilon$ is the minimum spacing between two robots and the minimum separation from the boundaries.}
+%\vspace{-1em}
+%}
+%\end{figure*}
+
+\subsection{3D model}
+
+Extending to 3D is straight forward, but possible only if the two particles do not have the same $x$ and $y$ positions.
+Fig.~\ref{fig:zaxis} shows a cylinder.
+The blue particle is in the blue plate and the red robot is in the red plate.
+There are two possible optimal paths shown with parallel arrows.
+Each arrow will cause one of the particles to touch the wall, enabling the other robot to move freely in $z$ axis to make the required relative position.
+
+\begin{figure}
 \centering
-\renewcommand{\figwid}{0.5\columnwidth}
-{\begin{overpic}[width =\figwid]{epsilonMin.pdf}%\put(20,20){$\epsilon = 0.001$}
-\put(20, 3.1){{\scriptsize$\swarrow$}~$\epsilon = 0.001$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMed.pdf}%\put(20,20){$\epsilon = 0.05$}
-\put(20, 3.9){{\tiny$\updownarrow$}~$\epsilon = 0.05$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{epsilonMax.pdf}%\put(20,20){$\epsilon = 0.1$}
-\put(20, 5){{\large$\updownarrow$}~$\epsilon = 0.1$}
-\end{overpic}
-}\\
-\caption{\label{fig:epsilon}{Changing the minimum spacing $\epsilon$ changes the path.   $\epsilon$ is the minimum spacing between two robots and the minimum separation from the boundaries.}
-\vspace{-1em}
-}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{2\columnwidth}
-{\begin{overpic}[width =\figwid]{contourDistnew.png}
-\end{overpic}
-
-\vspace{1em}
-\begin{overpic}[width =\figwid]{JustSimulationV6.png}
-%\put(97.5,21){5}
-%\put(97.5,17.25){4}
-%\put(97.5,13.5){3}
-%\put(97.5,24){6}
-%\put(97.5,27){7}
-%\put(97.5,10){2}
-%\put(97.5,7.5){1}
-%\put(97.5,5){0}
-\end{overpic}
-}
-\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and $\epsilon=0.001$.
- The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
-\vspace{-1em}
-}
-\end{figure*}
+\begin{overpic}[width=0.5\columnwidth]{zaxis.pdf}\end{overpic}
+\caption{\label{fig:zaxis}
+Extending the algorithm to position the particles in 3D.
+} \vspace{-1em}
+\end{figure}

### 2018-01-18 journalWallFriction/theory2robAlgs.tex
new abstract and 3D text

@@
 
-\subsection{3D model}
+\subsection{3D workspaces: Cylinders and Prisms}
 
-Extending to 3D is straight forward, but possible only if the two particles do not have the same $x$ and $y$ positions.
-Fig.~\ref{fig:zaxis} shows a cylinder.
-The blue particle is in the blue plate and the red robot is in the red plate.
-There are two possible optimal paths shown with parallel arrows.
-Each arrow will cause one of the particles to touch the wall, enabling the other robot to move freely in $z$ axis to make the required relative position.
+Extending path planning to 3D is possible only if the two particles do not initially have the same $x$ and $y$ positions.
+For ease of analysis, we assume the workspace boundaries extend in the $\pm z$ direction to form either right cylinders or right prisms.
+If the 3D projection is at a different angle, redefine the 2D workspace as a region perpendicular to the projection.
+ First, we move the closest particle to the boundary, which prevents its $z$ coordinate from changing.  
+ We next apply actuation in either the $\pm z$ direction to achieve the desired $\Delta z$.
+ Then the particles are actuated away from the boundary and to the appropriate $z$ positions.
+ Path planning continues using Alg. XX to position the particles to the desired $x$ and $y$ positions. 
+ As an example, consider Fig.~\ref{fig:zaxis} which shows a cylinder.
+ The blue particle starts in the blue disk and the red particle starts in the red disk. There are two possible optimal paths shown with parallel arrows. 
+Each arrow will cause one of the particles to touch the wall, enabling the other robot to move freely in the  $z$ axis to achieve the required relative position.
+This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.
+Other workspaces may be better handled by other path planners, such as \cite{MassiveUnifromManipulation}.
 

### 2018-01-24 journalWallFriction/theory2robAlgs.tex
Equations Added

--- 2018-01-18 journalWallFriction/theory2robAlgs.tex
@@
 \begin{align}
- \psi \in [\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}]
+ \psi \in \Big(\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big)
 % \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
@@
 d_\perp&= 2 \abs{(s_1.p_\psi - s_2.p_\psi)}\\
-\gamma &= \cos^{-1} (1-\frac{d_\perp}{r} )
+\gamma &= \cos^{-1} \Big(1-\frac{d_\perp}{r} \Big)
 \end{align}
@@
 The equation for the four lines outlining the reachable set can be found as follows:
-\todo{Insert the equations}
-%\begin{align}
-%l_1 = r{\cos{\psi_{\min}, 
-%   Sin[\[Theta] + \[Alpha]/2 + \[Pi]/2 ]} - 
-% 1/2 {Cos[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2], 
-%   Sin[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2]}
-%\end{align}
+
+\begin{align}\nonumber
+l_1 &=  r \Big((\cos\psi_{\min}, \sin\psi_{\min})\\ \nonumber
+& - (\cos(\gamma_c + \psi_{\min}), \sin(\gamma_c + \psi_{\min}))\Big) &  0<\gamma_c< \gamma_{\psi_{\min}},\\ \nonumber
+l_2 &=  r \Big((\cos\psi_{\max}, \sin\psi_{\max})\\ \nonumber
+& - (\cos(\gamma_c + \psi_{\max}), \sin(\gamma_c + \psi_{\max}))\Big) &  \gamma_{\psi_{\max}}<\gamma_c< 0,\\  \nonumber
+l_3 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
+& - (\cos( \psi+\gamma_c ), \sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+l_4 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
+& - (\cos( \psi-\gamma_c ), \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+\end{align}
+We combine these boundaries to make the reachable set. 
+The algorithm first makes the reachable set, and then checks if the goal relative position is in the reachable set.
+If the goal is inside the reachable set, we need to find a $\psi$ that would let us do the move.
+To do so, we first check if $\Delta g$ is in the correspondent ????
+
+\begin{align}
+&(p.x - c.x)^2 + (p.y - c.y)^2  > r^2\\
+ &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi <= -r \cos\gamma
+\end{align}
 % When a particle contacts a wall, the other particle can move to any position in the reachable set while the first robot is immobile. 

### 2018-01-24 journalWallFriction/SwarmShapeControl.tex

@@
 \begin{abstract}
-There are driving applications for large populations of tiny robots in robotics, biology, and chemistry.
-These robots often lack onboard computation, actuation, and communication.
-Instead, these ``robots''  are  particles carrying some payload and the particle swarm is controlled by a shared control input such as a uniform magnetic gradient or electric field.
+
+Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
+Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
+Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
+Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
+The walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
+In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
+However, in vivo environments are usually not square. 
+This work extends the previous work to convex workspaces and 3D positioning. 
+This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
+
+%There are driving applications for large populations of tiny robots in robotics, biology, and chemistry.
+%These robots often lack onboard computation, actuation, and communication.
+%Instead, these ``robots''  are  particles carrying some payload and the particle swarm is controlled by a shared control input such as a uniform magnetic gradient or electric field.
 %In previous works, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.
-In previous works, we showed that the individual 2D position of two particles in such a swarm is controllable in a rectangular workspace requiring non-slip wall contact.
+%In previous works, we showed that the individual 2D position of two particles in such a swarm is controllable in a rectangular workspace requiring non-slip wall contact.
 However, both in vivo and artificial environments are usually not rectangular.
@@
 Results are validated with simulations and with a magnetic setup using two steel half a millimeter balls.
-This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
+%This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
 %Requiring a small, rigid obstacle suspended in the middle of the workspace is a strong constraint, especially in 3D.

### 2018-01-30 journalWallFriction/theory2robAlgs.tex
One algorithm added

--- 2018-01-18 journalWallFriction/theory2robAlgs.tex
@@
 In our previous work we used a rectangular workspace.
-We use the same idea and algorithm here but we have to modify the algorithm to adjust it for a circular workspace.
+We use the same idea here but we modify the algorithm to handle a circular workspace.
 % Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
@@
 The equation for the four lines outlining the reachable set can be found as follows:
-\todo{Insert the equations}
-%\begin{align}
-%l_1 = r{\cos{\psi_{\min}, 
-%   Sin[\[Theta] + \[Alpha]/2 + \[Pi]/2 ]} - 
-% 1/2 {Cos[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2], 
-%   Sin[a2 + \[Theta] + \[Alpha]/2 + \[Pi]/2]}
-%\end{align}
+
+\begin{align}\label{eq:circlereachable}
+l_1 &=  r \Big((\cos\psi_{\min}, \sin\psi_{\min})\\ \nonumber
+& - (\cos(\gamma_c + \psi_{\min}), \sin(\gamma_c + \psi_{\min}))\Big) &  0<\gamma_c< \gamma_{\psi_{\min}},\\ \nonumber
+l_2 &=  r \Big((\cos\psi_{\max}, \sin\psi_{\max})\\ \nonumber
+& - (\cos(\gamma_c + \psi_{\max}), \sin(\gamma_c + \psi_{\max}))\Big) &  \gamma_{\psi_{\max}}<\gamma_c< 0,\\  \nonumber
+l_3 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
+& - (\cos( \psi+\gamma_c ), \sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+l_4 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
+& - (\cos( \psi-\gamma_c ), \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+\end{align}
+We combine these boundaries to make the reachable set. 
+The algorithm first makes the reachable set, and then checks if the goal relative position is in the reachable set.
+If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ would be our current goal.
+We need to find a $\psi$ that would enable us to reach to the required relative goal distance.
+To do so, we first check $\psi_{min}$ and $\psi_{max}$.
+\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the robots if the touching robot has the angle $\psi$.
+
+\begin{align}\label{eq:ifinchord}
+&(p.x - c.x)^2 + (p.y - c.y)^2  > r^2\\ \nonumber
+ &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi <= -r. \cos\gamma
+\end{align}
+
+Here $c$ is coordinate of the circle and $r$ is the radius of the workspace.
+ If $\Delta g$ is not in the region made by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g$ and the current relative position, $\Delta s$.
+This line is a chord of the circle and we find the $\psi$ that makes this region.
+Now that $\psi$ is found, we move the particles to make the current goal.
+If current goal is our final goal, we go to the final goal position.
+If it is not our final goal, we continue to set the closest point on the reachable set to our current goal until we reach the goal.
 % When a particle contacts a wall, the other particle can move to any position in the reachable set while the first robot is immobile. 
@@
 \begin{algorithm}[htb]
-\caption{ { \sc 2-ParticlePathFinder}($r_1,r_2,g_1,g_2,L$)}\label{alg:optimalAlg}
+\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P$)}\label{alg:optimalAlg}
 \begin{algorithmic}[1]
 %\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
- \State  \emph{PathList} $\gets \{\}$
- \State $P \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $P$ contains $h$, the admissible heuristic \eqref{eq:admissibleHeuristic}, the move sequence, and the current robot positions
-\While {$P.r_1 \ne g_1$ \textbf{and} $P.r_2 \ne g_2$}
-\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
-\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($P.r_1,P.r_2,g_1,g_2,\theta$)
-\State $\{d, $ \parbox[t]{.3\linewidth}{%
- \emph{moves,}$ r_1,r_2\} \gets$\\
- {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.\text{\emph{moves}}$)}
-\State $(\text{\emph{moves}}, r_1,r_2) \gets$ {\sc Rotate}(\emph{moves}$, r_1,r_2,-\theta$)
-\State   {\sc Push} $\{d, \text{\emph{moves}}, r_1,r_2\} $ onto \emph{PathList}
-\EndFor
-\State {\sc Sort}(\emph{PathList}) \Comment sort by admissible heuristic
-\State $P \gets $ {\sc Pop} first element of \emph{PathList}
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+%$(0,0)$ is bottom corner,
+ $P$ is a list of the vertices of a convex polygon. 
+% \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
+% \State  \emph{PathList} $\gets \{\}$
+ \State $(p_1,p_2) \gets (s_1,s_2) $
+ \State $R \gets   \{ p_1,p_2,g_1,g_2  ,moves\} $ \Comment $R$ contains the current robot positions, the goal positions, and the move sequence
+ \State $\Delta p \gets p_2-p1$
+ \State $\Delta g \gets g_2-g_1$
+\While {$|\Delta g| < |\Delta p|+ \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
+\State Calculate the reachable set using \eqref{eq:circlereachable} or Fig.~\ref{fig:polygonAlg}
+\State $ \Delta cg\gets $nearest point on the reachable set to $\Delta g$
+\State Add the required moves to $moves$ for the particles to achieve $\Delta cg$ 
+%\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
+%\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($R.r_1,R.r_2,g_1,g_2,\theta$)
+%\State $\{d, $ \parbox[t]{.3\linewidth}{%
+% \emph{moves,}$ r_1,r_2\} \gets$\\
+ %{\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.\text{\emph{moves}}$)}
+%\State $(\text{\emph{moves}}, r_1,r_2) \gets$ {\sc Rotate}(\emph{moves}$, r_1,r_2,-\theta$)
+%\State   {\sc Push} $\{d, \text{\emph{moves}}, r_1,r_2\} $ onto \emph{PathList}
+%\EndFor
+%\State {\sc Sort}(\emph{PathList}) \Comment sort by admissible heuristic
+%\State $R \gets $ {\sc Pop} first element of \emph{PathList}
 \EndWhile
+\State $moves \gets moves+ {g_2 - p2}$
 \State \Return \emph{moves}
@@
 \end{algorithm}
+
+\begin{algorithm}[htb]
+\caption{ { \sc CircularWorkspace}($s_1,s_2,g_1,g_2$)}\label{alg:circularReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[htb]
+\caption{ { \sc PolygonalWorkspace}($s_1,s_2,g_1,g_2, P$)}\label{alg:polygonReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$P$ is a list of the vertices of a convex polygon.
+
+\end{algorithmic}
+\end{algorithm}
+
+%\begin{algorithm}[htb]
+%\caption{ { \sc 2-ParticleConvexPoly}($r_1,r_2,g_1,g_2,P$)}\label{alg:optimalAlg}
+%\begin{algorithmic}[1]
+%%\scriptsize
+%\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+%$(0,0)$ is bottom corner,
+% $P$ is is a convex polygon. 
+% \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
+% \State  \emph{PathList} $\gets \{\}$
+% \State $R \gets   \{ h(\{\},r_1,r_2,g_1,g_2  ) ,\{\},r_1,r_2\} $ \Comment $R$ contains $h$, the admissible heuristic \eqref{eq:admissibleHeuristic}, the move sequence, and the current robot positions
+%\While {$R.r_1 \ne g_1$ \textbf{and} $R.r_2 \ne g_2$}
+%\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
+%\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($R.r_1,R.r_2,g_1,g_2,\theta$)
+%\State $\{d, $ \parbox[t]{.3\linewidth}{%
+% \emph{moves,}$ r_1,r_2\} \gets$\\
+% {\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.\text{\emph{moves}}$)}
+%\State $(\text{\emph{moves}}, r_1,r_2) \gets$ {\sc Rotate}(\emph{moves}$, r_1,r_2,-\theta$)
+%\State   {\sc Push} $\{d, \text{\emph{moves}}, r_1,r_2\} $ onto \emph{PathList}
+%\EndFor
+%\State {\sc Sort}(\emph{PathList}) \Comment sort by admissible heuristic
+%\State $R \gets $ {\sc Pop} first element of \emph{PathList}
+%\EndWhile
+%\State \Return \emph{moves}
+%\end{algorithmic}
+%\end{algorithm}
 

### 2018-01-30 journalWallFriction/experiment.tex
more todos on the paper.  It is missing things!

@@
 
-%\emph{Hardware Demonstration: Position Control of n Robots}\\
-%A hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used to implement  Alg. \ref{alg:PosControlNRobots}. 
-% % Designs for each are available at \cite{Arun2016Thingiverse}. 
-%  The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}.
-%    Only discrete, 1 cm moves in the $x$ and $y$ directions are used.
-The goal configuration highlighted in the top right corner represents a `U' made of seven particles.
-The dark red configuration is the current position of the particles.
-%Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.
-Free particles return to their start positions but particles on the boundary to move laterally, generating a net sliding motion.
-%
-%Fig.
-\ref{fig:construction2d} follows the motion of the particles through iterations  $k$=1, 2, and 7.
-All particles receive the same control inputs, but boundary interactions break the control symmetry.
-Particles reach their goal positions in a first-in, first-out arrangement beginning with the bottom-left particle from the staging zone occupying the top-right position of the build zone.
-%
+To demonstrate Alg. XX experimentally, we performed several tests.
+Each used the same magnetic setup.
+ Two different intestine models were employed, the first a 3D-printed cross-section representation of a small intestine, and the second bovine small intestine.
+
+ \subsection{Magnetic Manipulation Setup}
+ explain the magnetic field strength used, the camera system (briefly)
+
+todo: image of the magnetic setup with scale bar
+explain the particles used
+
+explain the fluid used in the model.
+
+\subsection{Intestine Phantom Model}
+
+methodology: we used xx material, built a model xx big
+
+todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.
+
+\subsection{Bovine Intestine Cross-section}
+todo: procedure for fixating the Bovine tissue sample,
+discussion of the challenges
+
+todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.

### 2018-01-30 journalWallFriction/theory2robAlgs.tex
more todos on the paper.  It is missing things!

@@
  We define a $\Delta$ configuration space as a circular shape that considers all possible $\Delta p$s.
-We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:deltaSpace}.
+We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.
 Reachable set is the part of $\Delta$ configuration space where if one robot touches a wall in a specific location, the other robot can make the required relative distance without causing touching robot to move.
@@
 \begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}\end{overpic}
-\caption{\label{fig:deltaSpace}Left:  the set of points where the red robot is the first to contact the boundary are drawn with a red arc.
+\caption{\label{fig:regionMove}Left:  the set of points where the red robot is the first to contact the boundary are drawn with a red arc.
 The  set of points where the blue robot is the first to contact the boundary are drawn with a blue arc.
@@
 \end{align}
-$\gamma$ is half the angle of the arc that the reachable set's chord has shown in Fig.~\ref{fig:deltaSpace} and is calculated by:
+$\gamma$ is half the angle of the arc that the reachable set's chord has shown in Fig.~\ref{fig:regionMove} and is calculated by:
 \begin{align}
@@
 The equation for the four lines outlining the reachable set can be found as follows:
-
+\todo{place x terms and y terms together (reduce the parenthesis)}
 \begin{align}\label{eq:circlereachable}
@@
 If it is not our final goal, we continue to set the closest point on the reachable set to our current goal until we reach the goal.
+
+ \todo{  how do we find the collision point }
+
 % When a particle contacts a wall, the other particle can move to any position in the reachable set while the first robot is immobile. 
@@
 %\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
 %$(0,0)$ is bottom corner,
@@
  \State $(p_1,p_2) \gets (s_1,s_2) $
- \State $R \gets   \{ p_1,p_2,g_1,g_2  ,moves\} $ \Comment $R$ contains the current robot positions, the goal positions, and the move sequence
- \State $\Delta p \gets p_2-p1$
+\State  moves $\gets \{\}$
+ \State $R \gets   \{ p_1,p_2,g_1,g_2  ,\textrm{moves}\} $ \Comment $R$ contains the current robot positions, the goal positions, and the move sequence
+ \State $\Delta p \gets p_2-p_1$
  \State $\Delta g \gets g_2-g_1$
 \While {$|\Delta g| < |\Delta p|+ \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State Calculate the reachable set using \eqref{eq:circlereachable} or Fig.~\ref{fig:polygonAlg}
-\State $ \Delta cg\gets $nearest point on the reachable set to $\Delta g$
-\State Add the required moves to $moves$ for the particles to achieve $\Delta cg$ 
+\State $R_s$ = Compute reachable set  \Comment use \eqref{eq:circlereachable} or Fig.~\ref{fig:polygonAlg}
+\State $ \Delta g_c\gets $nearest point in $R_s$ to $\Delta g$
+\State  add move to hit the wall
+\State  add move to reach  $ \Delta g_c$
+\State Add the required moves to $moves$ for the particles to achieve $\Delta g_c$ 
 %\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
@@
 \EndWhile
-\State $moves \gets moves+ {g_2 - p2}$
+\State $moves \gets moves+ {g_2 - p_2}$  \Comment translate to goal
 \State \Return \emph{moves}
@@
  Path planning continues using Alg. XX to position the particles to the desired $x$ and $y$ positions. 
- As an example, consider Fig.~\ref{fig:zaxis} which shows a cylinder.
- The blue particle starts in the blue disk and the red particle starts in the red disk. There are two possible optimal paths shown with parallel arrows. 
+ As an example, consider Fig.~\ref{fig:zaxis} which shows a cylindrical workspace.
+ The blue particle starts in the blue disk and the red particle starts in the red disk. 
+ The two candidate shortest-length paths that touch the wall are shown with parallel arrows. 
 Each arrow will cause one of the particles to touch the wall, enabling the other robot to move freely in the  $z$ axis to achieve the required relative position.
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.
-Other workspaces may be better handled by other path planners, such as \cite{MassiveUnifromManipulation}.
+Other workspaces may be better handled by other path planners, such as \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.
 

### 2018-04-12 journalWallFriction/experiment.tex
A lot of new images
Updated the paper with new images, algorithms, edits, etc

@@
 
-To demonstrate Alg. XX experimentally, we performed several tests.
+To demonstrate Alg.~\ref{alg:optimalAlg} experimentally, we performed several tests.
 Each used the same magnetic setup.
@@
  \subsection{Magnetic Manipulation Setup}
- explain the magnetic field strength used, the camera system (briefly)
+
+ The magnetic manipulation system consists of two pairs of electromagnetic coils that consist of iron cores at their centers, and arranged orthogonal to each other.
+The iron core at the center of each coil served to concentrate the magnetic field towards the workspace.
+To set up the magnetic manipulation system for this project, an Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
+Finally, a Basler ace black and white camera was attached to the top of the system focusing on the back-lit workspace.
+
+To obtain experimental data, the test samples which comprised of the Phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in fluid.
+Over the course of experimentation, corn syrup was used to have the best viscosity for the experiments.
+The velocities of particles immersed in it were dampened enough to control their movements with ease.
+Spherical $0.5 mm$ magnets were used as our particles.
+
+ %explain the magnetic field strength used, the camera system (briefly)
 
 todo: image of the magnetic setup with scale bar
-explain the particles used
+%explain the particles used
 
-explain the fluid used in the model.
+%explain the fluid used in the model.
 
@@
 
-methodology: we used xx material, built a model xx big
+The Intestine Phantom Model was used as the first test field of the project and was made to mimic an intestine and its villi.
+The model consists of a circular ring with an outer diameter of $50 mm$, an inner diameter of 46mm, and a thickness of 2mm created using a 3D printer and Fused filament fabrication.
+The model had some $2 mm$ long protrusions on its inner surface to mimic the effects of intestinal villi on the target particles.
 
-todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.
+%methodology: we used xx material, built a model xx big
+
+%todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  %Fake this series so we have an image for now.
 
 \subsection{Bovine Intestine Cross-section}
-todo: procedure for fixating the Bovine tissue sample,
-discussion of the challenges
+%todo: procedure for fixating the Bovine tissue sample,
+%discussion of the challenges
 
-todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.
+This phase of the project involved the use of beef intestines.
+Strips of intestine about 5mm thick were cut and placed in Neutrally buffered formalin for 24 hours for fixation.
+After fixation, each sample was transferred to 70\% ethanol for storage.
+For the experiments, a slice of fixed intestine was attached to the acrylic disc with cyanoacrylate (superglue) and then submerged in corn syrup.
+A drawback of fixing the tissue samples before experimentation is that they tended to shrivel and dry up a few minutes after being removed from the 70\% ethanol.
+%In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.
+This meant more care had to be taken in navigating the magnets to their goal locations to avoid getting them too close to each other.
+
+%todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.
+
+\begin{figure*}
+\centering
+\vspace{1.5em}
+%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
+\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
+\caption{\label{fig:story}
+Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+} \vspace{-1em}
+\end{figure*}

### 2018-04-12 journalWallFriction/intro.tex
One algorithm added
A lot of new images
Updated the paper with new images, algorithms, edits, etc

@@
 % Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
-As a current example, micro- and nano-robots can be manufactured in large numbers, see~\cite{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
-Someday large swarms of robots will be remotely guided
-to assemble structures in parallel and 
- through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
- For each task, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
-Instead, these robots are often controlled by a broadcast signal.
+%As a current example, micro- and nano-robots can be manufactured in large numbers, see~\cite{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
+%Someday large swarms of robots will be remotely guided
+%to assemble structures in parallel and 
+% through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
+% For each task, large numbers of micro robots are required to deliver sufficient payloads, but
+
+The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
@@
 \vspace{1.5em}
-\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
+%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
+\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\end{overpic}
+\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\end{overpic}
 \caption{\label{fig:IntroPic}
@@
 \end{figure}
+%\todo{add the picture of magnetic setup}
 

### 2018-04-12 journalWallFriction/theory.tex
One algorithm added
more todos on the paper.  It is missing things!
Updated the paper with new images, algorithms, edits, etc

@@
 The problem can be simplified by choosing the coordinate system carefully.
-We define the x axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point by a radius e and angle $\beta$, $E=e(\cos \beta,\sin \beta)$.
+We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point by a radius e and angle $\beta$, $E=e(\cos \beta,\sin \beta)$.
 Then define a symmetry point about S of line OP named T.
@@
 Note that the optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
-
- \subsection{Different Polygonal Workspaces}
-
- Fig.
-\ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
-Consider one robot touching each vertex of the workspace.
-For each pair of vertices, compute where the other robot will be if the first robot goes to that vertex.
-If the final position of the second robot is still inside the workspace, then its position is one of the vertices of the reachable set.
-If the point is not inside the polygon, then the intersection of the line that point and the second robot's position with the polygon is one vertex of the reachable set.
-Compute the distance to all the vertices of the workspace from this point.
-By subtracting the relative distance of the robots, then all the vertices of one reachable set are found.
-Doing this for all the vertices of the workspace will give us all the reachable sets.
- \begin{figure}
-\centering
-\renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
-\end{overpic}
-}
-\caption{\label{fig:polygon}{Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
-As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
-}
-\vspace{-1em}
-}
-\end{figure}
-
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}
-\end{overpic}
-\caption{\label{fig:polygonAlg}{ blah blah blah}
-\vspace{-1em}
-}
-\end{figure*}

### 2018-04-12 journalWallFriction/theory2robAlgs.tex
A lot of new images
Updated the paper with new images, algorithms, edits, etc

@@
 
-Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a circular workspace.
+Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a convex workspace.
 In our previous work we used a rectangular workspace.
-We use the same idea here but we modify the algorithm to handle a circular workspace.
+We use the same idea here but we modify the algorithm to handle any convex workspace, then we expand how the algorithm works in circular workspace.
+
+\subsection{Convex Polygonal Workspaces}
+
+ Fig.
+\ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
+Alg.~\ref{alg:polygonReachbale} shows how to compute reachable set for any convex polygonal workspace.
+Consider one robot touching each vertex of the workspace as shown in Fig.\ref{fig:polygonAlg}.
+For each pair of vertices, compute where the other robot will be if the first robot goes to that vertex.
+If the final position of the second robot is still inside the workspace, then its position is one of the vertices of the reachable set.
+If the point is not inside the polygon, then the intersection of the line that point and the second robot's position with the polygon is one vertex of the reachable set.
+Compute the distance to all the vertices of the workspace from this point.
+By subtracting the relative distance of the robots, then all the vertices of one reachable set are found.
+Doing this for all the vertices of the workspace will give us all the reachable sets.
+ \begin{figure}
+\centering
+\renewcommand{\figwid}{0.8\columnwidth}
+{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
+\end{overpic}
+}
+\caption{\label{fig:polygon}{Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
+As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
+}
+\vspace{-1em}
+}
+\end{figure}
+
+%\todo{latex line over the top of two vertices}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.66\columnwidth}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}
+\end{overpic}
+\caption{\label{fig:polygonAlg}{ Steps to generate  the reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
+\vspace{-1em}
+}
+\end{figure*}
+
+\subsection{Circular Workspaces}
+
 % Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
@@
 \centering
-%\begin{overpic}[width=0.5\columnwidth]{startPath.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\end{overpic}\\
+\vspace{1em}
+\centering
 \begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\end{overpic}
@@
 \caption{\label{fig:reachableSet}
-Left circle shows the workspace. %, and the corresponding reachable set when blue robot is touching the wall is shown as a blue region. 
+Top row shows a polygonal workspace with its reachable sets.
+Bottom row, left circle shows the workspace.
+%, and the corresponding reachable set when blue robot is touching the wall is shown as a blue region.
 Right shows the $\Delta$ configuration space and the reachable set that is shown in red is representative of the point we need to go to get to the goal relative distance in one move.%representative of the blue and pink reachable sets shown in the workspace.
@@
 \begin{align}
- \psi \in \Big(\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big)
+ \psi \in \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big]
 % \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
@@
 $\gamma$ is half the angle of the arc that the reachable set's chord has shown in Fig.~\ref{fig:regionMove} and is calculated by:
-\begin{align}
-p_\psi &= r[\cos(\psi ), \sin(\psi )]\\
-d_\perp&= 2 \abs{(s_1.p_\psi - s_2.p_\psi)}\\
+\begin{align}\label{eq:ppsi}
+p_\psi &= r[\cos(\psi ), \sin(\psi )]\\ \label{eq:dprep}
+d_\perp&= 2 \abs{(s_1.p_\psi - s_2.p_\psi)}\\ \label{eq:gamma}
 \gamma &= \cos^{-1} \Big(1-\frac{d_\perp}{r} \Big) 
@@
 The equation for the four lines outlining the reachable set can be found as follows:
-\todo{place x terms and y terms together (reduce the parenthesis)}
 \begin{align}\label{eq:circlereachable}
-l_1 &=  r \Big((\cos\psi_{\min}, \sin\psi_{\min})\\ \nonumber
-& - (\cos(\gamma_c + \psi_{\min}), \sin(\gamma_c + \psi_{\min}))\Big) &  0<\gamma_c< \gamma_{\psi_{\min}},\\ \nonumber
-l_2 &=  r \Big((\cos\psi_{\max}, \sin\psi_{\max})\\ \nonumber
-& - (\cos(\gamma_c + \psi_{\max}), \sin(\gamma_c + \psi_{\max}))\Big) &  \gamma_{\psi_{\max}}<\gamma_c< 0,\\  \nonumber
-l_3 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
-& - (\cos( \psi+\gamma_c ), \sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
-l_4 &=  r \Big((\cos\psi, \sin\psi)\\ \nonumber
-& - (\cos( \psi-\gamma_c ), \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+l_1 &=  r \Big((\cos\psi_{\min}- \cos(\gamma_c + \psi_{\min}) )\\ \nonumber
+& + (\sin\psi_{\min}- \sin(\gamma_c + \psi_{\min}))\Big) &  0<\gamma_c< \gamma_{\psi_{\min}},\\ \nonumber
+l_2 &=  r \Big((\cos\psi_{\max}- \cos(\gamma_c + \psi_{\max}))\\ \nonumber
+& + (\sin\psi_{\max}- \sin(\gamma_c + \psi_{\max}))\Big) &  \gamma_{\psi_{\max}}<\gamma_c< 0,\\  \nonumber
+l_3 &=  r \Big((\cos\psi- \cos( \psi+\gamma_c ) )\\ \nonumber
+& ( \sin\psi-\sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+l_4 &=  r \Big((\cos\psi- \cos( \psi-\gamma_c ))\\ \nonumber
+&  ( \sin\psi- \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
 \end{align}
@@
 &(p.x - c.x)^2 + (p.y - c.y)^2  > r^2\\ \nonumber
- &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi <= -r. \cos\gamma
+ &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r. \cos\gamma
 \end{align}
@@
 This line is a chord of the circle and we find the $\psi$ that makes this region.
+This equation finds this $\psi$:
+  \begin{equation}
+ \psi = \tan^{-1}(\Delta p - \Delta g ).
+ \end{equation}
  Now that $\psi$ is found, we move the particles to make the current goal.
@@
 
- \todo{  how do we find the collision point }
+% \begin{equation}
+% 
+% \end{equation}
 
@@
 
- %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction. 
- \subsection{Square workspace}
-Our algorithm uses an A*-like method to find the shortest path in each move. 
-If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
+% %This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
+% %employing the assumption we have made earlier about the walls' friction. 
+% \subsection{Square workspace}
+%Our algorithm uses an A*-like method to find the shortest path in each move. 
+%If $(\Delta g.x, \Delta g.y)$ is in the reachable set, one robot touches a wall and the other robot zeros the error in one move.
 This is shown as $m_2$ in Fig.
@@
 To find the best place to minimize $m_1$ and $m_3$, the touching robot's goal is reflected on that wall.
-The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
-If the goal configuration can be reached in three moves, then $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
+%The minimum distance to get to the goal in two moves when the robot should touch the wall, is the straight line between the robot and the reflection of the goal position on that wall. 
+%If the goal configuration can be reached in three moves, then $m_1$  makes one particle hit a wall, $m_2$ adjusts the relative spacing error $\Delta e$ to zero, and  $m_3$ takes the particles to their final positions, as shown in Fig.
 \ref{fig:shapeControlMathematica1}b.
-$m_2$ cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
+%$m_2$ cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+% Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
 \ref{fig:reflection}.
-That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
+%That point is selected when possible, but if this point would cause $m_2$ to push the moving robot out of the workspace, the hit point is translated until the moving robot will not leave the workspace.
 If $m_2$ causes the two particles to overlap, we add or subtract $\epsilon$ to $m_2.x$ to avoid collisions.
@@
 ~\ref{fig:epsilon} with three different $\epsilon$ values.
-
-\begin{figure}
-\centering
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
-\caption{\label{fig:TwoRegions}
-Boundary interaction is used to change the relative positions of the robots. Each robot gets the same control input. 
-(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-}
-\end{figure}
-If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
-%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
+%
+%
+%\begin{figure}
+%\centering
+%\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
+%\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+%\caption{\label{fig:TwoRegions}
+%Boundary interaction is used to change the relative positions of the robots. Each robot gets the same control input. 
+%(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
+%(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
+%}
+%\end{figure}
+%If  $\Delta g$ is not in the reachable set, we choose the nearest reachable $\Delta x$ and $\Delta y$ to $\Delta g$. 
+%%For simplicity in this step, we choose the straight line from the touching robot to the wall for the first move.
 This may cause the algorithm not to return the shortest path, but the difference is not significant.
-
-Alg.~\ref{alg:optimalAlg} uses an admissible heuristic that adds the current path length to the greatest distance from each robot to their goal.
+%
+%
+%Alg.~\ref{alg:optimalAlg} uses an admissible heuristic that adds the current path length to the greatest distance from each robot to their goal.
 This heuristic directs exploration by expanding favorable routes first.
-\begin{align}\label{eq:admissibleHeuristic}
-h(\text{\emph{moves}}, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|\text{\emph{moves}}|} \norm{\text{\emph{moves}}_i}  \\
-&+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
-\end{align}
-
-We  exploit symmetry in the solution by labeling the leftmost (or, if they have the same $x$ coordinate, the topmost) robot $r_1$. 
- If $r_1$ is not also the topmost robot, we mirror the coordinate frame about the right boundary. 
- As an example, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$. 
-  Because the leftmost robot is not the topmost robot, we mirror the coordinate frame about the right boundary giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$. 
- After the path is found, we undo the mirroring to the output path. 
-  Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
-   If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$, 180$^{\circ}$, or 270$^{\circ}$ counterclockwise and then push the robot to hit the top wall.
+%\begin{align}\label{eq:admissibleHeuristic}
+%h(\text{\emph{moves}}, r_1,r_2,g_1,g_2) =& \sum_{i=1}^{|\text{\emph{moves}}|} \norm{\text{\emph{moves}}_i}  \\
+%&+  \max( \norm{g_1-r_1}, \norm{g_2-r_2} ) \nonumber
+%\end{align}
+%
+%
+%We  exploit symmetry in the solution by labeling the leftmost (or, if they have the same $x$ coordinate, the topmost) robot $r_1$. 
+% If $r_1$ is not also the topmost robot, we mirror the coordinate frame about the right boundary. 
+% As an example, consider the two starting positions, $r_1 =  (0.2, 0.2) $ and $r_2 = (0.8, 0.8)$. 
+%  Because the leftmost robot is not the topmost robot, we mirror the coordinate frame about the right boundary giving $r_1 = (0.2, 0.8)$ and $r_2 = (0.8,0.2)$. 
+% After the path is found, we undo the mirroring to the output path. 
+%  Similarly, we exploit rotational symmetry and assume the command pushes a robot to hit the top wall.
+%   If a different wall is selected, we rotate the coordinate frame by 90$^{\circ}$, 180$^{\circ}$, or 270$^{\circ}$ counterclockwise and then push the robot to hit the top wall.
 After the path is found, we undo the rotation.
-   This symmetry allows us to use a single function, Alg. \ref{alg:Wallup},  for collisions with all four walls.
+%   This symmetry allows us to use a single function, Alg. \ref{alg:Wallup},  for collisions with all four walls. 
+%
+%
 
@@
 \While {$|\Delta g| < |\Delta p|+ \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_s$ = Compute reachable set  \Comment use \eqref{eq:circlereachable} or Fig.~\ref{fig:polygonAlg}
+\State $R_s$ = Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
 \State $ \Delta g_c\gets $nearest point in $R_s$ to $\Delta g$
@@
 \State  add move to reach  $ \Delta g_c$
-\State Add the required moves to $moves$ for the particles to achieve $\Delta g_c$ 
+\State Add the required moves to moves for the particles to achieve $\Delta g_c$ 
 %\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
@@
 \EndWhile
-\State $moves \gets moves+ {g_2 - p_2}$  \Comment translate to goal
-\State \Return \emph{moves}
+\State $\textrm{moves} \gets \textrm{moves}+ {g_2 - p_2}$  \Comment translate to goal
+\State \Return moves
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[htb]
+\caption{ { \sc PolygonalWorkspace}($s_1,s_2,g_1,g_2, P$)}\label{alg:polygonReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+$P$ is a list of the vertices of a convex polygon. RSet contains all the reachable polygons.
+\For {$i$ in $P$}
+\State $p_{i} \gets s_1 + s_2 - i$
+\State $p_{i+1} \gets s_1 + s_2 - (i+1)$
+\State $L \gets$ line with $(p_{11}, p_{12})$
+\State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon
+\If {$p_{i}\notin$ polygon}
+\State $p_{i} \gets l_i$
+\EndIf
+\State $v_i \gets$ All vertices $\in [ l_i , l_{i+1} ] $
+\State $d_i \gets$ All vectors from $p_i$ to $[ l_i , i]\cup  p_{i+1}$ to $ [i+1 , l_{i+1} ] $
+\State RSet $\gets$ RSet  + polygon ($s_2 - (s_1 +d_i)$)
+\EndFor
+\State Return RSet
 \end{algorithmic}
@@
 \Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[htb]
-\caption{ { \sc PolygonalWorkspace}($s_1,s_2,g_1,g_2, P$)}\label{alg:polygonReachbale}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$P$ is a list of the vertices of a convex polygon.
-
+\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
+\State Calculate $p_{\psi}$ \Comment use \eqref{eq:ppsi}
+\State Calculate $\gamma$ \Comment use \eqref{eq:gamma}
+\State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
+\State polygon $\gets$ Join($l_1, l_2, l_3, l_4$)
+\State return polygon
 \end{algorithmic}
@@
 
-\begin{algorithm}
-\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, $\emph{moves})}\label{alg:Wallup}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner,
- $L$ is length of the walls. 
- The array \emph{moves} is the current sequence of moves up to the current position.
- Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots, then undo the mirroring before returning.
- $\epsilon $ is a small, nonzero, user-specified value.
+%\begin{algorithm}
+%\caption{{\sc PlanMoveUp}($r_1,r_2,g_1,g_2,L, $\emph{moves})}\label{alg:Wallup}
+%\begin{algorithmic}[1]
+%%\scriptsize
+%\Require knowledge of current $(r_1,r_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+%$(0,0)$ is bottom corner,
+% $L$ is length of the walls. 
+% The array \emph{moves} is the current sequence of moves up to the current position.
+% Assume $r_1.x < r_2.x$ and $r_1.y \geq r_2.y$. If not, mirror the coordinate frame and swap the robots, then undo the mirroring before returning.
+% $\epsilon $ is a small, nonzero, user-specified value.
+% 
+%\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least $\epsilon$ distance from each other.  
+%$m_1$ is the first move toward the wall or goal.
+% $m_2$ is the second move adjusting $\Delta e$.
+%\State $\Delta e \gets (g_2 - g_1)- (r_2- r_1)$
+%\If {$\Delta e = (0,0)$} \Comment{base case}
+%\State $m_1  \gets g_2 -  r_2$
+%\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1 \}$
+%\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+%\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
+%\EndIf
+%\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
+%\State $m_1 \gets \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
+%\If {$r_2.x + m_1.x >L$}
+%\State $m_1.x \gets 1- r_2.x$
+%%\EndIf
+%\Else \textbf{ if} {$r_2.x + m_1.x < 0$}  \textbf{then}
+%\State $m_1.x \gets -r_2.x$
+%\EndIf
+%\Else 
+%\State $m_1 = (0, 1-r_1.y)$
+%\State $\Delta g \gets $ closest reachable $(\Delta x,\Delta y)$.
+%%Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
+%%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
+%\EndIf
+%\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1\}$
+%\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
+%\State $m_2 \gets \Delta g -  (r_2- r_1)$
+%\If {robots on each other \textbf{or} on the wall}
+%\State Add $\pm \epsilon$ to $m_2.x$ to avoid collision
+%\EndIf
+%\State \emph{moves} $ \gets \{ \text{\emph{moves}},   m_2\}$
+%\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
+%\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
+%\end{algorithmic}
+%\end{algorithm}
 
-\Ensure $(g_1, g_2) , (r_1, r_2)$ all at least $\epsilon$ distance from walls the goals and starting points have at least $\epsilon$ distance from each other.  
-$m_1$ is the first move toward the wall or goal.
- $m_2$ is the second move adjusting $\Delta e$.
-\State $\Delta e \gets (g_2 - g_1)- (r_2- r_1)$
-\If {$\Delta e = (0,0)$} \Comment{base case}
-\State $m_1  \gets g_2 -  r_2$
-\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1 \}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
-\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
-\EndIf
-\If {$r_2.x - r_1.x - 1 + 2 \epsilon \leq  \Delta g.x \leq  1 ~\textbf{and}~r_2.y - r_1.y \leq \Delta g.y \leq  0$} \Comment{$\Delta g \in $ reachable region}
-\State $m_1 \gets \left(\frac{1-r_1.y}{2-g_1.y-r_1.y} (g_1.x -r_1.x), 1-r_1.y \right)$
-\If {$r_2.x + m_1.x >L$}
-\State $m_1.x \gets 1- r_2.x$
-%\EndIf
-\Else \textbf{ if} {$r_2.x + m_1.x < 0$}  \textbf{then}
-\State $m_1.x \gets -r_2.x$
-\EndIf
-\Else 
-\State $m_1 = (0, 1-r_1.y)$
-\State $\Delta g \gets $ closest reachable $(\Delta x,\Delta y)$.
-%Adjust $\Delta g_x$ and $\Delta g_y$ %\Comment{To go to the nearest $\Delta g_x$ and $\Delta g_y$}
-%\State $\Delta g_x = r_2.x - r_1.x - 1 + 2 \epsilon
-\EndIf
-\State \emph{moves}$ \gets \{ \text{\emph{moves}},   m_1\}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_1,r_1,r_2)$
-\State $m_2 \gets \Delta g -  (r_2- r_1)$
-\If {robots on each other \textbf{or} on the wall}
-\State Add $\pm \epsilon$ to $m_2.x$ to avoid collision
-\EndIf
-\State \emph{moves} $ \gets \{ \text{\emph{moves}},   m_2\}$
-\State $(r_1,r_2) \gets ${\sc ApplyMove}$(m_2,r_1,r_2)$
-\State \Return $\{h(\text{\emph{moves}}, r_1,r_2,g_1,g_2), \text{\emph{moves}},r_1,r_2\}$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{figure}
-\centering
-\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
-\caption{\label{fig:reflection}
-If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
-The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
-} \vspace{-1em}
-\end{figure}
+%
+%\begin{figure}
+%\centering
+%\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+%\caption{\label{fig:reflection}
+%If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error $\Delta e$ to zero, and the third move takes the particles to their final positions.
+%The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+% Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location.
+%} \vspace{-1em}
+%\end{figure}
 
@@
  Then the particles are actuated away from the boundary and to the appropriate $z$ positions.
- Path planning continues using Alg. XX to position the particles to the desired $x$ and $y$ positions. 
+ Path planning continues using Alg.~\ref{alg:optimalAlg} to position the particles to the desired $x$ and $y$ positions. 
  As an example, consider Fig.~\ref{fig:zaxis} which shows a cylindrical workspace.
@@
 Extending the algorithm to position the particles in 3D.
-} \vspace{-1em}
+} %\vspace{-1em}
 \end{figure}

### 2018-04-17 IROS17/theory2robAlgs.tex
Edits to the paper and pictures.

@@
 \caption{\label{fig:TwoRegions}
-Boundary interaction is used to change the relative positions of the robots. Each robot gets the same control input. 
-(left) If robot 2 hits the bottom wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
-(right) Similarly, if robot 2 hits the right wall before robot 1 reaches a wall, robot 2 can reach anywhere along the green line, and  robot 1 can move to anywhere in the shaded area. 
+Boundary interaction is used to change the relative positions of the particles. Each particle gets the same control input. 
+(left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
+(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
 }

### 2018-04-17 journalWallFriction/SwarmShapeControl.tex
One algorithm added
A lot of new images
Edits to the paper and pictures.

@@
 % paper title
-\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using a Shared Input}
+\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using The Same Control Input}
 
 % You will get a Paper-ID when submitting a pdf file to the conference system
-\author{Shiva Shahrokhi, Jingang Shi, and Aaron T. Becker% <-this % stops a space
+\author{Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker% <-this % stops a space
 \thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
@@
 Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
+ This paper investigates control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
 Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
 Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
-The walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
+However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
 In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
-However, in vivo environments are usually not square. 
-This work extends the previous work to convex workspaces and 3D positioning. 
-This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
+Because in vivo environments are usually not square, this work extends the previous work to convex workspaces including circles and 3D positioning. 
+This paper also implements the algorithms using a hardware setup inspired by gastrointestinal tract.
 

### 2018-04-17 journalWallFriction/intro.tex
Edits to the paper and pictures.

@@
 %\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
-\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\end{overpic}
-\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\end{overpic}
+\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
+\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\put(28,-8){magnetic setup}\end{overpic}
+\vspace{1em}
 \caption{\label{fig:IntroPic}
-Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+Workspace and magnetic setup for an experiment of positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
 } \vspace{-1em}
@@
 
-This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which consider only square workspaces.
+This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces.
 This work extends the analysis to convex workspaces and 3D positioning.
-This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
+This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-04-17 journalWallFriction/theory.tex
Edits to the paper and pictures.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action.
+Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g.
+PRM and RRT* are not suited \cite{lavalle2006planning,kavraki1996probabilistic}.
+Instead, this paper employs a graph search.
+This section starts with a boundary interaction model in subsection \ref{subsec:WallFriction}.
 
+Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle.
+In subsections \ref{subsec:square} and \ref{subsec:circular} we provide shortest-path results for two representative workspaces, squares and circles.
 \subsection{Boundary Interaction Model}\label{subsec:WallFriction}
@@
  \\
- 0 & \textrm{else} 
+ 0 & \textrm{else}.
  \end{cases}\nonumber
  \end{align}
- Here $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$, and
- $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary.
+ Here  $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary, and
+ $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$.
 
@@
 
- \subsection{Shortest Path}
+ \subsection{Shortest Path in Square}\label{subsec:square}
+
+ If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error  to zero, and  the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
+\ref{fig:reflection}.
+
+\begin{figure}
+\centering
+\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\caption{\label{fig:reflection}
+The shortest path between two points in a square. Reflecting the goal position across the boundary wall and plotting a straight line gives the shortest path.
+} \vspace{-1em}
+\end{figure}
+
+ \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
  The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments shown in Fig.~\ref{fig:shortestpath}.
 The problem can be simplified by choosing the coordinate system carefully.
-We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point by a radius e and angle $\beta$, $E=e(\cos \beta,\sin \beta)$.
-Then define a symmetry point about S of line OP named T.
+We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point $E$ by a radius $e$ and angle $\beta$, $E=e(\cos \beta,\sin \beta)$.
 Then the length of the two line segments is
  \begin{equation}
-\scriptsize{ \sqrt{(s-\cos \theta)^2+(-\sin \theta)^2}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}}
+\scriptsize{ \sqrt{(s-\cos \theta)^2+\sin^2 \theta}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}},
  \end{equation}
@@
 \caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference.
-The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.}
+The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.
+See \cite{shortestPathMathematica}.}
 %\vspace{-1em}
@@
 
- The length of the two line segments as a function of $\theta$ is drawn in the right plot.
+ The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}.
 There are several simple solutions.
@@
 Label the origin $O$.
-The optimal solution shows that the angle $\angle{OPS}$ (from the origin to P to S) is the same as the angle $\angle{OPE}$ (from the origin to P to E).
+The optimal solution shows that the angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$).
 We name these angles $\alpha$.
-This can be proved by drawing an ellipse whose foci are S and E.
-When the ellipse is tangent to the circle, the point of tangency is exactly P.
-  Since the distance from the origin to P is always 1, we can set up three equalities using the law of sines:
- From triangle OSP: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{SP}$, and from triangle OEP: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{EP}$.
-If we mirror the point S about the $\theta$ axis and label this point C, from triangle CEO: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{CE}$.
+This can be proved by drawing an ellipse whose foci are $S$ and $E$.
+When the ellipse is tangent to the circle, the point of tangency is exactly $P$.
+  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
+ From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$.
+If we mirror the point $S$ about the $\theta$ axis and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.
 
  Simplifying this system of equations results in: $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$.
-Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi<=\beta<=\pi$.
+Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$.
 We evaluate each and select the solution that results in the shortest length path.
-Note that the optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+For an interactive mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.

### 2018-04-17 journalWallFriction/theory2robAlgs.tex
Edits to the paper and pictures.

@@
 
-Alg.~\ref{alg:optimalAlg} uses non-slip contacts with walls to arbitrarily position two robots in a convex workspace.
-In our previous work we used a rectangular workspace.
-We use the same idea here but we modify the algorithm to handle any convex workspace, then we expand how the algorithm works in circular workspace.
+This section presents algorithms that use non-slip contacts with walls to arbitrarily position two robots in a convex workspace.
+Our previous work used a square workspace.
+We modify the algorithm to handle any convex workspace, and address the special case of circular workspaces.
+In the last subsection we present techniques to extend the workspace to 3D.
 
-\subsection{Convex Polygonal Workspaces}
+Workspaces are 2D convex polygons with no internal obstacles.
+The configuration space for two particles is a four dimensional manifold.
+Translating both particles the same amount is a trivial operation, but changing the relative positions requires boundary interaction.
+For this reason, our algorithms use the two dimensional $\Delta$ configuration space, defined as the difference in position of the first particle and the second particle.
 
- Fig.
-\ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
-Alg.~\ref{alg:polygonReachbale} shows how to compute reachable set for any convex polygonal workspace.
-Consider one robot touching each vertex of the workspace as shown in Fig.\ref{fig:polygonAlg}.
-For each pair of vertices, compute where the other robot will be if the first robot goes to that vertex.
+\subsection{Two Particle Path Planning}
+
+Assume two particles are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+Denote the current positions of the particles  $p_1$ and $p_2$ and the current distance between the particles is $d$.
+Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
+The algorithm assigns a uniform control input at every instance.
+The goal is to move the particles to the goal positions using a shared control input.
+We do this by first moving them to the correct relative position and then translating the particles to the goal.
+The correct relative position is when
+ $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
+
+ We define a $\Delta$ configuration space as a circular shape that considers all possible $\Delta p$ values.
+We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.
+The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+We will discuss how to make reachable sets in the next subsections.
+ The algorithm first makes the reachable set, and then checks if the goal relative position is in the reachable set.
+If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ would be our current goal.
+We move the particles to make the current goal.
+If current goal is our final goal, we go to the final goal position.
+If it is not our final goal, we continue to set the closest point on the reachable set to our current goal until we reach the goal.
+\begin{algorithm}[htb]
+\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P$)}\label{alg:optimalAlg}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
+%$(0,0)$ is bottom corner,
+ $P$ is a description of the workspace.
+% \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
+% \State  \emph{PathList} $\gets \{\}$
+ \State $(p_1,p_2) \gets (s_1,s_2) $ \Comment $p_1 , p_2$ are current positions
+\State  moves $\gets \{\}$
+% \State $R \gets   \{ p_1,p_2,g_1,g_2  ,\textrm{moves}\} $ \Comment $R$ contains the current particle positions, the goal positions, and the move sequence
+ \State $\Delta p \gets p_2-p_1$
+ \State $\Delta g \gets g_2-g_1$
+\While {$||\Delta p - \Delta g|| > \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
+\State $R_{\textrm{SET}}$ = Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
+\State $ \Delta g_c\gets $nearest point in $R_{\textrm{SET}}$ to $\Delta g$
+\State $m \gets $move-to-wall
+\State Append($m$, moves)
+\State $p_1$ = ApplyMove($p_1,m$)
+\State $p_2$ = ApplyMove($p_2,m$)
+ \State $\Delta p \gets p_2-p_1$
+\EndWhile
+\State Append($g2-p2$, moves)  \Comment translate to goal
+\State \Return moves
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Convex Polygonal Workspaces: Reachable Set}
+
+ Fig. \ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
+
+% Consider one particle touching each vertex of the workspace as shown in Fig.~\ref{fig:polygonAlg}.
+%  For each pair of vertices, compute where the other robot will be if the first robot goes to that vertex.
 If the final position of the second robot is still inside the workspace, then its position is one of the vertices of the reachable set.
@@
 
+\begin{figure}
+\centering
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
+\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+\caption{\label{fig:TwoRegions}
+Boundary interaction is used to change the relative positions of the particles. Each particle gets the same control input. 
+(left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
+(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
+}
+\end{figure}
+If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
+We extend a square workspace to any convex workspace by computing the reachable set for all the sides of the polygon.
+Fig.~\ref{fig:polygonAlg} shows different cases with the particles and an arbitrary side.
+If particle 2 is touching the $i^{th}$ side, we compute if particle 1 will stay in the polygon or not.
+Then we generate the reachable set using all the vertices of the polygon and potential place of particle 1.
+We unify all the reachable sets corresponding to all the sides.
+Alg.~\ref{alg:polygonReachbale} computes reachable set for any convex polygonal workspace.
+ \begin{algorithm}[htb]
+\caption{ { \sc ReachableSetPolygon}($s_1,s_2,g_1,g_2, P$)}\label{alg:polygonReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
+$P$ is a list of the vertices of a convex polygon. %RSet contains all the reachable polygons.
+\For {$i$ in $P$}
+\State $p_{i} \gets s_1 + s_2 - i$
+\State $p_{i+1} \gets s_1 + s_2 - (i+1)$
+\State $L \gets$ line with $(p_{11}, p_{12})$
+\State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon $P$
+\If {$p_{i}\notin$ polygon}
+\State $p_{i} \gets l_i$
+\EndIf
+\State $v_i \gets$ All vertices $\in [ l_i , l_{i+1} ] $
+\State $d_i \gets$ All vectors from $p_i$ to $[ l_i , i]\cup  p_{i+1}$ to $ [i+1 , l_{i+1} ] $
+\State Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
+\EndFor
+\State Return $R_{\textrm{SET}}$
+\end{algorithmic}
+\end{algorithm}
 %\todo{latex line over the top of two vertices}
@@
 \renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){Case 1:$s_1$ does not hit boundary}\put(28,39){$s_2$}\put(63,50){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){Case 2:$s_1$ hits boundary}\put(20,40){$s_2$}\put(55,58){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){Generates polygon}
 \end{overpic}
@@
 
-\subsection{Circular Workspaces}
+\subsection{Circular Workspaces: Reachable Set}
 
@@
 
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-Denote the current positions of the robots  $p_1$ and $p_2$ and the current distance between the robots is $d$.
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
-The algorithm assigns a uniform control input at every instance.
-The goal is to move the particles to the goal positions using a shared control input.
-We do this by first moving them to the correct relative position and then translating the particles to the goal.
-The first step minimizes
- $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
-
- We define a $\Delta$ configuration space as a circular shape that considers all possible $\Delta p$s.
-We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.
-Reachable set is the part of $\Delta$ configuration space where if one robot touches a wall in a specific location, the other robot can make the required relative distance without causing touching robot to move.
-To compute reachable set for circular workspace, first we considered all passible hitting point locations in the workspace.
- The set of boundary points that a robot can touch before the  other robot  touches is an arc of angle $2(\pi - \frac{\arcsin{d}}{r})$, where $d= |s_1 - s_2|$ and $r$ is the radius of the circle.
+To compute reachable set for a circular workspace, first we consider all passible hitting point locations in the workspace.
+ The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d}}{r})$, where $d= |s_1 - s_2|$ and $r$ is the radius of the circle.
  %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
@@
 \begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}\end{overpic}
-\caption{\label{fig:regionMove}Left:  the set of points where the red robot is the first to contact the boundary are drawn with a red arc.
-The  set of points where the blue robot is the first to contact the boundary are drawn with a blue arc.
-The possible points for the blue and pink particles to touch the boundary is shown in blue and pink arcs.
-Right: When the blue particle is touching a wall (blue square) the other particle (pink square) can go anywhere in the reachable set (blue region).}
+\caption{\label{fig:regionMove}
+% the set of points where the red particle is the first to contact the boundary are drawn with a red arc.
+The  set of points where the blue particle is the first to contact the boundary are drawn with a blue arc.
+Left top: The possible points for the blue and red particles to touch the boundary is shown in blue and red arcs.
+Left bottom: When the blue particle is touching a wall (blue square) the other particle (red square) can go anywhere in the reachable set (blue region).
+Right: The $\Delta$ configuration space for the corresponded starting positions of the particles is shown.}
 \end{figure}
@@
 
-There are infinite reachable sets, parameterized by first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove}.
+A circle has infinite number of sides, thus infinite reachable sets. However, parameterized by the first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where 
 %We make the reachable sets by knowing the angle of the chord, $\psi$, shown in Fig.~\ref{fig:psigamma}. Minimum and maximum values of $\psi$ can be calculated in the following equation.
 \begin{align}
- \psi \in \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big]
+ \psi \in \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big].
 % \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
 \end{align}
-$\gamma$ is half the angle of the arc that the reachable set's chord has shown in Fig.~\ref{fig:regionMove} and is calculated by:
+The possible first contact locations line on an arc with interior angle $\gamma$ is calculated by:
 \begin{align}\label{eq:ppsi}
@@
 
-The equation for the four lines outlining the reachable set can be found as follows:
+The equations for the four lines outlining the reachable set are as follows:
 \begin{align}\label{eq:circlereachable}
@@
 l_3 &=  r \Big((\cos\psi- \cos( \psi+\gamma_c ) )\\ \nonumber
-& ( \sin\psi-\sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+&+ ( \sin\psi-\sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
 l_4 &=  r \Big((\cos\psi- \cos( \psi-\gamma_c ))\\ \nonumber
-&  ( \sin\psi- \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+&+  ( \sin\psi- \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
 \end{align}
 We combine these boundaries to make the reachable set. 
-The algorithm first makes the reachable set, and then checks if the goal relative position is in the reachable set.
-If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ would be our current goal.
 We need to find a $\psi$ that would enable us to reach to the required relative goal distance.
@@
 &(p.x - c.x)^2 + (p.y - c.y)^2  > r^2\\ \nonumber
- &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r. \cos\gamma
+ &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r\cdot \cos\gamma
 \end{align}
 
-Here $c$ is coordinate of the circle and $r$ is the radius of the workspace.
+Here $c$ is the coordinate of the workspace center and $r$ is the radius of the workspace.
  If $\Delta g$ is not in the region made by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g$ and the current relative position, $\Delta s$.
@@
  \end{equation}
- Now that $\psi$ is found, we move the particles to make the current goal.
+
+\begin{algorithm}[htb]
+\caption{ { \sc ReachableSetCircle}($s_1,s_2,g_1,g_2$)}\label{alg:circularReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
+\State Calculate $p_{\psi}$ \Comment use \eqref{eq:ppsi}
+\State Calculate $\gamma$ \Comment use \eqref{eq:gamma}
+\State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
+\State Return Join($l_1, l_2, l_3, l_4$)
+\end{algorithmic}
+\end{algorithm}
+ %Now that $\psi$ is found, we move the particles to make the current goal.
 If current goal is our final goal, we go to the final goal position.
@@
 %
-
-\begin{algorithm}[htb]
-\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P$)}\label{alg:optimalAlg}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
-%$(0,0)$ is bottom corner,
- $P$ is a list of the vertices of a convex polygon. 
-% \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
-% \State  \emph{PathList} $\gets \{\}$
- \State $(p_1,p_2) \gets (s_1,s_2) $
-\State  moves $\gets \{\}$
- \State $R \gets   \{ p_1,p_2,g_1,g_2  ,\textrm{moves}\} $ \Comment $R$ contains the current robot positions, the goal positions, and the move sequence
- \State $\Delta p \gets p_2-p_1$
- \State $\Delta g \gets g_2-g_1$
-\While {$|\Delta g| < |\Delta p|+ \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_s$ = Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
-\State $ \Delta g_c\gets $nearest point in $R_s$ to $\Delta g$
-\State  add move to hit the wall
-\State  add move to reach  $ \Delta g_c$
-\State Add the required moves to moves for the particles to achieve $\Delta g_c$ 
-%\For{ $\theta \in \{0^\circ, 90^\circ, 180^\circ, 270^\circ \}$ }
-%\State $(r_1,r_2,g_1,g_2) \gets$ {\sc Rotate}($R.r_1,R.r_2,g_1,g_2,\theta$)
-%\State $\{d, $ \parbox[t]{.3\linewidth}{%
-% \emph{moves,}$ r_1,r_2\} \gets$\\
- %{\sc  PlanMoveUp}($r_1,r_2,g_1,g_2,L, P.\text{\emph{moves}}$)}
-%\State $(\text{\emph{moves}}, r_1,r_2) \gets$ {\sc Rotate}(\emph{moves}$, r_1,r_2,-\theta$)
-%\State   {\sc Push} $\{d, \text{\emph{moves}}, r_1,r_2\} $ onto \emph{PathList}
-%\EndFor
-%\State {\sc Sort}(\emph{PathList}) \Comment sort by admissible heuristic
-%\State $R \gets $ {\sc Pop} first element of \emph{PathList}
-\EndWhile
-\State $\textrm{moves} \gets \textrm{moves}+ {g_2 - p_2}$  \Comment translate to goal
-\State \Return moves
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[htb]
-\caption{ { \sc PolygonalWorkspace}($s_1,s_2,g_1,g_2, P$)}\label{alg:polygonReachbale}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-$P$ is a list of the vertices of a convex polygon. RSet contains all the reachable polygons.
-\For {$i$ in $P$}
-\State $p_{i} \gets s_1 + s_2 - i$
-\State $p_{i+1} \gets s_1 + s_2 - (i+1)$
-\State $L \gets$ line with $(p_{11}, p_{12})$
-\State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon
-\If {$p_{i}\notin$ polygon}
-\State $p_{i} \gets l_i$
-\EndIf
-\State $v_i \gets$ All vertices $\in [ l_i , l_{i+1} ] $
-\State $d_i \gets$ All vectors from $p_i$ to $[ l_i , i]\cup  p_{i+1}$ to $ [i+1 , l_{i+1} ] $
-\State RSet $\gets$ RSet  + polygon ($s_2 - (s_1 +d_i)$)
-\EndFor
-\State Return RSet
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[htb]
-\caption{ { \sc CircularWorkspace}($s_1,s_2,g_1,g_2$)}\label{alg:circularReachbale}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
-\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
-\State Calculate $p_{\psi}$ \Comment use \eqref{eq:ppsi}
-\State Calculate $\gamma$ \Comment use \eqref{eq:gamma}
-\State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
-\State polygon $\gets$ Join($l_1, l_2, l_3, l_4$)
-\State return polygon
-\end{algorithmic}
-\end{algorithm}
 

### 2018-04-20 journalWallFriction/experiment.tex
Edits to the paper and pictures.
Edited a lot of parts of the paper

@@
 The Intestine Phantom Model was used as the first test field of the project and was made to mimic an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $50 mm$, an inner diameter of 46mm, and a thickness of 2mm created using a 3D printer and Fused filament fabrication.
+The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and a thickness of 2mm created using a 3D printer and Fused filament fabrication.
 The model had some $2 mm$ long protrusions on its inner surface to mimic the effects of intestinal villi on the target particles.
+Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.
 
@@
 
-\begin{figure*}
+\begin{figure*}\label{fig:3dPrinted}
 \centering

### 2018-04-20 journalWallFriction/simulation.tex
Updated the paper with new images, algorithms, edits, etc
Edits to the paper and pictures.
Edited a lot of parts of the paper

@@
 
-\subsection{Position Control of Two Robots}
+%\subsection{Position Control of Two Robots}
 
-Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using point robots (radius = $0$). 
+Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. 
 %An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
@@
 \begin{overpic}[width=0.49\columnwidth]{distcontour.pdf}\end{overpic}
+ \vspace{-2em}
 \caption{\label{fig:contour}
 Plots showing the algorithm with one goal on the boundary.
-} \vspace{-1em}
+}
 \end{figure}
@@
 \caption{\label{fig:deltanumdist}
-Plots showing decreasing error when the number of moves grows.
-} \vspace{-1em}
+Plots showing decreasing error as the number of moves grows.
+} 
+\vspace{-1em}
 \end{figure}
@@
 \end{overpic}
-}\caption{\label{fig:contourPlots}{Starting positions of robots $1$ and $2$ and goal position of robot $2$ are fixed, and $\epsilon=0.001$.
+}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
  The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}

### 2018-04-20 journalWallFriction/theory.tex
Edited a lot of parts of the paper

@@
 These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
- In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
-  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
+ In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
+  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of particles that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
   As in our model, fluid flowing in a pipe has zero velocity along the boundary.
@@
 
- \subsection{Shortest Path in Square}\label{subsec:square}
+ \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
 
- If the goal configuration can be reached in three moves, the first move makes one particle hit a wall, the second move adjusts the relative spacing error  to zero, and  the third move takes the particles to their final positions.
-The second move cannot be shortened, so optimization depends on choosing the location where the robot hits the wall. 
+ If the goal configuration can be reached in three moves, the shortest path has a simple solution.
+The first move makes one particle hit a wall, the second move adjusts the relative spacing error  to zero, and  the third move takes the particles to their final positions.
+The second move cannot be shortened, so optimization depends on choosing the location where the particle hits the wall. 
  Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
 \ref{fig:reflection}.
+There are four walls, and four candidate solutions.
 
@@
  \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
- The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments shown in Fig.~\ref{fig:shortestpath}.
+ The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}.
 The problem can be simplified by choosing the coordinate system carefully.
-We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis $P=(\cos \theta,\sin \theta)$, and the final point $E$ by a radius $e$ and angle $\beta$, $E=e(\cos \beta,\sin \beta)$.
+We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis: $P=(\cos \theta,\sin \theta)$.
+Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$.
 Then the length of the two line segments is
@@
  \begin{equation}
- \sqrt{1+e^2-2 e \cos(\beta-\theta)}+\sqrt{1+s^2-2 s \cos \theta}. 
+\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
  \end{equation}
@@
 
- Simplifying this system of equations results in: $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$.
+ Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$.
 Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$.
@@
 This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
-For an interactive mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.
+For an interactive Mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.

### 2018-04-20 journalWallFriction/theory2robAlgs.tex
Edited a lot of parts of the paper

@@
-\section{Position Control of Two Robots Using Boundary Interaction}\label{sec:PostionControl2Robots}
+\section{Position Control of Two Particles Using Boundary Interaction}\label{sec:PostionControl2Robots}
 
-This section presents algorithms that use non-slip contacts with walls to arbitrarily position two robots in a convex workspace.
-Our previous work used a square workspace.
-We modify the algorithm to handle any convex workspace, and address the special case of circular workspaces.
-In the last subsection we present techniques to extend the workspace to 3D.
+This section presents algorithms that use non-slip contacts with walls to arbitrarily position two particles in a convex workspace.
+Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
+The algorithm can now handle any convex workspace, inducing the special limit case of a circular workspace.
+In the last subsection we present techniques to control 3D positioning of two particles.
 
 Workspaces are 2D convex polygons with no internal obstacles. 
-The configuration space for two particles is a four dimensional manifold.
-Translating both particles the same amount is a trivial operation, but changing the relative positions requires boundary interaction.
-For this reason, our algorithms use the two dimensional $\Delta$ configuration space, defined as the difference in position of the first particle and the second particle.
+ Assume two particles are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
+Denote the current positions of the particles  $p_1$ and $p_2$
 
@@
 
-Assume two particles are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-Denote the current positions of the particles  $p_1$ and $p_2$ and the current distance between the particles is $d$.
+The configuration space for two particles is a four dimensional manifold.
+Translating both particles the same amount is a trivial operation, but changing the relative positions requires boundary interaction.
+For this reason, our algorithms use the two dimensional $\Delta$ configuration space, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
  Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
@@
 The correct relative position is when
- $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
+ $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||< \epsilon$, where $\epsilon$ is an arbitrary small number.
 
- We define a $\Delta$ configuration space as a circular shape that considers all possible $\Delta p$ values.
-We also show the starting and ending relative distance as $\Delta s$ and $\Delta e$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.
-The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+The $\Delta$ configuration space is a set of all possible $\Delta p$ values.
+ $\Delta$ configuration spaces for a representative set of workspaces are shown in Fig.~\ref{fig:polygon}.
+The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+ We also show the starting and ending relative distance as $\Delta s$ and $\Delta g$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
 We will discuss how to make reachable sets in the next subsections.
- The algorithm first makes the reachable set, and then checks if the goal relative position is in the reachable set.
-If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ would be our current goal.
-We move the particles to make the current goal.
-If current goal is our final goal, we go to the final goal position.
-If it is not our final goal, we continue to set the closest point on the reachable set to our current goal until we reach the goal.
+ The next subsections give procedures to compute the reachable set.
+ The algorithm first computes the reachable set, and then checks if the goal relative position is in the reachable set.
+If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ becomes the current goal.
+We move the particles to achieve the current goal.
+If current goal is the final goal, the correct \emph{relative} position has been achieved, and a final translation delivers both particles to their goal destinations.
+%we go to the final goal position.
+If it is not our final goal,
+ Otherwise, we iterate until we reach the goal. 
 \begin{algorithm}[htb]
@@
 \While {$||\Delta p - \Delta g|| > \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_{\textrm{SET}}$ = Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
+\State $R_{\textrm{SET}}\gets$  Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
 \State $ \Delta g_c\gets $nearest point in $R_{\textrm{SET}}$ to $\Delta g$
-\State $m \gets $move-to-wall
-\State Append($m$, moves)
-\State $p_1$ = ApplyMove($p_1,m$)
-\State $p_2$ = ApplyMove($p_2,m$)
+\State $m \gets $move-to-wall corresponding to $\Delta g_c$
+\State moves $\gets$ Append(moves, $m$)
+\State $(p_1, p_2)$ = ApplyMove($(p_1,p_2),m$)
+%\State $p_2$ = ApplyMove($p_2,m$)
  \State $\Delta p \gets p_2-p_1$
 \EndWhile
-\State Append($g2-p2$, moves)  \Comment translate to goal
+\State moves $\gets$ Append(moves, $g_2-p_2$)  \Comment translate to goal
 \State \Return moves
@@
 }
-\caption{\label{fig:polygon}{Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
+\caption{\label{fig:polygon}{Reachable sets are drawn with transparent blue polygons.
+A polygon with $n$ sides has $n$ reachable sets, but if $m$ is even and the polygon is regular, half the reachable sets overlap.
+%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
 As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
@@
 (left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
-(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
+(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area. 
 }
@@
 $P$ is a list of the vertices of a convex polygon. %RSet contains all the reachable polygons.
-\For {$i$ in $P$}
-\State $p_{i} \gets s_1 + s_2 - i$
-\State $p_{i+1} \gets s_1 + s_2 - (i+1)$
-\State $L \gets$ line with $(p_{11}, p_{12})$
+\State $R_{\textrm{SET}\gets \{\}}$
+\For {$p_i$ in $P$}
+\State $p_{i}' \gets s_1 + s_2 - p_i$
+\State $p_{i+1}' \gets s_1 + s_2 - p_{i+1}$
+\State $L \gets$ line with $(p_{i}', p_{i+1}')$
 \State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon $P$
-\If {$p_{i}\notin$ polygon}
-\State $p_{i} \gets l_i$
+\If {$p_{i}'$ not inside polygon $P$}
+\State $p_{i}' \gets l_i$
 \EndIf
-\State $v_i \gets$ All vertices $\in [ l_i , l_{i+1} ] $
-\State $d_i \gets$ All vectors from $p_i$ to $[ l_i , i]\cup  p_{i+1}$ to $ [i+1 , l_{i+1} ] $
-\State Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
+\If {$p_{i+1}'$ not inside polygon $P$}
+\State $p_{i+1}' \gets l_{i+1}$
+\EndIf
+%\State $D \gets$ All vertices $\in [ l_i , l_{i+1} ] $
+\State $D =([l_i, v_{\textrm{min}}, ..., p_i ] -p_i' $,
+
+$[p_{i+1} , o_{i+2}, ... , v_{\textrm{max}}, l_{i+1}] - p_{i+1}')$
+%\State $R_{\textrm{SET}} \gets$ Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
+\State Append polygon ($s_2 - (s_1 +d_i)$ to $R_{\textrm{SET}}$
 \EndFor
@@
 %The algorithm works by expanding the path with the shortest estimated length.
-Expanding a path means either moving directly to the goal, or pushing one robot to a wall and adjusting the relative position of the other robot.
+Expanding a path means either moving directly to the goal, or pushing one particle to a wall and adjusting the relative position of the other particle.
 As soon as the goal is reached, the algorithm returns this path.
@@
 To do so, we first check $\psi_{min}$ and $\psi_{max}$.
-\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the robots if the touching robot has the angle $\psi$.
+\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the particles if the touching particle has the angle $\psi$.
 
@@
 %\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two robots. 
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
 \State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
@@
 \State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
-\State Return Join($l_1, l_2, l_3, l_4$)
+\State Return $\Join$($l_1, l_2, l_3, l_4$)
 \end{algorithmic}
@@
  The two candidate shortest-length paths that touch the wall are shown with parallel arrows. 
-Each arrow will cause one of the particles to touch the wall, enabling the other robot to move freely in the  $z$ axis to achieve the required relative position.
+Each arrow will cause one of the particles to touch the wall, enabling the other particle to move freely in the  $z$ axis to achieve the required relative position.
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.

### 2018-04-24 journalWallFriction/simulation.tex
Edited algorithms parts.

@@
 The hardest point to achieve is the when the goals have $\pi$ difference and are very close to the boundary.
+Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace.
+Fig.~\ref{fig:contourPlots} top row considers the particles are interchangeable for three arbitrary starting and goal positions for the particles.
+Using interchangeable particles makes the path length significantly smaller.
+The middle and bottom row shows the same configurations with unique particles.
  %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
@@
 \renewcommand{\figwid}{1\columnwidth}
-{\begin{overpic}[width =\figwid]{identical.png}
-\end{overpic}
-\begin{overpic}[width =\figwid]{contourDistnew.png}
+{
+\begin{overpic}[width =\figwid]{contourDistnew.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
+\end{turn}}
+
 \end{overpic}
 \vspace{1em}
-\begin{overpic}[width =\figwid]{JustSimulationV6.png}
-%\put(97.5,21){5}
-%\put(97.5,17.25){4}
-%\put(97.5,13.5){3}
-%\put(97.5,24){6}
-%\put(97.5,27){7}
-%\put(97.5,10){2}
-%\put(97.5,7.5){1}
-%\put(97.5,5){0}
+\begin{overpic}[width =\figwid]{JustSimulationV6.png}\put(-2,10){\begin{turn}{90} \tiny{unique particles}
+\end{turn}}
+
+\end{overpic}
+\begin{overpic}[width =\figwid]{identical.png}\put(-2,6){\begin{turn}{90} \tiny{interchangeable particles}
+\end{turn}}
 \end{overpic}
@@
 \end{figure}
-
-   %The path given by our algorithm is shown with solid arrows.
-%Each row has five snapshots taken every quarter second. 
-%For the sake of brevity axis-aligned moves were replaced with oblique moves that combine 
-%jjkjk klklk klklklk klklk lklkl lklklklkl  klklklklk lklklklklk lklklkl klklklkl klklklk klklklk klklk klk klkl
-% THES IMAGES DON'T DO THIS ANYMORE 
-% $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot at $t = 0.25$. 
-% The following frames  adjust $\Delta r_y$ to $\Delta e_y$. 
-% $\Delta r_y$ is corrected by $t = 0.75$. 
-% Finally, the algorithm moves the robots to their corresponding destinations.
-
-%\subsection{Position Control of $n$ Robots}
-%Alg. \ref{alg:PosControlNRobots}  was simulated in {\sc Matlab} using square block robots with unity width. %Code is available at \cite{Arun2015}.
-% Simulation results are shown in Fig.~\ref{fig:4diagramsplots.pdf} for arrangements with an increasing number of robots,  $n$= [8, 46, 130, 390, 862]. 
-%The distance moved grows quadratically with the number of robots $n$. A best-fit line $210 n^2 + 1200n-10,000$ is overlaid by the data.
-%
-%In Fig.~\ref{fig:4diagramsplots.pdf}, the amount of clearance is $\epsilon=1$.
-%Control performance is sensitive to the desired clearance.
-As $\epsilon$ increases, the total distance decreases asymptotically, as shown in Fig.~\ref{fig:graphrobo.pdf}, because the robots have more room to maneuver and fewer $\operatorname{DriftMove}$s are required.
-%%We observed that the value of $\epsilon$, greatly affects the total number of moves of commanded moves on the robots.
-The smaller the value of ‘?’ the greater is the number of moves required to move a said distance.
-The second graph plot(fig roboplot) is that of ‘?’ vs ‘total commanded moves’ which is the number of moves the global controller commands onto all the robots.
-Here we take the pattern
-%%‘ROBO’ and calculate the ‘total commanded moves’ for ‘?’=[ 0.25,0.375,0.5,0.625,0.75,0.875,1].
-We can notice an exponential fall in the total command moves.
-It can be inferred that there is a tradeoff between area of bounded region and the number of moves.
-For smaller ‘?’, smaller is the required area to implement algorithm but the longer is the time taken to complete the algorithm.
-Fig robo
-%%Unit of ‘?’ is the ratio of minimum clearance value to the body length of a robot.
-%
-%
-%
-%
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=\columnwidth]{5diagramsplots2.pdf}
-%\end{center}
-%\vspace{-1em}
-%\caption{\label{fig:4diagramsplots.pdf}
-%The required number of moves under Alg.~\ref{alg:PosControlNRobots} using non-slip contact walls to rearrange $n$ square-shaped robots grows quadratically with $n$. % \cite{Arun2015}.
-% See code at \cite{Arun2015}.
-%}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{center}
-%	\includegraphics[width=\columnwidth]{graphrobo.pdf}
-%\end{center}
-%\vspace{-1em}
-%\caption{\label{fig:graphrobo.pdf}
-%Control performance is sensitive to the desired clearance $\epsilon$.  As $\epsilon$ increases, the total distance decreases asymptotically.
-%}\vspace{-1em}
-%\end{figure}

### 2018-04-24 journalWallFriction/theory2robAlgs.tex
Edited algorithms parts.

@@
 Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
-The algorithm can now handle any convex workspace, inducing the special limit case of a circular workspace.
+The algorithm can now handle any convex workspace, including the special limit case of a circular workspace.
 In the last subsection we present techniques to control 3D positioning of two particles.
@@
 The algorithm assigns a uniform control input at every instance.
-The goal is to move the particles to the goal positions using a shared control input.
-We do this by first moving them to the correct relative position and then translating the particles to the goal.
-The correct relative position is when
- $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||< \epsilon$, where $\epsilon$ is an arbitrary small number.
+The goal is to move the particles within $\epsilon$ of the goal positions using a shared control input where $\epsilon$ is an arbitrary small number.
+We do this by first moving them within $\epsilon$ of the correct relative position and then translating the particles to the goal.
+The relative position is $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
 
@@
 The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- We also show the starting and ending relative distance as $\Delta s$ and $\Delta g$ in $\Delta$ configuration space in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+ Fig.~\ref{fig:regionMove} shows the starting and ending relative distance as $\Delta s$ and $\Delta g$ in $\Delta$ configuration space% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
 We will discuss how to make reachable sets in the next subsections.
  The next subsections give procedures to compute the reachable set.
- The algorithm first computes the reachable set, and then checks if the goal relative position is in the reachable set.
-If it is not in the reachable set, the closest point on the reachable set from $\Delta g$ becomes the current goal.
-We move the particles to achieve the current goal.
-If current goal is the final goal, the correct \emph{relative} position has been achieved, and a final translation delivers both particles to their goal destinations.
-%we go to the final goal position.
-If it is not our final goal,
+ Alg.~\ref{alg:optimalAlg} first computes the reachable set.
+If the goal relative position is in the reachable set, we move particles to achieve relative position.
+If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
+ \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to push until one particle touches a wall, and the second to adjust the relative spacing.
+ Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
  Otherwise, we iterate until we reach the goal. 
 \begin{algorithm}[htb]
-\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P$)}\label{alg:optimalAlg}
+\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P,\epsilon$)}\label{alg:optimalAlg}
 \begin{algorithmic}[1]
@@
 %$(0,0)$ is bottom corner,
- $P$ is a description of the workspace.
+ $P$ is a description of the workspace. $\epsilon$ is a positive error bound.
 % \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
@@
 \While {$||\Delta p - \Delta g|| > \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_{\textrm{SET}}\gets$  Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or Alg.~\ref{alg:circularReachbale}
+\State $R_{\textrm{SET}}\gets$  Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or \ref{alg:circularReachbale}
 \State $ \Delta g_c\gets $nearest point in $R_{\textrm{SET}}$ to $\Delta g$
 \State $m \gets $move-to-wall corresponding to $\Delta g_c$
-\State moves $\gets$ Append(moves, $m$)
-\State $(p_1, p_2)$ = ApplyMove($(p_1,p_2),m$)
+\State moves $\gets$ Append $m$ to moves
+\State $(p_1, p_2)$ $\gets$ ApplyMove $m$ to $(p_1,p_2)$
 %\State $p_2$ = ApplyMove($p_2,m$)
@@
 \EndWhile
-\State moves $\gets$ Append(moves, $g_2-p_2$)  \Comment translate to goal
+\State moves $\gets$ Append $g_2-p_2$ to moves \Comment translate to goal
 \State \Return moves
@@
 If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
-We extend a square workspace to any convex workspace by computing the reachable set for all the sides of the polygon.
+ We compute the reachable set for any convex workspace.
+ Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{th}$ side.
+If one particle hist side $i$ before the other ($\overline{\rm s_1s_2 }\nparallel \overline{\rm p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in line 2-13 of Alg.~\ref{alg:polygonReachbale}.
+The union of these polygons for all $n$ sides is the reachable set of $\Delta$ configurations.
+ % by computing the reachable set for all the sides of the polygon.
 Fig.~\ref{fig:polygonAlg} shows different cases with the particles and an arbitrary side.
@@
 %\State $D \gets$ All vertices $\in [ l_i , l_{i+1} ] $
-\State $D =([l_i, v_{\textrm{min}}, ..., p_i ] -p_i' $,
+\State $D \gets$ $s_2 - s_1 -([l_i, v_{\textrm{min}}, ..., p_i ] -p_i' $,
 
@@
 %\State $R_{\textrm{SET}} \gets$ Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
-\State Append polygon ($s_2 - (s_1 +d_i)$ to $R_{\textrm{SET}}$
+\State $R_{\textrm{SET}}$ $\gets$ Append polygon $D$ to $R_{\textrm{SET}}$
 \EndFor
@@
 
-To compute reachable set for a circular workspace, first we consider all passible hitting point locations in the workspace.
- The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d}}{r})$, where $d= |s_1 - s_2|$ and $r$ is the radius of the circle.
+To compute the reachable set for a circular workspace, first we consider all possible first contact locations.
+ The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the circle.
  %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
@@
 %The algorithm works by expanding the path with the shortest estimated length.
-Expanding a path means either moving directly to the goal, or pushing one particle to a wall and adjusting the relative position of the other particle.
-As soon as the goal is reached, the algorithm returns this path.
+%Expanding a path means either moving directly to the goal, or pushing one particle to a wall and adjusting the relative position of the other particle.
+%As soon as the goal is reached, the algorithm returns this path.
 
-A circle has infinite number of sides, thus infinite reachable sets. However, parameterized by the first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where 
+A circle has an infinite number of sides, thus infinite reachable sets.
+However, the reachable set can be parameterized by the angle of first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where
 %We make the reachable sets by knowing the angle of the chord, $\psi$, shown in Fig.~\ref{fig:psigamma}. Minimum and maximum values of $\psi$ can be calculated in the following equation.
 \begin{align}
- \psi \in \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big].
+ \psi \in [\psi_{\min}, \psi_{\max}]= \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big].
 % \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
 \end{align}
-The possible first contact locations line on an arc with interior angle $\gamma$ is calculated by:
+The possible first contact locations are on an arc with interior angle $\gamma$ parameterized by $\psi$ is calculated by:
 \begin{align}\label{eq:ppsi}
-p_\psi &= r[\cos(\psi ), \sin(\psi )]\\ \label{eq:dprep}
-d_\perp&= 2 \abs{(s_1.p_\psi - s_2.p_\psi)}\\ \label{eq:gamma}
-\gamma &= \cos^{-1} \Big(1-\frac{d_\perp}{r} \Big) 
+p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )]\\ \label{eq:dprep}
+d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||\\ \label{eq:gamma}
+\gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big) 
 \end{align} 
-Reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose to immobilize the particle closest to a wall.
+Reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to immobilize the particle closest to a wall.
+The reachable set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
 
-The equations for the four lines outlining the reachable set are as follows:
+The equations for the four lines outlining the union of reachable sets are as follows:
 \begin{align}\label{eq:circlereachable}
-l_1 &=  r \Big((\cos\psi_{\min}- \cos(\gamma_c + \psi_{\min}) )\\ \nonumber
-& + (\sin\psi_{\min}- \sin(\gamma_c + \psi_{\min}))\Big) &  0<\gamma_c< \gamma_{\psi_{\min}},\\ \nonumber
-l_2 &=  r \Big((\cos\psi_{\max}- \cos(\gamma_c + \psi_{\max}))\\ \nonumber
-& + (\sin\psi_{\max}- \sin(\gamma_c + \psi_{\max}))\Big) &  \gamma_{\psi_{\max}}<\gamma_c< 0,\\  \nonumber
-l_3 &=  r \Big((\cos\psi- \cos( \psi+\gamma_c ) )\\ \nonumber
-&+ ( \sin\psi-\sin( \psi+ \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
-l_4 &=  r \Big((\cos\psi- \cos( \psi-\gamma_c ))\\ \nonumber
-&+  ( \sin\psi- \sin( \psi- \gamma_c))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+l_1 =  r \Big(&(\cos\psi_{\min}- \cos(\gamma + \psi_{\min}) )\\ \nonumber
+ + &(\sin\psi_{\min}- \sin(\gamma + \psi_{\min}))\Big) &  0<\gamma< \gamma(\psi_{\min}),\\ \nonumber
+l_2 =  r \Big(&(\cos\psi_{\max}- \cos(\gamma + \psi_{\max}))\\ \nonumber
+ + &(\sin\psi_{\max}- \sin(\gamma + \psi_{\max}))\Big) &  \gamma(\psi_{\max})<\gamma< 0,\\  \nonumber
+l_3 =  r \Big(&(\cos\psi- \cos( \psi+\gamma(\psi) ) )\\ \nonumber
++ &( \sin\psi-\sin( \psi+ \gamma(\psi)))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+l_4 =  r \Big(&(\cos\psi- \cos( \psi-\gamma(\psi) ))\\ \nonumber
++ & ( \sin\psi- \sin( \psi- \gamma(\psi)))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
 \end{align}
-We combine these boundaries to make the reachable set. 
-We need to find a $\psi$ that would enable us to reach to the required relative goal distance.
+We combine these boundaries to compute the reachable set. 
+Next, find a $\psi$ that would enable us to reach to the required relative goal distance.
 To do so, we first check $\psi_{min}$ and $\psi_{max}$.
-\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the particles if the touching particle has the angle $\psi$.
+Eq.~\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the particles if the touching particle has the angle $\psi$.
 
@@
 
-Here $c$ is the coordinate of the workspace center and $r$ is the radius of the workspace.
+%Here $c$ is the coordinate of the workspace center and $r$ is the radius of the workspace.
  If $\Delta g$ is not in the region made by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g$ and the current relative position, $\Delta s$.
@@
 \Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
-\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
+%\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
 \State Calculate $p_{\psi}$ \Comment use \eqref{eq:ppsi}
@@
 \State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
-\State Return $\Join$($l_1, l_2, l_3, l_4$)
+\State Return $\cup$($l_1, l_2, l_3, l_4$)
 \end{algorithmic}

### 2018-04-24 journalWallFriction/theory2robAlgs.tex
a lot of figures are corrected

@@
  Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{th}$ side.
-If one particle hist side $i$ before the other ($\overline{\rm s_1s_2 }\nparallel \overline{\rm p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in line 2-13 of Alg.~\ref{alg:polygonReachbale}.
+If one particle hist side $i$ before the other ($\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in line 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the reachable set of $\Delta$ configurations.
@@
 \renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){Case 1:$s_1$ does not hit boundary}\put(28,39){$s_2$}\put(63,50){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){\tiny{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){Case 2:$s_1$ hits boundary}\put(20,40){$s_2$}\put(55,58){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){\tiny{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){Generates polygon}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){\tiny{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
 \end{overpic}

### 2018-04-25 journalWallFriction/Conclusion.tex
Add files via upload

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Future Work}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This paper presented techniques for controlling the positions of two particles using uniform inputs and boundary friction forces.  
The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D delta configuration space.
Extending algorithms \ref{alg:optimalAlg}  to 3D is straightforward, but increases the complexity.
 Additionally, this paper assumed friction was sufficient to completely stop particles in contact with the boundary. 
  The algorithms would require retooling to handle small friction coefficients. Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Appendix}\label{sec:Appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 % With regard to technological advances, this includes designing controllers that efficiently regulate $\sigma_{xy}$, perhaps using Lyapunov-inspired controllers as in \cite{kim2015imparting}.

 % It may be possible to rank controllability as a function of friction.
%  In hardware, the wall friction can be varied by laser-cutting boundary walls with different of profiles.

%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles.

%    Inspired by large-scale human experiments with swarms of robots under global control,  this paper investigated controllers that use only the mean and variance of a robot swarm. We proved that the mean position is controllable, and provided conditions under which variance is controllable.  We derived automatic controllers for each and a hysteresis-based switching control that controls the mean and variance of a robot swarm.  We employed these controllers as primitives for a block-pushing task. 
%    
%    Future work should implement these controllers on a robot swarm and decrease completion time by avoiding counter-productive contact with the block while the swarm is lowering its variance.  We have also assumed the swarm is unimodal and has a straight-line path to the moveable block. Relaxing these assumptions requires solving the \emph{gathering problem}.  The gathering problem for a swarm with uniform inputs is largely unexplored, and must be examined probabilistically for nontrivial environments.
%    
    % We should also control the covariance $\sigma_xy$ and higher moments of the distribution

%Sensing is expensive, especially on the nanoscale. To see nanocars~\citep{Chiang2011}, scientists fasten molecules that fluoresce light when activated by a strong light source. Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. Photobleaching can be minimized by lowering the excitation light intensity, but this increases the probability of missed detections~\citep{Cazes2001}. A control methodology based on statistics of the robot swarm rather than the actual position of each robot, allows relaxing demands on imagine systems, controllers robust to tracking errors, and a simpler methodology.  In this work we...
%

% Additionally, as population characteristics, they are available even if only a percentage of the robots are detected each control cycle.
%Photobleaching: http://www.piercenet.com/browse.cfm?fldID=4DD9D52E-5056-8A76-4E6E-E217FAD0D86B
%
%Photobleaching is caused by the irreversible destruction of fluorophores due to either the prolonged exposure to the excitation source or exposure to high-intensity excitation light. Photobleaching can be minimized or avoided by exposing the fluor(s) to the lowest possible level of excitation light intensity for the shortest length of time that still yields the best signal detection; this requires optimization of the detection method using high sensitivity CCD cameras, high numerical aperture objective and/or the widest bandpass emission filter(s) available. Other approaches include using fluorophores that are more photostable than traditional fluorophores and/or using antifade reagents to protect the fluor(s) against photobleaching. Steps to avoid photobleaching are not feasible for all detection methods and should be optimized for each method used. For example, antifade reagents are toxic to live cells, and therefore they can only be used with fixed cells or tissue. Furthermore, some detection methods, such as flow cytometry, normally do not require steps to avoid photobleaching because of the extremely short exposure time of the fluorophore to the excitation source.

### 2018-04-25 journalWallFriction/Experiment.tex
Add files via upload

--- 2018-04-20 journalWallFriction/experiment.tex
@@
 To demonstrate Alg.~\ref{alg:optimalAlg} experimentally, we performed several tests.
-Each used the same magnetic setup.
- Two different intestine models were employed, the first a 3D-printed cross-section representation of a small intestine, and the second bovine small intestine.
+Each used the same magnetic setup shown in Fig.~\ref{fig:}.
+% The figure might need to be referenced 
+ Two different intestine models were employed, the first a 3D-printed cross-section representation of a small intestine, and the second a cross-section of a bovine stomach.
 
@@
 
- The magnetic manipulation system consists of two pairs of electromagnetic coils that consist of iron cores at their centers, and arranged orthogonal to each other.
-The iron core at the center of each coil served to concentrate the magnetic field towards the workspace.
-To set up the magnetic manipulation system for this project, an Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a Basler ace black and white camera was attached to the top of the system focusing on the back-lit workspace.
+ The magnetic manipulation system has two pairs of electromagnetic coils each with iron cores at their centers, and arranged orthogonal to each other.
+The iron core at the center of each coil concentrates the magnetic field towards the workspace.
+An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
+Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15w LED light strip.
 
-To obtain experimental data, the test samples which comprised of the Phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in fluid.
-Over the course of experimentation, corn syrup was used to have the best viscosity for the experiments.
+To obtain experimental data, the test samples which comprised of the phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in corn syrup.
+Corn syrup was used to have the best viscosity(12000 cP) for the experiments.
 The velocities of particles immersed in it were dampened enough to control their movements with ease.
-Spherical $0.5 mm$ magnets were used as our particles.
+Spherical $0.5 mm$ magnets(supermagnetman #SP0100-50) were used as our particles.
 
@@
 
-The Intestine Phantom Model was used as the first test field of the project and was made to mimic an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and a thickness of 2mm created using a 3D printer and Fused filament fabrication.
-The model had some $2 mm$ long protrusions on its inner surface to mimic the effects of intestinal villi on the target particles.
+The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi.
+The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $2 mm$ long protrusions on its inner surface cut out of $6 mm$ thick acrylic to model the geometry of intestinal villi.
 Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.
@@
 
-\subsection{Bovine Intestine Cross-section}
+\subsection{Bovine Stomach Cross-section}
 %todo: procedure for fixating the Bovine tissue sample,
@@
 
-This phase of the project involved the use of beef intestines.
-Strips of intestine about 5mm thick were cut and placed in Neutrally buffered formalin for 24 hours for fixation.
+Strips of cow tripe approximately 5mm thick were cut and placed in Neutrally buffered formalin for 24 hours for fixation.
 After fixation, each sample was transferred to 70\% ethanol for storage.

### 2018-04-25 journalWallFriction/experiment.tex
Added tons of new images!

@@
 \begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
+\\
+\vspace{1em}
+\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
 \caption{\label{fig:story}

### 2018-05-01 journalWallFriction/Experiment.tex
Added a lot of new pictures, edited a little bit.

@@
 An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15w LED light strip.
+Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15$w$ LED light strip.
 
@@
 The velocities of particles immersed in it were dampened enough to control their movements with ease.
-Spherical $0.5 mm$ magnets(supermagnetman #SP0100-50) were used as our particles.
+Spherical $\SI{0.5}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
 
@@
 The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $2 mm$ long protrusions on its inner surface cut out of $6 mm$ thick acrylic to model the geometry of intestinal villi.
+The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
 Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.

### 2018-05-01 journalWallFriction/SwarmShapeControl.tex
Edited a lot of parts of the paper
Edited algorithms parts.
Added a lot of new pictures, edited a little bit.

@@
 \usepackage{authblk}
-
+\usepackage[per-mode=symbol, detect-weight=true, binary-units=true]{siunitx}
+\usepackage{rotating}
 %\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
@@
 Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
- This paper investigates control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
+ This paper investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
 Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
@@
 In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
-Because in vivo environments are usually not square, this work extends the previous work to convex workspaces including circles and 3D positioning. 
-This paper also implements the algorithms using a hardware setup inspired by gastrointestinal tract.
+Because in vivo environments are usually not square, this paper extends the previous work to convex workspaces including circles and 3D positioning. 
+This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.
 

### 2018-05-01 journalWallFriction/intro.tex
Edited a lot of parts of the paper
Added a lot of new pictures, edited a little bit.

@@
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
-Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the particle swarm.
+Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the $n$-particle swarm.
  Techniques are needed that can handle this underactuation. 
- In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.
+ In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
 
@@
   Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
-We provide a shortest-path algorithm to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}.
+We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
 Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 

### 2018-05-01 journalWallFriction/theory.tex
Edited algorithms parts.
Added a lot of new pictures, edited a little bit.

@@
 Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g.
-PRM and RRT* are not suited \cite{lavalle2006planning,kavraki1996probabilistic}.
+PRM and RRT* are not suitable \cite{lavalle2006planning,kavraki1996probabilistic}.
 Instead, this paper employs a graph search.
@@
 
-In the absence of obstacles uniform inputs move a swarm identically.  
+In the absence of obstacles, uniform inputs move a swarm identically.  
 Independent control requires breaking this symmetry. 
@@
 Similar mechanical systems exist at larger scales, e.g.
-all tumblers of a combination look move uniformly unless obstructed by an obstacle.
+all tumblers of a combination lock move uniformly unless obstructed by an obstacle.
  Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
@@
  \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
- The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}.
+ The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}.
 The problem can be simplified by choosing the coordinate system carefully.
@@
 The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.
-See \cite{shortestPathMathematica}.}
+See \cite{BeckerShortestPath}.}
 %\vspace{-1em}
@@
 This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
-For an interactive Mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.
+For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}.

### 2018-05-01 journalWallFriction/theory2robAlgs.tex
slight update to the terminology
Added a lot of new pictures, edited a little bit.

@@
 Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
-The algorithm can now handle any convex workspace, including the special limit case of a circular workspace.
+Alg.~\ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
 In the last subsection we present techniques to control 3D positioning of two particles.
@@
  Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
-The algorithm assigns a uniform control input at every instance.
+Alg.~\ref{alg:optimalAlg} assigns a uniform control input at every instance.
 The goal is to move the particles within $\epsilon$ of the goal positions using a shared control input where $\epsilon$ is an arbitrary small number.
@@
 The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- Fig.~\ref{fig:regionMove} shows the starting and ending relative distance as $\Delta s$ and $\Delta g$ in $\Delta$ configuration space% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+ Fig.~\ref{fig:regionMove} shows the starting and ending relative distance as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
 We will discuss how to make reachable sets in the next subsections.
  The next subsections give procedures to compute the reachable set.
+
  Alg.~\ref{alg:optimalAlg} first computes the reachable set.
@@
 \caption{\label{fig:polygon}{Reachable sets are drawn with transparent blue polygons.
-A polygon with $n$ sides has $n$ reachable sets, but if $m$ is even and the polygon is regular, half the reachable sets overlap.
+A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
 %Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
@@
 If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
- We compute the reachable set for any convex workspace.
+Alg.~\ref{alg:polygonReachbale} computes the reachable set for any convex workspace.
  Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{th}$ side.
-If one particle hist side $i$ before the other ($\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in line 2-13 of Alg.~\ref{alg:polygonReachbale}.
+If one particle hits side $i$ before the other ($\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the reachable set of $\Delta$ configurations.
@@
 \State $p_{i+1}' \gets s_1 + s_2 - p_{i+1}$
-\State $L \gets$ line with $(p_{i}', p_{i+1}')$
+\State $L \gets \overline{ p_i p_{i+1}}$ \Comment{line $(p_{i}', p_{i+1}')$}
 \State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon $P$
@@
 
-$[p_{i+1} , o_{i+2}, ... , v_{\textrm{max}}, l_{i+1}] - p_{i+1}')$
+$[p_{i+1} , p_{i+2}, ... , v_{\textrm{max}}, l_{i+1}] - p_{i+1}')$
 %\State $R_{\textrm{SET}} \gets$ Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
@@
 \begin{figure*}
-\centering
-\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\end{overpic}\\
-\vspace{1em}
-\centering
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}\\
+
 \begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\end{overpic}
@@
 \begin{align}
- \psi \in [\psi_{\min}, \psi_{\max}]= \Big[\theta +\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2}, \theta -\frac{\sin^{-1}{d}}{2r} + \frac{\pi}{2}\Big].
+ \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d}}{2r} \Big].
 % \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
 \end{align}
-The possible first contact locations are on an arc with interior angle $\gamma$ parameterized by $\psi$ is calculated by:
+The possible first contact locations are on an arc with interior angle $\gamma$ parameterized by $\psi$:
 \begin{align}\label{eq:ppsi}
-p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )]\\ \label{eq:dprep}
-d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||\\ \label{eq:gamma}
-\gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big) 
+p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )],\\ \label{eq:dprep}
+d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||,\\ \label{eq:gamma}
+\gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big). 
 \end{align} 
@@
 The reachable set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
+One such chord is drawn in red in Fig.~\ref{fig:regionMove}.
 
@@
 We combine these boundaries to compute the reachable set. 
-Next, find a $\psi$ that would enable us to reach to the required relative goal distance.
-To do so, we first check $\psi_{min}$ and $\psi_{max}$.
-Eq.~\eqref{eq:ifinchord} shows if any point $p$, is in the region made by the particles if the touching particle has the angle $\psi$.
+Next, find a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the reachable set to $\Delta g$.
+Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
 
 \begin{align}\label{eq:ifinchord}
-&(p.x - c.x)^2 + (p.y - c.y)^2  > r^2\\ \nonumber
- &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r\cdot \cos\gamma
+&(p.x - c.x)^2 + (p.y - c.y)^2  > r^2,\\ \nonumber
+ &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r\cdot \cos\gamma.
 \end{align}
@@
 %Here $c$ is the coordinate of the workspace center and $r$ is the radius of the workspace.
- If $\Delta g$ is not in the region made by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g$ and the current relative position, $\Delta s$.
-This line is a chord of the circle and we find the $\psi$ that makes this region.
-This equation finds this $\psi$:
+We first check if $\Delta g_c$ is achievable with $\psi_{min}$ and $\psi_{max}$.
+ If $\Delta g_c$ is not achievable by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g_c$ and the current relative position, $\Delta s$.
+This line is a chord of the circle and we find the $\psi$ correspondent to this chord.
+The following equation finds this $\psi$:
   \begin{equation}
@@
  \end{equation}
+
+Having $\psi$ enable the particles to achieve $\Delta g_c$ in two moves. One move will cause one robot to touch the wall, the other move will make the required relative position.
 
@@
 \State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
-\State Return $\cup$($l_1, l_2, l_3, l_4$)
+\State Return the union of ($l_1, l_2, l_3, l_4$)
 \end{algorithmic}

### 2018-05-01 journalWallFriction/Experiment.tex
edited the paper!

@@
 An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15$w$ LED light strip.
+Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a $\SI{15}{\watt}$ LED light strip.
 
@@
 The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
-Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.
+Fig.~\ref{fig:story} shows an experiment with starting and ending positions drawn with marker on the workspace.
 
@@
 
-Strips of cow tripe approximately 5mm thick were cut and placed in Neutrally buffered formalin for 24 hours for fixation.
+Strips of cow stomach approximately 5mm thick were cut and placed in neutrally buffered formalin for 24 hours for fixation.
 After fixation, each sample was transferred to 70\% ethanol for storage.
 For the experiments, a slice of fixed intestine was attached to the acrylic disc with cyanoacrylate (superglue) and then submerged in corn syrup.
-A drawback of fixing the tissue samples before experimentation is that they tended to shrivel and dry up a few minutes after being removed from the 70\% ethanol.
+A drawback of fixing the tissue samples before experiments is that they tended to shrivel and dry up a few minutes after being removed from the 70\% ethanol.
 %In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.
@@
 \caption{\label{fig:story}
-Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+Frames showing particle positions before and after control inputs. Top row: small intestine phantom. Bottom row: cow's stomach tissue.
 } \vspace{-1em}

### 2018-05-01 journalWallFriction/theory2robAlgs.tex
edited the paper!

@@
 To compute the reachable set for a circular workspace, first we consider all possible first contact locations.
- The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the circle.
+ The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.
  %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
@@
 The  set of points where the blue particle is the first to contact the boundary are drawn with a blue arc.
-Left top: The possible points for the blue and red particles to touch the boundary is shown in blue and red arcs.
-Left bottom: When the blue particle is touching a wall (blue square) the other particle (red square) can go anywhere in the reachable set (blue region).
-Right: The $\Delta$ configuration space for the corresponded starting positions of the particles is shown.}
+Left top: The possible first contact points for the blue and red particles are shown with blue and red arcs.
+Left bottom: Once the blue particle touches the wall (blue square) the other particle (red square) can go anywhere in the reachable set (red region).
+Right: The $\Delta$ configuration space for the corresponded starting positions of the particles is shown.
+The possible reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
+Once the blue particle contacts the boundary as shown, the reachable $\Delta$ configuration is the red set.}
 \end{figure}
@@
 The possible first contact locations are on an arc with interior angle $\gamma$ parameterized by $\psi$:
-\begin{align}\label{eq:ppsi}
-p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )],\\ \label{eq:dprep}
-d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||,\\ \label{eq:gamma}
-\gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big). 
+\begin{align}\label{eq:gamma}
+\gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big), \textrm{ where:}\\ \label{eq:dprep}
+d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||,\\ \label{eq:ppsi}
+p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )].
 \end{align} 
 Reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to immobilize the particle closest to a wall.
-The reachable set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
+The reachable $\Delta$ configuration set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
 One such chord is drawn in red in Fig.~\ref{fig:regionMove}.
 
-The equations for the four lines outlining the union of reachable sets are as follows:
+The equations for the four lines outlining the union of reachable $\Delta$ configuration sets are as follows:
 \begin{align}\label{eq:circlereachable}
@@
 \end{align}
-We combine these boundaries to compute the reachable set. 
+We combine these boundaries to compute the reachable set summarized in Alg.~\ref{alg:circularReachbale}.
 Next, find a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the reachable set to $\Delta g$.
-Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
+%Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
+We first check if $\Delta g_c$ is in the $\Delta$ configuration space chords defined by either $\psi_{\min}$ or $\psi_{\max}$ using \ref{eq:ifinchord}.
 
 \begin{align}\label{eq:ifinchord}
-&(p.x - c.x)^2 + (p.y - c.y)^2  > r^2,\\ \nonumber
- &(p_x - c_x) \cos\psi + (p.y- c.y) \sin\psi \leq -r\cdot \cos\gamma.
+&(\Delta g_c.x - c.x)^2 + (\Delta g_c.y - c.y)^2  > r^2 \wedge\\ \nonumber
+ &( c.x-\Delta g_c.x ) \cos\psi + ( c.y-\Delta g_c.y) \sin\psi > r\cdot \cos\gamma.
 \end{align}
@@
 %Here $c$ is the coordinate of the workspace center and $r$ is the radius of the workspace.
-We first check if $\Delta g_c$ is achievable with $\psi_{min}$ and $\psi_{max}$.
- If $\Delta g_c$ is not achievable by $\psi_{\min}$ or $\psi_{\max}$, we draw a line from $\Delta g_c$ and the current relative position, $\Delta s$.
-This line is a chord of the circle and we find the $\psi$ correspondent to this chord.
-The following equation finds this $\psi$:
+%We first check if $\Delta g_c$ is achievable with $\psi_{min}$ and $\psi_{max}$.
+ If $\Delta g_c$ is not in either chord, we draw a line from $\Delta g_c$ to the current relative position, $\Delta p$.
+This line is a chord of the circle centered at $c$.
+The $\psi$ to this chord obeys:
   \begin{equation}
- \psi = \tan^{-1}(\Delta p - \Delta g ).
+ \psi = \tan^{-1}(\Delta p - \Delta g_c ).
  \end{equation}
 
-Having $\psi$ enable the particles to achieve $\Delta g_c$ in two moves. One move will cause one robot to touch the wall, the other move will make the required relative position.
+The particles achieve $\Delta g_c$ in two moves. The first move causes one particle to touch the wall, the second move achieves the required relative position.
 
@@
 \caption{\label{fig:zaxis}
-Extending the algorithm to position the particles in 3D.
+Illustration on how boundary contacts enable 3D positioning.
+Once one particle contacts a boundary, the other particle's reachable set is a prism formed by extending the 2D reachable set in the $z$-direction.
 } %\vspace{-1em}

### 2018-05-02 journalWallFriction/experiment.tex
A lot of the pictures have been changed!

@@
 An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15w LED light strip.
+Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a $\SI{15}{\watt}$ LED light strip.
 
@@
 The velocities of particles immersed in it were dampened enough to control their movements with ease.
-Spherical $0.5 mm$ magnets(supermagnetman #SP0100-50) were used as our particles.
+Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
 
@@
 The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $2 mm$ long protrusions on its inner surface cut out of $6 mm$ thick acrylic to model the geometry of intestinal villi.
-Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.
+The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
+Fig.~\ref{fig:story} shows an experiment with starting and ending positions drawn with marker on the workspace.
 

### 2018-05-02 journalWallFriction/intro.tex
A lot of the pictures have been changed!

@@
 %\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
-\begin{overpic}[width=0.4\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
-\begin{overpic}[width=0.5\columnwidth]{magneticsetup.pdf}\put(28,-8){magnetic setup}\end{overpic}
+\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
+\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(28,-8){magnetic setup}\end{overpic}
 \vspace{1em}

### 2018-05-02 journalWallFriction/theory2robAlgs.tex
A lot of the pictures have been changed!

@@
 \begin{figure*}
-\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(7,15){\tiny{\textbf{workspace}}}\end{overpic}\\
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(7,15){\small{workspace}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(7,15){\small{workspace}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(7,15){\small{workspace}}\end{overpic}\\
 
@@
 \centering
-\begin{overpic}[width=0.5\columnwidth]{zaxis.pdf}\end{overpic}
+\begin{overpic}[width=0.95\columnwidth]{zaxis.pdf}\end{overpic}
 \caption{\label{fig:zaxis}

### 2018-05-08 journalWallFriction/Experiment.tex
A bunch of figures and texts added!

@@
 To demonstrate Alg.~\ref{alg:optimalAlg} experimentally, we performed several tests.
-Each used the same magnetic setup shown in Fig.~\ref{fig:}.
+Each used the same magnetic setup shown in Fig.~\ref{fig:IntroPic}.
 % The figure might need to be referenced 
@@
 The velocities of particles immersed in it were dampened enough to control their movements with ease.
-Spherical $\SI{0.5}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
+Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
 

### 2018-05-08 journalWallFriction/simulation.tex
A bunch of figures and texts added!

@@
 \caption{\label{fig:contour}
-Plots showing the algorithm with one goal on the boundary.
+Plots show performance with one goal on the boundary.
 }
@@
 \caption{\label{fig:deltanumdist}
-Plots showing decreasing error as the number of moves grows.
+Particle swap poles as their goal is the worst case. It can never be achieved but can be asymptotically reached. Plots show decreasing error as the number of moves grows.
 } 

### 2018-05-08 journalWallFriction/theory.tex
A bunch of figures and texts added!

@@
 
- If the goal configuration can be reached in three moves, the shortest path has a simple solution.
-The first move makes one particle hit a wall, the second move adjusts the relative spacing error  to zero, and  the third move takes the particles to their final positions.
-The second move cannot be shortened, so optimization depends on choosing the location where the particle hits the wall. 
+ If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
+The first move, $m_1$, makes one particle hit a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
+$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle hits the wall. 
  Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
@@
 \caption{\label{fig:reflection}
-The shortest path between two points in a square. Reflecting the goal position across the boundary wall and plotting a straight line gives the shortest path.
+The shortest ?three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle.
 } \vspace{-1em}

### 2018-05-08 journalWallFriction/theory2robAlgs.tex
A bunch of figures and texts added!

@@
 Alg.~\ref{alg:polygonReachbale} computes the reachable set for any convex workspace.
- Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{th}$ side.
+ Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{\textrm{th}}$ side.
 If one particle hits side $i$ before the other ($\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
@@
 \gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big), \textrm{ where:}\\ \label{eq:dprep}
-d_\perp(\psi)&= 2 ||(s_1.p_\psi(\psi) - s_2.p_\psi(\psi))||,\\ \label{eq:ppsi}
+d_\perp(\psi)&= 2 ||s_1.p_\psi(\psi) - s_2.p_\psi(\psi)||,\\ \label{eq:ppsi}
 p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )].
@@
 %Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
-We first check if $\Delta g_c$ is in the $\Delta$ configuration space chords defined by either $\psi_{\min}$ or $\psi_{\max}$ using \ref{eq:ifinchord}.
+We first check if $\Delta g_c$ is in the $\Delta$ configuration space chords defined by either $\psi_{\min}$ or $\psi_{\max}$ using:
 
 \begin{align}\label{eq:ifinchord}
-&(\Delta g_c.x - c.x)^2 + (\Delta g_c.y - c.y)^2  > r^2 \wedge\\ \nonumber
- &( c.x-\Delta g_c.x ) \cos\psi + ( c.y-\Delta g_c.y) \sin\psi > r\cdot \cos\gamma.
+&(\Delta g_c.x - c.x)^2 + (\Delta g_c.y - c.y)^2  > r^2 \textrm{     \textbf{and}}\\ \nonumber
+ &( c.x-\Delta g_c.x ) \cos\psi + ( c.y-\Delta g_c.y) \sin\psi > r\cos\gamma.
 \end{align}
@@
   \begin{equation}
- \psi = \tan^{-1}(\Delta p - \Delta g_c ).
+ \psi = \tan^{-1}\Big(\frac{\Delta p.x - \Delta g_c.x}{\Delta p.y - \Delta g_c.y} \Big).
  \end{equation}
 
-The particles achieve $\Delta g_c$ in two moves. The first move causes one particle to touch the wall, the second move achieves the required relative position.
+The particles achieve $\Delta g_c$ in two moves. The first move causes one particle to touch the wall at $p_\psi$, \eqref{eq:ppsi}. The second move achieves the required relative position.
 

### 2018-05-08 journalWallFriction/theory.tex
A bunch of plots and figures changed.
Added New pictures to The paper

@@
 \centering
-\begin{overpic}[width=0.5\columnwidth]{Reflection.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
+\vspace{-2em}
 \caption{\label{fig:reflection}
-The shortest ?three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle.
+The shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle.
 } \vspace{-1em}

### 2018-05-14 journalWallFriction/Conclusion.tex
edited the paper!
A lot of edits to the pictures and other stuff

@@
 
-This paper presented techniques for controlling the positions of two particles using uniform inputs and boundary friction forces.  
-The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D delta configuration space.
-Extending algorithms \ref{alg:optimalAlg}  to 3D is straightforward, but increases the complexity.
- Additionally, this paper assumed friction was sufficient to completely stop particles in contact with the boundary. 
+This paper presented techniques for controlling the positions of two particles using uniform inputs and non-slip boundary contact.  
+The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D $\Delta$ configuration space.
+Extending Alg. \ref{alg:optimalAlg}  to 3D was straightforward, but increased the complexity.
+Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue.
+
+There are several avenues for future work. This paper assumed friction was sufficient to completely stop particles in contact with the boundary. 
   The algorithms would require retooling to handle small friction coefficients.
-Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue
+The techniques in \cite{shahrokhi2017algorithms} and  \cite{AaronManipulation2013} could be applied to extend the analysis to more than two particles.
 

### 2018-05-14 journalWallFriction/SwarmShapeControl.tex
A lot of edits to the pictures and other stuff

@@
 However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
-In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
+In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
 Because in vivo environments are usually not square, this paper extends the previous work to convex workspaces including circles and 3D positioning. 

### 2018-05-14 journalWallFriction/experiment.tex
A lot of edits to the pictures and other stuff

@@
 
- The magnetic manipulation system has two pairs of electromagnetic coils each with iron cores at their centers, and arranged orthogonal to each other.
+ The magnetic manipulation system has two pairs of electromagnetic coils, each with iron cores at their centers, and arranged orthogonal to each other.
 The iron core at the center of each coil concentrates the magnetic field towards the workspace.
 An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a $\SI{15}{\watt}$ LED light strip.
+Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system, focusing on the workspace which was backlit by a $\SI{15}{\watt}$ LED light strip.
 
-To obtain experimental data, the test samples which comprised of the phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in corn syrup.
-Corn syrup was used to have the best viscosity(12000 cP) for the experiments.
-The velocities of particles immersed in it were dampened enough to control their movements with ease.
+To obtain experimental data, the test samples (the phantom intestine model and the bovine cross section) were placed in laser-cut acrylic discs and then immersed in corn syrup.
+Corn syrup was used to increase the viscosity to 12000 cP for the experiments.
 Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
+The magnetic field in this setup is only approximately uniform.
+The magnetic force varies in both magnitude and orientation.
+As shown in the video attachment, the particle closer to the coil moves faster than the other particle.
+Also, magnetic forces are not exactly parallel, but point toward the activated coil.
+Algorithm~\ref{alg:optimalAlg} is robust to these non uniformities, but sometimes requires additional iterations.
 
 %explain the magnetic field strength used, the camera system (briefly)
-
-todo: image of the magnetic setup with scale bar
 %explain the particles used
@@
 
-The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi.
+The intestine phantom model was used first and was made to mimic the geometry of an intestine and its villi.
 The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
-Fig.~\ref{fig:story} shows an experiment with starting and ending positions drawn with marker on the workspace.
+Fig.~\ref{fig:story} top row shows an experiment.
+Starting and ending positions were printed beneath the workspace on transparency film.
 
@@
 
-Strips of cow stomach approximately 5mm thick were cut and placed in neutrally buffered formalin for 24 hours for fixation.
-After fixation, each sample was transferred to 70\% ethanol for storage.
-For the experiments, a slice of fixed intestine was attached to the acrylic disc with cyanoacrylate (superglue) and then submerged in corn syrup.
-A drawback of fixing the tissue samples before experiments is that they tended to shrivel and dry up a few minutes after being removed from the 70\% ethanol.
+Strips of cow stomach approximately $\SI{5}{\milli\metre}$ thick were cut and sewn to acrylic cylinder and then glued to an acrylic substrate using cyanoacrylate (superglue).
+This assemble was then filled with corn syrup.
+The experiment is shown in Fig.~\ref{fig:story} bottom row.
 %In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.

### 2018-05-14 journalWallFriction/intro.tex
A lot of edits to the pictures and other stuff

@@
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
-Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the $n$-particle swarm.
+Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
  Techniques are needed that can handle this underactuation. 
@@
 We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
-Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
+Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
  We end with directions for future research in Sec.  \ref{sec:conclusion}.

### 2018-05-14 journalWallFriction/simulation.tex
A lot of edits to the pictures and other stuff

@@
 
- The contour plots in Fig.~\ref{fig:contour} left shows the length of the path for given $s_1,s_2,g_1$ with $g_2$ ranging over all the workspace.
+ The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
+Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} and middle and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
 Fig.~\ref{fig:contour} right shows the total distance of the path.
-This plot clearly shows the nonlinear nature of the path planning.
-The hardest point to achieve is the when the goals have $\pi$ difference and are very close to the boundary.
+This plot shows the nonlinear nature of the path planning.
+The path length grows when the goals have $\pi$ difference and are very close to the boundary.
+ The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
+This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
  Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace.
-Fig.~\ref{fig:contourPlots} top row considers the particles are interchangeable for three arbitrary starting and goal positions for the particles.
-Using interchangeable particles makes the path length significantly smaller.
-The middle and bottom row shows the same configurations with unique particles.
+Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
+All of the discussed plots have considered the particles to be unique.
+If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller.
+The bottom row of Fig.~\ref{fig:contour} and Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
+
  %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
 
- The plots in Fig.~\ref{fig:deltanumdist} show the exponentially increasing number of moves and distance when the accuracy of reaching to the goal ($\delta$) is getting to zero when the goal positions have $\pi$ difference with each on the boundaries.
+% The plots in Fig.~\ref{fig:deltanumdist} show the exponentially increasing number of moves and distance when the accuracy of reaching to the goal ($\delta$) is getting to zero when the goal positions have $\pi$ difference with each on the boundaries.
 
@@
 \centering
-\begin{overpic}[width=0.49\columnwidth]{deltanum.pdf}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{deltadist.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
+\vspace{1em}
+\begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
 \caption{\label{fig:deltanumdist}
-Particle swap poles as their goal is the worst case. It can never be achieved but can be asymptotically reached. Plots show decreasing error as the number of moves grows.
+The worst case occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plots show decreasing error as the number of moves grows.
 } 

### 2018-05-14 journalWallFriction/theory.tex
A lot of edits to the pictures and other stuff

@@
 Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g.
-PRM and RRT* are not suitable \cite{lavalle2006planning,kavraki1996probabilistic}.
+PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
 Instead, this paper employs a graph search.
@@
 \mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
- \\
+\textrm{    ,} \\
  0 & \textrm{else}.
@@
 \ref{fig:reflection}.
-There are four walls, and four candidate solutions.
+There are four walls, and four candidate solutions, but some candidate solutions may be \emph{invalid} because they would cause both particles to hit the bounary in move $m_1$ (light grey regions) or \emph{not achievable} because they would cause both particles to hit the boundary in move $m_2$ (dark grey regions).
 
@@
  \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
+Changing the relative positions of particles in any workspace requires making one particles contact the boundary.
  The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}.
@@
 For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}.
+The closed form solution is long, and is in the paper attachments.

### 2018-05-14 journalWallFriction/theory2robAlgs.tex
A lot of edits to the pictures and other stuff

@@
 The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- Fig.~\ref{fig:regionMove} shows the starting and ending relative distance as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+ Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
 % in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
@@
 
- Alg.~\ref{alg:optimalAlg} first computes the reachable set.
-If the goal relative position is in the reachable set, we move particles to achieve relative position.
+ Algorithm ~\ref{alg:optimalAlg} first computes the reachable set.
+If the goal relative position is in the reachable set, we move particles to achieve that relative position.
 If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
- \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to push until one particle touches a wall, and the second to adjust the relative spacing.
+ \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
  Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
@@
 \end{figure}
+
+\begin{figure*}
+\centering
+\renewcommand{\figwid}{0.66\columnwidth}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){\tiny{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){\tiny{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){\tiny{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
+\end{overpic}
+\caption{\label{fig:polygonAlg}{ Steps to generate  the reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
+\vspace{-1em}
+}
+\end{figure*}
 If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
@@
 \State $p_{i+1}' \gets s_1 + s_2 - p_{i+1}$
-\State $L \gets \overline{ p_i p_{i+1}}$ \Comment{line $(p_{i}', p_{i+1}')$}
+\State $L \gets \overline{ p_i' p_{i+1}'}$ \Comment{line $(p_{i}', p_{i+1}')$}
 \State $l_i, l_{i+1} \gets $ intersections of $L$ and polygon $P$
@@
 
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){\tiny{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){\tiny{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){\tiny{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
-\end{overpic}
-\caption{\label{fig:polygonAlg}{ Steps to generate  the reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
-\vspace{-1em}
-}
-\end{figure*}
-
 \subsection{Circular Workspaces: Reachable Set}
@@
 % Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-To compute the reachable set for a circular workspace, first we consider all possible first contact locations.
- The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.
- %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
- We define the angle between two particles as $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
-
 \begin{figure}
@@
 \end{figure*}
+To compute the reachable set for a circular workspace, first we consider all possible first contact locations.
+ The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.
+ %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
+ We define the angle between two particles as $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
 
@@
 
+\begin{figure}
+\centering
+\begin{overpic}[width=0.95\columnwidth]{zaxis.pdf}\end{overpic}
+\caption{\label{fig:zaxis}
+Illustration on how boundary contacts enable 3D positioning.
+Once one particle contacts a boundary, the other particle's reachable set is a prism formed by extending the 2D reachable set in the $\pm z$-direction.
+} %\vspace{-1em}
+\end{figure}
+
 Extending path planning to 3D is possible only if the two particles do not initially have the same $x$ and $y$ positions.
@@
 Other workspaces may be better handled by other path planners, such as \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.
-
-\begin{figure}
-\centering
-\begin{overpic}[width=0.95\columnwidth]{zaxis.pdf}\end{overpic}
-\caption{\label{fig:zaxis}
-Illustration on how boundary contacts enable 3D positioning.
-Once one particle contacts a boundary, the other particle's reachable set is a prism formed by extending the 2D reachable set in the $z$-direction.
-} %\vspace{-1em}
-\end{figure}

### 2018-05-16 journalWallFriction/theory.tex
changed order of section on the model, removed one plot

@@
 
+ The system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
+ In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
  If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
@@
 
-These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
- In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
   The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of particles that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
@@
 \ref{fig:reflection}.
-There are four walls, and four candidate solutions, but some candidate solutions may be \emph{invalid} because they would cause both particles to hit the bounary in move $m_1$ (light grey regions) or \emph{not achievable} because they would cause both particles to hit the boundary in move $m_2$ (dark grey regions).
+There are four walls, and four candidate solutions, but some candidate solutions may be \emph{invalid} because they would cause both particles to hit the boundary in move $m_1$ (light grey regions) or \emph{not achievable} because they would cause both particles to hit the boundary in move $m_2$ (dark grey regions).
 
@@
 The problem can be simplified by choosing the coordinate system carefully.
-We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis: $P=(\cos \theta,\sin \theta)$.
+We define the $x$-axis along the line from the circle center to the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$-axis: $P=(\cos \theta,\sin \theta)$.
 Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$.
 Then the length of the two line segments is
- \begin{equation}
-\scriptsize{ \sqrt{(s-\cos \theta)^2+\sin^2 \theta}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}},
- \end{equation}
- which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to
+ \begin{align}\scalebox{.8}{$
+ \sqrt{{ (s-\cos \theta)^2+\sin^2 \theta}} +  \sqrt{(e \cos \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2},$}
+ \end{align}
+ which is minimized by choosing an appropriate $\theta$ value.
 
- \begin{equation}
-\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
- \end{equation}
-
+%  This equation can be simplified to 
+% 
+% \begin{equation}
+%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
+% \end{equation}
+% 
 \begin{figure}
@@
 Label the origin $O$.
-The optimal solution shows that the angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$).
-We name these angles $\alpha$.
-This can be proved by drawing an ellipse whose foci are $S$ and $E$.
-When the ellipse is tangent to the circle, the point of tangency is exactly $P$.
+ The optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+ The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
+ We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is exactly $P$. 
   Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
  From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$.
-If we mirror the point $S$ about the $\theta$ axis and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.
+If we mirror the point $S$ about line $\overline{OP}$ and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.
 

### 2018-05-16 journalWallFriction/theory2robAlgs.tex
changed order of section on the model, removed one plot

@@
  Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{\textrm{th}}$ side.
-If one particle hits side $i$ before the other ($\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $), the reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
+ % $\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $
+  If one particle hits side $i$ before the other (one particle will always hit before the other unless  the particles are parallel to the wall), the reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the reachable set of $\Delta$ configurations.
@@
  The two candidate shortest-length paths that touch the wall are shown with parallel arrows. 
-Each arrow will cause one of the particles to touch the wall, enabling the other particle to move freely in the  $z$ axis to achieve the required relative position.
+Each arrow will cause one of the particles to touch the wall, enabling the other particle to move freely in the  $z$-axis to achieve the required relative position.
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.

### 2018-05-17 journalWallFriction/simulation.tex
changed order of section on the model, removed one plot
A lot of changes to figures.

@@
 If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller.
-The bottom row of Fig.~\ref{fig:contour} and Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
+The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
 
@@
 \centering
-\begin{overpic}[width=0.49\columnwidth]{numcontour.pdf}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{distcontour.pdf}\end{overpic}
- \vspace{-2em}
+\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\end{overpic}
 \caption{\label{fig:contour}
@@
 \centering
-\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
-\vspace{1em}
+%\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
+%\vspace{1em}
 \begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
 \caption{\label{fig:deltanumdist}
-The worst case occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plots show decreasing error as the number of moves grows.
+The worst case occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
 } 

### 2018-05-22 journalWallFriction/experiment.tex
A lot of changes to figures.
Only requires the algorithm for irregular polygon

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[htb!]\label{fig:3dPrinted}
+\centering
+\vspace{1.5em}
+%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
+\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
+\\
+\vspace{1em}
+\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
+\caption{\label{fig:story}
+Frames showing particle positions before and after control inputs. Top row: small intestine phantom. Bottom row: cow stomach tissue.
+} \vspace{-1em}
+\end{figure*}
 
@@
 As shown in the video attachment, the particle closer to the coil moves faster than the other particle.
-Also, magnetic forces are not exactly parallel, but point toward the activated coil.
-Algorithm~\ref{alg:optimalAlg} is robust to these non uniformities, but sometimes requires additional iterations.
+Also, magnetic forces are not exactly parallel, but point toward the center of activated coil.
+Algorithm~\ref{alg:optimalAlg} is robust to these non-uniformities, but sometimes requires additional iterations.
 
@@
 Strips of cow stomach approximately $\SI{5}{\milli\metre}$ thick were cut and sewn to acrylic cylinder and then glued to an acrylic substrate using cyanoacrylate (superglue).
-This assemble was then filled with corn syrup.
+This assembly was then filled with corn syrup.
 The experiment is shown in Fig.~\ref{fig:story} bottom row.
@@
 %todo: snapshots of the the beads in the model, moving toward goal.  We can add lots of annotations.  Fake this series so we have an image for now.
-
-\begin{figure*}\label{fig:3dPrinted}
-\centering
-\vspace{1.5em}
-%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
-\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
-\\
-\vspace{1em}
-\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
-\caption{\label{fig:story}
-Frames showing particle positions before and after control inputs. Top row: small intestine phantom. Bottom row: cow's stomach tissue.
-} \vspace{-1em}
-\end{figure*}

### 2018-05-22 journalWallFriction/intro.tex
Only requires the algorithm for irregular polygon

@@
 To extend the range of possible shapes,
-  Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
+  Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
 We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
@@
 
-This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces.
+This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces.
 This work extends the analysis to convex workspaces and 3D positioning.

### 2018-05-22 journalWallFriction/simulation.tex
Only requires the algorithm for irregular polygon

@@
 %\subsection{Position Control of Two Robots}
-
-Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. 
-%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
-%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
- %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.
-
- The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
-Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} and middle and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
-Fig.~\ref{fig:contour} right shows the total distance of the path.
-This plot shows the nonlinear nature of the path planning.
-The path length grows when the goals have $\pi$ difference and are very close to the boundary.
- The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
-This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
- Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace.
-Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
-All of the discussed plots have considered the particles to be unique.
-If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller.
-The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
-
- %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
-
-% The plots in Fig.~\ref{fig:deltanumdist} show the exponentially increasing number of moves and distance when the accuracy of reaching to the goal ($\delta$) is getting to zero when the goal positions have $\pi$ difference with each on the boundaries.
-
 \begin{figure}
 \centering
-\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){a)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){b)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){c)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){d)}\end{overpic}
 \caption{\label{fig:contour}
@@
 \begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
+\vspace{-1em}
 \caption{\label{fig:deltanumdist}
-The worst case occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
+The worst case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
 } 
-\vspace{-1em}
 \end{figure}
@@
 \end{figure}
+Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. 
+%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
+%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
+ %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.
+
+ The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
+Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
+Fig.~\ref{fig:contour} right shows the total distance of the path.
+This plot shows the nonlinear nature of the path planning.
+The path length grows when the goals have $\pi$ difference and are very close to the boundary.
+TODO:????
+We need to discuss the new plots.
+ The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
+This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
+ Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace.
+Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
+All of the discussed plots have considered the particles to be unique.
+If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller.
+The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
+
+ %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
+
+% The plots in Fig.~\ref{fig:deltanumdist} show the exponentially increasing number of moves and distance when the accuracy of reaching to the goal ($\delta$) is getting to zero when the goal positions have $\pi$ difference with each on the boundaries.

### 2018-05-22 journalWallFriction/theory.tex
Only requires the algorithm for irregular polygon

@@
 }
-\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference.
-The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right.
+\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference.
+The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right.
 See \cite{BeckerShortestPath}.}
@@
  The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
- We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is exactly $P$. 
+ We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is $P$. 
   Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
@@
 We evaluate each and select the solution that results in the shortest length path.
-This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
-For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}.
-The closed form solution is long, and is in the paper attachments.
+%This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+ For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}. Because the closed form solution is long, it is included in the paper attachments.

### 2018-05-22 journalWallFriction/theory2robAlgs.tex
Only requires the algorithm for irregular polygon

@@
 Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
+Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of a square workspace for different cases.
 Alg.~\ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
@@
 
+\begin{figure*}
+\renewcommand{\figwid}{0.4\columnwidth}
+
+{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
+\put(10,85){a)}
+\put(30,70){$s_1$}
+\put(48,88){$g_1$}
+\put(65,17){$s_2$}
+\put(87,52){$g_2$}
+\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
+\put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
+\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
+%\put(47, 65){$\xleftrightarrow{1.8cm}$}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
+\put(10,85){b)}
+\put(28,70){$s_1$}
+\put(50,80){$g_1$}
+\put(62,15){$s_2$}
+\put(30,22){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
+\put(110,50){Three move sequence}
+\put(12,38){$m_3$}
+\put(48,52){$m_2$}
+\put(65,30){$m_1$}
+\end{overpic}
+%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
+%\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
+\put(10,85){c)}
+\put(28,70){$s_1$}
+\put(62,28){$g_1$}
+\put(87,50){$s_2$}
+\put(30,22){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+%\vspace{1em}
+{
+\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
+\put(10,85){d)}
+\put(25,30){$s_1$}
+\put(89,70){$g_1$}
+\put(67,20){$s_2$}
+\put(28,75){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
+\put(10,85){e)}
+\put(15,35){$s_1,g_2$}
+\put(65,35){$s_2,g_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
+\put(10,85){f)}
+\put(15,10){$s_1,g_2$}
+\put(65,88){$s_2,g_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
+\put(15, 85){worst case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
+\end{overpic}
+}\\
+
+%\vspace{-0.75em}
+%{
+%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
+%\end{overpic}
+%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
+%%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
+%\end{overpic}
+%}\\
+
+\vspace{-1em}
+\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two partivle positioning using walls with non-slip contacts. 
+Particle start positions are shown by a square, and goal positions by a circle.
+Dashed lines show the shortest route if partivles could be controlled independently.
+Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
+Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
+%The bottom row shows an extreme case where the robots must switch position.
+}
+\vspace{-1em}
+}
+\end{figure*}
+
 \subsection{Two Particle Path Planning}
@@
 For this reason, our algorithms use the two dimensional $\Delta$ configuration space, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
+% Each side of the $\Delta$ configuration space has the same length of each side of the workspace if the workspace has odd number of sides therefore, the $\Delta$ configuration space has two times the number of sides.
+$\Delta$ configuration space has the same number of sides for the polygons that have even number of sides, and each side of the $\Delta$ configuration space is twice as big as the workspace sides.
+
+The $\Delta$ configuration space is a set of all possible $\Delta p$ values.
+ $\Delta$ configuration spaces for a representative set of workspaces are shown in Fig.~\ref{fig:polygon}.
+The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+ Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+The next subsections give procedures to compute the reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+We will discuss how to make reachable sets in the next subsections.
+
   Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
@@
 
-The $\Delta$ configuration space is a set of all possible $\Delta p$ values.
- $\Delta$ configuration spaces for a representative set of workspaces are shown in Fig.~\ref{fig:polygon}.
-The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
-% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
-We will discuss how to make reachable sets in the next subsections.
- The next subsections give procedures to compute the reachable set.
-
+  \begin{figure}
+\centering
+\renewcommand{\figwid}{0.8\columnwidth}
+{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
+\end{overpic}
+}
+\caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
+The sets reachable in two moves, called \emph{reachable sets} are drawn with transparent blue polygons.
+A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
+If $\Delta g$ is in the reachable sets, we can achieve the required relative position in two moves.
+If $\Delta g$ is not in the reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
+We do it until we reach the relative goal position.
+%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
+As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
+}
+\vspace{-1em}
+}
+\end{figure}
  Algorithm ~\ref{alg:optimalAlg} first computes the reachable set.
@@
 Doing this for all the vertices of the workspace will give us all the reachable sets.
- \begin{figure}
-\centering
-\renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
-\end{overpic}
-}
-\caption{\label{fig:polygon}{Reachable sets are drawn with transparent blue polygons.
-A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
-%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
-As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
-}
-\vspace{-1em}
-}
-\end{figure}
 
@@
 \centering
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
-\begin{overpic}[width=0.47\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{twoRobotRegionH.pdf}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{DeltaConfigSquare.pdf}\end{overpic}
 \caption{\label{fig:TwoRegions}
@@
 (left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
-(right) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area. 
+(middle) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
+(right) The reachable sets in the $\Delta$ configuration space is shown.
 }
@@
 \renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(15,0){\tiny{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(17,-3){\small{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,0){\tiny{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,-3){\small{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(25,0){\tiny{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(17,-3){\small{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
 \end{overpic}
@@
 \begin{figure*}
-\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(7,15){\small{workspace}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(7,15){\small{workspace}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(7,15){\small{workspace}}\end{overpic}\\
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\end{overpic}\\
+
+\begin{overpic}[width=0.67\columnwidth]{regionMove4.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove5.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove6.pdf}\end{overpic}\\
 
@@
 \begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{finalPath.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move3.pdf}\end{overpic}\\
+
+\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
 \caption{\label{fig:reachableSet}

### 2018-05-22 journalWallFriction/theory2robAlgs.tex
small updates showing todos in red

@@
  $\Delta$ configuration spaces for a representative set of workspaces are shown in Fig.~\ref{fig:polygon}.
-The \emph{reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
  Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
@@
 
-\subsection{Convex Polygonal Workspaces: Reachable Set}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%\
+\subsection{$\Delta$ Configuration Space}
+
+ Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is an $2n$-sided convex polygon.
+ A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided.
+
+ \textcolor{red}{[INSERT PROCEDURE]}\\
+
+ \textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
+
+ \textcolor{red}{[INSERT ONE IMAGE OF A 4-sided irregular convex polygon]}\\
+
+\subsection{Convex Polygonal Workspaces: 2-Move Reachable Set}
 
@@
 
-\subsection{Circular Workspaces: Reachable Set}
+\subsection{Circular Workspaces: 2-Move Reachable Set}
 

### 2018-05-22 journalWallFriction/simulation.tex
small updates showing todos in red
Again :) full of image changes.

@@
 \centering
-\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){a)}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){b)}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){c)}\end{overpic}
-\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){d)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){a)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){b)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){c)}\end{overpic}
+\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){d)}\end{overpic}
 \caption{\label{fig:contour}
@@
  The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
-Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
-Fig.~\ref{fig:contour} right shows the total distance of the path.
+Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
+Fig.~\ref{fig:contour} right shows the number of moves and left shows the total distance of the path.
 This plot shows the nonlinear nature of the path planning.
+When the goal is in the middle of the workspace, a symmetry in the path length is expected as the top row shows.
+The bottom row shows a shift in the goal position which will break the symmetry of the path length in the workspace.
+
 The path length grows when the goals have $\pi$ difference and are very close to the boundary.
-TODO:????
-We need to discuss the new plots.
 The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).

### 2018-05-22 journalWallFriction/theory2robAlgs.tex
Again :) full of image changes.

@@
 Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
-Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
+%Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
 %The bottom row shows an extreme case where the robots must switch position.
@@
  Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
-The next subsections give procedures to compute the reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
 We will discuss how to make reachable sets in the next subsections.
@@
 
-  \begin{figure}
-\centering
-\renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
-\end{overpic}
-}
-\caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
-The sets reachable in two moves, called \emph{reachable sets} are drawn with transparent blue polygons.
-A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
-If $\Delta g$ is in the reachable sets, we can achieve the required relative position in two moves.
-If $\Delta g$ is not in the reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
-We do it until we reach the relative goal position.
-%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
-As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
-}
-\vspace{-1em}
-}
-\end{figure}
- Algorithm ~\ref{alg:optimalAlg} first computes the reachable set.
-If the goal relative position is in the reachable set, we move particles to achieve that relative position.
-If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
+ Algorithm ~\ref{alg:optimalAlg} first computes the 2-move reachable set.
+If the goal relative position is in the 2-move reachable set, we move particles to achieve that relative position.
+If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$.
  \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
@@
 \While {$||\Delta p - \Delta g|| > \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_{\textrm{SET}}\gets$  Compute reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or \ref{alg:circularReachbale}
+\State $R_{\textrm{SET}}\gets$  Compute 2-move reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or \ref{alg:circularReachbale}
 \State $ \Delta g_c\gets $nearest point in $R_{\textrm{SET}}$ to $\Delta g$
@@
 
- \textcolor{red}{[INSERT ONE IMAGE OF A 4-sided irregular convex polygon]}\\
+% \textcolor{red}{[INSERT ONE IMAGE OF A 4-sided irregular convex polygon]}\\
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{irregular.pdf}\end{overpic}
+\caption{\label{fig:irregular}
+Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-side reachable sets are drawn in transparent blue.
+}
+\end{figure}
 
 \subsection{Convex Polygonal Workspaces: 2-Move Reachable Set}
-
+   \begin{figure}
+\centering
+\renewcommand{\figwid}{0.8\columnwidth}
+{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
+\end{overpic}
+}
+\caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
+The sets reachable in two moves, called \emph{2-move reachable sets} are drawn with transparent blue polygons.
+A polygon with $n$ sides has $n$ 2-move reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
+If $\Delta g$ is in the 2-move reachable sets, we can achieve the required relative position in two moves.
+If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the 2-move reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
+We do it until we reach the relative goal position.
+%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
+As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
+}
+\vspace{-1em}
+}
+\end{figure}
  Fig. \ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
@@
 (middle) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
-(right) The reachable sets in the $\Delta$ configuration space is shown.
+(right) The 2-move reachable sets in the $\Delta$ configuration space is shown.
 }
@@
 \end{overpic}
-\caption{\label{fig:polygonAlg}{ Steps to generate  the reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
+\caption{\label{fig:polygonAlg}{ Steps to generate  the 2-move reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
 \vspace{-1em}
@@
 If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
-Alg.~\ref{alg:polygonReachbale} computes the reachable set for any convex workspace.
- Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the reachable set generated by collisions with the $i^{\textrm{th}}$ side.
+Alg.~\ref{alg:polygonReachbale} computes the 2-move reachable set for any convex workspace.
+ Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the 2-move reachable set generated by collisions with the $i^{\textrm{th}}$ side.
  % $\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $
-  If one particle hits side $i$ before the other (one particle will always hit before the other unless  the particles are parallel to the wall), the reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
-The union of these polygons for all $n$ sides is the reachable set of $\Delta$ configurations.
+  If one particle hits side $i$ before the other (one particle will always hit before the other unless  the particles are parallel to the wall), the 2-move reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
+The union of these polygons for all $n$ sides is the 2-move reachable set of $\Delta$ configurations.
  % by computing the reachable set for all the sides of the polygon.
@@
 Right: The $\Delta$ configuration space for the corresponded starting positions of the particles is shown.
-The possible reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
+The possible 2-move reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
 Once the blue particle contacts the boundary as shown, the reachable $\Delta$ configuration is the red set.}
@@
 \caption{\label{fig:reachableSet}
-Top row shows a polygonal workspace with its reachable sets.
+Top row shows a polygonal workspace with its 2-move reachable sets.
 Bottom row, left circle shows the workspace.
 %, and the corresponding reachable set when blue robot is touching the wall is shown as a blue region.
-Right shows the $\Delta$ configuration space and the reachable set that is shown in red is representative of the point we need to go to get to the goal relative distance in one move.%representative of the blue and pink reachable sets shown in the workspace.
+Right shows the $\Delta$ configuration space and the 2-move reachable set that is shown in red is representative of the point we need to go to get to the goal relative distance in one move.%representative of the blue and pink reachable sets shown in the workspace.
 }
 \end{figure*}
-To compute the reachable set for a circular workspace, first we consider all possible first contact locations.
+To compute the 2-move reachable set for a circular workspace, first we consider all possible first contact locations.
  The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.
@@
 A circle has an infinite number of sides, thus infinite reachable sets.
-However, the reachable set can be parameterized by the angle of first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where
+However, the 2-move reachable set can be parameterized by the angle of first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where
 %We make the reachable sets by knowing the angle of the chord, $\psi$, shown in Fig.~\ref{fig:psigamma}. Minimum and maximum values of $\psi$ can be calculated in the following equation.
@@
 \end{align} 
-Reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to immobilize the particle closest to a wall.
+2-move eachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to immobilize the particle closest to a wall.
 The reachable $\Delta$ configuration set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
@@
 \end{align}
-We combine these boundaries to compute the reachable set summarized in Alg.~\ref{alg:circularReachbale}.
-Next, find a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the reachable set to $\Delta g$.
+We combine these boundaries to compute the 2-move reachable set summarized in Alg.~\ref{alg:circularReachbale}.
+Next, find a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the 2-move reachable set to $\Delta g$.
 %Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
@@
 Illustration on how boundary contacts enable 3D positioning.
-Once one particle contacts a boundary, the other particle's reachable set is a prism formed by extending the 2D reachable set in the $\pm z$-direction.
+Once one particle contacts a boundary, the other particle's 2-move reachable set is a prism formed by extending the 2D 2-move reachable set in the $\pm z$-direction.
 } %\vspace{-1em}

### 2018-05-23 journalWallFriction/theory2robAlgs.tex
This draft contains all the required results.

@@
  Assume two particles are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-Denote the current positions of the particles  $p_1$ and $p_2$
+Denote the current positions of the particles  $p_1$ and $p_2$.
 
@@
 
- Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is an $2n$-sided convex polygon.
+ $\Delta$ configuration space is a space containing all the possible relative distances for $p_1$ and $p_2$.
+Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
  A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
+ To give an intuition what $\Delta$ configuration space is, consider the first particle on a vertex of the workspace.
+Now if the other particle moves on the boundary of the workspace, the relative distance between the two particles is maximized.
+Consider the first particle is fixed at (0,0).
+To generate the $\Delta$ configuration space we iterate through all the vertices, $i$ and consider that first particle is positioned at $i^{\textrm{th}}$ vertex.
+To find all possible $\Delta p$ values, compute $v_i - v_j$ where $j$ goes from 1 to $n$.
+For each vertex, $i$, we have made one $n$-sided polygon defined by its vertices.
+The outer boundary of all these polygons shows the boundary of the $\Delta$ configuration space.
+Fig.~\ref{fig:irregular} shows an arbitrary 4-sided workspace on the left.
+All the polygons made by $p_1$ and $p_2$ are shown in the middle.
+$\Delta$ configuration space is shown on the right with 2-move reachable sets drawn in transparent blue.
+We discuss how to generate 2-step reachable sets in the following subsection.
 
- \textcolor{red}{[INSERT PROCEDURE]}\\
+ %\textcolor{red}{[INSERT PROCEDURE]}\\
 
- \textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
+ %\textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
 
@@
 \caption{\label{fig:irregular}
-Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-side reachable sets are drawn in transparent blue.
+Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-move reachable sets are drawn in transparent blue.
 }

### 2018-05-24 journalWallFriction/theory2robAlgs.tex
added a todo

@@
 \vspace{-1em}
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two partivle positioning using walls with non-slip contacts. 
-Particle start positions are shown by a square, and goal positions by a circle.
-Dashed lines show the shortest route if partivles could be controlled independently.
+
+\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two particle positioning using walls with non-slip contacts. 
+Particle start positions are shown by squares, and goal positions by circles.
+Dashed lines show the shortest route if particles could be controlled independently.
 Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
-%Online demonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
+%Online dmonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
 %The bottom row shows an extreme case where the robots must switch position.
@@
 If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$.
+
+  \begin{figure}
+\centering
+\renewcommand{\figwid}{0.8\columnwidth}
+{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
+\end{overpic}
+}
+\caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
+The sets reachable in two moves, called \emph{reachable sets} are drawn with transparent blue polygons.
+A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
+If $\Delta g$ is in the reachable sets, we can achieve the required relative position in two moves.
+If $\Delta g$ is not in the reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
+We do it until we reach the relative goal position.
+%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
+As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
+}
+\vspace{-1em}
+}
+\end{figure}
+ Algorithm ~\ref{alg:optimalAlg} first computes the 2-move reachable set.
+If the goal relative position is in the reachable set, we move particles to achieve that relative position.
+If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
+>>>>>>> Stashed changes
  \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
@@
 
- $\Delta$ configuration space is a space containing all the possible relative distances for $p_1$ and $p_2$.
+The $\Delta$ configuration space is a space containing all the possible relative configurations for $p_1$ and $p_2$.
 Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
  A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
- To give an intuition what $\Delta$ configuration space is, consider the first particle on a vertex of the workspace.
-Now if the other particle moves on the boundary of the workspace, the relative distance between the two particles is maximized.
+ To give intuition about the $\Delta$ configuration space, consider placing the first particle on a vertex of the workspace. 
+ If the other particle moves on the boundary of the workspace, the relative distance between the two particles is maximized. 
  Consider the first particle is fixed at (0,0).
@@
 \caption{\label{fig:irregular}
+\textcolor{red}{Shiva, this image is nice.  A few suggestions: ?TODO: Label the middle image 'Translate workspace around (0,0)', and draw the workspace outlines with different colors and thicknesses}
 Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-move reachable sets are drawn in transparent blue.
@@
 Once the blue particle contacts the boundary as shown, the reachable $\Delta$ configuration is the red set.}
+
 \end{figure}

### 2018-05-24 journalWallFriction/theory2robAlgs.tex
Edited the procedure of Delta configuration space

--- 2018-05-22 journalWallFriction/theory2robAlgs.tex
@@
 
- Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is an $2n$-sided convex polygon.
+% $\Delta$ configuration space is a space containing all the possible relative distances for $p_1$ and $p_2$. 
+Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
  A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
+ To give an intuition, suppose that the first particle is positioned on a vertex of the workspace.
+The $\Delta$ configuration space is mainly determined by moving the second particle along the edges of the workspace and considering its relative distance to the first particle.
+%  what $\Delta$ configuration space is, consider the first particle on a vertex of the workspace.
+Now if the other particle moves on the boundary of the workspace, the relative distance between the two particles is maximized.
+In particular, to generate the $\Delta$ configuration space we iterate through all vertices $v_i$.  %In each iteration, we put the first particle at the $i^{\textrm{th}}$ vertex. 
+At each iteration, we compute $v_i - v_j$  ($1 \leq j \leq n$) and then make an $n$-sided polygon according to these values.
+Since our current workspaces are convex, the convex hull of these  $n$-sided polygons determines the boundary of the $\Delta$ configuration space.
+To give a concrete example, Fig.~\ref{fig:irregular} shows an arbitrary 4-sided workspace on the left.
+All the 4-sided polygons made by the computation of $(v_i-v_j)$s are shown in the middle and accordingly, $\Delta$ configuration space is shown on the right with 2-move reachable sets drawn in transparent blue.
+We will discuss how to generate 2-step reachable sets in the following subsection.
 
- \textcolor{red}{[INSERT PROCEDURE]}\\
+ %\textcolor{red}{[INSERT PROCEDURE]}\\
 
- \textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
+ %\textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
 

### 2018-05-24 journalWallFriction/theory2robAlgs.tex
Retried in writing the delta configuration procedure.

@@
 \caption{\label{fig:irregular}
-\textcolor{red}{Shiva, this image is nice.  A few suggestions: ?TODO: Label the middle image 'Translate workspace around (0,0)', and draw the workspace outlines with different colors and thicknesses}
+%\textcolor{red}{Shiva, this image is nice.  A few suggestions: ?TODO: Label the middle image 'Translate workspace around (0,0)', and draw the workspace outlines with different colors and thicknesses}
 Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-move reachable sets are drawn in transparent blue.

### 2018-05-28 journalWallFriction/Conclusion.tex
Text changes to improve readability

@@
 
-This paper presented techniques for controlling the positions of two particles using uniform inputs and non-slip boundary contact.  
+This paper presented techniques for controlling the positions of two particles using uniform inputs and non-slip boundary contacts.  
 The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D $\Delta$ configuration space.
-Extending Alg. \ref{alg:optimalAlg}  to 3D was straightforward, but increased the complexity.
+Extending Alg.~\ref{alg:optimalAlg}  to 3D was straightforward, but increased the complexity.
 Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue.

### 2018-05-28 journalWallFriction/intro.tex
Text changes to improve readability

@@
 \caption{\label{fig:IntroPic}
-Workspace and magnetic setup for an experiment of positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
 } \vspace{-1em}

### 2018-05-28 journalWallFriction/simulation.tex
Edited one caption.
Text changes to improve readability

@@
 \caption{\label{fig:deltanumdist}
-The worst case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
+The worst-case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
 } 
@@
 }\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
- The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
+ The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$.
+The middle row shows the number of moves required for the same configurations.
+The bottom row shows the same configuration but when the particles are interchangeable.}
 \vspace{-1em}
@@
 The path length grows when the goals have $\pi$ difference and are very close to the boundary.
-The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
+The worst-case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
 This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
@@
 Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
-All of the discussed plots have considered the particles to be unique.
-If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller.
+ Thus far, this paper has considered the particles to be unique.
+If particles are interchangeable, the path lengths often decrease.
 The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
+The worst-case path lengths decrease by ??\%, ??\%, and ??\% for the three test cases shown.
+%TODO: shiva, insert numbers.
 
- %If the length of each side of the square workspace is $L$, the worst case path length is $(\sqrt{2}+2)L$.
+ %If the length of each side of the square workspace is $L$, the worst-case path length is $(\sqrt{2}+2)L$.
 

### 2018-05-28 journalWallFriction/theory2robAlgs.tex
Text changes to improve readability

@@
 \begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
-\put(15, 85){worst case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
+\put(15, 85){worst-case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
 \end{overpic}
@@
   The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- Fig.~\ref{fig:regionMove} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+ Fig.~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
 The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
@@
 
-  \begin{figure}
-\centering
-\renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
-\end{overpic}
-}
-\caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
-The sets reachable in two moves, called \emph{reachable sets} are drawn with transparent blue polygons.
-A polygon with $n$ sides has $n$ reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
-If $\Delta g$ is in the reachable sets, we can achieve the required relative position in two moves.
-If $\Delta g$ is not in the reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
-We do it until we reach the relative goal position.
-%Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
-As the number of sides in the polygon increases, the total area of the $\Delta$ configuration space is four times of the workspace.
-}
-\vspace{-1em}
-}
-\end{figure}
  Algorithm ~\ref{alg:optimalAlg} first computes the 2-move reachable set.
@@
 If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
->>>>>>> Stashed changes
  \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
@@
 \subsection{$\Delta$ Configuration Space}
+% ANOTHER TRY
+The $\Delta$ configuration space is all possible configurations of $p_2-p_1$. 
+The  $\Delta$ configuration space for an $n$-sided convex polygon $P$ can be constructed by translating $n$ copies of $P$  so that each copy moves a different vertex of $P$ to $(0,0)$.
+Because $P$ is convex, the convex-hull of all these translated vertices is the boundary of the  $\Delta$ configuration space.
+ For an $n$-sided convex polygon, the $\Delta$ configuration space is a $2n$-sided convex polygon.
+  Even-sided regular polygons are a special case in which half the sides align and the $\Delta$ configuration space is $n$-sided. 
+An example $\Delta$ configuration space construction is shown in Fig.~\ref{fig:irregular}: 
+a four-sided workspace is on the left, 
+the four translated copies with dashed lines outlining the convex hull is in the middle,
+ and the resulting $\Delta$ configuration space is on the right.  
+ The 2-move reachable sets for the given $\Delta$s starting configuration are drawn in transparent blue.
+The following subsection describes how 2-move reachable sets are constructed.
 
-% $\Delta$ configuration space is a space containing all the possible relative distances for $p_1$ and $p_2$. 
-Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
- A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
- To give an intuition, suppose that the first particle is positioned on a vertex of the workspace.
+% ORIGINAL
+%Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
+% A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
+% To give an intuition, suppose that the first particle is positioned on a vertex of the workspace.
 The $\Delta$ configuration space is mainly determined by moving the second particle along the edges of the workspace and considering its relative distance to the first particle.
-%  what $\Delta$ configuration space is, consider the first particle on a vertex of the workspace.
-Now if the other particle moves on the boundary of the workspace, the relative distance between the two particles is maximized.
-In particular, to generate the $\Delta$ configuration space we iterate through all vertices $v_i$.  %In each iteration, we put the first particle at the $i^{\textrm{th}}$ vertex. 
-At each iteration, we compute $v_i - v_j$  ($1 \leq j \leq n$) and then make an $n$-sided polygon according to these values.
+%In particular, to generate the $\Delta$ configuration space we iterate through all vertices $v_i$.  %In each iteration, we put the first particle at the $i^{\textrm{th}}$ vertex. 
+%At each iteration, we compute $v_i - v_j$  ($1 \leq j \leq n$) and then make an $n$-sided polygon according to these values.
 Since our current workspaces are convex, the convex hull of these  $n$-sided polygons determines the boundary of the $\Delta$ configuration space.
@@
 \caption{\label{fig:polygon}{The $\Delta$ configuration space is all possible configurations of $p_2-p_1$.
-The sets reachable in two moves, called \emph{2-move reachable sets} are drawn with transparent blue polygons.
+The sets reachable in two moves, called \emph{2-move reachable sets}, are drawn with transparent blue polygons.
 A polygon with $n$ sides has $n$ 2-move reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
 If $\Delta g$ is in the 2-move reachable sets, we can achieve the required relative position in two moves.
-If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal, $\Delta g_c$, which is the closest point on the 2-move reachable set to $\Delta g$ and apply two moves that make the relative position closer to $\Delta g$.
-We do it until we reach the relative goal position.
+If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal $\Delta g_c$ (the closest point on the 2-move reachable set to $\Delta g$) and apply two moves to achieve $\Delta g_c$.
+We repeat this process until the relative goal position is achieved.
 %Workspace and $\Delta$ configuration spaces for different polygonal workspaces and their representative $\Delta$ configuration spaces and reachable sets.
@@
 \end{figure}
- Fig. \ref{fig:polygon} shows different workspaces and their representative $\Delta$ configuration spaces.
 
@@
 \end{figure*}
-If a particle is touching a wall, the other particle can move freely in the reachable set as shown in Fig.~\ref{fig:TwoRegions}.
+ Fig. \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
+  The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
 Alg.~\ref{alg:polygonReachbale} computes the 2-move reachable set for any convex workspace.
- Fig.~\ref{fig:polygonAlg} illustrates the procedure to construct the 2-move reachable set generated by collisions with the $i^{\textrm{th}}$ side.
- % $\overline{ s_1s_2 }\nparallel \overline{ p_i p_{i+1}} $
-  If one particle hits side $i$ before the other (one particle will always hit before the other unless  the particles are parallel to the wall), the 2-move reachable set is defined by a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
+The set is constructed by considering each edge of the workspace.
+  If one particle hits edge $i$ before the other (one particle will always hit before the other unless the particles are parallel to the wall), the corresponding 2-move reachable set is a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the 2-move reachable set of $\Delta$ configurations.
- % by computing the reachable set for all the sides of the polygon.
-Fig.~\ref{fig:polygonAlg} shows different cases with the particles and an arbitrary side.
-If particle 2 is touching the $i^{th}$ side, we compute if particle 1 will stay in the polygon or not.
-Then we generate the reachable set using all the vertices of the polygon and potential place of particle 1.
-We unify all the reachable sets corresponding to all the sides.
-Alg.~\ref{alg:polygonReachbale} computes reachable set for any convex polygonal workspace.
+Figure~\ref{fig:polygonAlg} illustrates the procedure to construct the  2-move reachable set generated by collisions with the $i^{\textrm{th}}$ edge.
+% 
  \begin{algorithm}[htb]
@@
 \caption{\label{fig:reachableSet}
-Top row shows a polygonal workspace with its 2-move reachable sets.
-Bottom row, left circle shows the workspace.
-%, and the corresponding reachable set when blue robot is touching the wall is shown as a blue region.
-Right shows the $\Delta$ configuration space and the 2-move reachable set that is shown in red is representative of the point we need to go to get to the goal relative distance in one move.%representative of the blue and pink reachable sets shown in the workspace.
+Frames from reconfiguring two particles. 
+Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
+The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
+See multimedia attachment for animations of each.
 }
 \end{figure*}
-To compute the 2-move reachable set for a circular workspace, first we consider all possible first contact locations.
+To compute the 2-move reachable set for a circular workspace, first consider all possible first contact locations.
  The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.

### 2018-05-29 journalWallFriction/theory2robAlgs.tex
updated pictures

@@
 \begin{figure*}
-\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\end{overpic}\\
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,  - , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
 
-\begin{overpic}[width=0.67\columnwidth]{regionMove4.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove5.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove4.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove5.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
 \begin{overpic}[width=0.67\columnwidth]{regionMove6.pdf}\end{overpic}\\
 
-\begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move3.pdf}\end{overpic}\\
+\begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move3.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
 
-\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
 \begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}

### 2018-05-29 journalWallFriction/SwarmShapeControl.tex
updated pictures
removed contradictory text after read through of paper.

@@
 \usepackage{rotating}
+\usepackage{tikz}
+\usepgflibrary{arrows}% for more options on arrows
 %\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
@@
 In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
-Because in vivo environments are usually not square, this paper extends the previous work to convex workspaces including circles and 3D positioning. 
+Because in vivo environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
 This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.

### 2018-05-29 journalWallFriction/intro.tex
removed contradictory text after read through of paper.

@@
 The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
- The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
+ The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
 Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.

### 2018-05-29 journalWallFriction/theory.tex
small updates showing todos in red
removed contradictory text after read through of paper.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action.
-Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g.
-PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
-Instead, this paper employs a graph search.
-This section starts with a boundary interaction model in subsection \ref{subsec:WallFriction}.
-
-Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle.
-In subsections \ref{subsec:square} and \ref{subsec:circular} we provide shortest-path results for two representative workspaces, squares and circles.
-\subsection{Boundary Interaction Model}\label{subsec:WallFriction}
-
  In the absence of obstacles, uniform inputs move a swarm identically.  
@@
 The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.  
+ Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle. 
+ We start with a boundary interaction model in subsection \ref{subsec:WallFriction}.
 
+If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
+ Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
+  Instead, this paper employs a greedy search strategy. 
+For some configurations, we can obtain the optimal solution. 
+Subsections \ref{subsec:square} and \ref{subsec:circular} provide shortest-path results for two representative workspaces, squares and disks.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Boundary Interaction Model}\label{subsec:WallFriction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  The system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
@@
 all tumblers of a combination lock move uniformly unless obstructed by an obstacle.
- Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
+ Our control problem is to design the control inputs $\mathbf{u}(t)$ to deliver two particles to goal positions.
 
@@
 \ref{fig:reflection}.
-There are four walls, and four candidate solutions, but some candidate solutions may be \emph{invalid} because they would cause both particles to hit the boundary in move $m_1$ (light grey regions) or \emph{not achievable} because they would cause both particles to hit the boundary in move $m_2$ (dark grey regions).
+  There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
 
@@
  \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
-Changing the relative positions of particles in any workspace requires making one particles contact the boundary.
- The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}.
+Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
+ The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments and has an optimal contact point, as shown in Fig.~\ref{fig:shortestpath}.
  The problem can be simplified by choosing the coordinate system carefully.

### 2018-05-29 journalWallFriction/theory2robAlgs.tex
removed contradictory text after read through of paper.

@@
 Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
-Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of a square workspace for different cases.
+Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a Mathematica implementation in a square workspace for six representative configurations.
 Alg.~\ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
@@
 
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%\
+\subsection{$\Delta$ Configuration Space}
+The configuration space for two particles is a four dimensional manifold.
+Translating both particles the same amount is a trivial operation, but changing the relative positions requires boundary interaction.
+For this reason, our algorithms use the two dimensional $\Delta$ configuration space.
+The $\Delta$ configuration space is a set of all possible $\Delta p$ values, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
+
+The  $\Delta$ configuration space for an $n$-sided convex polygon $P$ can be constructed by translating $n$ copies of $P$  so that each copy moves a different vertex of $P$ to $(0,0)$.
+Because $P$ is convex, the convex-hull of all these translated vertices is the boundary of the  $\Delta$ configuration space.
+ For an $n$-sided convex polygon, the $\Delta$ configuration space is a $2n$-sided convex polygon.
+  Even-sided regular polygons are a special case in which half the sides align and the $\Delta$ configuration space is $n$-sided. 
+An example $\Delta$ configuration space construction is shown in Fig.~\ref{fig:irregular}: 
+a four-sided workspace is on the left, 
+the four translated copies with dashed lines outlining the convex hull is in the middle,
+ and the resulting $\Delta$ configuration space is on the right.  
+% The 2-move reachable sets for the given $\Delta$s starting configuration are drawn in transparent blue.
+%The following subsection describes how 2-move reachable sets are constructed.
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%\
 \subsection{Two Particle Path Planning}
 
-The configuration space for two particles is a four dimensional manifold.
-Translating both particles the same amount is a trivial operation, but changing the relative positions requires boundary interaction.
-For this reason, our algorithms use the two dimensional $\Delta$ configuration space, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
-% Each side of the $\Delta$ configuration space has the same length of each side of the workspace if the workspace has odd number of sides therefore, the $\Delta$ configuration space has two times the number of sides.
-$\Delta$ configuration space has the same number of sides for the polygons that have even number of sides, and each side of the $\Delta$ configuration space is twice as big as the workspace sides.
-
-The $\Delta$ configuration space is a set of all possible $\Delta p$ values.
- $\Delta$ configuration spaces for a representative set of workspaces are shown in Fig.~\ref{fig:polygon}. 
   The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+  The rightmost image of Fig.~\ref{fig:irregular} draws the 2-move reachable sets for the given $\Delta$s starting configuration in transparent blue.
  Fig.~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
@@
 
- Algorithm ~\ref{alg:optimalAlg} first computes the 2-move reachable set.
-If the goal relative position is in the reachable set, we move particles to achieve that relative position.
-If it is not in the reachable set, we move particles to achieve the closest point on the reachable set from $\Delta g$.
  \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
@@
 
- %%%%%%%%%%%%%%%%%%%%%%%%%%%\
-\subsection{$\Delta$ Configuration Space}
-% ANOTHER TRY
-The $\Delta$ configuration space is all possible configurations of $p_2-p_1$. 
-The  $\Delta$ configuration space for an $n$-sided convex polygon $P$ can be constructed by translating $n$ copies of $P$  so that each copy moves a different vertex of $P$ to $(0,0)$.
-Because $P$ is convex, the convex-hull of all these translated vertices is the boundary of the  $\Delta$ configuration space.
- For an $n$-sided convex polygon, the $\Delta$ configuration space is a $2n$-sided convex polygon.
-  Even-sided regular polygons are a special case in which half the sides align and the $\Delta$ configuration space is $n$-sided. 
-An example $\Delta$ configuration space construction is shown in Fig.~\ref{fig:irregular}: 
-a four-sided workspace is on the left, 
-the four translated copies with dashed lines outlining the convex hull is in the middle,
- and the resulting $\Delta$ configuration space is on the right.  
- The 2-move reachable sets for the given $\Delta$s starting configuration are drawn in transparent blue.
-The following subsection describes how 2-move reachable sets are constructed.
-
-% ORIGINAL
-%Given an $n$-sided convex polygon $P$, the $\Delta$ configuration space is a $2n$-sided convex polygon.
-% A special case is for even-sided regular polygons, in which half the sides align and the $\Delta$ configuration space  is $n$-sided. 
-% To give an intuition, suppose that the first particle is positioned on a vertex of the workspace.
-The $\Delta$ configuration space is mainly determined by moving the second particle along the edges of the workspace and considering its relative distance to the first particle.
-%In particular, to generate the $\Delta$ configuration space we iterate through all vertices $v_i$.  %In each iteration, we put the first particle at the $i^{\textrm{th}}$ vertex. 
-%At each iteration, we compute $v_i - v_j$  ($1 \leq j \leq n$) and then make an $n$-sided polygon according to these values.
-Since our current workspaces are convex, the convex hull of these  $n$-sided polygons determines the boundary of the $\Delta$ configuration space.
-To give a concrete example, Fig.~\ref{fig:irregular} shows an arbitrary 4-sided workspace on the left.
-All the 4-sided polygons made by the computation of $(v_i-v_j)$s are shown in the middle and accordingly, $\Delta$ configuration space is shown on the right with 2-move reachable sets drawn in transparent blue.
-We will discuss how to generate 2-step reachable sets in the following subsection.
-
- %\textcolor{red}{[INSERT PROCEDURE]}\\
-
- %\textcolor{red}{    Describe how to generate a $\Delta$ Configuration Space for polygon $P$.}
-
-% \textcolor{red}{[INSERT ONE IMAGE OF A 4-sided irregular convex polygon]}\\
 \begin{figure}
@@
 To compute the 2-move reachable set for a circular workspace, first consider all possible first contact locations.
- The set of boundary points that a particle can touch before the  other particle  touches is an arc of angle $2(\pi - \frac{\arcsin{d_{12}}}{r})$, where $d_{12}= ||s_1 - s_2||$ and $r$ is the radius of the workspace.
- %Here $\alpha$ is the arc of the circle that it is not possible to keep the current relative distance while one robot is touching the wall in that point and is defined by $\alpha = 2 \sin ^{-1}(d)$. 
- We define the angle between two particles as $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
+ The set of boundary points that a particle can touch before the  other particle  touches are two arcs from $\psi_{\min}$ to $\psi_{\max}$  and from $\pi+ \psi_{\min}$ to $\pi+ \psi_{\max}$:
+ \begin{align}
+ \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
+\end{align}
+where $d_{12}= ||s_1 - s_2||$, $r$ is the radius of the workspace,
+and the angle between two particles is $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
 
@@
 A circle has an infinite number of sides, thus infinite reachable sets.
-However, the 2-move reachable set can be parameterized by the angle of first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove} where
-%We make the reachable sets by knowing the angle of the chord, $\psi$, shown in Fig.~\ref{fig:psigamma}. Minimum and maximum values of $\psi$ can be calculated in the following equation.
-\begin{align}
- \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d}}{2r} \Big].
-% \psi_{\max} &= \theta -\frac{\alpha}{2} + \frac{\pi}{2}
-\end{align}
-The possible first contact locations are on an arc with interior angle $\gamma$ parameterized by $\psi$:
+However, the 2-move reachable set can be parameterized by the angle of first contact location $\psi$, as shown in Fig.~\ref{fig:regionMove}.
+
+Each $\phi$ value generates a 2-move reachable set that is a chord of the disk, with interior angle $\gamma$ parameterized by $\psi$:
 \begin{align}\label{eq:gamma}
@@
 \end{align} 
-2-move eachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to immobilize the particle closest to a wall.
+The 2-move reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to minimize the total distance.
+%immobilize the particle closest to a wall. 
 The reachable $\Delta$ configuration set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
@@
 
-The equations for the four lines outlining the union of reachable $\Delta$ configuration sets are as follows:
+The equations for the four lines outlining the union of  two-move reachable sets are as follows:
 \begin{align}\label{eq:circlereachable}
@@
 We combine these boundaries to compute the 2-move reachable set summarized in Alg.~\ref{alg:circularReachbale}.
-Next, find a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the 2-move reachable set to $\Delta g$.
+The  motion-planner finds a $\psi$ that would enable us to reach $\Delta g_c$, the nearest point in the 2-move reachable set to $\Delta g$.
 %Eq.~\eqref{eq:ifinchord} checks if an arbitrary point $p$, is in the reachable set corresponding to $\psi$.
-We first check if $\Delta g_c$ is in the $\Delta$ configuration space chords defined by either $\psi_{\min}$ or $\psi_{\max}$ using:
+We first check if $\Delta g_c$ is in the $\Delta$ configuration space chords defined by either $\psi_{\min}$ or $\psi_{\max}$ using the following two tests:
 

### 2018-05-29 journalWallFriction/experiment.tex
Updated with the latest comments.

@@
 Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
-The magnetic field in this setup is only approximately uniform.
+Our experimental setup did not perfectly implement the system dynamics in \eqref{swarmDynamicsAndFric}.
+In particular, the magnetic field in this setup is only approximately uniform.
 The magnetic force varies in both magnitude and orientation.
-As shown in the video attachment, the particle closer to the coil moves faster than the other particle.
-Also, magnetic forces are not exactly parallel, but point toward the center of activated coil.
+As shown in the video attachment, this non-uniformity causes the particle closer to the coil to move faster than the other particle.
+Also, magnetic forces are not exactly parallel, but point toward the center of the activated coil.
 Algorithm~\ref{alg:optimalAlg} is robust to these non-uniformities, but sometimes requires additional iterations.

### 2018-05-29 journalWallFriction/simulation.tex
Updated with the latest comments.

@@
  The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
-Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
+Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
 Fig.~\ref{fig:contour} right shows the number of moves and left shows the total distance of the path.
@@
 When the goal is in the middle of the workspace, a symmetry in the path length is expected as the top row shows.
-The bottom row shows a shift in the goal position which will break the symmetry of the path length in the workspace.
+The bottom row shows a shift in the goal position which breaks the symmetry of the path length in the workspace.
 
-The path length grows when the goals have $\pi$ difference and are very close to the boundary.
+%The path length grows when the goals have $\pi$ difference and are very close to the boundary. 
  The worst-case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
 This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
- Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace.
-Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
+ Figure \ref{fig:contourPlots} shows the same concepts in a square workspace.
+Figure \ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
  Thus far, this paper has considered the particles to be unique.

### 2018-05-29 journalWallFriction/theory2robAlgs.tex
Updated with the latest comments.

@@
 \begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
-\put(110,50){Three move sequence}
+\put(110,50){Three-move sequence}
 \put(12,38){$m_3$}
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(17,-3){\small{Generates polygon defined by $\overline{ p_i p_{i+1}}$ }}
+\begin{overpic}[width =\figwid]{anypolygon3.pdf}\put(-6,-3){\small{Polygon generated by first contacts along $\overline{ p_i p_{i+1}}$ }}
 \end{overpic}
@@
  The set of boundary points that a particle can touch before the  other particle  touches are two arcs from $\psi_{\min}$ to $\psi_{\max}$  and from $\pi+ \psi_{\min}$ to $\pi+ \psi_{\max}$:
- \begin{align}
+ \begin{align}\label{eq:psiMinMax}
  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
@@
 \begin{align}\label{eq:circlereachable}
-l_1 =  r \Big(&(\cos\psi_{\min}- \cos(\gamma + \psi_{\min}) )\\ \nonumber
- + &(\sin\psi_{\min}- \sin(\gamma + \psi_{\min}))\Big) &  0<\gamma< \gamma(\psi_{\min}),\\ \nonumber
-l_2 =  r \Big(&(\cos\psi_{\max}- \cos(\gamma + \psi_{\max}))\\ \nonumber
- + &(\sin\psi_{\max}- \sin(\gamma + \psi_{\max}))\Big) &  \gamma(\psi_{\max})<\gamma< 0,\\  \nonumber
-l_3 =  r \Big(&(\cos\psi- \cos( \psi+\gamma(\psi) ) )\\ \nonumber
-+ &( \sin\psi-\sin( \psi+ \gamma(\psi)))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
-l_4 =  r \Big(&(\cos\psi- \cos( \psi-\gamma(\psi) ))\\ \nonumber
-+ & ( \sin\psi- \sin( \psi- \gamma(\psi)))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
+l_1 =  r \Big(&\cos\psi_{\min}- \cos(\gamma + \psi_{\min} )\\ \nonumber
+ + &\sin\psi_{\min}- \sin(\gamma + \psi_{\min})\Big) &  0<\gamma< \gamma(\psi_{\min}),\\ \nonumber
+l_2 =  r \Big(&\cos\psi_{\max}- \cos(\gamma + \psi_{\max})\\ \nonumber
+ + &\sin\psi_{\max}- \sin(\gamma + \psi_{\max})\Big) &  \gamma(\psi_{\max})<\gamma< 0,\\  \nonumber
+l_3 =  r \Big(&\cos\psi- \cos( \psi+\gamma(\psi) )\\ \nonumber
++ & \sin\psi-\sin( \psi+ \gamma(\psi))\Big) &  \psi_{\min}<\psi< \psi_{\max},\\ \nonumber
+l_4 =  r \Big(&\cos\psi- \cos( \psi-\gamma(\psi) )\\ \nonumber
++ &  \sin\psi- \sin( \psi- \gamma(\psi))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
 \end{align}
@@
 %\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
-\State Calculate $p_{\psi}$ \Comment use \eqref{eq:ppsi}
-\State Calculate $\gamma$ \Comment use \eqref{eq:gamma}
+\State Calculate $\psi_{\min}$ and $\psi_{\max}$ \Comment use \eqref{eq:psiMinMax}
+\State Calculate $\gamma(\psi)$ \Comment use \eqref{eq:gamma}
 \State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
@@
 If the 3D projection is at a different angle, redefine the 2D workspace as a region perpendicular to the projection.
- First, we move the closest particle to the boundary, which prevents its $z$ coordinate from changing.  
+ First, we move the closest particle to the boundary, which prevents its $z$-coordinate from changing.  
  We next apply actuation in either the $\pm z$ direction to achieve the desired $\Delta z$.

### 2018-05-29 journalWallFriction/experiment.tex
A full read through-Shiva

@@
 Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
-Our experimental setup did not perfectly implement the system dynamics in \eqref{swarmDynamicsAndFric}.
+Our experimental setup did not perfectly implement the system dynamics in \eqref{eq:swarmDynamicsAndFric}.
 In particular, the magnetic field in this setup is only approximately uniform.
@@
 Also, magnetic forces are not exactly parallel, but point toward the center of the activated coil.
-Algorithm~\ref{alg:optimalAlg} is robust to these non-uniformities, but sometimes requires additional iterations.
+Algorithm~\ref{alg:optimalAlg} can still work with  these non-uniformities, but sometimes requires additional iterations.
 
@@
 The experiment is shown in Fig.~\ref{fig:story} bottom row.
+Video of this experiment is available in the supplementary materials.
 %In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.

### 2018-05-29 journalWallFriction/intro.tex
Updated with the latest comments.
A full read through-Shiva

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics. 
+Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
 %If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
@@
  Techniques are needed that can handle this underactuation.
+
+ Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
  In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
-
-Positioning is a foundational capability for a robotic system, e.g.
-placement of brachytherapy seeds.
  However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
@@
 \begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
-\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(28,-8){magnetic setup}\end{overpic}
+\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
 \vspace{1em}
@@
   Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
-We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
+We provide an algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
 Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 

### 2018-05-29 journalWallFriction/simulation.tex
A full read through-Shiva

@@
 %Two simulations were implemented using non-slip contact walls for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.  
+\begin{figure*}
+\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,  - , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
+
+\begin{overpic}[width=0.67\columnwidth]{regionMove4.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove5.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{regionMove6.pdf}\end{overpic}\\
+
+\begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move3.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
+
+\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
+\caption{\label{fig:reachableSet}
+Frames from reconfiguring two particles. 
+Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
+The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
+See multimedia attachment for animations of each.
+}
+\end{figure*}
 
@@
 Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius.
+Figure \ref{fig:reachableSet} shows frames of the algorithm in two representative workspaces, square and disk, with two arbitrary starting and goal configurations.
 %An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.

### 2018-05-29 journalWallFriction/theory.tex
A full read through-Shiva

@@
 
-If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
- Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
-  Instead, this paper employs a greedy search strategy. 
 For some configurations, we can obtain the optimal solution. 
@@
  \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
-
+ Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
  If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
@@
  \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
-Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
+
  The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments and has an optimal contact point, as shown in Fig.~\ref{fig:shortestpath}.

### 2018-05-29 journalWallFriction/theory2robAlgs.tex
A full read through-Shiva

@@
  Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a Mathematica implementation in a square workspace for six representative configurations.
-Alg.~\ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
-In the last subsection we present techniques to control 3D positioning of two particles.
-
  Workspaces are 2D convex polygons with no internal obstacles. 
  Assume two particles are initialized at $s_1$ and $s_2$ with corresponding goal destinations $g_1$ and $g_2$. 
-Denote the current positions of the particles  $p_1$ and $p_2$.
+ Denote the current positions of the particles  $p_1$ and $p_2$. Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
+ Algorithm \ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
+In the last subsection we present techniques to control 3D positioning of two particles.
 
@@
   The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
-  The rightmost image of Fig.~\ref{fig:irregular} draws the 2-move reachable sets for the given $\Delta$s starting configuration in transparent blue.
- Fig.~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+ For the given $\Delta s$ (starting configuration), the rightmost image of Fig.~\ref{fig:irregular} draws the 2-move reachable sets  in transparent blue.
+ Figure ~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
 The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
@@
 
-  Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
-Alg.~\ref{alg:optimalAlg} assigns a uniform control input at every instance.
-The goal is to move the particles within $\epsilon$ of the goal positions using a shared control input where $\epsilon$ is an arbitrary small number.
-We do this by first moving them within $\epsilon$ of the correct relative position and then translating the particles to the goal.
+The goal is to move the particles within $\delta$ of the goal positions using a shared control input where $\delta$ is an arbitrary small number.
+We do this by first moving them within $\delta$ of the correct relative position and then translating the particles to the goal.
 The relative position is $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
 
- Algorithm ~\ref{alg:optimalAlg} first computes the 2-move reachable set.
+ Algorithm \ref{alg:optimalAlg} assigns a uniform control input at every instance.
+% Algorithm ~\ref{alg:optimalAlg} 
+ It first computes the 2-move reachable set.
 If the goal relative position is in the 2-move reachable set, we move particles to achieve that relative position.
-If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$.
+If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$, which is $\Delta g_c$.
 
@@
  Fig. \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
-  The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
-Alg.~\ref{alg:polygonReachbale} computes the 2-move reachable set for any convex workspace.
-The set is constructed by considering each edge of the workspace.
-  If one particle hits edge $i$ before the other (one particle will always hit before the other unless the particles are parallel to the wall), the corresponding 2-move reachable set is a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
+ % The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+Algorithm \ref{alg:polygonReachbale} computes the 2-move reachable set for any convex workspace.
+The set is constructed by considering each edge of the workspace. We name each vertex as $p_i$ where $1\leq i \leq n$.
+  If one particle hits edge $\overline{ p_i p_{i+1}}$ before the other (one particle will always hit before the other unless the particles are parallel to the wall), the corresponding 2-move reachable set is a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the 2-move reachable set of $\Delta$ configurations.
-Figure~\ref{fig:polygonAlg} illustrates the procedure to construct the  2-move reachable set generated by collisions with the $i^{\textrm{th}}$ edge.
+Figure~\ref{fig:polygonAlg} illustrates the procedure to construct the  2-move reachable set generated by collisions with the $\overline{ p_i p_{i+1}}$ edge.
 % 
@@
 
+Figure \ref{fig:TwoRegions} shows reachable sets for a square workspace.
+There are four reachable sets, the horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
+
 \subsection{Circular Workspaces: 2-Move Reachable Set}
@@
 
-\begin{figure*}
-\begin{overpic}[width=0.67\columnwidth]{regionMove1.pdf}\put(55.5,26){\begin{turn}{94} 
-{\begin{tikzpicture}[thick]
-\draw [red,  - , dotted      ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove2.pdf}\put(55.5,26){\begin{turn}{94} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -  , dotted     ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove3.pdf}\put(55.5,26){\begin{turn}{94} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}\\
-
-\begin{overpic}[width=0.67\columnwidth]{regionMove4.pdf}\put(55.5,26){\begin{turn}{94} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove5.pdf}\put(55.5,26){\begin{turn}{94} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{regionMove6.pdf}\end{overpic}\\
-
-\begin{overpic}[width=0.67\columnwidth]{Move1.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move2.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move3.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}\\
-
-\begin{overpic}[width=0.67\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
-{\begin{tikzpicture}[thick]
-\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
-\end{tikzpicture}}\end{turn}}\end{overpic}
-\begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
-\caption{\label{fig:reachableSet}
-Frames from reconfiguring two particles. 
-Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
-Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
-For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
-The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
-See multimedia attachment for animations of each.
-}
-\end{figure*}
 To compute the 2-move reachable set for a circular workspace, first consider all possible first contact locations.
@@
 
-Each $\phi$ value generates a 2-move reachable set that is a chord of the disk, with interior angle $\gamma$ parameterized by $\psi$:
+Each $\psi$ value generates a 2-move reachable set that is a chord of the disk, with interior angle $\gamma$ parameterized by $\psi$:
 \begin{align}\label{eq:gamma}
@@
 \end{align} 
-The 2-move reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space, so we can plan in this space and choose between the two options to minimize the total distance.
+The 2-move reachable sets with $\pi$ difference in $\psi$ value are equivalent in the  $\Delta$ configuration space. 
 %immobilize the particle closest to a wall. 
 The reachable $\Delta$ configuration set for any first contact point defined by $\psi$ is the area under a chord from angle $\psi- \frac{\gamma(\psi)}{2}$ to $\psi+ \frac{\gamma(\psi)}{2}$, for a circle of radius $r$ centered at $c = r(\cos(\psi-\pi), \sin(\psi-\pi))$.
-One such chord is drawn in red in Fig.~\ref{fig:regionMove}.
+Two such chords are drawn in red and green in Fig.~\ref{fig:regionMove}.
 

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
Fixed caption for fig. 9

@@
 \begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}\end{overpic}
+\vspace{-1em}
 \caption{\label{fig:regionMove}
@@
 Left top: The possible first contact points for the blue and red particles are shown with blue and red arcs. 
-Left bottom: Once the blue particle touches the wall (blue square) the other particle (red square) can go anywhere in the reachable set (red region).
-Right: The $\Delta$ configuration space for the corresponded starting positions of the particles is shown.
+Left bottom: if the blue particle touches the wall at $\psi_{\min}$ (blue square) the other particle (red square) can move anywhere in the red region. 
+Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the green region. 
+Right: The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
 The possible 2-move reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
-Once the blue particle contacts the boundary as shown, the reachable $\Delta$ configuration is the red set.}
-
+ If the blue particle contacts the boundary at $\psi_{\min}$, the reachable $\Delta$ configuration is the red set, or the green set if $\psi = \frac{17 \pi}{10}$.}
 \end{figure}

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
edited one caption.

@@
 \end{overpic}
-\caption{\label{fig:polygonAlg}{ Steps to generate  the 2-move reachable set when  one particle collides with edge $i,i+1$ of a convex polygonal workspace.}
+\caption{\label{fig:polygonAlg}{ Steps to generate  the 2-move reachable set when  one particle collides with edge $\overline{ p_i p_{i+1}}$ of a convex polygonal workspace.}
 \vspace{-1em}

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
Small equation issues

@@
 Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
- Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a Mathematica implementation in a square workspace for six representative configurations.
+ Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a \emph{Mathematica} implementation in a square workspace for six representative configurations.
  Workspaces are 2D convex polygons with no internal obstacles. 
@@
 %\textcolor{red}{Shiva, this image is nice.  A few suggestions: ?TODO: Label the middle image 'Translate workspace around (0,0)', and draw the workspace outlines with different colors and thicknesses}
-Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides. The 2-move reachable sets are drawn in transparent blue.
+Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides.
+The 2-move reachable sets for this initial configuration (shown by square markers) are drawn in transparent blue.
 }
@@
  \begin{align}\label{eq:psiMinMax}
- \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
+  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\frac{d_{12}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\frac{d_{12}}{2r} \Big],
+  %  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\left(\frac{d_{12}}{2r} \right)- \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\left(\frac{d_{12}}{2r}\right) \Big],
+% \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
 \end{align}

### 2018-05-30 journalWallFriction/experiment.tex
added a reference

--- 2018-05-29 journalWallFriction/experiment.tex
@@
 As shown in the video attachment, this non-uniformity causes the particle closer to the coil to move faster than the other particle.
+This phenomenon makes it easier to increase particle separation than to decrease separation, but boundary collisions can be used to decrease separation.
 Also, magnetic forces are not exactly parallel, but point toward the center of the activated coil.
@@
 Starting and ending positions were printed beneath the workspace on transparency film.
+Our algorithm successfully delivered the particles to goal positions in ??
+out of ??
+trials.
 
@@
 The experiment is shown in Fig.~\ref{fig:story} bottom row.
+Our algorithm successfully delivered the particles to goal positions in ??
+out of ??
+trials.
+A video showing one trial of this experiment is available in the supplementary materials. 
 %In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.

### 2018-05-30 journalWallFriction/theory.tex
added a reference

@@
   - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
-\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
+\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
 \textrm{    ,} \\
@@
  Here  $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary, and
- $\mathbf{N}(\textrm{boundary$(\mathbf{x}_i(t))$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$.
+ $\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$.
 
@@
   There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
-
+ %
 \begin{figure}
@@
 \end{figure}
+%
+
+This optimization result was implemented in our previous work using a square workspace \cite{shahrokhi2017algorithms}.
+ Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a \emph{Mathematica} implementation in a square workspace for six representative configurations.
+\begin{figure*}
+\renewcommand{\figwid}{0.4\columnwidth}
+
+{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
+\put(10,85){a)}
+\put(30,70){$s_1$}
+\put(48,88){$g_1$}
+\put(65,17){$s_2$}
+\put(87,52){$g_2$}
+\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
+\put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
+\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
+%\put(47, 65){$\xleftrightarrow{1.8cm}$}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
+\put(10,85){b)}
+\put(28,70){$s_1$}
+\put(50,80){$g_1$}
+\put(62,15){$s_2$}
+\put(30,22){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
+\put(110,50){Three-move sequence}
+\put(12,38){$m_3$}
+\put(48,52){$m_2$}
+\put(65,30){$m_1$}
+\end{overpic}
+%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
+%\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
+\put(10,85){c)}
+\put(28,70){$s_1$}
+\put(62,28){$g_1$}
+\put(87,50){$s_2$}
+\put(30,22){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+%\vspace{1em}
+{
+\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
+\put(10,85){d)}
+\put(25,30){$s_1$}
+\put(89,70){$g_1$}
+\put(67,20){$s_2$}
+\put(28,75){$g_2$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
+\put(10,85){e)}
+\put(15,35){$s_1,g_2$}
+\put(65,35){$s_2,g_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
+\end{overpic}
+}\\
+
+\vspace{-0.75em}
+{
+\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
+\put(10,85){f)}
+\put(15,10){$s_1,g_2$}
+\put(65,88){$s_2,g_1$}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
+\end{overpic}
+\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
+\put(15, 85){worst-case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
+\end{overpic}
+}\\
+
+%\vspace{-0.75em}
+%{
+%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
+%\end{overpic}
+%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
+%%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
+%\end{overpic}
+%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
+%\end{overpic}
+%}\\
+
+\vspace{-1em}
+
+\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two particle positioning using walls with non-slip contacts. 
+Particle start positions are shown by squares, and goal positions by circles.
+Dashed lines show the shortest route if particles could be controlled independently.
+Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
+%Online dmonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
+%The bottom row shows an extreme case where the robots must switch position.
+}
+\vspace{-1em}
+}
+\end{figure*}
 

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
added a reference

@@
 This section presents algorithms that use non-slip contacts with walls to arbitrarily position two particles in a convex workspace. 
-Our previous work used a square workspace \cite{shahrokhi2017algorithms}.
- Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a \emph{Mathematica} implementation in a square workspace for six representative configurations.
  Workspaces are 2D convex polygons with no internal obstacles. 
@@
 In the last subsection we present techniques to control 3D positioning of two particles.
-
-\begin{figure*}
-\renewcommand{\figwid}{0.4\columnwidth}
-
-{\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
-\put(10,85){a)}
-\put(30,70){$s_1$}
-\put(48,88){$g_1$}
-\put(65,17){$s_2$}
-\put(87,52){$g_2$}
-\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
-\put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
-\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
-%\put(47, 65){$\xleftrightarrow{1.8cm}$}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
-\put(10,85){b)}
-\put(28,70){$s_1$}
-\put(50,80){$g_1$}
-\put(62,15){$s_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves2.pdf}\put(10,10){Move 1}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves3.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story3Moves4.pdf}\put(10,10){Move 3}
-\put(110,50){Three-move sequence}
-\put(12,38){$m_3$}
-\put(48,52){$m_2$}
-\put(65,30){$m_1$}
-\end{overpic}
-%\begin{overpic}[width =\figwid]{s5}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{S6.pdf}\put(50,80){Move 5}
-%\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
-\put(10,85){c)}
-\put(28,70){$s_1$}
-\put(62,28){$g_1$}
-\put(87,50){$s_2$}
-\put(30,22){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move2.pdf}\put(10,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move3.pdf}\put(10,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move4.pdf}\put(10,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story5move5.pdf}\put(10,10){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-%\vspace{1em}
-{
-\begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
-\put(10,85){d)}
-\put(25,30){$s_1$}
-\put(89,70){$g_1$}
-\put(67,20){$s_2$}
-\put(28,75){$g_2$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move5.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move6.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{story7move7.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
-\put(10,85){e)}
-\put(15,35){$s_1,g_2$}
-\put(65,35){$s_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame2.pdf}\put(50,80){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame3.pdf}\put(50,80){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame4.pdf}\put(50,80){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storySame5.pdf}\put(50,80){Move 5}
-\end{overpic}
-}\\
-
-\vspace{-0.75em}
-{
-\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
-\put(10,85){f)}
-\put(15,10){$s_1,g_2$}
-\put(65,88){$s_2,g_1$}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
-\end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
-\put(15, 85){worst-case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
-\end{overpic}
-}\\
-
-%\vspace{-0.75em}
-%{
-%\begin{overpic}[width =\figwid]{storyCorner1.pdf}\put(50,80){Move 1}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner2.pdf}\put(50,80){Move 2}
-%\end{overpic}
-%%\begin{overpic}[width =\figwid]{storyCorner3.pdf}\put(50,80){Move 3}
-%%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner4.pdf}\put(50,80){Move 4}
-%\end{overpic}
-%\begin{overpic}[width =\figwid]{storyCorner5.pdf}\put(50,80){Move 5}
-%\end{overpic}
-%}\\
-
-\vspace{-1em}
-
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two particle positioning using walls with non-slip contacts. 
-Particle start positions are shown by squares, and goal positions by circles.
-Dashed lines show the shortest route if particles could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
-%Online dmonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-\vspace{-1em}
-}
-\end{figure*}
 
@@
 The $\Delta$ configuration space is a set of all possible $\Delta p$ values, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
+We use the $\Delta$ configuration space to plan move sequences that achieve the desired relative spacing.
+Once the particles have the correct relative spacing, they can be delivered to the goal configuration in one move.
 
-The  $\Delta$ configuration space for an $n$-sided convex polygon $P$ can be constructed by translating $n$ copies of $P$  so that each copy moves a different vertex of $P$ to $(0,0)$.
+The  $\Delta$ configuration space for an $n$-sided convex polygon $P$ can be constructed in a method analogous to computing configuration space obstacles for polygons~\cite{lozano1983spatial}. 
+ Translate $n$ copies of $P$  so that each copy moves a different vertex of $P$ to $(0,0)$.
 Because $P$ is convex, the convex-hull of all these translated vertices is the boundary of the  $\Delta$ configuration space.
@@
 
- \emph{Achieving} a $\Delta$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
+ \emph{Achieving} a $\Delta g_c$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
  Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
@@
  \begin{align}\label{eq:psiMinMax}
-  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\frac{d_{12}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\frac{d_{12}}{2r} \Big],
-  %  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\left(\frac{d_{12}}{2r} \right)- \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\left(\frac{d_{12}}{2r}\right) \Big],
+%  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\frac{d_{12}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\frac{d_{12}}{2r} \Big],\\
+\scalebox{.9}{$   \psi \in [\psi_{\min}, \psi_{\max}]= \theta \!+  \left[ \sin^{-1} \left(\frac{d_{12}}{2r}  \right)- \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1} \left(\frac{d_{12}}{2r} \right) \right]\!\!,$}
 % \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
@@
 l_4 =  r \Big(&\cos\psi- \cos( \psi-\gamma(\psi) )\\ \nonumber
-+ &  \sin\psi- \sin( \psi- \gamma(\psi))\Big) &  \psi_{\min}<\psi< \psi_{\max}.\\ \nonumber
++ &  \sin\psi- \sin( \psi- \gamma(\psi))\Big) &  \psi_{\min}<\psi< \psi_{\max}. \nonumber
 \end{align}

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
Changed Figure 7 description.

@@
 \end{figure*}
- Fig. \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
+ Figure \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
+ Figure \ref{fig:TwoRegions} shows reachable sets for a square workspace.
+There are four reachable sets, the horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
  % The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
@@
 %\todo{latex line over the top of two vertices}
-
-Figure \ref{fig:TwoRegions} shows reachable sets for a square workspace.
-There are four reachable sets, the horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
 

### 2018-05-30 journalWallFriction/experiment.tex
Changed Figure 7 description.
cleaned results

--- 2018-05-29 journalWallFriction/experiment.tex
@@
 As shown in the video attachment, this non-uniformity causes the particle closer to the coil to move faster than the other particle.
+This phenomenon makes it easier to increase particle separation than to decrease separation, but this can be compensated because boundary collisions easily decrease the separation.
 Also, magnetic forces are not exactly parallel, but point toward the center of the activated coil.
-Algorithm~\ref{alg:optimalAlg} can still work with  these non-uniformities, but sometimes requires additional iterations.
+Algorithm~\ref{alg:optimalAlg} still works despite these non-uniformities, but sometimes requires additional iterations.
 
@@
 The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
-Fig.~\ref{fig:story} top row shows an experiment.
+Figure \ref{fig:story} top row shows an experiment.
 Starting and ending positions were printed beneath the workspace on transparency film.
+Our algorithm successfully delivered the particles to goal positions in 10 out of 10 trials.
 
@@
 The experiment is shown in Fig.~\ref{fig:story} bottom row.
-Video of this experiment is available in the supplementary materials.
+Our algorithm successfully delivered the particles to goal positions in 5 out of 5 trials.
+A video showing one trial of this experiment is available in the supplementary materials. 
 %In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines.

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
cleaned results

@@
  Figure \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
- Figure \ref{fig:TwoRegions} shows reachable sets for a square workspace.
-There are four reachable sets, the horizontal (and vertical) reachable sets are equivalent in the  $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
+ Figure \ref{fig:TwoRegions} highlights the construction of the 2-move reachable sets for a square workspace.
+There are four 2-move reachable sets, but the horizontal (and vertical) reachable sets are equivalent in the $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
  % The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
@@
 Illustration on how boundary contacts enable 3D positioning.
-Once one particle contacts a boundary, the other particle's 2-move reachable set is a prism formed by extending the 2D 2-move reachable set in the $\pm z$-direction.
+Once one particle contacts a boundary, the other particle's 2-move reachable set is a prism formed by extending the 2D 2-move reachable set in the $\pm z$ direction.
 } %\vspace{-1em}

### 2018-05-30 journalWallFriction/simulation.tex
Added numbers.
changed hits to contacts

@@
 }\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
- The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$.
+ The top row of contour plots show the distance if particle $1$'s goal position is varied in $x$ and $y$.
 The middle row shows the number of moves required for the same configurations.
@@
 The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.
-The worst-case path lengths decrease by ??\%, ??\%, and ??\% for the three test cases shown.
+The worst-case path lengths decrease by 33\%, 60\%, and 30\% for the three test cases shown.
 %TODO: shiva, insert numbers.

### 2018-05-30 journalWallFriction/theory.tex
changed hits to contacts

@@
  If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
-The first move, $m_1$, makes one particle hit a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
-$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
+The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
+$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal contact location, as shown in Fig.
 \ref{fig:reflection}.

### 2018-05-30 journalWallFriction/theory2robAlgs.tex
changed hits to contacts

@@
 Boundary interaction is used to change the relative positions of the particles. Each particle gets the same control input. 
-(left) If particle 2 hits the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
-(middle) Similarly, if particle 2 hits the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
+(left) If particle 2 contacts the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
+(middle) Similarly, if particle 2 contacts the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
 (right) The 2-move reachable sets in the $\Delta$ configuration space is shown.
@@
 \renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(17,-3){\small{Case 1: $s_1$ does not hit boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(12,-3){\small{Case 1: $s_1$ does not contact boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
 \end{overpic}
-\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(25,-3){\small{Case 2: $s_1$ hits boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
+\begin{overpic}[width =\figwid]{anypolygon2.pdf}\put(20,-3){\small{Case 2: $s_1$ contacts boundary}}\put(20,40){$s_2$}\put(55,58){$s_1$}
 \end{overpic}
@@
 The set is constructed by considering each edge of the workspace. We name each vertex as $p_i$ where $1\leq i \leq n$.
-  If one particle hits edge $\overline{ p_i p_{i+1}}$ before the other (one particle will always hit before the other unless the particles are parallel to the wall), the corresponding 2-move reachable set is a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
+  If one particle contacts edge $\overline{ p_i p_{i+1}}$ before the other (one particle will always contact before the other unless the particles are parallel to the wall), the corresponding 2-move reachable set is a polygon, constructed in lines 2-13 of Alg.~\ref{alg:polygonReachbale}.
 The union of these polygons for all $n$ sides is the 2-move reachable set of $\Delta$ configurations.

### 2018-05-30 journalWallFriction/SwarmShapeControl.tex
Small equation issues
previous work!

@@
 \usepackage{multicol}
-\usepackage[bookmarks=true]{hyperref}
+%\usepackage[bookmarks=true]{hyperref}
 
@@
 \usepackage{transparent}
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks =false,
-  urlcolor = black,
-  linkcolor = black
-}
+\usepackage[bookmarks=true,hidelinks]{hyperref}
+
 \usepackage{graphicx}
@@
 However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
-In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
+In previous work, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
 Because in vivo environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 

### 2018-05-31 journalWallFriction/SwarmShapeControl.tex
updated pdf info, moved submitted version to folder.

@@
 \pdfinfo{
-   /Author (Shiva Shahrokhi and Aaron T. Becker)
-   /Title  (Algorithms for Shaping a Particle Swarm with a Shared Input by Exploiting Non-Slip Wall Contacts)
+   /Author (Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker)
+   /Title  (Exploiting Non-Slip Wall Contacts  to Position Two Particles Using The Same Control Input)
    /CreationDate (D:20160129120000)
    /Subject (Simple Robots)
-   /Keywords (Robots;Uniform Control Inputs)
+   /Keywords (Robots;Uniform Control Inputs, Underactuated Robots,Motion Control,Path Planning for Multiple Mobile Robot Systems)
 }
@@
 
-%There are driving applications for large populations of tiny robots in robotics, biology, and chemistry.
-%These robots often lack onboard computation, actuation, and communication.
-%Instead, these ``robots''  are  particles carrying some payload and the particle swarm is controlled by a shared control input such as a uniform magnetic gradient or electric field.
-%In previous works, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.
-%In previous works, we showed that the individual 2D position of two particles in such a swarm is controllable in a rectangular workspace requiring non-slip wall contact.
-However, both in vivo and artificial environments are usually not rectangular.
-%This work extends that work to any polygonal workspace, and focus on extending the algorithm for circular workspaces.
-Then it extends the algorithm to 3D.
-Results are validated with simulations and with a magnetic setup using two steel half a millimeter balls.
-%This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by intestine anatomy.
-%Requiring a small, rigid obstacle suspended in the middle of the workspace is a strong constraint, especially in 3D.
-%Often particles are placed into an em
-%This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contact in 2D.
-%Both in vivo and artificial environments often have such boundaries.
-%We assume that particles in contact with the boundaries have zero velocity if the shared control input pushes the particle into the wall.
-%This paper provides a shortest-path algorithm for positioning a two-particle swarm, and a generalization to positioning an $n$-particle swarm.
-%Results are validated with simulations and a hardware demonstration.
-
-%Consider a swarm of particles controlled by global inputs. 
-%This paper presents algorithms for shaping such swarms in 2D using boundary walls.
-%The range of configurations created by conforming a swarm to a boundary wall is limited. 
-%We describe the set of stable configurations of a particle swarm in two canonical workspaces, a circle and a square. 
-%To increase the diversity of configurations, we add boundary interaction to our model.  
-%We provide algorithms using friction with walls to place two robots at arbitrary locations in a rectangular workspace.
-%Next, we extend this algorithm to place $n$ agents at desired locations. 
-%We conclude with efficient techniques to generate correlations of a particle swarm not possible without wall-friction. 
-%Simulations %and hardware implementations with 100 robots 
-%validate these results.
-
-%These methods may have particular relevance for micro- and nano-robots controlled by global inputs.
-%, whose small size limits onboard computation and power.
-For this reason they are usually powered and controlled by global inputs, such as a uniform external electric or magnetic field, and every robot receives the same control inputs.
-%Due to their small size, large numbers of micro-robots are required to deliver sufficient payloads.
-% Nevertheless, these applications require precision control of the shape and position of the robot swarm. Precision control requires breaking the symmetry caused by the global input.
-
-% KEYWORDS:   uniform control, under-actuation, particle swarm
+% KEYWORDS:   
+%Underactuated Robots
+%Motion Control
+%Path Planning for Multiple Mobile Robot Systems
+%	
+%This is an extension of work presented in a preliminary conference paper:
+%S.
+Shahrokhi, A.
+Mahadev, and A.
+T.
+Becker, ÒAlgorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts,Ó in Intelligent Robots and Systems (IROS), 2017 IEEE/RSJ International Conference on, 2017.
 \end{abstract}

### 2018-06-11 JournalTorqueControl/intro.tex
Shiva PhD dissertation

\section{Introduction}\label{sec:Intro}
Micro-robots can be manufactured in large numbers.
Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 Future implementations require control techniques that can reliably exploit large populations despite significant under-actuation.

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{CoverPhoto.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:FirstImage}
Torque control of an object is essential for manipulating objects to their goal position  when there are narrow passageways and for aligning sensors, emitters, or other objects. 
This paper provides feedback control laws to apply torques and forces using a highly under-actuated system where all 
robots are controlled globally by the same input. 
(a) Simulation of robots exerting torque on a hinged ``door''.
(b) Simulation of swarm orientation manipulation.
(c) 97 hardware robots applying torque to an object. These robots have light sensors and are programmed to move toward the brightest light in the room.  This light is a shared control input that operates globally on the swarm. A full resolution video is available at:~\cite{ShahrokhiTorqueVideo}
}
\vspace{-1em}
\end{figure}

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\cite{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Previous work proves that the mean position of a swarm is controllable and that, with an obstacle, the swarm's position variance orthogonal to rectangular boundary walls  is also controllable
(these are $\sigma_x^2$ and $\sigma_y^2$ for a workspace with axis-aligned walls). 
The usefulness of these techniques was demonstrated by several automatic controllers. One controller steered a swarm of robots to push a larger block through a 2D maze~\cite{ShahrokhiIROS2015}. 
We also showed methods to control a swarm's position covariance to enable the swarm to navigate through a workspaces with narrow corridors.  
However, object manipulation often also requires controlling torque on an object, for example changing the orientation of an object with a large aspect ratio to thread through narrow corridors.
Torque control  is also necessary for a variety of alignment tasks including retroreflectors, and targeted radiation therapy.

Accurate torque control is difficult.
A swarm with $n$ agents has $2n$ degrees of freedom.  
The swarm, when steered toward an object, begins interacting with the object at different times. 
The number of robots touching this object as a function of time is difficult to predict and often impossible to directly measure.
Stochastic effects make long-term prediction challenging.
Even when it is possible to predict which agents will hit the object first, as agents interact with the object, the swarm's configuration changes.
The challenge is not only limited to swarm-object interaction, but also to swarm-swarm interactions when the swarm self-collides  or is split into multiple components. 
As a result, the force the swarm will exert on the object is not easy to predict.
%Predicting the force a swarm exerts on an object requires knowing the location of every robot in contact with the object. 
This information is often hard to gather, because remote sensing of tiny particles is difficult.
Particles may be smaller than the minimum resolution of MRI, PET, and ultrasound, but these sensing modalities can still return aggregate data.
From this aggregate data,  some statistics---including mean and variance---are easy to obtain. 
Rather than design open-loop algorithms, this paper focuses on feedback control strategies using just two statistics of the swarm, the mean and variance of the swarm's position.

Fig. \ref{fig:FirstImage} illustrates the torque control using 97 kilobots, as well as two simulation snapshots of torque and orientation control.

%This paper first discusses the ways that we can control torque of an object in Section ~\ref{sec:theory}. Then it introduces algorithms for torque control in Section ~\ref{sec:simulation}. We show that algorithms work in Section ~\ref{sec:expResults} on 100 kilobots.

%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.

%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles.

% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2018-06-11 RA-L/theory.tex
moved epsilon description into assumptions
Changed the image of the algorithm
Added the full algorithm to the paper
Previous changes
Shiva PhD dissertation

--- 2016-12-09 RA-L/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
-\label{sec:theory}
+\section{Covariance Control}
+\label{sec:covControl}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{fig:friction}
-}
+}\vspace{-1em}
 \end{center}
 \end{figure}
-
-\section{Algorithms}\label{sec:algorithms}
-\subsection{Position Control of Two Robots Using Wall Friction}\label{sec:PostionControl2Robots}
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.4\columnwidth}
-{\begin{overpic}[width =\figwid]{S1_1.png}\put(10,15){$t$ = 0 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_2.png}\put(10,15){$t$ = 0.14 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_3.png}\put(10,15){$t$  = 0.29 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_4.png}\put(10,15){$t$  = 0.43 s}
-\end{overpic}
-\begin{overpic}[width =\figwid]{S1_5.png}\put(10,15){$t$  = 0.76 s}
-\end{overpic}}\\
-
-\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
-The algorithm only requires friction along the bottom and right walls.
-Robot initial positions are shown by a crosshair, and final positions by a circled crosshair.
-Dashed lines show the shortest route if robots could be controlled independently.
-Solid arrows show path given by  Alg.\ \ref{alg:PosControl2Robots}.
-Online demonstration and source code at \citep{Shahrokhi2015mathematicaParticle}.
-%The bottom row shows an extreme case where the robots must switch position.
-}
-%\vspace{-2em}
-}
-\end{figure*}
-
-Alg.~\ref{alg:PosControl2Robots} uses wall-friction to arbitrarily position two robots in a rectangular workspace.
-This algorithm  introduces concepts that will be used for multi-robot positioning.
-Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
-
-Assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
-Denote the current positions of the robots  $r_1$ and $r_2$. 
-Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $s_1.x$ and $s_1.y$ denote the $x$ and $y$ locations of $s_1$. Also the sign function is:
-\begin{align}
-\sign (x) = \begin{cases}  1, &  x > 0 \\
-0, & x=0\\
--1, & x<0 \end{cases} 
-\end{align}
-The algorithm assigns a global control input at every instance.
-The goal is to adjust 
- $\Delta r_x = r_2.x-r_1.x$ from $\Delta s_x = s_2.x-s_1.x$ to $\Delta e_x = e_2.x-e_1.x$ and  adjust 
- $\Delta r_y = r_2.y-r_1.y$ from $\Delta s_y = s_2.y-s_1.y$ to $\Delta e_y = e_2.y-e_1.y$ using a shared global control input. 
- This algorithm exploits the position-dependent friction model \eqref{eq:frictionmodel}.
- %employing the assumption we have made earlier about the walls' friction.
-
-Our algorithm solves the positioning problem in four steps: 
-First, it adjusts $\Delta r_y , \Delta r_x$ as much as possible with the left wall.
-Second, $\Delta r_x -\Delta e_x$ is reduced to zero with the bottom wall.
-Third, if the robots were not correctly positioned relative to each other, $\Delta r_y -\Delta e_y$ is reduced to zero with the right wall.
-%First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
-%Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
-Lastly, the robots, now correctly positioned relative to each other, are moved to their goal locations.
-
-In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ needs all four steps. The worst case path length is $2(\sqrt{2}+1)L$.
-
-\begin{algorithm}
-\caption{GenerateDesiredSpacing($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
-\begin{algorithmic}[1]
-\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
-$(0,0)$ is bottom corner, $s_1$ is leftmost robot, 
- $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
- Assume $s_1.x < s_2.x$ and $s_1.y \geq s_2.y$. If not, rotate workspace coordinates $90^{\circ}$.
-\Ensure $(e_1, e_2) , (s_1, s_2)$ all at least $\epsilon$ distance from walls.
-\State $\epsilon \gets$ robot diameter
-\State $\Delta e_y = e_2.y - e_1.y$
-\State $\Delta e_x = e_2.x - e_1.x$
-\State Move $\left(-r_1.x , -\min \left(r_2.y , r_1.y+ \Delta e_y \right)+\epsilon \right)$ \Comment{Touch left wall}
-%Second Move
-%(-\min(|-r_2.x+ \Delta e_x|, r_2.x),
-\State Move $ (r_2.x, |\sign(\Delta e_y)| \, \min(r_1.y - r_2.y + \Delta e_y-\epsilon,L-r_2.y-\epsilon)+ \epsilon)$
-\Statex \Comment{Adjust  $y$}
-%Third Move
-\State Move $(\max(\epsilon, -\sign(r_2.y - r_1.y)\Delta e_x), -\min(r_1.y , r_2.y))$ 
-\Statex \Comment{Touch bottom wall}
-%Fourth Move
-\State Move $(\sign(r_2.y - r_1.y)\Delta e_x , (|\sign(\Delta e_y)|-1) |r_2.y - r_1.y|)$ \Comment{Adjust $x$}
-\If {$\Delta e_y \neq r_2.y - r_1.y$}
-%Fifth Move
-%TODO This step is not correct.
-\State Move $(L- \max(r_1.x, r_2.x) , \epsilon)$ \Comment{Touch right wall}
-%Sixth Move
-\State Move $(0, -\sign(\Delta e_x)\Delta e_y -\max(r_1.y, r_2.y)- \min(r_1.y, r_2.y)$ 
-\Statex \Comment{Adjust $y$}
-\EndIf
-%Seventh Move
-\State Move $(e_2.x - r_2.x, e_2.y - r_2.y)$ \Comment{Go to goal}
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Position Control of $n$ Robots Using Wall Friction}\label{sec:PostionControlnRobots}
-Alg.
-\ref{alg:PosControl2Robots}  can be extended to control the position of $n$ robots using wall friction under several constraints.
-The solution described here is an iterative procedure with $n$ loops.
-The $k$th loop moves the $k$th robot from a \emph{staging zone} to the desired position in a \emph{build zone}.
-All robots move according to the global input, but due to wall friction, at the end the $k$th loop, robots 1 through $k$ are in their desired final configuration in the build zone, and robots $k+1$ to $n$ are in the staging zone.
-See Fig.~\ref{fig:construction2d} for a schematic of the build and staging zones.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:construction2d}
-Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
-}
-\end{figure}
-
-Assume an open workspace with four axis-aligned walls with infinite friction.
-The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots.
-Without loss of generality, assume the build zone  is above the staging zone.
-Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.
-The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.
-
-The $n$ robot position control algorithm relies on a $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ control input, shown in Fig.\  \ref{fig:driftmove}.
-A drift move consists of repeating a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-The robot touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:driftmove}
-A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
-Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
-\end{figure}
-
-Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot.
-Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.
-
-\begin{algorithm}
-\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
-\begin{algorithmic}[1]
-\State move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom
-
-\While{ $p_{kx} > r$} 
-\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
-\EndWhile
-
-\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
-\State $\beta \gets \frac{f_{ky}-r}{m}$
-\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
-\For{ $m$ iterations}
-\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
-\EndFor
-
-\State Move ($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
-\State Move ($f_{kx}-r, 0$)
-
-\end{algorithmic}
-\end{algorithm}
-
-Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
-First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
-Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
-Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
-Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
-Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.
 

### 2018-06-11 Shiva Dissertation/chap/abs.tex
Shiva PhD dissertation

% abstract for thesis

%This is the abstract.  It has a limit of 150 words for a masters thesis and 350 words for a PhD dissertation.

Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion. 
Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
 Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
 Rather than focusing on a specific microrobotic system, this dissertation designs control laws and algorithms for steering many particles controlled by global fields.

 First, we identify key parameters for particle manipulation by using a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges. Inspired by techniques where human operators performed well, we investigate controllers that only use the mean and variance of the swarm. We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution. Torque control is also necessary for manipulating objects as well as for aligning sensors, emitters, or redirecting an incoming signal.

 Second, this dissertation proves that swarm torque control is possible, then presents algorithms to automate the task. 
 Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).

 Finally, this dissertation investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
 We provide position control algorithms that only require non-slip wall contact in 2D.
 The walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
We assume that particles in contact with the boundaries have zero velocity if the shared control input pushes the particle into the wall. 
All the results are validated with simulations and hardware implementations.

### 2018-06-11 Shiva Dissertation/chap/ack.tex
Shiva PhD dissertation

% acknowledgements for thesis

%This is where you offer thanks to anyone you feel deserves it.

I thank God that I spent some of my best years during the Ph.D. program at the University of Houston. I grew up and became more capable. I would like to take this opportunity to express my gratitude to those who were part of this journey.

First, I would like to express my appreciation to my advisor, Dr. Aaron T. Becker. I truly appreciate his support, kindness, intellectual comments and caring for not only about my research but also many aspects of my life. Dr. Becker was a perfect match for me as an advisor. He is a nice person and full of energy and he makes the lab a happy and energetic place. I thank him for everything he has done for me.

I would like to thank my thesis committee members: Prof. Lydia Kavraki, Dr. David Mayerich, Dr. Rose Faghih, Dr. Stanko Brankovic and Dr. Nicaolas Tsekos for their insightful comments and suggestions and their kindness and supportiveness. I also enjoyed their classes and learned a lot from them.

I was very fortunate to be the first student of swarm control lab. I am thankful to my great collaborators: Lillian Lin, Arun Mahadev, Mable Wan, Chris Ertel, Jingang Shi, and Benedict Isichei. I have always benefited from our discussions. I also would like to thank my other fabulous lab mates: Mary Burbage, Haoran Zhao, Julien Leclerc, Li Huang, Sheryl Manzoor, Mohammad Sultan, An Nguyen, Shriya Bhatnagar, Daniel Bao, Srikanth K.V.S, Parth Joshi, Javier Garcia, Steban Soto, Victor Montano, and Jarret Lonsford. I enjoyed being with all of them.

I would also like to express my sincere appreciation to my parents, Nasrin and Hadi. Their belief in me made me brave enough to go for my dreams. I appreciate them for my whole life. I miss my father who was a great role model for me so much and hope he rests in peace. I believe he is looking at me and hope that I have made him proud. I am very fortunate to have my mother and deeply appreciate her encouragement and self-sacrifice for me. I also like to thank my brothers and my sister, Ali, Amir and Shima, who always backed me up and encouraged me. I appreciate them all for being such a great family.

Finally, I would like to express my most gratitude and appreciation to the love of my life, my husband, Ali. I cannot imagine how I could finish my Ph.D. if he was not there to encourage and support me. Each dark day he was there to show me the tiny light at the end of the road. Each bright day he was there to make my smile bigger and make my heart warmer. He made the path possible for me, and never left me alone in the path. He is smart and hardworking and looking at him motivated me to work smarter and harder. I am forever grateful for each and every day of my life with him.

%I would like to thank my thesis committee members. I appreciate all the insightful comments Prof. Lydia Kavraki made for me. I also thank her for the great class I had with her where I enjoyed so much and had unique experiences. I appreciate all the insightful suggestions Dr. David Mayerich and Dr. Rose Faghih made and their kindness and supportiveness. I appreciate Dr. Stanko Brankovic's insightful suggestions and comments and my other wonderful class I had with him where I enjoyed and learned many things I did not know before.

### 2018-06-11 Shiva Dissertation/chap/conclusion.tex
Shiva PhD dissertation

% Thesis

\chapter[Conclusion]{Conclusion}
\label{chap-conc}

%This is where you draw your conclusions.

  The small size of micro and nano particles makes individual control and autonomy challenging, so currently these particles are steered by global control inputs such as magnetic fields or chemical gradients.

Inspired by the human players in \href{www.swarmcontrol.net}{SwarmControl.net}, this dissertation designed controllers and controllability results using only the mean and variance of a particle swarm. 
We developed a hysteresis-based controller to regulate the position and variance of a swarm. We designed a controller for object manipulation using value iteration for path planning, regions for outlier rejection, and potential fields for minimizing moving the object backwards. 
All automatic controllers were implemented using 100 kilobots steered by the direction of a global light source.
These experiments culminated in an object manipulation task in a workspace with obstacles.
This dissertation also presented techniques for controlling the orientation of an object by manipulating it using a swarm of simple robots with global inputs.
It provided algorithms for precise orientation control, as well as demonstrations of orientation control. 
 This dissertation finally presented techniques for controlling the positions of two particles using uniform inputs and non-slip boundary contacts.  
 It provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D $\Delta$ configuration space.
 Extending Alg.~\ref{alg:optimalAlg}  to 3D was straightforward, but increased the complexity.
 Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue.

 \section{Future Work}
 This dissertation provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.
 Our future goal is to perform assembly using particle swarms to manipulate and attach components. Future efforts should be directed toward examining the effects of Brownian noise, pose control for multiple-part assembly, trajectory prediction, and manipulation in crowded workspaces.
 We assumed friction was sufficient to completely stop particles in contact with the boundary. 
  The algorithms would require retooling to handle small friction coefficients.

  Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network. 
  In out experiments, size of particles were always in millimeter scale or bigger. Future work should consider smaller particles and apply the proposed methods inside a fluid in an MRI machine. That would open up possibilities for non-invasive surgeries and drug delivery.

### 2018-06-11 Shiva Dissertation/chap/intro.tex
Shiva PhD dissertation

% Thesis Introduction

\chapter[Introduction]{Introduction}
\label{chap-intro}

%Here begins the body of the thesis.
\section{Controlling a swarm of particles with global inputs}

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =0.2\figwid]{first.png}\put(3,97){a)}
\end{overpic}
\begin{overpic}[width =0.2\figwid]{second.png}\put(3,95){b)}
\end{overpic}
\begin{overpic}[width =0.25\figwid]{third.png}\put(3,85){c)}
\end{overpic}
\begin{overpic}[width =0.25\figwid]{fourth.png}\put(3,85){d)}
\end{overpic}
\caption{\label{fig:model} %Controlling a swarm of robots with global inputs. 
a) The robots are actually simple particles that can be controlled by a global force. b) All the particles get the same control input. c,d) Breaking the symmetry is only possible with external boundary or obstacle.
}
\end{figure}
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =\figwid]{MicroRobotics.png}
\end{overpic}
\caption{\label{fig:microrobotics} %Controlling a swarm of robots with global inputs. 
Microrobots has the potential to revolutionize medical applications. Figure courtesy:  \cite{nelson2010microrobots}.
}
\end{figure}

Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}. These activities require robots that behave intelligently.
Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel \cite{nelson2010microrobots}. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads.
 Often particles are difficult or impossible to sense individually due to their size and location. 
For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance. 
 Limited computation and communication at small scales also makes autonomous operation or direct control over individual robots difficult. 
  Instead, this dissertation treats the robots that are steered by a global control signal broadcast to the entire population. 
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
 Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}. 
% The transportation methodology is similar to that in~\cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm. Figure \ref{fig:model} shows an example of simple particles when they all get the same control input.
 Techniques are needed that can handle this underactuation.

 To make progress in automatic control with global inputs, this dissertation presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particles' positions. 
We prove that the mean position of a swarm is controllable and that, with an obstacle, the swarm's position variance orthogonal to rectangular boundary walls  is also controllable
(these are $\sigma_x^2$ and $\sigma_y^2$ for a workspace with axis-aligned walls). 
The usefulness of these techniques is demonstrated by several automatic controllers. One controller steers a swarm of robots to push a larger block through a 2D maze~\cite{ShahrokhiIROS2015}.

 Meanwhile, object manipulation often also requires controlling torque on an object for a variety of alignment tasks including retroreflectors, and targeted radiation therapy.
 Accurate torque control is difficult.
 The swarm, when steered toward an object, begins interacting with the object at different times. 
The number of robots touching this object as a function of time is difficult to predict and often impossible to directly measure.
Stochastic effects make long-term prediction challenging.  
Even when it is possible to predict which agents will hit the object first, as agents interact with the object, the swarm's configuration changes.
The challenge is not only limited to swarm-object interaction, but also to swarm-swarm interactions when the swarm self-collides  or is split into multiple components.
 As a result, the force the swarm will exert on the object is hard to predict.
 This dissertation studies the torque applied by a swarm of particles on a long aspect-ratio rod. We calculate the force and torque generated by a swarm that has assumed a characteristic shape as a function of the direction the swarm moves.
We generalize these results for three canonical position distributions of a swarm: uniform, triangular, and normal. The model shows that for a pivoted rod the uniform distribution produces the maximum torque for small swarm standard deviations, but the normal distribution maximizes torque for large standard deviations.

 Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
 However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
 This dissertation relaxes that constraint, and provides position control algorithms for two particles that only require non-slip wall contacts.
 We assume that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.

\section{Dissertation organization}

The dissertation is arranged as follows: Chapter 2 discusses steering a swarm of particles using global inputs and swarm statistics. Chapter 3 goes deep into the torque control of the swarm and orientation control of the object. Chapter 4 demonstrates algorithms to control position of two robots in any convex workspace when the walls have non-slip contact. We conclude in Chapter 5. Figure \ref{fig:chapters} shows a demonstration of each chapter.

\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{MainExpFig.pdf}\end{overpic}
\begin{overpic}[width=0.465\columnwidth]{CoverPhoto.pdf}
\end{overpic}\\

\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
\vspace{3em}
\caption{\label{fig:chapters} 
Top left: Chapter 2 discusses steering a swarm of particles using global inputs and swarm statistics. Top right: Chapter 3 goes deep into the torque control of the swarm and orientation control of the object. Bottom Chapter 4 demonstrates algorithms to control position of two robots in any convex workspace when the walls have non-slip contact. %Bottom right: the magnetic setup that is used to manipulate magnetic particles when they both get the same control input.
}
\end{figure}

### 2018-06-11 Shiva Dissertation/chap2/Introduction.tex
Shiva PhD dissertation

--- 2017-10-26 2017-tro-SteeringSwarmParticles/Introduction.tex
@@
 Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}.
-These activities require robots that behave intelligently.
+%These activities require robots that behave intelligently.
 Limited computation and communication at small scales makes autonomous operation or direct control over individual robots difficult. 
-Instead, this paper treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
-This paper examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
+Instead, this chapter treats the robots as particles that are steered by a global control signal broadcast to the entire population.    
+This chapter examines object manipulation by a swarm of particles, as illustrated in Fig.~{\ref{fig:bigPictureMeanAndVarianceForSwarm}. 
 The transportation methodology is similar to that in~\cite{sugawara2014object}, but rather than using onboard computation or sensing, the particles all move in the same direction.
@@
 Many promising applications for particle swarms require direct human control, but user interfaces to thousands---or millions---of particles is a daunting human-swarm interaction (HSI) challenge.
-Our early work with over a hundred hardware robots and thousands of simulated particles demonstrated that direct human control of large swarms is possible, \cite{Becker2013b}.
-Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
-There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
-One contribution of this paper is a tool for investigating HSI methods through statistically significant numbers of experiments.
+Human control of large swarms is possible over a hundred hardware robots and thousands of simulated particles \cite{Becker2013b}.
+%Unfortunately, the logistical challenges of repeated experiments with over one hundred robots prevented large-scale tests. 
+%There is currently no comprehensive understanding of user interfaces for controlling multi-robot systems with massive populations.  
+%One contribution of this paper is a tool for investigating HSI methods through statistically significant numbers of experiments.
 
@@
 \caption{\label{fig:bigPictureMeanAndVarianceForSwarm} 
-A swarm of particles, all actuated by a uniform control input where each particle gets the same control input, can be effectively manipulated by a control law that uses only the mean and variance of the robot distribution.
-Here a swarm of particles (kilobot robots) pushes a green hexagon toward the goal (see video attachment).
+A swarm of particles, all actuated by a uniform control input %where each particle gets the same control input, 
+can be effectively manipulated by a control law that uses only the mean and variance of the robot distribution.  
+Here a swarm of kilobots pushes a green hexagon toward the goal.
 }
@@
 However, it is often possible to sense global properties of the group, such as mean position and density.
-Often particles are difficult or impossible to sense individually due to their size and location. 
-For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
-To make progress in automatic control with global inputs, this paper presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the particle's positions.
+%Often particles are difficult or impossible to sense individually due to their size and location. 
+%For example, microrobots are smaller than the minimum resolution of a clinical MRI-scanner, see \cite{martel2014computer}, however it is often possible to sense global properties of the group such as mean position and variance.
+ This chapter presents swarm manipulation controllers inspired by our online experiments that require only mean and variance measurements of the positions of the particles. 
  To perform the object manipulation task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}, we use these controllers  as primitives, policy iteration for path planning, handle outliers by partitioning the workspace, and minimize pushing the object backwards with potential field navigation.
@@
  %function and implementation function
-
- Our paper is organized as follows.
-After a discussion of related work in \S \ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in \S \ref{sec:expMethods}.
-Next we prove that the mean and variance of a particle swarm are controllable in \S \ref{sec:theory}, and present automatic controllers in \S \ref{sec:simulation}.
-We use these controllers as primitives and present a framework for manipulating an object through a maze in \S \ref{sec:exp}.
- We implement these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in \S \ref{sec:realExperiment}, and conclude in \S \ref{sec:conclusion}.
+This chapter is an extended version of \cite{shahrokhi2018tro} and is organized as follows.
+After a discussion of related work in Sec.
+\ref{sec:RelatedWork},  we describe our experimental methods for an online human-user experiment and their results in Sec.
+\ref{sec:expMethods}.
+Next we prove that the mean and variance of a particle swarm are controllable in Sec.
+\ref{sec:theory}, and present automatic controllers in Sec.
+\ref{sec:simulation}.
+We use these controllers as primitives and present a framework for manipulating an object through a maze in Sec.
+\ref{sec:exp}.
+ We implement these controllers in our hardware robots and use them to complete an object manipulation task with 100 kilobots in Sec.
+\ref{sec:realExperiment}, and conclude in Sec.
+\ref{sec:conclusion}.

### 2018-06-11 Shiva Dissertation/chap2/conclusion.tex
Shiva PhD dissertation

--- 2017-10-26 2017-tro-SteeringSwarmParticles/conclusion.tex
@@
   The small size of micro and nano particles makes individual control and autonomy challenging, so currently these particles are steered by global control inputs such as magnetic fields or chemical gradients.
-To investigate this control challenge, this paper introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
+To investigate this control challenge, this chapter introduced \href{http://www.swarmcontrol.net}{SwarmControl.net}, an open-source tool for large-scale user experiments where human users steer swarms of robots to accomplish tasks.
 Analysis of the gameplay results revealed benefits of measuring and controlling statistics of the swarm rather than full state feedback, robustness to IID noise, and small effects of varying population size of large swarms.
@@
 
-Our future goal is to perform assembly using particle swarms to manipulate and attach components.
-This task requires applying force and torque to components and manipulating them through obstacles and each other.
-This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.
+Our goal in the next chapter is to apply force and torque to components and manipulating them through obstacles and each other.
+This work provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields.%, but there are many opportunities to extend the work.
 
-Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network.
+%Topics of interest include control with nonuniform flow such as fluid in an artery, gradient control fields like that of an MRI, competitive playing, multi-modal control, flexible workspaces, optimal-control, and targeted drug delivery in a vascular network.
 

### 2018-06-11 Shiva Dissertation/chap2/experiment.tex
Shiva PhD dissertation

--- 2017-11-15 journalSwarmControl/experiment.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Particle Swarm Object manipulation}\label{sec:exp}
+\section{Particle swarm object manipulation}\label{sec:exp}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
-%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
-We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
-%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
- To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.
-
-The table below summarizes the simulation results for each 10 successful trials:
+%\vspace{10em}
+\begin{table}[h!]
   \begin{center}
+  %\vspace{-1em}
+    \caption{\label{tbl:resultsSum}Summary, Object Manipulation Results}
 \begin{tabular}{ c c }
 \hline
-Method & Result, mean$\pm$std (s)\\ [0.5ex] 
+Method & Result (10 trials), mean$\pm$std (s)\\ [0.5ex] 
 \hline \hline
@@
 \end{center}
+%\vspace{-1em}
+\end{table}
+
+This section analyzes an \emph{object manipulation} task attempted by our hybrid, hysteresis-based controllers. The swarm must deliver the object to the goal region. 
+%We assume that particles pushing the object move slower than unencumbered particles, and that pushing with more particles increases the object speed
+We assume Coulomb and viscous friction parameters such that the object can be moved by particle motion.  Increasing the number of pushing particles increases the object speed.
+%Although one particle might be able to move the object, it requires a long time for it to complete the task and probability of getting stuck when there is not enough particles is high.
+ To solve this object manipulation task we divide the task into three components: 1) designing a policy for the object, 2) pushing the object with a compliant swarm, and 3) managing outliers.
+Table \ref{tbl:resultsSum} summarizes results for successful simulation trials.
+
 \subsection{Learning a policy for the object}\label{subsec:objectpolicy}
@@
 \centering
-\renewcommand{\figwid}{3cm}
-\begin{overpic}[height=\figwid]{GradientView.png}
+%\renewcommand{\figwid}{5cm}
+\begin{overpic}[width=0.33\columnwidth]{GradientView.png}
 \end{overpic}
-\begin{overpic}[height=\figwid]{PolicyIter.png}
+\begin{overpic}[width=0.33\columnwidth]{PolicyIter.png}
 \end{overpic}
-\begin{overpic}[height=\figwid]{MDPmap.pdf}
+\begin{overpic}[width=0.255\columnwidth]{MDPmap.pdf}
 \end{overpic}
@@
 (Right) The vision algorithm detects obstacles in the hardware setup.
-This map is used to produce the value function and control policy shown.
+%This map is used to produce the value function and control policy shown.
 %\vspace{-2em}
@@
 \centering
-\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{DifferentShapes.pdf}\end{overpic}
 \vspace{-1em}
@@
 \centering
-\renewcommand{\figwid}{0.5\columnwidth}
+\renewcommand{\figwid}{0.49\columnwidth}
 \begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
@@
 \begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
-\end{overpic}
+\end{overpic}\\
+\vspace{2em}
 \begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
@@
 \end{overpic}
-\vspace{-0.5em}
+%\vspace{-0.5em}
 \caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.

### 2018-06-11 Shiva Dissertation/chap2/realExperiment.tex
Shiva PhD dissertation

--- 2017-11-03 journalSwarmControl/realExperiment.tex
@@
 \begin{figure*}
-\renewcommand{\figwid}{5.25cm}
+%\renewcommand{\figwid}{5.25cm}
 \begin{center}
 	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
-	\begin{overpic}[height=\figwid]{SetUp.pdf}%\put(1,75){A}
-	\end{overpic}\begin{overpic}[height=5.5cm]{AllShapesR.pdf}%\put(-0,85){B}
-	\end{overpic}\begin{overpic}[height=\figwid]{PotentialExp.pdf}%\put(-0,85){C}
+	\begin{overpic}[width=0.7\columnwidth]{SetUp.pdf}%\put(1,75){A}
+	\end{overpic}\\
+	\vspace{3em}
+	\begin{overpic}[width=0.33\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
+	\end{overpic}
+	\begin{overpic}[width=0.6\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
 	\end{overpic}
@@
 The moveable object was placed in the lower center of the workspace.
-{\sc Matlab} code for vision processing, the value iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
+{\sc Matlab} code for vision processing, the value iteration of Sec.
+\ref{subsec:objectpolicy} and the algorithm of Sec.
+\ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
 Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

### 2018-06-11 Shiva Dissertation/chap2/relatedWork.tex
Shiva PhD dissertation

--- 2017-10-03 journalSwarmControl/relatedWork.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:RelatedWork}
+\section{Related work}\label{sec:RelatedWork}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This paper investigates global control of particles that have no onboard computation.
+This dissertation investigates global control of particles that have no onboard computation.
 This prevents us from applying controllers that require computation on the agents, as in \cite{milutinovic2006modeling,prorok2011multi,demir2015probabilistic}.
@@
 Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};  magnetic structures  with different cross-sections that can be independently steered by \cite{Floyd2011,Diller2013};  \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
-\cite{Tottori2012} and \cite{Peyer2013}. 
-Similarly, our previous work, \cite{Becker2012,Becker2012k}, focused on exploiting inhomogeneity between robots.
+\cite{Tottori2012} and \cite{Peyer2013}, similarly, exploiting inhomogeneity between robots~\cite{Becker2012,Becker2012k}.
 These control algorithms theoretically apply to any number of robots, even robotic continuums.
@@
  We desire control algorithms that extend to many thousands of robots.
-Limited position control was achieved by \cite{bretl2007control} and our previous work \cite{beckerIJRR2014}, but both used robots commanded in their local coordinate frame.
-Our new submission focuses on a more common paradigm: particles commanded in a global coordinate frame.
+Limited position control was achieved by \cite{bretl2007control} and \cite{beckerIJRR2014}, but both used robots commanded in their local coordinate frame.
+This chapter focuses on a more common paradigm: particles commanded in a global coordinate frame.
 
@@
 \paragraph{Computation---calculating the control law}
-In our previous work the controllers required at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
+There are controllers that require at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
 These operations become intractable for large populations of robots.
@@
   Most humans are able to, with practice, steer a swarm of robots controlled by a global input.
-Prior to our paper, no algorithm existed.
+Prior to this work, no algorithm existed.
 Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.
-This paper exploits insights gained from \href{http://www.swarmcontrol.net}{\emph{SwarmControl.net}}, particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
+This chapter exploits insights gained from \href{http://www.swarmcontrol.net}{\emph{SwarmControl.net}}, particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.
 
  A user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism is designed in \cite{ayanian2014controlling}.
-Our paper shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.
+This chapter shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.
 
@@
 They hypothesized the location of this plateau is highly dependent on the underlying task.
-Indeed, our paper indicates there are tasks without plateaus.
+Indeed, this chapter indicates there are tasks without plateaus.
 % (Lemmings is an example of this as well, for systems with autonomy
@@
 Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in~\cite{odhner2014compliant} or silicone in~\cite{deimel2014novel}.
-Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in \S \ref{sec:VarianceControl}.
+Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in Sec.
+\ref{sec:VarianceControl}.
 
-\subsection{Relationship to authors' prior work}
- This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper.
+%\subsection{Relationship to authors' prior work}
+% This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper.
 One paper covered the first three months of SwarmControl.net experiments~\cite{swarmcontrol2013}, and the second presented simulations of object manipulation~\cite{ShahrokhiIROS2015}.
-This paper presents three years of results from SwarmControl.net.
-For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parameter sweeps over key variables.
-All hardware validation experiments are new.
 

### 2018-06-11 Shiva Dissertation/chap3/Previous/conclusion.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This chapter presented techniques for controlling the orientation of an object by manipulating it using a swarm of simple robots with global inputs.
The chapter provided algorithms for precise orientation control, as well as demonstrations of orientation control.

%Future efforts should be directed toward optimizing torque control, examining the effects of Brownian noise, applying the techniques to hardware robots, pose control for multiple part assembly, and manipulation in a crowded workspace.
The control laws in this chapter used only the mean and variance of the swarm.  The control techniques may be optimized using high-order moments, or by stochastic modeling of the collisions between swarm members and the object.

### 2018-06-11 Shiva Dissertation/chap3/Previous/experiment.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiment}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Hardware System}

Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.  These allow us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, from \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open source platform or commercially from~\cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments, as shown in Fig.~\ref{fig:setup}, we used $n$=97 kilobots, a glass-covered 1.5 m$\times$1.2 m whiteboard as the workspace, and eight 30W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,NE,E,SE,S,SW,W,NW\}$ vertices of a 6 m square centered on the workspace. The lights were controlled using an Arduino Uno board connected to an 8-relay shield.  Above  the table, an overhead machine vision system tracks the position of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=.8\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:setup}
Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system.
}
\vspace{-1.5em}
\end{figure}

The experiments from Section \ref{sec:simulation}.a were manually demonstrated using this physical swarm.
Fig. \ref{fig:expBig} illustrates an experiment showing pure torque control with a swarm of robots. In this figure a large-aspect-ratio rectangle  (91$\times$2 cm, colored pink in the image) was hinged to one side of the table.  Like a door, this object could only be moved around this pivot. 
Two trials were performed.  In each trial the swarm was initialized in the lower right side of the table, and then commanded to push the object with the mean position of the swarm directed toward a point distance $C$ from the pivot point. Data was recorded for 150 seconds.
In the first trial, $C = L$, so the robots were commanded to push the door at the extreme edge of the door from the pivot.  
In the second trial $C = 1/2 L$, and so the swarm pushed the object in the center of the rectangle.
As discussed in Section \ref{sec:simulation}, the robots spread when commanded to push the object at the extreme end, and half of the robots flowed past the end of the rectangle without engaging the rectangle.
 This illustrates a key difference between robotic swarms and a single pusher robot. The swarm exerts the most torque when  \eqref{eq:swarmtorque} is maximized.
  \eqref{eq:swarmtorque} is maximized when the majority of the swarm engages the object.
For this reason in Fig. \ref{fig:expBig}, the trial in the second row of screenshots moves the door further than the  trial in the  first row.
\begin{figure*}
\centering

\includegraphics[width=\columnwidth]{Experiment.png}
\vspace{-1em}
\caption{\label{fig:expBig}{Snapshots showing the effect of pushing a pivoted rectangular object at different distances from the pivot point. 
 97 robots were programmed to move toward the brightest light in the room, and   controlled by choosing which of 8 lights were on at any given instant. The top row of snapshots illustrate the swarm pushing at the end of the object.  In this case,  the swarm flows past the object and the force decreases. The bottom row illustrates that when the swarm pushes at the middle of the object the force provided by the swarm remains constant. In this case the swarm does not flow past the object. See video attachment for recordings of these experiments at\href{https://www.youtube.com/watch?v=ZiKEa5UwukM&feature=youtu.be}{\cite{ShahrokhiTorqueVideo}}.
 }
%\vspace{-2em}
}
\end{figure*}

### 2018-06-11 Shiva Dissertation/chap3/Previous/intro.tex
Shiva PhD dissertation

--- 2018-06-11 JournalTorqueControl/intro.tex
@@
 \section{Introduction}\label{sec:Intro}
-Micro-robots can be manufactured in large numbers.
-Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
- For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
+%Micro-robots can be manufactured in large numbers.
+%Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
+% For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
 Instead, these robots are often controlled by a global, broadcast signal.
- Future implementations require control techniques that can reliably exploit large populations despite significant under-actuation.
+% Future implementations require control techniques that can reliably exploit large populations despite significant under-actuation.
 
@@
 This light is a shared control input that operates globally on the swarm.
-A full resolution video is available at:~\cite{ShahrokhiTorqueVideo}
+%A full resolution video is available at:~\cite{ShahrokhiTorqueVideo}
 }
@@
 Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze.
-Previous work proves that the mean position of a swarm is controllable and that, with an obstacle, the swarm's position variance orthogonal to rectangular boundary walls  is also controllable
+Previous chapter proves that the mean position of a swarm is controllable and that, with an obstacle, the swarm's position variance orthogonal to rectangular boundary walls  is also controllable
 (these are $\sigma_x^2$ and $\sigma_y^2$ for a workspace with axis-aligned walls).
-The usefulness of these techniques was demonstrated by several automatic controllers. One controller steered a swarm of robots to push a larger block through a 2D maze~\cite{ShahrokhiIROS2015}. 
-We also showed methods to control a swarm's position covariance to enable the swarm to navigate through a workspaces with narrow corridors.  
+
 However, object manipulation often also requires controlling torque on an object, for example changing the orientation of an object with a large aspect ratio to thread through narrow corridors.
@@
 
-Accurate torque control is difficult.
-A swarm with $n$ agents has $2n$ degrees of freedom.  
-The swarm, when steered toward an object, begins interacting with the object at different times. 
-The number of robots touching this object as a function of time is difficult to predict and often impossible to directly measure.
-Stochastic effects make long-term prediction challenging.
-Even when it is possible to predict which agents will hit the object first, as agents interact with the object, the swarm's configuration changes.
-The challenge is not only limited to swarm-object interaction, but also to swarm-swarm interactions when the swarm self-collides  or is split into multiple components. 
-As a result, the force the swarm will exert on the object is not easy to predict.
-%Predicting the force a swarm exerts on an object requires knowing the location of every robot in contact with the object. 
-This information is often hard to gather, because remote sensing of tiny particles is difficult.
-Particles may be smaller than the minimum resolution of MRI, PET, and ultrasound, but these sensing modalities can still return aggregate data.
-From this aggregate data,  some statistics---including mean and variance---are easy to obtain. 
-Rather than design open-loop algorithms, this paper focuses on feedback control strategies using just two statistics of the swarm, the mean and variance of the swarm's position.
+With a single agent, torque control is straightforward: the agent simply maximizes the length of the movement arm to maximize torque.
+To make an agent push open a door, it should push on the edge furthest from the hinge.
+The optimal solution for a swarm of particles is not straightforward because they cannot all push at one position.
 
+This paper focuses on optimizing control strategies using the swarm's position distribution. 
 Fig. \ref{fig:FirstImage} illustrates the torque control using 97 kilobots, as well as two simulation snapshots of torque and orientation control.
+This chapter first discusses the ways that we can control torque of an object in Section ~\ref{sec:theory}.
+Then it introduces algorithms for torque control in Section ~\ref{sec:simulation}.
+We show that algorithms work in Section ~\ref{sec:expResults} on up to 100 kilobots.
+%Representative results are shown in Fig. \ref{fig:FirstImage}.
+
+%Accurate torque control is difficult.
+%A swarm with $n$ agents has $2n$ degrees of freedom.  
+%The swarm, when steered toward an object, begins interacting with the object at different times. 
+%The number of robots touching this object as a function of time is difficult to predict and often impossible to directly measure.
+%Stochastic effects make long-term prediction challenging.
+%Even when it is possible to predict which agents will hit the object first, as agents interact with the object, the swarm's configuration changes.
+%The challenge is not only limited to swarm-object interaction, but also to swarm-swarm interactions when the swarm self-collides  or is split into multiple components. 
+%As a result, the force the swarm will exert on the object is not easy to predict.
+%%Predicting the force a swarm exerts on an object requires knowing the location of every robot in contact with the object. 
+%This information is often hard to gather, because remote sensing of tiny particles is difficult.
+%Particles may be smaller than the minimum resolution of MRI, PET, and ultrasound, but these sensing modalities can still return aggregate data.
+%From this aggregate data,  some statistics---including mean and variance---are easy to obtain. 
+%Rather than design open-loop algorithms, this paper focuses on feedback control strategies using just two statistics of the swarm, the mean and variance of the swarm's position.
 

### 2018-06-11 Shiva Dissertation/chap3/Previous/relatedWork.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:RelatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object~\cite{Sudsang2002,Fink2007}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object.

Robotic manipulation by pushing has a long and successful history\cite{Lynch1999,Lynch1996,Akella2000,Bernheisel2006}.  Key developments introduced the notion of stable pushes and a friction cone.  A \emph{stable push} is a pushing operation by a robot with a flat-plate pushing element in which the object does not change orientation relative to the pushing robot\cite{Lynch1999}.  The \emph{friction cone} is the set of vector directions a robot in contact with an object can push that object with a stable push.
Stable pushes can be used as primitives in an rapidly-expanding random tree to form motion plans.
A key difference is that our robots are compliant and tend to flow around the object, making this similar to fluidic trapping~\cite{Armani2006,Becker2009}.  
%\subsection{Block-pushing and Compliant Manipulation}

While ferrous particles tend to clump in a magnetic field, the magnetotactic bacteria of~\cite{martel2015magnetotactic,ou2012motion} are directed by the orientation of the magnetic field and do not suffer from magnetic aggregation.

Controlling the \emph{shape}, or relative positions, of a swarm of robots is a key ability for a myriad of applications.  Correspondingly, it has been studied from a control-theoretic perspective in  both centralized, e.g. virtual leaders in \cite{egerstedt2001formation}, and decentralized approaches, e.g. decentralized control-Lyapunov function controllers in~\cite{hsieh2008decentralized}. Most approaches assume a level of intelligence and autonomy in the individual robots that exceeds the capabilities of current micro- and nano-robots~\cite{martel2015magnetotactic,Xiaohui2015magnetiteMicroswimmers}.
Instead, this chapter focuses on centralized techniques that apply the same control input to each member of the swarm, as in~\cite{Becker2013b}.

### 2018-06-11 Shiva Dissertation/chap3/covernote.tex
Shiva PhD dissertation

Dear Editors,
 We are submitting our manuscript "Maximizing Torque on a Rigid Body by Pushing, Using Many Particles Steered by Shared Control Inputs", to be considered for publishing in IEEE Transaction on Automation Science and Engineering  (T-ASE). Authors are Shiva Shahrokhi, Lillian Lin, and Aaron T. Becker.

This paper springboards from the content of a preliminary conference paper, extending its substance and providing full details in a single journal paper. This paper is in IEEE International Conference on Automation Science and Engineering  (CASE), 2016, pp. 561--566. That paper covered techniques for orientation control of an object pushed by a swarm of particles, when the particles all have the same shared control input.
This paper provides analytical models and optimal solutions for three common swarm distributions, and new hardware experiments.

The primary Methodology is 3. Materials/Parts Handling and secondary Methodology is 9. Theoretical Foundations.
The primary Application is I. Construction and secondary Application is C. Healthcare and Life sciences.

We believe that the topic is a perfect fit for the scope of your journal and are hopeful that editors and reviewers share the opinion that this submission will be of interest to the readers of IEEE Transaction on Automation Science and Engineering.

With Best Regards,
Shiva Shahrokhi (on behalf of all authors)
713 743 6671, sshahrokhi2@uh.edu.
PhD Candidate, Electrical and Computer Engineering, University of Houston

### 2018-06-11 Shiva Dissertation/chap3/experiment.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
\begin{center}
	\includegraphics[width=.8\columnwidth]{SetUp.pdf}
\end{center}
%\vspace{-1em}
\caption{\label{fig:setup}
Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered LED floodlights, with an overhead machine vision system.
}
%\vspace{-1.5em}
\end{figure}

%The experiments from Section \ref{sec:simulation}.a were manually demonstrated using this physical swarm.
%Fig. \ref{fig:expBig} illustrates an experiment showing pure torque control with a swarm of robots. In this figure a large-aspect-ratio rectangle  (91$\times$2 cm, colored pink in the image) was hinged to one side of the table.  Like a door, this object could only be moved around this pivot. 
%Two trials were performed.  In each trial the swarm was initialized in the lower right side of the table, and then commanded to push the object with the mean position of the swarm directed toward a point distance $C$ from the pivot point. Data was recorded for 150 seconds.
%In the first trial, $C = L$, so the robots were commanded to push the door at the extreme edge of the door from the pivot.  
%In the second trial $C = 1/2 L$, and so the swarm pushed the object in the center of the rectangle.
%As discussed in Section \ref{sec:simulation}, the robots spread when commanded to push the object at the extreme end, and half of the robots flowed past the end of the rectangle without engaging the rectangle.
% This illustrates a key difference between robotic swarms and a single pusher robot. The swarm exerts the most torque when  \eqref{eq:swarmtorque} is maximized.
%  \eqref{eq:swarmtorque} is maximized when the majority of the swarm engages the object.
%For this reason in Fig. \ref{fig:expBig}, the trial in the second row of screenshots moves the door further than the  trial in the  first row.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{ComparisonAllPivoted.eps}
\end{center}
%\vspace{-1em}
\caption{\label{fig:PivotedKilobot}
Pivoted object using uniformly distributed Kilobots with $\sigma = 0.05 L = \SI{6.8}{\centi \metre}$, but different mean positions. Mean positions are normalized along a rod of length 1.42 m.
}
%\vspace{-1.5em}
\end{figure}

The previous sections assumed perfect transmission of force from each particle to the object. 
Accurate swarm torque control is difficult.
In 2D, a swarm of $n$ particles has $2n$ degrees of freedom.  
The swarm, when steered toward an object, begins interacting with the object at different times. 
The number of particles touching this object as a function of time is difficult to predict and often impossible to directly measure.
Stochastic effects make long-term prediction challenging.
Even when it is possible to predict which particles will hit the object first, as particles interact with the object, the swarm's configuration changes.
The challenge is not only limited to swarm-object interaction, but also to swarm-swarm interactions when the swarm self-collides  or is split into multiple components. 
As a result, the force the swarm will exert on the object is not easy to predict. 
%Predicting the force a swarm exerts on an object requires knowing the location of every robot in contact with the object. 
To demonstrate the analytical results from Section \ref{sec:torqueDist} hold, we performed experiments using centimeter-scale hardware robots called \emph{Kilobots}.

%\subsection{Hardware System}

%Our experiments are on 
  These allow us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The Kilobot, from \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available in an open source platform or commercially from~\cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.

%3 robots were donated to another lab during this test.
These experiments used up to $n = 97$ Kilobots, a glass-covered 1.5 m$\times$1.2 m whiteboard as the workspace, and 30 W and 50 W LED floodlights arranged 1.5 m above the plane of the table at the midpoint of the east side of a 6 m square centered on the workspace, as shown in Fig.~\ref{fig:setup}. Above  the table, an overhead machine vision system tracks the position of the swarm. For the last two experiments, we designed the object to rigidly hold the swarm, and programmed the robots to go straight.

\subsection{Pushing a pivoted object}

Fig. \ref{fig:expBig} shows snapshots of the experiment for a pivoted object at different mean positions for the swarm when they are uniformly distributed. The object was designed such that the Kilobots keep the starting uniform distribution with $\sigma = 0.05 L = \SI{6.8}{\centi \metre}$. Kilobots were programmed to go straight so that they always apply force perpendicular to the object. The results of different mean positions are shown in Fig.~\ref{fig:PivotedKilobot} for four trials at $\mu=[0.5, 0.66, 0.8]$. Mean positions near the optimal solution of \eqref{eq:maxMuPivot} increase torque.

The experiments from Section \ref{sec:simulation}.a were manually demonstrated using this physical swarm.
Fig. \ref{fig:expBig} illustrates an experiment showing pure torque control with a swarm of robots. In this figure a large-aspect-ratio rectangle  (91$\times$2 cm, colored pink in the image) was hinged to one side of the table.  Like a door, this object could only be moved around this pivot. 
Two trials were performed.  In each trial the swarm was initialized in the lower right side of the table, and then commanded to push the object with the mean position of the swarm directed toward a point distance $C$ from the pivot point. Data was recorded for 150 seconds.
In the first trial, $C = L$, so the robots were commanded to push the door at the extreme edge of the door from the pivot.  
In the second trial $C = 1/2 L$, and so the swarm pushed the object in the center of the rectangle.
As discussed in Section \ref{sec:simulation}, the robots spread when commanded to push the object at the extreme end, and half of the robots flowed past the end of the rectangle without engaging the rectangle.
 This illustrates a key difference between robotic swarms and a single pusher robot. The swarm exerts the most torque when  \eqref{eq:swarmtorque} is maximized.
  \eqref{eq:swarmtorque} is maximized when the majority of the swarm engages the object.
For this reason in Fig. \ref{fig:expBigNormal}, the trial in the second row of screenshots moves the door further than the  trial in the  first row.

\begin{figure*}
\centering
\renewcommand{\figwid}{0.24\columnwidth}
\begin{overpic}[width =\figwid]{PivotedEnd1of4.pdf}%\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedEnd2of4.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedEnd3of4.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedEnd4of4.pdf}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{PivotedMiddle1of4.pdf}%\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedMiddle2of4.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedMiddle3of4.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{PivotedMiddle4of4.pdf}
\end{overpic}
%\vspace{-1em}
\caption{\label{fig:expBig}{Snapshots showing the effect of pushing a pivoted rectangular object at different distances from the pivot point. 
%56 robots were programmed to move straight. The object is designed to maintain the initial distribution of the robots. 
%The top row of snapshots illustrate the swarm pushing at the end of the object.
 The top row shows the frames with the mean position is at 0.75 of the object length. The bottom row show frames with swarm mean position $\mu = 1/2$. %See attachment for experimental videos.
%\textcolor{red}{??Why are there 4 robots per row in the pictures but 3 robots per row in the closeup? }
 }
%\vspace{-2em}
}
\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\columnwidth]{Experiment.pdf}
\vspace{-1em}
\caption{\label{fig:expBigNormal}{Snapshots showing the effect of pushing a pivoted rectangular object at different distances from the pivot point. 
% 97 robots were programmed to move toward the brightest light in the room, and controlled by choosing which of 8 lights were on at any given instant.
 The top row of snapshots illustrate the swarm pushing at the end of the object.  %In this case,  the swarm flows past the object and the force decreases.
 The bottom row illustrates that when the swarm pushes at the middle of the object the force provided by the swarm remains constant.% In this case the swarm does not flow past the object. %See video attachment for recordings of these experiments~{\cite{ShahrokhiTorqueVideo}}.
%\textcolor{red}{??Can you get higher resolution? }
 }
%\vspace{-2em}
}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation control of a free object}

\begin{figure*}
\centering
\begin{overpic}[width =0.32\columnwidth]{orient1.png}\put(70,5){\emph{t} = 0 s}
\end{overpic}
\begin{overpic}[width =0.32\columnwidth]{orient3.png}\put(65,5){\emph{t} = 10 s}
\end{overpic}
\begin{overpic}[width =0.32\columnwidth]{orient2.png}\put(65,5){\emph{t} = 60 s}
\end{overpic}\\
\vspace{1em}
\begin{overpic}[width =0.32\columnwidth]{orient4.png}\put(65,5){\emph{t} = 90 s}
\end{overpic}
\begin{overpic}[width =0.32\columnwidth]{orient5.png}\put(65,5){\emph{t} = 120 s}
\end{overpic}
\begin{overpic}[width =0.32\columnwidth]{orient6.png}\put(65,5){\emph{t} = 150 s}
\end{overpic}
\caption{\label{fig:expOrient}{Snapshots showing orientation control of a free object using 97 hardware robots that all get the same control input. Light direction is the global control input and the robots are programmed to move toward the brightest light in the environment.
 }
%\vspace{-2em}
}
\end{figure*}
Figure \ref{fig:expOrient} shows snapshots of orientation control of a free rod using 97 Kilobots. Kilobots were programmed to move toward the brightest light in the room. Therefore, the light is the global input. The goal orientation of the object changed periodically, and the robots will go to a corner when the goal orientation is achieved.

%\textcolor{Red}{TODO?? insert content from orientation control experiment, number of robots, timing, show a series of images, etc.}

### 2018-06-11 Shiva Dissertation/chap3/intro.tex
Shiva PhD dissertation

\section{Introduction}\label{sec:Intro}

This chapter investigates maximizing torque applied by a large number of particles, hereafter called a \emph{swarm}, when the swarm has non-slip contact with a rigid, 2D body. 
 The under-actuated swarm is steered by a shared signal that consists of a vector direction for movement. 
  The robotic system is comprised of the swarm of particles, the shared control signal, and an external sensor that measures the swarm position.
   This chapter examines analytically two representative aspects of swarm torque control: first, pushing a pivoted rod, and second pushing a free body. 
   We conclude with hardware experiments with centimeter-scale robots. Maximizing torque improves the efficiency of a particle swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{CoverPhoto.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:FirstImage}
%Torque control of an object is essential for manipulation unless objects are homogeneous discs, especially when there are narrow passageways or the objects must be aligned, e.g. sensors and emitters. 
%This paper provides the optimal position for a swarm to push to maximize torque production using a highly under-actuated system where all robots are controlled uniformly by the same input. 
(a) Simulation of robots exerting torque on a hinged ``door".
(b) Orientation control of a free long rod.
(c) Hardware robots applying torque to an object. See video attachment.% A full resolution video is available at \url{https://youtu.be/7Q5lu_ZFbxI}.
}
\vspace{-1em}
\end{figure}

With a single agent, torque control is straightforward: the agent simply maximizes the length of the movement arm to maximize torque. To make an agent push open a door, it should push on the edge furthest from the hinge. 
The optimal solution for a swarm of particles is not straightforward because they cannot all push at one position.

This chapter focuses on maximizing torque using the swarm's position distribution. 
Representative results are shown in Fig. \ref{fig:FirstImage}:  torque control using 56 mobile robots on a pivoted and a free object.

### 2018-06-11 Shiva Dissertation/chap3/simulation.tex
Shiva PhD dissertation

--- 2016-11-18 JournalTorqueControl/simulation.tex
@@
 \section{Simulation}\label{sec:simulation}
-This section examines four main challenges for pose and torque control of an object, arranged in increasing difficulty.
+This section examines three challenges for torque control of an object, arranged in increasing difficulty.
 Each task uses a PD controller that regulates the swarm's mean position,  as in \cite{ShahrokhiIROS2015}.
@@
 
-\paragraph{Pure torque control} \label{sec:simulationPureTorque}
+\paragraph{Maximizing torque on a pivoted body} \label{sec:simulationPureTorque}
 An object with a pivot point can rotate, but not translate. A door is an example.
  If there was only one robot touching the object, the robot should push at the point which maximizes the moment arm, at the extreme end of the object furthest from the pivot point.
-The optimal pushing location provides the maximum force, because it maximizes  $r$ in \eqref{eq:torque}.
+% \begin{align}
+%\tau = F \times (P - O )\label{eq:torque}
+% = F \cdot (P - O )\label{eq:torque}
+%\end{align}
+The optimal pushing location provides the maximum force, because it maximizes the distance $r = \norm{P - O}$ in \eqref{eq:torque}.
 However, given a swarm of robots, maximizing $r$ is no longer the optimal solution.  
@@
 \end{align}
+%\textcolor{red}{??The coordinate frames here are different from the rest of the paper}
 $O_{\theta}$ = the orientation of the object's major axis, measured from the world $x$-axis.
@@
 
-\begin{figure}
-\begin{center}
-	\includegraphics[width=0.8\columnwidth]{LFig.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:LFig}
-Simulation results from a swarm applying force to a hinged door. 
-The swarm mean is steered toward a point $C$ units along the object from the pivot point. The red dashed line indicates the times that the swarm was in variance control mode.
- Simulation used 144 robots of diameter $0.2$ m with a standard deviation of less than $1.5$ m and an object length of $6$ m.
-}
-\vspace{-1em}
-\end{figure}
+%\textcolor{red}{??TODO: PUT THE NEW REAL SIMULATIONS HERE
+%SIm 1: include  push at the optimal location
+%}
+%
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{LFig.pdf}
+%\end{center}
+%\vspace{-1em}
+%\caption{\label{fig:LFig}
+%Simulation results from a swarm applying force to a hinged door. 
+%The swarm mean is steered toward a point $C$ units along the object from the pivot point. The red dashed line indicates the times that the swarm was in variance control mode.
+% Simulation used 144 robots of diameter $0.2$ m with a standard deviation of less than $1.5$ m and an object length of $6$ m.
+%}
+%\vspace{-1em}
+%\end{figure}
 
@@
  \begin{align}\nonumber
-goal_x = O_x +  K_{orient}  (O_{\theta} - goal_\theta ) \cos(O_{\theta}) \\
-goal_y = O_y +  K_{orient}  ( O_{\theta} -goal_\theta  ) \sin(O_{\theta})
+goal_x = O_x +  K_{\textrm{orient}}  (O_{\theta} - goal_\theta ) \cos(O_{\theta}) \\
+goal_y = O_y +  K_{\textrm{orient}}   ( O_{\theta} -goal_\theta  ) \sin(O_{\theta})
 \end{align}
 Here $K_{orient}$ is a positive gain on the control input.
+
+%\textcolor{red}{??TODO: PUT THE REAL SIMULATIONS HERE.}
+%
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=0.8\columnwidth]{Orientation.pdf}
+%\end{center}
+%\vspace{-1em}
+%\caption{\label{fig:OrientCont}
+%Plot demonstrating  orientation control of a rectangular object.
+The green line is the goal orientation.
+Other lines show different random starting position of the swarm.
+When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace until the variance is below a desired threshold.
+%}
+%\vspace{-1em}
+%\end{figure}
+
+Fig.
+\ref{fig:OrientCont} illustrates this controller with different starting positions.
+When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace.
+Code is available at \cite{Shahrokhi16Orient}.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=0.8\columnwidth]{LFig.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:LFig}
+Simulation results from a swarm applying force to a hinged door. 
+The swarm mean is steered toward a point $C$ units along the object from the pivot point. The red dashed line indicates the times that the swarm was in variance control mode.
+% Simulation used 144 robots of diameter $0.2$ m with a standard deviation of less than $1.5$ m and an object length of $6$ m.
+}
+\vspace{-1em}
+\end{figure}
 
@@
 The green line is the goal orientation.
-Other lines show different random starting position of the swarm.
-When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace until the variance is below a desired threshold.
+Other lines show different random starting position of the swarm.% When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace until the variance is below a desired threshold.
 }
@@
 \end{figure}
-
-Fig.
-\ref{fig:OrientCont} illustrates this controller with different starting positions.
-When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace.
-Code is available at \cite{Shahrokhi16Orient}.

### 2018-06-11 Shiva Dissertation/chap3/theory.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torque control}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Controlling Torque}

We derive inspiration from recent work on pulling with a swarm \cite{pulltogether}. The contribution of this work is to map swarm distributions to torque production.
%The orientation of an object's major axis is important when a swarm is manipulating a non-symmetric object through narrow corridors. 
%Orientation is controllable by applying torque to the object. 
To change the output torque $\tau$ in~\eqref{eq:torque}, we can choose the direction and magnitude of the force applied $F$, and the moment arm from the object's center of mass (COM) $O$ to the point of contact $P$.
We define a coordinate frame rooted at the COM, with the $x$-axis parallel to the object's longest axis. The resulting torque is:
%
\begin{equation}
\tau = F \times (P - O ).\label{eq:torque}
\end{equation}
%
We assume that the forces produced by individual particles add linearly. 
 As \cite{pulltogether} indicates, this is often a simplification of the true dynamics. 
 The swarm version of \eqref{eq:torque} is the summation of the forces contributed by $n$ individual particles:
\begin{align}
\tau_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O ),   \label{eq:swarmtorque}\\
F_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i.  \label{eq:swarmforce}
\end{align}
Here $F_i$ is the force that the $i^{\textrm{th}}$ particle applies. 
Not all particles are in contact with the object.  
$\rho_i$ is an indicator variable: $\rho_i=1$ if the particle is in direct contact with the object or touching a chain of particle where at least one particle is in contact with the object. 
 Otherwise $\rho_i = 0$.
The moment arm is the particle's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$. 
 If all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ equals some constant.
%
\section{Instantiating torque from swarm distribution}\label{sec:torqueDist}
%
Consider a swarm of particles whose marginal distribution along $x$ has probability density $p(x)$, where $x$ is defined as perpendicular to the object's long axis.
This section considers three canonical probability distributions: uniform, triangular, and normal, all parameterized by mean $\mu$ and standard deviation $\sigma$. 
They are plotted in Fig.~\ref{fig:PDF} and described by:
%
%uniform distribution
\begin{align}
p_u(x) &=  \left\{
\begin{array}{ll}
    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu+\sqrt{3}\sigma\\
     0, & \textrm{otherwise}\\
\end{array} 
\right. ,\\
%triangular distribution
p_t(x) &=  \left\{
\begin{array}{ll}
    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
     0, & \textrm{otherwise}\\
\end{array} 
\right. ,\\
%normal distribution
p_n(x) &= \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right. \kern-\nulldelimiterspace} {2\sigma ^2 }}}.
\end{align}
The next sections examine where to steer the mean of the probability distribution to maximize torque. We discuss two problems: pivoted object torque and free object torque. All the results are summarized in Table~\ref{tbl:MaximizingTorque}, and the calculations are included in the Mathematica file in the multimedia attachment.
%
%%%%%%%%%%% HUGE TABLE %%%%%%%%%%%%%%%%%%%%

\begin{table}
\begin{center}
	\caption{Main results from Section \ref{sec:torqueDist} for maximizing torque with three common distributions. \label{tbl:MaximizingTorque} }
\resizebox{\columnwidth}{!}{
\begin{tabular}{ c rl  rl }
Distribution  & \multicolumn{2}{c}{The best $\mu$ to push} & \multicolumn{2}{c}{Maximum possible torque}\\ [0.5ex] 
\hline \hline
Pivoted Uniform & $
\mu_{u_\max}=$&$\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &    \,~~\qquad \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
1-\sqrt{3}\sigma<\mu<\sqrt{3}\sigma &    \,~~\qquad \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right.
$& $
\tau_{u_{\max}} =$&$\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &  ~~\quad\qquad\qquad \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
\frac{1}{4 \sqrt{3}\sigma} &~~\quad\qquad\qquad \textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right.
$\\ 
\hline
Pivoted Triangular &$\mu_{t_\max}=$&$ \left\{
\begin{array}{ll}
 \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \qquad\qquad\textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
 \frac{\sqrt{2}}{2} &   					\qquad\qquad\textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
\end{array} 
\right. $&$ \tau_{t_\max} =$&$\left\{
\begin{array}{ll}
\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
\frac{\sqrt{2}-2+3\sqrt{6}\sigma}{36\sigma^2} &   \textrm{for     }\sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right. $\\  
\hline \hline
Free Uniform & $\mu_{u_{f_\max}}=$&$\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &  ~\qquad\qquad\qquad\qquad \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
\sqrt{3}\sigma &    ~\qquad\qquad\qquad\qquad \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right.$ & $\tau_{u_{f_{\max}}} =$&$\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &   ~~\quad\qquad\qquad\textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
\frac{1}{4 \sqrt{3}\sigma} & ~~\quad\qquad\qquad\textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right.$\\
\hline
Free Triangular & $\mu_{t_{f_\max}}=$&$ \left\{
\begin{array}{ll}
 \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     }  \sigma  <\frac{2}{\sqrt{6}} \\
1<\mu<  \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma&   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}} 
\end{array} 
\right. $& $\tau_{t_{f_\max}} =$&$\left\{
\begin{array}{ll}
\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma < \frac{2}{\sqrt{6}}\\
\frac{1}{9\sigma^2} &   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}}\\
\end{array} 
\right.
$\\%[0.5ex]
\hline
\end{tabular}
}
\vspace{0.5em}
\end{center}
\end{table}

%%%%%%%%%%%%%%%% END HUGE TABLE %%%%%%%%%%%%%%%%
%
\subsection{Pivoted object torque}
%
In this problem, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is
\begin{equation}
\tau_{p} = \int_0^1 x\,p(x)\, dx.
\end{equation}

\paragraph{Defining torque for distributions}
\begin{figure*}
\centering
\renewcommand{\figwid}{0.66\columnwidth}
\begin{overpic}[width =\figwid]{Uniform.pdf}%\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{Triangular.pdf}
\end{overpic}
\begin{overpic}[width =\figwid]{Normal.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:PDF} Three distributions are examined in this work: uniform, triangular, and normal. Each is plotted above with $\mu=1$ for representative $\sigma$ values.
%\vspace{-2em}
}
\end{figure*}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.66\columnwidth}
\begin{overpic}[width =\figwid]{TorqueUni.pdf}\put(1,30){Uniform}
\end{overpic}
\begin{overpic}[width =\figwid]{TorqueTri.pdf}\put(1,30){Triangular}
\end{overpic}
\begin{overpic}[width =\figwid]{TorqueNormal.pdf}\put(1,30){Normal}
\end{overpic}
\vspace{0.5em}
\caption{\label{fig:torque} Torque as a function of mean position $\mu$. Mean position is the pushing location along a rod extending from 0 to 1. For all distributions pushing at $\mu = 1$ is not optimal unless $\sigma = 0$ or the swarm is uniformly distributed with $\sigma > \frac{1}{2\sqrt{3}}$.
%\vspace{-2em}
}
\end{figure*}

First,  for simplicity of the following derivations, the lower $l$ and upper bound $u$ are defined for the uniform distribution as
%defining torque uniform distribution
\begin{align}
l &= \max(0,\mu -\sqrt{3} \sigma) \, , u = \min({1,\mu+\sqrt{3}\sigma}).
\end{align}
Torque by a uniformly distributed swarm is
\begin{align}
\tau_{u_p} &=  \frac{u^2-l^2}{4\sqrt{3}\sigma}. %\textrm{  for    }  u>l
\end{align}
To simplify derivations for the triangular distribution, we define the bounds of integration as
%define torque for triangular:
\begin{align}
l_1 &= \max({0,\mu-\sqrt{6}\sigma})\, , l_2 = \max({0,\mu}),\\ \nonumber
u_1 &= \min({1,\mu})\, , u_2 = \min({1,\mu+\sqrt{6}\sigma}). \nonumber
\end{align}
Torque by a triangularly distributed swarm is
\begin{align}
\tau_{t_p} &= \frac{\tau_{l} + \tau_{r}}{36\sigma^2},
\end{align}
where $\tau_{l}$ and $\tau_{r}$ are defined as:
\begin{align}
\tau_{l} &= {l_1}^2(2l_1-3(\mu-\sqrt{6}\sigma)) \\ \nonumber
&+{u_1}^2(2u_1 - 3(\mu-\sqrt{6}\sigma)), \\ \nonumber%& \textrm{for     } u_1 > l_1\\ \nonumber
\tau_{r} &= {l_2}^2(2{l_2}-3(\mu+\sqrt{6}\sigma))\\ \nonumber
&-{u_2}^2(2u_2 - 3(\mu+\sqrt{6}\sigma)).\\ \nonumber % & \textrm{for     } u_2 > l_2\\ \nonumber
\end{align}
Torque by a normally distributed swarm is:
%defining torque normal distribution
\begin{align} \label{eq:normalTorque}
\tau_{n_p} &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\ \nonumber
&+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right) .
\end{align}
The $\erf(\cdot)$ is the error function. Torque is plotted as a function of $\mu$ for representative $\sigma$ values in Fig.~\ref{fig:torque}.
\paragraph{Maximum torque for distributions}
For a uniformly distributed swarm, the mean position that maximizes torque on a pivoted object is:
% best place to push for uniform
\begin{align}\label{eq:maxMuPivot}
\mu_{u_\max}=\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
1-\sqrt{3}\sigma<\mu<\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right. .
\end{align}

As soon as the width of the uniformly distributed swarm is longer than the rod ($\sigma > \frac{1}{2\sqrt{3}}$), there exists a range of optimal solutions: any $\mu$ such that the swarm covers 0 to 1.

The resulting torque is:
\begin{align}
\label{eq:MaxTorqueUniformPivot}
\tau_{u_{\max}} =\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
\frac{1}{4 \sqrt{3}\sigma} & \textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right. .
\end{align}
%best place to push for triangular:
For a triangularly distributed swarm, the mean position that maximizes torque is:

\begin{align}
\mu_{t_\max}= \left\{
\begin{array}{ll}
 \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
 \frac{\sqrt{2}}{2} &   \textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
\end{array} 
\right. .
\end{align}

Thus the torque is:
\begin{align}\label{eq:triMaxTorque}
\tau_{t_\max} =\left\{
\begin{array}{ll}
\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
\frac{\sqrt{2}-2+3\sqrt{6}\sigma}{36\sigma^2} &   \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right. .
\end{align}

%best place to push for normal distribution:
For all distributions, the $\mu$ that maximizes torque if $\sigma = 0$ is 1. For the $\tau_{p}$ case, the optimal $\mu$ moves in the $-x$ direction, and for some distributions approaches a limit as $\sigma$ increases. This limit is $\mu = \frac{\sqrt{2}}{2}$ for a triangularly distributed swarm. %We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity. It was because in the first derivative numerator and denominator were both going to infinity and we were not able to calculate the limit. The result is:
In a normal distribution, the first derivative of \eqref{eq:normalTorque} is:
\begin{align}
\frac{d_{\tau}}{d_{\mu}}= \frac{1}{2}\left(\erf{\frac{1-\mu}{\sqrt{2}\sigma}}+ \erf{\frac{\mu}{\sqrt{2}\sigma}}\right)- \frac{e^{\frac{-(\mu-1)^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma}.
\end{align}

The limit as $\sigma\to\infty$ is $\frac{2 - 3 \mu}{\sigma^3 6 \sqrt{2 \pi}}$ which goes to zero as the swarm variance increases. However, the zero of the numerator is $\mu = \frac{2}{3}$ which means as $\sigma$ increases the best position to push the rod asymptotically reaches $\mu = \frac{2}{3}$.

The central limit theorem explains why systems where all the particles have independent noise form a normal distribution. However, optimizing torque for a normal distribution involves two error functions ($\erf$). Instead, we have solved all the equations for the triangular distribution. The numerical study illustrated in the left plot of Fig.~\ref{fig:bestLoc} shows the triangular and normal distributions perform similarly.

%\begin{align}
%\lim_{\sigma\to\infty} \frac{d^3\tau_n}{d\mu}= \frac{3\mu}{2}-1
%\end{align}

%which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.

%pivoted comparisons:
\begin{figure*}
\centering
%\renewcommand{\figwid}{0.66\columnwidth}
\begin{overpic}[width=0.4\columnwidth]{PivotDrawing.pdf}
\end{overpic}\\
\begin{overpic}[width =0.75\columnwidth]{BestLocationPivot.pdf}%\put(1,55){a)}
\end{overpic} 
\begin{overpic}[width =0.61\columnwidth]{MaxTorquePivot.pdf}%\put(1,55){a)}
\end{overpic}
%\vspace{-0.5em}
\caption{\label{fig:bestLoc} Best location to push and maximum torque plots for a pivoted object of length 1, pivoted at 0. Generating code is in the attachment.
%\vspace{-2em}
}
\end{figure*}

\subsection{Free object torque}
In this problem, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
\begin{equation}
\tau_{f} = \int_{-1}^1 x\,p(x)\, dx .
\end{equation}
\emph{a) Defining torque to a free rod for distributions:}

To calculate free object torque for a uniformly distributed swarm, for simplicity of derivations, $l_f$ and $u_f$ are defined: 
%torque uniform free
\begin{align}
l_f &= \max({-1,\mu -\sqrt{3} \sigma}),\, u_f = \min({1,\mu+\sqrt{3}\sigma}).
\end{align}

The torque applied by a uniformly distributed swarm is:
\begin{align}
\tau_{u_f} &= \frac{u_f^2-l_f^2}{4\sqrt{3}\sigma} \textrm{  for    }  u_f>l_f .
\end{align}

%torque triangular free
To simplify derivations for a triangularly distributed swarm, we define bounds of integration as:
\begin{align}
l_{f_1} &= \max({-1,\mu-\sqrt{6}\sigma}), \,l_{f_2} = \max({-1,\mu}),\\ \nonumber
u_{f_1} &= \min({1,\mu}), \, u_{f_2} = \min({1,\mu+\sqrt{6}\sigma}) . \nonumber
\end{align}
The torque applied by a triangularly distributed swarm is:
\begin{align}
\tau_{t_f} = \frac{\tau_{f_l}+ \tau_{f_r}}{36\sigma^2},
\end{align}
where $\tau_{f_l}$ and $\tau_{f_r}$ is defined as
\begin{align}\nonumber
\tau_{f_l} &=  {u_{f_1}}^2(2u_{f_1} - 3(\mu-\sqrt{6}\sigma)) \\ \nonumber
&-{l_{f_1}}^2(2{l_{f_1}}-3(\mu-\sqrt{6}\sigma)) , \\ \nonumber %&   \textrm{for     } u_{f1} > l_{f1}\\ \nonumber
\tau_{fr} &=-{u_{f_2}}^2(2u_{f_2} -3(\mu+\sqrt{6}\sigma) , \\ \nonumber
&+ {l_{f_2}}^2(2{l_{f_2}}-3(\mu+\sqrt{6}\sigma)) . \\ \nonumber%&   \textrm{for     } u_{f2} > l_{f2}\\ \nonumber
\end{align}

%torque normal free
The torque applied by a normally distributed swarm is:
\begin{align} \nonumber
\tau_{n_f} &= \frac{\left(e^{\frac{-(\mu+1)^2}{2\sigma^2}}-e^{\frac{-(\mu-1)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\
 &+ \frac{1}{2}\mu\left(\erf\Big(\frac{1-\mu}{\sqrt{2}\sigma}\Big)+\erf\Big(\frac{1+\mu}{\sqrt{2}\sigma}\Big)\right) .
\end{align}

%free torque comparison
\begin{figure*}
\centering
%\renewcommand{\figwid}{0.49\columnwidth}
\begin{overpic}[width =0.4\columnwidth]{FreeDrawing.pdf}
\end{overpic}\\
\begin{overpic}[width =0.75\columnwidth]{TorqueFreeBestPush.pdf}
\end{overpic}
\begin{overpic}[width =0.61\columnwidth]{TorqueFreeComparison.pdf}
\end{overpic}
%\vspace{-0.5em}
\caption{\label{fig:maxTorque} Best location to push and maximum torque plots for a free object of length 2, located from $x=-1$ to $1$.
%\vspace{-2em}
}
\end{figure*}

\emph{b) Maximum torque for distributions}
Maximum torque of a uniformly distributed swarm to a free object has the same upper solution as with a pivoted object:

\begin{align}\label{eq:maxMuFree}
\mu_{u_{f_\max}}=\left\{
\begin{array}{ll}
1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
\end{array} 
\right. .
\end{align}

The major difference is the existence of only one solution for $\sigma$ values for torque on a free object. The left bound of the distribution must never be less than 0, or torque applied left of the center will cancel torque to the right of the center. 
 The maximum torque produced is given by \eqref{eq:MaxTorqueUniformPivot}.
%The difference is because the pivoted object even when the swarm is pushing all the length of the object, they are applying torque, but for the free object, if the swarm pushes the whole length of the object, they would just apply force.

For a triangularly distributed swarm the optimal mean is:

\begin{align}
\mu_{t_{f_\max}}= \left\{
\begin{array}{ll}
 \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     }  \sigma  <\frac{2}{\sqrt{6}} \\
1<\mu<  \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma&   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}} 
\end{array} 
\right. .
\end{align}
%limits and equations

As with the uniform distribution in \eqref{eq:maxMuPivot} the optimal solution with $\sigma > \frac{2}{\sqrt{6}}$ has a range of solutions. Setting $\mu=1$ maximizes force for the optimal torque. However, moving the mean right reduces the negative torque applied by particles to the left of 0, but this gain is canceled by a corresponding reduction in positive torque.

Therefore, the maximum torque is:

\begin{align}\label{eq:triFreeMaxTorque}
\tau_{t_{f_\max}} =\left\{
\begin{array}{ll}
\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma < \frac{2}{\sqrt{6}}\\
\frac{1}{9\sigma^2} &   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}}\\
\end{array} 
\right. .
\end{align}

Again, for a normally distributed swarm, we have numerically found the maximum torque and optimal pushing location $\mu$. The results are closely approximated by the upper bound solution for the triangular system. 
 The results are plotted in Fig.~\ref{fig:maxTorque}.

\section{Angle of repose}\label{sec:angle}

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =0.6\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
%Particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
% A thin gray line extends upwards from the rod COM, showing how more particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
%  This uneven particulate generates a restoring torque. 
 Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
 %The rod length $\ell=1$. 
   The maximum torque values are shown with black dots.
% Generating code is in the attachment. 
%\vspace{-2em}
}
\end{figure}

Consider a swarm of granular particles applying force to a rod. 
If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle. This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}.% for the media, see \cite{angleofRepose}.
By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
 We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:

\begin{align}
F(\theta,\alpha,\ell) =\left\{
\begin{array}{ll}
\frac{-\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos\alpha\sin{\theta}} &   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise}\\
\end{array} 
\right.
\end{align}

%\begin{align}
%F(\theta,\alpha,l)  = \frac{l^2}{8\cos\alpha\sin{\theta}} \Big(\cos(2\theta)-\cos(2\alpha)\Big).
%\end{align}

The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}. 
Torque will also be similarly defined as:
\begin{align}
\tau(\theta, \alpha, \ell) =\left\{
\begin{array}{ll}
\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin{\theta} }{48\sin^2\alpha}&   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce}.
 Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$:
\begin{align}
\theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.
\end{align}
To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.

### 2018-06-11 Shiva Dissertation/chap4/covControl.tex
Shiva PhD dissertation

--- 2018-06-11 RA-L/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Covariance Control}
+\section{Covariance control}
 \label{sec:covControl}
@@
 Left plot shows locus of possible mean positions for five values of $A$.
-The locus morphs from a square to a circle as $A$ increases.
+%The locus morphs from a square to a circle as $A$ increases.
 The covariance ellipse for each $A$ is shown with a dashed line.
  Center shows two corresponding arrangements of kilobots.
-At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
-See online interactive demonstration at \citep{Zhao2016mathematicaSquare}.}
+% At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
+See online interactive demonstration at \cite{Zhao2016mathematicaSquare}.
+ }
 \label{fig:SquareFill}
@@
 Left plot shows locus of possible mean positions for four values of $h$.
-The locus of possible mean positions are concentric circles.
-See online interactive demonstration at \citep{Zhao2016mathematica}.}
+%The locus of possible mean positions are concentric circles.
+%See online interactive demonstration at \cite{Zhao2016mathematica}.
+}  
 \label{fig:CircleFill}
@@
 \end{figure*} 
+Covariance control as well as mean position and variance control is needed when the swarm is moving through narrow passageways.
+This section discusses ideas to control covariance of the swarm using boundaries.
 
-\subsection{Using Boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
+\subsection{Using boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
 One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
@@
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
-Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \citep{Zhao2016mathematicaSquare}.
+Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \cite{Zhao2016mathematicaSquare}.
 
@@
 \end{align}
-The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
+The full equations are included in the appendix~\cite{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
 A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
@@
 Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}, see also \citep{Zhao2016mathematica}.
+Results are displayed in Fig.~\ref{fig:CircleFill}, see also \cite{Zhao2016mathematica}.
 
@@
 
-\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
+\subsection{Using boundaries: friction and boundary layers}\label{subsec:WallFriction}
 Global inputs move a swarm uniformly.  
@@
 Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \citep{fluidMechanics}.
+Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.
 \begin{align}
@@
 
-\subsection{Maximizing Correlation Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
+\subsection{Maximizing correlation using wall friction}\label{subsec:ClosedLoopCovarianceControl}
 Using the environment to individually move $n$ particles to goal positions in Alg.
 \ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg.
-\ref{alg:PosControl2Robots} can only control two particles.
+\ref{alg:optimalAlg} can only control two particles.
 The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited.
@@
  Assume that the swarm is initialized in the lower-left corner, in a rectangle of width $w$ and height $A/w$. 
- Such a rectangular configuration can be accomplished using the variance controllers from \citep{ShahrokhiIROS2015}. 
+ Such a rectangular configuration can be accomplished using the variance controllers from \cite{ShahrokhiIROS2015}. 
   If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
@@
   Negative correlations can be generated by initializing the swarm in the upper left, or lower right.
+
+\subsection{Efficient control of correlation}
+%A set of simulations were conducted to demonstrate the importance of boundary friction.  
+This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
+from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
+The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
+For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
+As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
+Larger boundary layers $h$ enable more control of correlation.
+The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:ControlledCorrelation}
+Analytical results comparing maximum correlation $\rho_{xy}$ under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+}\vspace{-1em}
+\end{figure}
+\subsection{Hardware experiment: control of covariance}
+To demonstrate covariance control, up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
+The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{newExperiment.pdf}
+\end{center}
+%\vspace{-1em}
+\caption{\label{fig:covExperiment}
+Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance.
+The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.
+The actual covariance is shown in different trials.
+Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
+%\vspace{-2.5 em}
+}
+\end{figure}

### 2018-06-11 Shiva Dissertation/chap4/intro.tex
Shiva PhD dissertation

--- 2018-05-29 journalWallFriction/intro.tex
@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
+%Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
 %If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
@@
 
-The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
- The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
-Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
- Techniques are needed that can handle this underactuation.
+%The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
+ %The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
+%Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
+% Techniques are needed that can handle this underactuation.
 
  Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
- In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
+ 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
  However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
-This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contacts.
+This chapter relaxes that constraint, and provides position control algorithms that only require non-slip wall contacts.
 We assume that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
@@
 \begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
-\vspace{1em}
+\vspace{2em}
 \caption{\label{fig:IntroPic}
 Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
-} \vspace{-1em}
+}% \vspace{-1em}
 \end{figure}
@@
 
-This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces.
-This work extends the analysis to convex workspaces and 3D positioning.
-This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.
+%This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces. 
+This work analyses any convex workspaces and 3D positioning. This chapter also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-06-11 Shiva Dissertation/chap4/relatedWork.tex
Shiva PhD dissertation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}\label{sec:RelatedWork}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Controlling the \emph{shape}, or relative positions, of a swarm of robots is a key ability for a range of applications.  Correspondingly, it has been studied from a control-theoretic perspective in  both centralized and decentralized approaches. For examples of each, see the centralized virtual leaders in \cite{egerstedt2001formation}, and the  gradient-based decentralized controllers  using control-Lyapunov functions in~\cite{hsieh2008decentralized}. However, these approaches assume a level of intelligence and autonomy in individual robots that exceeds the capabilities of many systems, including current micro- and nano-robots.  Current micro- and nano-robots, such as those in~\cite{Chowdhury2015,martel2015magnetotactic,Xiaohui2015magnetiteMicroswimmers} lack onboard computation.

This chapter focuses on centralized techniques that apply the same control input to both particles. 
Precision control requires breaking the symmetry caused by the uniform input.  
Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}. 
 The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
 This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013}, and breaks the control symmetry using obstacles in the workspace.

%The techniques in this paper are inspired by artificial force-fields.

%\emph{Fluid-flow:} 
%Fluid flow along boundaries generates a shear force that pushes different parts of a body in opposing directions. 
%Most introductory fluid dynamics textbooks provide models~\citep{Munson2013}.
%Similarly, a swarm of robots under global control pushed along a boundary will experience shear forces.  
%This is a position-dependent force, and so can be exploited to control the configuration or shape of the swarm.  
% \citep{spears2006physics}~used these forces to disperse a swarm's spatial position for coverage for physics-based swarm simulations.

%\emph{Artificial Force-fields:}
%Much research has focused on generating non-uniform artificial force-fields that can be used to rearrange passive components. 
Alternative techniques rely on non-uniform inputs, such as artificial force-fields.
Applications have included techniques to design shear forces for sensorless manipulation of a single object by~\cite{lamiraux+2001:ra}.  
\cite{vose2012sliding} demonstrated a collection of 2D force fields generated by six degree-of-freedom vibration inputs to a rigid plate.  These force fields, including shear forces, could be used as a set of primitives for motion control to steer the formation of multiple objects. %However unlike the uniform control model in this paper, their control was multi-modal and position-dependent.
%\todo{talk about obstacles, think about adding goldberg}

%This paper develops control algorithms using uniform control fields, such as the magnetic resonance navigation \cite{nosrati2018development}.%field in a clinical MRI [insert a recent reference from Sylvain Martel using MRI].
Similarly, much recent work in magnet control has focused on exploiting inhomogeneities in the magnetic field to control multiple micro particles  using gradient-based pulling~\cite{Salmanipour2018EightDOF,Denasi2018independent}.  
Unfortunately, using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the  distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six, Denasi2018independent, Salmanipour2018EightDOF }. In contrast, % to methods that exploit inhomogeneities in the magnetic field to control multiple micro particles, e.g. \cite{Denasi2018independent}, that exploited nonlinearities generated by four magnetic coils in close proximity to the workspace to achieve trajectory control of two microspheres, 
 this chapter requires only a controllable constant gradient in orthogonal directions to position the particles.
% Systems like this one are poorly suited for PRM and RRT*-type methods\cite{lavalle2006planning} because if during a movement a collision occurs, that movement is irreversible.
   %Flow in a pipe or body lume

If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
 Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
  Instead, this chapter employs a greedy search strategy.

### 2018-06-11 Shiva Dissertation/chap4/simulation.tex
Shiva PhD dissertation

--- 2018-05-29 journalWallFriction/simulation.tex
@@
 %Two simulations were implemented using non-slip contact walls for position control.  The first controls the position of two robots, the second controls the position of $n$ robots.  
+\begin{figure*}
+\begin{overpic}[width=0.32\columnwidth]{regionMove1.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,  - , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{regionMove2.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{regionMove3.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
+
+\begin{overpic}[width=0.32\columnwidth]{regionMove4.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{regionMove5.pdf}\put(55.5,26){\begin{turn}{94} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{regionMove6.pdf}\end{overpic}\\
+
+\begin{overpic}[width=0.32\columnwidth]{Move1.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{Move2.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{Move3.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted    ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}\\
+
+\begin{overpic}[width=0.32\columnwidth]{Move4.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -  , dotted      ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{Move5.pdf}\put(55.5,32.5){\begin{turn}{50.5} 
+{\begin{tikzpicture}[thick]
+\draw [red,   -   , dotted     ] (1,14.0) -- (0,14.0);
+\end{tikzpicture}}\end{turn}}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{finalMove.pdf}\end{overpic}
+\caption{\label{fig:reachableSet}
+Frames from reconfiguring two particles. 
+Top six images show a polygonal workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
+%For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
+%The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
+%See multimedia attachment for animations of each.
+}
+\end{figure*}
 

### 2018-06-11 Shiva Dissertation/chap4/theory.tex
Shiva PhD dissertation

--- 2018-05-30 journalWallFriction/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Boundary Interaction Model}\label{subsec:WallFriction}
+\subsection{Boundary interaction model}\label{subsec:WallFriction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
- \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
+ \subsection{Example: shortest path in a square workspace}\label{subsec:square}
  Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
@@
 \begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
-\vspace{-2em}
+%\vspace{-2em}
 \caption{\label{fig:reflection}
 The shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle.
-} \vspace{-1em}
+} %\vspace{-1em}
 \end{figure}
@@
 \begin{figure*}
-\renewcommand{\figwid}{0.4\columnwidth}
+\renewcommand{\figwid}{0.19\columnwidth}
 
@@
 \put(87,52){$g_2$}
-\put(60, 3.9){{\tiny$\updownarrow$}~$\epsilon$}
+\put(60, 3.5){{\tiny$\updownarrow$}~$\epsilon$}
 \end{overpic}
@@
 \put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
-\put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
+\put(65, 47){{$\xupdownarrow{1.5cm}$}$L$}
 %\put(47, 65){$\xleftrightarrow{1.8cm}$}
@@
 {
-\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
+\begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(55,10){Move 1}
 \put(10,85){f)}
@@
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(60,10){Move 2}
+\begin{overpic}[width =\figwid]{storyWorst2.pdf}\put(55,10){Move 2}
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(60,10){Move 3}
+\begin{overpic}[width =\figwid]{storyWorst4.pdf}\put(55,10){Move 3}
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(60,10){Move 4}
+\begin{overpic}[width =\figwid]{storyWorst6.pdf}\put(55,10){Move 4}
 \end{overpic}
-\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(60,10){Move 5}
+\begin{overpic}[width =\figwid]{storyWorst7.pdf}\put(55,10){Move 5}
 \put(15, 85){worst-case}\put(15,75){path length}\put(15, 62){$(\sqrt{2}+2)L$}
@@
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two particle positioning using walls with non-slip contacts. 
-Particle start positions are shown by squares, and goal positions by circles.
+%Particle start positions are shown by squares, and goal positions by circles.
 Dashed lines show the shortest route if particles could be controlled independently.
@@
 
- \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
+ \subsection{Shortest path in unit disk that intersects circumference}\label{subsec:circular}
 
@@
 The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right.
-See \cite{BeckerShortestPath}.}
+%See \cite{BeckerShortestPath}.
+}
 %\vspace{-1em}

### 2018-06-11 Shiva Dissertation/chap4/theory2robAlgs.tex
Shiva PhD dissertation

--- 2018-05-30 journalWallFriction/theory2robAlgs.tex
@@
-\section{Position Control of Two Particles Using Boundary Interaction}\label{sec:PostionControl2Robots}
+\section{Position control of two particles using boundary interaction}\label{sec:PostionControl2Robots}
 
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%\
-\subsection{$\Delta$ Configuration Space}
+\subsection{$\Delta$ configuration space}
 The configuration space for two particles is a four dimensional manifold.
@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%\
-\subsection{Two Particle Path Planning}
+\subsection{Two particle path planning}
 
@@
 
-\subsection{Convex Polygonal Workspaces: 2-Move Reachable Set}
+\subsection{Convex polygonal workspaces: 2-move reachable set}
    \begin{figure}
 \centering
-\renewcommand{\figwid}{0.8\columnwidth}
+\renewcommand{\figwid}{0.48\columnwidth}
 {\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
@@
 The sets reachable in two moves, called \emph{2-move reachable sets}, are drawn with transparent blue polygons.
-A polygon with $n$ sides has $n$ 2-move reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
+%A polygon with $n$ sides has $n$ 2-move reachable sets, but if $n$ is even and the polygon is regular, half the reachable sets overlap.
 If $\Delta g$ is in the 2-move reachable sets, we can achieve the required relative position in two moves.
@@
 }
-\vspace{-1em}
+%\vspace{-1em}
 }
@@
 \caption{\label{fig:TwoRegions}
-Boundary interaction is used to change the relative positions of the particles. Each particle gets the same control input. 
-(left) If particle 2 contacts the bottom wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area. 
-(middle) Similarly, if particle 2 contacts the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
+%Boundary interaction is used to change the relative positions of the particles. %Each particle gets the same control input. 
+(left and middle) If particle 2 contacts the bottom or right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and  particle 1 can move to anywhere in the shaded area.
+%(middle) Similarly, if particle 2 contacts the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
 (right) The 2-move reachable sets in the $\Delta$ configuration space is shown.
@@
 \centering
-\renewcommand{\figwid}{0.66\columnwidth}
+\renewcommand{\figwid}{0.5\columnwidth}
 \begin{overpic}[width =\figwid]{anypolygon1.pdf}\put(12,-3){\small{Case 1: $s_1$ does not contact boundary}}\put(28,39){$s_2$}\put(63,50){$s_1$}
@@
 \end{overpic}
+\vspace{2em}
 \caption{\label{fig:polygonAlg}{ Steps to generate  the 2-move reachable set when  one particle collides with edge $\overline{ p_i p_{i+1}}$ of a convex polygonal workspace.}
-\vspace{-1em}
+%\vspace{-1em}
 }
@@
 
-\subsection{Circular Workspaces: 2-Move Reachable Set}
+\subsection{Circular workspaces: 2-move reachable set}
 
@@
 
-\subsection{3D workspaces: Cylinders and Prisms}
+\subsection{3D workspaces: cylinders and prisms}
 

### 2018-06-11 Shiva Dissertation/chap4/theorynrobAlgs.tex
Shiva PhD dissertation

--- 2017-07-27 ICRA-RA-L2018/theorynrobAlgs.tex
@@
-\section{Position Control of $n$ Robots Using Boundary Interaction}\label{sec:PostionControlnRobots}
+\section{Position control of $n$ robots using boundary interaction}\label{sec:PostionControlnRobots}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \caption{\label{fig:driftmove}
-A  $\text{\sc DriftMove}(\alpha, \beta, \epsilon,0^\circ)$ repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$.
+A  $\text{\sc DriftMove}(\alpha, \beta, \epsilon,0^\circ)$ repeats a triangular movement sequence $\{ (\beta/2,-\epsilon)$, $(\beta/2,\epsilon)$, $(-\alpha,0)\}$.
 At the sequence end, robot $A$ has moved $\beta$ units right, and robot $B$ has moved $\beta-\alpha$ units right.}

### 2018-06-11 Shiva Dissertation/chap/abs.tex
Still requires some edits.

@@
 The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
-Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
- Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
+%Robotic manipulation usually assumes intelligent agents, but particle systems manipulated by a global signal.
+ When there are more particles than control inputs the system is underactuated and requires new control techniques.
  Rather than focusing on a specific microrobotic system, this dissertation designs control laws and algorithms for steering many particles controlled by global fields.
@@
 
- Second, this dissertation proves that swarm torque control is possible, then presents algorithms to automate the task. 
- Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
+ Second, this dissertation proves that swarm torque control is possible, then presents algorithms to automate the task. Torque control enables us to control the position and orientation of an object.
+% Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 

### 2018-06-11 Shiva Dissertation/chap/conclusion.tex
Still requires some edits.

@@
 
- \section{Future Work}
+ \section{Future work}
+
+ Today there are more robots than ever before.
  This dissertation provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.

### 2018-06-11 Shiva Dissertation/chap/intro.tex
Still requires some edits.

@@
 
+\begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{MicroRobotics.png}
+\end{overpic}
+\caption{\label{fig:microrobotics} %Controlling a swarm of robots with global inputs. 
+Microrobots has the potential to revolutionize medical applications. Used with permission from \cite{nelson2010microrobots}.
+}
+\end{figure}
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-\centering
-\renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{MicroRobotics.png}
-\end{overpic}
-\caption{\label{fig:microrobotics} %Controlling a swarm of robots with global inputs. 
-Microrobots has the potential to revolutionize medical applications. Figure courtesy:  \cite{nelson2010microrobots}.
-}
-\end{figure}
-
 Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction, see \cite{Peyer2013,Shirai2005,Chiang2011}.
 These activities require robots that behave intelligently.
-Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel \cite{nelson2010microrobots}.
+Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection shown in Fig.
+\ref{fig:microrobotics} \cite{nelson2010microrobots} and 2) ex vivo to assemble structures in parallel.
  For each application, large numbers of micro robots are required  to deliver sufficient payloads.
@@
 Top left: Chapter 2 discusses steering a swarm of particles using global inputs and swarm statistics.
-Top right: Chapter 3 goes deep into the torque control of the swarm and orientation control of the object.
-Bottom Chapter 4 demonstrates algorithms to control position of two robots in any convex workspace when the walls have non-slip contact.
+Top right: Chapter 3 goes deep into the torque control.
+Bottom: Chapter 4 demonstrates algorithms to position two robots.
 %Bottom right: the magnetic setup that is used to manipulate magnetic particles when they both get the same control input.

### 2018-06-12 Shiva Dissertation/chap/conclusion.tex
Made few edits and added the new pictures.

@@
  This dissertation provides foundational algorithms and techniques for steering swarms, object manipulation, and addressing obstacle fields, but there are many opportunities to extend the work.
- Our future goal is to perform assembly using particle swarms to manipulate and attach components.
+ Extensions to Chapter 2 and 3 is to perform assembly using particle swarms to manipulate and attach components.
 Future efforts should be directed toward examining the effects of Brownian noise, pose control for multiple-part assembly, trajectory prediction, and manipulation in crowded workspaces.
- We assumed friction was sufficient to completely stop particles in contact with the boundary. 
+ In Chapter 4 We assumed friction was sufficient to completely stop particles in contact with the boundary. 
   The algorithms would require retooling to handle small friction coefficients.

### 2018-06-12 Shiva Dissertation/chap/intro.tex
Made few edits and added the new pictures.

@@
 \end{figure}
+\begin{figure}
+\centering
+\renewcommand{\figwid}{0.49\columnwidth}
+\begin{overpic}[width =\figwid]{physicalPuzzle.pdf}
+\end{overpic}
+\begin{overpic}[width =\figwid]{physicalDelivery.pdf}
+\end{overpic}
+\caption{\label{fig:assembly} %Controlling a swarm of robots with global inputs. 
+Assembling and delivering different parts of an object using a swarm of simple robots.
+}
+\end{figure}
+
 \begin{figure}
@@
 Our vision is for large swarms of robots to be remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection shown in Fig.
-\ref{fig:microrobotics} \cite{nelson2010microrobots} and 2) ex vivo to assemble structures in parallel.
+\ref{fig:microrobotics} \cite{nelson2010microrobots} and 2) ex vivo to assemble structures in parallel shown in Fig.
+\ref{fig:assembly}.
  For each application, large numbers of micro robots are required  to deliver sufficient payloads.

### 2018-06-12 Shiva Dissertation/chap3/theory.tex
Still requires some edits.
Made few edits and added the new pictures.

@@
 %\renewcommand{\figwid}{0.66\columnwidth}
+%\hbox{\hspace{-0.5em} \includegraphics[scale=0.6]{bar}}
 \begin{overpic}[width=0.4\columnwidth]{PivotDrawing.pdf}
@@
 \end{overpic} 
-\begin{overpic}[width =0.61\columnwidth]{MaxTorquePivot.pdf}%\put(1,55){a)}
-\end{overpic}
+\hbox{\hspace{1.5em} \begin{overpic}[width =0.68\columnwidth]{MaxTorquePivot.pdf}%\put(1,55){a)}
+\end{overpic}}
 %\vspace{-0.5em}
@@
 \begin{figure*}
-\centering
+\begin{center}
 %\renewcommand{\figwid}{0.49\columnwidth}
@@
 \end{overpic}
-\begin{overpic}[width =0.61\columnwidth]{TorqueFreeComparison.pdf}
-\end{overpic}
+\hbox{\hspace{1.75em}\begin{overpic}[width =0.68\columnwidth]{TorqueFreeComparison.pdf}
+\end{overpic}}
+
 %\vspace{-0.5em}
@@
 }
+\end{center}
 \end{figure*}
@@
 
+Consider a swarm of granular particles applying force to a rod. 
+If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle.
+This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}.
+Three different values of angle of repose is shown in Fig.~\ref{fig:angle}.
+The center of mass of the rod is in the middle of the rod, but center of mass of the granular particles changes for different values of angle of repose.
+% for the media, see \cite{angleofRepose}.
+\begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{Angle.pdf}%\put(1,55){a)}
+\end{overpic}
+\caption{\label{fig:Angle} Different values of angle of repose is shown when granular particles move faster than the rod.
+}
+\end{figure}
+
+By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
+ We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
+By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:
 \begin{figure}
@@
 \end{figure}
-
-Consider a swarm of granular particles applying force to a rod. 
-If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle.
-This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}.% for the media, see \cite{angleofRepose}.
-By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
- We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
-By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:
 

### 2018-07-12 JournalTorqueControl/Review/Response.tex
so many changes to the dissertation, and Other!

\documentclass[conference, onecolumn, 10pt]{../IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{times}

\usepackage{microtype,xparse,tcolorbox}
\usepackage{mathtools, cuted}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{hyperref}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}

\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

\setlength\parindent{0pt}

\renewcommand{\thepoint}{[R\,\thereviewer.\arabic{point}] }

\newcommand{\response}[1]{\refstepcounter{point}  \smallbreak \noindent
	\hangindent=3.5em
	\hangafter=1
	\thepoint \space ``\textcolor{blue}{#1}" \par\smallskip}

\newcommand{\letter}[2]{\vspace*{\fill}

	\begin{quote}

		\includegraphics[width=8cm]{engineeringLogo.eps}

		\begin{flushright}
		Aaron T.\ Becker\\Department of Electrical Engineering\\University of Houston\\Houston, TX 77004, USA\\Email: atbecker@central.uh.edu\\Web: youTube.com/aabecker5\\

		\today \vspace{2cm} \\
		\end{flushright}

		#1\medskip\\ 
		\medskip\\
		Dear #2 Editorial Office,\medskip\\
		Please find attached the revised paper, #1,
		along with the document containing a response to the reviewers. We are grateful to the reviewers for helping us improve our manuscript through their comments and questions. Please let us know if further information is required.\medskip\\

		Sincerely,\\

			\includegraphics[width=4cm]{SignatureAaronTBeckerWhite.png}
			\medskip\\
		Aaron T.\ Becker (on behalf of all the authors) 
	\end{quote}
	\vspace*{\fill}
	\break }

\NewDocumentCommand \Reviewer { m } {
	\stepcounter{reviewer}
	\subsection*{Comments~by~Reviewer~#1}
}

\begin{document}
	\letter{\emph{Object Manipulation and Position Control Using a Swarm With Global Inputs}}{IEEE Transactions on Automation Science and Engineering}

	\section*{Response to Reviewers}
	In the following document, we have provided detailed responses to the comments and questions of the reviewers. Comments and questions by reviewers are in \textcolor{blue}{blue}, our responses are in black.

	\emph{Comments by Associate Editor}
	\response{The paper is interesting and technically sound; however, a thorough revision is required before being considered for publication.
The main concern to be addressed is the quality of the writing. Reviewers have provided detailed comments on how to improve the quality of this manuscript. Authors are urged to thoroughly revise the paper according to reviewers' comments and suggestions.}

	\Reviewer{\#1}\response{This paper studies the pushing effect of a robot swarm over a rectangular object. In order to study the generated torque, the authors represent the swarm as a probability distribution. Three different probability distributions are presented. The ideas in the paper are very interesting and relevant for the community. However it requires a major revision. The comments are the following.}
	Thank you!

	\response{At the end of Section III, What is the motivation to consider that all robots apply the same force? An intuitive efficient approach would make the farther robots apply a larger force.}
	Roughly the applied force for the particles are the same, the applied torque would be different for the farther robots.

	\response{Saturation is not considered. An efficient approach would be the same as using a bigger robot but a particle instead. Obviously, this approach cannot be done because of the small maximum force that each particle can generate, but it is not mentioned in the paper.}

	\response{The authors claim that one of the main differences is that their robots are compliant and tend to flow over the object. However, some of the experiments present an adaptation on the object to avoid the robots flow over the object.}

	\response{This "flow over the object" is not mentioned at all in the methodology (even for the triangular distribution). The authors propose static distributions, but the the distributions variate on time due to the flow.}

	\response{Based on the point above, the most efficient way (and obvious) to generate the torque is just sending all robots as far as possible from the center of mass.}
	Please pay attention that all the robots are getting the same control input and we are only controlling the mean position.

	\response{In contrast to the title, none of the probability distributions maximizes the torque. At least, the maximization problem is not stated in the manuscript.}

	\response{The authors do not mention how to control the robots. It is one of the main points that call the attention in this type of works.} 
	The particles are getting the same control input.

	\response{In equations (2) and (3), there is a variable n that is not defined. I suppose it is the number of the robots, but it is not explicitly denoted. In these equations, it is not clear that only the surface robots can apply force to the bar. Does $F_i$ takes into account the force that all the robots behind $F_i$ are applying? In this way, (2) and (10) are not consistent.  Assuming that the individual forces are linear is a very strong assumption, in contrast to \[15\], the particles move in a very unorganized way (even more when the particles flow over the object).}

	\response{I recommend to separate Fig. 2 in sub-panels to organize the caption. In the top figure, the particle is not easy to observe and it is not in all the examples. The caption says "Particulate is moving in the -y direction", but a coordinate frame is not defined (same for the x-axis). Is the particle moving with respect to the inertia frame or to the bar frame? Additionally, there is not such a gray line. Are the authors referring to the thin black line? 
In general, this figure is not helping to picture what the authors want.}

	\response{The derivation of (4) and (5) is no shown or extending the description is required.}

	\response{I like the idea of representing the swarm as a probability distribution. It is easier for the observer and the global controller to drive the swarm as a whole. Again in equations (7)-(9), the coordinate frames are not defined and not consistent. The COM $[O_x, O_y]$ is defined in the inertia frame, but the distribution of the robots is defined on the object frame (where the x-axis is parallel to the long side of the rectangle). Although these equations are intuitive, the way the authors obtained the constants (e.g. the sqrt(3) in Eq. (7), only in the next section is assumed that the length is 1) is not obvious.}

	\response{In (16), the function erf is not defined. It is only defined at the end of the section.}

	\response{Based on the granular effect, there is a maximum number of robots that can apply forces without flowing. What is this number? or how can you describe this final configuration?}

	\response{The experiments section need to be extended, there granular effect need to be extensively studied.
Again, it is necessary to describe how the robots are driven to maintain the distribution and to generate the torque.}

	\response{According to the format guidelines, the equations are part of the text and should be punctuated. Most of them are not.}

	\response{Caption in Fig 2. says "Particulate" instead of "Particle"}

	\Reviewer{\#2}
	\response{This paper studies the torque applied by a swarm of particles on a long aspect-ratio rod. Generally, the topic of this paper is interesting. Experiments are well designed to illustrate the theory in practice. Here are some suggestions to be considered for the possible revision. }

	\response{Please add x-y-z axes in all the experimental figures such as Figs.1 and 8.}

	\response{What does cyan/white star points mean in Fig.1?}

	\response{In this paper, two different cases are considered. One is a pivoted object, the other is a free object. Differences between two experimental results are encouraged to be highlighted. }

	\response{Please highlight the material and the size of the rod used in the experiments. Contacts between robots or between robot and rod should be detailed, and their effects should be further discussed if it is possible.}

	\response{The definition of 'kilobots' should be given when it is used for the first time. }

	\Reviewer{\#3}	
	\response{The articles discusses a variant of swarm manipulation where a large number of particle like robots are steered using a common force input.  The article primarily analyses the force and torque they exert on movable objects they come in contact with; as well as the shape the particles make as they accumulate around these objects (shown to be a triangle whole apex angle is called the angle of repose.
The strengths of the article are:
The analysis provided for several manipulation scenarios appears to be sound
The results are novel to my knowledge.
Experimental results on a hardware test bed are shown.
The weaknesses are:
The analysis can be difficult to follow because in some cases the variables are not well defined.
There are many, many grammatical and punctuation errors. (see below)
The abstract is poorly written.}

	\response{Question:  It would seem that the angle of repose might depend on the particle size or the coefficient of friction between particles or the object-particle.   Can you explain why it doesn?t?   I understand this isn?t the primary contribution of the article but it seems those parameters would play a role.}

	\response{The abstract is repetitive and has a non-standard format (note to practitioners?) and should be restructured.   It doesn?t make it clear that all the particles are steered with a common force (different from other swarm scenarios).   It uses the term angle of repose without defining.   The last paragraph about a single robot does not belong in the abstract at all.}

	\response{Introduction - paragraph 1: Last sentence doesn?t flow. }

	\response{The term angle of repose should be defined earlier than the end of page 2, because it is used many times before that and while it is defined in [2] it is not a standard term in the robotics community.}

	\response{ Many of the equations are not properly punctuated.  For example, eq 4 and 5 need periods as they end a sentence.  Same with 7,8,9, 12,13, 18, 19, 20 }

	\response{The last sentence of Setcion 4 doesn?t make sense.  Should the ? -? be an ?=? symbol?}

	\response{This sentence could be worded better: (page 3, col. 1) ?Given sufficient particles to pile up to the angle of repose??   also it raises an interesting point that these calculations assume a sufficient number of particles.   Is there an expression that would estimate the number of particles, given the particle size?}

	\response{Page 3, col 2, top:
It?s not clear what $l_p$ and $u_p$ are. (lower and upper bounds?).   Earlier l is used to mean the length of the rod.
In the first sentence you mention a variable $up_p$, perhaps you mean $u_p$?}

	\response{On page 6 you use an inconsistent style when referring to equations.   At one point you use ?Eq. 16?  and later on the page you use the style ?(18)?}

	\response{Caption of figure 3 needs an extra comma depending on the meaning you are trying to convey.
?For all distributions *,*  pushing at $\mu = 1$ ...?}

	\response{I don?t think the subsection headings should end in ?:?}

	\response{The annotations in Figures 4-5 are small and hard to see.}

	\response{Section 6 paragraph 1:  The word Sections should be capitalized when referring to previous sections by number. }

	\response{Should Kilobots be capitalized? (proper noun)}

	\response{Reference section:   the way you cite ICRA is inconsistent.}

\end{document}

### 2018-07-12 ICRA19/covControl.tex
so many changes to the dissertation, and Other!

--- 2018-06-11 Shiva Dissertation/chap4/covControl.tex
@@
 \bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \\
+\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \text{, }\\
 %\end{align}
@@
 \sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\\
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\text{, and}\\
 %\end{align}
@@
 \sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}
+\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}.
 \end{align}
@@
 \bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
-	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
+	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}.
 \end{align}
@@
 \begin{align}
-r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}
+r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}.
 \end{align}
@@
 F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
-\text{and } N &= F \cos(\theta) \nonumber
+\text{and } N &= F \cos(\theta) .\nonumber
 \end{align}
@@
 Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.
+Force in such a system can be calculated as
 \begin{align}
@@
 F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
-0, & \text{else} \end{cases}
+0, & \text{else} \end{cases}.
 \end{align}
@@
  , & \frac{1}{2} \le A\le 1
-\end{cases}
+\end{cases}.
 \end{align}
@@
 \begin{align}
-r_{\text{left}} &= \min (L,1-w)\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right)\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right)
+r_{\text{left}} &= \min (L,1-w),\nonumber \\
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right) \text{and}\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right).
 \end{align}
@@
 &+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
-&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy \nonumber
+&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy .\nonumber
 \end{align}

### 2018-07-12 ICRA19/relatedWork.tex
so many changes to the dissertation, and Other!

--- 2018-06-11 Shiva Dissertation/chap4/relatedWork.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}\label{sec:RelatedWork}
+\section{Related Work}\label{sec:RelatedWork}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This chapter focuses on centralized techniques that apply the same control input to both particles. 
+This paper focuses on centralized techniques that apply the same control input to both particles. 
 Precision control requires breaking the symmetry caused by the uniform input.  
@@
 \cite{Denasi2018independent}, that exploited nonlinearities generated by four magnetic coils in close proximity to the workspace to achieve trajectory control of two microspheres,
- this chapter requires only a controllable constant gradient in orthogonal directions to position the particles.
+ this paper requires only a controllable constant gradient in orthogonal directions to position the particles.
 % Systems like this one are poorly suited for PRM and RRT*-type methods\cite{lavalle2006planning} because if during a movement a collision occurs, that movement is irreversible.
@@
  Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
-  Instead, this chapter employs a greedy search strategy.
+  Instead, this paper employs a greedy search strategy.

### 2018-07-12 ICRA19/theory.tex
so many changes to the dissertation, and Other!

--- 2018-06-12 Shiva Dissertation/chap3/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Torque control}
-\label{sec:theory}
+%\section{Torque control}
+%\label{sec:theory}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We derive inspiration from recent work on pulling with a swarm \cite{pulltogether}. The contribution of this work is to map swarm distributions to torque production.
+%We derive inspiration from recent work on pulling with a swarm \cite{pulltogether}. The contribution of this work is to map swarm distributions to torque production.
 %The orientation of an object's major axis is important when a swarm is manipulating a non-symmetric object through narrow corridors. 
 %Orientation is controllable by applying torque to the object. 
-To change the output torque $\tau$ in~\eqref{eq:torque}, we can choose the direction and magnitude of the force applied $F$, and the moment arm from the object's center of mass (COM) $O$ to the point of contact $P$.
-We define a coordinate frame rooted at the COM, with the $x$-axis parallel to the object's longest axis. The resulting torque is:
+%To change the output torque $\tau$ in~\eqref{eq:torque}, we can choose the direction and magnitude of the force applied $F$, and the moment arm from the object's center of mass (COM) $O$ to the point of contact $P$.
+%We define a coordinate frame rooted at the COM, with the $x$-axis parallel to the object's longest axis. The resulting torque is:
 %
-\begin{equation}
-\tau = F \times (P - O ).\label{eq:torque}
-\end{equation}
+%\begin{equation}
+%\tau = F \times (P - O ).\label{eq:torque}
+%\end{equation}
 %
-We assume that the forces produced by individual particles add linearly. 
- As \cite{pulltogether} indicates, this is often a simplification of the true dynamics. 
- The swarm version of \eqref{eq:torque} is the summation of the forces contributed by $n$ individual particles:
-\begin{align}
-\tau_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O ),   \label{eq:swarmtorque}\\
-F_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i.  \label{eq:swarmforce}
-\end{align}
-Here $F_i$ is the force that the $i^{\textrm{th}}$ particle applies. 
-Not all particles are in contact with the object.  
-$\rho_i$ is an indicator variable: $\rho_i=1$ if the particle is in direct contact with the object or touching a chain of particle where at least one particle is in contact with the object. 
- Otherwise $\rho_i = 0$.
-The moment arm is the particle's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$. 
- If all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ equals some constant.
+%We assume that the forces produced by individual particles add linearly. 
+% As \cite{pulltogether} indicates, this is often a simplification of the true dynamics. 
+% The swarm version of \eqref{eq:torque} is the summation of the forces contributed by $n$ individual particles:
+%\begin{align}
+%\tau_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O ) \textrm{      and}   \label{eq:swarmtorque}\\
+%F_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i.  \label{eq:swarmforce}
+%\end{align}
+%Here $F_i$ is the force that the $i^{\textrm{th}}$ particle applies. 
+%Not all particles are in contact with the object.  
+%$\rho_i$ is an indicator variable: $\rho_i=1$ if the particle is in direct contact with the object or touching a chain of particle where at least one particle is in contact with the object. 
+% Otherwise $\rho_i = 0$.
+%The moment arm is the particle's position $P_i$ to the object's COM $O=[O_x,O_y]^{\top}$. 
+% If all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ equals some constant.
 %
-\section{Instantiating torque from swarm distribution}\label{sec:torqueDist}
-%
-Consider a swarm of particles whose marginal distribution along $x$ has probability density $p(x)$, where $x$ is defined as perpendicular to the object's long axis.
-This section considers three canonical probability distributions: uniform, triangular, and normal, all parameterized by mean $\mu$ and standard deviation $\sigma$. 
-They are plotted in Fig.~\ref{fig:PDF} and described by:
-%
-%uniform distribution
-\begin{align}
-p_u(x) &=  \left\{
-\begin{array}{ll}
-    \frac{1}{2\sqrt{3}\sigma}, &  \textrm{for   } \mu-\sqrt{3}\sigma \leq x \leq \mu+\sqrt{3}\sigma\\
-     0, & \textrm{otherwise}\\
-\end{array} 
-\right. ,\\
-%triangular distribution
-p_t(x) &=  \left\{
-\begin{array}{ll}
-    \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
-     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
-     0, & \textrm{otherwise}\\
-\end{array} 
-\right. ,\\
-%normal distribution
-p_n(x) &= \frac{1}{{\sigma \sqrt {2\pi } }}e^{{{ - \left( {x - \mu } \right)^2 } \mathord{\left/ {\vphantom {{ - \left( {x - \mu } \right)^2 } {2\sigma ^2 }}} \right.
-\kern-\nulldelimiterspace} {2\sigma ^2 }}}.
-\end{align}
-The next sections examine where to steer the mean of the probability distribution to maximize torque.
-We discuss two problems: pivoted object torque and free object torque.
-All the results are summarized in Table~\ref{tbl:MaximizingTorque}, and the calculations are included in the Mathematica file in the multimedia attachment.
-%
-%%%%%%%%%%% HUGE TABLE %%%%%%%%%%%%%%%%%%%%
-
-\begin{table}
-\begin{center}
-	\caption{Main results from Section \ref{sec:torqueDist} for maximizing torque with three common distributions. \label{tbl:MaximizingTorque} }
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{ c rl  rl }
-Distribution  & \multicolumn{2}{c}{The best $\mu$ to push} & \multicolumn{2}{c}{Maximum possible torque}\\ [0.5ex] 
-\hline \hline
-Pivoted Uniform & $
-\mu_{u_\max}=$&$\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &    \,~~\qquad \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
-1-\sqrt{3}\sigma<\mu<\sqrt{3}\sigma &    \,~~\qquad \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right.
-$& $
-\tau_{u_{\max}} =$&$\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &  ~~\quad\qquad\qquad \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
-\frac{1}{4 \sqrt{3}\sigma} &~~\quad\qquad\qquad \textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right.
-$\\ 
-\hline
-Pivoted Triangular &$\mu_{t_\max}=$&$ \left\{
-\begin{array}{ll}
- \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \qquad\qquad\textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
- \frac{\sqrt{2}}{2} &   					\qquad\qquad\textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
-\end{array} 
-\right. $&$ \tau_{t_\max} =$&$\left\{
-\begin{array}{ll}
-\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
-\frac{\sqrt{2}-2+3\sqrt{6}\sigma}{36\sigma^2} &   \textrm{for     }\sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right. $\\  
-\hline \hline
-Free Uniform & $\mu_{u_{f_\max}}=$&$\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &  ~\qquad\qquad\qquad\qquad \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
-\sqrt{3}\sigma &    ~\qquad\qquad\qquad\qquad \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right.$ & $\tau_{u_{f_{\max}}} =$&$\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &   ~~\quad\qquad\qquad\textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
-\frac{1}{4 \sqrt{3}\sigma} & ~~\quad\qquad\qquad\textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right.$\\
-\hline
-Free Triangular & $\mu_{t_{f_\max}}=$&$ \left\{
-\begin{array}{ll}
- \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     }  \sigma  <\frac{2}{\sqrt{6}} \\
-1<\mu<  \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma&   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}} 
-\end{array} 
-\right. $& $\tau_{t_{f_\max}} =$&$\left\{
-\begin{array}{ll}
-\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma < \frac{2}{\sqrt{6}}\\
-\frac{1}{9\sigma^2} &   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}}\\
-\end{array} 
-\right.
-$\\%[0.5ex]
-\hline
-\end{tabular}
-}
-\vspace{0.5em}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%% END HUGE TABLE %%%%%%%%%%%%%%%%
-%
-\subsection{Pivoted object torque}
-%
-In this problem, the torque applied to a rod of length 1 pivoted at 0 when $\theta = 0$ is
-\begin{equation}
-\tau_{p} = \int_0^1 x\,p(x)\, dx.
-\end{equation}
-
-\paragraph{Defining torque for distributions}
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{Uniform.pdf}%\put(1,55){a)}
-\end{overpic}
-\begin{overpic}[width =\figwid]{Triangular.pdf}
-\end{overpic}
-\begin{overpic}[width =\figwid]{Normal.pdf}
-\end{overpic}
-\vspace{-0.5em}
-\caption{\label{fig:PDF} Three distributions are examined in this work: uniform, triangular, and normal. Each is plotted above with $\mu=1$ for representative $\sigma$ values.
-%\vspace{-2em}
-}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\renewcommand{\figwid}{0.66\columnwidth}
-\begin{overpic}[width =\figwid]{TorqueUni.pdf}\put(1,30){Uniform}
-\end{overpic}
-\begin{overpic}[width =\figwid]{TorqueTri.pdf}\put(1,30){Triangular}
-\end{overpic}
-\begin{overpic}[width =\figwid]{TorqueNormal.pdf}\put(1,30){Normal}
-\end{overpic}
-\vspace{0.5em}
-\caption{\label{fig:torque} Torque as a function of mean position $\mu$.
-Mean position is the pushing location along a rod extending from 0 to 1.
-For all distributions pushing at $\mu = 1$ is not optimal unless $\sigma = 0$ or the swarm is uniformly distributed with $\sigma > \frac{1}{2\sqrt{3}}$.
-%\vspace{-2em}
-}
-\end{figure*}
-
-First,  for simplicity of the following derivations, the lower $l$ and upper bound $u$ are defined for the uniform distribution as
-%defining torque uniform distribution
-\begin{align}
-l &= \max(0,\mu -\sqrt{3} \sigma) \, , u = \min({1,\mu+\sqrt{3}\sigma}).
-\end{align}
-Torque by a uniformly distributed swarm is
-\begin{align}
-\tau_{u_p} &=  \frac{u^2-l^2}{4\sqrt{3}\sigma}. %\textrm{  for    }  u>l
-\end{align}
-To simplify derivations for the triangular distribution, we define the bounds of integration as
-%define torque for triangular:
-\begin{align}
-l_1 &= \max({0,\mu-\sqrt{6}\sigma})\, , l_2 = \max({0,\mu}),\\ \nonumber
-u_1 &= \min({1,\mu})\, , u_2 = \min({1,\mu+\sqrt{6}\sigma}). \nonumber
-\end{align}
-Torque by a triangularly distributed swarm is
-\begin{align}
-\tau_{t_p} &= \frac{\tau_{l} + \tau_{r}}{36\sigma^2},
-\end{align}
-where $\tau_{l}$ and $\tau_{r}$ are defined as:
-\begin{align}
-\tau_{l} &= {l_1}^2(2l_1-3(\mu-\sqrt{6}\sigma)) \\ \nonumber
-&+{u_1}^2(2u_1 - 3(\mu-\sqrt{6}\sigma)), \\ \nonumber%& \textrm{for     } u_1 > l_1\\ \nonumber
-\tau_{r} &= {l_2}^2(2{l_2}-3(\mu+\sqrt{6}\sigma))\\ \nonumber
-&-{u_2}^2(2u_2 - 3(\mu+\sqrt{6}\sigma)).\\ \nonumber % & \textrm{for     } u_2 > l_2\\ \nonumber
-\end{align}
-Torque by a normally distributed swarm is:
-%defining torque normal distribution
-\begin{align} \label{eq:normalTorque}
-\tau_{n_p} &= \frac{\left(e^{\frac{\mu^2}{2\sigma^2}}-e^{\frac{(1-\mu)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\ \nonumber
-&+ \frac{\mu}{2}\left(\erf\left(\frac{1-\mu}{\sqrt{2}\sigma}\right)+\erf\left(\frac{\mu}{\sqrt{2}\sigma}\right)\right) .
-\end{align}
-The $\erf(\cdot)$ is the error function. Torque is plotted as a function of $\mu$ for representative $\sigma$ values in Fig.~\ref{fig:torque}.
-\paragraph{Maximum torque for distributions}
-For a uniformly distributed swarm, the mean position that maximizes torque on a pivoted object is:
-% best place to push for uniform
-\begin{align}\label{eq:maxMuPivot}
-\mu_{u_\max}=\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
-1-\sqrt{3}\sigma<\mu<\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right. .
-\end{align}
-
-As soon as the width of the uniformly distributed swarm is longer than the rod ($\sigma > \frac{1}{2\sqrt{3}}$), there exists a range of optimal solutions: any $\mu$ such that the swarm covers 0 to 1.
-
-The resulting torque is:
-\begin{align}
-\label{eq:MaxTorqueUniformPivot}
-\tau_{u_{\max}} =\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
-\frac{1}{4 \sqrt{3}\sigma} & \textrm{for     }  \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right. .
-\end{align}
-%best place to push for triangular:
-For a triangularly distributed swarm, the mean position that maximizes torque is:
-
-\begin{align}
-\mu_{t_\max}= \left\{
-\begin{array}{ll}
- \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     } 0< \sigma  <\frac{1}{2\sqrt{3}} \\
- \frac{\sqrt{2}}{2} &   \textrm{for     } \sigma \geq \frac{1}{2\sqrt{3}} 
-\end{array} 
-\right. .
-\end{align}
-
-Thus the torque is:
-\begin{align}\label{eq:triMaxTorque}
-\tau_{t_\max} =\left\{
-\begin{array}{ll}
-\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma \leq \frac{1}{2\sqrt{3}}\\
-\frac{\sqrt{2}-2+3\sqrt{6}\sigma}{36\sigma^2} &   \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right. .
-\end{align}
-
-%best place to push for normal distribution:
-For all distributions, the $\mu$ that maximizes torque if $\sigma = 0$ is 1.
-For the $\tau_{p}$ case, the optimal $\mu$ moves in the $-x$ direction, and for some distributions approaches a limit as $\sigma$ increases.
-This limit is $\mu = \frac{\sqrt{2}}{2}$ for a triangularly distributed swarm.
-%We used L'H\^opital's rule and found third derivative of the torque to find the limit when standard deviation goes to infinity.
-It was because in the first derivative numerator and denominator were both going to infinity and we were not able to calculate the limit.
-The result is:
-In a normal distribution, the first derivative of \eqref{eq:normalTorque} is:
-\begin{align}
-\frac{d_{\tau}}{d_{\mu}}= \frac{1}{2}\left(\erf{\frac{1-\mu}{\sqrt{2}\sigma}}+ \erf{\frac{\mu}{\sqrt{2}\sigma}}\right)- \frac{e^{\frac{-(\mu-1)^2}{2\sigma^2}}}{\sqrt{2\pi}\sigma}.
-\end{align}
-
-The limit as $\sigma\to\infty$ is $\frac{2 - 3 \mu}{\sigma^3 6 \sqrt{2 \pi}}$ which goes to zero as the swarm variance increases.
-However, the zero of the numerator is $\mu = \frac{2}{3}$ which means as $\sigma$ increases the best position to push the rod asymptotically reaches $\mu = \frac{2}{3}$.
-
-The central limit theorem explains why systems where all the particles have independent noise form a normal distribution.
-However, optimizing torque for a normal distribution involves two error functions ($\erf$).
-Instead, we have solved all the equations for the triangular distribution.
-The numerical study illustrated in the left plot of Fig.~\ref{fig:bestLoc} shows the triangular and normal distributions perform similarly.
-
-%\begin{align}
-%\lim_{\sigma\to\infty} \frac{d^3\tau_n}{d\mu}= \frac{3\mu}{2}-1
-%\end{align}
-
-%which means the best position to push the rod moves left until it reaches $\mu = \frac{2}{3}$.
-
-%pivoted comparisons:
-\begin{figure*}
-\centering
-%\renewcommand{\figwid}{0.66\columnwidth}
-%\hbox{\hspace{-0.5em} \includegraphics[scale=0.6]{bar}}
-\begin{overpic}[width=0.4\columnwidth]{PivotDrawing.pdf}
-\end{overpic}\\
-\begin{overpic}[width =0.75\columnwidth]{BestLocationPivot.pdf}%\put(1,55){a)}
-\end{overpic} 
-\hbox{\hspace{1.5em} \begin{overpic}[width =0.68\columnwidth]{MaxTorquePivot.pdf}%\put(1,55){a)}
-\end{overpic}}
-%\vspace{-0.5em}
-\caption{\label{fig:bestLoc} Best location to push and maximum torque plots for a pivoted object of length 1, pivoted at 0. Generating code is in the attachment.
-%\vspace{-2em}
-}
-\end{figure*}
-
-\subsection{Free object torque}
-In this problem, the torque applied to a free rod of length 2 from $[-1,0]$ to $[1,0]$, is:
-\begin{equation}
-\tau_{f} = \int_{-1}^1 x\,p(x)\, dx .
-\end{equation}
-\emph{a) Defining torque to a free rod for distributions:}
-
-To calculate free object torque for a uniformly distributed swarm, for simplicity of derivations, $l_f$ and $u_f$ are defined: 
-%torque uniform free
-\begin{align}
-l_f &= \max({-1,\mu -\sqrt{3} \sigma}),\, u_f = \min({1,\mu+\sqrt{3}\sigma}).
-\end{align}
-
-The torque applied by a uniformly distributed swarm is:
-\begin{align}
-\tau_{u_f} &= \frac{u_f^2-l_f^2}{4\sqrt{3}\sigma} \textrm{  for    }  u_f>l_f .
-\end{align}
-
-%torque triangular free
-To simplify derivations for a triangularly distributed swarm, we define bounds of integration as:
-\begin{align}
-l_{f_1} &= \max({-1,\mu-\sqrt{6}\sigma}), \,l_{f_2} = \max({-1,\mu}),\\ \nonumber
-u_{f_1} &= \min({1,\mu}), \, u_{f_2} = \min({1,\mu+\sqrt{6}\sigma}) . \nonumber
-\end{align}
-The torque applied by a triangularly distributed swarm is:
-\begin{align}
-\tau_{t_f} = \frac{\tau_{f_l}+ \tau_{f_r}}{36\sigma^2},
-\end{align}
-where $\tau_{f_l}$ and $\tau_{f_r}$ is defined as
-\begin{align}\nonumber
-\tau_{f_l} &=  {u_{f_1}}^2(2u_{f_1} - 3(\mu-\sqrt{6}\sigma)) \\ \nonumber
-&-{l_{f_1}}^2(2{l_{f_1}}-3(\mu-\sqrt{6}\sigma)) , \\ \nonumber %&   \textrm{for     } u_{f1} > l_{f1}\\ \nonumber
-\tau_{fr} &=-{u_{f_2}}^2(2u_{f_2} -3(\mu+\sqrt{6}\sigma) , \\ \nonumber
-&+ {l_{f_2}}^2(2{l_{f_2}}-3(\mu+\sqrt{6}\sigma)) . \\ \nonumber%&   \textrm{for     } u_{f2} > l_{f2}\\ \nonumber
-\end{align}
-
-%torque normal free
-The torque applied by a normally distributed swarm is:
-\begin{align} \nonumber
-\tau_{n_f} &= \frac{\left(e^{\frac{-(\mu+1)^2}{2\sigma^2}}-e^{\frac{-(\mu-1)^2}{2\sigma^2}}\right)\sigma}{\sqrt{2\pi}}\\
- &+ \frac{1}{2}\mu\left(\erf\Big(\frac{1-\mu}{\sqrt{2}\sigma}\Big)+\erf\Big(\frac{1+\mu}{\sqrt{2}\sigma}\Big)\right) .
-\end{align}
-
-%free torque comparison
-\begin{figure*}
-\begin{center}
-%\renewcommand{\figwid}{0.49\columnwidth}
-\begin{overpic}[width =0.4\columnwidth]{FreeDrawing.pdf}
-\end{overpic}\\
-\begin{overpic}[width =0.75\columnwidth]{TorqueFreeBestPush.pdf}
-\end{overpic}
-\hbox{\hspace{1.75em}\begin{overpic}[width =0.68\columnwidth]{TorqueFreeComparison.pdf}
-\end{overpic}}
-
-%\vspace{-0.5em}
-\caption{\label{fig:maxTorque} Best location to push and maximum torque plots for a free object of length 2, located from $x=-1$ to $1$.
-%\vspace{-2em}
-}
-\end{center}
-\end{figure*}
-
-\emph{b) Maximum torque for distributions}
-Maximum torque of a uniformly distributed swarm to a free object has the same upper solution as with a pivoted object:
-
-\begin{align}\label{eq:maxMuFree}
-\mu_{u_{f_\max}}=\left\{
-\begin{array}{ll}
-1-\sqrt{3}\sigma &   \textrm{for     } \sigma < \frac{1}{2\sqrt{3}}\\
-\sqrt{3}\sigma &   \textrm{for     } \sigma > \frac{1}{2\sqrt{3}}\\
-\end{array} 
-\right. .
-\end{align}
-
-The major difference is the existence of only one solution for $\sigma$ values for torque on a free object.
-The left bound of the distribution must never be less than 0, or torque applied left of the center will cancel torque to the right of the center.
- The maximum torque produced is given by \eqref{eq:MaxTorqueUniformPivot}.
-%The difference is because the pivoted object even when the swarm is pushing all the length of the object, they are applying torque, but for the free object, if the swarm pushes the whole length of the object, they would just apply force.
-
-For a triangularly distributed swarm the optimal mean is:
-
-\begin{align}
-\mu_{t_{f_\max}}= \left\{
-\begin{array}{ll}
- \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma &   \textrm{for     }  \sigma  <\frac{2}{\sqrt{6}} \\
-1<\mu<  \sqrt{12\sigma^2 +1}  -\sqrt{6}\sigma&   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}} 
-\end{array} 
-\right. .
-\end{align}
-%limits and equations
-
-As with the uniform distribution in \eqref{eq:maxMuPivot} the optimal solution with $\sigma > \frac{2}{\sqrt{6}}$ has a range of solutions.
-Setting $\mu=1$ maximizes force for the optimal torque.
-However, moving the mean right reduces the negative torque applied by particles to the left of 0, but this gain is canceled by a corresponding reduction in positive torque.
-
-Therefore, the maximum torque is:
-
-\begin{align}\label{eq:triFreeMaxTorque}
-\tau_{t_{f_\max}} =\left\{
-\begin{array}{ll}
-\frac{ (1+12\sigma^2)^{\frac{3}{2}}-1}{18\sigma^2} - \sqrt{6}\sigma &   \textrm{for     } \sigma < \frac{2}{\sqrt{6}}\\
-\frac{1}{9\sigma^2} &   \textrm{for     } \sigma \geq \frac{2}{\sqrt{6}}\\
-\end{array} 
-\right. .
-\end{align}
-
-Again, for a normally distributed swarm, we have numerically found the maximum torque and optimal pushing location $\mu$.
-The results are closely approximated by the upper bound solution for the triangular system.
- The results are plotted in Fig.~\ref{fig:maxTorque}.
 
@@
 % for the media, see \cite{angleofRepose}.
+ By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
+ We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
  \begin{figure}
@@
 \end{overpic}
-\caption{\label{fig:Angle} Different values of angle of repose is shown when granular particles move faster than the rod.
+\caption{\label{fig:angle} Different values of angle of repose is shown when granular particles move faster than the rod.
 }
 \end{figure}
-
-By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
- We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
 By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:
@@
 \frac{-\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos\alpha\sin{\theta}} &   -\alpha<\theta<\alpha\\
-0 &    \textrm{otherwise}\\
+0 &    \textrm{otherwise .}\\
 \end{array} 
-\right.
+\right .\\
 \end{align}
@@
 The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}. 
-Torque will also be similarly defined as:
+Torque will also be similarly defined as
 \begin{align}
@@
 \frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin{\theta} }{48\sin^2\alpha}&   -\alpha<\theta<\alpha\\
-0 &    \textrm{otherwise}\\
+0 &    \textrm{otherwise .}\\
 \end{array} 

### 2018-07-12 Qualifying Exam/chap/abs.tex
so many changes to the dissertation, and Other!

% abstract for thesis

%This is the abstract.  It has a limit of 150 words for a masters thesis and 350 words for a PhD dissertation.

Micro-robots are small enough to move through the passageways of the body, therefore they are suited for targeted drug delivery and micro-scale manufacturing. Due to their small size, a single robot does not have enough force to deliver payloads, and it is prohibitively difficult to have onboard computation. Therefore, these robots are usually controlled by global inputs such as a uniform external magnetic field. This thesis presents controllers and algorithms for steering such an under-actuated swarm. This work first proves that the mean position of the swarm is controllable, and shows how an obstacle can make the variance controllable. Then it derives automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the swarm distribution. Finally, this work uses friction with boundary walls to break the symmetry caused by the global input and uses it to steer two particles to arbitrary positions.

### 2018-07-12 Qualifying Exam/chap/conclusion.tex
so many changes to the dissertation, and Other!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Conclusion]{Conclusion and Future Work}
\label{chap-conc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    Inspired by large-scale human experiments with swarms of robots under global control,  this thesis investigated controllers that use only the mean and variance of a robot swarm. We proved that the mean position is controllable, and provided conditions under which variance is controllable.  We derived automatic controllers for each and a hysteresis-based switching control that controls the mean and variance of a robot swarm.  We employed these controllers as primitives for a block-pushing task. 
%These parts were done as my masters work, and has been published or submitted into two conference papers and is in draft in one IJRR journal paper. Fig.~\ref{fig:timeline} shows the timeline plan I have. The green parts in the timeline has already been done, and I will be working on the others during my PhD years.

%    \begin{figure}
%\centering
%\begin{overpic}[width=1.0\columnwidth]{ProjectTimeline.pdf}\end{overpic}
%%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%%Please use the same color and line style for the variance min and max as you use in Figure 4.
%%}
%%{blockpushingImageWithMeanAndVarianceOverlay.png}
%\caption{\label{fig:timeline} This is a timeline for my plan for research. The green parts are already done and presented in this document, I would like to pursue the others for my PhD. }
%\end{figure}

  \section{Future Work: Gathering Problem}
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction. These activities require robots that behave intelligently. Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population. It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. 
However, it is often possible to sense global properties of the group, such as mean position and variance. In our experiments, we showed that controlling mean is possible, and controlling variance with some constraints is possible. One of the constraints was that we had assumed that the swarm is clustered together and the swarm was trying to keep the variance small. If we did not assume that and we had a world like Fig.~\ref{fig:sshaped} or if during the experiment, a fraction of the swarm makes a new branch and put the swarm to two or more clusters by hitting the obstacles in their path, we could not gather them together again, and our algorithm for controlling variance was failed. So the next problem is to design an algorithm that will gather all the robots to an specific region with global input so that we can have the minimum variance. \\

For a \emph{S} shaped maze it is easy to implement. Alg. ~\ref{alg:S-Shaped} shows how to gather all the particles of the swarm when we have a \emph{S} shaped maze. With the same algorithm, we have a  solution for a mirrored \emph{S} shaped maze. It is not an optimal way for a \emph{T} shaped map also, but works for it. The problem is: Is there a general way to gather all the particles in \emph{any} map?
\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{SShaped.png}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:sshaped} If the robots were not grouped initially, how can we gather them all together and what are the features of the map which is gather-able? }
\end{figure}
\begin{algorithm}
\caption{Gather Robots in One Corner of an S-shaped or inverse S-shape Maze}\label{alg:S-Shaped}
\begin{algorithmic}[1]
\Ensure   The maze is \emph{S}-shaped
\While{$\sigma_x^2 > \sigma_{min}^2 \vee \sigma_y^2 > \sigma_{min}^2$}
\State Go Maximal Left
\State Go Maximal Down
\State Go Maximal Right
\State Go Maximal Down
\EndWhile
\end{algorithmic}
\end{algorithm}

\section{Future Work: Torque Control}

Another remaining problem is that sometimes we do not want the block to rotate during the experiment, or we want to rotate it to some angle. The problem statement is how to control torque of the object, and discussing what features of the object make torque controllable. Fig.~\ref{fig:torque} shows an example of torque control.
    \begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{torque.png}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:torque} How can we control torque of the object with a swarm of robots and a shared input? }
\end{figure}
    % We should also control the covariance $\sigma_xy$ and higher moments of the distribution

%Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011}, scientists fasten molecules that fluoresce light when activated by a strong light source. Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. Photobleaching can be minimized by lowering the excitation light intensity, but this increases the probability of missed detections~\cite{Cazes2001}. A control methodology based on statistics of the robot swarm rather than the actual position of each robot, allows relaxing demands on imagine systems, controllers robust to tracking errors, and a simpler methodology.  In this work we...
%

% Additionally, as population characteristics, they are available even if only a percentage of the robots are detected each control cycle.
%Photobleaching: http://www.piercenet.com/browse.cfm?fldID=4DD9D52E-5056-8A76-4E6E-E217FAD0D86B
%
%Photobleaching is caused by the irreversible destruction of fluorophores due to either the prolonged exposure to the excitation source or exposure to high-intensity excitation light. Photobleaching can be minimized or avoided by exposing the fluor(s) to the lowest possible level of excitation light intensity for the shortest length of time that still yields the best signal detection; this requires optimization of the detection method using high sensitivity CCD cameras, high numerical aperture objective and/or the widest bandpass emission filter(s) available. Other approaches include using fluorophores that are more photostable than traditional fluorophores and/or using antifade reagents to protect the fluor(s) against photobleaching. Steps to avoid photobleaching are not feasible for all detection methods and should be optimized for each method used. For example, antifade reagents are toxic to live cells, and therefore they can only be used with fixed cells or tissue. Furthermore, some detection methods, such as flow cytometry, normally do not require steps to avoid photobleaching because of the extremely short exposure time of the fluorophore to the excitation source.

### 2018-07-12 Qualifying Exam/chap/intro.tex
so many changes to the dissertation, and Other!

\chapter[Introduction]{Introduction}
\label{chap-intro}

\section{Shaping a Swarm Using a Shared Control Input}\label{sec:Intro}
Micro- and nanorobotics can be manufactured in large numbers.
Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite incredible under-actuation.  
Additionally, it is not always possible to gather pose information on each robot for feedback control. 
Robots might be difficult or impossible to sense individually due to their size and location. 
For example, micro-robots are smaller than the minimum resolution of a clinical MRI-scanner~\cite{martel2014computer}.
However, it is often possible to sense global properties of the group, such as mean position and variance. 
To make progress in automatic control with global inputs, we present swarm manipulation controllers requiring only mean and variance measurements of the robot's positions.  These controllers are used as primitives to perform a block-pushing task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.
\begin{figure}
\centering
\begin{overpic}[width=1.0\columnwidth]{BlockBotView2.eps}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.  Here is a mockup of a swarm of hardware robots(kilobots) that pushes a green block toward the goal. %See video attachment~\cite{ShivaVideo2015}.
}
\end{figure}
A limitation was that variance control could only compress a swarm along the world $x$ and $y$ axes.  This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$. 
This thesis continues to prove  that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), implements these position control algorithms in simulation (Section \ref{sec:simulation}) and on a hardware setup with up to 64 robots (Section \ref{sec:expResults}), and ends with directions for future research.

\begin{figure}
\centering
\begin{overpic}[width=0.9\columnwidth]{Leaf2.jpg}\end{overpic}
\caption{\label{fig:vascularNetwork}%Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.  
Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging due to many obstacles. This thesis demonstrates how friction with walls can be used to change the shape of a swarm.} %TODO: save this as a pdf
\end{figure}

%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles.

\begin{figure}[t]
\centering
\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
\vspace{-1em}
\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
}\vspace{-1em}
\end{figure}

% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

### 2018-07-12 Qualifying Exam/conclusion.tex
so many changes to the dissertation, and Other!

--- 2018-07-12 Qualifying Exam/chap/conclusion.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter[Conclusion]{Conclusion and Future Work}
-\label{chap-conc}
+\section{Conclusion and Future Work}\label{sec:conclusion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-  \section{Future Work: Gathering Problem}
+  \subsection{Gathering Problem}
 Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction.
@@
 One of the constraints was that we had assumed that the swarm is clustered together and the swarm was trying to keep the variance small.
-If we did not assume that and we had a world like Fig.~\ref{fig:sshaped} or if during the experiment, a fraction of the swarm makes a new branch and put the swarm to two or more clusters by hitting the obstacles in their path, we could not gather them together again, and our algorithm for controlling variance was failed.
+If we did not assume that and we had a world like Fig.
+~\ref{fig:sshaped} or if during the experiment, a fraction of the swarm makes a new branch and put the swarm to two or more clusters by hitting the obstacles in their path, we could not gather them together again, and our algorithm for controlling variance was failed.
 So the next problem is to design an algorithm that will gather all the robots to an specific region with global input so that we can have the minimum variance.
@@
 
-\section{Future Work: Torque Control}
+\subsection{Torque Control}
 
-Another remaining problem is that sometimes we do not want the block to rotate during the experiment, or we want to rotate it to some angle.
-The problem statement is how to control torque of the object, and discussing what features of the object make torque controllable.
+Another problem that still remains, is that sometimes we don't want the block to rotate during the experiment, or we want to rotate it in some angles.
+The problem statement is how to control torque of the object, and what are the features of the object that torque is controllable?
 Fig.~\ref{fig:torque} shows an example of torque control.

### 2018-07-12 Qualifying Exam/experiment.tex
so many changes to the dissertation, and Other!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Block-pushing results}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section analyzes a \emph{block-pushing} task attempted by both our hybrid, hysteresis-based controller and by human users.

\subsection{Human-controlled block-pushing}

In previous work over 1000 human users completed an online version of this task using varying levels of feedback.
 The original experiment explored manipulation with varying amounts of sensing information: {\bf full-state} sensing showed the position of all robots; {\bf convex-hull} drew a convex hull around the outermost robots; {\bf mean} displayed the average position of the population; and {\bf mean + variance} added a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same robot swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ robots. Convex-hull requires at worst $2n$, but usually a smaller number.  Mean requires two, and variance three, data points.  Mean and mean + variance are convenient even with millions of robots. We hypothesized a steady decay in performance as the amount of visual feedback decreased.

\begin{figure}[b!]
\renewcommand{\figwid}{0.24\columnwidth}
\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
\caption{\label{fig:Visualization}Screenshots from a block-pushing task with human users. This experiment challenged players to quickly steer 100 robots (blue discs) to push an object (green hexagon) into a goal region. 
%\vspace{-1em}
}
\end{figure}

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth]{ResVaryVis.pdf}\end{overpic}
\vspace{-2em}
\caption{\label{fig:ResVaryVis} Completion-time results for the four levels of visual feedback shown in Fig.~\ref{fig:Visualization}. 
%\vspace{-2em}
}
\end{figure}

To our surprise, the results indicated the opposite: players with just the mean completed the task faster than those with full-state feedback.  As Fig.~\ref{fig:ResVaryVis} shows, the levels of feedback arranged by increasing completion time are [mean, mean + variance, full-state, convex-hull].  Interviews with  beta-testers suggests that tracking 100 robots was overwhelming---similar to schooling phenomenons that confuse predators---while working with just the mean + variance was like using a ``spongy'' manipulator. Convex-hull feedback was confusing and irritating because a single robot left behind an obstacle would distort the entire hull, obscuring the information about the majority of the swarm.
%obscuring what the rest of the swarm is doing.

\subsection{Automated block-pushing}
Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm. To solve this block-pushing task, we discretized the environment. On this discretized grid we used breadth-first search to determine $\mathbf{M}$, the shortest distance from any grid cell to the goal, and generated a gradient map $\nabla \mathbf{M}$ toward the goal as shown in Fig.~\ref{fig:BFSGradient}.  The block's center of mass is at $\mathbf{b}$ and has radius $r_b$. 
Three constants are needed, where $k_1>k_2>1$ and $1>k_2>0$. All experiments used $[k_1,k_2,k_3] = [2.5,1.5,0.1]$.
The robots were directed to assemble behind the block at  $\mathbf{b} - k_2 r_b \nabla \mathbf{M}(\mathbf{b})$, then move to  $\mathbf{b} - k_3 r_b \nabla \mathbf{M}(\mathbf{b})$ to push the block toward the goal location. We use the hybrid hysteresis-based controller in Alg.~\ref{alg:MeanVarianceControl}  to track the desired position, while maintaining sufficient robot density to move a block by switching to minimize variance whenever variance exceeds a set limit. The minimize variance control law \eqref{eq:PDcontrolVariance} is slightly modified to choose the nearest corner further from the goal than $\mathbf{b}$ with an obstacle-free straight-line path to $\mathbf{b}$. 
The control algorithm  for block-pushing is listed in Alg.~\ref{alg:BlockPushing}. 
Experimental results are summarized in Fig.~\ref{fig:AutoControlVaryN}.  Although larger populations of robots can apply more force, minimizing the variance requires more time with larger populations and dominates task completion time.

\begin{algorithm}
\caption{Block-pushing controller for a robotic swarm.}\label{alg:BlockPushing}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$,  moveable block's center of mass $\mathbf{b}$, map of the environment, and the locations of all convex corners $\mathbf{C}$
\Require Robot distribution is unimodal
\Require Obstacle-free, straight-line path from swarm to moveable block
\State Compute $\mathbf{M}$, the distance to goal, with breadth-first search
\State Compute the gradient, $\nabla \mathbf{M}$
\State $\mathbf{C} \gets \mathrm{sort(\mathbf{C})}$ according to $-\mathbf{M}$
\While{$\mathbf{b}$ is not in goal region}
\State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
\If {$\sigma^2 > \sigma_{max}^2$}
\While{$\sigma^2 > \sigma_{min}^2$}
\State $\mathbf{c}_i \gets$ the nearest corner in $\mathbf{C}$ to $[\bar{x},\bar{y}]$
\State $ [x_{goal}, y_{goal}] \gets \mathbf{c}_i $
\If {$\mathbf{M}(\mathbf{b}) > \mathbf{M} (\mathbf{c}_i)$}
\State  $[x_{goal}, y_{goal}] \gets  \mathbf{c}_{i-1}$ 
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndIf
\EndWhile
\Else  
\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
	\Else
	\State$r_p \gets k_3 r_b$  \Comment{pushing move}
	\EndIf
\State $[x_{goal}, y_{goal}] \gets \mathbf{b} - r_p \nabla \mathbf{M}(\mathbf{b})$ 
\EndIf
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndWhile
\end{algorithmic}
\end{algorithm}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.3\columnwidth}
\href{http://youtu.be/tCej-9e6-4o}{\begin{overpic}[width =\figwid]{story1.png}\put(6,15){T = 5 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story2.png}\put(6,15){T = 12 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story3.png}\put(6,15){T = 20 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story4.png}\put(6,15){T = 25 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story5.png}\put(6,15){T = 33 s}
\end{overpic}}
\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing the block-pushing experiment with 200 robots under automatic control.  See the video attachment for an animation~\cite{ShivaVideo2015}.}
%\vspace{-2em}
}
\end{figure*}

\begin{figure}
\centering
\begin{overpic}[scale=0.3]{BFSMode.png}
\end{overpic}
\begin{overpic}[scale=0.3]{GradientView.png}
\end{overpic}
\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable block (left), which is used to compute gradient vectors (right).
%\vspace{-2em}
}
\end{figure}

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth]{AutoControlVaryN.pdf}\end{overpic}
\vspace{-2em}
\caption{\label{fig:AutoControlVaryN} Completion-time results using the automatic controller from Alg.~\ref{alg:BlockPushing} for different numbers of robots.  Each bar is labelled with the number of trials.
}
\end{figure}

Algorithm \ref{alg:BlockPushing} is an imperfect solution and has a failure mode if the robot swarm becomes multi-modal with modes separated by an obstacle, as shown in Fig.~\ref{fig:Failure}.  In this case, moving toward a corner will never reduce the variance below $\sigma_{min}^2$.

  The first challenge is to identify when the distribution has become multi-modal.  Measuring just the mean and variance is insufficient to determine if a distribution is no longer unimodal, but if the swarm is being directed to a corner, and the variance does not reduce below $\sigma_{min}^2$, the swarm has become separated. In this case, we must either manipulate with a partial swarm, or run a gathering algorithm.  For the  `{\sffamily S}'-shaped workspace in this study, an open-loop input that commands the swarm to move in succession \{{\sc left, down, right, down}\} will move the swarm to the bottom right corner.
This is not true for all obstacle fields. In a  `{\sffamily T}'-shaped workspace, it is not possible to find an open-loop input that will move the entire swarm to the bottom of the `{\sffamily T}'.

  Using only the mean and variance may be overly restrictive.  Many heuristics using high-order moments have been developed to test if a distribution is multimodal~\cite{haldane1951simple}.  Often the sensor data itself, though it may not resolve individual robots, will indicate multi-modality.  For instance CCD images reveal clusters of bacteria, and MRI scans show agglomerations of particles~\cite{stuber2007positive}.  This data can be fitted with $k$-means or expectation maximization algorithms, and manipulation could be performed with the nearest swarm of sufficient size.

\begin{figure}
\centering
\begin{overpic}[scale=0.3]{FailureBlockPush.png}
\end{overpic}
\begin{overpic}[scale=0.3]{FailureBlockPushing}
\end{overpic}
\caption{Algorithm \ref{alg:BlockPushing} fails when some robots are separated by the maze and the swarm can not achieve $\sigma^2 < \sigma_{min}^2$.  These failures occured during 14\% of trials.\label{fig:Failure}
%\vspace{-2em}
}
\end{figure}

\subsection{Hardware system}

Our experiments are on centimeter-scale hardware systems called \emph{kilobots}.  These allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot \cite{Rubenstein2012,rubenstein2014programmable} is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available commercially or as an open source platform~\cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source. 
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=64 kilobots, a 1.5 m$\times$1.2 m whiteboard as the workspace, and four 30W LED floodlights arranged 1.5 m above the plane of the table at the $\{N,E,S,W\}$ vertices of a 6 m square centered on the workspace. The lights were controlled using an Arduino Uno board connected to an 8 relay shield board.  At top of the table, an overhead machine vision system was added to track the position of the swarm. Laser-cut patterns for our neon green fiducial markers and our {\sc Matlab} tracking code are available at our github repository~\cite{Shahrokhi2015GitHubShapeControl}.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{SetUp.pdf}
\end{center}
\caption{\label{fig:setup}
Hardware platform:  table with 1.5$\times$1.2 m workspace, surrounded by eight remotely triggered 30W LED floodlights, with an overhead machine vision system.
}
\end{figure}

The walls of the hardware platform have almost infinite friction, due to the three legged design of the kilobots. When a kilobot is steered into the wall, they pin themselves to the wall until the light changes direction and they begin turning in the other direction.  This wall friction is sufficient to enable independent position control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.

\begin{figure*}
\centering
\renewcommand{\figwid}{0.3\columnwidth}
{
\begin{overpic}[width =0.315\columnwidth]{twoR_1.pdf}\put(15,65){$t$  = 30 s}\end{overpic}\hspace{-.5em}
\begin{overpic}[width =\figwid]{twoR_2.jpg}\put(15,65){$t$  = 60 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR_3.jpg}\put(15,65){$t$  = 90 s}
\end{overpic}\\
\vspace{1em}
\begin{overpic}[width =\figwid]{twoR_4.jpg}\put(15,65){$t$  = 120 s}
\end{overpic}
\begin{overpic}[width =\figwid]{twoR_5.jpg}\put(15,65){$t$  = 150 s}
\end{overpic}}
\caption{\label{fig:storyReal}{Two robot positioning using the hardware setup and two kilobot robots.  The walls have nearly infinite friction, as illustrated by this fig, the robot with the blue path that is stopped by the wall until the light changes orientation, while the orange robot in free-space is unhindered.}
%\vspace{-2em}
}
\end{figure*}

To demonstrate covariance control $n=64$ robots were placed on the workspace and manually steered with a single light source, using friction with the boundary walls to vary the covariance from  -5000 to 10,000.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{experiment.pdf}
\end{center}
\caption{\label{fig:covExperiment}
Hardware demonstration steering 64 kilobot robots to desired covariance.  Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
}
\end{figure}

### 2018-07-12 Qualifying Exam/intro.tex
so many changes to the dissertation, and Other!

--- 2018-07-12 Qualifying Exam/chap/intro.tex
@@
-\chapter[Introduction]{Introduction}
-\label{chap-intro}
-
-\section{Shaping a Swarm Using a Shared Control Input}\label{sec:Intro}
+\section{Introduction}\label{sec:Intro}
 Micro- and nanorobotics can be manufactured in large numbers.
@@
 Here is a mockup of a swarm of hardware robots(kilobots) that pushes a green block toward the goal.
-%See video attachment~\cite{ShivaVideo2015}.
-}
+See video attachment~\cite{ShivaVideo2015}.}
 \end{figure}
@@
 Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$. 
-This thesis continues to prove  that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), implements these position control algorithms in simulation (Section \ref{sec:simulation}) and on a hardware setup with up to 64 robots (Section \ref{sec:expResults}), and ends with directions for future research.
+Because of this reason, this paper continues to prove  that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), implements these position control algorithms in simulation (Section \ref{sec:simulation}) and on a hardware setup with up to 64 robots (Section \ref{sec:expResults}), and ends with directions for future research.
+
+Micro- and nano-robots can be manufactured in large numbers.
+Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
+ For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.
+Instead, these robots are often controlled by a global, broadcast signal.
+ The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite incredible under-actuation.
 
@@
 \begin{overpic}[width=0.9\columnwidth]{Leaf2.jpg}\end{overpic}
-\caption{\label{fig:vascularNetwork}%Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.  
-Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging due to many obstacles.
-This thesis demonstrates how friction with walls can be used to change the shape of a swarm.} %TODO: save this as a pdf
+\caption{\label{fig:vascularNetwork}Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.
+Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging
+due to the many obstacles. This paper demonstrates how friction with walls can be used to change the shape of a swarm.} %TODO: save this as a pdf
 \end{figure}

### 2018-07-12 Qualifying Exam/simulation.tex
so many changes to the dissertation, and Other!

\section{Simulation}\label{sec:simulation}
% Simulation:
%%  Mean control results   (image and plot(s) )
%%  Variance Control      (image and plot(s) )
%%  Hybrid control   (image and plot(s) )

Our simulations use a Javascript port of \href{http://box2d.org/}{Box2D}, a popular 2D physics engine with support for rigid-body dynamics and fixed-time step simulation~\cite{catto2010box2d}.  All experiments ran on a Chrome web browser on a 2.6 GHz Macbook.  \href{https://github.com/aabecker/SwarmControlSandbox/blob/master/exampleControllers/BlockPushingIROS2015.html}{All code is available at}~\cite{Shahrokhi2015}.

\subsection{Controlling the mean position}

We control mean position with a PD controller that uses the mean position and mean velocity. Our control input is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\bar{v}_x) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\bar{v}_y)  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain. We performed a parameter sweep to identify the best values.  Representative experiments are shown in Fig.~\ref{fig:gainvalues}. 100 robots were used and the maximum speed was 3 meters per second. As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52 s with $K_{p}= 4$, and  $K_{d} = 1$. Fig.~\ref{fig:meanVideo} shows an example of controlling mean position by making it trace the word ``SWARM".

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth ]{gains.eps}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:gainvalues} Tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots.
%\vspace{-2em}
}
\end{figure}

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth* 2/3 ]{meanVideoSnapShot.png}
\end{overpic}
\vspace{0 em}
\caption{\label{fig:meanVideo} A frame from video attachment showing mean position control on a swarm of 200 robots. The mean position of the swarm traces ``SWARM".%~\cite{ShivaVideo2015}. 
%\vspace{-2em}
}
\end{figure}

\subsection{Controlling the variance}
\begin{figure}
\centering
\begin{overpic}[width = \columnwidth] {brownianWpublish.eps}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:varyBrownian} Increased noise results in more responsive variance control because stronger Brownian noise causes a faster increase of variance.
%\vspace{-2em}
}
\end{figure}

%cite the control law, explain experiment (number of robots, maximum speed, ).

For variance control we use the control law discussed in Section~\ref{sec:VarianceControl}.  Moving away from the wall and waiting is sufficient to increase variance because Brownian noise naturally disperses the swarm in such a way that the variance increases linearly~\cite{einstein1956investigations}.  If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board~\cite{Becker2013b}. %no need to worry about blocks in this controller.

The variance control law to regulate the variance to $\sigma^2_{ref}$ has three gains:
\begin{align}
u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}) \nonumber\\
u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
\end{align}
In a slight abuse of notation we call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{goal}(\sigma^2_{ref}) = r + \sqrt{3\sigma^2_{ref}}.
\end{align}
 If another wall is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.  Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.

%\todo{image showing control x variance and y-variance out of phase}

\subsection{Hybrid control of mean and variance}

Fig. ~\ref{fig:hybrid} shows a simulation run of the hybrid controller in Alg.~\ref{alg:MeanVarianceControl} with 100 robots in a square workspace containing no internal obstacles. Fig.~\ref{fig:videoVar} shows the experiment of running the hybrid controller.

%\todo{plot showing 1.5 cycles of mean position, and a variance goal.  We might need a longer time}
\begin{figure}
\centering
\begin{overpic}[width = \columnwidth]{meanVariance4.eps}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:hybrid} Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom). For ease of analysis, only $x$ position and variance are shown.
%\vspace{-2em}
}
\end{figure}

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth * 2/3]{VideoSnapShot.png}
\end{overpic}
\caption{\label{fig:videoVar} A frame from video, using Alg.~\ref{alg:MeanVarianceControl} to control variance and mean position of a swarm of 200 robots.%~\cite{ShivaVideo2015}.
%\vspace{-2em}
}
\end{figure}

Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, \ref{alg:YControl}, were implemented in Mathematica using point robots (radius = $0$).  All code is available at a public github repository~\cite{Shahrokhi2015GitHubShapeControl}.  Figs~\ref{fig:shapeControlMathematica1} and \ref{fig:shapeControlMathematica2} show the examples of the implementation of our algorithm. In both of these figures we have denoted the starting points and the destinations by small circles. However, destination points are surrounded by larger circles so as to be distinct from starting points.

In each of these figures we have five snapshots of the running of our algorithm taken every quarter second. For the sake of brevity we have replaced straight moves (e.g. upward, downward, etc) with oblique moves that shows a combination of two moves simultaneously (e.g. left and down together).

As we can see, in Fig.~\ref{fig:shapeControlMathematica1} $\Delta r_x$ is adjusted to $\Delta e_x$ in the second snapshot, i.e., at $t = t_1$ where $t_1<0.25$. The rest of the steps in this figure is dedicated to adjusting the $\Delta r_y$ to $\Delta e_y$. As it is clear from Fig.~\ref{fig:shapeControlMathematica1}, $\Delta r_y$ is also adjusted at $t = t_2$ where $0.75<t_2<1$. Finally, once $\Delta r_x$ and $\Delta r_y$ are adjusted, the algorithm gives a global input both of the robots so as to move them toward their corresponding destinations. This is happening in the time interval of $(t_2,1]$.

Similarly, in Fig.~\ref{fig:shapeControlMathematica2} we can see that the $\Delta r_x$ is adjusted in the third snapshot, i.e., at $t=t_3$ where $0.25<t_3<0.5$ and $\Delta r_y$ is adjusted in the last snapshot at $t=t_4$ where $0.75<t_4<1$. The final positioning steps are happening in the time interval of $(t_4,1]$.

As we pointed out earlier, adjusting each of $\Delta r_x$ and $\Delta r_y$ needs two iterations in the worst case. In other words, both of the Alg. \ref{alg:XControl} and Alg. \ref{alg:YControl} are executed two times in the worst case in positioning process of the robots. It is easy to see that we need two iterations of Alg. \ref{alg:XControl} only if $|\Delta e_x - \Delta s_x|>L$. Similarly we need two iterations of Alg. \ref{alg:YControl} only if $|\Delta e_y - \Delta s_y|>L$.

%\textcolor{red}{Shiva:  what is the complexity of this algorithm?  How many steps in the worst case?  When does worst case happen?}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.3\columnwidth}
{\begin{overpic}[width =\figwid]{story7move1.pdf}%\put(45,75){$t$  = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move2.pdf}%\put(45,75){$t$  = 0.25 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move3.pdf}%\put(45,75){$t$  = 0.5 s}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{story7move5.pdf}%\put(45,75){$t$  = 0.75 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move6.pdf}%\put(45,75){$t$  = 1 s}
\end{overpic}
\begin{overpic}[width =\figwid]{story7move7.pdf}%\put(45,75){$t$  = 1 s}
\end{overpic}
}
\caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:PosControl2Robots}: two robot positioning using walls with infinite friction.
Robot initial positions are shown by a crosshair, and final positions by a circled square.  Dashed lines show the shortest route if robots could be controlled independently.  The path given by  Alg. \ref{alg:PosControl2Robots} is shown with solid arrows.
}
%\vspace{-2em}
}
\end{figure*}

\begin{figure*}
\centering
\renewcommand{\figwid}{0.3\columnwidth}
{\begin{overpic}[width =\figwid]{storyWorst1.pdf}%\put(45,75){$t$ = 0 s}
\end{overpic}
\begin{overpic}[width =\figwid]{storyWorst2.pdf}%\put(45,75){$t$ = 0.25 s}
\end{overpic}
\begin{overpic}[width =\figwid]{storyWorst4.pdf}%\put(45,75){$t$  = 0.5 s}
\end{overpic}\\
\vspace{1em}
\begin{overpic}[width =\figwid]{storyWorst6.pdf}%\put(45,75){$t$  = 0.75 s}
\end{overpic}
\begin{overpic}[width =\figwid]{storyWorst7.pdf}%\put(45,75){$t$  = 1 s}
\end{overpic}}
\caption{\label{fig:shapeControlMathematica2}{Two robot positioning: switching positions using walls with infinite friction.}
%\vspace{-2em}
}
\end{figure*}

### 2018-07-12 Qualifying Exam/theory.tex
so many changes to the dissertation, and Other!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Theory:
%% Models (of each robot, of the global input)
%% Impossibility result
%% Controlling Mean proof
%% Controlling Variance proof (CLF)
%% A Hybrid controller using hysteresis
\subsection{Models}
Consider holonomic robots that move in the 2D plane. We want to control position and velocity of the robots. 
First, assume a noiseless system containing one robot with mass $m$.
 Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) \\
\mathbf{y}(t) &= C \mathbf{x}(t) + D \mathbf{u}(t)\nonumber 
\end{align}
and our state space representation as:
\begin{equation}
\begin{bmatrix}
\dot{x}_1\\ 
\dot{x}_2\\
\dot{x}_3\\
\dot{x}_4
\end{bmatrix} = \begin{bmatrix}
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0\\
0 & 0 & 0 & 1\\
0 & 0 & 0 & 0
\end{bmatrix}  \begin{bmatrix}
x_1\\
x_2\\
x_3\\
x_4
\end{bmatrix} + \begin{bmatrix}
0 & 0 \\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{bmatrix}  [u_x,u_y].
\end{equation}

We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
\begin{equation}
\mathcal{C} = [ B, AB, A^2B, ... , A^{n-1}B ].
\end{equation}

\begin{equation}
\textrm{Here }
\mathcal{C}=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],
\end{equation}
and thus all four states are controllable.

\subsection{Independent control with multiple robots is impossible}
A single robot is fully controllable, but what happens with $n$ robots? For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
%\begin{equation}\left[ x_{1,1},x_{2,1},\ldots, x_{2n-1},x_{2n}\right]^\intercal = \left[ p_{x,1},v_{x,1},\ldots,p_{x,n},v_{x,n}\right]^\intercal. \nonumber\end{equation}
%\begin{align}
%\dot{p}_{x1} &= v_{x1}\\\nonumber
%\dot{v}_{x1} &= a_{x1}\\\nonumber
%\dot{p}_{x2} &= v_{x2}\\\nonumber
%\dot{v}_{x2} &= a_{x2}\\\nonumber
%&\vdots\\\nonumber
%\dot{p}_{xn} &= v_{xn}\\\nonumber
%\dot{v}_{xn} &= a_{xn}\nonumber
%\end{align}
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1,1}\\ 
\dot{x}_{2.1}\\
\vdots\\
\dot{x}_{1,n}\\
\dot{x}_{2,n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1,1}\\
x_{2,1}\\
\vdots \\
x_{1,n}\\
x_{2,n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{equation}
 However, just as with one robot, we can only control two states because $\mathcal{C}$ has rank two:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,, 
\ldots \right].
\end{equation}  
\subsection{Controlling mean position}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We can not control the position of all the robots, but what states are controllable? To answer this question we create the following reduced order system that represents the average $x$ position and $x$ velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}}_1 \\
\dot{\bar{x}}_2
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1,1}\\
x_{2,1}\\
\vdots\\
x_{1,n}\\
x_{2,n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{align}
Thus we have
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}}_1 \\
\dot{\bar{x}}_2
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}_1\\
\bar{x}_2
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x.
\end{equation}

We again analyze $\mathcal{C}$:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right].
\end{equation}
This matrix has rank two, and thus the average position and average velocity are controllable.

Due to symmetry, only the mean position and mean velocity are controllable. However, there are several techniques for breaking symmetry, for example by allowing independent noise sources \cite{beckerIJRR2014}, or by using obstacles \cite{Becker2013b}.

\subsection{Controlling the variance of many robots}\label{sec:VarianceControl}

The variance, $\sigma_x^2,\sigma_y^2$, of the swarm's position is computed:
\begin{align}
 \overline{x}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n x_{1,i}, \qquad  %\nonumber \\ 
\sigma_x^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (x_{1,i} - \overline{x})^2,  \nonumber \\ 
 \overline{y}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n x_{3,i}, \qquad  %\nonumber \\ 
\sigma_y^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (x_{3,i} - \overline{y})^2.  
%NOT 1/(N-1) because we are measuring the actual Variance, not the variance of samples drawn from a distribution
\end{align}

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. Both conditions are readily found at the micro and nanoscale. 
Real systems, especially at the micro scale, are affected by unmodelled dynamics. These unmodelled dynamics are dominated by Brownian noise. To model this~\eqref{eq:stdform} must be modified as
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t)\\
 \mathbf{y}(t) &= C  \mathbf{x}(t) + D  \mathbf{u}(t),\nonumber
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise. Given a large free workspace, a \emph{Brownian noise} process increases the variance linearly with time:
\begin{equation}\dot{\sigma_x^2}(\mathbf{x}(t), \mathbf{u}(t) = 0)  = W \bm{\varepsilon}. \end{equation}
%After some time, Gaussian distribution shapes the outline of the robots because of the Brownian noise feature:\\
%\begin{equation}
%P(x) = \frac{1}{\sigma \sqrt{2\pi } }e^{ -  \frac{\left( x - \mu  \right)^2 } {2\sigma ^2 }}
%\end{equation}
If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

 A flat obstacle can be used to decrease variance. Pushing a group of dispersed robots against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached. For large $n$, Graham and Sloan showed that the minimum-variance packing  $\sigma^2_{optimal}(n,r)$ for $n$ circles with radius $r$ is $\approx 
   \frac{\sqrt{3}}{\pi} (n r)^2\approx 0.55(n r)^2 $~\cite{graham1990penny}. 
% Sloan minimized second moment U = 1/d^2 \sum_{i=1}^{n} || P_i - P ||^2 = (\sqrt{3} n^2)/(4 \pi), where d = 2r.
% Therefore the variance is d^2 (\sqrt{3} n^2)/(4 \pi), or 4^r^2 (\sqrt{3} n^2)/(4 \pi) = r^2 (\sqrt{3} n^2)/(\pi)If we only need to reduce variance in a single axis, the variance can be reduced to zero, given sufficient space.

We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function \cite{Lyapunov1992}.  A suitable Lyapunov function is squared variance error:
\begin{align}
\label{eq:LyapunovVariance}
V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2\nonumber\\
\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma^2}(\mathbf{x}).
\end{align}
We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{goal}$.
To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
\begin{align}
u(t) &=   \begin{cases}
	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{goal} \\ 
	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal}.
\end{cases} 
\end{align}
 For such a $u(t)$,
 \begin{align}
\dot{\sigma^2}(\mathbf{x}) &=   \begin{cases}
	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{goal}, \sigma^2_{optimal}(n,r))  \\ 
	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal},
\end{cases} 
\end{align} and thus
$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$

\subsection{Controlling both mean and variance of many robots}

The mean and variance of the swarm cannot be controlled simultaneously, however if the dispersion due to Brownian motion is much less than the maximum controlled speed, we can adopt a hybrid, hysteresis-based controller to regulate the mean and variance shown in Alg.~\ref{alg:MeanVarianceControl}.  Such a controller normally controls the mean position according to \eqref{eq:PDcontrolPosition}, but switches to minimizing variance if the variance exceeds some $\sigma_{max}^2$.  The variance is lowered to less than $\sigma_{min}^2$, and the system returns to controlling the mean position. This is a standard technique for dealing with control objectives that evolve at different rates~\cite{Sadraddini2015,kloetzer2007temporal}, and the hysteresis avoids rapid switching between control modes. The process is illustrated in Fig.~\ref{fig:hysteresis}.

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ 
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$

\If {$\sigma_x^2 > \sigma_{max}^2$}
\State $x_{goal}  \gets x_{min}$
\State $flag_x  \gets \TRUE$
\Else {\bf~if} {$flag_x$ \textbf{and} $\sigma_x^2 < \sigma_{min}^2$}
\State $x_{goal}  \gets  x_{target}$
\State $flag_x  \gets  \FALSE$
\EndIf

\If{$\sigma_y^2 > \sigma_{max}^2$}
\State $y_{goal}  \gets y_{min}$
\State $flag_y  \gets \TRUE$
\Else {\bf~if} {$flag_y$ \textbf{and} $\sigma_y^2 < \sigma_{min}^2$}
\State $y_{goal}  \gets  y_{target}$
\State $flag_y  \gets \FALSE$
\EndIf
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndLoop
\end{algorithmic}
\end{algorithm}
\begin{figure}
\centering
\begin{overpic}[scale=.4]{hysteresis.png}
\put(39,21){$\sigma^2 < \sigma^2_{min}$ }
\put(39,5){$\sigma^2 > \sigma^2_{max}$}\end{overpic}
\caption{\label{fig:hysteresis} Two states for controlling the mean and variance of a robot swarm.
%\vspace{-2em}
}
\end{figure}

A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance is 
$ \sigma^2_{optimal}(n,r) = \frac{\sqrt{3}}{ \pi}  n r^2 $.  
The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:VarianceMinMaxBand}:
\begin{align}
 \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\
  \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf

\begin{figure}
\centering
\begin{overpic}[width = \columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:VarianceMinMaxBand} The switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The above plot uses robot radius $r=1/10$.
}\vspace{-1em}
\end{figure}

%what images should I show here?
%hysteresis control law, \cite{sadra2014}

\subsection{Controlling covariance using friction}
Global inputs move a swarm uniformly.  
Controlling covariance requires breaking this uniform symmetry.  A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. Directly controlling covariance by pushing the swarm into a boundary requires changing the boundary.  An obstacle in the lower-right corner is enough to generate positive covariance.  Generating both positive and negative covariance requires additional obstacles.  Requiring special obstacle configuration also makes covariance control dependent on the local environment. 
  Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  Robots touching a wall have a negative friction force that opposes movement along the boundary, as shown in Eq.\ \eqref{eq:frictionmodel}.  This  causes robots along the boundary to slow down compared to robots in free-space. This enables covariance control using  boundaries with arbitrary orientations.

 Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$.  The force of friction $F_f$ is
\begin{align}
N &= F \cos(\theta) \nonumber\\
F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel} \\ 
F \sin(\theta), & \text{else} \end{cases}  \\%\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f. \nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. As illustrated, bot experiences different net forces although each receive the same inputs.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, if the control input is parallel or into the wall, the touching robot will not move. 
The next section shows how a system with friction model \eqref{eq:frictionmodel} and two walls are sufficient to arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{friction.pdf} 
\caption{Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot.}
\label{fig:friction}
\end{center}
\end{figure}

\section{Position control of $2$ robots using wall friction}\label{sec:PostionControl2Robots}
This section describes an algorithm for positioning two robots and introduces concepts that will be used for multi-robot positioning.
%Although the generalization of the proposed positioning algorithm in here is rather straightforward for multi robot systems, for the sake of simplicity we describe the algorithm designed for two robots. 
As we can see in Alg.~\ref{alg:PosControl2Robots}, assume two robots are initialized at $s_1$ and $s_2$ with corresponding goal destinations $e_1$ and $e_2$. 
Denote the current positions of the robots  $r_1$ and $r_2$. 
Let the subscripts $x$ and $y$ denote the $x$ and $y$ coordinates, i.e., $s_{1x}$ and $s_{1y}$ denote the $x$ and $y$ locations of $s_1$. 
The algorithm assigns a global control input at every instance.
 As a result, our goal is to adjust 
 $\Delta r_x = r_{2x}-r_{1x}$ from $\Delta s_x = s_{2x}-s_{1x}$ to $\Delta e_x = e_{2x}-e_{1x}$ and similarly adjust 
 $\Delta r_y = r_{2y}-r_{1y}$ from $\Delta s_y = s_{2y}-s_{1y}$ to $\Delta e_y = e_{2y}-e_{1y}$ with one global input at every instance. 
 The key to the algorithm is the position-dependent friction model \eqref{eq:frictionmodel}.
 %employing the assumption we have made earlier about the walls' friction.

Our algorithm uses a divide and conquer method to solve the positioning problem. 
It finds the final position of the robots in two steps: (i) First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
(ii) Having fixed $\Delta r_x$ to $\Delta e_x$ as desired, the  algorithm next keeps $\Delta r_x$ constant and adjusts $\Delta r_y$ to $\Delta e_y$, as desired in Alg..~\ref{alg:YControl}. 
Though steps (i) and (ii) are similar from an algorithmic point of view, the following subsections describe the process in detail.

\subsection{Step (i): Fixing $\Delta r_x$}
\label{theory:step1}
\begin{itemize}
\item Define $e'_1=(e_{1x},s_{1y})$ and $e'_2=(e_{2x},s_{2y})$. Our goal for defining $e'_1$ and $e'_2$ is to understand the direction to which robots should move in order to adjust $\Delta r_x$. Let $e'_{\rm{top}} = \arg \max_i e'_{iy}$ and $e'_{\rm{bottom}} = \arg \min_i e'_{iy}$. Now if $e'_{\rm{top},x}-e'_{\rm{bottom},x}>0$, then the global input to both robots would be toward left direction and if $e'_{\rm{top},x}-e'_{\rm{bottom},x}<0$, then the global input to both robots would be toward right direction. The two robots continue their horizontal path until one of them reaches the $\epsilon$-neighborhood of one of the left or right walls.
\item At this step, let $y_{\min} = \min_i r_{iy}$, i.e., $y_{\min}$ is the minimum height of the two robots. We move both robots downward by the amount of $y_{\min}$ such that one of the robots would touch the bottom wall and hence friction force will not let that robot to move left or right.
\item The fact that the friction force of the bottom wall would not let the lower robot to move right or left will let the other robot to move to right and left freely to adjust $\Delta r_x $ according to $\Delta e_x$.
\item Finally, even if with the free move of the upper robot $\Delta r_x$ is not set to the $\Delta e_x$, we can run the Step (i) (as described in the previous paragraphs) again to adjust the $\Delta r_x$. It is easy to show that it is guaranteed that we can adjust $\Delta r_x$ to $\Delta e_x$ in only two iterations.
\end{itemize}

\subsection{Step (ii): Fixing $\Delta r_y$}
Now that we have adjusted the difference in robots' positions along one axis, we focus to do the same on the other axis as well. Therefore, similar to Section \ref{theory:step1}, we employ the following steps:
\begin{itemize}
\item Let $s'_1$ and $s'_2$ be the points we derived at the end of the steps in Section \ref{theory:step1}. 
\item Define $e''_1=(s'_{1x},e_{1y})$ and $e''_2=(s'_{2x},e_{2y})$. We define $e''_1$ and $e''_2$ to understand the direction to which robots should move in order to adjust $\Delta r_y$. Let $e''_{\rm{right}} = \arg \max_i e''_{ix}$ and $e''_{\rm{left}} = \arg \min_i e'_{ix}$. Now if $e''_{\rm{right},y}-e''_{\rm{left},y}>0$, then the global input to both robots would be toward down direction and if $e''_{\rm{right},y}-e''_{\rm{left},y}<0$, then the global input to both robots would be toward up direction. The two robots continue their vertical path until one of them reaches the $\epsilon$-neighborhood of one of the top or bottom walls.
\item At this step, let $x_{\min} = \min_i r_{ix}$, i.e., $x_{\min}$ is the minimum distance of the two robots from the origin along the $x$-axis. We move both robots to the left by the amount of $x_{\min}$ such that one of the robots would touch the left wall and hence friction force will not let that robot to move up or down.
\item The fact that the friction force of the left wall would not let one of the robots to move up or down will let the other robot to move to up or down freely to adjust $\Delta r_y $ according to $\Delta e_y$.
\item Finally, even if with the free move of the robot which is not touching the wall  $\Delta r_y$ is not set to the $\Delta e_y$, we can run the Step (i) (as described in the previous paragraphs) again to adjust the $\Delta r_y$. It is easy to show that it is guaranteed that we can adjust $\Delta r_y$ to $\Delta e_y$ in only two iterations.
\end{itemize}
Once $\Delta r_x$ and $\Delta r_y$ are set to $\Delta e_x$ and $\Delta e_y$, we can use global input to easily move both robots from $r_1$ and $r_2$ toward $e_1$ and $e_2$.

\begin{algorithm}
\caption{WallFrictionArrange2Robots($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\Require 
Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. 
 Current position of the robots are $(r_1,r_2)$.

\State ($r_1,r_2$) = GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
\State GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)

\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $\epsilon \gets $ small number
\State $ \Delta s_x  \gets s_{1x} - s_{2x} $
\State $ \Delta e_x \gets e_{1x} - e_{2x} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_x < 0 $ }
\State $ m \gets ( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Move to right wall}
\Else 
\State  $ m \gets ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
\State $ m \gets (\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
\Else
\State $ m \gets (-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move up}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_x = r_{1x} - r_{2x}$
\If {$\Delta r_x \equiv \Delta e_x$} 
%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{GenerateDesired$y$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:YControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. Current position of the robots are $(r_1,r_2)$.
\Ensure   $ r_{1x} - r_{2x}  \equiv s_{1x} - s_{2x} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $ \Delta s_y  \gets s_{1y} - s_{2y} $
\State $ \Delta e_y \gets e_{1y} - e_{2y} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_y < 0 $ }
\State $ m \gets ( L-\max( r_{1y},r_{2y}) ,0)   $ \Comment{Move to top wall}
\Else 
\State  $ m \gets ( -\min( r_{1y},r_{2y}),0 )    $ \Comment{Move to bottom wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1x},r_{2x} ))$ \Comment{Move to left}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_y - (r_{1y} - r_{2y} ) > 0 $}
\State $ m \gets (\min(|\Delta e_y - \Delta s_y |, L- r_{1y}), 0)$  \Comment{Move top}
\Else
\State $ m \gets (-\min(|\Delta e_y - \Delta s_y |, r_{1y}), 0)$\Comment{Move bottom}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move right}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_y = r_{1y} - r_{2y}$
\If {$\Delta r_y \equiv \Delta e_y$} 
\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

### 2018-07-12 Shiva Dissertation/chap/intro.tex
so many changes to the dissertation, and Other!

@@
 \centering
-\begin{overpic}[width=0.5\columnwidth]{MainExpFig.pdf}\end{overpic}
-\begin{overpic}[width=0.465\columnwidth]{CoverPhoto.pdf}
+\begin{overpic}[width=0.39\columnwidth]{MainExpFig.pdf}\end{overpic}
+\begin{overpic}[width=0.365\columnwidth]{CoverPhoto.pdf}
 \end{overpic}\\
 
-\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
-\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
+\begin{overpic}[width=0.35\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
+\begin{overpic}[width=0.35\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
 \vspace{3em}

### 2018-07-12 Shiva Dissertation/chap2/experiment.tex
so many changes to the dissertation, and Other!

@@
 \begin{array}{ll}
-     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
-      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
+     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state,}\\
+      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state,} \\
       -1, & \textrm{otherwise}\\
@@
 &\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
-\text{end}&  %page 502
+\text{end}& \textrm{.}  %page 502
 \end{align}
@@
 $\mathbf{D}(\mathbf{b})$ is the desired direction of motion from~\eqref{eq:OptimalPolicy}.
+Therefore, we compute the goal locations with the following equations:
 \begin{align}

### 2018-07-12 Shiva Dissertation/chap2/realExperiment.tex
so many changes to the dissertation, and Other!

@@
 \subsection{Environmental setup}  
-In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots and a 1.5 m$\times$1.2 m whiteboard as the workspace. 
+In these experiments as shown in Fig.~\ref{fig:setup2}, we used $n$=100 kilobots and a 1.5 m$\times$1.2 m whiteboard as the workspace. 
 LED floodlights were placed 1.5 m above the table on the sides and corners of a square with 6 m sides. An Arduino Uno connected to an 8-relay shield controlled the lights.
@@
 \end{center}
-\caption{\label{fig:setup}
+\caption{\label{fig:setup2}
 Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
@@
 
-\paragraph{Swarm mean control (hardware experiment)}
+\subsection{Swarm mean control (hardware experiment)}
 
@@
 The kilobots run a phototaxis routine where they search for an orientation that aligns them with the light source, and then move with an approximately constant velocity toward this light.
-The kilobots oscillate along this orientation because they only have one light detector.
+% The kilobots oscillate along this orientation because they only have one light detector.
 

### 2018-07-12 Shiva Dissertation/chap2/theory.tex
Shiva PhD dissertation
Still requires some edits.
dissertation as I sent to committee members.
so many changes to the dissertation, and Other!

--- 2017-10-03 journalSwarmControl/theory.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Global Control Laws for a holonomic swarm}
+\section{Global control laws for a holonomic swarm}
 \label{sec:theory}
@@
 
-We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
+We represent particles as holonomic robots (robots that can move freely in any direction) that move in the 2D plane.
+ We want to control position and velocity of the particles. 
 First, assume a noiseless system containing one particle with mass $m$.
@@
 \begin{align}\label{eq:stdform}
-\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t).
+\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t),
 \end{align}
@@
 where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$.
-Given a large obstacle-free workspace with $\mathbf{u}(t)= 0$, a \emph{Brownian noise} process increases the variance linearly with time.
+Given a large obstacle-free workspace with $\mathbf{u}(t)= 0$, a \emph{Brownian noise} process increases the variance linearly with time and can be calculated by
 \begin{equation}
@@
 \begin{align}
-u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
+u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}) \textrm{       and}  \label{eq:PDcontrolVariance}\\
 u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}

### 2018-07-12 Shiva Dissertation/chap3/simulation.tex
so many changes to the dissertation, and Other!

@@
   We use the hysteresis variance control from \cite{ShahrokhiIROS2015}  to gather the swarm when its variance grows too large. 
-  The following control law chooses a goal position to regulate the orientation of the object.
+  The control law that chooses a goal position to regulate the orientation of the object is
 
@@
 \ref{fig:OrientCont} illustrates this controller with different starting positions.
-When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace.
-Code is available at \cite{Shahrokhi16Orient}.
+When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace.% Code is available at \cite{Shahrokhi16Orient}.
 

### 2018-07-12 Shiva Dissertation/chap3/theory.tex
dissertation as I sent to committee members.
so many changes to the dissertation, and Other!

@@
 \begin{align}
-\tau_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O ),   \label{eq:swarmtorque}\\
+\tau_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i \times (P_i - O ) \textrm{      and}   \label{eq:swarmtorque}\\
 F_{\text{total}} &= \sum\limits_{i=1}^n \rho_i F_i.  \label{eq:swarmforce}
@@
     \frac{x-\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu-\sqrt{6}\sigma \leq x \leq \mu\\
-     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma\\
+     \frac{-x+\mu + \sqrt{6} \sigma}{6\sigma^2}, &  \textrm{for   } \mu < x \leq \mu+ \sqrt{6}\sigma \\ 
      0, & \textrm{otherwise}\\
 \end{array} 
-\right. ,\\
+\right.  \textrm{and} \\
 %normal distribution

### 2018-07-12 Shiva Dissertation/chap4/relatedWork.tex
so many changes to the dissertation, and Other!

@@
 If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
- Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
+ Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g., PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
   Instead, this chapter employs a greedy search strategy.

### 2018-07-12 Shiva Dissertation/chap4/theory2robAlgs.tex
so many changes to the dissertation, and Other!

@@
 Left top: The possible first contact points for the blue and red particles are shown with blue and red arcs. 
-Left bottom: if the blue particle touches the wall at $\psi_{\min}$ (blue square) the other particle (red square) can move anywhere in the red region. 
-Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the green region. 
-Right: The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
+Bottom: Two representative 2-move reachable sets are shown.
+% if the blue particle touches the wall at $\psi_{\min}$ or $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the red or green region.
+%Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the green region. 
+ Right: %The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
 The possible 2-move reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
- If the blue particle contacts the boundary at $\psi_{\min}$, the reachable $\Delta$ configuration is the red set, or the green set if $\psi = \frac{17 \pi}{10}$.}
+% If the blue particle contacts the boundary at $\psi_{\min}$, the reachable $\Delta$ configuration is the red set, or the green set if $\psi = \frac{17 \pi}{10}$.
+}
 \end{figure}
@@
 \gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big), \textrm{ where:}\\ \label{eq:dprep}
-d_\perp(\psi)&= 2 ||s_1.p_\psi(\psi) - s_2.p_\psi(\psi)||,\\ \label{eq:ppsi}
+d_\perp(\psi)&= 2 ||s_1.p_\psi(\psi) - s_2.p_\psi(\psi)|| \textrm{    and}\\ \label{eq:ppsi}
 p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )].

### 2018-07-13 JournalTorqueControl/Review/Response.tex
Added Response.

@@
 
+	Sure.
+
 	\Reviewer{\#1}\response{This paper studies the pushing effect of a robot swarm over a rectangular object.
@@
 Obviously, this approach cannot be done because of the small maximum force that each particle can generate, but it is not mentioned in the paper.}
+	??? I don't understand this sentence completely.
 
@@
 However, some of the experiments present an adaptation on the object to avoid the robots flow over the object.}
+	The shape of the particles distribution swarm is not rigid.
+This is good when we need to move the swarm in narrow passageways and crowded environment.
+However, for applying force on the object, we need enough particles to make the required force.
+This is why we try to keep the particles together when we are applying force and torque.
+It is good to mention that we are only controlling the mean position of swarm.
 
@@
 The authors propose static distributions, but the the distributions variate on time due to the flow.}
+	This is a nice future work, but we have discussed in the experiment section that the stochasticity of the system is high and calculating the dynamic distributions is hard.
+Therefore, we approximate the final distribution with normal distribution.
 
@@
 	\response{In contrast to the title, none of the probability distributions maximizes the torque. At least, the maximization problem is not stated in the manuscript.}
+	We have changed the title so it better matches the paper. However, all the torques calculated in the theory section, are the maximized torque.
 
@@
 	\response{In (16), the function erf is not defined. It is only defined at the end of the section.} 
+	It is defined now.
 
 	\response{Based on the granular effect, there is a maximum number of robots that can apply forces without flowing. What is this number? or how can you describe this final configuration?}
+	We have moved granular effect from this paper.
 
@@
 Again, it is necessary to describe how the robots are driven to maintain the distribution and to generate the torque.}
+	We have moved granular effect from this paper.
 
 	\response{According to the format guidelines, the equations are part of the text and should be punctuated. Most of them are not.}
+	We corrected them.
 

### 2018-07-17 JournalTorqueControl/Review/Response.tex
response is updated

--- 2018-07-12 JournalTorqueControl/Review/Response.tex
@@
 We are grateful to the reviewers for helping us improve our manuscript through their comments and questions.
-Please let us know if further information is required.\medskip\\
+Please let us know if further information is required.\medskip
+
+		There seemed to be confusion from the reviewers about the {\bf note to Practitioners} section.
+We included this section, as per the instructions at \href{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}, but are happy to remove it if the standards have changed.
+		\\
 
@@
 
+	We have thoroughly revised this paper and included new experimental data. Please see the detailed responses below.
+
 	\Reviewer{\#1}\response{This paper studies the pushing effect of a robot swarm over a rectangular object.
@@
 The comments are the following.}
-	Thank you!
+	Thank you! We have completed a major revision, as explained below.
 
 	\response{At the end of Section III, What is the motivation to consider that all robots apply the same force? An intuitive efficient approach would make the farther robots apply a larger force.}
-	Roughly the applied force for the particles are the same, the applied torque would be different for the farther robots.
+
+	The reviewer is right that the further particles apply more {\bf torque}, as explained by the new equations 2 and 3. The longer the moment arm, the greater the torque.
+	As explained in section II, ``all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ is a constant for all $i$.''
 
@@
 
+	The note to practitioner compares using one particle to using many ``Torque control with only one steered particle is easy: maximize torque by pushing on the object at a location as far from the pivot point as possible.
+However, a swarm of particles contributes force at different places on the object."  Due to space limitations, this paper does not explore saturation effects.
+
 	\response{The authors claim that one of the main differences is that their robots are compliant and tend to flow over the object.
@@
 
+	We added several experiments to better explore torque control with uniform inputs.
+Figs 11 and 12 both feature a free-moving swarm, which as predicted by the central limit theorem, converges to a normal distribution.
+	  To explore other distributions we used an adaptor to constrain the particles.  Fig 9 shows the setup for a swarm forced to assume a uniform distribution.
+
 	\response{This "flow over the object" is not mentioned at all in the methodology (even for the triangular distribution).
 The authors propose static distributions, but the the distributions variate on time due to the flow.}
+	This is a nice future work, but we have discussed in the experiment section that the stochasticity of the system is high and calculating the dynamic distributions is hard.
+Therefore, we approximate the final distribution with normal distribution.
 
@@
 
+	Please refer to Fig 4c and Fig. 5c, which show the maximum torques as a function of standard deviation. 
+	The third column of Table I lists the maximum torques for the three canonical distributions examined.
+
+	We have changed the title so it better matches the paper.
+
 	\response{The authors do not mention how to control the robots. It is one of the main points that call the attention in this type of works.} 
-	The particles are getting the same control input.
+	The particles are getting the same control input, as shown in Eq. 28.
 
@@
 Assuming that the individual forces are linear is a very strong assumption, in contrast to \[15\], the particles move in a very unorganized way (even more when the particles flow over the object).}
+	We have added definition of n and explained more about our mathematics.
+(2) considers the basic torque when all the particles are touching the object, in (10) we are giving a probability to each particle regarding the swarm distribution.
 
@@
 
+We removed this section to better focus the paper.
+
 	\response{The derivation of (4) and (5) is no shown or extending the description is required.}
+	We removed this section to better focus the paper.
 
@@
 (7), only in the next section is assumed that the length is 1) is not obvious.}
+	We defined coordinate frames in each section to better represent the equations.
 
 	\response{In (16), the function erf is not defined. It is only defined at the end of the section.} 
+	We defined erf function.
 
 	\response{Based on the granular effect, there is a maximum number of robots that can apply forces without flowing. What is this number? or how can you describe this final configuration?}
+	We have moved granular effect from this paper to better focus the paper.
 
@@
 Again, it is necessary to describe how the robots are driven to maintain the distribution and to generate the torque.}
+	We have moved granular effect from this paper to better focus the paper.
 
 	\response{According to the format guidelines, the equations are part of the text and should be punctuated. Most of them are not.}
+	We have rewritten the parts that were not punctuated correctly and they are now part of the sentences.
 
 	\response{Caption in Fig 2. says "Particulate" instead of "Particle"}
+	We have removed Fig. 2 to better focus the paper.
 
@@
 	\response{Please add x-y-z axes in all the experimental figures such as Figs.1 and 8.}
+	We have added coordinate frames to the experimental figures.
 
 	\response{What does cyan/white star points mean in Fig.1?}
+	We have changed Fig. 1 and added text on the figure to define cyan star points.
 
 	\response{In this paper, two different cases are considered. One is a pivoted object, the other is a free object. Differences between two experimental results are encouraged to be highlighted. }
+	We have rewritten the paper and added simulation results to better highlight the experimental results.
 
@@
 Contacts between robots or between robot and rod should be detailed, and their effects should be further discussed if it is possible.}
+	We have discussed the material and size of the rod.
+Although in the theory of this paper we have not considered the robots' contacts with each other to simplify the equations, the experimental results of the last section validated the result of the simplified equations.
 
 	\response{The definition of 'kilobots' should be given when it is used for the first time. }
+	We have given the description of the kilobots earlier.
 
@@
 The abstract is poorly written.}
+Thank you. We have rewritten the paper to make it better organized.
 
 	\response{Question:  It would seem that the angle of repose might depend on the particle size or the coefficient of friction between particles or the object-particle.
-Can you explain why it doesn?t?
+Can you explain why it doesn't?
 I understand this isn?t the primary contribution of the article but it seems those parameters would play a role.}
+	We removed this section to better focus the paper.
 
 	\response{The abstract is repetitive and has a non-standard format (note to practitioners?) and should be restructured.
-It doesn?t make it clear that all the particles are steered with a common force (different from other swarm scenarios).
+It doesn't make it clear that all the particles are steered with a common force (different from other swarm scenarios).
 It uses the term angle of repose without defining.
 The last paragraph about a single robot does not belong in the abstract at all.}
+	We used authors' guidelines to write note to practitioners. If the standards of this journal has changed, we will happily remove that part.
 
-	\response{Introduction - paragraph 1: Last sentence doesn?t flow. }
+	\response{Introduction - paragraph 1: Last sentence doesn't flow. }
+	We rewrote the introduction to make it flow better.
 
 	\response{The term angle of repose should be defined earlier than the end of page 2, because it is used many times before that and while it is defined in [2] it is not a standard term in the robotics community.}
+	We removed this section to better focus the paper.
 
 	\response{ Many of the equations are not properly punctuated.  For example, eq 4 and 5 need periods as they end a sentence.  Same with 7,8,9, 12,13, 18, 19, 20 }
+	We have rewrote all the equations to make them part of the sentences.
 
-	\response{The last sentence of Setcion 4 doesn?t make sense.  Should the ? -? be an ?=? symbol?}
+	\response{The last sentence of Setcion 4 doesn't make sense.  Should the ? -? be an ?=? symbol?}
+	We removed this section to better focus the paper.
 
@@
 Is there an expression that would estimate the number of particles, given the particle size?}
+	We removed this section to better focus the paper.
 
 	\response{Page 3, col 2, top:
-It?s not clear what $l_p$ and $u_p$ are. (lower and upper bounds?).   Earlier l is used to mean the length of the rod.
+It's not clear what $l_p$ and $u_p$ are. (lower and upper bounds?).   Earlier l is used to mean the length of the rod.
 In the first sentence you mention a variable $up_p$, perhaps you mean $u_p$?}
+	The reviewer is right. There was a typo and we have made it clear now.
 
 	\response{On page 6 you use an inconsistent style when referring to equations.   At one point you use ?Eq. 16?  and later on the page you use the style ?(18)?}
+	We have now used only eqref style and we don't mention Eq. anymore.
 
@@
 ?For all distributions *,*  pushing at $\mu = 1$ ...?}
+	We added comma in that place.
 
-	\response{I don?t think the subsection headings should end in ?:?}
+	\response{I don't think the subsection headings should end in ``:"}
+	We have removed ``:" from subsections.
 
 	\response{The annotations in Figures 4-5 are small and hard to see.}
+	We have made these figures bigger.
 
 	\response{Section 6 paragraph 1:  The word Sections should be capitalized when referring to previous sections by number. }
+	We corrected this word to be capitalized.
 
 	\response{Should Kilobots be capitalized? (proper noun)}
+	We have capitalized Kilobots.
 
 	\response{Reference section:   the way you cite ICRA is inconsistent.}
+	We have fixed the citation for ICRA.
 

### 2018-08-02 Qualifying Exam/experiment.tex
Dissertation and Master Final Version

@@
 \end{overpic}}
-\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing the block-pushing experiment with 200 robots under automatic control.
-See the video attachment for an animation~\cite{ShivaVideo2015}.}
+\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing the block-pushing experiment with 200 robots under automatic control.}
 %\vspace{-2em}
@@
 \end{center}
-\caption{\label{fig:covExperiment}
-Hardware demonstration steering 64 kilobot robots to desired covariance.  Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
+\caption{\label{fig:covExperiment}Hardware demonstration steering 64 kilobot robots to desired covariance.
+Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
 }

### 2018-08-02 Qualifying Exam/simulation.tex
Dissertation and Master Final Version

@@
 Algorithms \ref{alg:PosControl2Robots}, \ref{alg:XControl}, \ref{alg:YControl}, were implemented in Mathematica using point robots (radius = $0$).
-All code is available at a public github repository~\cite{Shahrokhi2015GitHubShapeControl}.
+% All code is available at a public github repository~\cite{Shahrokhi2015GitHubShapeControl}.
  Figs~\ref{fig:shapeControlMathematica1} and \ref{fig:shapeControlMathematica2} show the examples of the implementation of our algorithm.
@@
 For the sake of brevity we have replaced straight moves (e.g.
-upward, downward, etc) with oblique moves that shows a combination of two moves simultaneously (e.g.
-left and down together).
+upward, downward, etc) with oblique moves that shows a combination of two moves simultaneously (e.g., left and down together).
 

### 2018-08-02 Shiva Dissertation/chap/abs.tex
Dissertation and Master Final Version

@@
 
-Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.
+Microrobotics has the potential to revolutionize many applications, including targeted material delivery, assembly, and surgery.
 The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.
 %Robotic manipulation usually assumes intelligent agents, but particle systems manipulated by a global signal.
- When there are more particles than control inputs the system is underactuated and requires new control techniques.
+ When there are more particles than control inputs, the system is underactuated and requires new control techniques.
  Rather than focusing on a specific microrobotic system, this dissertation designs control laws and algorithms for steering many particles controlled by global fields.

### 2018-08-02 Shiva Dissertation/chap2/experiment.tex
Dissertation and Master Final Version

@@
   %\vspace{-1em}
-    \caption{\label{tbl:resultsSum}Summary, Object Manipulation Results}
+    \caption{\label{tbl:resultsSum}Summary, object manipulation results}
 \begin{tabular}{ c c }
@@
 \end{center}
-\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
+\caption{\label{fig:Emaze} The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
 }

### 2018-08-02 Shiva Dissertation/chap4/covControl.tex
Dissertation and Master Final Version

@@
 \begin{align}
-r_{\text{left}} &= \min (L,1-w)\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right)\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right)
+r_{\text{left}} &= \min (L,1-w),\nonumber \\
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{\text{left}}}{L}  \right) \text{and}\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{\text{left}}}{L}  \right).
 \end{align}

### 2018-08-14 ICRA19/covControl.tex
New ICRA paper

@@
 The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
-Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \cite{Zhao2016mathematicaSquare}.
+Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill}.
+% see also \cite{Zhao2016mathematicaSquare}.
 
@@
 
-Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm and are calculated as%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
@@
 \bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
-	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
+	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )},\\
 \bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
@@
 \end{align}
-The full equations are included in the appendix~\cite{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}.
+The full equations are summarized in Fig.~\ref{fig:SquareFill}.
 A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
@@
 \end{align}
-For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
+For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center which is
 \begin{align}
@@
 Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}, see also \cite{Zhao2016mathematica}.
+Results are displayed in Fig.~\ref{fig:CircleFill}.%, see also \cite{Zhao2016mathematica}.
 
@@
 \begin{align}
-F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
+F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f , \nonumber \\
 \text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
@@
 \subsection{Maximizing correlation using wall friction}\label{subsec:ClosedLoopCovarianceControl}
-Using the environment to individually move $n$ particles to goal positions in Alg.
+%Using the environment to individually move $n$ particles to goal positions in Alg.
 \ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg.
@@
  This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1.
-By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
+By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is
 \begin{align} \label{eq:GravityCorrelation}
@@
 The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
-For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
+For swarms with small area, this method enables generating the full range of correlations $\pm 1$.%, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
  As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.

### 2018-08-14 ICRA19/intro.tex
New ICRA paper

--- 2018-05-29 journalWallFriction/intro.tex
@@
  Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
- In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
- However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
-This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contacts.
-We assume that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
+ In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}
+We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
+This work analyses interactions between the boundary and the particle swarm to control shape of the swarm.
+In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including mean position and variance.
+This work extends that work to covariance control.
 
-\begin{figure}
-\centering
-\vspace{1.5em}
-%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
-\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
-\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
-\vspace{1em}
-\caption{\label{fig:IntroPic}
-Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
-} \vspace{-1em}
-\end{figure}
+%    \begin{figure}
+%    \centering
+%    \vspace{1.5em}
+%    %\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
+%    \begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
+%    \begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
+%    \vspace{1em}
+%    \caption{\label{fig:IntroPic}
+%    Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+%    } \vspace{-1em}
+%    \end{figure}
 %\todo{add the picture of magnetic setup}
@@
 To extend the range of possible shapes,
-  Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
-We provide an algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
-Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
+  Sec.  \ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
+Section  \ref{sec:covControl} describes implementations of the algorithms of covariance control in simulation and hardware experiments. 
+ % Sec.  \ref{sec:expResults} 
  We end with directions for future research in Sec.  \ref{sec:conclusion}.
-
-This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces.
-This work extends the analysis to convex workspaces and 3D positioning.
-This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-08-14 ICRA19/relatedWork.tex
New ICRA paper

@@
 
-If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
- Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
-  Instead, this paper employs a greedy search strategy.
+%If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
+% Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
+%  Instead, this paper employs a greedy search strategy.

### 2018-08-14 ICRA19/theory.tex
New ICRA paper

@@
 \end{figure}
-By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:
+By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
 \begin{figure}
@@
  Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
-Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$:
+Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$ and is defined as
 \begin{align}

### 2018-08-24 journalWallFriction/simulation.tex
Initial paper surgery.  Much work to do, must be careful to answer all reviewers questions

@@
+\section{Two Optimal Results}\label{sec:optimalResults}
+  Algorithm ??? provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.  
+  In this section we present two results that can be incorporated into Algorithms ??? and ??? to generate shorter motion paths.
+
+ \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
+ Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
+ If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
+The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
+$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 
+ Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal contact location, as shown in Fig.
+\ref{fig:reflection}.
+  There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
+ %
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
+\vspace{-2em}
+\caption{\label{fig:reflection}
+\textcolor{red}{add 0.2 increments to both, add green dot on left image, draw incident and reflected angle}
+The shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle, as shown at left.
+(right) The first contact is colored red if the red particle is the first to touch a boundary, and colored blue if the blue particle is the first to touch.
+} \vspace{-1em}
+\end{figure}
+%
+
+ \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
+
+ The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments and has an optimal contact point, as shown in Fig.~\ref{fig:shortestpath}.
+ The problem can be simplified by choosing the coordinate system carefully.
+We define the $x$-axis along the line from the circle center to the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$-axis: $P=(\cos \theta,\sin \theta)$.
+Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$.
+Then the length of the two line segments is
+ \begin{align}\scalebox{.8}{$
+ \sqrt{{ (s-\cos \theta)^2+\sin^2 \theta}} +  \sqrt{(e \cos \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2},$}
+ \end{align}
+ which is minimized by choosing an appropriate $\theta$ value.
+
+%  This equation can be simplified to 
+% 
+% \begin{equation}
+%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
+% \end{equation}
+% 
+\begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+{\begin{overpic}[width =\figwid]{shortestpath.pdf}
+\end{overpic}
+}
+\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference.
+The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right.
+See \cite{BeckerShortestPath}.}
+%\vspace{-1em}
+}
+\end{figure}
+
+ The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}.
+There are several simple solutions.
+If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0.
+If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$.
+Label the origin $O$.
+ The optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+ The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
+ We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is $P$. 
+  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
+ From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$.
+If we mirror the point $S$ about line $\overline{OP}$ and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.
+
+ Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$.
+Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$.
+We evaluate each and select the solution that results in the shortest length path.
+%This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
+ For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}. Because the closed form solution is long, it is included in the paper attachments.
+
 \section{Simulation}\label{sec:simulation}

### 2018-08-24 journalWallFriction/theory.tex
Initial paper surgery.  Much work to do, must be careful to answer all reviewers questions

--- 2018-05-30 journalWallFriction/theory.tex
@@
 
- \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
- Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
- If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
-The first move, $m_1$, makes one particle hit a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
-$m_2$ cannot be shortened, so optimization depends on choosing the location where the particle hits the wall. 
- Since the shortest distance between two points is a straight line, reflecting the goal position across the boundary wall and plotting a straight line gives the optimal hit location, as shown in Fig.
-\ref{fig:reflection}.
-  There are four walls, and four candidate solutions, but some candidate solutions may be invalid because a different boundary is hit before the desired first contact position in move $m_1$ (light grey regions) or  invalid because $m_2$ cannot generate the goal relative spacing (dark grey regions).
- %
-\begin{figure}
-\centering
-\begin{overpic}[width=\columnwidth]{Reflection.pdf}\end{overpic}
-\vspace{-2em}
-\caption{\label{fig:reflection}
-The shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle.
-} \vspace{-1em}
-\end{figure}
-%
-
-This optimization result was implemented in our previous work using a square workspace \cite{shahrokhi2017algorithms}.
+We implemented a solution to this problem for  square workspaces in our previous work, \cite{shahrokhi2017algorithms}.
  Fig.~\ref{fig:shapeControlMathematica1} shows solutions from a \emph{Mathematica} implementation in a square workspace for six representative configurations.
@@
 \end{figure*}
-
- \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
-
- The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments and has an optimal contact point, as shown in Fig.~\ref{fig:shortestpath}.
- The problem can be simplified by choosing the coordinate system carefully.
-We define the $x$-axis along the line from the circle center to the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$-axis: $P=(\cos \theta,\sin \theta)$.
-Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$.
-Then the length of the two line segments is
- \begin{align}\scalebox{.8}{$
- \sqrt{{ (s-\cos \theta)^2+\sin^2 \theta}} +  \sqrt{(e \cos \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2},$}
- \end{align}
- which is minimized by choosing an appropriate $\theta$ value.
-
-%  This equation can be simplified to 
-% 
-% \begin{equation}
-%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
-% \end{equation}
-% 
-\begin{figure}
-\centering
-\renewcommand{\figwid}{\columnwidth}
-{\begin{overpic}[width =\figwid]{shortestpath.pdf}
-\end{overpic}
-}
-\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference.
-The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right.
-See \cite{BeckerShortestPath}.}
-%\vspace{-1em}
-}
-\end{figure}
-
- The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}.
-There are several simple solutions.
-If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0.
-If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$.
-Label the origin $O$.
- The optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
- The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
- We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is $P$. 
-  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
- From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$.
-If we mirror the point $S$ about line $\overline{OP}$ and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.
-
- Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$.
-Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$.
-We evaluate each and select the solution that results in the shortest length path.
-%This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
- For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}. Because the closed form solution is long, it is included in the paper attachments.

### 2018-08-24 journalWallFriction/theory2robAlgs.tex
Initial paper surgery.  Much work to do, must be careful to answer all reviewers questions

@@
  %%%%%%%%%%%%%%%%%%%%%%%%%%%\
-\subsection{Two Particle Path Planning}
+\subsection{Two Particle Path-Planning}
 
@@
 
+\subsection{Accessible Workspace}
+
+The $\Delta$ configuration makes it easy to compute the accessible workspace.
+Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachabable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
+
+??? INSERT NEW PLOT SHOWING FRACTION OF WORKSPACE REACHABLE IN 2 moves as a function of $d_{12}$  \\
+
+Text \\
+
+??? INSERT NEW IMAGE SHOWING FRACTION OF WORKSPACE REACHABLE IN 2 moves and 4 moves for a specific initial Delta config \\
+
+MORE TEXT \\
+
 \subsection{3D workspaces: Cylinders and Prisms}

### 2018-08-29 ICRA19/intro.tex
Made some todos.  Shiva, can you add the missing citations?

@@
 \section{Introduction}\label{sec:Intro}
+
+   \textcolor{red}{TODO: lets do some experiments with salt in 3D printed shapes (or laser cut).
+Salt is micro scale, and roughly uniform (I think) we can get different magnitudes of global force by varying the tilt angle.}
+
 Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.

### 2018-09-04 journalWallFriction/simulation.tex
updated reachable plots and paper

@@
 \section{Two Optimal Results}\label{sec:optimalResults}
-  Algorithm ??? provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.  
-  In this section we present two results that can be incorporated into Algorithms ??? and ??? to generate shorter motion paths.
+  Algorithm \ref{alg:optimalAlg} provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.  
+  In this section we present two results that can be incorporated into Algorithms \ref{alg:polygonReachbale} and \ref{alg:circularReachbale} to generate shorter motion paths.
 
@@
  If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
-The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
+The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final ositions.
 $m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 

### 2018-09-04 journalWallFriction/theory2robAlgs.tex
updated reachable plots and paper

@@
 The $\Delta$ configuration makes it easy to compute the accessible workspace.  
-Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachabable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
+ Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
+ The angle $\theta$ between the initial particle positions simply rotates the reachable $\Delta$ configuration space.
 
 ??? INSERT NEW PLOT SHOWING FRACTION OF WORKSPACE REACHABLE IN 2 moves as a function of $d_{12}$  \\
+
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{reachableSetsFuncDistance}\end{overpic}
+\caption{\label{fig:reachableSetsFuncDistance}
+  Reachable fraction of  $\Delta$ configuration space  as a function of initial distance $d_{12}$.  
+  As long as the initial configurations are distinct, the reachable set grows quickly. 
+   Only epipolar locations are unreachable, but can be asymptotically reached.
+}
+\end{figure}
 
@@
 MORE TEXT \\
+
+\begin{figure}
+\centering
+\begin{overpic}[width=.8\columnwidth]{ReachableSets0p1}\end{overpic}
+\begin{overpic}[width=.8\columnwidth]{ReachableSets0p3}\end{overpic}
+\caption{\label{fig:reachableSets}
+  Plots showing the 2, 4, 6, 8, 10, 12, and 14-move reachable sets in the  $\Delta$ configuration space for  $d_{12} = 0.1$ and $0.3$.  
+  Shown are reachable sets with initial particle-to-particle angle $\theta=0$, all other sets for other angle values can be formed by rotating by $\theta$.
+}
+\end{figure}
 

### 2018-09-05 ICRA19/covControl.tex
added more todos and blank photos

@@
 %u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
+F_{\text{\emph{forward}}}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
 0, & \text{else} \end{cases}.
@@
 r_{\text{left}} &= \min (L,1-w),\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{left}}{L}  \right) \text{and}\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{left}}{L}  \right).
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{\text{left}}}{L}  \right).
 \end{align}
@@
 The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
-For swarms with small area, this method enables generating the full range of correlations $\pm 1$.%, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
+For swarms with small area, this method enables generating the full range of correlations $\pm 1$.
+%, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.
   As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
@@
 \end{figure}
+
+\section{Experiments}
+
+\subsection{Hardware experiment: angle of repose}
+In this section we vary the force and approach angle to shape a swarm with a characteristic angle of repose.
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{angleOfReposeExp.pdf}
+\end{center}
+\caption{\label{fig:angleOfReposeExp}
+Experimental data
+}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{ironAndSaltCloseup.pdf}
+\end{center}
+\caption{\label{fig:ironAndSaltCloseup}
+Magnified image of salt and iron at angle of repose.
+}
+\end{figure}
+
+\textcolor{red}{Explain hardware experiment}
+
 \subsection{Hardware experiment: control of covariance}

### 2018-09-05 ICRA19/theory.tex
small updates to paper
Made some todos.  Shiva, can you add the missing citations?
added more todos and blank photos

@@
 
+ \begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{Angle.pdf}%\put(1,55){a)}
+\end{overpic}
+\caption{\label{fig:angle} 
+%Different values of angle of repose is shown when granular particles move faster than the rod.
+}
+\end{figure}
+
 Consider a swarm of granular particles applying force to a rod. 
@@
  We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$.
- \begin{figure}
-\centering
-\renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{Angle.pdf}%\put(1,55){a)}
-\end{overpic}
-\caption{\label{fig:angle} Different values of angle of repose is shown when granular particles move faster than the rod.
-}
-\end{figure}
+
 By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
@@
 \renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =0.6\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
 \end{overpic}\\
 \vspace{0.5em}
-\begin{overpic}[width =0.6\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
 \end{overpic}\\
 \vspace{0.5em}
-\begin{overpic}[width =0.6\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
 \end{overpic}
@@
    The maximum torque values are shown with black dots.
+   \textcolor{red}{Dr B: I don't believe that this fit line is really linear.  Someone must look at the mathematica and figure out the fit}
 % Generating code is in the attachment. 
-%\vspace{-2em}
+\vspace{-2em}
 }

### 2018-09-06 ICRA19/theory.tex
Checked the math for angle of repose, small updates

@@
 Consider a swarm of granular particles applying force to a rod. 
-If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle.
+If the rod moves slower than the particles, some particles will slide past the rod, but other particles will build up behind the rod in a characteristic triangular shape %defined by an apex angle. 
+ defined by a steepest angle of descent perpendicular to the direction of particle motion. 
 This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}.
@@
  %The rod length $\ell=1$. 
-   The maximum torque values are shown with black dots.
-   \textcolor{red}{Dr B: I don't believe that this fit line is really linear.  Someone must look at the mathematica and figure out the fit}
+   The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
+  % \textcolor{red}{Dr B: I don't believe that this fit line is really linear.  Someone must look at the mathematica and figure out the fit.  The fit is linear, but it is }
 % Generating code is in the attachment. 
@@
 \begin{array}{ll}
-\frac{-\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos\alpha\sin{\theta}} &   -\alpha<\theta<\alpha\\
+\frac{\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos(\alpha)\sin(\theta)} &   -\alpha<\theta<\alpha\\
 0 &    \textrm{otherwise .}\\
@@
 \begin{array}{ll}
-\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin{\theta} }{48\sin^2\alpha}&   -\alpha<\theta<\alpha\\
+\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin(\theta) }{48\sin^2(\alpha)}&   -\alpha<\theta<\alpha\\
 0 &    \textrm{otherwise .}\\
@@
 Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$ and is defined as
-\begin{align}
+\begin{align}\label{eq:maxTorqueAngleGivenAOR}
 \theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.

### 2018-09-06 ICRA19/covControl.tex
Checked the math for angle of repose, small updates
Some more todos for Shiva & Haoran

@@
 
-\subsection{Hardware experiment: angle of repose}
+\subsection{Hardware experiment: particle angle of repose}
 In this section we vary the force and approach angle to shape a swarm with a characteristic angle of repose.
@@
 
-\textcolor{red}{Explain hardware experiment}
+\textcolor{red}{TODO: Haoran will Explain hardware experiment}
+
+\subsection{Hardware experiment: kilobot angle of repose}
+In this section we vary the force and approach angle to shape a swarm with a characteristic angle of repose.
+
+\textcolor{red}{TODO: Shiva will insert the kilobot images and a couple of sentences, and a couple of sentences about kilobots}
 

### 2018-09-06 ICRA19/intro.tex
added more todos and blank photos
Some more todos for Shiva & Haoran

@@
 Salt is micro scale, and roughly uniform (I think) we can get different magnitudes of global force by varying the tilt angle.}
+
+   \textcolor{red}{TODO: this text is from a previous paper}
 
@@
  Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
- In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}
+ In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}. 
 We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
@@
 This work extends that work to covariance control.
+
+ \begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{leadfigure.pdf}%\put(1,55){a)}
+\end{overpic}
+\caption{\label{fig:leadfigure} 
+      \textcolor{blue}{TODO: we need a lead figure in column 1:   Combine kilobots from Fig 3 and 4 with photos of angle of repose by Haoran of iron with 4 rod angles. }
+}
+\end{figure}
 

### 2018-09-06 journalWallFriction/SwarmShapeControl.tex
Initial paper surgery.  Much work to do, must be careful to answer all reviewers questions
made todos for Shiva

@@
  This paper investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
-Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
+%Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
+Given a magnetic field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial. 
 Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
@@
 Because in vivo environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
+We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that are applied uniformly to all particles in the workspace.
 This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.

### 2018-09-06 journalWallFriction/TROsubmission01andreview/ReviewsTROv01.tex
made todos for Shiva

Associate Editor's Report:

The Associate Editor (AE) was able to obtain two high-quality reviews
from researchers working directly in this field. After considering
their reviews, it is the opinion of the AE that this paper is not
currently suitable for publication in T-RO. However, the authors are
encouraged to prepare a revised draft, addressing the concerns of the
reviewers.

Please also address how the technique could be practically applied to
navigate inside an actual 3D intestine as opposed to a cross section in
which the particles move on a planar surface. Would magnetic forces be
sufficient to overcome gravity? Would the particles need to be made
neutrally buoyant?

## add wording about  neutrally buoyant

REVIEWER 1
This paper explains an underactuated control strategy for small teams
of magnetic microparticles which move under the influence of a single
broadcast signal. This underactuated control problem is relevant to
microrobotic swarms moving under magnetic field input. In previous
work, the authors have explored interesting control methods relying on
obstacles in the environment, but the case studied here is more
application-realistic. The control method here is clever and
successful. The work is interesting to read and I think that anyone
with a basic background in controls will find this work interesting.

The main concept used here to achieve multi-particle control is that
the particles experience a no-slip condition when pushed against the
environment wall. The particles can be removed from the wall by pulling
them.

The paper is an extension of an IROS2017 paper which introduced the
idea for square workspace only. This TRO submission generalizes to a
non-square workspace, which is a notable improvement. This version of
the work will thus be a helpful addition to the literature.

The paper gives a very nice review of previous work in using broadcast
signals for multi-particle control.

Questions to consider:
Is it expected that the no-slip condition will be a strongly satisfied
assumption for in-vivo biomedical applications? In cases where there
were some slip, how could that be handled?

##  To maintain clarity, this paper focused on systems with no-slip boundaries.  While several environments with no slip boundaries exist (e.g. tripe, and intestines with villi), extending to lower friction values is a intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.

This control method requires advance knowledge of the workspace
boundary geometry, in addition to micro-particle state feedback. It
would be interesting to hear about how accurate the boundary geometry
must be known for this control to work well, and perhaps compare this
will medical imaging techniques such as CT or MRI which could
potentially be used to generate such maps. For motion through 3D
lumens, the cross-sectional geometry will change throughout the lumen.

##  add text about systems with large numbers of protrusions or concave workspaces are best handled by motion planners such as RRT.

Abstract: It is stated that “given 3 orthogonal magnetic fields …”.
This statement should be made more precise, because the magnetic
pulling here uses field gradients, not fields. In addition, these
gradients are not orthogonal. An accurate statement would need to be
more complex.

# made fix

Fig 2: “first contact point” is not clear. There is no way given to
tell where these line up with actual positions on the workspace.

# FIg 2 is now fig 11.  We updated this figure to clarify (tell what we changed) we rearranged the order because this optimization results....

Fig 3: the meaning of grey areas should be mentioned in the fig or
caption.

# now Figure 2, Gray areas denote regions unaccessible by our motion planner. The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and the goal positions must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.

Figures are hard to read because there is a lot going on, the font is
often too small and text sometimes is obscured.

# We fixed figures ?????

III.C. It’s not clear why this section is relevant. Why is the shortest
path which intersects the wall needed? Some writing structure could
help the reader understand what the goal is here.

# To better present the motion planner, we moved this section, which covers an optimization result, to the end, and altered the framing text
# TODO: alter the framing text

REVIEWER 6  (1 and 6 )

In this manuscript algorithms to position two particles in arbitrary
locations under uniform actuation are proposed.  The algorithms rely on
differentiating between particles by contacting them with walls or
boundaries of the domain, at which they are assumed to not move unless
actuated in a direction away from the wall.  The paper presents some
results on shortest paths that contact surfaces (but see below on some
presentation issues), then describes the algorithm, which basically
entails moving both particles so that one particle contacts a boundary,
adjusting the relative displacement between particles while the one
particle remains at the boundary, then translating both particles to
the desired location.
Simulations of the algorithm are presented in square and circular
domains, and experiments are described in circular cross sections
inspired by intestinal and stomach scenarios.

Although similar ideas were presented earlier by the authors using
either an obstacle in the workspace or square workspaces, this paper
extends that work to circular and convex polygonal workspaces.	The
overall idea is a novel contribution as other swarm control techniques
focus on either heterogeneity of microrobots or of actuation (such as
fields) rather than distinguishing particles by their proximity to
boundaries.  However, in my opinion the manuscript requires major
revisions before being suitable for publication, mainly to address the
theoretical effectiveness of the algorithm (as detailed below), and
less so to address clarity of presentation.

1.  The main weakness of the paper in my opinion is that the authors
never really prove that their algorithm works, or alternatively how
general their algorithm is, i.e. what are its limits of applicability. 
To be more specific, in their Algorithm 1, if the desired configuration
is not within the 2-move reachable set, then the algorithm targets
instead the closest point in the 2-move reachable set, and then
"iterates until we reach the goal."  When does such iteration actually
achieve the goal?  What are the attainable final goals that can be
reached after iteration?  Does the space of attainable goals depend on
the initial positions of the particles?  While there are simulations,
the ones presented do not explore the entire possible space (which is
understandable since they it is quite large).  I think an analysis of
the attainable space of their algorithm is needed.

# An excellent suggestion!  We rearranged the paper and added section ## which analyzes the reachable set.  We added figures # and # which...

2.  There are issues with the presentation that can be easily improved.
 At the beginning of the paper many terms are not defined which makes
figures and discussion hard to follow. Definitions do come later, but
they should be moved up.  For example, in Fig 2 which symbols are
targets and which are initial conditions are not defined.  In Fig 3
epsilon is shown but not defined nor is its significance explained.  s
and g and \Delta s and \Delta g are not defined. In Fig 12, where is
the goal on the boundary? Can that be indicated on the figures?

## Fig 2 is now 12, and we added
## Figure 3 is now 2, we did ...
## Figure 12 is now
Add captions for start and end (colored squares and circles) to each drawing  (including fig 12)

3.  (more presentation) At the beginning of section IV it wasn't clear
to me whether Algorithm 1 had been described yet when it was first
mentioned, and whether that sentence was a description of what was to
come, or was as assertion that logically followed from the previous
parts of the paper.

4. A minor comment, sections IIIB and C are a little confusing since it
is not motivated why shortest path is being considered in the overall
argument of the paper.	Furthermore, it seems that shortest path in
IIIB is used for the situation moving two particles, while in IIIC it
is used for the situation of moving only one particle.	Consistency in
the presentation would help.

# we moved these optimization results to ## to better focus on the

### 2018-09-06 journalWallFriction/theory2robAlgs.tex
made todos for Shiva

@@
  The angle $\theta$ between the initial particle positions simply rotates the reachable $\Delta$ configuration space.
-
-??? INSERT NEW PLOT SHOWING FRACTION OF WORKSPACE REACHABLE IN 2 moves as a function of $d_{12}$  \\
+  As long as the initial configurations are distinct $(s_1 \neq s_2)$, the reachable set grows quickly. 
+  This relationship is shown in Fig.~\ref{fig:reachableSetsFuncDistance}.
+   Only antipodal locations are unreachable $(\norm{g_2-g_1} = 1)$, but can be asymptotically approached.
+   Indeed even with a tiny initial separation of $d_{12} = 0.001$, after 14 moves 90\% of the $\Delta$ configuration space is reachable.
 
@@
   Reachable fraction of  $\Delta$ configuration space  as a function of initial distance $d_{12}$.  
-  As long as the initial configurations are distinct, the reachable set grows quickly. 
-   Only epipolar locations are unreachable, but can be asymptotically reached.
 }
@@
 
-Text \\
-
-??? INSERT NEW IMAGE SHOWING FRACTION OF WORKSPACE REACHABLE IN 2 moves and 4 moves for a specific initial Delta config \\
-
-MORE TEXT \\
+Two example sets of the reachable $\Delta$ configuration space for  $d_{12} = 0.1$ and $d_{12} = 0.3$ are shown in Fig.~\ref{fig:reachableSets}.
+After two moves, $d_{12} = 0.1$ reaches only 6.3\% of the $\Delta$ configuration space, but 30\% in four moves, 55\% in six moves, 75\% in eight, 86\% in ten, 93\% in twelve, and 96\% in fourteen moves.
+%{0.1`, 0.06275233581754532`, 0.2996219528008345`, 0.5531978612123086`, 0.7453936085358723`, 0.8631378723622217`, 0.9280921766555345`, 0.9614346262938109`}
+ Though these images show reachable sets with initial particle-to-particle angle $\theta=0$, all other sets for other $\theta$ values can be formed by rotating by $\theta$.
 
@@
   Plots showing the 2, 4, 6, 8, 10, 12, and 14-move reachable sets in the  $\Delta$ configuration space for  $d_{12} = 0.1$ and $0.3$.  
-  Shown are reachable sets with initial particle-to-particle angle $\theta=0$, all other sets for other angle values can be formed by rotating by $\theta$.
 }
@@
 Extending path planning to 3D is possible only if the two particles do not initially have the same $x$ and $y$ positions.
-For ease of analysis, we assume the workspace boundaries extend in the $\pm z$ direction to form either right cylinders or right prisms.
+For ease of analysis, we assume neutrally buoyant particles with workspace boundaries that extend in the $\pm z$ direction to form either right cylinders or right prisms.
 If the 3D projection is at a different angle, redefine the 2D workspace as a region perpendicular to the projection.
@@
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.
-Other workspaces may be better handled by other path planners, such as \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.
+Other workspaces may be better handled by other path planners, such as RRT \cite{LavalleTODO-Shiva}, or \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.

### 2018-09-09 ICRA19/theory.tex
added a caption
added Fig description for angle of repose

@@
 \end{overpic}
-\caption{\label{fig:angle} 
+\caption{\label{fig:angle}  If particles move faster than the (pink) rod, some particles slide past the rod, but the ones that remain pile up in a shape characterized by the angle of repose $\alpha$, which is particular to the particle type.
 %Different values of angle of repose is shown when granular particles move faster than the rod.
@@
  By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
+ In this plot, particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
+  A thin black line extends upwards from the rod COM, and the COM of the particulate is shown by a white and black disk.   
+  More particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
+  This uneven particulate generates a restoring torque. 
  We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$.
@@
 \caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
-%Particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
-% A thin gray line extends upwards from the rod COM, showing how more particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
-%  This uneven particulate generates a restoring torque. 
  Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
- %The rod length $\ell=1$. 
    The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
-  % \textcolor{red}{Dr B: I don't believe that this fit line is really linear.  Someone must look at the mathematica and figure out the fit.  The fit is linear, but it is }
 % Generating code is in the attachment. 
@@
 
-The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}. 
+The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}, with the rod length $\ell=1$. 
 Torque will also be similarly defined as
@@
 \end{align}
-Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce}.
+Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce} with the rod length $\ell=1$. 
  Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.

### 2018-09-09 ICRA19/covControl.tex
changed section name.. we need to read this paper carefully

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Covariance control}
+\section{Shape  Control using Rigid Boundaries}
 \label{sec:covControl}
@@
 \end{figure*} 
-Covariance control as well as mean position and variance control is needed when the swarm is moving through narrow passageways.
-This section discusses ideas to control covariance of the swarm using boundaries.
+Shape control using angle of repose only considers the particles that pile up on the object, but most of the particles pass by.
+ This section instead focuses on shape control of the entire group when all the particles are pushed against rigid boundaries, starting by examining analytically all possible configurations in two canonical configuration spaces, and then examining the effect of boundary layer friction.
 

### 2018-09-09 ICRA19/theory.tex
Restructuring paper -- it doesn't hold together well.

@@
 
+\subsection{Force and Torque}
+
 By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
@@
 To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.
+
+\subsection{Shape Control}
+
+Often the angle of the rigid  is given, but we can choose the desired approach direction $\beta$. 
+ This section examines the possible shapes that can be generated given a rigid rod of length $\ell$, an angle of repose $\alpha$, and an approach angle where the swarm moves at angle relative to the long axis of the rod $\beta$.
+
+Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm and are calculated as%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
+%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
+%
+\begin{align}
+A &=\iint_R \,dx\,dy \text{, }
+\bar{x} =\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
+\text{, } \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \text{, }\\
+%\end{align}
+%\begin{align}
+\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\text{, and}\\
+%\end{align}
+%\begin{align}
+\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
+\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}.
+\end{align}

### 2018-09-10 ICRA19/covControl.tex
Restructuring paper -- it doesn't hold together well.
added variance results

@@
 Like fluid settling in a tank, the stable final configuration minimizes potential energy.
+The set of final configurations are parametrized by a single degree of freedom, the global input angle $\beta$.
 \paragraph{Square workspace}
@@
 
-Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm and are calculated as%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else.
-%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:
-
-\begin{align}
-\bar{x} &=\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, }\qquad \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \text{, }\\
-%\end{align}
-%\begin{align}
-\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\text{, and}\\
-%\end{align}
-%\begin{align}
-\sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}.
-\end{align}
-
 The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
@@
 The full equations are summarized in Fig.~\ref{fig:SquareFill}.
+For the full equations and demonstration code, see~\cite{Haoran2016SwarminSquare}.
 A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape.
@@
 Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$.
-Results are displayed in Fig.~\ref{fig:CircleFill}.%, see also \cite{Zhao2016mathematica}.
+Results are displayed in Fig.~\ref{fig:CircleFill}.
+For the full equations and demonstration code, see~\cite{Haoran2016SwarminCircle}.%, see also \cite{Zhao2016mathematica}.
 

### 2018-09-10 ICRA19/theory.tex
added variance results

@@
 \end{array} 
-\right .\\
+\right . \label{eq:ForceAOR}
 \end{align}
@@
 
-Often the angle of the rigid  is given, but we can choose the desired approach direction $\beta$. 
+Often the angle of the rigid rod is given, but we can choose the desired approach direction $\beta$. 
  This section examines the possible shapes that can be generated given a rigid rod of length $\ell$, an angle of repose $\alpha$, and an approach angle where the swarm moves at angle relative to the long axis of the rod $\beta$.
@@
 \sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}\text{, and}\\
+\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}  \text{, }\\
 %\end{align}
@@
 \sigma_{xy} &= \frac{\iint_R  \left(x-\bar{x}\right) \left(y-\bar{y}\right) \, dx \,dy}{A} \label{eq:covAndcorrInSquareWorkspace}
-\text{, }\rho_{xy} = \frac{\sigma^2_x}{\sigma_x\sigma_y}.
+\text{, }\rho_{xy} = \frac{\sigma_{xy}}{\sqrt{\sigma^2_x\sigma^2_y}}.
 \end{align}
+
+Using these equations, we can calculate the representative shape statistics as a function of approach angle $\beta$, a one degree-of-freedom set. 
+For instance, the area $A$ of the particles is \eqref{eq:ForceAOR} with $\theta = \beta + \pi/2$, and the $y$-variance is
+\begin{align}
+\sigma_y^2(\alpha, \beta,\ell) &= \frac{1}{72}   \left( \cot(2\alpha) + \frac{\cos(2\beta)}{ \cos(2\alpha)} \right)^2.
+\end{align}
+Representative results are shown in Fig.~\ref{fig:AngleOfReposeStatistics}.
+
+\begin{figure}
+\centering
+\renewcommand{\figwid}{\columnwidth}
+\begin{overpic}[width =\figwid]{aorYmean}%\put(1,55){a)}
+\end{overpic}\\
+\vspace{0.5em}
+\begin{overpic}[width =\figwid]{aorVarX}%\put(1,55){a)}
+\end{overpic}\\
+\vspace{0.5em}
+\begin{overpic}[width =\figwid]{aorCovariance}%\put(1,55){a)}
+\end{overpic}
+\vspace{-0.5em}
+\caption{\label{fig:AngleOfReposeStatistics} 
+Each plot shows a statistic of the swarm shape for several angle of repose $\alpha$ values as a function of the approach angle $\beta$ with $\ell = 1$.
+Top plot shows the mean y coordinate $\bar{y}$,
+ middle plot shows $x$-variance $\sigma^2_x$, and
+ bottom plot shows the covariance $\sigma_{xy}$.
+\vspace{-2em}
+}
+\end{figure}

### 2018-09-10 ICRA19/covControl.tex
added lead figure, and kilobot angle of repose section

@@
 \subsection{Hardware experiment: kilobot angle of repose}
-In this section we vary the force and approach angle to shape a swarm with a characteristic angle of repose.
 
-\textcolor{red}{TODO: Shiva will insert the kilobot images and a couple of sentences, and a couple of sentences about kilobots}
+In this section we vary the force and approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose.
+Kilobot is a three centimeter low cost robot that has light sensor.
+In our experiments, these robots were programmed to go toward the brightest light in the room.
+The experimental setup is shown in Fig.
+\ref{fig:SetUp}.
+Figure \ref{fig:kilobotAngle} shows angle of repose of the Kilobots with three different angles of the pink rod when the rod has 0$^\circ$, 22.5$^\circ$ and 45$^\circ$.
+The green lines show the triangle that the robots pile up.
+The cyan star is the center of mass of the rod, and mean position and variance of the robots are shown with a red star and red ellipse.
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{SetUp.pdf}
+\end{center}
+\caption{\label{fig:SetUp}
+The platform where we control Kilobots with lights.
+}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{KilobotAngleThreeAngles.pdf}
+\end{center}
+\caption{\label{fig:kilobotAngle}
+Angle of repose of Kilobots controlled by light.
+}
+\end{figure}
+
+%\textcolor{red}{TODO: Shiva will insert the kilobot images and a couple of sentences, and a couple of sentences about kilobots}
 

### 2018-09-10 ICRA19/intro.tex
added lead figure, and kilobot angle of repose section

@@
 \renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{leadfigure.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{firstpic.pdf}%\put(1,55){a)}
 \end{overpic}
 \caption{\label{fig:leadfigure} 
-      \textcolor{blue}{TODO: we need a lead figure in column 1:   Combine kilobots from Fig 3 and 4 with photos of angle of repose by Haoran of iron with 4 rod angles. }
+     Reshaping particles by collisions with rigid bodies. Top pictures illustrates using closed boundaries to reshape the particles. Bottom pictures illustrates angle of repose of different particles.
 }

### 2018-09-10 ICRA19/intro.tex
updated introduction

@@
 
-   \textcolor{red}{TODO: this text is from a previous paper}
+  % \textcolor{red}{TODO: this text is from a previous paper}
 
-Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
+%Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
 %If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
@@
 
-The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
- The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
-Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
- Techniques are needed that can handle this underactuation.
+Shape control of many particles is essential for navigation, construction, and conveying information. 
+ The small size of the particles makes it hard or even impossible to control each particle position individually. Instead, these particles are controlled with a single global control input.
+ In previous work, we illustrated algorithms that could adjust the position of two particles in an environment with walls that have non-slip contacts \cite{shahrokhi2017}.
+We assumed that if a particle is touching a wall, it will stay still until the uniform input pull it from the wall.
+This work analyses interactions between the boundary and the particle swarm to shape shape of the swarm.
+In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including mean position and variance.
+This work extends that work to shape control.
 
- Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
- In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}. 
-We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
+%The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
+ %The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
+%Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
+% Techniques are needed that can handle this underactuation.
+
+% Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
+% In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}. 
+%We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.
 This work analyses interactions between the boundary and the particle swarm to control shape of the swarm.

### 2018-09-10 ICRA19/relatedWork.tex
edited related work a little bit.

@@
 
-Controlling the \emph{shape}, or relative positions, of a swarm of robots is a key ability for a range of applications.
+Shape, or relative position, control of many particles is necessary for a range of applications.
+Therefore, it has been a topic of research from a control perspective in both centralized and distributed approaches.
+ This paper focuses on centralized approach that apply the uniform control input to all the particles.
+To change the relative position of the particles, we need to break the symmetry caused by the uniform input.
+Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}.
+ The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
+%Controlling the \emph{shape}, or relative positions, of many particles is a key ability for a range of applications.
 Correspondingly, it has been studied from a control-theoretic perspective in  both centralized and decentralized approaches.
-For examples of each, see the centralized virtual leaders in \cite{egerstedt2001formation}, and the  gradient-based decentralized controllers  using control-Lyapunov functions in~\cite{hsieh2008decentralized}.
-However, these approaches assume a level of intelligence and autonomy in individual robots that exceeds the capabilities of many systems, including current micro- and nano-robots.
+% For examples of each, see the centralized virtual leaders in \cite{egerstedt2001formation}, and the  gradient-based decentralized controllers  using control-Lyapunov functions in~\cite{hsieh2008decentralized}.
+%However, these approaches assume a level of intelligence and autonomy in individual robots that exceeds the capabilities of many systems, including current micro- and nano-robots.
 Current micro- and nano-robots, such as those in~\cite{Chowdhury2015,martel2015magnetotactic,Xiaohui2015magnetiteMicroswimmers} lack onboard computation.
 
-This paper focuses on centralized techniques that apply the same control input to both particles. 
-Precision control requires breaking the symmetry caused by the uniform input.  
-Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}.
- The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
+%This paper focuses on centralized techniques that apply the same control input to both particles. 
+%Precision control requires breaking the symmetry caused by the uniform input.  
+%Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}.
+% The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
  This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013}, and breaks the control symmetry using obstacles in the workspace.
@@
 %Much research has focused on generating non-uniform artificial force-fields that can be used to rearrange passive components.
+
 Alternative techniques rely on non-uniform inputs, such as artificial force-fields. 
-Applications have included techniques to design shear forces for sensorless manipulation of a single object by~\cite{lamiraux+2001:ra}.  
-\cite{vose2012sliding} demonstrated a collection of 2D force fields generated by six degree-of-freedom vibration inputs to a rigid plate.
+Much recent work in magnet control has focused on exploiting inhomogeneities in the magnetic field to control multiple micro particles using gradient-based pulling~\cite{Salmanipour2018EightDOF,Denasi2018independent}.
+Using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six, Denasi2018independent, Salmanipour2018EightDOF }.
+%Applications have included techniques to design shear forces for sensorless manipulation of a single object by~\cite{lamiraux+2001:ra}.  
+%\cite{vose2012sliding} demonstrated a collection of 2D force fields generated by six degree-of-freedom vibration inputs to a rigid plate.
 These force fields, including shear forces, could be used as a set of primitives for motion control to steer the formation of multiple objects.
@@
 %This paper develops control algorithms using uniform control fields, such as the magnetic resonance navigation \cite{nosrati2018development}.%field in a clinical MRI [insert a recent reference from Sylvain Martel using MRI].
-Similarly, much recent work in magnet control has focused on exploiting inhomogeneities in the magnetic field to control multiple micro particles  using gradient-based pulling~\cite{Salmanipour2018EightDOF,Denasi2018independent}.
-Unfortunately, using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the  distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six, Denasi2018independent, Salmanipour2018EightDOF }.
-In contrast, % to methods that exploit inhomogeneities in the magnetic field to control multiple micro particles, e.g.
+ %In contrast, % to methods that exploit inhomogeneities in the magnetic field to control multiple micro particles, e.g.
 \cite{Denasi2018independent}, that exploited nonlinearities generated by four magnetic coils in close proximity to the workspace to achieve trajectory control of two microspheres,
- this paper requires only a controllable constant gradient in orthogonal directions to position the particles.
+% this paper requires only a controllable constant gradient in orthogonal directions to position the particles.
 % Systems like this one are poorly suited for PRM and RRT*-type methods\cite{lavalle2006planning} because if during a movement a collision occurs, that movement is irreversible.

### 2018-09-10 ICRA19/covControl.tex
removed word robots.  If it shows up again we will know who to blame.

@@
  Center shows two corresponding arrangements of kilobots.
-% At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
-See online interactive demonstration at \cite{Zhao2016mathematicaSquare}.
+At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
+See online interactive demonstration at \cite{Haoran2016SwarminSquare}.
  }
@@
 Left plot shows locus of possible mean positions for four values of $h$.
-%The locus of possible mean positions are concentric circles.
-%See online interactive demonstration at \cite{Zhao2016mathematica}.
+The locus of possible mean positions are concentric circles.
+See online interactive demonstration at \cite{Haoran2016SwarminCircle}.
 }  
@@
 %The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
-We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
-The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
+We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of particles as they move inside a square workspace under the influence of a force pulling in the direction $\beta$.
+The swarm is large, but the particles are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$.
 Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill}.
@@
 \end{align}
-The full equations are summarized in Fig.~\ref{fig:SquareFill}.
+The set of configurations are summarized in Fig.~\ref{fig:SquareFill}.
 For the full equations and demonstration code, see~\cite{Haoran2016SwarminSquare}.
@@
 For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
-Under a global input, the robot swarm fills the region under a chord with area
+Under a global input, the particle swarm fills the region under a chord with area
 \begin{align}
@@
 Eq.
-~\ref{eq:Wall} shows the relation between number of robots and area provided and the correlation we can reach with it.
-%%TODO: ADD the equation for showing the correlation Vs area, vs number of robots
+~\ref{eq:Wall} shows the relation between number of particles and area provided and the correlation we can reach with it.
+%%TODO: ADD the equation for showing the correlation Vs area, vs number of particles
 % This equation shows that we are not able to reach high correlations bigger than $|0.5|$.
@@
 %Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
-Instead of pushing our robots directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the robots. 
+Instead of pushing our particles directly into a wall, the following sections examine an oblique approach  using boundaries that generate friction with the particles. 
  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  
-Robots touching a wall have a friction force that opposes movement along the boundary.  
-This causes robots along the boundary to move more slowly than robots in free-space.
+ Particles touching a wall have a friction force that opposes movement along the boundary.  
+This causes particles along the boundary to move more slowly than particles in free-space.
 
 Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary.
-$N$ is the normal or perpendicular force between the robot and the boundary.
-The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$.
-The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
+$N$ is the normal or perpendicular force between the particle and the boundary.
+The force of friction $F_f$ is nonzero if the particle is in contact with the boundary and  $\sin(\theta) < 0$.
+The resulting net force on the particle, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
 \begin{align}
@@
 \end{align}
- Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
+ Fig.~\ref{fig:friction} shows the resultant forces on two particles when one is touching a wall. 
 Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
-This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
+  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and particles touching the wall are prevented from sliding along the wall.
+This means that if one particle is touching the wall and another particle is free, the touching particle will not move when the control input is parallel or into the wall. 
 There are many alternate models of friction that also break control symmetry.
@@
 
-The next section shows how a system in a rectangularly bounded workspace with friction model \eqref{eq:frictionmodel} can arbitrarily position two robots. 
 \begin{figure}[h]
@@
 \vspace{-1em}
-\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
+\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a particle near a wall, but not for a free particle.
+(c) velocity of a fluid reduces to zero at the boundary.
 \label{fig:friction}
@@
 Larger boundary layers $h$ enable more control of correlation.
-The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
+%  The video attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
 

### 2018-09-10 ICRA19/covControl.tex
made the file not 17mb

@@
  If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
- If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
+ If the swarm size is smaller than $A \approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
 
@@
 \begin{align}
-r_{\text{left}} &= \min (L,1-w),\nonumber \\
+r_{\text{left}} &= \min \left(L,1-w \right),\nonumber \\
 r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
@@
 \caption{\label{fig:ControlledCorrelation}
-Analytical results comparing maximum correlation $\rho_{xy}$ under the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
-}\vspace{-1em}
-\end{figure}
+Analytical results comparing maximum correlation $\rho_{xy}$.
+Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+}\end{figure}
 
-\section{Experiments}
+\section{Experiments}\label{sec:Experiments}
 
@@
 In this section we vary the force and approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
-Kilobot is a three centimeter low cost robot that has light sensor.
+ A Kilobot is a three-centimeter diameter,  low-cost robot with a light sensor.
 In our experiments, these robots were programmed to go toward the brightest light in the room.

### 2018-09-10 ICRA19/intro.tex
removed word robots.  If it shows up again we will know who to blame.
made the file not 17mb

@@
 \section{Introduction}\label{sec:Intro}
-
-   \textcolor{red}{TODO: lets do some experiments with salt in 3D printed shapes (or laser cut).
-Salt is micro scale, and roughly uniform (I think) we can get different magnitudes of global force by varying the tilt angle.}
 
@@
 Shape control of many particles is essential for navigation, construction, and conveying information. 
- The small size of the particles makes it hard or even impossible to control each particle position individually. Instead, these particles are controlled with a single global control input.
- In previous work, we illustrated algorithms that could adjust the position of two particles in an environment with walls that have non-slip contacts \cite{shahrokhi2017}.
-We assumed that if a particle is touching a wall, it will stay still until the uniform input pull it from the wall.
-This work analyses interactions between the boundary and the particle swarm to shape shape of the swarm.
-In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including mean position and variance.
-This work extends that work to shape control.
+ The small size of the particles makes it hard or even impossible to control each particle's position individually. Instead, these particles are often controlled with a single global control input.
+ In previous work, we gave algorithms that could adjust the position of two particles in an environment with walls that have non-slip contacts~\cite{shahrokhi2017}.
+We assumed that if a particle is pushed into a wall, it will not move until a uniform input pulls it from the wall.
+   This work analyses interactions between the rigid obstacles and the particles to control the configuration of the particles.
+    % In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including the mean position and variance. 
+    % This work extends that work to shape control.
 
@@
 The paper is arranged as follows. 
-After a review of recent related work in Sec.  \ref{sec:RelatedWork},
- %Sec.
-\ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
+After a review of recent related work in Sec.~\ref{sec:RelatedWork},
+  Sec.~\ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
+Section \ref{sec:covControl} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
 These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes,
-  Sec.  \ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
-Section  \ref{sec:covControl} describes implementations of the algorithms of covariance control in simulation and hardware experiments. 
+To extend the range of possible shapes, the section explores using linear boundary friction.
+%describes implementations of the algorithms of covariance control in simulation and hardware experiments. 
  % Sec.  \ref{sec:expResults} 
+ Hardware experiments and results are presented in Sec.~\ref{sec:Experiments}.
  We end with directions for future research in Sec.  \ref{sec:conclusion}.

### 2018-09-12 ICRA19/covControl.tex
small clarification of level sets
Arun gave us some TODOs

@@
  Center shows two corresponding arrangements of kilobots.
-At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho(A)$ for a range of $\beta$ values.
+At right is $\bar{x}(A), \sigma_{xy}(A), \sigma_x^2(A),$ and $\rho_{xy}(A)$ for a range of $\beta$ values.
 See online interactive demonstration at \cite{Haoran2016SwarminSquare}.
@@
 The locus of possible mean positions are concentric circles.
+At right is $\sigma_{xy}(h), \sigma_x^2(h),$ and $\rho_{xy}(h)$ for a range of $\beta$ values.
 See online interactive demonstration at \cite{Haoran2016SwarminCircle}.
@@
 One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary.
-Like fluid settling in a tank, the stable final configuration minimizes potential energy.
+Like fluid settling in a tank, the stable final configuration minimizes potential energy by forming a configuration with a level surface perpendicular to the push direction.
 The set of final configurations are parametrized by a single degree of freedom, the global input angle $\beta$.
@@
 Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
-For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
+For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h\in[0,2]$.  
 Under a global input, the particle swarm fills the region under a chord with area
@@
 \subsection{Efficient control of correlation}
+\begin{figure}
+\begin{center}
+	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
+\end{center}
+\vspace{-1em}
+\caption{\label{fig:ControlledCorrelation}
+Analytical results comparing maximum correlation $\rho_{xy}$.
+Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+\textcolor{red}{??label rbottom, tleft, rtop}
+}\end{figure}
+
 %A set of simulations were conducted to demonstrate the importance of boundary friction.  
@@
 %  The video attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
-\end{center}
-\vspace{-1em}
-\caption{\label{fig:ControlledCorrelation}
-Analytical results comparing maximum correlation $\rho_{xy}$.
-Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
-}\end{figure}
 
@@
 
-\textcolor{red}{TODO: Haoran will Explain hardware experiment}
+\textcolor{red}{TODO: Haoran will Explain hardware experiment
+These experiements show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.
+We can also apply varying levels of force and torque, as shown in  \eqref{?}.
+}
 

### 2018-09-12 ICRA19/intro.tex
Arun gave us some TODOs

@@
 
+This paper investigates the set of configurations that can be stably achieved for a  large group of particles, all controlled by the same global force, in a workspace with rigid obstacles.
 Shape control of many particles is essential for navigation, construction, and conveying information. 
  The small size of the particles makes it hard or even impossible to control each particle's position individually. Instead, these particles are often controlled with a single global control input.
- In previous work, we gave algorithms that could adjust the position of two particles in an environment with walls that have non-slip contacts~\cite{shahrokhi2017}.
-We assumed that if a particle is pushed into a wall, it will not move until a uniform input pulls it from the wall.
    This work analyses interactions between the rigid obstacles and the particles to control the configuration of the particles.

### 2018-09-12 ICRA19/relatedWork.tex
a tiny edit
made the file not 17mb
Arun gave us some TODOs

@@
 
-Shape, or relative position, control of many particles is necessary for a range of applications.
+Controlling the relative position of many particles is necessary for a range of applications.
 Therefore, it has been a topic of research from a control perspective in both centralized and distributed approaches.
@@
 Current micro- and nano-robots, such as those in~\cite{Chowdhury2015,martel2015magnetotactic,Xiaohui2015magnetiteMicroswimmers} lack onboard computation.
-
 %This paper focuses on centralized techniques that apply the same control input to both particles. 
@@
  This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013}, and breaks the control symmetry using obstacles in the workspace. 
+  While in \cite{shahrokhi2018TRO} we showed how to control the mean position and variance through repeated collisions with the workspace, this paper focuses on stable configurations. 
+   In \cite[?UniformManip??} we showed that single obstacle can be used to rearrange $n$ particles using an $O(n^2)$-time algorithm.
+   In other previous work, we gave algorithms that could adjust the position of $n$ particles in an environment with walls that have non-slip contacts~\cite{shahrokhi2017} in $O(n^2)$-time.
+We assumed that if a particle is pushed into a wall, it will not move until a uniform input pulls it from the wall.
 

### 2018-09-12 ICRA19/theory.tex
removed word robots.  If it shows up again we will know who to blame.
Arun gave us some TODOs

@@
  defined by a steepest angle of descent perpendicular to the direction of particle motion. 
-This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}.
+This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack  is the \emph{angle of repose}.
 Three different values of angle of repose is shown in Fig.~\ref{fig:angle}.
@@
 By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
+\textcolor{red}{?? TODO: Aaron INsert the integral}
 \begin{figure}
@@
 Each plot shows a statistic of the swarm shape for several angle of repose $\alpha$ values as a function of the approach angle $\beta$ with $\ell = 1$.
-Top plot shows the mean y coordinate $\bar{y}$,
+Top plot shows the mean $y$ coordinate $\bar{y}$,
  middle plot shows $x$-variance $\sigma^2_x$, and

### 2018-09-12 ICRA19/covControl.tex
killed all 'h' for boundary layer, replaced with \ell
thanks arun..

@@
 Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and particles touching the wall are prevented from sliding along the wall.
+%  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and particles touching the wall are prevented from sliding along the wall.
 This means that if one particle is touching the wall and another particle is free, the touching particle will not move when the control input is parallel or into the wall.
@@
 Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.
+Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer of thickness $\ell$ \cite{fluidMechanics}.
 Force in such a system can be calculated as
@@
 %u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{\text{\emph{forward}}}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
+F_{\text{\emph{forward}}}(y) &= F - F_f\begin{cases}  \frac{\ell -y}{\ell}   , &  y<\ell \label{eq:boundarylayerflow} \\
 0, & \text{else} \end{cases}.
@@
 \end{align}
- If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
+ If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $\ell$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
  If the swarm size is smaller than $A \approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.
@@
  Such a rectangular configuration can be accomplished using the variance controllers from \cite{ShahrokhiIROS2015}. 
-  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
+  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction thickness $\ell$ move further than components near the boundary. 
    The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top.
@@
 r_{\text{left}} &= \min \left(L,1-w \right),\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{\text{left}}}{L}  \right).
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, \ell \frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-\ell \frac{r_{\text{left}}}{L}  \right).
 \end{align}
@@
 The mean, variance, and correlation are calculated  using \eqref{eq:meanInSquareWorkspace}, \eqref{eq:varInSquareWorkspace}, and \eqref{eq:covAndcorrInSquareWorkspace} over the region $R$:
-\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{h}y}^{\frac{L}{h}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
+%
+\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{\ell}y}^{\frac{L}{\ell}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
 &+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
-&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy .\nonumber
+&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{\ell}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{\ell}} f(x,y)   dx \, dy .\nonumber
 \end{align}
 
- Given an environment parameterized by $A$ and $h$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
+ Given an environment parameterized by $A$ and $\ell$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
   Negative correlations can be generated by initializing the swarm in the upper left, or lower right.
@@
 Analytical results comparing maximum correlation $\rho_{xy}$.
-Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
 \textcolor{red}{??label rbottom, tleft, rtop}

### 2018-09-13 ICRA19/covControl.tex
tiny caption update
Added Haoran's images

--- 2018-09-12 ICRA19/covControl.tex
@@
 \begin{center}
-	\includegraphics[width=\columnwidth]{angleOfReposeExp.pdf}
+	\includegraphics[width=\columnwidth]{aorPlot.pdf}
 \end{center}
 \caption{\label{fig:angleOfReposeExp}
-Experimental data
+Experimental data using iron particles and garvity.
 }
@@
 
-\begin{figure}
-\begin{center}
-	\includegraphics[width=\columnwidth]{ironAndSaltCloseup.pdf}
-\end{center}
-\caption{\label{fig:ironAndSaltCloseup}
-Magnified image of salt and iron at angle of repose.
-}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%	\includegraphics[width=\columnwidth]{ironAndSaltCloseup.pdf}
+%\end{center}
+%\caption{\label{fig:ironAndSaltCloseup}
+%Magnified image of salt and iron at angle of repose.
+%}
+%\end{figure}
 
@@
 \subsection{Hardware experiment: kilobot angle of repose}
-
-In this section we vary the force and approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
- A Kilobot is a three-centimeter diameter,  low-cost robot with a light sensor.
+We varied the approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
+ A Kilobot is a three-centimeter diameter,  low-cost robot.
 In our experiments, these robots were programmed to go toward the brightest light in the room.
 The experimental setup is shown in Fig.
-\ref{fig:SetUp}.
-Figure \ref{fig:kilobotAngle} shows angle of repose of the Kilobots with three different angles of the pink rod when the rod has 0$^\circ$, 22.5$^\circ$ and 45$^\circ$.
-The green lines show the triangle that the robots pile up.
-The cyan star is the center of mass of the rod, and mean position and variance of the robots are shown with a red star and red ellipse.
+\ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$.
+The green lines show the triangle shape the robots pile up into.
+The cyan star is the center of mass of the rod, and the mean position and variance of the robots are shown with  red stars and red ellipses.
 \begin{figure}
@@
 \caption{\label{fig:SetUp}
-The platform where we control Kilobots with lights.
-}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-	\includegraphics[width=\columnwidth]{KilobotAngleThreeAngles.pdf}
-\end{center}
-\caption{\label{fig:kilobotAngle}
-Angle of repose of Kilobots controlled by light.
+Hardware platform for steering Kilobots with lights and characteristic shape formed by kilobot angle of repose for three angle of attacks.
 }

### 2018-09-13 ICRA19/covControl.tex
update the images and Haoran's experiment

@@
    The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top.
-These regions can be defined by the rectangle's left side, bottom, and top:
+These regions can be defined by the rectangle's left side, bottom, and top (see Fig.~\ref{fig:ControlledCorrelation}):
 \begin{align}
@@
 Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
-\textcolor{red}{??label rbottom, tleft, rtop}
 }\end{figure}
@@
 \section{Experiments}\label{sec:Experiments}
-
+This section reviews three demonstrations of efficient shape control of many particles.
 \subsection{Hardware experiment: particle angle of repose}
-In this section we vary the force and approach angle to shape a swarm with a characteristic angle of repose.
-
+In this section we varied the approach angle to shape a swarm with a characteristic angle of repose.  
+We then showed how the angle of repose can be increased by decreasing the magnitude of the force that moves the particles.
+%
 \begin{figure}
@@
 \caption{\label{fig:angleOfReposeExp}
-Experimental data using iron particles and garvity.
+Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
 }
 \end{figure}
-
+%
 %\begin{figure}
@@
 %\end{figure}
-
-\textcolor{red}{TODO: Haoran will Explain hardware experiment
-These experiements show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.
-We can also apply varying levels of force and torque, as shown in  \eqref{?}.
-}
+%
+The particles used are iron filings (?Company name and particle size?), and the motive force is gravity. 
+The particles were poured into a frame composed of two sheets of acrylic with ?? mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
+ $\{0^\circ,15^\circ, 30^\circ, 45^\circ, 60^\circ \}$, generating angle of attacks $\beta = \{90^\circ,75^\circ, 60^\circ, 45^\circ, 30^\circ \}$.
+ The frame was tested at three tilt angles with respect to horizontal: $\beta = \{90^\circ,60^\circ, 45^\circ \}$, resulting in $\{100,87, 71\}\%$ of the gravitational force.
+ For each test the bottom of the frame was removed to allow unsupported particles to escape and the shape of the remaining particles was measured from photographs taken by a tripod-mounted camera inclined at the same tilt angle as the frame.
+  The decreasing tilt angles resulting in increasing angle of repose values: $\alpha = \{42^\circ,50^\circ, 63^\circ \}$
+These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.
+We can also apply varying levels of force and torque to the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.
 
@@
 \caption{\label{fig:SetUp}
-Hardware platform for steering Kilobots with lights and characteristic shape formed by kilobot angle of repose for three angle of attacks.
+Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks.
 }

### 2018-09-13 ICRA19/relatedWork.tex
killed all 'h' for boundary layer, replaced with \ell
thanks arun..
Added new references.

@@
 Controlling the relative position of many particles is necessary for a range of applications.
-Therefore, it has been a topic of research from a control perspective in both centralized and distributed approaches.
+Therefore, it has been a topic of research from a control perspective in both centralized and decentralized approaches.
+
  This paper focuses on centralized approach that apply the uniform control input to all the particles.
@@
 Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}.
- The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
+Some examples of global inputs include magnetic fields and acoustic radiation forces.
+The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace \cite{nosrati2018development}.
+The periodicity of the acoustic vibration can be exploited for simultaneous positioning \cite{oberti2007manipulation,zhou2016controlling,lipkens2017acoustic}.
+
 %Controlling the \emph{shape}, or relative positions, of many particles is a key ability for a range of applications.
@@
   While in \cite{shahrokhi2018TRO} we showed how to control the mean position and variance through repeated collisions with the workspace, this paper focuses on stable configurations. 
-   In \cite[?UniformManip??} we showed that single obstacle can be used to rearrange $n$ particles using an $O(n^2)$-time algorithm.
+   In \cite{AaronManipulation2013} we showed that single obstacle can be used to rearrange $n$ particles using an $O(n^2)$-time algorithm.
    In other previous work, we gave algorithms that could adjust the position of $n$ particles in an environment with walls that have non-slip contacts~\cite{shahrokhi2017} in $O(n^2)$-time.
@@
 Using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six, Denasi2018independent, Salmanipour2018EightDOF }.
+
+ Decentralized and distributed approaches are also studied.
+An example of distributed approach to shape control a large number of agents has probabilistic swarm guidance using inhomogeneous Markov chains\cite{bandyopadhyay2017probabilistic}.
+Another work considers agents are not communicating and give probabilistic results by only measuring a scalar distance between the agents \cite{li2017decentralized}.
+  Same agents in a group that might behave differently due to their place in the environment inspired by birds or ants \cite{fine2013eliciting}.
+This paper also make use of the workspace to shape the particles.
+However, using autonomous approaches for positioning a large group of particles is not practical in our approach because the particles do not have that level of intelligence. 
 %Applications have included techniques to design shear forces for sensorless manipulation of a single object by~\cite{lamiraux+2001:ra}.  

### 2018-09-14 ICRA19/covControl.tex
reviewed the intro and related work, added info on experiments.

@@
 \begin{center}
-\includegraphics[width=\linewidth]{SquareFill.pdf} 
-\vspace{-1em}
+\includegraphics[width=.8\linewidth]{SquareFill.pdf} 
+\vspace{-.8em}
 \caption{Pushing the swarm against a square boundary wall allows limited control of the shape of the swarm, as a function of swarm area $A$ and the commanded movement direction $\beta$.
@@
 \vspace{-1em}
-\includegraphics[width=\linewidth]{CircleFill.pdf} 
-\vspace{-1em}
+\includegraphics[width=.8\linewidth]{CircleFill.pdf} 
+\vspace{-1.em}
 \caption{Pushing the swarm against a circular boundary wall allows limited control of the shape of the swarm, as a function of the fill level $h$ and the commanded movement direction $\beta$.
@@
 
-\begin{figure}[h]
+\begin{figure}[tb]
 \begin{center}
@@
 \label{fig:friction}
-}\vspace{-1em}
+}\vspace{-2em}
 \end{center}
@@
 \subsection{Efficient control of correlation}
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}
@@
 \end{center}
+\vspace{-1.5em}
 \caption{\label{fig:angleOfReposeExp}
 Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
-}
+}\vspace{-1.em}
 \end{figure}
@@
 %
-The particles used are iron filings (?Company name and particle size?), and the motive force is gravity. 
-The particles were poured into a frame composed of two sheets of acrylic with ?? mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
+The particles used are iron filings (Dowling Magnets, \#731019, $\approx0.5$~mm diameter), and the motive force is gravity. 
+The particles were poured into a frame composed of two sheets of acrylic with 4.5 mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
  $\{0^\circ,15^\circ, 30^\circ, 45^\circ, 60^\circ \}$, generating angle of attacks $\beta = \{90^\circ,75^\circ, 60^\circ, 45^\circ, 30^\circ \}$.
@@
 These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.
-We can also apply varying levels of force and torque to the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.
+These result in varying levels of force and torque on the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.
 
@@
 \end{center}
+\vspace{-1.5em}
 \caption{\label{fig:SetUp}
@@
 
-\begin{figure}
+\begin{figure}[!htb]
 \begin{center}
@@
 \end{center}
-%\vspace{-1em}
+\vspace{-2em}
 \caption{\label{fig:covExperiment}

### 2018-09-14 ICRA19/relatedWork.tex
removed patent citation
reviewed the intro and related work, added info on experiments.

--- 2018-09-10 ICRA19/relatedWork.tex
@@
 
-Shape, or relative position, control of many particles is necessary for a range of applications.
-Therefore, it has been a topic of research from a control perspective in both centralized and distributed approaches.
- This paper focuses on centralized approach that apply the uniform control input to all the particles.
+Controlling the relative position of many particles is necessary for a range of applications.
+Therefore, it has been a topic of research from a control perspective in both centralized and decentralized approaches.
+
+ This paper focuses on centralized approaches that apply uniform control inputs to all the particles.
 To change the relative position of the particles, we need to break the symmetry caused by the uniform input.
 Symmetry can be broken using particles that respond differently to the uniform control signal, either through agent-agent reactions \cite{bertozzi2015ring}, or engineered inhomogeneity  \cite{Donald2013,bretl2007,beckerIJRR2014}.
- The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
+Some examples of global inputs include magnetic fields and acoustic radiation forces.
+The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace \cite{nosrati2018development}.
+Similarly, a single acoustic vibration source can be exploited for simultaneous positioning many particles~\cite{oberti2007manipulation,zhou2016controlling}.
+
 %Controlling the \emph{shape}, or relative positions, of many particles is a key ability for a range of applications.
@@
  This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013}, and breaks the control symmetry using obstacles in the workspace. 
+  While in \cite{shahrokhi2018TRO} we showed how to control the mean position and variance through repeated collisions with the workspace, this paper focuses on stable configurations using few pushes.
+   In \cite{AaronManipulation2013} we showed that single obstacle can be used to rearrange $n$ particles using an $O(n^2)$-time algorithm.
+   In~\cite{shahrokhi2017}, we gave algorithms that could adjust the position of $n$ particles in an environment with walls that have non-slip contacts in $O(n^2)$-time.
+%We assumed that if a particle is pushed into a wall, it will not move until a uniform input pulls it from the wall.
 
 %The techniques in this paper are inspired by artificial force-fields. 
-
 %\emph{Fluid-flow:} 
@@
 % \citep{spears2006physics}~used these forces to disperse a swarm's spatial position for coverage for physics-based swarm simulations.
-
 %\emph{Artificial Force-fields:}
@@
 
-Alternative techniques rely on non-uniform inputs, such as artificial force-fields. 
+Alternative techniques often rely on non-uniform inputs.
  Much recent work in magnet control has focused on exploiting inhomogeneities in the magnetic field to control multiple micro particles using gradient-based pulling~\cite{Salmanipour2018EightDOF,Denasi2018independent}.
-Using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six, Denasi2018independent, Salmanipour2018EightDOF }.
+Using large-scale external magnetic fields makes it challenging to independently control more than one microrobot unless the distance between the electromagnetic coils is at the same length scales as the robot workspace~\cite{diller2016six}.
+
+ Many other shape control techniques rely on agent intelligence.
+ These include distributed approaches where each agent has a map of the desired shape and its position~\cite{bandyopadhyay2017probabilistic,rubenstein2014programmable}; %bandyopadhyay2017probabilistic to shape control a large number of agents has probabilistic swarm guidance using inhomogeneous Markov chains
+  agents that do not communicate, but sense local environmental cues to determine where to spend more time~\cite{li2017decentralized};
+ % Another work considers agents are not communicating and give probabilistic results by only measuring a scalar distance between the agents \cite{li2017decentralized}. 
+ or workspace designs that direct or trap agents as in~\cite{fine2013eliciting}.
+ % Same agents in a group that might behave differently due to their place in the environment inspired by birds or ants \cite{fine2013eliciting}. 
+ %As in \cite{li2017decentralized,fine2013eliciting}, this paper uses the workspace to shape the particles.
+%Because this paper assumes the particles can neither communicate nor compute, a centralized controller is necessary.
+%However, using autonomous approaches for positioning a large group of particles is not practical in our approach because the particles do not have that level of intelligence. 
 %Applications have included techniques to design shear forces for sensorless manipulation of a single object by~\cite{lamiraux+2001:ra}.  

### 2018-09-14 ICRA19/theory.tex
update the images and Haoran's experiment
reviewed the intro and related work, added info on experiments.

@@
 \end{overpic}
+\vspace{-2em}
 \caption{\label{fig:angle}  If particles move faster than the (pink) rod, some particles slide past the rod, but the ones that remain pile up in a shape characterized by the angle of repose $\alpha$, which is particular to the particle type.
 %Different values of angle of repose is shown when granular particles move faster than the rod.
-}
+}\vspace{-1em}
 \end{figure}
@@
  defined by a steepest angle of descent perpendicular to the direction of particle motion. 
-This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack  is the \emph{angle of repose}.
+This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack  is the \emph{angle of repose}\footnote{this paper only considers the 2D projection of the piled particles.}.
+%\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}. 
  Three different values of angle of repose is shown in Fig.~\ref{fig:angle}.
@@
 \subsection{Force and Torque}
-
+%
 By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
-\textcolor{red}{?? TODO: Aaron INsert the integral}
 \begin{figure}
 \centering
-\renewcommand{\figwid}{\columnwidth}
+\renewcommand{\figwid}{0.8\columnwidth}
 \begin{overpic}[width =\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
@@
 \end{figure}
-
+%
 \begin{align}
-F(\theta,\alpha,\ell) =\left\{
+F(\theta,\alpha,\ell) &=  \!\!\!\!\!\!\!\! \int\limits_{-\cos{\theta}\ell/2}\limits^{\cos{\theta}\ell/2} \!\!\!\!\!\!\!\! x\left(
+ \min\!\! \left(\!\! 
+    \frac{ \ell\sin{\alpha \pm\theta }}{\cos{\alpha}}  \mp x \tan{\alpha} \!
+    %, \frac{ \ell\sin{\alpha-\theta }}{\cos{\alpha}}  + x \tan{\alpha}
+ \right)
+ \!-\! x \tan{\theta}\!
+ \right)  dx
+ \nonumber \\
+F(\theta,\alpha,\ell) &=\left\{
 \begin{array}{ll}
 \frac{\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos(\alpha)\sin(\theta)} &   -\alpha<\theta<\alpha\\
-0 &    \textrm{otherwise .}\\
+0 &    \textrm{otherwise .}
 \end{array} 
@@
 \begin{array}{ll}
-\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin(\theta) }{48\sin^2(\alpha)}&   -\alpha<\theta<\alpha\\
-0 &    \textrm{otherwise .}\\
+\frac{\ell^3\Big(\cos{2\alpha}-\cos{2\theta}\Big)\sin{\theta} }{48\sin^2(\alpha)}&   -\alpha<\theta<\alpha\\
+0 &    \textrm{otherwise .} 
 \end{array} 
-\right.
+\right.  \label{eq:TorqueAOR}
 \end{align}
@@
 \begin{align}
-A &=\iint_R \,dx\,dy \text{, }
-\bar{x} =\frac{\iint_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
-\text{, } \bar{y}=\frac{\iint_R y \,dx\,dy}{A} \text{, }\\
+A &=\iint_R \!dx\,dy \text{, }
+\bar{x} =\frac{\iint\limits_R x \,dx\,dy}{A} \label{eq:meanInSquareWorkspace}
+\text{, } \bar{y}=\frac{\iint\limits_R y \,dx\,dy}{A} \text{, }\\
 %\end{align}
 %\begin{align}
-\sigma^2_x &=\frac{\iint_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
-\text{, } \sigma^2_y =\frac{\iint_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}  \text{, }\\
+\sigma^2_x &=\frac{\iint\limits_R \left(x-\bar{x}\right)^2  \,dx \,dy}{A}  \label{eq:varInSquareWorkspace}
+\text{, } \sigma^2_y =\frac{\iint\limits_R  \left(y-\bar{y}\right)^2 \,dx \,dy}{A}  \text{, }\\
 %\end{align}
@@
 \centering
-\renewcommand{\figwid}{\columnwidth}
-\begin{overpic}[width =\figwid]{aorYmean}%\put(1,55){a)}
-\end{overpic}\\
-\vspace{0.5em}
-\begin{overpic}[width =\figwid]{aorVarX}%\put(1,55){a)}
-\end{overpic}\\
-\vspace{0.5em}
-\begin{overpic}[width =\figwid]{aorCovariance}%\put(1,55){a)}
-\end{overpic}
+%\renewcommand{\figwid}{\columnwidth}
+%\begin{overpic}[width =\figwid]{aorYmean}%\put(1,55){a)}
+%\end{overpic}\\
+%\vspace{0.5em}
+%\begin{overpic}[width =\figwid]{aorVarX}%\put(1,55){a)}
+%\end{overpic}\\
+%\vspace{0.5em}
+%\begin{overpic}[width =\figwid]{aorCovariance}%\put(1,55){a)}
+%\end{overpic}
+\begin{overpic}[width =.8\columnwidth]{aorPlots}\end{overpic}
 \vspace{-0.5em}

### 2018-09-14 ICRA19/intro.tex
Added Haoran's images
reviewed the intro and related work, added info on experiments.
updated the first 2 images

@@
 \end{overpic}
+\vspace{-2em}
 \caption{\label{fig:leadfigure} 
-     Reshaping particles by collisions with rigid bodies. Top pictures illustrates using closed boundaries to reshape the particles. Bottom pictures illustrates angle of repose of different particles.
+     Reshaping particles by collisions with rigid bodies. Top pictures illustrate using closed boundaries to reshape the particles. 
+      Bottom pictures illustrate with different particles how varying the the angle of attack modifies the shape of particles remaining on a rigid object and how modifying the applied force changes the angle of repose.
 }
+\vspace{-1em}
 \end{figure}

### 2018-09-15 ICRA19/covControl.tex
cleaned up 'h' vs 'ell' and worked on image 3

@@
 Though each receives the same inputs,  they experience different net forces.
-  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and particles touching the wall are prevented from sliding along the wall.
-This means that if one particle is touching the wall and another particle is free, the touching particle will not move when the control input is parallel or into the wall. 
+%  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and particles touching the wall are prevented from sliding along the wall.
+% This means that if one particle is touching the wall and another particle is free, the touching particle will not move when the control input is parallel or into the wall. 
 There are many alternate models of friction that also break control symmetry.
 Fig.~\ref{fig:friction}c shows fluid flow along a boundary.
-Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.
+Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer of thickness $\ell$ \cite{fluidMechanics}.
 Force in such a system can be calculated as
@@
 %u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
-F_{\text{\emph{forward}}}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
+F_{\text{\emph{forward}}}(y) &= F - F_f\begin{cases}  \frac{\ell-y}{\ell}   , &  y<\ell \label{eq:boundarylayerflow} \\
 0, & \text{else} \end{cases}.
@@
 \caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a particle near a wall, but not for a free particle.
-(c) velocity of a fluid reduces to zero at the boundary.
+(c) Velocity of a fluid reduces to zero at the boundary.
 \label{fig:friction}
@@
  Such a rectangular configuration can be accomplished using the variance controllers from \cite{ShahrokhiIROS2015}. 
-  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer of height $h$ move further than components near the boundary. 
+  If the swarm is then commanded to move a distance $L$ to the right, components of the swarm outside the boundary friction layer thickness $\ell$ move further than components near the boundary. 
    The swarm is contained in a region $R$ composed of no more than three stacked components: at bottom a parallelogram inclined to the right top, at middle a rectangle, and at top a parallelogram inclined to the left top.
@@
 r_{\text{left}} &= \min \left(L,1-w \right),\nonumber \\
-r_{\text{bottom}} &=\min \left(\frac{A}{w}, h\frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
-r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-h\frac{r_{\text{left}}}{L}  \right).
+r_{\text{bottom}} &=\min \left(\frac{A}{w}, \ell\frac{r_{\textrm{left}}}{L}  \right) \text{and}\nonumber \\
+r_{\text{top}}  &= \min \left(\frac{A}{w}, 1-\ell\frac{r_{\text{left}}}{L}  \right).
 \end{align}
@@
 The mean, variance, and correlation are calculated  using \eqref{eq:meanInSquareWorkspace}, \eqref{eq:varInSquareWorkspace}, and \eqref{eq:covAndcorrInSquareWorkspace} over the region $R$:
-\begin{align} \iint_R f(x,y) \, dx \,dy &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{h}y}^{\frac{L}{h}y+w}  f(x,y)  dx \, dy \label{eq:correlationFriction} \\
-&+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)    dx \, dy \nonumber\\
-&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{h}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{h}} f(x,y)   dx \, dy .\nonumber
+\begin{align}
+ \iint_R f(x,y) \, dx \,dy 
+ &=  \int_0^{r_{\text{bottom}}}  \int_{\frac{L}{\ell}y}^{\frac{L}{\ell}y+w}  f(x,y)  \, dx \, dy \label{eq:correlationFriction} \\
+&+\int_{r_{\text{bottom}}}^{r_{\text{top}}} \int_{r_{\text{left}}}^{r_{\text{left}} +w} f(x,y)   \,  dx \, dy \nonumber\\
+&+\int_{r_{\text{top}}}^{\frac{A}{w}} \int_{-\frac{L (y-r_{\text{top}})}{\ell}+r_{\text{left}}}^{r_{\text{left}}+w-\frac{L (y-r_{\text{top}})}{\ell}} f(x,y)  \,  dx \, dy .\nonumber
 \end{align}
 
- Given an environment parameterized by $A$ and $h$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
+ Given an environment parameterized by $A$ and $\ell$, efficient correlation control consists of choosing the $w,L$ pair  that generates the desired positive correlation.
   Negative correlations can be generated by initializing the swarm in the upper left, or lower right.
@@
 Analytical results comparing maximum correlation $\rho_{xy}$.
-Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $h$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
 }\end{figure}
@@
 from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
-The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
+The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho_{xy}(A,\ell,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}.
 For swarms with small area, this method enables generating the full range of correlations $\pm 1$.
@@
   As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
-Larger boundary layers $h$ enable more control of correlation.
+Thicker boundary layers $\ell$ enable more control of correlation.
 %  The video attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.
@@
 
-\subsection{Hardware experiment: kilobot angle of repose}
+\subsection{Hardware experiment: Kilobot angle of repose}
 We varied the approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
@@
 \subsection{Hardware experiment: control of covariance}
-To demonstrate covariance control, up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
+To demonstrate covariance control, up to 100 Kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
@@
 \begin{center}
+\vspace{-1em}
 	\includegraphics[width=\columnwidth]{newExperiment.pdf}
@@
 \caption{\label{fig:covExperiment}
-Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance.
+Hardware demonstration steering $\approx 50$ Kilobot robots to desired covariance.
 The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.

### 2018-09-15 ICRA19/theory.tex
cleaned up 'h' vs 'ell' and worked on image 3

@@
  defined by a steepest angle of descent perpendicular to the direction of particle motion. 
-This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack  is the \emph{angle of repose}\footnote{this paper only considers the 2D projection of the piled particles.}.
-%\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}. 
+This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack $\alpha$ is the \emph{angle of repose}\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}.
  Three different values of angle of repose is shown in Fig.~\ref{fig:angle}.
@@
 \renewcommand{\figwid}{0.8\columnwidth}
-\begin{overpic}[width =\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
+\begin{overpic}[width =\figwid]{AngleOfRepose.pdf}
+\put(39.1,12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}
+\rotatebox{22.5}{\put(55,-12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}}
 \end{overpic}\\

### 2018-09-15 ICRA19/covControl.tex
arun's suggestion about the V.c experiment

@@
 Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
-}\vspace{-1.em}
+}\vspace{-1.5em}
 \end{figure}
@@
 Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks.
-}
+}\vspace{-1.5em}
 \end{figure}
@@
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
+The system has an approximately 10 s lag time before the covariance switches direction.
 
@@
 The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.
-The actual covariance is shown in different trials.
-Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
-%\vspace{-2.5 em}
+The measured covariance is shown for five trials (different line colors).
+Frames above the plot show output from machine vision system and overlaid goal and measured covariance ellipses.
+\vspace{-1.5 em}
 }

### 2018-09-15 ICRA19/covControl.tex
shortened abstract, made experiments correspond with rest of paper

@@
 
-\subsection{Efficient control of correlation}
+\subsection{Efficient control of correlation  $\rho_{xy}$}\label{subsec:efficientCorrelation}
 \begin{figure}[htb]
@@
  For each test the bottom of the frame was removed to allow unsupported particles to escape and the shape of the remaining particles was measured from photographs taken by a tripod-mounted camera inclined at the same tilt angle as the frame.
-  The decreasing tilt angles resulting in increasing angle of repose values: $\alpha = \{42^\circ,50^\circ, 63^\circ \}$
-These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.
-These result in varying levels of force and torque on the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.
+  The decreasing tilt angles resulted in increasing angle of repose values: $\alpha = \{42^\circ,50^\circ, 63^\circ \}$, corresponding to Fig.~\ref{fig:AngleOfReposeForce}(a).
+%%% We removed these because they seemed unnecessary
+% These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle. 
+% These result in varying levels of force and torque on the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.
 
@@
 We varied the approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
- A Kilobot is a three-centimeter diameter,  low-cost robot.
+ A Kilobot is a 3 cm diameter,  low-cost robot.
 In our experiments, these robots were programmed to go toward the brightest light in the room.
 The experimental setup is shown in Fig.
-\ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$.
+\ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$, generating three different shapes, as in Fig.~\ref{fig:angle}.
 The green lines show the triangle shape the robots pile up into.
@@
 
-\subsection{Hardware experiment: control of covariance}
-To demonstrate covariance control, up to 100 Kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
+\subsection{Hardware experiment: control of correlation}
+To demonstrate the correlation control of Sec.~\ref{subsec:efficientCorrelation}, 34 Kilobots were placed on a workspace with boundary layer thickness $\ell \approx 2$ cm and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.
 The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.
@@
 \caption{\label{fig:covExperiment}
-Hardware demonstration steering $\approx 50$ Kilobot robots to desired covariance.
+Hardware demonstration steering $34$ Kilobot robots to desired covariance.
 The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds.

### 2018-09-15 ICRA19/theory.tex
shortened abstract, made experiments correspond with rest of paper

@@
 This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack $\alpha$ is the \emph{angle of repose}\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}.
- Three different values of angle of repose is shown in Fig.~\ref{fig:angle}.
-The center of mass of the rod is in the middle of the rod, but center of mass of the granular particles changes for different values of angle of repose.
+$\alpha$ is a function  of many variables, including the physical shape and composition of the particles and the external force applied.  
+ Three different values of angle of repose are shown in Fig.~\ref{fig:angle}.
+The center of mass (COM) of the rod is in the middle of the rod, but  the COM of the granular particles changes for different $\alpha$ values.
 % for the media, see \cite{angleofRepose}.
- By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
- In this plot, particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
+ The $\alpha$ values can be measured experimentally.   Four types of particles with different angles of repose are shown in Fig.~\ref{fig:AngleOfReposeForce}(a). 
+  Once the angle of repose is known, we can estimate the force and torque that the particles are applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
+ In this plot, particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$-axis. 
   A thin black line extends upwards from the rod COM, and the COM of the particulate is shown by a white and black disk.   
@@
 \begin{overpic}[width =\figwid]{AngleOfRepose.pdf}
+\put(0,55){(a)}
 \put(39.1,12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}
@@
 \begin{overpic}[width =\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
+\put(0,45){(b)}
 \end{overpic}\\
@@
 \begin{overpic}[width =\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
+\put(0,55){(c)}
 \end{overpic}
 \vspace{-0.5em}
-\caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
- Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
-   The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
+\caption{\label{fig:AngleOfReposeForce} (a) plots colored particulate heaped up on pink-colored long rods. 
+ (b) shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
+  (c)  The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
 % Generating code is in the attachment. 
@@
 
-The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}, with the rod length $\ell=1$. 
+The force for different angle of repose values are shown in Fig.~\ref{fig:AngleOfReposeForce}(a), with the rod length $\ell=1$. 
 Torque will also be similarly defined as
@@
 \end{align}
-Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce} with the rod length $\ell=1$. 
+Torque is shown in Fig.~\ref{fig:AngleOfReposeForce}(c) with the rod length $\ell=1$. 
  Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.

### 2018-09-18 journalWallFriction/resubmitv02/Response.tex
Added response document (that is blank -- use the google doc to finish)
Lots of images for Shiva to make wonderful.

--- 2018-07-17 JournalTorqueControl/Review/Response.tex
@@
 
-		\today \vspace{2cm} \\
+		\today \vspace{.1cm} \\
 		\end{flushright}
@@
 		#1\medskip\\
-		\medskip\\
+
 		Dear #2 Editorial Office,\medskip\\
@@
 We are grateful to the reviewers for helping us improve our manuscript through their comments and questions.
-Please let us know if further information is required.\medskip
+Please let us know if further information is required.
+This work is extending the preliminary conference paper, ``Object Manipulation and Position Control Using a Swarm With Global Inputs'', presented at IEEE International Conference on Automation Science and Engineering (CASE) 2016.
+The original preliminary conference paper was a phenological study, whereas the journal version developed analytical models and tested these in a series of new simulations and experiments.\medskip
 
 		There seemed to be confusion from the reviewers about the {\bf note to Practitioners} section. 
+		 This type of note is unique to  IEEE Transactions on Automation Science and Engineering. 
 		 We included this section, as per the instructions at \href{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}, but are happy to remove it if the standards have changed.
-		\\
+		\medskip
+
+		As requested, we have updated our Related Work section to include relevant references that have appeared in the past two years in conferences and journals.  These additions include:
+		\begin{enumerate}
+			\item		F. Ruggiero, V. Lippiello, and B. Siciliano, Nonprehensile dynamic manipulation: A survey, IEEE Robotics and Automation Letters, vol. 3, no. 3, pp. 17111718, July 2018.
+			\item D.
+Loghin, C.
+Tremblay, M.
+Mohammadi, and S.
+Martel, Exploiting the responses of magnetotactic bacteria robotic agents to enhance displacement control and swarm formation for drug delivery platforms, The International Journal of Robotics Research, vol.
+36, no.
+11, pp.
+11951210, 2017.
+			\item S.
+Shahrokhi, L.
+Lin, C.
+Ertel, M.
+Wan, and A.
+T.
+Becker, Steering a swarm of particles using global inputs and swarm statistics, IEEE Transactions on Robotics, vol.
+34, no.
+1, pp.
+207219, 2018.
+			\item  J.
+Alonso-Mora, S.
+Baker, and D.
+Rus, Multi-robot formation control and object transport in dynamic environments via constrained optimization, The International Journal of Robotics Research, vol.
+36, no.
+9, pp.
+10001021, 2017.
+			\item J.
+Zhang, O.
+Onaizah, K.
+Middleton, L.
+You, and E.
+Diller, ÒReliable grasping of three-dimensional untethered mobile magnetic microgripper for autonomous pick-and-place,Ó IEEE Robotics and Automation Letters, vol.
+2, no.
+2, pp.
+835Ð840, 2017.
+		\end{enumerate}
+		\medskip
+
+		The new revision of our paper includes multi-media so that others can build on our results.
+These include illustrative videos for the simulations and experiments: (1) \emph{TorqueSoftware.mp4} and (2) \emph{TorqueHardware.mp4}.
+We also include a Mathematica Notebook containing all the analytical derivations and numerical solutions and corresponding plots \emph{SwarmTorqueControlSubmitV02.nb}, and github.com links to simulation code.\\
 
@@
 \begin{document}
-	\letter{\emph{Object Manipulation and Position Control Using a Swarm With Global Inputs}}{IEEE Transactions on Automation Science and Engineering}
+	\letter{\emph{Exploiting Non-Slip Wall Contacts  to Position Two Particles Using The Same Control Input}}{IEEE Transactions on Robotics}
 
@@
 
-	We have thoroughly revised this paper and included new experimental data. Please see the detailed responses below.
+	We are grateful for the efforts of our reviewers.
+We have thoroughly revised this paper with (1) a more complete derivation section and included (2) new simulation and (3) new hardware experimental data.
+Please see the detailed responses below.
 
@@
 	The reviewer is right that the further particles apply more {\bf torque}, as explained by the new equations 2 and 3. The longer the moment arm, the greater the torque.
-	As explained in section II, ``all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ is a constant for all $i$.''
+	As explained in Section II, `` $\ldots$ all particles are identical and the control input is uniform, the force is equivalent for every particle and so $F_i $ is a constant for all $i$.''
 
@@
 
-	The note to practitioner compares using one particle to using many ``Torque control with only one steered particle is easy: maximize torque by pushing on the object at a location as far from the pivot point as possible.
+	The note to practitioner compares using one particle to using many, ``Torque control with only one steered particle is easy: maximize torque by pushing on the object at a location as far from the pivot point as possible.
 However, a swarm of particles contributes force at different places on the object."  Due to space limitations, this paper does not explore saturation effects.
@@
 
+	The term ``compliant'' has several meanings and was ambiguous, so instead the paper now states ``A key difference is that our swarm of particles do not form a rigid structure and instead tend to flow around the object''.
+This flow is not necessarily a desired characteristic.
+The flowing behavior spreads out the particles, so for some experiments we add a controller that periodically regathers the particles in a workspace corner.
+
 	We added several experiments to better explore torque control with uniform inputs.
-Figs 11 and 12 both feature a free-moving swarm, which as predicted by the central limit theorem, converges to a normal distribution.
-	  To explore other distributions we used an adaptor to constrain the particles.  Fig 9 shows the setup for a swarm forced to assume a uniform distribution.
+Figures 11 and 12 both feature a free-moving swarm, which as predicted by the central limit theorem, converges to a normal distribution.
+	  To explore other distributions we used an adaptor to constrain the particles.  Figure 9 shows the setup for a swarm forced to assume a uniform distribution.
 
@@
 The authors propose static distributions, but the the distributions variate on time due to the flow.}
-	This is a nice future work, but we have discussed in the experiment section that the stochasticity of the system is high and calculating the dynamic distributions is hard.
-Therefore, we approximate the final distribution with normal distribution.
+
+	  For some of our experiments the distributions do vary with respect to time.
+By updating our control law using the current variance, we can generate greater torque.
+Fig.
+6 reports on an experiment using a control law that calculated where to steer the swarm mean as  a function of the current distribution.
 
 	\response{Based on the point above, the most efficient way (and obvious) to generate the torque is just sending all robots as far as possible from the center of mass.}
-	Please pay attention that all the robots are getting the same control input and we are only controlling the mean position.
+If each robot could be steered separately,  sending all robots as far as possible from the center of mass would generate the maximum torque.
+However, since the swarm is controlled by global inputs (we are only controlling the mean position) and the particles have non-zero diameter, we cannot have them all push at the point with the largest moment arm.
 
@@
 
+	We have changed the title so it better matches the paper. 
 	Please refer to Fig 4c and Fig. 5c, which show the maximum torques as a function of standard deviation. 
 	The third column of Table I lists the maximum torques for the three canonical distributions examined.
-
-	We have changed the title so it better matches the paper.
+	Section IV.A.b and IV.B.b are titled ``Maximum torque for distributions''.
 
 	\response{The authors do not mention how to control the robots. It is one of the main points that call the attention in this type of works.} 
-	The particles are getting the same control input, as shown in Eq. 28.
+	Paragraph 2 of the manuscript now states, ``The under-actuated swarm is steered by a shared signal such that the same force is applied to each particle.''
+	Section V. makes this more clear stating `` The control input is the global force applied to each particle:''
+	The equations of motion are given in  (25).
 
@@
 In this way, (2) and (10) are not consistent.
-Assuming that the individual forces are linear is a very strong assumption, in contrast to \[15\], the particles move in a very unorganized way (even more when the particles flow over the object).}
-	We have added definition of n and explained more about our mathematics.
-(2) considers the basic torque when all the particles are touching the object, in (10) we are giving a probability to each particle regarding the swarm distribution.
+Assuming that the individual forces are linear is a very strong assumption, in contrast to [15], the particles move in a very unorganized way (even more when the particles flow over the object).}
+	We have added definition of $n$ as the number of particles in Section III  and explained more about our mathematics. 
+	Equation (2) uses an indicator variable $\rho_i$ to determine which particles contribute to the  torque, ``
+	Not all particles are in contact with the object.  
+The indicator variable $\rho_i$ is 1 if the particle is in direct contact with the object or touching a chain of particle where at least one particle is in contact with the object,  otherwise $\rho_i = 0$.''
+
+	 The equations in section IV integrate over all the particles that contribute to the torque.
 
@@
 	We defined coordinate frames in each section to better represent the equations.
+``Let the marginal distribution of the swarm along $x$ have probability density $p(x)$, where $x$ is defined as perpendicular to the object's long axis.''
 
 	\response{In (16), the function erf is not defined. It is only defined at the end of the section.} 
-	We defined erf function.
+	We now define the erf function as the error function and give the integral that generates it in Section IV-A .
 
 	\response{Based on the granular effect, there is a maximum number of robots that can apply forces without flowing. What is this number? or how can you describe this final configuration?}
-	We have moved granular effect from this paper to better focus the paper.
+	We have removed granular effect from this paper to better focus the paper.
 
@@
 Again, it is necessary to describe how the robots are driven to maintain the distribution and to generate the torque.}
-	We have moved granular effect from this paper to better focus the paper.
+	Correct, but we have removed granular effect from this paper to better focus the paper.
 
@@
 	\response{Caption in Fig 2. says "Particulate" instead of "Particle"}
-	We have removed Fig. 2 to better focus the paper.
+	Correct, but we have removed Fig. 2 to better focus the paper.
 
@@
 }
+	Thank you, and we appreciate the detailed editing suggestions.
 
@@
 	\response{What does cyan/white star points mean in Fig.1?}
-	We have changed Fig. 1 and added text on the figure to define cyan star points.
+	We have changed Fig. 1 and added annotations to the figure to define the cyan markers. The cyan markers are the goal position for the swarm mean.
 
 	\response{In this paper, two different cases are considered. One is a pivoted object, the other is a free object. Differences between two experimental results are encouraged to be highlighted. }
-	We have rewritten the paper and added simulation results to better highlight the experimental results.
+	We have rewritten the paper and added simulation and experimental results to highlight the differences.  Please see Section VI.
 
@@
 Contacts between robots or between robot and rod should be detailed, and their effects should be further discussed if it is possible.}
-	We have discussed the material and size of the rod.
-Although in the theory of this paper we have not considered the robots' contacts with each other to simplify the equations, the experimental results of the last section validated the result of the simplified equations.
+	We now discuss the material and size of the rod in Section VI.A. 
+	Although Section III explains why our analytical section abstracts away the effect of inter robot contacts to simplify the derivations,  the experimental results of the last section validated the result of the simplified equations.
 
 	\response{The definition of 'kilobots' should be given when it is used for the first time. }
-	We have given the description of the kilobots earlier.
+	We have given the description of the Kilobots in Section VI. EXPERIMENTS, when it is used for the first time.
 
@@
 The abstract is poorly written.}
-Thank you. We have rewritten the paper to make it better organized.
+Thank you. We have rewritten the paper to make it better organized, have added experiments, and have performed extensive editing.
 
@@
 The last paragraph about a single robot does not belong in the abstract at all.}
-	We used authors' guidelines to write note to practitioners. If the standards of this journal has changed, we will happily remove that part.
+	We followed the authors' guidelines to write note to practitioners as per the instructions at \href{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}{http://www.ieee-ras.org/publications/t-ase/information-for-authors/50-publications/t-ase/236-t-ase-note-to-practitioners}.
+	The format for a ``note to practitioners'' is deliberately different than an abstract, and have a specialized style.
+	 If the standards of this journal have changed, we will remove this section.
 
@@
 	\response{The term angle of repose should be defined earlier than the end of page 2, because it is used many times before that and while it is defined in [2] it is not a standard term in the robotics community.}
-	We removed this section to better focus the paper.
+	Yes, but we  removed this section to better focus the paper.
 
@@
 
-	\response{The last sentence of Setcion 4 doesn't make sense.  Should the ? -? be an ?=? symbol?}
-	We removed this section to better focus the paper.
+	\response{The last sentence of Section 4 doesn't make sense.  Should the ? -? be an ?=? symbol?}
+	Yes, but we removed this section to better focus the paper.
 
@@
 Is there an expression that would estimate the number of particles, given the particle size?}
-	We removed this section to better focus the paper.
+	Yes, but we removed this section to better focus the paper.
 
@@
 	\response{On page 6 you use an inconsistent style when referring to equations.   At one point you use ?Eq. 16?  and later on the page you use the style ?(18)?}
-	We have now used only eqref style and we don't mention Eq. anymore.
+	We now use only the eqref style and  do not use ``Eq.'' anymore.
 
@@
 ?For all distributions *,*  pushing at $\mu = 1$ ...?}
-	We added comma in that place.
+	We added a comma as suggested.
 

### 2018-09-18 journalWallFriction/TROsubmission01andreview/ReviewsTROv01.tex
Added response document (that is blank -- use the google doc to finish)
Lots of images for Shiva to make wonderful.

@@
 
-## add wording about  neutrally buoyant
+## we added  wording about  neutrally buoyant  ``QUOTE CHANGE"
 
@@
 tripe, and intestines with villi), extending to lower friction values is a intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.
+The reachable sets are smaller -- for example,
+The conclusion now explains `` This paper assumed friction was sufficient to completely stop particles in contact with the boundary.
+The algorithms would require retooling to handle small friction coefficients."
 

### 2018-09-18 journalWallFriction/simulation.tex
Added response document (that is blank -- use the google doc to finish)
Lots of images for Shiva to make wonderful.

@@
 The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
-See multimedia attachment for animations of each.
+See multimedia attachment for animations of each.  \textcolor{red}{$\bigcirc, \blacksquare, \square, \circ$ }
 }
@@
 \caption{\label{fig:contour}
-Plots show performance with one goal on the boundary.
+%Plots show performance with one goal on the boundary.
+Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
+Starting positions of red and blue particles   ??TODO DRAW TICKS Circle and Square?? and goal position of red particle are fixed.
+The top row has the blue particles' goal position  at the origin, generating symmetric contour plots.
+Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.
 }

### 2018-09-18 journalWallFriction/theory2robAlgs.tex
updated fig 8 and 9
Added response document (that is blank -- use the google doc to finish)
Lots of images for Shiva to make wonderful.

@@
 
-This section presents algorithms that use non-slip contacts with walls to arbitrarily position two particles in a convex workspace. 
+This section presents an algorithm, Alg.~\ref{alg:optimalAlg}, that uses non-slip contacts with walls to arbitrarily position two particles in a convex workspace. 
  Workspaces are 2D convex polygons with no internal obstacles. 
@@
  Denote the current positions of the particles  $p_1$ and $p_2$. Values $.x$ and $.y$ denote the $x$ and $y$ coordinates, i.e., $p_1.x$ and $p_1.y$ denote the $x$ and $y$ locations of $p_1$. 
- Algorithm \ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
+  As an improvement over~\cite{shahrokhi2017algorithms}, Alg.~\ref{alg:optimalAlg} can now handle any convex workspace, including the special limit case of a circular workspace.
 In the last subsection we present techniques to control 3D positioning of two particles.
@@
 Figure~\ref{fig:polygonAlg} illustrates the procedure to construct the  2-move reachable set generated by collisions with the $\overline{ p_i p_{i+1}}$ edge.
+%
+
+% worst case analysis
+ This algorithm allows two particles to be steered to arbitrary positions as long as the initial particle positions are separated by at least $\epsilon$, and at least one goal position is $\epsilon$ distance from a wall, where $\epsilon$ is a small, positive, user-defined number.
+For a square workspace where the length of each side is $L$, the worst case path length is $(\sqrt{2}+2)L$, and requires at most five moves.
 % 
@@
   Reachable fraction of  $\Delta$ configuration space  as a function of initial distance $d_{12}$.
+Red vertical lines correspond to the reachable sets for $d_{12} = 0.1$ and $0.3$ in Fig.~\ref{fig:reachableSets}.
 }
@@
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.
-Other workspaces may be better handled by other path planners, such as RRT \cite{LavalleTODO-Shiva}, or \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.
+Workspaces that are tortuous or with many obstacles are better handled by other path planners, such as RRT \cite{LavalleTODO-Shiva}, or \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.

### 2018-10-06 journalWallFriction/SwarmShapeControl.tex
Cleaned up the paper -- full pass

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\definecolor{myMagenta}{RGB}{233,63,246}
+\definecolor{myDarkGreen}{RGB}{97,173,89}
 
@@
 Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
-Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
- This paper investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
+For example, magnetic actuation of particles has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
+ This paper investigates particle control with uniform forces (the same force is applied everywhere in the workspace).
 %Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
-Given a magnetic field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial. 
+Given a controllable field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial. 
 Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
-However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
-In previous work, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
-Because in vivo environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
+However, the walls of \emph{in vivo} and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
+In previous work, we showed that the individual 2D position of two particles is controllable using global inputs in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
+Because \emph{in vivo} environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
 We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that are applied uniformly to all particles in the workspace.

### 2018-10-06 journalWallFriction/experiment.tex
Cleaned up the paper -- full pass

@@
 An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils.
-Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system, focusing on the workspace which was backlit by a $\SI{15}{\watt}$ LED light strip.
+Finally, a FOculus F0134SB $659 \times 494$ pixel camera was attached to the top of the system, focusing on the workspace which was backlit by a $\SI{15}{\watt}$ LED light strip.
 
 To obtain experimental data, the test samples (the phantom intestine model and the bovine cross section) were placed in laser-cut acrylic discs and then immersed in corn syrup.
-Corn syrup was used to increase the viscosity to 12000 cP for the experiments.
+Corn syrup was used to increase the viscosity to $12000$ cP for the experiments.
 Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.
@@
 In particular, the magnetic field in this setup is only approximately uniform.
-The magnetic force varies in both magnitude and orientation.
+The magnetic force is increasingly nonuniform as distance from the center increases in both magnitude and orientation.
 As shown in the video attachment, this non-uniformity causes the particle closer to the coil to move faster than the other particle.
@@
 The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
-Figure \ref{fig:story} top row shows an experiment.
+The top row of Fig.~\ref{fig:story}  shows one experiment.
 Starting and ending positions were printed beneath the workspace on transparency film.
 Our algorithm successfully delivered the particles to goal positions in 10 out of 10 trials.
+A video showing one trial of this experiment is available in the supplementary materials.
 

### 2018-10-06 journalWallFriction/intro.tex
Cleaned up the paper -- full pass

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by a uniform field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
-%If the $i$th particle has position $[x_i,y_i]$ and velocity $[\dot{x}_i, \dot{y}_i]$, then
-%\begin{align}
-%[\dot{x}_i, \dot{y}_i]^\top = \mathbf{u}(t), \qquad i \in [1,n]. \label{eq:swarmDynamics} %[u_x, u_y]^\top
-%\end{align}
-%These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\citep{Purcell1977}. 
-% In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the robots.  
-%  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of robots that move at a constant speed in a direction specified by a global input $\mathbf{u}(t)$~\citep{Rubenstein2012}.
-% Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.
-%As a current example, micro- and nano-robots can be manufactured in large numbers, see~\cite{Chowdhury2015,martel2014computer,kim2015imparting,Donald2013,Ghosh2009,Ou2013,qiu2015magnetic}.
-%Someday large swarms of robots will be remotely guided
-%to assemble structures in parallel and 
-% through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
-% For each task, large numbers of micro robots are required to deliver sufficient payloads, but
-
+Particle swarms propelled by an external field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
+%
 The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
+  Consider a system of point-particles in a 2D, planar workspace.
 Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
@@
 After a review of recent related work in Sec.  \ref{sec:RelatedWork},
- %Sec.
-\ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
-These results are limited in the set of shapes that can be generated.
-To extend the range of possible shapes,
-  Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
-We provide an algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
+  Sec.~\ref{sec:theory} introduces a  model for boundary interaction.   
+We provide an algorithm to arbitrarily position two particles in Sec.~\ref{sec:PostionControl2Robots},  and two shortest path results for representative workspaces in Sec.~\ref{sec:optimalResults}.
  Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 

### 2018-10-06 journalWallFriction/simulation.tex
Cleaned up the paper -- full pass

@@
   Algorithm \ref{alg:optimalAlg} provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.
+Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
   In this section we present two results that can be incorporated into Algorithms \ref{alg:polygonReachbale} and \ref{alg:circularReachbale} to generate shorter motion paths.
@@
  \subsection{Example: Shortest Path in a Square Workspace}\label{subsec:square}
- Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
  If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
@@
 \vspace{-2em}
-\caption{\label{fig:reflection}
-\textcolor{red}{add 0.2 increments to both, add green dot on left image, draw incident and reflected angle}
-The shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle, as shown at left.
+\caption{\label{fig:reflection} In a square workspace, the shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle, as shown at left.
 (right) The first contact is colored red if the red particle is the first to touch a boundary, and colored blue if the blue particle is the first to touch.
@@
 The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
-See multimedia attachment for animations of each.  \textcolor{red}{$\bigcirc, \blacksquare, \square, \circ$ }
+See multimedia attachment for animations of each.
 }
@@
 Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
-Starting positions of red and blue particles   ??TODO DRAW TICKS Circle and Square?? and goal position of red particle are fixed.
-The top row has the blue particles' goal position  at the origin, generating symmetric contour plots.
+Starting positions of red and blue particles  
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
+ and goal position of blue particle  \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex); are fixed.
+The top row has the blue particle's goal position  at the origin, generating symmetric contour plots.
 Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.
@@
 %\vspace{1em}
-\begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{deltadist.pdf}
+\put(65,22){\scriptsize
+$\Delta s$
+\protect\tikz \protect\draw[myDarkGreen,fill=myDarkGreen,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
+\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
+\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
+\end{overpic}
 \vspace{-1em}
@@
 The worst-case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
+ Red fit line is $8.66/(\textrm{distance}^3)$, which has an R-squared value of 0.77.
 } 
@@
 \end{turn}}
-
+%
 \end{overpic}
@@
 \end{overpic}
-}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
+}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ 
+ (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+ \protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
+and goal position of particle $2$ (\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed, and $\epsilon=0.001$.
  The top row of contour plots show the distance if particle $1$'s goal position is varied in $x$ and $y$.
@@
  The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings.
-Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
+The top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and the bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\}, each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
 Fig.~\ref{fig:contour} right shows the number of moves and left shows the total distance of the path.
@@
 Figure \ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles.
+
  Thus far, this paper has considered the particles to be unique.
-If particles are interchangeable, the path lengths often decrease.
+If particles are interchangeable, the path lengths often decrease, which can be computed by running Alg.~\ref{alg:optimalAlg} twice,  but swap the goal positions for the second run and select the shortest path.
 The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.

### 2018-10-06 journalWallFriction/theory.tex
made todos for Shiva
Added response document (that is blank -- use the google doc to finish)
Lots of images for Shiva to make wonderful.
Cleaned up the paper -- full pass

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory}
+\section{Boundary Interaction Model}
 \label{sec:theory}
@@
  Our algorithms rely on holding one particle stationary by pushing it into the boundary while moving the other particle. 
- We start with a boundary interaction model in subsection \ref{subsec:WallFriction}.
-
-For some configurations, we can obtain the optimal solution. 
-Subsections \ref{subsec:square} and \ref{subsec:circular} provide shortest-path results for two representative workspaces, squares and disks.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Boundary Interaction Model}\label{subsec:WallFriction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- The system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
+ These system dynamics can represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
  In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
@@
 \dot{\mathbf{x}}_i(t)
- &=
- \mathbf{u}(t)
- +F \left( \mathbf{x}_i(t), \mathbf{u}(t) \right), ~i \in [1,n].\\
- F(\mathbf{x}_i(t), \mathbf{u}(t)) &= \begin{cases}
-  - \mathbf{u}(t) &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
+ &= \begin{cases}
+  0 &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
 \mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
 \textrm{    ,} \\
- 0 & \textrm{else}.
- \end{cases}\nonumber
+ \mathbf{u}(t) & \textrm{else}.
+ \end{cases}
  \end{align}
- Here  $F(\mathbf{x}_i(t), \mathbf{u}(t)) $ is the frictional force provided by the boundary, and
- $\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$.
+ Here
+ $\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$ and the frictional force provided by the boundary cancels any control force $ \mathbf{u}(t)$ that pushes into the boundary.
 
@@
 \put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
+\put(110,40){ 
+start $s_1$ \tikz \draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+goal $g_1$ \tikz \draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
+start $s_2$ \tikz \draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+goal $g_2$ \tikz \draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);
+}
 \put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
@@
 Solid arrows show path given by  Alg.\ \ref{alg:optimalAlg}.
+Gray areas denote regions unaccessible by our motion planner.
+The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.
 %Online dmonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.

### 2018-10-06 journalWallFriction/theory2robAlgs.tex
Cleaned up the paper -- full pass

@@
 
+\begin{figure}
+\centering
+\begin{overpic}[width=\columnwidth]{irregular.pdf}
+\put(0,10){\scriptsize
+$\Delta s$ \tikz \draw[green,fill=green,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
+\tikz \draw[red,fill=red,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
+\tikz \draw[blue,fill=blue,line width=0.3mm] (-1ex,0) rectangle (0,1ex); 
+}
+\put(0,5){\scriptsize
+$\Delta g$ \tikz \draw[green,fill=green,line width=0.3mm] (0,0) circle (0.5ex); $=$
+\tikz \draw[red,fill=red,line width=0.3mm] (0,0) circle (0.5ex); $-$
+\tikz \draw[blue,fill=blue,line width=0.3mm] (0,0) circle (0.5ex); 
+}
+\end{overpic}
+\caption{\label{fig:irregular}
+Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides.
+The 2-move reachable sets for this initial configuration ($\Delta s= s_2-  s_1$) are drawn in transparent blue.}
+\end{figure}
+
  Algorithm \ref{alg:optimalAlg} assigns a uniform control input at every instance.
@@
 
- \emph{Achieving} a $\Delta g_c$ configuration requires two-moves, the first to move until one particle touches a wall, and the second to adjust the relative spacing.
+ \emph{Achieving} a $\Delta g_c$ configuration requires two moves, the first to move until one particle touches a boundary, and the second to adjust the relative spacing by moving only the particle not touching a boundary.
  Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
@@
 %$(0,0)$ is bottom corner,
- $P$ is a description of the workspace. $\epsilon$ is a positive error bound.
+ $P$ is a description of the workspace. $\epsilon$ is an error bound ($\epsilon>0$).
 % \emph{PathList} contains all the paths sorted by their path length plus an admissible heuristic. 
@@
 \end{algorithm}
-
-\begin{figure}
-\centering
-\begin{overpic}[width=\columnwidth]{irregular.pdf}\end{overpic}
-\caption{\label{fig:irregular}
-%\textcolor{red}{Shiva, this image is nice.  A few suggestions: ?TODO: Label the middle image 'Translate workspace around (0,0)', and draw the workspace outlines with different colors and thicknesses}
-Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides.
-The 2-move reachable sets for this initial configuration (shown by square markers) are drawn in transparent blue.
-}
-\end{figure}
 
@@
 \begin{overpic}[width=0.32\columnwidth]{twoRobotRegionV.pdf}\end{overpic}
-\begin{overpic}[width=0.32\columnwidth]{DeltaConfigSquare.pdf}\end{overpic}
+\begin{overpic}[width=0.32\columnwidth]{DeltaConfigSquare.pdf}
+\put(27,-5){\scriptsize
+$\Delta s$
+\protect\tikz \protect\draw[myDarkGreen,fill=myDarkGreen,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
+\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
+\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
+\end{overpic}
 \caption{\label{fig:TwoRegions}
@@
 (middle) Similarly, if particle 2 contacts the right wall before particle 1 reaches a wall, particle 2 can reach anywhere along the green line, and then particle 1 can move to anywhere in the shaded area.
-(right) The 2-move reachable sets in the $\Delta$ configuration space is shown.
+(right) All 2-move reachable sets in the $\Delta$ configuration space.
 }
@@
  Figure \ref{fig:TwoRegions} highlights the construction of the 2-move reachable sets for a square workspace.
-There are four 2-move reachable sets, but the horizontal (and vertical) reachable sets are equivalent in the $\Delta$ configuration space so we can plan in this space and choose between the two options to minimize the total distance.
+There are four 2-move reachable sets, but the horizontal (and vertical) reachable sets are equivalent in the $\Delta$ configuration space so we can plan in this space and choose between the  options to minimize the total distance.
  % The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
@@
 %\State $R_{\textrm{SET}} \gets$ Append (polygon ($s_2 - (s_1 +d_i)$),$R_{\textrm{SET}}$ )
-\State $R_{\textrm{SET}}$ $\gets$ Append polygon $D$ to $R_{\textrm{SET}}$
+\State $R_{\textrm{SET}}$ $\gets$ Union of polygon $D$ and $R_{\textrm{SET}}$
 \EndFor
@@
 \end{align}
-where $d_{12}= ||s_1 - s_2||$, $r$ is the radius of the workspace,
+where $d_{12}= \Vert s_1 - s_2\Vert_{\scriptscriptstyle 2}$, $r$ is the radius of the workspace,
 and the angle between two particles is $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
@@
 \gamma(\psi) &= \cos^{-1} \Big(1-\frac{d_\perp(\psi)}{r} \Big), \textrm{ where:}\\ \label{eq:dprep}
-d_\perp(\psi)&= 2 ||s_1.p_\psi(\psi) - s_2.p_\psi(\psi)||,\\ \label{eq:ppsi}
+d_\perp(\psi)&= 2 \Vert s_1.p_\psi(\psi) - s_2.p_\psi(\psi)  \Vert_{\scriptscriptstyle 2},\\ \label{eq:ppsi}
 p_\psi(\psi) &= r[\cos(\psi ), \sin(\psi )].
@@
 
-The $\Delta$ configuration makes it easy to compute the accessible workspace.  
+The $\Delta$ configuration enables an iterative method to compute the accessible workspace.  
  Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
@@
    Only antipodal locations are unreachable $(\norm{g_2-g_1} = 1)$, but can be asymptotically approached.
-   Indeed even with a tiny initial separation of $d_{12} = 0.001$, after 14 moves 90\% of the $\Delta$ configuration space is reachable.
+   Indeed, even with a tiny initial separation of $d_{12} = 0.001$, after 14 moves 90\% of the $\Delta$ configuration space is reachable. 
+   In two moves, the maximum reachable fraction of 0.373 is achieved with $d_{12} \approx 0.81$.
 
@@
 \caption{\label{fig:reachableSetsFuncDistance}
-  Reachable fraction of  $\Delta$ configuration space  as a function of initial distance $d_{12}$.
+  Reachable fraction of  $\Delta$ configuration space  as a function of initial distance $d_{12}$ for different numbers of total moves.
 Red vertical lines correspond to the reachable sets for $d_{12} = 0.1$ and $0.3$ in Fig.~\ref{fig:reachableSets}.
@@
 %{0.1`, 0.06275233581754532`, 0.2996219528008345`, 0.5531978612123086`, 0.7453936085358723`, 0.8631378723622217`, 0.9280921766555345`, 0.9614346262938109`}
- Though these images show reachable sets with initial particle-to-particle angle $\theta=0$, all other sets for other $\theta$ values can be formed by rotating by $\theta$.
+ Though these images show reachable sets with initial particle-to-particle angle $\theta=0$, all sets for other $\theta$ values can be formed by rotating these solutions by $\theta$.
 
@@
   Plots showing the 2, 4, 6, 8, 10, 12, and 14-move reachable sets in the  $\Delta$ configuration space for  $d_{12} = 0.1$ and $0.3$.
+The numeric method used for plotting strictly underestimates the reachable set.
 }
@@
 This can be extended to other 3D workspaces if the workspace can be locally approximated as a 3D prism or cylinder.
-Workspaces that are tortuous or with many obstacles are better handled by other path planners, such as RRT \cite{LavalleTODO-Shiva}, or \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.
+Workspaces that are tortuous or with many obstacles are better handled by other path planners, such as RRT \cite{lavalle1998rapidly}, or \cite{AaronManipulation2013}, which used  collisions with  protrusions of the workspace to rearrange particles.

### 2018-10-07 journalWallFriction/resubmitv02/Response.tex
updated letter

--- 2018-07-13 JournalTorqueControl/Review/Response.tex
@@
 \usepackage{hyperref}
+\usepackage{tikz}
+\definecolor{myMagenta}{RGB}{233,63,246}
+\definecolor{myDarkGreen}{RGB}{97,173,89}
 \providecommand{\norm}[1]{\left\lVert#1\right\rVert}
@@
 We are grateful to the reviewers for helping us improve our manuscript through their comments and questions.
-Please let us know if further information is required.\medskip\\
+Please let us know if further information is required.
+This work is extending the preliminary conference paper, ``Algorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts'', presented at the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS).
+The conference paper considered only square workspaces.
+This work extends the analysis to convex workspaces and 3D positioning.
+This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.\medskip
+		\medskip
+
+		The new revision of our paper includes multi-media so that others can build on our results. These include illustrative videos for the simulations and experiments: 
+		??? SHIVA, WHERE ARE THE VIDEOS???
+		(1) \emph{TorqueSoftware.mp4},
+		 (2) \emph{TorqueHardware.mp4},
+		 (3) \emph{video name}.
+		   We also include  Mathematica Notebooks (.nb files) containing the simulation code.\\
 
@@
 	\emph{Comments by Associate Editor}
-	\response{The paper is interesting and technically sound; however, a thorough revision is required before being considered for publication.
-The main concern to be addressed is the quality of the writing.
-Reviewers have provided detailed comments on how to improve the quality of this manuscript.
-Authors are urged to thoroughly revise the paper according to reviewers' comments and suggestions.}
+	\response{The Associate Editor (AE) was able to obtain two high-quality reviews
+from researchers working directly in this field. After considering
+their reviews, it is the opinion of the AE that this paper is not
+currently suitable for publication in T-RO. However, the authors are
+encouraged to prepare a revised draft, addressing the concerns of the
+reviewers.  Please also address how the technique could be practically applied to
+navigate inside an actual 3D intestine as opposed to a cross section in
+which the particles move on a planar surface. Would magnetic forces be
+sufficient to overcome gravity? Would the particles need to be made
+neutrally buoyant?}
 
-	Sure.
+	We are grateful for the efforts of our reviewers.
+We have thoroughly revised this paper by (1) restructuring the paper as suggested by the reviewers (2) constructing reachability sets for all initial conditions, proving controllability as long as the initial positions are distinct and the final positions are not   antipodes, and (3) updating almost all the figures and captions.
 
-	\Reviewer{\#1}\response{This paper studies the pushing effect of a robot swarm over a rectangular object.
-In order to study the generated torque, the authors represent the swarm as a probability distribution.
-Three different probability distributions are presented.
-The ideas in the paper are very interesting and relevant for the community.
-However it requires a major revision.
-The comments are the following.}
-	Thank you!
+	We made the model for both 2D and 3D more clear -- the particles are assumed to be neutrally buoyant so that the same force is applied to each particle.
+	Please see the detailed responses below.
 
-	\response{At the end of Section III, What is the motivation to consider that all robots apply the same force? An intuitive efficient approach would make the farther robots apply a larger force.}
-	Roughly the applied force for the particles are the same, the applied torque would be different for the farther robots.
+	\Reviewer{\#1}\response{This paper explains an underactuated control strategy for small teams
+of magnetic microparticles which move under the influence of a single
+broadcast signal. This underactuated control problem is relevant to
+microrobotic swarms moving under magnetic field input. In previous
+work, the authors have explored interesting control methods relying on
+obstacles in the environment, but the case studied here is more
+application-realistic. The control method here is clever and
+successful. The work is interesting to read and I think that anyone
+with a basic background in controls will find this work interesting.  
+The main concept used here to achieve multi-particle control is that
+the particles experience a no-slip condition when pushed against the
+environment wall. The particles can be removed from the wall by pulling
+them. 
+The paper is an extension of an IROS2017 paper which introduced the
+idea for square workspace only. This TRO submission generalizes to a
+non-square workspace, which is a notable improvement. This version of
+the work will thus be a helpful addition to the literature. 
+The paper gives a very nice review of previous work in using broadcast
+signals for multi-particle control.  }
+	Thank you! We have completed a major revision, as explained below.
 
-	\response{Saturation is not considered.
-An efficient approach would be the same as using a bigger robot but a particle instead.
-Obviously, this approach cannot be done because of the small maximum force that each particle can generate, but it is not mentioned in the paper.}
-	??? I don't understand this sentence completely.
+	\response{Is it expected that the no-slip condition will be a strongly satisfied
+assumption for in-vivo biomedical applications? In cases where there
+were some slip, how could that be handled?}
 
-	\response{The authors claim that one of the main differences is that their robots are compliant and tend to flow over the object.
-However, some of the experiments present an adaptation on the object to avoid the robots flow over the object.}
-	The shape of the particles distribution swarm is not rigid.
-This is good when we need to move the swarm in narrow passageways and crowded environment.
-However, for applying force on the object, we need enough particles to make the required force.
-This is why we try to keep the particles together when we are applying force and torque.
-It is good to mention that we are only controlling the mean position of swarm.
+	To maintain clarity, this paper focused on systems with non-slip boundaries.
+While several environments with non-slip boundaries exist (e.g.
+tripe, and intestines with villi), extending to lower friction values is a intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.
+The reachable sets with non-infinite friction are smaller.
+The conclusion now explains `` This paper assumed friction was sufficient to completely stop particles in contact with the boundary.
+The algorithms would require retooling to handle small friction coefficients."
 
-	\response{This "flow over the object" is not mentioned at all in the methodology (even for the triangular distribution).
-The authors propose static distributions, but the the distributions variate on time due to the flow.}
-	This is a nice future work, but we have discussed in the experiment section that the stochasticity of the system is high and calculating the dynamic distributions is hard.
-Therefore, we approximate the final distribution with normal distribution.
+	\response{This control method requires advance knowledge of the workspace
+boundary geometry, in addition to micro-particle state feedback. It
+would be interesting to hear about how accurate the boundary geometry
+must be known for this control to work well, and perhaps compare this
+will medical imaging techniques such as CT or MRI which could
+potentially be used to generate such maps. For motion through 3D
+lumens, the cross-sectional geometry will change throughout the lumen.}
 
-	\response{Based on the point above, the most efficient way (and obvious) to generate the torque is just sending all robots as far as possible from the center of mass.}
-	Please pay attention that all the robots are getting the same control input and we are only controlling the mean position.
+	These are excellent points.
+This paper assumes perfect knowledge of the boundary and initial conditions of the particles.
+We added text explaining that systems with large numbers of protrusions or concave workspaces are best handled by motion planners such as RRT.
 
-	\response{In contrast to the title, none of the probability distributions maximizes the torque. At least, the maximization problem is not stated in the manuscript.}
-	We have changed the title so it better matches the paper. However, all the torques calculated in the theory section, are the maximized torque.
+	Section IV-F: ``Workspaces that are tortuous or with many obstacles are better handled by other path
+planners, such as RRT [29], or [5], which used collisions with protrusions of the workspace to rearrange particles.''.
 
-	\response{The authors do not mention how to control the robots. It is one of the main points that call the attention in this type of works.} 
-	The particles are getting the same control input.
+	Also, during the hardware experiments, the particle that was pushed into the com stomach boundary disappeared from the camera, but control still was successful, indicating that perfect sensing is not necessary: ``Our algorithm successfully delivered the particles to goal positions in 5 out of 5 trials.''
 
-	\response{In equations (2) and (3), there is a variable n that is not defined.
-I suppose it is the number of the robots, but it is not explicitly denoted.
-In these equations, it is not clear that only the surface robots can apply force to the bar.
-Does $F_i$ takes into account the force that all the robots behind $F_i$ are applying?
-In this way, (2) and (10) are not consistent.
-Assuming that the individual forces are linear is a very strong assumption, in contrast to \[15\], the particles move in a very unorganized way (even more when the particles flow over the object).}
+	\response{Abstract: It is stated that ``given 3 orthogonal magnetic fields''.
+This statement should be made more precise, because the magnetic
+pulling here uses field gradients, not fields. In addition, these
+gradients are not orthogonal. An accurate statement would need to be
+more complex.}
 
-	\response{I recommend to separate Fig.
-2 in sub-panels to organize the caption.
-In the top figure, the particle is not easy to observe and it is not in all the examples.
-The caption says "Particulate is moving in the -y direction", but a coordinate frame is not defined (same for the x-axis).
-Is the particle moving with respect to the inertia frame or to the bar frame?
-Additionally, there is not such a gray line.
-Are the authors referring to the thin black line?
-In general, this figure is not helping to picture what the authors want.}
+	Yes.
+To make this work as broadly applicable as possible (magnetic inputs are just one example) we changed the abstract to say ``This paper investigates particle control with uniform forces (the same force is applied everywhere in the workspace).
+Given a controllable field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial.
+''
 
-	\response{The derivation of (4) and (5) is no shown or extending the description is required.}
+	\response{Fig 2: ``first contact point'' is not clear. There is no way given to
+tell where these line up with actual positions on the workspace. }
 
-	\response{I like the idea of representing the swarm as a probability distribution.
-It is easier for the observer and the global controller to drive the swarm as a whole.
-Again in equations (7)-(9), the coordinate frames are not defined and not consistent.
-The COM $[O_x, O_y]$ is defined in the inertia frame, but the distribution of the robots is defined on the object frame (where the x-axis is parallel to the long side of the rectangle).
-Although these equations are intuitive, the way the authors obtained the constants (e.g.
-the sqrt(3) in Eq.
-(7), only in the next section is assumed that the length is 1) is not obvious.}
+	  Fig 2 is now Fig. 11. We rearranged the order because the optimization results can be applied to Alg.~1, but are not necessary to implement or understand the algorithm.
+	    We updated this figure to clarify what the contact points are by adding and labelling tick marks from 0 to 4 along the boundary of the square workspace and the distance plot.
+We also added angle markings and green points showing the optimal solution in both the workspace and the distance plots.
 
-	\response{In (16), the function erf is not defined. It is only defined at the end of the section.} 
-	It is defined now.
+	\response{Fig 3: the meaning of grey areas should be mentioned in the fig or
+caption.}
+This is now Figure 2.  We added a legend and caption text stating:
+``Gray areas denote regions unaccessible by our motion planner.
+The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.''
 
-	\response{Based on the granular effect, there is a maximum number of robots that can apply forces without flowing. What is this number? or how can you describe this final configuration?}
-	We have moved granular effect from this paper.
+	\response{Figures are hard to read because there is a lot going on, the font is
+often too small and text sometimes is obscured.}
 
-	\response{The experiments section need to be extended, there granular effect need to be extensively studied.
-Again, it is necessary to describe how the robots are driven to maintain the distribution and to generate the torque.}
-	We have moved granular effect from this paper.
+There were a number of problems with the initial figures.
+We have added legends to Figs.
+2, 3, 5, 15, and added captions with visual elements to 2, 14, 16.
+We standardized our variable names and icons, and added Figs~8 and 9.
 
-	\response{According to the format guidelines, the equations are part of the text and should be punctuated. Most of them are not.}
-	We corrected them.
+	\response{III.C. It’s not clear why this section is relevant. Why is the shortest
+path which intersects the wall needed? Some writing structure could
+help the reader understand what the goal is here.}
 
-	\response{Caption in Fig 2. says "Particulate" instead of "Particle"}
+	To better present the motion planner, we moved this section, which covers an optimization result, to the end, and altered the framing text:
+	``Algorithm 1 provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.
+Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
+In this section we present two results that can be incorporated into Algorithms 2 and 3 to generate shorter motion paths.''
 
-	\Reviewer{\#2}
-	\response{This paper studies the torque applied by a swarm of particles on a long aspect-ratio rod.
-Generally, the topic of this paper is interesting.
-Experiments are well designed to illustrate the theory in practice.
-Here are some suggestions to be considered for the possible revision.
-}
+	\Reviewer{\#6}
+	\response{In this manuscript algorithms to position two particles in arbitrary
+locations under uniform actuation are proposed.  The algorithms rely on
+differentiating between particles by contacting them with walls or
+boundaries of the domain, at which they are assumed to not move unless
+actuated in a direction away from the wall.  The paper presents some
+results on shortest paths that contact surfaces (but see below on some
+presentation issues), then describes the algorithm, which basically
+entails moving both particles so that one particle contacts a boundary,
+adjusting the relative displacement between particles while the one
+particle remains at the boundary, then translating both particles to
+the desired location.
+Simulations of the algorithm are presented in square and circular
+domains, and experiments are described in circular cross sections
+inspired by intestinal and stomach scenarios.
+%
+Although similar ideas were presented earlier by the authors using
+either an obstacle in the workspace or square workspaces, this paper
+extends that work to circular and convex polygonal workspaces.	The
+overall idea is a novel contribution as other swarm control techniques
+focus on either heterogeneity of microrobots or of actuation (such as
+fields) rather than distinguishing particles by their proximity to
+boundaries.  However, in my opinion the manuscript requires major
+revisions before being suitable for publication, mainly to address the
+theoretical effectiveness of the algorithm (as detailed below), and
+less so to address clarity of presentation. }
+	Thank you. We were pleased to get this review, because it led us to think more deeply on how to present the results and led to the major revisions detailed below.
 
-	\response{Please add x-y-z axes in all the experimental figures such as Figs.1 and 8.}
+	\response{  The main weakness of the paper in my opinion is that the authors
+never really prove that their algorithm works, or alternatively how
+general their algorithm is, i.e. what are its limits of applicability. 
+To be more specific, in their Algorithm 1, if the desired configuration
+is not within the 2-move reachable set, then the algorithm targets
+instead the closest point in the 2-move reachable set, and then
+"iterates until we reach the goal."  When does such iteration actually
+achieve the goal?  What are the attainable final goals that can be
+reached after iteration?  Does the space of attainable goals depend on
+the initial positions of the particles?  While there are simulations,
+the ones presented do not explore the entire possible space (which is
+understandable since they it is quite large).  I think an analysis of
+the attainable space of their algorithm is needed.}
 
-	\response{What does cyan/white star points mean in Fig.1?}
+	An excellent suggestion!  This was a major oversight in the first version.  It is  possible to explore the possible space, and we can be quite clear about the locations that are attainable.
+	We rearranged the paper and added Section IV-F, which analyzes the reachable set:
 
-	\response{In this paper, two different cases are considered. One is a pivoted object, the other is a free object. Differences between two experimental results are encouraged to be highlighted. }
+	``The $\Delta$ configuration enables an iterative method to compute the accessible workspace.  
+ Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
+ The angle $\theta$ between the initial particle positions simply rotates the reachable $\Delta$ configuration space.
+As long as the initial configurations are distinct $(s_1 \neq s_2)$, the reachable set grows quickly.
+  This relationship is shown in Fig.~15.
+   Only antipodal locations are unreachable $(\norm{g_2-g_1} = 1)$, but can be asymptotically approached.
+   Indeed, even with a tiny initial separation of $d_{12} = 0.001$, after 14 moves 90\% of the $\Delta$ configuration space is reachable. 
+   In two moves, the maximum reachable fraction of 0.373 is achieved with $d_{12} \approx 0.81$."
 
-	\response{Please highlight the material and the size of the rod used in the experiments.
-Contacts between robots or between robot and rod should be detailed, and their effects should be further discussed if it is possible.}
+   We added Figs. 8 and 9 which illustrate the effects of 2-move reachable sets, as well as 4-move, 6, 8, 10 , 12, and 14-moves.
 
-	\response{The definition of 'kilobots' should be given when it is used for the first time. }
+	Fig.
+15 analyzes the worst case: when the goal locations are at the antipodes.
+Even in this situation we can approach arbitrarily close to the desired configuration, but can never achieve it.
+This figure shows that the approximation error decrease can be fit by a 1/(distance travelled$^3$).
 
-	\Reviewer{\#3}	
-	\response{The articles discusses a variant of swarm manipulation where a large number of particle like robots are steered using a common force input.
-The article primarily analyses the force and torque they exert on movable objects they come in contact with; as well as the shape the particles make as they accumulate around these objects (shown to be a triangle whole apex angle is called the angle of repose.
-The strengths of the article are:
-The analysis provided for several manipulation scenarios appears to be sound
-The results are novel to my knowledge.
-Experimental results on a hardware test bed are shown.
-The weaknesses are:
-The analysis can be difficult to follow because in some cases the variables are not well defined.
-There are many, many grammatical and punctuation errors. (see below)
-The abstract is poorly written.}
+	\response{There are issues with the presentation that can be easily improved.
+ At the beginning of the paper many terms are not defined which makes
+figures and discussion hard to follow. Definitions do come later, but
+they should be moved up.  For example, in Fig 2 which symbols are
+targets and which are initial conditions are not defined.  In Fig 3
+epsilon is shown but not defined nor is its significance explained.  s
+and g and $\Delta s$ and $\Delta g$ are not defined. In Fig 12, where is
+the goal on the boundary? Can that be indicated on the figures?}
 
-	\response{Question:  It would seem that the angle of repose might depend on the particle size or the coefficient of friction between particles or the object-particle.
-Can you explain why it doesn?t?
-I understand this isn?t the primary contribution of the article but it seems those parameters would play a role.}
+Thank you for detailed points. We performed a large restructuring. 
+Fig 2 is now 11.
+We updated this figure to clarify what the contact points are by adding and labelling tick marks from 0 to 4 along the boundary of the square workspace and the distance plot.
+We also added angle markings and green points showing the optimal solution in both the workspace and the distance plots.
+We also added visual icons to the caption: ``from starting positions
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
+ to goal positions  
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);)''
 
-	\response{The abstract is repetitive and has a non-standard format (note to practitioners?) and should be restructured.
-It doesn?t make it clear that all the particles are steered with a common force (different from other swarm scenarios).
-It uses the term angle of repose without defining.
-The last paragraph about a single robot does not belong in the abstract at all.}
+ Figure 3 is now 2, we added both a legend to the figure and a descriptive caption: 
+ ``Particles move from start positions 
+(\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);), to goal positions (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
+\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);).
+Dashed lines show the shortest route if particles could be controlled independently.
+Solid arrows show path given by  Alg.~1.
+Gray areas denote regions unaccessible by our motion planner.
+The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.''
 
-	\response{Introduction - paragraph 1: Last sentence doesn?t flow. }
+ Figure 12 had the wrong caption.  It now is Fig. 14, and has the caption: ``Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
+Starting positions of red and blue particles  
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
+ and goal position of blue particle  \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex); are fixed.
+The top row has the blue particle's goal position  at the origin, generating symmetric contour plots.
+Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.''
 
-	\response{The term angle of repose should be defined earlier than the end of page 2, because it is used many times before that and while it is defined in [2] it is not a standard term in the robotics community.}
+We also added captions for start and end (colored squares and circles) to each drawing.
 
-	\response{ Many of the equations are not properly punctuated.  For example, eq 4 and 5 need periods as they end a sentence.  Same with 7,8,9, 12,13, 18, 19, 20 }
+	\response{(more presentation) At the beginning of section IV it wasn't clear
+to me whether Algorithm 1 had been described yet when it was first
+mentioned, and whether that sentence was a description of what was to
+come, or was as assertion that logically followed from the previous
+parts of the paper.  }
 
-	\response{The last sentence of Setcion 4 doesn?t make sense.  Should the ? -? be an ?=? symbol?}
+	We  rewrote this section. It now begins with:
+	``This section presents an algorithm, Alg.~1, that uses non-slip contacts with walls to arbitrarily position two particles in a convex workspace. 
+ Workspaces are 2D convex polygons with no internal obstacles. ''
 
-	\response{This sentence could be worded better: (page 3, col.
-1) ?Given sufficient particles to pile up to the angle of repose??
-also it raises an interesting point that these calculations assume a sufficient number of particles.
-Is there an expression that would estimate the number of particles, given the particle size?}
+	\response{A minor comment, sections IIIB and C are a little confusing since it
+is not motivated why shortest path is being considered in the overall
+argument of the paper.	Furthermore, it seems that shortest path in
+IIIB is used for the situation moving two particles, while in IIIC it
+is used for the situation of moving only one particle.	Consistency in
+the presentation would help..}
 
-	\response{Page 3, col 2, top:
-It?s not clear what $l_p$ and $u_p$ are. (lower and upper bounds?).   Earlier l is used to mean the length of the rod.
-In the first sentence you mention a variable $up_p$, perhaps you mean $u_p$?}
+	We rearranged the order because the optimization results can be applied to Alg.
+1, but are not necessary to implement or understand the algorithm.
+The results were combined into a new Section V, which opens by stating:
 
-	\response{On page 6 you use an inconsistent style when referring to equations.   At one point you use ?Eq. 16?  and later on the page you use the style ?(18)?}
-
-	\response{Caption of figure 3 needs an extra comma depending on the meaning you are trying to convey.
-?For all distributions *,*  pushing at $\mu = 1$ ...?}
-
-	\response{I don?t think the subsection headings should end in ?:?}
-
-	\response{The annotations in Figures 4-5 are small and hard to see.}
-
-	\response{Section 6 paragraph 1:  The word Sections should be capitalized when referring to previous sections by number. }
-
-	\response{Should Kilobots be capitalized? (proper noun)}
-
-	\response{Reference section:   the way you cite ICRA is inconsistent.}
+	``Algorithm 1 provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.
+Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
+  In this section we present two results that can be incorporated into Algorithms 2 and 3 to generate shorter motion paths.''
 

### 2018-10-07 journalWallFriction/simulation.tex
updated letter

@@
 \vspace{-2em}
-\caption{\label{fig:reflection} In a square workspace, the shortest three-move path that reconfigures two particles has the property that the incident angle equals the reflected angle, as shown at left.
+\caption{\label{fig:reflection} In a square workspace, the shortest three-move path that reconfigures two particles from starting positions 
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
+ to goal positions  
+ (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed.
+ has the property that the incident angle equals the reflected angle, as shown at left.
 (right) The first contact is colored red if the red particle is the first to touch a boundary, and colored blue if the blue particle is the first to touch.

### 2018-10-07 journalWallFriction/theory.tex
updated letter

@@
 \caption{\label{fig:shapeControlMathematica1}{Frames from an implementation of Alg.\ \ref{alg:optimalAlg}: two particle positioning using walls with non-slip contacts. 
-Particle start positions are shown by squares, and goal positions by circles.
+Particles move from start positions 
+(\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
+\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);), to goal positions (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
+\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);).
 Dashed lines show the shortest route if particles could be controlled independently.

### 2018-10-09 journalWallFriction/resubmitv02/Response.tex
updated response letter

@@
+% difference between two pdf files: https://draftable.com/
+% insert a pdf: https://tex.stackexchange.com/questions/8662/merge-two-pdf-files-output-by-latex
+%   \usepackage{pdfpages}
+%   \includepdf[pages=-]{ChangesFromV01toV02.pdf}
+%
 \documentclass[conference, onecolumn, 10pt]{../IEEEtran}
@@
 \usepackage{tikz}
+\usepackage{pdfpages}
 \definecolor{myMagenta}{RGB}{233,63,246}
@@
 \providecommand{\norm}[1]{\left\lVert#1\right\rVert}
-
+\pagestyle{plain}
 \newcounter{reviewer}
@@
 \setcounter{point}{0}
+\usepackage{lastpage}
 
@@
 		Dear #2 Editorial Office,\medskip\\
-		Please find attached the revised paper, #1,
+		Please find attached the revised paper, #1, paper number 18-0277,
 		along with the document containing a response to the reviewers.
@@
 Please let us know if further information is required.
-This work is extending the preliminary conference paper, ``Algorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts'', presented at the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS).
+This work extends the preliminary conference paper, ``\emph{Algorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts}'', presented at the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS).
 The conference paper considered only square workspaces.
@@
 		The new revision of our paper includes multi-media so that others can build on our results. These include illustrative videos for the simulations and experiments: 
-		??? SHIVA, WHERE ARE THE VIDEOS???
-		(1) \emph{TorqueSoftware.mp4},
-		 (2) \emph{TorqueHardware.mp4},
-		 (3) \emph{video name}.
-		   We also include  Mathematica Notebooks (.nb files) containing the simulation code.\\
+	\begin{itemize}	 
+ \item Video  \emph{01Model.mov} animates the concepts of uniform control inputs and non-slip boundary contacts.
+ \item Video  \emph{02DeltaConfigurationSpace.mp4} demonstrates how the $\Delta$ configuration space (the configuration space showing the difference in position for two particles) is constructed for a variety of workspaces.
+ \item Video  \emph{03SimulationWorkSpaces.mp4} shows demonstrations of motion planning in square and disc-shaped workspaces. 
+ \item Video  \emph{04Hardware Experiments.mp4} shows an experiment trial moving two particles to goal positions in a small intestine phantom, then an experiment trial using cow stomach tissue.
+		\end{itemize}
+
+		   We also include four  Mathematica Notebooks (.nb files) containing the simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.\\
+
+		   At the end of this letter, pages~\pageref{sec_lastpage}--\pageref{LastPage} is the revised manuscript, formatted so all changes can be easily identified by the reviewers, by using green colored text to indicate revised passages.
 
@@
 
+\label{sec_lastpage}\includepdf[pages=-]{ChangesFromV01toV02.pdf}
+
 \end{document}

### 2018-10-09 journalWallFriction/theory2robAlgs.tex
updated response letter

@@
 For this reason, our algorithms use the two dimensional $\Delta$ configuration space.
-The $\Delta$ configuration space is a set of all possible $\Delta p$ values, defined as the difference in position of the first particle from the second particle: $\Delta p = p_2 - p_1$.
+The $\Delta$ configuration space is a set of all possible $\Delta p$ values, defined as the difference in position of the particles: $\Delta p = p_2 - p_1$.
 We use the $\Delta$ configuration space to plan move sequences that achieve the desired relative spacing.

### 2018-10-11 journalWallFriction/intro.tex
updated thanks to Lillian's comments!

@@
 \caption{\label{fig:IntroPic}
-Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
+Workspace and magnetic setup for an experiment to move one particle from $s_1$ to $g_1$ and a second particle from $s_2$ to $g_2$ when all particles receive the same control inputs, but cannot move while a control input pushes them into a boundary.
 } \vspace{-1em}

### 2018-10-11 journalWallFriction/theory.tex
updated thanks to Lillian's comments!

@@
 {\begin{overpic}[width =\figwid]{story1mov0.pdf}\put(10,10){Start}
-\put(10,85){a)}
+\put(10,85){(a)}
 \put(30,70){$s_1$}
@@
 \begin{overpic}[width =\figwid]{story1move.pdf}\put(10,10){Move 1}
-\put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed}
-\put(110,40){ 
+%\put(110,50){If $(\Delta g- \Delta s) = (0,0)$, only one move is needed.}
+\put(110,70){Legend:}
+\put(120,60){ 
 start $s_1$ \tikz \draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
@@
 }
+\put(110,30){ If $ g_1-  s_1 = g_2-  s_2$, only one move is needed.}
 \put(65, 47){{$\xupdownarrow{1.8cm}$}$L$}
@@
 {\begin{overpic}[width =\figwid]{story3Moves1.pdf}\put(10,10){Start}
-\put(10,85){b)}
+\put(10,85){(b)}
 \put(28,70){$s_1$}
@@
 \begin{overpic}[width =\figwid]{story5move1.pdf}\put(10,10){Move 1}
-\put(10,85){c)}
+\put(10,85){(c)}
 \put(28,70){$s_1$}
@@
 \begin{overpic}[width =\figwid]{story7move1.pdf}\put(50,80){Move 1}
-\put(10,85){d)}
+\put(10,85){(d)}
 \put(25,30){$s_1$}
@@
 \begin{overpic}[width =\figwid]{storySame1.pdf}\put(50,80){Move 1}
-\put(10,85){e)}
+\put(10,85){(e)}
 \put(15,35){$s_1,g_2$}
@@
 \begin{overpic}[width =\figwid]{storyWorst1.pdf}\put(60,10){Move 1}
-\put(10,85){f)}
+\put(10,85){(f)}
 \put(15,10){$s_1,g_2$}
@@
 The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.
+The required number of moves increases from (a) to (f).
 %Online dmonstration and source code at \cite{Shahrokhi2015mathematicaParticle}.

### 2018-10-11 journalWallFriction/theory2robAlgs.tex
updated thanks to Lillian's comments!

@@
 %The following subsection describes how 2-move reachable sets are constructed.
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%\
-\subsection{Two Particle Path-Planning}
-
-  The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
- For the given $\Delta s$ (starting configuration), the rightmost image of Fig.~\ref{fig:irregular} draws the 2-move reachable sets  in transparent blue.
- Figure ~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
-The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
-We will discuss how to make reachable sets in the next subsections.
-
-The goal is to move the particles within $\delta$ of the goal positions using a shared control input where $\delta$ is an arbitrary small number.
-We do this by first moving them within $\delta$ of the correct relative position and then translating the particles to the goal.
-The relative position is $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
-
 \begin{figure}
@@
 
- Algorithm \ref{alg:optimalAlg} assigns a uniform control input at every instance.
-% Algorithm ~\ref{alg:optimalAlg} 
- It first computes the 2-move reachable set.
-If the goal relative position is in the 2-move reachable set, we move particles to achieve that relative position.
-If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$, which is $\Delta g_c$.
-
- \emph{Achieving} a $\Delta g_c$ configuration requires two moves, the first to move until one particle touches a boundary, and the second to adjust the relative spacing by moving only the particle not touching a boundary.
- Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
- Otherwise, we iterate until we reach the goal. 
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%\
+\subsection{Two Particle Path-Planning}
 \begin{algorithm}[htb]
@@
 \end{algorithm}
+
+  The \emph{2-move reachable set} is the locus of points in the $\Delta$ configuration space corresponding to any two-move sequence where the first move brings one particle into contact with the boundary, and the second move translates the second particle without moving the first.
+ For the given $\Delta s$ (starting configuration), the rightmost image of Fig.~\ref{fig:irregular} draws the 2-move reachable sets  in transparent blue.
+ Figure ~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
+The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
+We will discuss how to make reachable sets in the next subsections.
+
+The goal is to use a shared control input to move the particles within $\delta$ of the goal positions, where $\delta$ is an arbitrary small number.
+We do this by first moving them within $\delta$ of the correct relative position and then translating the particles to the goal.
+The relative position is $||\Delta g - \Delta p || = ||(g_2-g_1)- (p_2-p_1)||$.
+
+ Algorithm \ref{alg:optimalAlg} assigns a uniform control input at every instance.
+% Algorithm ~\ref{alg:optimalAlg} 
+ It first computes the 2-move reachable set.
+If the goal relative position is in the 2-move reachable set, we move particles to achieve that relative position.
+If it is not in the 2-move reachable set, we move particles to achieve the closest point on this reachable set from $\Delta g$, which is $\Delta g_c$.
+
+ \emph{Achieving} a $\Delta g_c$ configuration requires two moves, the first to move until one particle touches a boundary, and the second to adjust the relative spacing by moving only the particle not touching a boundary.
+ Once the correct \emph{relative} position has been achieved, a final translation delivers both particles to their goal destinations. %we go to the final goal position. If it is not our final goal, 
+ Otherwise, we iterate until we reach the goal.
 
@@
 \subsection{Circular Workspaces: 2-Move Reachable Set}
+\begin{algorithm}[htb]
+\caption{ { \sc ReachableSetCircle}($s_1,s_2,g_1,g_2$)}\label{alg:circularReachbale}
+\begin{algorithmic}[1]
+%\scriptsize
+\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
+%\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
+\State Calculate $\psi_{\min}$ and $\psi_{\max}$ \Comment use \eqref{eq:psiMinMax}
+\State Calculate $\gamma(\psi)$ \Comment use \eqref{eq:gamma}
+\State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
+\State Return the union of ($l_1, l_2, l_3, l_4$)
+\end{algorithmic}
+\end{algorithm}
 
 % Fig.~\ref{fig:shapeControlMathematica1} shows a Mathematica implementation of the algorithm, and is useful as a visual reference for the following description.
+
+To compute the 2-move reachable set for a circular workspace, first consider all possible first contact locations.
+ The set of boundary points that a particle can touch before the  other particle  touches are two arcs from $\psi_{\min}$ to $\psi_{\max}$  and from $\pi+ \psi_{\min}$ to $\pi+ \psi_{\max}$:
+ \begin{align}\label{eq:psiMinMax}
+%  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\frac{d_{12}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\frac{d_{12}}{2r} \Big],\\
+\scalebox{.9}{$   \psi \in [\psi_{\min}, \psi_{\max}]= \theta \!+  \left[ \sin^{-1} \left(\frac{d_{12}}{2r}  \right)- \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1} \left(\frac{d_{12}}{2r} \right) \right]\!\!,$}
+% \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
+\end{align}
+where $d_{12}= \Vert s_1 - s_2\Vert_{\scriptscriptstyle 2}$, $r$ is the radius of the workspace,
+and the angle between two particles is $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
+
 \begin{figure}
@@
 \end{figure}
-
-To compute the 2-move reachable set for a circular workspace, first consider all possible first contact locations.
- The set of boundary points that a particle can touch before the  other particle  touches are two arcs from $\psi_{\min}$ to $\psi_{\max}$  and from $\pi+ \psi_{\min}$ to $\pi+ \psi_{\max}$:
- \begin{align}\label{eq:psiMinMax}
-%  \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\sin^{-1}\frac{d_{12}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1}\frac{d_{12}}{2r} \Big],\\
-\scalebox{.9}{$   \psi \in [\psi_{\min}, \psi_{\max}]= \theta \!+  \left[ \sin^{-1} \left(\frac{d_{12}}{2r}  \right)- \frac{\pi}{2},  \frac{\pi}{2} -\sin^{-1} \left(\frac{d_{12}}{2r} \right) \right]\!\!,$}
-% \psi \in [\psi_{\min}, \psi_{\max}]= \theta + \Big[\frac{\sin^{-1}{d_{12}}}{2r} - \frac{\pi}{2},  \frac{\pi}{2} -\frac{\sin^{-1}{d_{12}}}{2r} \Big],
-\end{align}
-where $d_{12}= \Vert s_1 - s_2\Vert_{\scriptscriptstyle 2}$, $r$ is the radius of the workspace,
-and the angle between two particles is $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$.
 
@@
 
-\begin{algorithm}[htb]
-\caption{ { \sc ReachableSetCircle}($s_1,s_2,g_1,g_2$)}\label{alg:circularReachbale}
-\begin{algorithmic}[1]
-%\scriptsize
-\Require knowledge of starting $(s_1,s_2)$ and goal $(g_1,g_2)$ positions of  two particles. 
-%\State $\theta = \arctan(\frac{p_1.x-p_2.x}{p_1.y - p_2.y})$
-\State Calculate $\psi_{\min}$ and $\psi_{\max}$ \Comment use \eqref{eq:psiMinMax}
-\State Calculate $\gamma(\psi)$ \Comment use \eqref{eq:gamma}
-\State Calculate $l_1, l_2, l_3, l_4$ \Comment use \eqref{eq:circlereachable} 
-\State Return the union of ($l_1, l_2, l_3, l_4$)
-\end{algorithmic}
-\end{algorithm}
  %Now that $\psi$ is found, we move the particles to make the current goal.
@@
 
-Extending path planning to 3D is possible only if the two particles do not initially have the same $x$ and $y$ positions.
-For ease of analysis, we assume neutrally buoyant particles with workspace boundaries that extend in the $\pm z$ direction to form either right cylinders or right prisms.
-If the 3D projection is at a different angle, redefine the 2D workspace as a region perpendicular to the projection.
+This section presents a method for extending Alg.~\ref{alg:optimalAlg} to 3D in workspaces that are right cylinders or right prisms.
+For ease of analysis, we assume neutrally buoyant particles, and choose a coordinate frame so that the cylinder or prism has end caps in the $xy$ plane.
+ The following method assumes the two particles do not initially have the same $x$ and $y$ positions.
+ The method does not use any contacts with the end caps, and so may be suitable for tubular lumens. 
  First, we move the closest particle to the boundary, which prevents its $z$-coordinate from changing.  

### 2018-10-13 journalWallFriction/resubmitv02/Response.tex
Haoran's changes

@@
 
-		   We also include four  Mathematica Notebooks (.nb files) containing the simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.\\
+		   We also include four  Mathematica Notebooks (.nb files) containing the simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.
+		   	\begin{itemize}	 
+ \item   \emph{SquareWorkSpace.nb} generates paths in a square workspace for two particles.
+ \item  \emph{CircularWorkSpace.nb} generates paths in a circular workspace for two particles.
+ \item   \emph{DeltaConfigurationSpacePolygon.nb} generates the $\Delta$ configuration space for convex polygons
+  \item   \emph{ShortestPathForADisk.nb} analytical solution for shortest path that touches a boundary from one position to another position in a circular workspace.
+		\end{itemize}
 

### 2018-10-13 journalWallFriction/Conclusion.tex
updated pdf info, moved submitted version to folder.
Cleaned up the paper -- full pass
Haoran's changes

@@
 This paper presented techniques for controlling the positions of two particles using uniform inputs and non-slip boundary contacts.  
-The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D $\Delta$ configuration space.
+The paper provided algorithms for precise position control. The algorithms relied on calculating reachable sets in a 2D, $\Delta$ configuration space.
 Extending Alg.~\ref{alg:optimalAlg}  to 3D was straightforward, but increased the complexity.
-Hardware experiments illustrated the algorithms in ex vivo and in artificial workspaces that mimic the geometry of biological tissue.
+Hardware experiments illustrated the algorithms in \emph{ex vivo} and in artificial workspaces that mimic the geometry of biological tissue.
 
-There are several avenues for future work. This paper assumed friction was sufficient to completely stop particles in contact with the boundary. 
+There are several avenues for future work beyond those mentioned previously. This paper assumed friction was sufficient to completely stop particles in contact with the boundary. 
   The algorithms would require retooling to handle small friction coefficients.
 The techniques in \cite{shahrokhi2017algorithms} and  \cite{AaronManipulation2013} could be applied to extend the analysis to more than two particles.
+
+  \section*{Appendix}\label{sec:appendix}
+ The supplemental material includes four videos:
+  	\begin{itemize}	 
+ \item Video  \emph{01Model.mov} animates the concepts of uniform control inputs and non-slip boundary contacts.
+ \item Video  \emph{02DeltaConfigurationSpace.mp4} demonstrates how the $\Delta$ configuration space (the configuration space showing the difference in position for two particles) is constructed for a variety of workspaces.
+ \item Video  \emph{03SimulationWorkSpaces.mp4} shows demonstrations of motion planning in square and disc-shaped workspaces. 
+ \item Video  \emph{04Hardware Experiments.mp4} shows an experiment trial moving two particles to goal positions in a small intestine phantom, then an experiment trial using cow stomach tissue.
+		\end{itemize}	 	
+We also include four  Mathematica Notebooks (.nb files) containing simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.
+		   	\begin{itemize}	 
+ \item   \emph{SquareWorkSpace.nb} generates paths in a square workspace for two particles.
+ \item  \emph{CircularWorkSpace.nb} generates paths in a circular workspace for two particles.
+ \item   \emph{DeltaConfigurationSpacePolygon.nb} generates the $\Delta$ configuration space for convex polygons
+  \item   \emph{ShortestPathForADisk.nb} analytical solution for shortest path that touches a boundary from one position to another position in a circular workspace.
+		\end{itemize}
+
+  %TODO: we also did not do the optimization  (some sort of graph search would be nice)
 

### 2018-10-13 journalWallFriction/experiment.tex
Haoran's changes

@@
 The intestine phantom model was used first and was made to mimic the geometry of an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
+The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and sixty $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
 The top row of Fig.~\ref{fig:story}  shows one experiment.

### 2018-10-13 journalWallFriction/simulation.tex
Haoran's changes

@@
 \caption{\label{fig:deltanumdist}
-The worst-case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
+The worst-case path length occurs when particles must swap antipodes.
+This can never be achieved but can be asymptotically approached.
+Plot shows decreasing error (radius $\delta$ around goal positions) as the number of moves grows.
  Red fit line is $8.66/(\textrm{distance}^3)$, which has an R-squared value of 0.77.
@@
  The worst-case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance).
-This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}.
+This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}, which plots the smallest achievable $\delta$ radius about each goal position as a function of path length.
  Figure \ref{fig:contourPlots} shows the same concepts in a square workspace.

### 2018-10-13 journalWallFriction/theory2robAlgs.tex
Haoran's changes

@@
  This algorithm allows two particles to be steered to arbitrary positions as long as the initial particle positions are separated by at least $\epsilon$, and at least one goal position is $\epsilon$ distance from a wall, where $\epsilon$ is a small, positive, user-defined number.
-For a square workspace where the length of each side is $L$, the worst case path length is $(\sqrt{2}+2)L$, and requires at most five moves.
+For a square workspace where the length of each side is $L$, the worst case path length is $(\sqrt{2}+2)L$, and requires at most five moves (see Fig.~\ref{fig:shapeControlMathematica1}f).
 % 
@@
 \centering
-\begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}\end{overpic}
+\begin{overpic}[width=\columnwidth]{reachableSetCircle.pdf}
+\put(2,96){(a)}
+\put(2,32){(b)}
+\put(48,32){(c)}
+\end{overpic}
 \vspace{-1em}
@@
 The  set of points where the blue particle is the first to contact the boundary are drawn with a blue arc.
-Left top: The possible first contact points for the blue and red particles are shown with blue and red arcs. 
-Left bottom: if the blue particle touches the wall at $\psi_{\min}$ (blue square) the other particle (red square) can move anywhere in the red region. 
-Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the green region. 
-Right: The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
+(a) The possible first contact points for the blue and red particles are shown with blue and red arcs. 
+(b) if the blue particle touches the wall at $\psi_{\min}$ (blue square) the other particle (red square) can move anywhere in the pink region. 
+Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the lightgreen region. 
+(c) The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
 The possible 2-move reachable sets before contact are shown in the $\Delta$ configuration as a blue region.
@@
 \caption{\label{fig:zaxis}
-Illustration on how boundary contacts enable 3D positioning.
+Illustration on how boundary contacts on the sides of a tube enable 3D positioning.
 Once one particle contacts a boundary, the other particle's 2-move reachable set is a prism formed by extending the 2D 2-move reachable set in the $\pm z$ direction.

### 2018-10-13 journalWallFriction/resubmitv02/Response.tex
submitted v 02

@@
-% difference between two pdf files: https://draftable.com/
+% difference between two pdf files: https://draftable.com/compare
 % insert a pdf: https://tex.stackexchange.com/questions/8662/merge-two-pdf-files-output-by-latex
@@
 %   \includepdf[pages=-]{ChangesFromV01toV02.pdf}
+% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=ResponseCompressed.pdf Response.pdf
 %
@@
 		Dear #2 Editorial Office,\medskip\\
-		Please find attached the revised paper, #1, paper number 18-0277,
+		Please find attached the revised paper, #1, previously submitted as paper number 18-0277,
 		along with the document containing a response to the reviewers.
@@
 
-		   At the end of this letter, pages~\pageref{sec_lastpage}--\pageref{LastPage} is the revised manuscript, formatted so all changes can be easily identified by the reviewers, by using green colored text to indicate revised passages.
+		%   At the end of this letter, pages~\pageref{sec_lastpage}--\pageref{LastPage} is the revised manuscript, formatted so all changes can be easily identified by the reviewers, by using green colored text to indicate revised passages.
+		 The revised manuscript is formatted so all changes can be easily identified by the reviewers by using green colored text to indicate revised passages.
 
@@
 While several environments with non-slip boundaries exist (e.g.
-tripe, and intestines with villi), extending to lower friction values is a intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.
+tripe, and intestines with villi), extending to lower friction values is an intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.
 The reachable sets with non-infinite friction are smaller.
@@
 To make this work as broadly applicable as possible (magnetic inputs are just one example) we changed the abstract to say ``This paper investigates particle control with uniform forces (the same force is applied everywhere in the workspace).
-Given a controllable field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial.
+Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial.
 ''
@@
 
-\label{sec_lastpage}\includepdf[pages=-]{ChangesFromV01toV02.pdf}
+%\label{sec_lastpage}\includepdf[pages=-]{ChangesFromV01toV02.pdf}
 

### 2018-10-13 journalWallFriction/Conclusion.tex
submitted v 02

@@
  \item Video  \emph{01Model.mov} animates the concepts of uniform control inputs and non-slip boundary contacts.
- \item Video  \emph{02DeltaConfigurationSpace.mp4} demonstrates how the $\Delta$ configuration space (the configuration space showing the difference in position for two particles) is constructed for a variety of workspaces.
+ \item Video  \emph{02DeltaConfigurationSpace.mp4} demonstrates how the $\Delta$ configuration space is constructed for a variety of workspaces.
  \item Video  \emph{03SimulationWorkSpaces.mp4} shows demonstrations of motion planning in square and disc-shaped workspaces. 
@@
 		\end{itemize}	 	
-We also include four  Mathematica Notebooks (.nb files) containing simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.
+We also include four  Mathematica Notebooks (.nb files) containing simulation code.  
 		   	\begin{itemize}	 

### 2018-10-13 journalWallFriction/SwarmShapeControl.tex
updated response letter
Haoran's changes
submitted v 02

@@
 
+% And here is the link to the videos and summary and read me:
+%  https://www.dropbox.com/home/BoundaryInteractionVideos
+% difference between two pdf files: https://draftable.com/compare
 %\documentclass[conference]{IEEEtran}
@@
 For example, magnetic actuation of particles has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
- This paper investigates particle control with uniform forces (the same force is applied everywhere in the workspace).
-%Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
-Given a controllable field that can generate forces in three orthogonal directions, steering one particle in 3D is trivial. 
+ This paper investigates position control of particles using uniform forces (the same force is applied everywhere in the workspace).
+Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial. 
 Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
@@
 % KEYWORDS:   
-%Underactuated Robots
-%Motion Control
-%Path Planning for Multiple Mobile Robot Systems
+% Underactuated Robots
+% Motion Control
+% Path Planning for Multiple Mobile Robot Systems
+% Configuration Space
 %	

### 2018-10-13 journalWallFriction/experiment.tex
submitted v 02

--- 2018-10-06 journalWallFriction/experiment.tex
@@
 The intestine phantom model was used first and was made to mimic the geometry of an intestine and its villi.
-The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi.
+To model the geometry of intestinal villi, the model consists of a circular ring cut out of $\SI{6}{\milli\metre}$ thick acrylic.
+The ring has an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and sixty $\SI{2}{\milli\metre}$ long protrusions on its inner surface.
 The top row of Fig.~\ref{fig:story}  shows one experiment.

### 2018-11-20 journalWallFriction/ReviewConditionalAccept/Response.tex
added minor revision Review for TRO

--- 2018-10-13 journalWallFriction/resubmitv02/Response.tex
@@
 	\emph{Comments by Associate Editor}
-	\response{The Associate Editor (AE) was able to obtain two high-quality reviews
-from researchers working directly in this field. After considering
-their reviews, it is the opinion of the AE that this paper is not
-currently suitable for publication in T-RO. However, the authors are
-encouraged to prepare a revised draft, addressing the concerns of the
-reviewers.  Please also address how the technique could be practically applied to
-navigate inside an actual 3D intestine as opposed to a cross section in
-which the particles move on a planar surface. Would magnetic forces be
-sufficient to overcome gravity? Would the particles need to be made
-neutrally buoyant?}
+	\response{Both reviewers were pleased with the authors' revised manuscript. It is
+the opinion of the Associate Editor that this paper will be publishable
+in T-RO in some form. However, both reviewers have lingering concerns
+about discrepancies between the purported contribution of the paper and
+what is actually demonstrated. In the next draft of the manuscript, the
+authors are encouraged to be explicit in making sure that the stated
+contributions match what has been shown. This could involve removing or
+softening certain claims, or adding new material to strengthen certain
+claims.}
 
 	We are grateful for the efforts of our reviewers.
-We have thoroughly revised this paper by (1) restructuring the paper as suggested by the reviewers (2) constructing reachability sets for all initial conditions, proving controllability as long as the initial positions are distinct and the final positions are not   antipodes, and (3) updating almost all the figures and captions.
 
-	We made the model for both 2D and 3D more clear -- the particles are assumed to be neutrally buoyant so that the same force is applied to each particle.
-	Please see the detailed responses below.
+	\Reviewer{\#1}\response{In the revised version most of my previous concerns have been
+addressed.  I think with minor revisions the manuscript could be
+suitable for publication. }
+	Thank you! We have completed a minor revision, as explained below.
 
-	\Reviewer{\#1}\response{This paper explains an underactuated control strategy for small teams
-of magnetic microparticles which move under the influence of a single
-broadcast signal. This underactuated control problem is relevant to
-microrobotic swarms moving under magnetic field input. In previous
-work, the authors have explored interesting control methods relying on
-obstacles in the environment, but the case studied here is more
-application-realistic. The control method here is clever and
-successful. The work is interesting to read and I think that anyone
-with a basic background in controls will find this work interesting.  
-The main concept used here to achieve multi-particle control is that
-the particles experience a no-slip condition when pushed against the
-environment wall. The particles can be removed from the wall by pulling
-them. 
-The paper is an extension of an IROS2017 paper which introduced the
-idea for square workspace only. This TRO submission generalizes to a
-non-square workspace, which is a notable improvement. This version of
-the work will thus be a helpful addition to the literature. 
-The paper gives a very nice review of previous work in using broadcast
-signals for multi-particle control.  }
-	Thank you! We have completed a major revision, as explained below.
+	\response{The remaining issue still comes back to whether the iterative procedure
+is able to reach all target configurations.  The new manuscript shows
+numerically that the accesible space approaches the entire space for a
+circular 2D domain.  The paper, however, is meant to also cover any
+convex domains.  Can the authors also prove that the accessible space
+approaches the entire domain for any convex domain?  Or at least other
+polygonal cases?   I see that on p4 in the last paragraph of section C,
+there is a statement that particles can be steered to arbitrary
+positions, but without an explanation of how that result is obtained,
+which should be included.  I also think that paragraph should be
+highlighted a little more by a section title (to be parallel with the
+development of the circular workspace).  Assuming this can be explained
+then I would be satisfied with the manuscript.}
 
-	\response{Is it expected that the no-slip condition will be a strongly satisfied
-assumption for in-vivo biomedical applications? In cases where there
-were some slip, how could that be handled?}
-
-	To maintain clarity, this paper focused on systems with non-slip boundaries.
-While several environments with non-slip boundaries exist (e.g.
-tripe, and intestines with villi), extending to lower friction values is an intriguing avenue that unfortunately does not fit within the 12 page limit for this journal, but is something we are investigating.
-The reachable sets with non-infinite friction are smaller.
-The conclusion now explains `` This paper assumed friction was sufficient to completely stop particles in contact with the boundary.
-The algorithms would require retooling to handle small friction coefficients."
-
-	\response{This control method requires advance knowledge of the workspace
-boundary geometry, in addition to micro-particle state feedback. It
-would be interesting to hear about how accurate the boundary geometry
-must be known for this control to work well, and perhaps compare this
-will medical imaging techniques such as CT or MRI which could
-potentially be used to generate such maps. For motion through 3D
-lumens, the cross-sectional geometry will change throughout the lumen.}
-
-	These are excellent points.
-This paper assumes perfect knowledge of the boundary and initial conditions of the particles.
-We added text explaining that systems with large numbers of protrusions or concave workspaces are best handled by motion planners such as RRT.
-
-	Section IV-F: ``Workspaces that are tortuous or with many obstacles are better handled by other path
-planners, such as RRT [29], or [5], which used collisions with protrusions of the workspace to rearrange particles.''.
-
-	Also, during the hardware experiments, the particle that was pushed into the com stomach boundary disappeared from the camera, but control still was successful, indicating that perfect sensing is not necessary: ``Our algorithm successfully delivered the particles to goal positions in 5 out of 5 trials.''
-
-	\response{Abstract: It is stated that ``given 3 orthogonal magnetic fields''.
-This statement should be made more precise, because the magnetic
-pulling here uses field gradients, not fields. In addition, these
-gradients are not orthogonal. An accurate statement would need to be
-more complex.}
-
-	Yes.
-To make this work as broadly applicable as possible (magnetic inputs are just one example) we changed the abstract to say ``This paper investigates particle control with uniform forces (the same force is applied everywhere in the workspace).
-Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial.
-''
-
-	\response{Fig 2: ``first contact point'' is not clear. There is no way given to
-tell where these line up with actual positions on the workspace. }
-
-	  Fig 2 is now Fig. 11. We rearranged the order because the optimization results can be applied to Alg.~1, but are not necessary to implement or understand the algorithm.
-	    We updated this figure to clarify what the contact points are by adding and labelling tick marks from 0 to 4 along the boundary of the square workspace and the distance plot.
-We also added angle markings and green points showing the optimal solution in both the workspace and the distance plots.
-
-	\response{Fig 3: the meaning of grey areas should be mentioned in the fig or
-caption.}
-This is now Figure 2.  We added a legend and caption text stating:
-``Gray areas denote regions unaccessible by our motion planner.
-The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.''
-
-	\response{Figures are hard to read because there is a lot going on, the font is
-often too small and text sometimes is obscured.}
-
-There were a number of problems with the initial figures.
-We have added legends to Figs.
-2, 3, 5, 15, and added captions with visual elements to 2, 14, 16.
-We standardized our variable names and icons, and added Figs~8 and 9.
-
-	\response{III.C. It’s not clear why this section is relevant. Why is the shortest
-path which intersects the wall needed? Some writing structure could
-help the reader understand what the goal is here.}
-
-	To better present the motion planner, we moved this section, which covers an optimization result, to the end, and altered the framing text:
-	``Algorithm 1 provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.
-Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
-In this section we present two results that can be incorporated into Algorithms 2 and 3 to generate shorter motion paths.''
-
-	\Reviewer{\#6}
-	\response{In this manuscript algorithms to position two particles in arbitrary
-locations under uniform actuation are proposed.  The algorithms rely on
-differentiating between particles by contacting them with walls or
-boundaries of the domain, at which they are assumed to not move unless
-actuated in a direction away from the wall.  The paper presents some
-results on shortest paths that contact surfaces (but see below on some
-presentation issues), then describes the algorithm, which basically
-entails moving both particles so that one particle contacts a boundary,
-adjusting the relative displacement between particles while the one
-particle remains at the boundary, then translating both particles to
-the desired location.
-Simulations of the algorithm are presented in square and circular
-domains, and experiments are described in circular cross sections
-inspired by intestinal and stomach scenarios.
-%
-Although similar ideas were presented earlier by the authors using
-either an obstacle in the workspace or square workspaces, this paper
-extends that work to circular and convex polygonal workspaces.	The
-overall idea is a novel contribution as other swarm control techniques
-focus on either heterogeneity of microrobots or of actuation (such as
-fields) rather than distinguishing particles by their proximity to
-boundaries.  However, in my opinion the manuscript requires major
-revisions before being suitable for publication, mainly to address the
-theoretical effectiveness of the algorithm (as detailed below), and
-less so to address clarity of presentation. }
+	\Reviewer{\#2}
+	\response{The authors did a great job in addressing my and the other reviewer's
+comments in a set of meaningful changes to the manuscript, including
+new analysis to investigate many of the questions. 
+The only remaining comment is that the abstract and introduction imply
+that 3D results are obtained in this work. While this is shown
+theoretically, no 3D experimental results are shown. The abstract and
+introduction should state this explicitly as this is viewed as a
+notable limitation in what is actually demonstrated. }
 	Thank you. We were pleased to get this review, because it led us to think more deeply on how to present the results and led to the major revisions detailed below.
 
-	\response{  The main weakness of the paper in my opinion is that the authors
-never really prove that their algorithm works, or alternatively how
-general their algorithm is, i.e. what are its limits of applicability. 
-To be more specific, in their Algorithm 1, if the desired configuration
-is not within the 2-move reachable set, then the algorithm targets
-instead the closest point in the 2-move reachable set, and then
-"iterates until we reach the goal."  When does such iteration actually
-achieve the goal?  What are the attainable final goals that can be
-reached after iteration?  Does the space of attainable goals depend on
-the initial positions of the particles?  While there are simulations,
-the ones presented do not explore the entire possible space (which is
-understandable since they it is quite large).  I think an analysis of
-the attainable space of their algorithm is needed.}
-
-	An excellent suggestion!  This was a major oversight in the first version.  It is  possible to explore the possible space, and we can be quite clear about the locations that are attainable.
-	We rearranged the paper and added Section IV-F, which analyzes the reachable set:
-
-	``The $\Delta$ configuration enables an iterative method to compute the accessible workspace.  
- Due to symmetry of the workspace, the fraction of the $\Delta$ configuration space reachable in $2 k$ moves is a function of only the initial separation distance $d_{12}$.
- The angle $\theta$ between the initial particle positions simply rotates the reachable $\Delta$ configuration space.
-As long as the initial configurations are distinct $(s_1 \neq s_2)$, the reachable set grows quickly.
-  This relationship is shown in Fig.~15.
-   Only antipodal locations are unreachable $(\norm{g_2-g_1} = 1)$, but can be asymptotically approached.
-   Indeed, even with a tiny initial separation of $d_{12} = 0.001$, after 14 moves 90\% of the $\Delta$ configuration space is reachable. 
-   In two moves, the maximum reachable fraction of 0.373 is achieved with $d_{12} \approx 0.81$."
-
-   We added Figs. 8 and 9 which illustrate the effects of 2-move reachable sets, as well as 4-move, 6, 8, 10 , 12, and 14-moves.
-
-	Fig.
-15 analyzes the worst case: when the goal locations are at the antipodes.
-Even in this situation we can approach arbitrarily close to the desired configuration, but can never achieve it.
-This figure shows that the approximation error decrease can be fit by a 1/(distance travelled$^3$).
-
-	\response{There are issues with the presentation that can be easily improved.
- At the beginning of the paper many terms are not defined which makes
-figures and discussion hard to follow. Definitions do come later, but
-they should be moved up.  For example, in Fig 2 which symbols are
-targets and which are initial conditions are not defined.  In Fig 3
-epsilon is shown but not defined nor is its significance explained.  s
-and g and $\Delta s$ and $\Delta g$ are not defined. In Fig 12, where is
-the goal on the boundary? Can that be indicated on the figures?}
-
-Thank you for detailed points. We performed a large restructuring. 
-Fig 2 is now 11.
-We updated this figure to clarify what the contact points are by adding and labelling tick marks from 0 to 4 along the boundary of the square workspace and the distance plot.
-We also added angle markings and green points showing the optimal solution in both the workspace and the distance plots.
-We also added visual icons to the caption: ``from starting positions
- (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
- \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
- to goal positions  
- (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
- \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);)''
-
- Figure 3 is now 2, we added both a legend to the figure and a descriptive caption: 
- ``Particles move from start positions 
-(\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
-\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);), to goal positions (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
-\protect\tikz \protect\draw[myMagenta,fill=white,line width=0.3mm] (0,0) circle (.5ex);).
-Dashed lines show the shortest route if particles could be controlled independently.
-Solid arrows show path given by  Alg.~1.
-Gray areas denote regions unaccessible by our motion planner.
-The particle start positions must be distinct $(\norm{s_2-s_1} \geq \epsilon)$, and at least one goal position must be farther than  $\epsilon$  from the boundary, where $\epsilon$ is a small but nonzero user-specified constant.''
-
- Figure 12 had the wrong caption.  It now is Fig. 14, and has the caption: ``Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
-Starting positions of red and blue particles  
- (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
- \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);)
- and goal position of blue particle  \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex); are fixed.
-The top row has the blue particle's goal position  at the origin, generating symmetric contour plots.
-Moving the  blue particles' goal position  to $(-0.2,0)$, generates non-symmetric contour plots.''
-
-We also added captions for start and end (colored squares and circles) to each drawing.
-
-	\response{(more presentation) At the beginning of section IV it wasn't clear
-to me whether Algorithm 1 had been described yet when it was first
-mentioned, and whether that sentence was a description of what was to
-come, or was as assertion that logically followed from the previous
-parts of the paper.  }
-
-	We  rewrote this section. It now begins with:
-	``This section presents an algorithm, Alg.~1, that uses non-slip contacts with walls to arbitrarily position two particles in a convex workspace. 
- Workspaces are 2D convex polygons with no internal obstacles. ''
-
-	\response{A minor comment, sections IIIB and C are a little confusing since it
-is not motivated why shortest path is being considered in the overall
-argument of the paper.	Furthermore, it seems that shortest path in
-IIIB is used for the situation moving two particles, while in IIIC it
-is used for the situation of moving only one particle.	Consistency in
-the presentation would help..}
-
-	We rearranged the order because the optimization results can be applied to Alg.
-1, but are not necessary to implement or understand the algorithm.
-The results were combined into a new Section V, which opens by stating:
-
-	``Algorithm 1 provided a technique to bring two particles to goal positions using global inputs, but did not optimize path length.
-Changing the relative positions of particles in any workspace requires making one particle contact the boundary.
-  In this section we present two results that can be incorporated into Algorithms 2 and 3 to generate shorter motion paths.''
-
-%\label{sec_lastpage}\includepdf[pages=-]{ChangesFromV01toV02.pdf}
-
 \end{document}

### 2018-11-20 journalWallFriction/SwarmShapeControl.tex
made some work for Shiva, put all the text revision we think we need, except to refer to new Fig. 5

@@
 
+\newcommand{\revision}[1]{
+  \textcolor{Blue}{#1}
+}
+
 \pdfinfo{
@@
 In previous work, we showed that the individual 2D position of two particles is controllable using global inputs in a square workspace with non-slip wall contact \cite{shahrokhi2017algorithms}.
-Because \emph{in vivo} environments are usually not square, this paper extends the previous work to all convex workspaces and to 3D positioning. 
+Because \emph{in vivo} environments are usually not square, this paper extends the previous work to all convex workspaces, \revision{ and shows how this could be extended} to 3D positioning \revision{of neutrally buoyant particles.}
 We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that are applied uniformly to all particles in the workspace.
-This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.
+This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the gastrointestinal tract.
 

### 2018-11-20 journalWallFriction/intro.tex
Haoran's changes
made some work for Shiva, put all the text revision we think we need, except to refer to new Fig. 5

@@
  The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the robot as the \emph{system} that consists of particles, a uniform control field, and sensing.
-  Consider a system of point-particles in a 2D, planar workspace.
+  Consider a system of point-particles in a 2D planar workspace.
 Such systems are severely underactuated, having 2 degrees of freedom in the shared planar control input, but $2n$ degrees of freedom for the $n$-particle swarm.
@@
 This work extends the analysis to convex workspaces and 3D positioning.
-This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.
+This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-11-20 journalWallFriction/theory2robAlgs.tex
made some work for Shiva, put all the text revision we think we need, except to refer to new Fig. 5

@@
 
+\subsection{\revision{Convex Workspaces: Accessible Region}}
 % worst case analysis
@@
 For a square workspace where the length of each side is $L$, the worst case path length is $(\sqrt{2}+2)L$, and requires at most five moves (see Fig.~\ref{fig:shapeControlMathematica1}f).
+As the corner angles increase, the number of moves required to access the entire configuration space increases.
+\revision{For a circular workspace, with corner angles of 180$^\circ$, the worst case configuration can only be approached asymptotically, as explained in Section~\ref{sec:simulation}.}
+
+\textcolor{red}{Insert figure 6 which shows the reachable delta config space for a square for 1,2,3,4,5 moves, maybe show two examples.}
+
 % 

### 2018-12-06 journalWallFriction/resubmitv03/Responsev03.tex
Wrote the letter!

% difference between two pdf files: https://draftable.com/compare
% insert a pdf: https://tex.stackexchange.com/questions/8662/merge-two-pdf-files-output-by-latex
%   \usepackage{pdfpages}
%   \includepdf[pages=-]{ChangesFromV01toV02.pdf}
% gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dNOPAUSE -dQUIET -dBATCH      -sOutputFile=ResponseCompressed.pdf Response.pdf
%
\documentclass[conference, onecolumn, 10pt]{../IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{times}

\usepackage{microtype,xparse,tcolorbox}
\usepackage{mathtools, cuted}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{pdfpages}
\definecolor{myMagenta}{RGB}{233,63,246}
\definecolor{myDarkGreen}{RGB}{97,173,89}
\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
\pagestyle{plain}
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}
\usepackage{lastpage}

\newcommand{\revision}[1]{\textcolor{blue}{#1}}
\setlength\parindent{0pt}

\renewcommand{\thepoint}{[R\,\thereviewer.\arabic{point}] }

\newcommand{\response}[1]{\refstepcounter{point}  \smallbreak \noindent
	\hangindent=3.5em
	\hangafter=1
	\thepoint \space ``\textcolor{red}{#1}" \par\smallskip}

\newcommand{\letter}[2]{\vspace*{\fill}

	\begin{quote}

		\includegraphics[width=8cm]{engineeringLogo.eps}

		\begin{flushright}
		Aaron T.\ Becker\\Department of Electrical Engineering\\University of Houston\\Houston, TX 77004, USA\\Email: atbecker@central.uh.edu\\Web: youTube.com/aabecker5\\

		\today \vspace{.1cm} \\
		\end{flushright}

		#1\medskip\\

		Dear #2 Editorial Office,\medskip\\
		Please find attached the revised paper, #1, of the conditionally accepted  paper,
		along with the document containing a response to the reviewers. We are grateful to the reviewers for helping us improve our manuscript through their comments and questions. Please let us know if further information is required. This work extends the preliminary conference paper, ``\emph{Algorithms for shaping a particle swarm with a shared input by exploiting non-slip wall contacts}'', presented at the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS). The conference paper considered only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.\medskip
		\medskip

		The new revision of our paper includes multi-media so that others can build on our results. These include illustrative videos for the simulations and experiments: 
	\begin{itemize}	 
 \item Video  \emph{01Model.mov} animates the concepts of uniform control inputs and non-slip boundary contacts.
 \item Video  \emph{02DeltaConfigurationSpace.mp4} demonstrates how the $\Delta$ configuration space (the configuration space showing the difference in position for two particles) is constructed for a variety of workspaces.
 \item Video  \emph{03SimulationWorkSpaces.mp4} shows demonstrations of motion planning in square and disc-shaped workspaces. 
 \item Video  \emph{04Hardware Experiments.mp4} shows an experiment trial moving two particles to goal positions in a small intestine phantom, then an experiment trial using cow stomach tissue.
		\end{itemize}

		   We also include four  Mathematica Notebooks (.nb files) containing the simulation code.  This code generates plots for the paper and animates the path planner algorithm for arbitrary inputs.
		   	\begin{itemize}	 
 \item   \emph{SquareWorkSpace.nb} generates paths in a square workspace for two particles.
 \item  \emph{CircularWorkSpace.nb} generates paths in a circular workspace for two particles.
 \item   \emph{DeltaConfigurationSpacePolygon.nb} generates the $\Delta$ configuration space for convex polygons
  \item   \emph{ShortestPathForADisk.nb} analytical solution for shortest path that touches a boundary from one position to another position in a circular workspace.
		\end{itemize}

		%   At the end of this letter, pages~\pageref{sec_lastpage}--\pageref{LastPage} is the revised manuscript, formatted so all changes can be easily identified by the reviewers, by using green colored text to indicate revised passages. 
%		 The revised manuscript is formatted so all changes can be easily identified by the reviewers by using green colored text to indicate revised passages.

		Sincerely,\\

			\includegraphics[width=4cm]{SignatureAaronTBeckerWhite.png}
			\medskip\\
		Aaron T.\ Becker (on behalf of all the authors) 
	\end{quote}
	\vspace*{\fill}
	\break }

\NewDocumentCommand \Reviewer { m } {
	\stepcounter{reviewer}
	\subsection*{Comments~by~Reviewer~#1}
}

\begin{document}
	\letter{\emph{Exploiting Non-Slip Wall Contacts  to Position Two Particles Using The Same Control Input}}{IEEE Transactions on Robotics}

	\section*{Response to Reviewers}
	In the following document, we have provided detailed responses to the comments and questions of the reviewers. Comments and questions by reviewers are in \textcolor{red}{red}, our responses are in black.

	\emph{Comments by Associate Editor}
	\response{Both reviewers were pleased with the authors' revised manuscript. It is
the opinion of the Associate Editor that this paper will be publishable
in T-RO in some form. However, both reviewers have lingering concerns
about discrepancies between the purported contribution of the paper and
what is actually demonstrated. In the next draft of the manuscript, the
authors are encouraged to be explicit in making sure that the stated
contributions match what has been shown. This could involve removing or
softening certain claims, or adding new material to strengthen certain
claims.}

	Thank you for this opportunity!

	\Reviewer{\#1}\response{In the revised version most of my previous concerns have been
addressed.  I think with minor revisions the manuscript could be
suitable for publication. 
The remaining issue still comes back to whether the iterative procedure is able to reach all target configurations.  The new manuscript shows numerically that the accessible space approaches the entire space for a circular 2D domain.  The paper, however, is meant to also cover any convex domains.  Can the authors also prove that the accessible space approaches the entire domain for any convex domain?  Or at least other polygonal cases?   I see that on p4 in the last paragraph of section C, there is a statement that particles can be steered to arbitrary positions, but without an explanation of how that result is obtained, which should be included.  I also think that paragraph should be highlighted a little more by a section title (to be parallel with the development of the circular workspace).  Assuming this can be explained then I would be satisfied with the manuscript.
}
	Thank you! This was a fun extension to work on.  In a similar form to the work with circular domain we now show that the accessible space rapidly grows to include the entire $\Delta$ configuration space.  For finite-sided polygons, this convergence requires finite time (assuming the particles have at least $\epsilon$ separation).  We have augmented Fig.~4 to show (as needed) the 4, 6, and 8-move reachable sets for polygons up to 20 sides.  These sets reach the entire $\Delta$ configuration space.

	\response{ I also think that paragraph should be highlighted a little more by a section title (to be parallel with the development of the circular workspace).  Assuming this can be explained then I would be satisfied with the manuscript.}
	Great idea! We added Section 
	\revision{IV.-D Convex Workspaces: Accessible Region}.

	This section includes the new text:

	``\revision{
As the corner angles increase, the number of moves required to access the entire configuration space increases.  
As shown in Fig.~4,  four moves are sufficient for 3-sided polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided polygons.  Eight moves is sufficient for 20-sided polygons, but we have not checked polygons with more sides. For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~VI.}''

	\Reviewer{\#2}
	\response{The authors did a great job in addressing my and the other reviewer's comments in a set of meaningful changes to the manuscript, including new analysis to investigate many of the questions. 
The only remaining comment is that the abstract and introduction imply that 3D results are obtained in this work. While this is shown theoretically, no 3D experimental results are shown. The abstract and introduction should state this explicitly as this is viewed as a notable limitation in what is actually demonstrated.
}
	Thank you. This has been a fun paper to work on. 
	 You are correct that no 3D experiments were performed.  
	 We corrected the abstract and introduction as follows:

	``{\bf Because \emph{in vivo} environments are usually not square, this paper extends the previous work to all convex workspaces, \revision{ and shows how this could be extended} to 3D positioning \revision{of neutrally buoyant particles.} 
We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that are applied uniformly to all particles in the workspace.
This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the gastrointestinal tract.}''

and at the end of the introduction:
``This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the anatomy of the gastrointestinal tract.''

	 We hope in the future to use our new magnetic manipulator to test 3D movement, but are not yet ready for these experiments (and S. Shahrokhi, J. Shi, and B. Isichei have all graduated!)

\end{document}

### 2018-12-06 journalWallFriction/theory2robAlgs.tex
Wrote the letter!

--- 2018-10-13 journalWallFriction/theory2robAlgs.tex
@@
 For a square workspace where the length of each side is $L$, the worst case path length is $(\sqrt{2}+2)L$, and requires at most five moves (see Fig.~\ref{fig:shapeControlMathematica1}f).
+\revision{
+As the corner angles increase, the number of moves required to access the entire configuration space increases.  
+As shown in Fig.~\ref{fig:polygon},  four moves are sufficient for 3-sided polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided polygons.
+Eight moves is sufficient for 20-sided polygons, but we have not checked polygons with more sides.
+For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~\ref{sec:simulation}.}
+
 % 

### 2018-12-06 journalWallFriction/theory2robAlgs.tex
Added new Fig 4, the caption needs to be changed, but just for Dr. Becker to see if it is what he wants.

@@
 \renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,100){Workspace}\put(20,100){$\Delta$ configuration space}
+{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,102){Workspace}\put(20,102){$\Delta$ configuration space}
 \end{overpic}
@@
 If $\Delta g$ is in the 2-move reachable sets, we can achieve the required relative position in two moves.
-If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal $\Delta g_c$ (the closest point on the 2-move reachable set to $\Delta g$) and apply two moves to achieve $\Delta g_c$.
+%If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal $\Delta g_c$ (the closest point on the 2-move reachable set to $\Delta g$) and apply two moves to achieve $\Delta g_c$.
 We repeat this process until the relative goal position is achieved.

### 2018-12-07 journalWallFriction/theory2robAlgs.tex
fixed figure 3  (whew!)

@@
 \begin{overpic}[width=\columnwidth]{irregular.pdf}
-\put(0,10){\scriptsize
-$\Delta s$ \tikz \draw[green,fill=green,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
-\tikz \draw[red,fill=red,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
-\tikz \draw[blue,fill=blue,line width=0.3mm] (-1ex,0) rectangle (0,1ex); 
-}
-\put(0,5){\scriptsize
-$\Delta g$ \tikz \draw[green,fill=green,line width=0.3mm] (0,0) circle (0.5ex); $=$
-\tikz \draw[red,fill=red,line width=0.3mm] (0,0) circle (0.5ex); $-$
-\tikz \draw[blue,fill=blue,line width=0.3mm] (0,0) circle (0.5ex); 
-}
+%\put(0,10){\scriptsize
+%$\Delta s$ \tikz \draw[green,fill=green,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
+%\tikz \draw[red,fill=red,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
+%\tikz \draw[blue,fill=blue,line width=0.3mm] (-1ex,0) rectangle (0,1ex); 
+%}
+%\put(0,5){\scriptsize
+%$\Delta g$ \tikz \draw[green,fill=green,line width=0.3mm] (0,0) circle (0.5ex); $=$
+%\tikz \draw[red,fill=red,line width=0.3mm] (0,0) circle (0.5ex); $-$
+%\tikz \draw[blue,fill=blue,line width=0.3mm] (0,0) circle (0.5ex); 
+%}
 \end{overpic}
@@
 Workspace and $\Delta$ configuration space is shown for an arbitrary convex polygon with $n=4$ sides.
-The 2-move reachable sets for this initial configuration ($\Delta s= s_2-  s_1$) are drawn in transparent blue.}
+% The 2-move reachable sets for this initial configuration ($\Delta s= s_2-  s_1$) are drawn in transparent blue.
+}
 \end{figure}

### 2018-12-07 journalWallFriction/theory2robAlgs.tex
corrected statement about regular polygons

@@
 As the corner angles increase, the number of moves required to access the entire configuration space increases.  
-As shown in Fig.~\ref{fig:polygon},  four moves are sufficient for 3-sided polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided polygons.
-Eight moves is sufficient for 20-sided polygons, but we have not checked polygons with more sides.
-For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~\ref{sec:simulation}.}
+As shown in Fig.~\ref{fig:polygon},  four moves are sufficient for 3-sided regular polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided regular polygons.
+Eight moves is sufficient for polygons with less than 14-sides, but we have not checked polygons with more sides.
+For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~\ref{sec:simulation}.
+If the polygon is irregular, more moves may be required.  The entire configuration space is reachable in four moves for an acute triangle, but obtuse triangles require six moves.
+}
 

### 2018-12-11 journalWallFriction/resubmitv03/Responsev03.tex
cleaned images, proofed test, and prepped letter

@@
 
-	Thank you for this opportunity!
+	Thank you for this opportunity! In addition to the changes listed below, we used this opportunity to remove a few typos and clean up several figures.
 
@@
 For finite-sided polygons, this convergence requires finite time (assuming the particles have at least $\epsilon$ separation).
-We have augmented Fig.~4 to show (as needed) the 4, 6, and 8-move reachable sets for polygons up to 20 sides.
+We have augmented Fig.~4 to show (as needed) the 4, 6, and 8-move reachable sets for polygons up to 10 sides.
 These sets reach the entire $\Delta$ configuration space.
@@
 As the corner angles increase, the number of moves required to access the entire configuration space increases.  
-As shown in Fig.~4,  four moves are sufficient for 3-sided polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided polygons.
-Eight moves is sufficient for 20-sided polygons, but we have not checked polygons with more sides.
-For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~VI.}''
+As shown in Fig.~4,  four moves are sufficient for 3-sided regular polygons, and six moves accesses the entire $\Delta$ configuration space for up to 8-sided regular polygons.
+Eight moves is sufficient for polygons with less than 14-sides, but we have not checked polygons with more sides.
+For a circular workspace, with corner angles of 180$^\circ$, the worst-case configuration can only be approached asymptotically, as explained in Section~VI.
+If the polygon is irregular, more moves may be required.  The entire configuration space is reachable in four moves for an acute triangle, but obtuse triangles require six moves.
+}''
 

### 2018-12-11 journalWallFriction/simulation.tex
cleaned images, proofed test, and prepped letter

@@
  If the goal configuration cannot be reached in one move but can be reached in three moves, the shortest path has a simple solution.
-The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final ositions.
+The first move, $m_1$, makes one particle contact a wall, $m_2$ adjusts the relative spacing error  to zero, and $m_3$ takes the particles to their final positions.
 $m_2$ cannot be shortened, so optimization depends on choosing the location where the particle contacts the wall. 
@@
  (\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (0,0) circle (.5ex);,
- \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);) are fixed.
+ \protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (0,0) circle (.5ex);) 
  has the property that the incident angle equals the reflected angle, as shown at left.
@@
 \renewcommand{\figwid}{\columnwidth}
-{\begin{overpic}[width =\figwid]{shortestpath.pdf}
+{\begin{overpic}[width =\figwid]{shortestpath3.pdf}
 \end{overpic}
@@
 Bottom six images show a disk-shaped workspace and the corresponding $\Delta$ configuration space with its 2-move reachable sets. 
-For each the move 1 and 3 are simple translations of both particles and so the reachable sets do not change.
+For each, moves 1 and 3 are simple translations of both particles and so the reachable sets do not change.
 The reachable set morphs during move 2 because one particle is held stationary by the boundary. 
@@
 %Plots show performance with one goal on the boundary.
-Contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
+Circular workspace: contour plots showing the number of moves and distance commanded if red particle's goal position is varied in $x$ and $y$. 
 Starting positions of red and blue particles  
@@
 \protect\tikz \protect\draw[myDarkGreen,fill=myDarkGreen,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $=$
-\protect\tikz \protect\draw[red,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
-\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
+\protect\tikz \protect\draw[red,fill=red,line width=0.3mm] (-1ex,0) rectangle (0,1ex); $-$
+\protect\tikz \protect\draw[blue,fill=blue,line width=0.3mm] (-1ex,0) rectangle (0,1ex);}
 \end{overpic}
@@
 \caption{\label{fig:deltanumdist}
-The worst-case path length occurs when particles must swap antipodes.
+Circular workspace: the worst-case path length occurs when particles must swap antipodes.
 This can never be achieved but can be asymptotically approached.
@@
 \end{overpic}
-}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ 
+}\caption{\label{fig:contourPlots}{Square workspace: starting positions of particles $1$ and $2$ 
  (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,
@@
 The top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} and the bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\}, each in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace.
-Fig.~\ref{fig:contour} right shows the number of moves and left shows the total distance of the path.
+Fig.~\ref{fig:contour} left shows the number of moves and right shows the total distance of the path.
 This plot shows the nonlinear nature of the path planning.

### 2018-12-11 journalWallFriction/theory.tex
cleaned images, proofed test, and prepped letter

@@
  In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
- If the $i^{\textrm{th}}$ particle has position $\mathbf{x}_i(t)$ and velocity $\dot{\mathbf{x}}_i(t)$,  we assume the following system model:
+ If the $i^{\textrm{th}}$ particle has position $p_i(t)$ and velocity $\dot{p}_i(t)$,  we assume the following system model:
   \begin{align}\label{eq:swarmDynamicsAndFric} 
-\dot{\mathbf{x}}_i(t)
+\dot{p}_i(t)
  &= \begin{cases}
-  0 &\begin{matrix} \mathbf{x}_i(t) \in  \textrm{boundary \textbf{and}}\\
-\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix}
-\textrm{    ,} \\
- \mathbf{u}(t) & \textrm{else}.
- \end{cases}
+  0 &\begin{matrix} p_i(t) \in  \textrm{boundary \textbf{and}}\\
+\mathbf{N}(\textrm{boundary$_{p_i(t)}$})\cdot   \mathbf{u}(t) \le 0 \end{matrix} \\
+ \mathbf{u}(t) & \textrm{else}
+ \end{cases}.
  \end{align}
  Here
- $\mathbf{N}(\textrm{boundary$_{\mathbf{x}_i(t)}$})$ is the normal to the boundary at position $\mathbf{x}_i(t)$ and the frictional force provided by the boundary cancels any control force $ \mathbf{u}(t)$ that pushes into the boundary.
+ $\mathbf{N}(\textrm{boundary$_{p_i(t)}$})$ is the normal to the boundary at position $p_i(t)$ and the frictional force provided by the boundary cancels any control force $ \mathbf{u}(t)$ that pushes into the boundary.
 

### 2018-12-11 journalWallFriction/theory2robAlgs.tex
cleaned images, proofed test, and prepped letter

@@
 \begin{algorithm}[htb]
-\caption{ { \sc 2-ParticlePathPlanning}($s_1,s_2,g_1,g_2,P,\epsilon$)}\label{alg:optimalAlg}
+\caption{ { \sc 2-ParticlePathPlan}($s_1,s_2,g_1,g_2,P,\epsilon$)}\label{alg:optimalAlg}
 \begin{algorithmic}[1]
@@
 \While {$||\Delta p - \Delta g|| > \epsilon$}  %R.p_1 \ne g_1$ \textbf{and} $R.p_2 \ne g_2$}
-\State $R_{\textrm{SET}}\gets$  Compute 2-move reachable set  \Comment use Alg.~\ref{alg:polygonReachbale} or \ref{alg:circularReachbale}
+\State $R_{\textrm{SET}}\gets$  Compute 2-move reachable set \newline
+ \Comment  use Alg.~\ref{alg:polygonReachbale} or \ref{alg:circularReachbale}
 \State $ \Delta g_c\gets $nearest point in $R_{\textrm{SET}}$ to $\Delta g$
@@
  Figure ~\ref{fig:polygon} shows the starting and ending relative positions as $\Delta s$ and $\Delta g$ in the $\Delta$ configuration space.
-The next subsections give procedures to compute the 2-move reachable set.% in Fig.~\ref{fig:regionMove}.% The reachable set is the part of $\Delta$ configuration space where if one particle touches a wall in a specific location, the other particle can make the required relative distance without causing the touching particle to move.
-We will discuss how to make reachable sets in the next subsections.
+The next subsections give procedures to compute the 2-move reachable set.
+ \revision{The $k+2$-move reachable set is constructed by the union of 2-move reachable sets starting from the boundary of the $k$-move reachable set.}
 
@@
 If $\Delta g$ is in the 2-move reachable sets, we can achieve the required relative position in two moves.
+\revision{If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal $\Delta g_c$ (the closest point on the 2-move reachable set to $\Delta g$) and apply two moves to achieve $\Delta g_c$.
+We repeat this process until the relative goal position is achieved.
+Higher-number reachable sets are drawn in distinct colors.}
 %If $\Delta g$ is not in the 2-move reachable set, we define a temporary goal $\Delta g_c$ (the closest point on the 2-move reachable set to $\Delta g$) and apply two moves to achieve $\Delta g_c$.
@@
 \end{figure*}
- Figure \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the 2-move reachable sets that correspond to representative initial conditions.
+ Figure \ref{fig:polygon} shows six workspaces, their $\Delta$ configuration spaces, and the \revision{$k$}-move reachable sets that correspond to representative initial conditions.
  Figure \ref{fig:TwoRegions} highlights the construction of the 2-move reachable sets for a square workspace.
@@
 $P$ is a list of the vertices of a convex polygon. %RSet contains all the reachable polygons.
-\State $R_{\textrm{SET}\gets \{\}}$
+\State $R_{\textrm{SET}}\gets \{\}$
 \For {$p_i$ in $P$}
@@
 (b) if the blue particle touches the wall at $\psi_{\min}$ (blue square) the other particle (red square) can move anywhere in the pink region. 
-Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the lightgreen region. 
+Right bottom: if the blue particle touches the wall at $\psi = \frac{17 \pi}{10}$ (blue square) the other particle (red square) can move anywhere in the light green region. 
 (c) The $\Delta$ configuration space for the corresponding starting positions of the particles is shown. 
@@
 \centering
-\begin{overpic}[width=0.95\columnwidth]{zaxis.pdf}\end{overpic}
+\begin{overpic}[width=0.95\columnwidth]{zaxis.pdf}
+\scriptsize
+\put(1,51){Move 1}
+\put(50,51){Move 2}
+\end{overpic}
 \caption{\label{fig:zaxis}

### 2018-12-20 journalWallFriction/SwarmShapeControl.tex
Added code for the plots
the code makes really big images.  You can rasterize them afterwards.
Wrote the letter!
Sheryl's next paper (?)
s

@@
 
-\newcommand{\revision}[1]{
-  \textcolor{Blue}{#1}
-}
+\newcommand{\revision}[1]{\textcolor{blue}{#1}}
 
@@
 \author{Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker% <-this % stops a space
-\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]} and \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}.}% <-this % stops a space
+\thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}, \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}, and  \href{https://nsf.gov/awardsearch/showAward?AWD_ID=1646566}{[CNS-1646566]}.}% <-this % stops a space
 \thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
@@
 This paper also implements the algorithms \revision{in 2D} using a hardware setup inspired by the gastrointestinal tract.
+
+% SHORTER ABSTRACT:
+%Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
+This paper investigates position control of particles using uniform forces (the same force is applied everywhere in the workspace).
+Given a controllable field that can generate bidirectional forces in three orthogonal directions, steering one particle in 3D is trivial.
+Adding additional particles to steer makes the system underactuated because there are more states than control inputs.
+However, the walls of in vivo and artificial environments often have surface roughness such that particles do not move unless actuation pulls them away from the wall.
+This paper shows that the individual 2D position of two particles is controllable using global inputs in all convex workspaces with non-slip wall contact, and shows how this could be extended to 3D positioning of neutrally buoyant particles.
+We investigate analytically an idealized variant of this problem with non-slip boundaries and control inputs that apply uniformly to all particles in the workspace.
+This paper implements the algorithms in 2D using a hardware setup inspired by the gastrointestinal tract.
 

### 2019-01-03 journalWallFriction/Conclusion.tex
added stubs for the final journal version.  Still need pictures

@@
   \section*{Appendix}\label{sec:appendix}
+We thank our anonymous reviewers for their advice which improved the presentation and content of this paper, especially the analysis of the reachable sets.
  The supplemental material includes four videos:

### 2019-01-03 journalWallFriction/SwarmShapeControl.tex
added stubs for the final journal version.  Still need pictures

@@
 %\documentclass[conference]{IEEEtran}
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
+\documentclass[journal]{IEEEtran}
 \IEEEoverridecommandlockouts% This command is only needed if 
@@
 
-\newcommand{\revision}[1]{\textcolor{blue}{#1}}
+\newcommand{\revision}[1]{\textcolor{black}{#1}}  %changed to black for final
 
@@
 % paper title
-\title{\LARGE \bf Exploiting Non-Slip Wall Contacts \\  to Position Two Particles Using The Same Control Input}
+\title{Exploiting Non-Slip Wall Contacts  to Position\\ Two Particles Using The Same Control Input}
 
 % You will get a Paper-ID when submitting a pdf file to the conference system
-\author{Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker% <-this % stops a space
+\author{Shiva Shahrokhi,~\IEEEmembership{Member,~IEEE,} Jingang Shi,  Benedict Isichei, and Aaron T. Becker,~\IEEEmembership{Senior Member,~IEEE,}% <-this % stops a space
 \thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}, \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}, and  \href{https://nsf.gov/awardsearch/showAward?AWD_ID=1646566}{[CNS-1646566]}.}% <-this % stops a space
 \thanks{Authors are with the Department of Electrical and Computer Engineering,  University of Houston, Houston, TX 77204 USA        {\tt\small  \{sshahrokhi2,atbecker\}@uh.edu}}%
-}
+\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
 %\affil{Department of Electrical and Computer Engineering, \\
@@
 %} %\end thanks
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
+{Shahrokhi \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
 
@@
 \end{abstract}
+\begin{IEEEkeywords}
+Underactuated Robots, Motion Control, Path Planning for Multiple Robot Systems, Configuration Space.
+\end{IEEEkeywords}
 
@@
 }
-
-% Uncomment to add appendix:
-%\input{appendix}
+%Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker
+%
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) received M.S.
+and PhD degrees in the Electrical and Computer Engineering Department at the University of Houston in 2018.
+She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
+Tehran, Iran. She is involved in studies of swarm manipulation and control for robotics applications.
+These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.
+\end{IEEEbiography}
+%
+\begin{IEEEbiography}{Jingang Shi}
+Biography text here.
+\end{IEEEbiography}
+%
+\begin{IEEEbiography}{Benedict Isichei}
+received an M.S. degree in the Electrical and Computer Engineering Department at the University of Houston in 2018.
+\end{IEEEbiography}
+%
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T.
+Becker}(S`06-M`12-SM`17) received M.S.
+and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012.
+He was a post doctoral research scholar at the Multi-Robot Systems Lab at Rice University and at Boston Children's Hospital \& Harvard Medical School, before joining the Electrical and Computer Engineering Department at the University of Houston as an Assistant Professor in 2014.
+He won the best paper at IROS 2013 and received the NSF CAREER award in 2016.
+%He is active in developing control laws and algorithms for robot swarms.
+\end{IEEEbiography}
 

### 2019-01-03 journalWallFriction/intro.tex
added stubs for the final journal version.  Still need pictures

@@
 \section{Introduction}\label{sec:Intro}
-Particle swarms propelled by an external field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
+\IEEEPARstart{P}{article} swarms propelled by an external field, where each particle  receives the same control input, are common in applied mathematics, biology, and computer graphics \cite{Peyer2013,Shirai2005,Chiang2011}.
 %

### 2019-01-04 journalWallFriction/SwarmShapeControl.tex
Added pictures and refs, shrank to 12 pages

@@
 \usepackage[ruled]{algorithm}
-\usepackage{algpseudocode}
+\usepackage[noend]{algpseudocode}
 %\usepackage{algorithmic}
@@
 % You will get a Paper-ID when submitting a pdf file to the conference system
-\author{Shiva Shahrokhi,~\IEEEmembership{Member,~IEEE,} Jingang Shi,  Benedict Isichei, and Aaron T. Becker,~\IEEEmembership{Senior Member,~IEEE,}% <-this % stops a space
+\author{Shiva Shahrokhi,~\IEEEmembership{Member,~IEEE,} Jingang Shi,  Benedict Isichei, and Aaron T. Becker,~\IEEEmembership{Senior Member,~IEEE}% <-this % stops a space
 \thanks{*This work was supported by the National Science Foundation under Grant No.\ \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1553063}{ [IIS-1553063]}, \href{http://nsf.gov/awardsearch/showAward?AWD_ID=1619278}{[IIS-1619278]}, and  \href{https://nsf.gov/awardsearch/showAward?AWD_ID=1646566}{[CNS-1646566]}.}% <-this % stops a space
@@
 }
-%Shiva Shahrokhi, Jingang Shi,  Benedict Isichei, and Aaron T. Becker
-%
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) received M.S.
@@
 %
-\begin{IEEEbiography}{Jingang Shi}
-Biography text here.
+\vspace{-1em}
+\begin{IEEEbiography}
+[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{JingangShi.jpg}}]{Jingang Shi}
+Jingang Shi received the B.Eng. degree in the School of Materials Science and Engineering at Huazhong University of Science and Technology in 2018. 
+ He studied robot control during a four-month internship at the University of Houston in 2017.
+ He is studying for the M.E. degree in the School of Mechanical Science and Engineering at Huazhong University of Science and Technology.
 \end{IEEEbiography}
 %
-\begin{IEEEbiography}{Benedict Isichei}
+\vspace{-1em}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Benedict.jpg}}]{Benedict Isichei}
 received an M.S. degree in the Electrical and Computer Engineering Department at the University of Houston in 2018. 
+ His thesis was titled ``Robotic Manipulators: Cryogenic Magnetic Manipulation System, Two-body Magnetic Manipulation, and Low-cost Robotic Manipulators'', and features low cost robotics.
+See his designs at https://www.thingiverse.com/Bisichei.
 \end{IEEEbiography}
 %
+\vspace{-1em}
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T.
@@
 \end{IEEEbiography}
-
+%
 \end{document}

### 2019-01-04 journalWallFriction/simulation.tex
Added pictures and refs, shrank to 12 pages

@@
 }
+ \vspace{-1em}
 \caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference.
@@
 \begin{overpic}[width=0.67\columnwidth]{finalMove.pdf}\end{overpic}
+\vspace{-1em}
 \caption{\label{fig:reachableSet}
@@
 \begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){d)}\end{overpic}
+\vspace{-1em}
 \caption{\label{fig:contour}
@@
 \end{overpic}
-}\caption{\label{fig:contourPlots}{Square workspace: starting positions of particles $1$ and $2$ 
+}\vspace{-1em}
+\caption{\label{fig:contourPlots}{Square workspace: starting positions of particles $1$ and $2$ 
  (\protect\tikz \protect\draw[blue,fill=white,line width=0.3mm] (-1ex,0) rectangle (0,1ex);,

### 2019-02-11 ICRA19/covControl.tex
smaller edits done.

@@
 
-The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
+The Region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
 \begin{align}\label{eq:meanInSquareWorkspaceLL}
@@
 %
-The particles used are iron filings (Dowling Magnets, \#731019, $\approx0.5$~mm diameter), and the motive force is gravity. 
+The particles used are iron filings (Dowling Magnets, \#731019, $\approx0.5$~mm diameter), and the motive force is gravity as shown in Fig.~\ref{fig:angleOfReposeExp}. 
 The particles were poured into a frame composed of two sheets of acrylic with 4.5 mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
@@
 The experimental setup is shown in Fig.
-\ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$, generating three different shapes, as in Fig.~\ref{fig:angle}.
+\ref{fig:SetUp} also showing angle of repose of 56 Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$, generating three different shapes, as in Fig.~\ref{fig:angle}.
 The green lines show the triangle shape the robots pile up into.

### 2019-02-11 ICRA19/intro.tex
smaller edits done.

@@
 After a review of recent related work in Sec.~\ref{sec:RelatedWork},
-  Sec.~\ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
-Section \ref{sec:covControl} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.
+  Sec.~\ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control, shown in Fig.~\ref{fig:leadfigure} bottom.
+Section \ref{sec:covControl} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls, shown in Fig.~\ref{fig:leadfigure} top.
 These results are limited in the set of shapes that can be generated.

### 2019-02-13 ICRA19/covControl.tex
Made small updates to ICRA paper

@@
 Analytical results comparing maximum correlation $\rho_{xy}$.
-Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
+Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four different boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
 }\end{figure}
@@
 \caption{\label{fig:angleOfReposeExp}
-Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
+Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity (see \cite{shahrokhiICRAvideo2019}).
 }\vspace{-1.5em}
@@
 \caption{\label{fig:SetUp}
-Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks.
+Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks (see \cite{shahrokhiICRAvideo2019}).
 }\vspace{-1.5em}

### 2019-02-13 ICRA19/relatedWork.tex
Made small updates to ICRA paper

@@
 % The magnetic gradients of MRI scanners are \emph{uniform}, meaning the same force is applied everywhere in the workspace\cite{nosrati2018development}.
- This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013}, and breaks the control symmetry using obstacles in the workspace. 
+ This work assumes a uniform control with homogenous particles, as in~\cite{AaronManipulation2013,shahrokhiTRO2019}, and breaks the control symmetry using obstacles in the workspace. 
   While in \cite{shahrokhi2018TRO} we showed how to control the mean position and variance through repeated collisions with the workspace, this paper focuses on stable configurations using few pushes.

### 2019-02-13 journalWallFriction/FinalSubmission/theory2robAlgs.tex
Made small updates to ICRA paper

--- 2018-12-11 journalWallFriction/theory2robAlgs.tex
@@
 \end{overpic}
+\vspace{-2em}
 \caption{\label{fig:irregular}
@@
 \renewcommand{\figwid}{0.8\columnwidth}
-{\begin{overpic}[width =\figwid]{differentNumSides.pdf}\put(5,102){Workspace}\put(20,102){$\Delta$ configuration space}
+{\begin{overpic}[width =\figwid]{differentNumSidesNew.pdf}\put(5,102){Workspace}\put(20,102){$\Delta$ configuration space}
 \end{overpic}
@@
 \end{overpic}
-\vspace{-1em}
+\vspace{-2em}
 \caption{\label{fig:regionMove}
@@
 \begin{overpic}[width=\columnwidth]{reachableSetsFuncDistance}\end{overpic}
+\vspace{-1em}
 \caption{\label{fig:reachableSetsFuncDistance}
@@
 \centering
-\begin{overpic}[width=.8\columnwidth]{ReachableSets0p1}\end{overpic}
-\begin{overpic}[width=.8\columnwidth]{ReachableSets0p3}\end{overpic}
+\begin{overpic}[width=.6\columnwidth]{ReachableSets0p1}\end{overpic}
+\begin{overpic}[width=.6\columnwidth]{ReachableSets0p3}\end{overpic}
 \caption{\label{fig:reachableSets}

### 2019-02-25 ICRA19/covControl.tex
added references to the video

@@
 \caption{\label{fig:angleOfReposeExp}
-Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity (see \cite{shahrokhiICRAvideo2019}).
+Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity (see video \cite{shahrokhiICRAvideo2019}).
 }\vspace{-1.5em}
@@
 \caption{\label{fig:SetUp}
-Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks (see \cite{shahrokhiICRAvideo2019}).
+Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks (see video \cite{shahrokhiICRAvideo2019}).
 }\vspace{-1.5em}

### 2019-02-25 ICRA19/intro.tex
added references to the video

@@
      Reshaping particles by collisions with rigid bodies. Top pictures illustrate using closed boundaries to reshape the particles. 
-      Bottom pictures illustrate with different particles how varying the the angle of attack modifies the shape of particles remaining on a rigid object and how modifying the applied force changes the angle of repose.
+      Bottom pictures illustrate with different particles how varying the the angle of attack modifies the shape of particles remaining on a rigid object and how modifying the applied force changes the angle of repose (see video \cite{shahrokhiICRAvideo2019}).
 }
